\section{Related Work}
\label{sec:related}
\noindent\textbf{Context window extension.}
Extending the context window has become a key direction for improving long-context performance, with recent LLMs supporting 128K and even 1M context windows~\citep{yang2025qwen3,hurst2024gpt,team2024gemini_1_5,google2024gemini_2,google2025gemini2_5}.
Existing approaches either extend context windows through lightweight methods, such as positional extrapolation~\citep{chen2023extending,pengyarn,ding2024longrope,zhang2024extending,zhu2023pose} and attention modifications~\citep{chen2023longlora,xiao2023efficient,xiao2024infllm,bertsch2023unlimiformer,jin2024llm}, or rely on continued pre-training to build more robust long-context capability~\citep{xiong2024effective,fu2024data,gao2025prolong,Yang2025Qwen251MTR}.
Our work follows the continued pre-training method, but studies it in multimodal settings where long contexts contain interleaved image and text tokens.

\noindent\textbf{Long-context vision-language models.}
As LLM context windows have expanded, recent LVLMs such as Gemini 3.1 Pro~\citep{gemini_3_1_pro_2026}, Claude Sonnet 4.7~\citep{anthropic2026introducing_claude_opus_4_7}, and Qwen3-VL~\citep{bai2025qwen3} have also supported substantially longer contexts.
However, recent LVLM technical reports~\citep{bai2025qwen3,hong2025glm} reveal limited details about how long-context capability is actually built, leaving practical LongPT recipes underexplored.
Concurrent work~\citep{veselka2026longvlm} studies long-document data construction for LVLMs, but mainly builds on backbones that already support 128K or longer contexts, such as Qwen3-VL~\citep{bai2025qwen3} and Mistral 3.1~\citep{mistralai2025mistral_small_3_1}; thus, its findings may reflect context alignment rather than true context extension.
For example, they find that 1B-token LongPT outperforms its 10B-token counterpart, and that LongSFT outperforms LongPT.
In contrast, we study LongPT on Qwen2.5-VL~\citep{bai2025qwen2}, whose native context window is only 32K, allowing us to directly examine how to extend LVLMs to longer context.
Another line of work studies long-video understanding~\citep{chen2024longvila,shen2025longvita,yang2025eea,zhang2024longva,yang2025longvt,liu2025bolt,shen2024longvu}, but these methods are often specialized for temporal redundancy and video token reduction rather than general long-context LVLM training.

\noindent\textbf{Multimodal long-context evaluation.}
Recent benchmarks evaluate multimodal long-context understanding from diverse perspectives, including long-document VQA~\citep{ma2024mmlongbenchdoc,deng2025longdocurl}, multimodal needle-in-a-haystack~\citep{wang2024multimodal,wu2024visual}, vision-text compression~\citep{zhao2025vtcbench}, and long-video understanding~\citep{fu2025video,zhou2025mlvu,wang2025lvbench}.
Among them, MMLongBench~\citep{wang2025mmlongbench} provides a comprehensive evaluation across five task categories with standardized context lengths up to 128K.
Our evaluation covers MMLongBench, VTCBench, and long-video benchmarks, demonstrating the broad generalization of our model \modelname.