\section{More Visualizations}
We include extended 24-second multi-agent rollouts, qualitative comparisons against baselines, and real-world robotic coordination videos in the supplementary video, offering a more intuitive demonstration of our method.

\section{Proof of Simplex Equidistance}
\label{app:simplex-equidistance}

We provide the derivation for the equidistance property of Simplex Rotary Agent Encoding. 
Let $V$ denote the simplex pool size, matching the notation of the main paper.
We first construct the simplex vertices in $\mathbb{R}^{V}$ and then embed them into the $(d_p/2)$-dimensional agent angle space, where $d_p$ is the agent rotary band size and $V\leq d_p/2+1$.

Let $\mathbf{e}_p\in\mathbb{R}^{V}$ denote the one-hot vector of agent $p$, and let $\mathbf{1}\in\mathbb{R}^{V}$ denote the all-one vector. 
We define the centered vector
\begin{equation}
\bar{\mathbf{s}}_p
=
\mathbf{e}_p-\frac{1}{V}\mathbf{1}.
\end{equation}
These vectors lie in the zero-mean subspace because
\begin{equation}
\mathbf{1}^{\top}\bar{\mathbf{s}}_p=0.
\end{equation}
For any agent $p$, the squared norm of $\bar{\mathbf{s}}_p$ is
\begin{align}
\|\bar{\mathbf{s}}_p\|_2^2
&=
\left(
\mathbf{e}_p-\frac{1}{V}\mathbf{1}
\right)^{\top}
\left(
\mathbf{e}_p-\frac{1}{V}\mathbf{1}
\right) \\
&=
1-\frac{2}{V}+\frac{1}{V} \\
&=
\frac{V-1}{V}.
\end{align}
For two distinct agents $p\neq q$, their inner product is
\begin{align}
\bar{\mathbf{s}}_p^{\top}\bar{\mathbf{s}}_q
&=
\left(
\mathbf{e}_p-\frac{1}{V}\mathbf{1}
\right)^{\top}
\left(
\mathbf{e}_q-\frac{1}{V}\mathbf{1}
\right) \\
&=
0-\frac{1}{V}-\frac{1}{V}+\frac{1}{V} \\
&=
-\frac{1}{V}.
\end{align}
After normalization, we obtain
\begin{equation}
\mathbf{s}_p
=
\sqrt{\frac{V}{V-1}}
\bar{\mathbf{s}}_p.
\end{equation}
Therefore,
\begin{align}
\|\mathbf{s}_p\|_2^2 &= 1, \\
\mathbf{s}_p^{\top}\mathbf{s}_q
&=
-\frac{1}{V-1},
\qquad p\neq q.
\end{align}
The squared distance between any two distinct simplex vertices ($V\geq 2 $) is then
\begin{align}
\|\mathbf{s}_p-\mathbf{s}_q\|_2^2
&=
\|\mathbf{s}_p\|_2^2
+
\|\mathbf{s}_q\|_2^2
-
2\mathbf{s}_p^{\top}\mathbf{s}_q \\
&=
2+\frac{2}{V-1} \\
&=
\frac{2V}{V-1},
\qquad p\neq q.
\end{align}
Thus, all distinct agent pairs are exactly equidistant in the simplex angle space.

In Simplex Rotary Agent Encoding, the agent angle is defined as
\begin{equation}
\boldsymbol{\theta}_p
=
\alpha\mathbf{s}_p,
\end{equation}
where $\alpha$ is a scalar scale factor. 
Therefore, for any $p\neq q$,
\begin{align}
\|\boldsymbol{\theta}_p-\boldsymbol{\theta}_q\|_2^2
&=
\alpha^2
\|\mathbf{s}_p-\mathbf{s}_q\|_2^2 \\
&=
\alpha^2
\frac{2V}{V-1}.
\end{align}
Hence, all different-agent pairs receive the same amount of separation in the underlying agent angle space. We further analyze the induced distance in the complex RoPE space. 
The complex rotary representation of agent $p$ is
\begin{equation}
\boldsymbol{\Phi}_p
=
\exp(i\boldsymbol{\theta}_p).
\end{equation}
For two agents $p$ and $q$, the squared distance between their complex rotations is
\begin{align}
\|\boldsymbol{\Phi}_p-\boldsymbol{\Phi}_q\|_2^2
&=
\left\|
\exp(i\boldsymbol{\theta}_p)
-
\exp(i\boldsymbol{\theta}_q)
\right\|_2^2 \\
&=
\sum_{r=1}^{d_p/2}
\left|
e^{i\theta_p^r}
-
e^{i\theta_q^r}
\right|^2 \\
&=
\sum_{r=1}^{d_p/2}
2\left(
1-\cos(\theta_p^r-\theta_q^r)
\right).
\end{align}
For sufficiently small $\alpha$ such that the coordinate-wise angle differences $|\theta_p^r-\theta_q^r|$ are small for all $r$, we use the approximation $1-\cos x\approx x^2/2$. 
Thus,
\begin{align}
\|\boldsymbol{\Phi}_p-\boldsymbol{\Phi}_q\|_2^2
&\approx
\sum_{r=1}^{d_p/2}
(\theta_p^r-\theta_q^r)^2 \\
&=
\|\boldsymbol{\theta}_p-\boldsymbol{\theta}_q\|_2^2 \\
&=
\alpha^2
\frac{2V}{V-1},
\qquad p\neq q.
\end{align}
This shows that Simplex Rotary Agent Encoding guarantees exact equidistance in the agent angle space and yields approximately equal pairwise separation in the complex RoPE space. 
In the common implementation where the centered one-hot simplex is embedded with zero padding and $d_p/2\geq V$, the complex-space distance is also identical across distinct agent pairs because all pairwise difference vectors share the same non-zero coordinate pattern up to permutation.

\section{More Architecture}
\label{app:more-architecture}

\subsection{Action Design}
\label{app:action-design}

We use explicit per-agent action traces as additional conditioning signals.
For both domains, actions are synchronized with the video frames and provided separately for different agents.
The action specification is domain-specific: the game domain uses player control commands, while the robot domain uses continuous end-effector state.

\noindent\textbf{Game action format.}
For Minecraft-style game videos, each agent action at a frame contains $25$ fields: $23$ discrete player controls and $2$ continuous camera controls.
The discrete controls cover inventory interaction, hotbar selection, movement, item manipulation, and mouse-button actions.
The continuous camera controls describe horizontal and vertical view motion.
Table~\ref{tab:game-action-format} summarizes the game action fields.

\begin{table*}[t]
  \centering
  \footnotesize
  \caption{Game action format. Each frame stores a $25$-field action vector for each agent.}
  \label{tab:game-action-format}
  \vspace{4pt}
  \setlength{\tabcolsep}{4pt}
  \renewcommand{\arraystretch}{1.12}
  \begin{tabular}{@{}lll@{}}
    \toprule
    \textbf{Index} & \textbf{Field} & \textbf{Description} \\
    \midrule
    $0$ & \texttt{inventory} & Open inventory \\
    $1$ & \texttt{ESC} & Exit or cancel current menu \\
    $2$--$10$ & \texttt{hotbar.1}--\texttt{hotbar.9} & Select hotbar slot 1--9 \\
    $11$--$14$ & \texttt{forward}, \texttt{back}, \texttt{left}, \texttt{right} & Locomotion controls \\
    $15$--$17$ & \texttt{jump}, \texttt{sneak}, \texttt{sprint} & Movement modifiers \\
    $18$ & \texttt{swapHands} & Swap held item between hands \\
    $19$--$22$ & \texttt{attack}, \texttt{use}, \texttt{pickItem}, \texttt{drop} & Interaction and item manipulation \\
    $23$--$24$ & \texttt{cameraX}, \texttt{cameraY} & Horizontal yaw and vertical pitch motion \\
    \bottomrule
  \end{tabular}
\end{table*}

\noindent\textbf{Robot action format.}
For robot videos, each agent action at a frame contains $10$ continuous fields describing the robot end-effector and gripper.
The action contains the 3D end-effector position, a 6D orientation field, and the gripper opening value.
The same format is used for the left and right robots, producing one temporally aligned action sequence per robot.
Table~\ref{tab:robot-action-format} summarizes the robot action fields.

\begin{table}[t]
  \centering
  \footnotesize
  \caption{Robot action format. Each frame stores a $10$-field continuous action vector for each robot.}
  \label{tab:robot-action-format}
  \vspace{4pt}
  \setlength{\tabcolsep}{4pt}
  \renewcommand{\arraystretch}{1.12}
  \begin{tabular}{@{}lll@{}}
    \toprule
    \textbf{Index} & \textbf{Field} & \textbf{Description} \\
    \midrule
    $0$--$2$ & \texttt{pos\_x}, \texttt{pos\_y}, \texttt{pos\_z} & End-effector position \\
    $3$--$8$ & \texttt{rot\_6d\_0}--\texttt{rot\_6d\_5} & End-effector orientation \\
    $9$ & \texttt{gripper} & Gripper opening value \\
    \bottomrule
  \end{tabular}
\end{table}


\section{Additional Implementation Details}
\label{app:additional_implementation_details}

\noindent\textbf{Architecture.}
Both bidirectional teacher model and causal student model are based on Cosmos-Predict2.5-2B~\cite{ali2025world-simulation},  with hidden dimension $D=2048$, $28$ transformer blocks, $16$ attention heads (head dimension $128$), MLP ratio $4$, and AdaLN-LoRA of rank $256$. Player identity is injected via our Simplex Rotary Agent Encoding (Sec.~\ref{sec:simplex-rotary-agent-encoding}), partitioning the head dimension as $(64, 32, 16, 16)$ across the $(t, p, h, w)$ axes and assigning each agent a vertex of a regular simplex as its rotation phase. We use a simplex pool of size $4$ over $2$ active runtime slots: at every training step we randomly sample $2$ of the $4$ vertices and additionally permute the agent slot order, forcing the model to disambiguate players exclusively through the simplex marker and allowing the same checkpoint to serve up to $4$ players at inference. Per-player actions (a $23$-dimensional keyboard one-hot and a $2$-dimensional camera vector) are encoded by a single shared action encoder: two MLP branches that lift each modality to $128$ dimensions, a fusion MLP that maps the concatenated $256$-dimensional vector through a $4\times$-stride 1D temporal convolution, and a final projection to the model dimension $D=2048$. The encoder output is added as a per-block bias to the self-attention input of every transformer layer. The causal student additionally enables (i) \emph{Sparse Hub Attention}: each player attends only to its own past tokens and to $K=8$ learnable global hub tokens per latent frame; and (ii) \emph{local windowed attention}: each query attends to the most recent $24$ latent frames per view, bounding the inference KV cache independently of generation length. The bidirectional teacher uses standard dense self-attention.

\noindent\textbf{Stages 1 and 2 -- Bidirectional teacher and causal student pretraining.}
Both the teacher and the student are initialized from the publicly released Cosmos-Predict2.5-2B TI2V checkpoint and trained on 2-agent gameplay at $320 \times 480$ per view. The teacher is first pretrained on $93$-frame clips (latent length $24$) for $10{,}000$ iterations and then fine-tuned on $189$-frame clips (latent length $48$) for $6{,}000$ iterations, while the causal student is pretrained on $93$-frame clips for $15{,}000$ iterations. Both stages use AdamW with learning rate $3{\times}10^{-5}$, weight decay $10^{-3}$, $(\beta_1, \beta_2) = (0.0, 0.999)$, $100$-step linear warm-up, and gradient clipping at $0.1$. The teacher is trained on $32$ NVIDIA GB200s and the student on $32$ NVIDIA GB200s. We do not apply classifier-free guidance (CFG) at inference, as we empirically observe that unguided sampling produces more accurate results.

\noindent\textbf{Stage 3 -- Self-Forcing distillation.}
The distillation model couples three networks: the student (trainable, initialized from Stage 2), a frozen \emph{real score} (the Stage-1 teacher), and a trainable \emph{fake score} also initialized from the Stage-1 teacher. We follow Self-Forcing~\cite{huang2025self} to optimize the student model with the DMD~\cite{yin2024one} loss on $189$-frame clips. At each generator step, each block is denoised with timesteps $\{1000, 750, 500, 250\}$ (warped by the flow shift $5.0$). After each block, the model re-forwards the block under context-noise level $128$ and writes the result into the per-layer KV cache before proceeding. Generator and critic are updated alternately at a $1{:}4$ ratio: the student is updated once every $5$ iterations, while on the remaining iterations the fake critic is updated with a flow-velocity prediction loss; the real score is kept frozen throughout. Distillation runs for $400$ iterations on $32$ NVIDIA GB200s, using AdamW with learning rates $2{\times}10^{-6}$ (student) and $4{\times}10^{-7}$ (critic), weight decay $10^{-2}$, $(\beta_1, \beta_2) = (0.0, 0.999)$.

\noindent\textbf{Inference.}
At inference time, the student generates each per-view latent block autoregressively with the same $4$-step denoising schedule and block size as in training. The KV cache uses the rolling local-attention window of $24$ latent frames per view, decoupling the generated sequence length from cache memory.


\section{Additional Ablations}
\label{app:additional-ablations}

We compare three variants of our model: causal, bidirectional, and distilled.
As shown in Table~\ref{tab:model-variant-ablation}, we report FVD, FID, LPIPS, PSNR, and SSIM to evaluate perceptual quality and reconstruction fidelity.

\begin{table}[t]
  \centering
  \footnotesize
  \caption{Comparison of causal, bidirectional, and distilled variants of our model.}
  \label{tab:model-variant-ablation}
  \vspace{4pt}
  \setlength{\tabcolsep}{4pt}
  \renewcommand{\arraystretch}{1.12}
  \begin{tabular}{@{}lccccc@{}}
    \toprule
    \textbf{Variant}
      & FVD $\downarrow$
      & FID $\downarrow$
      & LPIPS $\downarrow$
      & PSNR $\uparrow$
      & SSIM $\uparrow$ \\
    \midrule
    Bidirectional
      & $227.3$ & $31.0$ & $0.272$ & $27.7$ & $0.828$ \\
    Causal
      & $266.4$ & $34.4$ & $0.277$ & $26.2$ & $0.805$ \\
    Distilled
      & $239.7$ & $30.9$ & $0.273$ & $26.8$ & $0.811$ \\
    \bottomrule
  \end{tabular}
\end{table}

We provide an additional ablation on the number of hub tokens $K$ used in Sparse Hub Attention.
In our design, hub tokens serve as the compact shared state through which agents exchange information, so $K$ controls the capacity of the cross-agent communication bottleneck.
When $K$ is too small, the hub has limited capacity to summarize multi-agent state, which can hurt generation quality.
Increasing $K$ provides a richer shared state by allowing the hub to summarize multi-agent interactions with greater capacity.
As shown in Table~\ref{tab:hub-token-ablation}, we sweep $K\in\{1,8,32,128\}$ and report FVD, FID, LPIPS, PSNR, and SSIM.
This ablation helps identify a practical operating point where the hub has enough capacity for cross-agent interaction.

\input{tables/table_4}

\section{More Experimental Results}
\label{app:more-experimental-results}

\subsection{Training Stage Comparison}
\label{app:training-stage-comparison}

We compare three training stages of our model: the bidirectional teacher, the causal student, and the distilled model.
As shown in Table~\ref{tab:model-variant-ablation}, the bidirectional teacher achieves the best overall performance due to its access to full temporal context during generation.
The causal variant, while enabling streaming inference with KV caching, shows degraded performance because it can only attend to past frames.
The distilled model recovers much of the teacher's quality while retaining the causal structure, demonstrating that knowledge distillation effectively transfers bidirectional modeling capacity into a streaming-compatible architecture.
