% \vspace{-3mm}
\section{Introduction}
\label{sec:introduction}
% \vspace{-2mm}
% \begin{center}
%     \emph{``\textbf{Gamma-World} extends generative world modeling from \textbf{It Takes Two} to \textbf{It Takes Many}.''}
% \end{center}

The worlds we wish to simulate are populated, not solitary.
Players cooperate and compete in the same game, robot arms coordinate around shared objects, and embodied agents act under mutual physical and visual constraints.
Controllable multi-agent world modeling is therefore a necessary step toward multiplayer game generation~\cite{bruce2024genie,savva2026solaris}, interactive simulation~\cite{ren2025cosmos-drive-dreams,ali2025world-simulation}, and embodied AI~\cite{fung2025embodied-ai-agents,feng2026multi-agent-embodied-ai}.

Despite this need, most video world models~\cite{alonso2024diffusion,valevski2024diffusion,he2025matrix,xiao2025worldmem,hyworld2025,worldplay2025,yu2025cam} remain single-agent simulators: they roll out future observations conditioned on one action stream, one user input, or one controllable viewpoint.
Moving from single-agent to multi-agent simulation raises a new consistency requirement on top of standard video generation: generated observations must be consistent not only across time, but also across agent perspectives, since all agents share and act upon the same evolving world.

A concurrent effort in this direction is Solaris~\cite{savva2026solaris}, which builds a multiplayer Minecraft world model by combining a dense joint attention block over all agent tokens with a learned per-player ID embedding.
While effective in the two-player setting, this design has two structural limitations.
First, dense joint attention couples every agent token to every other agent token, so its cost grows quadratically with the number of agents, and becomes increasingly restrictive for real-time rollouts beyond two players.
Second, agents in a shared world are intrinsically exchangeable: two agents with identical capabilities should not be treated differently simply because they occupy different slots.
A learned per-slot ID embedding violates this symmetry and ties the model to a fixed player roster that cannot be extended without retraining.
This leaves open the question of whether a video world model can represent multiple agents in a way that is individually controllable, permutation-symmetric, and scalable beyond two players.

We address this question with \ours, a generative multi-agent world model for interactive simulation that scales beyond two players.
\ours extends standard 3D RoPE with an explicit agent axis through \emph{Simplex Rotary Agent Encoding}.
Rather than assigning agents scalar indices or learned identity vectors, we place them at the vertices of a regular simplex in rotary angle space.
Geometrically, this puts all agents at equal pairwise distances, so every pair is permutation-equivalent while each agent retains a distinct rotary phase.
The encoding is parameter-free, does not rely on a fixed learned slot identity, and can be instantiated for different numbers of agents without changing the video transformer architecture.


To complement this symmetric agent representation with efficient cross-agent interaction, we introduce \emph{Sparse Hub Attention}.
Within each causal block, agent tokens attend to their own stream and to a small set of learnable hub tokens.
The hub tokens aggregate information across agents and broadcast it back, providing a shared communication pathway without dense pairwise attention.
This hub-mediated topology preserves a shared communication pathway among agents without dense pairwise interaction, reducing the dominant cross-agent cost from quadratic to linear in the number of agents and substantially relieving the scaling pressure of going beyond two players.
To make the model deployable as a real-time interactive simulator, we further distill a bidirectional multi-agent teacher into a block-causal student with KV caching, enabling 24-FPS streaming autoregressive rollouts that respond to newly issued actions.

We evaluate \ours on multiplayer virtual environments with two and four players spanning movement, mining, combat, and building scenarios.
Across these settings, \ours improves video fidelity, action controllability, and inter-agent consistency over slot-based and dense-attention baselines. 
Ablations validate the contributions of simplex rotary encoding and sparse hub communication. Scaling studies further show that, benefiting from the permutation-symmetric simplex agent encoding, the model generalizes from two to four players without additional training. 

In summary, our main contributions are:
\begin{itemize}
    \setlength{\itemsep}{0pt}
    \setlength{\parsep}{0pt}
    \setlength{\topsep}{0pt}
    \item We identify permutation symmetry as a fundamental property of multi-agent world models and propose \emph{Simplex Rotary Agent Encoding}, a parameter-free rotary identity encoding that preserves permutation symmetry while maintaining distinct agent identities.
    \item We propose \emph{Sparse Hub Attention}, a cross-agent communication mechanism that reduces cross-agent attention cost from quadratic to linear in the number of agents.
    \item We demonstrate effective multi-agent simulation in multiplayer environments with two and four players, with ablations validating the proposed agent encoding and efficiency design.
\end{itemize}
