\input{figures/fig-mobileforge-teaser}
\section{Introduction}
\label{sec:introduction}

MLLM-based mobile GUI agents have made substantial progress in UI understanding and action execution~\cite{gao2026ui,zhou2025mai,hong2024cogagent,xu2026mobile,liu2025llm}. However, real deployment requires adaptation beyond fixed benchmarks. Mobile apps are numerous and fast-changing, making human-written tasks, expert demonstrations, and manual reward labels costly and quickly stale \citep{zhang2025tongui,wang2025mobilea3gent,sun2025genesis,yang2025zerogui}. This motivates annotation-free adaptation, where an agent discovers target-app functions, attempts executable tasks, evaluates behavior, and improves from the experience.

Recent annotation-free GUI agent work has reduced human supervision. TongUI mines supervision from web tutorials \citep{zhang2025tongui}, MobileA3gent uses decentralized user-phone trajectories \citep{wang2025mobilea3gent}, OS-Genesis synthesizes GUI tasks through reverse task construction \citep{sun2025genesis}, and GUI-explorer mines transition-aware interaction knowledge through autonomous exploration \citep{xie2025gui}. ZeroGUI and MobileGUI-RL study automatic reward estimation and online reinforcement learning \citep{yang2025zerogui,shi2025mobilegui}, while SEAgent, ACuRL, and UI-Oceanus explore self-evolving, continual, or synthetic-environment scaling for broader computer-use agents \citep{sun2025seagent,xue2026acurl,wu2026uioceanus}. Appendix~\ref{sec:related} provides a taxonomy.

Despite these advances, two bottlenecks still limit mobile target-app adaptation. \textbf{(I)} Existing methods \textbf{lack a unified mobile substrate} connecting target-app exploration, curriculum mining, rollout execution, and feedback, so generated tasks may be weakly grounded and evaluator feedback may remain detached from policy learning. \textbf{(II)} Policy optimization often treats rollouts as \textbf{isolated experiences with sparse reward}; even with step-level assessment, current loops rarely combine outcomes, process feedback, and corrective hints to accumulate reusable experience beyond the initial policy's capability boundary. Figure~\ref{fig:mobileforge-teaser} illustrates these bottlenecks and our solution.

\obsbox{\textbf{Research question.} Can we build an annotation-free adaptation system for mobile GUI agents that grounds tasks in target-app interaction, generates fine-grained feedback, and converts self-collected experience into policy-improvement signals without human-written tasks, demonstrations, or reward labels?}

We introduce \textbf{\ourmethod}, an annotation-free adaptation system with hierarchical feedback-guided policy optimization. \textbf{\emph{(i)}} \textbf{\mobilegym} is the interaction and evaluation substrate: \curriculum mines executable tasks from target-app traces, while \critic evaluates rollouts with outcome feedback, step-level feedback, and corrective hints. \textbf{\emph{(ii)}} \textbf{\underline{Hi}}erarchical \textbf{\underline{F}}eedback-Guided \textbf{\underline{P}}olicy \textbf{\underline{O}}ptimization (\textbf{\hifpo}) schedules hint-guided attempts, filters tasks and steps with hierarchical feedback, and trains the policy with hint-contextualized step-level GRPO.

We evaluate \ourmethod on AndroidWorld \citep{rawles2024androidworld} as the in-domain setting and MobileWorld GUI-only \citep{kong2026mobileworld} as the out-of-domain setting, with no MobileWorld rollout used for training. As shown in Figure~\ref{fig:main-performance}, annotation-free adaptation narrows the gap to closed-data GUI-specialized agents: \llmname{Qwen3-VL-8B}~\citep{bai2025qwen3} reaches 67.2\% Pass@3 on AndroidWorld using generated adaptation data, close to the \llmname{GUI-Owl-1.5-8B}~\citep{xu2026mobile} base model at 69.0\%. \ourmethod further improves GUI-Owl, yielding \llmname{ForgeOwl-8B} with 77.6\% Pass@3 on AndroidWorld and 41.0\% success on MobileWorld GUI-only, the strongest open-data mobile GUI agent in our evaluation.

\paragraph{Contributions.}
\textbf{(1)} We identify two bottlenecks in annotation-free mobile GUI adaptation: the lack of a unified target-app interaction and evaluation substrate, and isolated rollouts with coarse feedback for policy optimization.
\textbf{(2)} We propose \textbf{\mobilegym}, which grounds exploration, curriculum mining, rollout execution, and hierarchical evaluation in real mobile app interaction.
\textbf{(3)} We propose \textbf{\underline{Hi}}erarchical \textbf{\underline{F}}eedback-Guided \textbf{\underline{P}}olicy \textbf{\underline{O}}ptimization (\textbf{\hifpo}), which transforms multi-attempt feedback and corrective hints into hint-contextualized step-level GRPO updates.
\textbf{(4)} We show that \ourmethod improves generalist and GUI-specialized agents, transfers from AndroidWorld to MobileWorld GUI-only, and yields \llmname{ForgeOwl-8B}, the strongest open-data mobile GUI agent in our evaluation.
