\section{Detailed Related Work}
\label{sec:related}

This appendix expands the concise related-work discussion in Section~\ref{sec:introduction}. We organize prior work around the two bottlenecks emphasized in the main paper: the need for a mobile interaction and evaluation substrate, and the need to convert rollout feedback into effective policy optimization signals.

\paragraph{External GUI data construction and trajectory synthesis.}
Several recent systems reduce manual annotation by constructing GUI supervision from external or automatically synthesized sources. TongUI mines multimodal web tutorials and converts them into generalized GUI trajectories across platforms and applications \citep{zhang2025tongui}. MobileA3gent instead uses decentralized user-phone trajectories, automatic intent annotation, and federated training to exploit self-sourced mobile data while preserving user privacy \citep{wang2025mobilea3gent}. OS-Genesis reverses the usual task-first pipeline: it first explores GUI transitions, synthesizes instructions from observed state changes, and filters constructed trajectories with a trajectory reward model \citep{sun2025genesis}. These works substantially reduce the cost of obtaining GUI training data. However, they mainly address data construction. Their supervision is not necessarily grounded in the current state of a target mobile app, and they do not directly connect an agent's own rollout failures to step-level policy improvement.

\paragraph{Autonomous exploration and knowledge mining.}
GUI-explorer is closely related to the exploration side of \ourmethod. It autonomously explores target apps, mines transition-aware knowledge from observation-action-outcome triples, and retrieves that knowledge at inference time to improve GUI grounding and decision making \citep{xie2025gui}. This shows that direct target-app exploration can expose app-specific affordances that static model knowledge misses. The key distinction is the role of the explored knowledge. GUI-explorer is training-free and uses mined knowledge as prompt-side guidance, whereas \ourmethod uses target-app exploration to build an executable curriculum and then continues through rollout execution, hierarchical evaluation, and policy optimization.

\paragraph{Online GUI reinforcement learning.}
ZeroGUI moves GUI learning toward zero-human-cost online RL by automatically generating tasks, estimating trajectory-level success with a VLM evaluator, and updating the GUI policy through online reinforcement learning \citep{yang2025zerogui}. MobileGUI-RL further studies online mobile GUI RL with self-exploration, task filtering, trajectory-aware advantages, and composite rewards that combine success and execution efficiency \citep{shi2025mobilegui}. These works demonstrate that GUI agents can improve through interaction rather than relying only on offline demonstrations. Their optimization signals, however, remain centered on trajectory-level or composite rewards. Such feedback is useful for deciding whether a rollout is good overall, but it provides limited credit assignment for long-horizon mobile tasks where a failed trajectory may still contain useful local decisions and a successful trajectory may include redundant or accidental actions. \ourmethod addresses this by using trajectory outcome feedback, step-level process feedback, and corrective hints for task filtering, trajectory selection, step extraction, and hint-contextualized GRPO.

\paragraph{Self-evolving and continual computer-use agents.}
SEAgent and ACuRL share \ourmethod's motivation that computer-use agents should adapt from their own experience. SEAgent studies self-evolving computer-use agents with a world-state model, curriculum generation, step-wise assessment, guidebook-style experience accumulation, and policy improvement through GRPO and imitation-style objectives \citep{sun2025seagent}. ACuRL formulates environment adaptation as autonomous continual learning: the agent explores an environment, generates curriculum tasks, obtains automatic evaluator feedback, and improves over continual training rounds \citep{xue2026acurl}. These systems are broader computer-use frameworks, primarily designed around desktop or general digital environments. \ourmethod is mobile-specific: \mobilegym provides target mobile app interaction, curriculum mining, rollout execution, and hierarchical mobile rollout evaluation, while \hifpo schedules multi-attempt rollout, reuses corrective hints across attempts, and performs hint-contextualized step-level GRPO.

\paragraph{Synthetic GUI environmental dynamics.}
UI-Oceanus scales GUI agents by synthesizing environmental dynamics and training on transition-oriented objectives \citep{wu2026uioceanus}. This line improves the coverage of GUI state transitions without requiring every transition to be manually collected in the target environment. It is complementary to \ourmethod: synthetic dynamics can broaden pretraining or supervised training data, while \ourmethod focuses on annotation-free adaptation inside target mobile apps, where the agent must discover executable tasks, evaluate its own rollouts, and update from the resulting experience.

\paragraph{Positioning of \ourmethod.}
Taken together, prior work has made GUI learning less dependent on human-written demonstrations, but two bottlenecks remain for mobile target-app adaptation. External data construction provides scalable supervision but may not reflect the current target app. Exploration methods expose app-specific knowledge but often stop at inference-time guidance. Online RL methods update the policy but usually rely on coarse trajectory feedback. Continual computer-use methods study self-evolution but are not designed as mobile-specific interaction and evaluation substrates. \ourmethod addresses these gaps by separating the system into two roles. \mobilegym supplies the mobile substrate: target-app exploration, executable curriculum mining, rollout execution, and hierarchical feedback. \hifpo supplies the feedback-to-optimization mechanism: it schedules multi-attempt rollout, accumulates corrective hints, filters mastered tasks, selects informative trajectories and steps, and optimizes the policy with hint-contextualized step-level GRPO. Table~\ref{tab:related_work_comparison} summarizes this comparison.

\input{tables/table-related-work-comparison}
