\subsubsection{Action}
\label{subsubsec::data_gen_action}

Actions provide the causal variables that connect observed world states across time. While video-only training teaches the generator to extrapolate likely motion, it does not expose the model to controllable interventions: the same initial observation may evolve differently under different robot commands, camera trajectories, vehicle routes, or human hand motions. We therefore introduce paired text-video-action data during mid-training so that Cosmos 3 can learn both directions of the world-action relationship: predicting future observations conditioned on actions, inferring the actions that explain an observed trajectory, and jointly generating actions and future video.

\paragraph{Data statistics.} We focus action mid-training on four physical-AI pillars: egocentric motion, robotics, autonomous vehicles, and camera motion. The final curated data contains $8.4$M episodes and $61.3$K hours across these pillars, as summarized in \cref{fig:action_data_distribution}.

\input{figures/data/action/action_data_distribution}

\begin{itemize}
  \item \textit{Egocentric motion.} Egocentric motion data contributes $41.3$K hours ($67.4\%$), making it the largest component. It comprises $1.7$M episodes from a proprietary dataset of bimanual hand manipulation captured with a head-mounted RGB camera.   Each frame is annotated with the synchronized head-camera pose and, for each hand, a 21-keypoint 3D pose~\citep{zimmermann2017learning,simon2017hand} that provides per-joint position and orientation in the camera coordinate frame, enabling the model to jointly learn egocentric ego-motion and fine-grained dexterous hand motion.
  \item \textit{Autonomous vehicle.} Autonomous vehicle data contributes $10.0$K hours ($16.3\%$), derived from high-quality, in-house driving logs collected using the NVIDIA Hyperion platform. The dataset is constructed by mining a large-scale corpus to match a target distribution spanning diverse driving scenarios. The selected scenarios cover a broad range of conditions, including diverse weather, lighting, and road conditions, as well as varied longitudinal and lateral maneuvers, rather than being limited to predominantly near-straight cruising. To align with other domains, we transform driving trajectories from the vehicle coordinate frame to the front-wide camera coordinate frame.
  \item \textit{Robotics.} Robotics data contributes $5.4$K hours ($8.7\%$), aggregated from open-source datasets. The subset contains $90.4$K tasks and $516.7$K episodes, as broken down by embodiment and source in \cref{tab:robotics_breakdown}. To avoid embodiment-specific controller details such as PID parameters or low-level actuation interfaces, we use pseudo-actions derived from state differences. We curate data from both successful and failed episodes so the model observes not only intended completions but also off-nominal action effects.
  \item \textit{Camera motion.} Camera motion data contributes $4.6$K hours ($7.5\%$), mined from our pre-training video dataset. We convert these videos into action trajectories by estimating camera poses with ViPE~\citep{huang2025vipe} and DepthAnything3~\citep{lin2025depthanything3}. To ensure data quality, we rigorously filter the dataset to remove clips with unreliable pose estimation, such as those exhibiting excessive jitter or abnormal camera intrinsics. All camera poses are kept in metric scale and converted to the unified action coordinate convention.
  This curation process yields a dataset of $1.9$M clips.
\end{itemize}

\paragraph{Data processing pipeline.}
We convert each source using the unified action tokenization described in \cref{sec:action}. To balance action magnitudes across embodiments after this conversion, we compute per-dimension normalizers from the training data and scale action channels to a comparable range of roughly $[-1,1]$. For data with multiple synchronized viewpoints, we concatenate the views into a canvas and store the camera layout in metadata, as shown in \cref{fig:action_multiview_packaging}. Rather than filtering out idle operations, we retain them and record the idle-step count in metadata, allowing downstream sampling to explicitly balance active and inactive segments.
