\subsubsection{Audio}
\label{subsubsec::data_gen_audio}

Audio-video paired data teaches the Generator not only what sound should be present, but also when that sound should occur relative to visible events. Raw web video audio is challenging for this purpose: narration and voiceover often describe the video without being caused by it, while manually added background music (BGM) can mask the physical sounds produced by on-screen events. We therefore use audio differently across training stages. Pre-training preserves broad acoustic coverage, while mid-training constructs higher-precision audio-video pairs through an explicit selection policy for speech and non-speech audio.

\paragraph{Pre-training.}
The audio pre-training corpus is derived entirely from the pre-training video pool. In total, 138.9M pre-training clips contain usable audio tracks, covering a broad mixture of diegetic and non-diegetic speech, voiceover, BGM, ambient sound, music, and physical events. Of these clips, 62.5M are shorter than 30 seconds; for this subset, we use Qwen3-Omni-Captioner~\citep{xu2025qwen3omni} to generate synthetic audio descriptions. This stage favors scale and diversity, exposing the model to the long-tailed distribution of real video audio before applying stricter curation in mid-training.

\paragraph{Mid-training.}
The mid-training audio pool is filtered from the pre-training audio-video corpus to improve causal audio-visual alignment. The final pool contains 18.8M clips: 12.8M non-speech clips for environmental and physical sound generation, and 6M speech-synchronized clips for visually grounded speech generation. The curation pipeline is organized around a simple principle: keep speech only when it is synchronized with a visible face, remove off-screen speech from non-speech examples, and remove non-instrumental BGM when it would dominate the target audio.

\begin{itemize}
    \item \textit{Source separation.} For every candidate clip, SAM-Audio~\citep{shi2025samaudio} separates the original audio into a speech stem and a remaining stem. The speech stem is used to identify visually grounded speech, while the remaining stem provides a vocal-suppressed candidate for non-speech audio generation.
    \item \textit{Lip-sync scoring.} SyncNet~\citep{chung2016out} is run on the speech stem and original video to produce \texttt{has\_face} and \texttt{lip\_sync\_confidence}. We define \texttt{speech\_synced} as \texttt{has\_face} $=$ \texttt{True} and \texttt{lip\_sync\_confidence} $\geq 3.0$ (as shown in LatentSync~\citep{li2024latentsync}).
    \item \textit{Audio event detection.} FireRedASR2S~\citep{xu2026fireredasr2s} runs on the original audio to estimate \texttt{speech\_ratio}, the fraction of time labeled as speech or singing, and \texttt{music\_ratio}, the fraction labeled as music. We define \texttt{high\_music} as \texttt{music\_ratio} $\geq 0.1$.
    \item \textit{Instrument detection.} For all \texttt{high\_music} clips, including speech-synchronized ones, Qwen3-VL~\citep{qwen3vl2025} predicts \texttt{is\_music\_instrument}. This protects instrument-performance videos whose music is part of the visible event rather than removable BGM.
    \item \textit{Speech branch.} Clips satisfying \texttt{speech\_synced} form the speech mid-training pool. We keep the original audio unless the clip is \texttt{high\_music} and \texttt{is\_music\_instrument} is \texttt{False}; in that case, SAM-Audio removes music from the original waveform. This edited path is retained only if a second FireRedASR2S pass on the candidate audio reports \texttt{speech\_ratio} $\geq 0.05$ and \texttt{music\_ratio} $= 0$, and only if the candidate is not near-silent according to \texttt{max\_abs} $\geq 0.007$, \texttt{p50\_db} $\geq -80$, and \texttt{active\_ratio} $\geq 0.2$. This preserves lip-synchronized speech while suppressing non-instrumental BGM.
    \item \textit{Non-speech branch.} Clips not assigned to the speech-synchronized branch are curated for non-speech audio-visual generation. We first choose a base waveform: if \texttt{speech\_ratio} $\geq 0.05$, we use the SAM-Audio remaining stem to remove vocals; otherwise, we keep the original audio to preserve physical sounds. If the clip is \texttt{high\_music} and \texttt{is\_music\_instrument} is \texttt{False}, SAM-Audio music removal is applied to the chosen base waveform. Candidates derived from the remaining stem must pass a second FireRedASR2S check with \texttt{speech\_ratio} $< 0.05$; their \texttt{music\_ratio} must be $= 0$ when music removal was applied, and $< 0.1$ otherwise. Candidates derived from the original waveform with music removal must have \texttt{music\_ratio} $= 0$. Processed candidates must also pass the same non-silence thresholds used for the speech branch.
    \item \textit{Caption annotation.} Captions are tied to the final waveform used for training. If the selected waveform is the original audio, we retain the original audio description. If source separation or music removal changes the waveform, we re-caption the final candidate audio so that the text does not describe removed speech or accompaniment. For the speech-synchronized pool, we transcribe each selected speech track with Qwen3-ASR~\citep{shi2026qwen3asr}, then use GPT-OSS-120B~\citep{agarwal2025gptoss} to merge the transcript with the audio description. The resulting caption specifies both spoken content and non-linguistic acoustic context while remaining faithful to the audio paired with the video.
\end{itemize}