\subsection{Reasoner Data}
\label{subsec::data_reasoner}

Our reasoner data curriculum contains approximately $24.2$M samples: $22.0$M for pre-training and $2.2$M for supervised fine-tuning from domain-specific Physical AI datasets and synthetically generated data. \Cref{tab:data_reas_summary} summarizes the data modalities used in both stages. The pre-training stage is dominated by image--text and text-only data, providing broad general visual understanding. In contrast, the supervised fine-tuning stage shifts toward Physical AI specialization, with video--text samples comprising 50\% of the mixture to strengthen spatiotemporal understanding and capabilities in robotics, smart infrastructure, and autonomous vehicle domains. \Cref{fig:reasoner_category_mix} summarizes this mixture by capability category for the two stages.  

\input{figures/data/reasoner/reasoner_stats}
\input{tables/data/reasoner/data_stats}

\subsubsection{Pre-Training}
\label{subsubsec::data_reas_pretrain}

We build our pre-training data mixture with 19.7M samples sub-selected from the Nemotron Nano 2 data collection~\citep{nvidia2025nvidianemotronnanov2} and 2.3M additional samples curated to enhance math, video, spatial grounding, and instruction-following capabilities. See~\cref{tab:data_reas_summary} for details.
The datasets we source are fed through a two-stage data curation pipeline consisting of semantic deduplication followed by AI-judge quality filtering before inclusion in the final training mixture.

\paragraph{Semantic deduplication.}
The first stage removes multimodal near-duplicates at the conversation level, where a conversation denotes the complete training example: an image or video paired with its instruction-response text, or a text-only instruction-response sample when no media is present. For each conversation, we compute a joint embedding that combines the media representation, when available, with the associated instruction-response text representation. Image-text and text-only conversations are embedded using Qwen3-VL-Embedding-8B~\citep{qwen3vlembedding}, while video-text conversations are embedded using the Perception Encoder PE-Core-G14-448~\citep{bolya2025PerceptionEncoder}. The resulting concatenated embedding representation jointly captures visual and linguistic semantics, enabling the pipeline to distinguish visually similar samples with different task intents from truly redundant supervision. 

To scale duplicate detection to production-scale datasets, we use clustering. Image-text, video-text and text-only samples are first partitioned using K-means clustering. Near-duplicate groups are then identified within each cluster using cosine similarity in the conversation-embedding space. Samples with similarity above a high threshold of $0.95$ are removed. This hierarchical design makes large-scale duplicate detection tractable while preserving sensitivity to redundancy in both visual content and task semantics.

\paragraph{AI-judge quality filtering.}
The second stage applies an AI judge to assess annotation quality on the deduplicated corpus. We use Gemma-4 as the vision-language judge~\citep{gemma4modelcard}, specifically the Gemma-4-31B-it model~\citep{gemma4hf31bit}. The judge is prompted as a training-data auditor and assigns rubric-based integer scores from $1$ to $5$ across three primary quality dimensions:

\begin{itemize}
    \item \textit{Faithfulness:} whether all response claims are grounded in the provided image, video, or textual context.
    \item \textit{Completeness:} whether the response fully addresses the instruction without important omissions.
    \item \textit{Correctness:} whether the response is factually, logically, and task-level accurate.
\end{itemize}

Faithfulness is particularly important for Cosmos 3 because unsupported visual claims can teach the model to hallucinate physical states, object attributes, or temporal events. Meanwhile, Completeness filters under-specified or partial responses, while Correctness removes supervision whose final answer or reasoning is inconsistent with the input. In addition to scalar scores, the judge also produces short evidence-based rationales, enabling targeted spot checks and auditing of the filtering behavior.

\paragraph{Threshold-based dataset construction.}
We construct multiple judge-filtered dataset variants from the same deduplicated base corpus using a minimum-threshold rule over the three quality dimensions. A sample is retained only if its Completeness, Correctness, and Faithfulness scores all meet or exceed a specified threshold. In other words, every retained example must simultaneously satisfy the minimum quality requirement across all three dimensions.

This filtering strategy is intentionally stricter than averaging the scores. Samples with a severe failure mode in any single dimension are removed even if they score highly on the remaining criteria. For example, a response that is highly detailed and logically correct but contains unsupported visual claims will still be filtered out due to low Faithfulness. In practice, we use conservative thresholding to eliminate clearly low-quality supervision while minimizing excessive distribution shift across capability domains.

Multimodal deduplication removes $4.23\%$ of the data as near-duplicate supervision. The AI-judge score distribution reveals that quality failures are not uniform across dimensions. We also analyze retention by capability category to ensure that quality filtering improves the corpus without unintentionally collapsing the skill distribution. At stricter thresholds, pruning becomes strongly category-selective: for example, referring-expression grounding is removed most aggressively, while image captioning and visual question answering also decline substantially, as the threshold increases from $2$ to $5$. Here, the threshold denotes the minimum acceptable AI-judge score on each quality dimension: a sample is retained only if its Completeness, Correctness, and Faithfulness scores are all above the threshold. OCR data is comparatively robust with a threshold of $2$ but drops materially at $5$. These trends motivate using the lowest non-trivial judge threshold, \ie, $2$ for the primary mixture: it filters clear annotation failures while preserving the original coverage of reasoning, grounding, OCR, captioning, and VQA capabilities, thereby creating an optimal quality--quantity trade-off on the pre-training dataset. However, in the SFT stage, we use a threshold of $5$ to retain only the highest-confidence supervision examples, where annotation precision and response reliability are more critical than broad coverage. The AI-judge filter retains $78\%$ and $46\%$ of data at a threshold of $2$ and $5$, respectively.

The final pre-training mixture contains approximately 22M samples spanning OCR, grounding, question answering, reasoning, captioning, and instruction-following data, as shown in~\cref{tab:data_reas_summary}. Regarding the composition, OCR is the largest component, contributing $9.44$M samples ($42.9\%$), followed by 2D grounding with $3.62$M samples ($16.5\%$), visual QA with $2.48$M samples ($11.3\%$), and image reasoning with $1.66$M samples ($7.5\%$). The remaining mixture provides broad multimodal coverage through text QA ($1.35$M, $6.1\%$), image captioning ($1.30$M, $5.9\%$), video QA ($0.99$M, $4.5\%$), text instruction data ($0.82$M, $3.7\%$), visual instruction data ($0.34$M, $1.5\%$), and small amounts of video captioning and video reasoning data ($0.01$M each). This composition emphasizes strong image-text alignment, reading, and spatial grounding while retaining a lightweight video component that prepares the model for later supervised fine-tuning on temporal and video-reasoning tasks.


\subsubsection{Supervised Fine-Tuning}
\label{subsubsec::data_reas_posttrain}

In the supervised fine-tuning stage, we enhance the general spatial and temporal understanding capabilities and focus on curating data for the following three domains: autonomous vehicle, robotics, and smart infrastructure. In total, we train with 2.2M samples in this stage.

\paragraph{General spatial understanding.}
We enhance general spatial understanding through 2D and 3D grounding, augmented with both real and simulation data.

\textit{2D and 3D grounding.} 2D grounding supports Physical AI tasks that require localization (objects, regions, parts, and landmarks), pointing, counting, and multi-image correspondences. We curate samples spanning detection, referring expressions, OCR/layout, visual-prompted description and VQA, pointing, trajectories, counting, and dense grounding. Boxes and points are normalized and converted into a unified JSON format. We curate 2D grounding data from synthetic and existing data, with a majority coming from LocateAnything~\citep{wang2025locateanything} data, and we sample a high-quality subset from them. For 3D grounding data, we convert 3D scanned scenes into instruction-following training samples with camera-relative 3D boxes. Each object is annotated with label, center, dimensions, and orientation after canonicalization and intrinsic normalization~\citep{brazil2023omni3d}. Instead of chaining 2D grounding and 3D inference~\citep{gr3d,man2025locateanything3d}, we directly supervise the final structured boxes.

\textit{Real-world spatial understanding and grounding.} We combine several complementary QA types. Image-referring QA task requires pointing to a target object or to empty free space as normalized image coordinates given a natural-language expression~\citep{zhou2025roborefer}. Robotics spatial QA requires predicting placement points in free space and answers binary relative-position and reachability questions across ego-, world-, and object-centric reference frames~\citep{song2025robospatial}. We further include posed-scene multi-image QA, video-frame spatial QA, and multiple-choice questions that predict the relation between two objects from a fixed set (\eg, left of, behind, on/above), including perspective-substituted viewpoints~\citep{liu-etal-2025-multimodal-large}. Together, these cover object references, free space, cross-view correspondence, camera motion, size, distance, direction, routes, counting, and room-scale reasoning.

\textit{Simulator-grounded embodied spatial reasoning.} We bridge visual spatial QA and embodied action by deriving labels from executable simulator state. 
Answers are computed from cameras, object poses/boxes, depth, masks, visibility, and feasible regions, then checked by programmatic and VLM critics. The curriculum spans metric geometry, spatial frame, physical semantics, actionable grounding, viewpoint dynamics, and embodied composition, with multiple-choice question, numerical, point, box, binary, and text outputs.

\paragraph{General temporal understanding.}
We augment temporal capabilities along three axes: temporal event understanding, physical plausibility judgment, and structured spatiotemporal scene upsampling.

\textit{Temporal event understanding.}
We strengthen temporal and motion understanding with three complementary supervised data sources. First, human annotators create dense temporal captions: egocentric videos of everyday indoor tasks are labeled with atomic human-action descriptions and start/end timestamps, with actions averaging 1.8 seconds and 14.2 words. In addition, a broader video corpus of 55K videos (2.6K hours) provides 743K event triplets $(t_{\text{start}}, t_{\text{end}}, \text{caption})$ for event enumeration and query-conditioned localization. Second, to increase question diversity, we curate training data with the FoundationMotion pipeline~\citep{gan2025foundationmotion}, producing ten four-way multiple-choice questions per clip that probe action identity, temporal evolution, and fine-grained motion differences. Finally, we annotate camera motion patterns such as panning and zooming so the model can learn ego-camera movement.

\textit{Physical plausibility judgment.}
We strengthen Cosmos 3 Reasoner's ability to judge physical plausibility in generated videos with two complementary supervised sources. 
First, we incorporate annotations from the Cosmos human evaluation described in Appendix~\ref{sec::cosmos_hue_bench}, which contains 13.5K human-graded $(\text{video}, \text{question}, \text{answer})$ tuples from 1K generated videos, covering visual integrity, temporal stability, geometry, anatomy, motion plausibility, and physical commonsense with categorical Yes/No/Unclear answers. Second, we adopt VideoPhy-2~\citep{bansal2025videophy}, providing 3.4K action-centric videos spanning $200$ actions, each rated from $1$--$5$ for adherence to physical laws; its annotations cover conservation laws, gravity, collision dynamics, temporal causality, and spatial constraints, and are converted into supervised QA pairs where the model predicts a physics-adherence score. 

\textit{Structured spatiotemporal scene upsampling.} 
We curate paired data of under-specified user inputs and densely-structured captions. The input  may be a text prompt, an image, or both, and the target is the paired structured-caption annotation described in \Cref{subsubsection::imagevideo}. The Upsampler capability recovers spatial, temporal, and visual details that are implicit in the input, such as subject attributes, scene layout, camera framing, object interactions, and plausible future motion. To make the model robust to different request formats, we synthesize various instruction variants, with one variant being the canonical form, sampled more often, and by crossing different input-prompt lengths (\eg, how detailed a description should be; brief, detailed, \etc)  with several prompting styles (\eg, how the description should detail the scene; request, declarative, \etc). The resulting supervision encourages the model to follow compact, direct, and stylized user requests and generate detailed scene descriptions while preserving the same structured output contract. \Cref{sec::prompt_upsampling} describes using the capability for Cosmos 3 image/video generation.

\paragraph{Autonomous vehicle (AV).} We incorporate AV datasets spanning human-labeled and auto-labeled chain-of-thought reasoning, temporal event understanding, and 3D vehicle grounding.

\textit{Action CoT.} Human-labeled CoT data from internal driving logs contains more than 10K videos with explicit driving decisions, covering weather, lighting, road conditions, traffic rules, ego-vehicle behaviors, critical objects, and causal links between scene elements and ego behavior. To scale this signal, we auto-label about 1.1M additional decision-rich videos from internal logs. For each video, we identify the meta-action transition keyframe as the decision moment and use state-of-the-art VLMs, raw video, ego trajectory, dynamic states, and meta actions to produce structured decisions, critical components, and concise reasoning traces.

\textit{Temporal event localization.} We derive temporal event localization data from Nexar dashcam footage~\citep{dmoura2025nexar}, with more than 24K videos covering collisions, near-collisions, hard braking, harsh acceleration, and sharp cornering. Clips are sampled at 6 FPS and capped at 300 seconds. We augment human-/auto-labeled data with dense captions of the scene, agents, interactions, ego behavior, and spatiotemporal context.

\textit{3D vehicle grounding.} We train metric 3D vehicle grounding from the MADS dataset~\citep{ren2025cosmos}, which provides synchronized multi-camera sequences, world-scenario-map 3D annotations, camera intrinsics/extrinsics, and ego poses. We sample frames at ${\sim}1$\,FPS and create open-vocabulary detection and referring-grounding QA, asking the model to enumerate vehicles or localize instances by relative position, lane, motion state, or distance. Answers are camera-frame 3D boxes parameterized by position, size, roll, pitch, yaw, and category label, filtered to visible, lightly occluded objects within $100$\,m across diverse regions, lighting, and weather.

\paragraph{Robotics and embodied AI.} We curate data for action CoT in robot manipulation, embodied reasoning, and healthcare robotic surgery understanding.

\textit{Robot action Chain-of-Thought.} 
Action-CoT teaches Cosmos 3 Reasoner to turn a high-level embodied instruction and current frame into a 2D image-plane motion plan for robot end-effector control. Instead of free-form rationales, it structures reasoning through task-relevant locations, grounding points, move reasoning, and 2D waypoints, moving from perception to action in a compact, inspectable trace. The trace identifies manipulated objects, context objects, affordance points, and collision-free regions, localizes them as coordinates, and resolves the plan into pixel-space end-effector waypoints. Data is built with a modular pipeline because grounding, localization, and manipulation planning require different skills: Qwen3-VL-72B-Instruct~\citep{bai2025qwen3} generates grounding rationales and move reasoning, Molmo-7B~\citep{deitke2024molmo} localizes referring expressions, and motion-plan targets come from MolmoAct~\citep{lee2025molmoact} or tracked DROID~\citep{khazatsky2024droid} episodes. 

\textit{Embodied reasoning.} 
We strengthen Cosmos for embodied reasoning with temporal localization, task planning, and robotics embodied QA data. For robot manipulation, we target the MimicGen~\citep{mandlekar2023mimicgen} bottleneck of segmenting demonstrations into object-centric subtasks with precise timestamp boundaries, using 60 held-out videos for zero-shot evaluation and a supervised fine-tuning set of 3.6K Omniverse-rerendered videos across six tasks, with timestamps derived from object trajectories and joint kinematics and manually verified. For long-horizon planning, we curate 83K BEHAVIOR-1K~\citep{li2023behavior1k} samples that map a scene frame and candidate action list to the ground-truth action. We also add ERQA robotics QA from EO-Data-1.5M~\citep{eo1}, covering task planning, affordances, failure detection, physical commonsense, localization, referring, relations, and trajectory prediction.

\textit{Healthcare robotic surgery understanding.}
We curate a robotic-assisted surgery VQA dataset with 398K multi-turn conversations over 2.2M images. The data is collected from exocentric operating-room cameras, an egocentric robotic detail camera, and console displays inspired by ORQA~\citep{ozsoy2025specializedfoundationmodelsintelligent}. Most samples combine multiple viewpoints and include tracker metadata (tool states, 3D translations, Euler rotations) and robot metadata (surgical phase and step) as contextual hints. The tasks cover tool recognition, localization, yes/no classification, scene graphs, monitor-text transcription, personnel counting, distance/time estimation, action labeling, and surgical-step recognition.

\paragraph{Smart infrastructure.}
We curate three complementary smart-infrastructure data sources, covering warehouse spatial intelligence, dense pedestrian localization, and traffic and anomaly reasoning. 

\textit{Warehouse spatial intelligence.} We use PhysicalAI-Spatial-Intelligence-Warehouse~\citep{tang20259thaicitychallenge}, a synthetic Omniverse corpus spanning 44 warehouse scene collections and 40 camera views. From 93K RGB--D images and 873K QA pairs, we subsample 80K balanced examples covering object counting, metric distance, grounding, and binary spatial relations over pallets, boxes, forklifts, shelves, and operator zones.

\textit{Dense pedestrian localization.} We curate annotations with 208K images from 44 scenes and 5.6M manually labeled person boxes. All person bounding boxes are manually labeled by human annotators, and personally identifiable information is redacted via blurring prior to annotation and release to ensure subject anonymity.

\textit{Traffic and anomaly reasoning.} We combine synthetic ITS collision supervision, real traffic-event reasoning, and surveillance anomaly verification. CARLA~\citep{Dosovitskiy17} clips train binary collision prediction between marked vehicle pairs in unprotected-left-turn and T-bone scenarios, with Cosmos-Transfer2.5~\citep{cosmos_predict2p5} augmentation yielding 3.4K labeled pair queries. TAR~\citep{nvidia2026tar} contributes 3{,}6K traffic-camera videos (26 hours) and 44K annotations spanning QA, temporal reasoning, causal linkage, scene description, and summarization, with hierarchical CoT auto-labels cross-checked against human annotations. To broaden anomaly coverage beyond traffic, we also curate 1K internal surveillance clips for binary tailgating verification.