\section{Introduction}

\paragraph{Background and Demand.}
Recent video generation models, such as Sora, Veo, Wan, and Seedance~\citep{videoworldsimulators2024,veo,wan2025wan,seedance2026seedance}, have shown strong ability in generating high-quality videos. However, most of them still follow an offline, one-shot generation paradigm: a user enters a prompt, waits for minutes or even tens of minutes, and receives the complete video only after generation finishes. 
During this process, the user can only passively wait to receive information and cannot actively initiate any interaction.
This limitation comes from the offline diffusion paradigm, where the model denoises the entire video synchronously over many steps and only produces an entire clean video at the end. Such a paradigm works well for offline content creation, but human visual entertainment is not limited to pre-generated videos. People also enjoy face-to-face communication, live streaming, games, talking with someone, and other interactive visual experiences, where content must respond immediately to the user. From a demand perspective, suppose that each user has an average demand of $\alpha \in [0,1]$ for real-time interactive visual content, e.g., $\alpha = 0.5$. Then the total demand scales with $\alpha \times N$, where $N$ is the number of users. In contrast, suppose that each user has an average demand of $\beta \in [0,1]$ for offline-generated visual content, e.g., $\beta = 0.5$. Since offline-generated videos can be replayed and shared, their generation demand scales more like $\beta \times N / m$, where $m$ is the average number of views per generated video. If we assume $\alpha \approx \beta$ and $m > 100$, then the demand for real-time interactive video generation is much greater than that for offline-generated videos.


\begin{center}
\textbf{Real-time interactive video generation is one of the most important future directions for industry.}
\end{center}
\vspace{-6pt}

\paragraph{Limitation.}
Despite the importance of this direction, existing autoregressive video generation models still have several key limitations. (1) Many methods~\citep{DF,kim2024fifo,henschel2025streamingt2v,song2025history} only change the generation process from offline generation to autoregressive generation, but they still do not support real-time user interaction during generation. (2) Most methods~\citep{liu2025rolling,sun2025ar,xie2025progressive,feng2025streamdiffusionv2} do not use speech as a direct and explicit control signal for future video content, making it difficult for users to control the generation process through natural spoken instructions. (3) Long-horizon video generation remains challenging because small errors can accumulate over time, causing drift, instability, and eventually visual collapse~\citep{yin2025slow,huang2026self}. As a result, many models~\citep{huang2026self,liu2025rolling,yin2025slow,zhu2026causal,zhao2026causal,yang2025longlive,teng2025magi,chen2025skyreels,zhang2025packing,gao2025wan} can only generate videos within a limited duration and cannot maintain stable generation over an open-ended stream. (4) Real-time video generation is not only a modeling problem. It also requires strong inference kernels, serving systems, scheduling strategies, and large-scale deployment infrastructure. Without such infrastructure, even a capable model may be too slow or too expensive for practical use.

\vspace{-1pt}
\paragraph{Vidu S1.}
To address these limitations, we introduce \ours, a real-time interactive video generation model designed for continuous user interaction. 
(1)\ours allows users to actively interact with the generation process at any moment, intervening in real time instead of merely specifying all controls before generation begins.
(2) \ours treats user speech as a direct and explicit control signal, so spoken instructions can continuously guide what the video should generate next.
(3) \ours supports stable long-horizon real-time generation by mitigating error accumulation during streaming generation, allowing the video to continue indefinitely without rapid drift or collapse.
(4) It is also built together with an efficient inference and serving stack, including TurboDiffusion~\citep{zhang2025turbodiffusion} and TurboServe~\citep{jiang2026turboserve}, which enables low-cost real-time inference on commodity GPUs at 540p resolution and 42 FPS. 

\vspace{-3pt}
\paragraph{Contribution.}
We summarize our contributions as follows.

\begin{enumerate}
\item We introduce \ours, a real-time interactive video generation model that supports continuous user interaction and enables explicit speech control over future video content during generation.

\item \ours supports indefinitely long video generation with high quality and consistency. By mitigating error accumulation during streaming, it allows video streams to run without drift or collapse.

\item \ours achieves real-time video generation at 540p resolution and 42 FPS. With an efficient inference and serving stack, \ours makes high-quality real-time generation practical on low-cost GPUs.

\item Experiments on Vidu-StreamBench and HDTF demonstrate that \ours combines leading quality and real-time interactivity, achieving the best CSIM (0.9192), Sync-D (7.847), and DOVER (0.5660), strong human preference while supporting speech-guided control and stable long-horizon streaming.

\end{enumerate}


