\section{Results}
\label{sec:results}

\subsection{Per-axiom analysis}
\label{sec:results:per-axiom}

\input{tables/headline_causality}
\textbf{Causality.} We measure the divergence between the source LLM's continuation distribution and the distribution induced by substituting a candidate TR, reporting the KL divergence (\cref{eq:causality-error}) for which lower is better. Each table in this subsection shows the family-best variant per source LLM, with cells shaded by the $|z|$-score of the bootstrap-paired gap to IE (blue: above IE, red: below). Every TR yields KL substantially below the information-free RV baseline (\cref{tab:causality-headline}), establishing that the representations encode continuation-relevant information. None of the TRs consistently exceeds the IE reference, indicating that the thought representations carry no additional causal information beyond the prompt.

\input{tables/headline_minimality}
\textbf{Minimality.} We test whether the representation compresses input-specific information and retains what is needed to predict the output, reporting the IB residual gap $\Delta_{\text{IB}}$ (\cref{eq:delta-ib}) for which higher is better. On every source LLM, the OE family ranks above IE but falls outside the interpretable range of the decomposition by construction (see \cref{methodology}). Among the remaining candidates, results are mixed (\cref{tab:minimality-headline}). LIT falls below IE on most source models, soft-thinking candidates at or above IE, and LT almost the same as IE. This indicates that no candidate consistently encodes more output-relevant compression than the prompt already provides. The absolute scale shifts across source LLMs because the cross-entropy decomposition discards a constant tied to each LLM's entropies. Therefore, ranking is comparable only on each row. \Cref{fig:results_summary} applies normalization on each LLM to recover a shared axis.

\input{tables/headline_separability}
\textbf{Separability.} We test whether $\mathbf{T}$ encodes per-question identity (\cref{eq:separability-axiom,tab:separability-headline}). Cross-task accuracy is usually near saturation for every candidate, including the IE reference. This shows that representations encode task-related information required to distinguish one task from another. Same-task (or within-task) accuracy, however, illustrates that every candidate except OE is very close to the random baseline. Panel (a) of \cref{fig:results_summary} shows the joint view of both modes for Separability. It is important to note that the highest same-task accuracy OE achieved is $73\%$. We view OE as an upper-bound to what the candidate representations can encode about the outputs in this metric. However, the collapse is structural rather than a probe-capacity artifact, with \cref{appendix:geometric} tracing it to too few effective dimensions in the within-task geometry for any probe to recover.

\input{figures/results_summary}

\input{tables/headline_stability}
\textbf{Stability.} The DCS is the AUROC of a probe that predicts from the candidate whether a question's beam outputs span more than one semantic equivalence class (\cref{tab:stability-headline}). On the four LLMs that produce non-singleton beam clusters on a sizeable fraction of questions, the candidates clear the Random Vector baseline by a wide margin, the Output Embedding family ranks highest, and the iterative thinking families exhibit decreasing DCS as the step count grows, with the largest decrease for Latent Thinking. The Input Embedding reference matches or exceeds the iterative candidates and is the highest-scoring representation on Llama-70B, reflecting that distributional uncertainty is largely predictable from the question text alone. GPT-OSS-20B is an MoE outlier whose forced beam-generation protocol yields almost no divergent semantic outputs across questions, leaving the DCS estimates uninformative rather than indicative of DCS failure. Per-variant tables across all candidates and step counts, the per-step DCS trajectory, and the projection- and length-control ablations are reported in \cref{appendix:sub:detailed_results,appendix:causality_length,appendix:sub:causality_minproj,appendix:sub:dcs_tau,appendix:capacity_ablation,appendix:sub:minimality_ib_derivation}.

\subsection{Per-family analysis}
\label{sec:results:per-family}

\textbf{Output Embedding as anchor.} The Output Embedding family serves as the upper anchor on axes that reward information about the generated continuation, since these candidates are computed from the output directly. On Causality and Separability the Output Embedding is near the maximum achievable score on every source LLM. In the within-task discrimination setting it is the only family exceeding the random baseline by a meaningful margin, with the Exact variant above the Pooled variant. The Output Embedding is not a meaningful Minimality reference, as its construction violates the chain-rule simplification that the Minimality decomposition assumes. \Cref{appendix:geometric} traces this advantage reaching content related to the  question encoded along the Output Embedding directions that other candidates sharing its cluster coordinates do not carry.

\textbf{Cross-axis pattern.} \Cref{fig:results_summary}(b) reports the per-axiom score of each family relative to the Input Embedding reference, averaged across LLMs and computed from the family-best variant per (LLM, axis) pair, with a Random Vector row as a sanity-check anchor. The heatmap recovers the Separability collapse from \cref{sec:results:per-axiom} and exposes three additional patterns. On within-task Separability, only the Output Embedding clears the random baseline. On Minimality, the iterative thinking families are mostly above the Input Embedding reference whereas Last Input Token is below it, since Last Input Token encodes input-specific information that is not used in the explicit output but is intrinsic to the input. On Stability, the iterative families and Last Input Token are close to the reference, with Soft Thinking with Gumbel noise as the only outlier. The decrease after adding noise is expected and reinforces why noise is a critical step for exploration~\citep{deng2025latentreasoningllmsvocabularyspace, wu2026llms}.

The Output Embedding's Causality cell is within bootstrap noise of the Input Embedding reference, reflecting a substantive empirical observation rather than a normalization artifact. Output Embedding and Input Embedding project into regions of the model's space that perturb the continuation distribution by comparable amounts, leaving neither dominant on this axis. Taken together, no iterative thinking family achieves the simultaneous strong clustering, wide within-task spread, and high within-question similarity that the framework rewards. The per-family geometric trajectories and the uniform failure pattern across architectures and sizes are reported in \cref{appendix:geometric}.

\subsection{Joint behavior and synthesis}
\label{sec:results:joint}
\input{tables/c4_per_llm_consistency}
No candidate thought representation consistently exceeds the Input Embedding reference on any of the four axes when results are averaged across LLMs. The per-LLM advantages reported by \cref{tab:causality-headline,tab:minimality-headline,tab:separability-headline,tab:stability-headline} for individual candidates do not generalize (\cref{tab:per-llm-consistency}). In addition, the evaluation does not match what benchmark accuracy alone would predict. A model can score well on a reasoning benchmark while its thought representations fail to discriminate two questions drawn from the same task. The per-task correlation between Separability and downstream accuracy in panel (c) of \cref{fig:results_summary} confirms that this collapse is not explained by task difficulty. The framework's contribution is a measurement protocol that exposes these gaps at the representation level, without retraining the source model and without dependence on a downstream benchmark.

\textbf{Takeaway.} \textit{No candidate beats the Input Embedding reference on every axis across the tested LLMs, and the iterative thinking variants degrade as the step count grows.}

\FloatBarrier
