\section{Background}
\subsection{Autoregressive Diffusion for Interactive Video Generation}
\label{sec: diffusion}

\paragraph{Generative modeling through diffusion.}
Score-based diffusion models~\cite{ho2020denoising, song2020score} typically perturb the data $\vx_0\sim p_{\mathrm{data}}(\vx_0)$ through the diffusion process $\vx_t = \alpha(t)\vx_0 + \sigma(t)\vepsilon, \, \vepsilon \sim \gN(\vzero, \mI), t\in [0,1]$, where $\alpha(\cdot),\sigma(\cdot)$ are functions of the timestep $t$, namely the noise schedules. Trained in this process, the model learns score-related information about $p_{\mathrm{data}}(\vx_0)$. A recent trend is flow matching~\cite{liu2022flow, lipman2022flow}, which sets $\alpha(t)=1-t$ and $\sigma(t)=t$, and parameterizes the model using velocity prediction, denoted by $\vv_\theta$. The model is trained by minimizing $\E_{\vx_0,\vepsilon, t}[||\vv_\theta(\vx_t,t)-\vv_t||^2]$, where $\vv_t := \frac{\mathrm{d}\vx_t}{\mathrm{d}t} = \vepsilon - \vx_0$. By using the optimal model $\vv_\theta$ to solve the probability flow ordinary differential equation (PF-ODE)~\cite{song2020score} $\mathrm{d}\vx_t = \vv_\theta(\vx_t,t)\mathrm{d}t,\,t:1\rightarrow 0$, one can generate samples $\vx_0 \sim p_{\mathrm{data}}(\vx_0)$. Owing to their remarkable expressive power, diffusion models have achieved widespread success in image and video generation~\cite{zhao2022egsde,zhao2024identifying,bao2024vidu,zhao2025controlvideo,zhao2025riflex,zhao2025ultravico,zhao2025ultraimage}.

\paragraph{Autoregressive diffusion for video generation.}
Despite the remarkable success of diffusion models in video generation~\cite{videoworldsimulators2024,lin2024open,zheng2024open,yang2024cogvideox,wan2025wan,kong2024hunyuanvideo,bao2024vidu,li2025radial}, they are typically bidirectional, generating the entire video content in a single pass. This leads to high latency, namely a long delay from the start of generation to the completion of the first frame, and also makes them non-interactive, since the user must specify the full conditioning signals in advance. By contrast, autoregressive (AR) generation~\cite{wu2021godiva,hong2022cogvideo,wu2022nuwa,weissenborn2019scaling,yan2021videogpt,zhao2025ultraimage,zhao2025riflex,deng2024autoregressive,kondratyuk2023videopoet}, operates on much smaller generation units, thereby offering lower latency. It also enables interactive generation, as users can provide feedback based on the content already generated and adjust subsequent conditioning signals accordingly. To combine the high generation quality of diffusion models with the low latency and interactivity of AR models, a recent trend is AR diffusion~\cite{jin2024pyramidal,teng2025magi,chen2025skyreels}, which performs autoregression across frames (or chunks) and diffusion within each frame (or chunk). These models typically adopt a causal attention mask to be trained with teacher forcing or its variants~\cite{teng2025magi, chen2024diffusion, wu2025pack, guo2025end, po2025bagger}, and perform self-rollout inference with a KV cache.


\subsection{Autoregressive Diffusion Distillation}
\label{sec:bg-distillation}
Although AR diffusion enables interactivity, its multi-step generation process still hinders real-time generation. This has motivated the development of AR diffusion distillation~\cite{lin2025autoregressive,lin2025diffusion,yang2025towards,lu2025reward,wang2026worldcompass,yang2025longlive,liu2025rolling,cui2025self,sun2025worldplay}.

\paragraph{CausVid.}
CausVid~\cite{yin2025slow}, one of the the earliest representatives of the current AR diffusion distillation paradigm, adopts a two-stage framework:
\emph{\textbf{(1) ODE initialization}}, which samples PF-ODE trajectories from a bidirectional diffusion model and trains an AR student via regression;
\emph{\textbf{(2) asymmetric DMD}}~\cite{wang2023prolificdreamer,luo2023diff,yin2024one}, which keeps the teacher and critic bidirectional while using the ODE-initialized AR model as the student, trained under diffusion forcing on real data.
The rationale is that bidirectional diffusion models currently achieve stronger performance than AR diffusion models, so a stronger bidirectional teacher is expected to transfer a better generative distribution to the AR student.

\paragraph{Self Forcing.}
Self Forcing~\cite{huang2025self} improves upon CausVid by \textbf{correcting the asymmetric DMD stage}. Specifically, it observes that CausVid trains the student under diffusion forcing on real data, so each generated frame is conditioned on ground-truth context rather than on self-generated prefixes. Consequently, the resulting frames do not form a valid generated video when concatenated, leading to a substantial gap between training and inference-time self-rollout. To resolve this issue, Self Forcing replaces diffusion forcing generation in the DMD stage with student self-rollout, thereby aligning training with inference and substantially improving performance.

\paragraph{Causal Forcing.}
Causal Forcing~\cite{zhu2026causal} \textbf{corrects the ODE initialization stage} used in CausVid and Self Forcing, while retaining the asymmetric DMD design of Self Forcing. It points out that, unlike DMD which merely pursues distribution matching, ODE distillation is intended to match the generation trajectories. Consequently, an AR student is theoretically incapable of fitting the ODE trajectories induced by a bidirectional teacher, and thus cannot properly bridge the architectural gap. Motivated by this observation, Causal Forcing first fine-tunes a bidirectional diffusion model into an AR diffusion model, and then uses the resulting AR teacher to generate ODE trajectories for initializing the AR student. The initialized student is subsequently optimized with asymmetric DMD, while the teacher and critic remain bidirectional. 

In summary, the training pipeline consists of the following three stages, whose terminology we adopt below: \emph{\textbf{(1) Stage 1: multi-step AR diffusion training} via teacher forcing; \textbf{(2) Stage 2: causal ODE initialization} with the AR teacher; and \textbf{(3) Stage 3: asymmetric DMD} with student self-rollout.} 
