\section{Experiments}
\subsection{Setup}
\label{sec: setup}

\paragraph{Implementation details.} Following Causal Forcing~\cite{zhu2026causal}, we adopt Wan2.1-1.3B~\cite{wan2025wan} as the base model from which the AR few-step model is derived. The model generates videos at a resolution of 480 $\times$ 832 with 81 frames, using frame-wise autoregressive generation. We follow the three-stage distillation pipeline of Causal Forcing, namely: Stage 1, teacher forcing AR diffusion training; Stage 2, few-step initialization; and Stage 3, asymmetric DMD with self rollout. For Stage 2 causal CD, we use the square norm, adopt 48 discretized timesteps, and employ the Euler solver.
 For the number of sampling steps, in Stage 3 we train 4-step ($t=1,0.9375,0.8333,0.625$), 2-step ($t=1,0.8333$), and 1-step settings. For settings with fewer than 4 steps, we follow ASD~\cite{yang2025towards} and keep the generation of the first latent frame at 4 steps, while using 2 steps or 1 step, respectively, for the subsequent 20 latent frames. Note that this trick is used only in Stage 3 and is not involved in the first two stages. All teacher and student models used throughout the pipeline are based on Wan2.1-1.3B, except for the two score models in the Stage 3 DMD, which are based on Wan2.1-14B. For Stage 1,2, we utilize an 80K dataset including videos sampled from OpenVid~\cite{nan2024openvid}; for Stage 3, we use the  VidProM~\cite{wang2024vidprom} dataset. The numbers of training steps for the three stages are 20K, 5K, and 1K, respectively, and the batch size is 64. All other hyperparameters are kept the same as in Causal Forcing. 

\paragraph{Evaluation.} Following Causal Forcing, we adopt two benchmarks: VBench~\cite{huang2024vbench} and VisionReward~\cite{xu2024visionreward}. For VBench, we report the overall metrics and additionally evaluate dynamic degree separately using the 100 prompts from Causal Forcing. For VisionReward, we also use the 100 prompts from Causal Forcing and report both the overall score and instruction-following performance. For readability, all metrics are multiplied by 100. All other evaluation details follow Causal Forcing. In addition, we report first-frame latency and throughput. \textit{\textbf{These efficiency metrics are measured \emph{on the single A800 GPU without the VAE-related time cost}, rather than on H100 as in the Self Forcing and Causal Forcing papers.}}

\subsection{Results}
\input{Tables/performance_comparison}
\input{Figures_tex/performance_comparison}
\paragraph{Performance comparison with existing methods.} We compare Causal Forcing++ with prior AR diffusion distillation methods, including CausVid~\cite{yin2025slow}, Self Forcing~\cite{huang2025self}, and Causal Forcing~\cite{zhu2026causal}, which are all chunk-wise models with 4-step sampling. 

As illustrated in Tab. \ref{tab:performance-comparison}, our frame-wise Causal Forcing++ (2-step generation) achieves the best VBench total score. Its Total score (84.14) and Quality score (84.89) surpasses all previous methods, while its Semantic score remains comparable to Causal Forcing. Notably, while maintaining overall generation quality comparable to, and in some aspects better than, prior state-of-the-art (SOTA) methods, our 2-step frame-wise Causal Forcing++ improves throughput by about 1.4× and reduces latency by 50\%. Furthermore, Causal Forcing++ under 4-step generation attains even higher Quality and VisionReward scores, and achieves a Dynamic Degree beyond all previous SOTA methods, while still maintaining 50\% lower first-frame latency.

These quantitative results are consistent with the visualization in Fig. \ref{fig:performance-comparison}. CausVid and Self Forcing both exhibit weaker dynamics; moreover, CausVid shows over-saturation artifacts, while Self Forcing suffers from object inconsistency, as illustrated by the spurious horizontal structure growing from the building next to the waterfall in the right-hand demo. In contrast, under fewer-step frame-wise generation, Causal Forcing++ attains quality and dynamics comparable to those of Causal Forcing, and even shows advantages in aesthetic aspects such as color vividness and brightness.


\paragraph{Ablation studies.} 
In this section, we provide a detailed method ablation. We experiment under the frame-wise 1, 2, and 4-step settings, respectively, and evaluate Self Forcing~\cite{huang2025self} ODE initialization, multi-step AR diffusion initialization, causal ODE initialization, causal DMD initialization, and our causal CD initialization. Note that the difference in the number of generation steps only applies to Stage 3, while Stage 2 remains identical across settings, with special note that the causal ODE is trained under the 4-step setting. In addition to the quality metrics in Sec.~\ref{sec: setup}, we also list here the training time (measured in A800 GPU$\cdot$ hours,  rounded to the nearest hundred) for Stage 2 (with ODE including the 80K data curation time) and the extra storage for the ODE paired data (measured in GiB). The quantitative results and visual comparisons are illustrated in Tab. \ref{tab:ablation} and Fig. \ref{fig:ablation}, respectively, from which we draw the following conclusions.

\paragraph{Self Forcing ODE initialization performs consistently poorly in the frame-wise setting.} Specifically, under the 2-step and 4-step settings, Self Forcing ODE initialization underperforms all other methods across all metrics. Its VBench Total score remains below 80, and its Dynamic Degree reaches at most 2. Beyond its poor quality, the ODE stage also costs 5,000 A800 GPU$\cdot$ hours and requires 1,500 GB of storage.


\paragraph{A strong few-step initialization is indispensable, especially in aggressive low-step settings.}
Using the multi-step AR diffusion model directly as initialization yields consistently weak performance. This issue is most severe in the 1-step setting, where the model nearly collapses in dynamics and instruction following, achieving only 0 Dynamic Degree, 1.101 VisionReward, and even a -14 Instruction Following score. Although the performance improves as the number of steps increases, it still remains clearly inferior even under 4-step generation, where its VisionReward is only 5.166, still far below all explicit few-step initialization methods. This shows that AR diffusion initialization is insufficient, and that explicit few-step adaptation before asymmetric DMD is crucial, corroborating Sec. \ref{sec:necessity}.

\paragraph{Causal CD is a stronger and more efficient substitute for causal ODE.}
Causal CD matches or surpasses causal ODE across all step settings while dramatically reducing Stage 2 cost. In 1-step generation, it improves Total, Quality, and Dynamic Degree over causal ODE, with comparable VisionReward and Instruction Following. In the 2-step setting, it achieves the best overall performance, with the highest Total, Quality, and VisionReward scores, while maintaining comparable Semantic score. In the 4-step setting, causal CD again gives the strongest overall results. Meanwhile, causal ODE requires around 11,600 A800 GPU$\cdot$ hours and 1,900\,GiB extra storage for ODE-paired data curation and training, whereas causal CD requires only around 2,900 A800 GPU$\cdot$ hours and no extra storage. Thus, causal CD reduces Stage 2 time cost by roughly 4$\times$, completely removes storage overhead, and serves as a better substitute for causal ODE initialization, consistent with Sec.~\ref{sec:substitute}.

\paragraph{Causal DMD is not an ideal replacement for few-step initialization.} Although causal DMD improves substantially over AR diffusion initialization, it is weaker than causal CD and is usually also worse than causal ODE. For example, its VisionReward remains about 0.5 lower than that of causal CD initialization across all settings. This suggests that the causal DMD initialization is suboptimal, which is we discuss the underlying reason in the analysis below.

These ablation results are highly consistent with Sec. \ref{sec: method}, and provide strong evidence for the validity and superiority of replacing causal ODE with causal CD in our Causal Forcing++.


\input{Tables/ablation}
