\section{Introduction}

In this paper, we study whether we can learn novel manipulation skills from human data on a bi-manual robotic platform.
% 
Benefiting from the robot-free setting, the collection of human manipulation data can be conducted at a much lower cost and almost anywhere~\cite{generalist2025gen0,grauman2022ego4d}.
% 
Therefore, in terms of environmental and skill diversity, human data is considered one of the most scalable, economical, and promising ways to scale up robot learning~\cite{hoque2025egodex,generalist2025gen0,zheng2026egoscale}.
% 
Along with the growing popularity of ego-centric human data collection~\cite{grauman2022ego4d,hoque2025egodex,grauman2024ego}, researchers have flocked to the exploration of how human data can improve embodied policies.


Prior research has proposed various types of human action representations for ego-centric human data.
% 
A large proportion of work adopts wrist poses~\cite{kareer2025egomimic,zhu2026emma,punamiya2025egobridge,cheang2025gr3,kareer2025emergence} with optional fingertip positions~\cite{cai2025n,qiu2025humanoid}, parameterized hands~\cite{romero2022embodied,luo2025being,luo2026being,yang2025egovla}, or finger joints~\cite{zheng2026egoscale} as human actions.
% 
In addition to explicit action representation, there are also attempts to learn latent action representations from ego-centric human videos~\cite{chen2025moto,fan2025xr,bruce2024genie,ye2024latent}.
% 
Incorporating human actions when building robot policies has shown great potential for improving generalization~\cite{kareer2025egomimic,punamiya2025egobridge,cai2025n}, reducing the reliance on robot data~\cite{zheng2026egoscale,bi2026h}, and steering pre-trained policies to manipulate novel objects and follow new instructions~\cite{cheang2025gr3,kareer2025emergence,yuan2025motiontrans} via human-robot co-training.




Different from cross-embodiment robot data~\cite{o2024open,wu2024robomind,hou2025robomind,wu2025robocoin}, learning from human data encounters unique challenges. 
% 
Human actions are typically derived from hand pose estimators~\cite{kareer2025emergence,cheang2025gr3,kareer2025egomimic,yang2025egovla}.
% 
Though human hands can be constrained by parameterization~\cite{luo2025being,luo2026being,pavlakos2024reconstructing,romero2022embodied,yang2025egovla} or re-targeting~\cite{zheng2026egoscale}, the extracted actions are inevitably noisy.
% 
Additionally, contact patterns differ greatly between human hands and robot grippers~\cite{wei2026one}: the additional degrees of freedom in fingers make wrist rotations less meaningful when reflecting manipulation behaviors.
% 
Therefore, we argue that \emph{learning rotation-inclusive robotic manipulation signals from human data is highly challenging and sub-optimal}.
% 
Instead, we propose to learn only the wrist translations as a shared action space between humans and robots.
% 
Since humans and robots both act on what they perceive, we extract the relative wrist translations under the head camera frame.
% 
To this end, we end up with two different action spaces: 1) the rotation-inclusive end-effector action to control the robot~\cite{zheng2026egoscale,cheang2025gr3,kareer2025emergence} and 2) the wrist-translation action that can be shared between humans and robots.



To enable efficient manipulation knowledge transfer from human to robot, we are required to design a unified training objective for a multi-embodiment model.
% 
We build our method upon a $\pi_0$-like~\cite{black2024pi_0} vision-language-action model, which processes the observation and instruction inputs with pre-trained vision-language priors and generates actions via flow matching~\cite{liu2022flow,lipman2022flow}.
% 
Since we are required to handle the potential missing action components across different embodiments, we propose to adopt interleaved action tokens to model wrist translation, 6DoF end-effector action, and gripper signals.
% 
By masking the missing action components within the attention layers, we can handle actions from different embodiments via variable length inputs.
% 
Critically, we find that randomly adding and substituting wrist translations for 6DoF end-effector actions on robot action data is essential for transferring wrist translation knowledge to executable robot end-effector actions.



We adopt a three-stage training strategy by first pre-training the model on a large-scale collection of human action data with only wrist translation actions.
% 
Then, we co-train the model by incorporating generalized pick-and-place robot actions~\cite{cheang2025gr3} with task-specific human actions (\textit{e.g.}, opening microwaves, wiping, stacking, \textit{etc.}).
% 
Empirically, we find that the robot is able to complete certain tasks without task-specific robot demonstrations.
% 
We can also improve the overall performance of the model by post-training on an additional set of few-shot real-robot data.
% 
To conclude, our contributions include:

\begin{itemize}
    \item We propose a bridging action representation based on wrist translations, which is shared between robots and humans and is robust to noisy human actions.
    % 
    We show that the bridging action facilitates superior manipulation skill transfer from human to robot than simply adopting noisy 6DoF human actions.

    \item To cope with the heterogeneous action components across data sources, we design a unified, interleaved action representation that handles missing action components through attention masks.
    % 
    This interleaved design also enables the model to benefit from large-scale human action pre-training.

    \item We further characterize the upper bound of our bridging representation in robotic skill transfer, suggesting its potential to scale up to broader cross-embodiment learning settings.
\end{itemize}