\section{Evaluating Conversational Tool-Using Agents}
\label{sec:evaluation}

We focus on conversational tool-using agents, a particularly complex and interesting special case of general tool use, in which static, single-turn prompts are replaced by continuous interactions between two agents. This setup naturally subsumes single-turn scenarios.
Our setup involves a conversation between two agents: the evaluated agent and the simulated user. The interaction begins with the evaluated agent receiving a task. At each turn, an agent takes an \emph{action}, either sending a message to the other agent or performing a \emph{tool call}. If the action is sending a message, the turn passes to the other agent. If the action is a tool call, the environment may modify the world state or return an observation, visible only to the calling agent, after which the agent continues its turn. This setup makes evaluation fundamentally more challenging than in standard input-output systems. Evaluation must be carried out over trajectories of actions that are shaped jointly by the agents and the environment, and also evolve the world state \citep{yehudai2025surveyevaluationllmbasedagents}. As a result, correctness must be determined by the resulting world state. A natural solution adopted by the $\tau\text{-Bench}$ benchmark \citep{taubench_24_yao} is \emph{final-state evaluation}:\footnote{Final-state evaluation reduces assessment to a single binary check on the end state, overlooking trajectory-level properties such as efficiency, intermediate rewards, and interaction quality. $\tau\text{-Bench}$ supports additional rewards, such as \emph{action checks} or \emph{communication checks}, but in practice these are excluded from the main reward in most tasks. We adopt the final-state reward alone, both for simplicity and for consistency with how results are standardly reported in $\tau\text{-Bench}$ evaluations.} a task is solved if the interaction terminates in a world state equivalent to a predefined \emph{target state}, induced by a gold tool-call sequence. Accordingly, many different trajectories may be correct, as long as they lead to this target state.

Formally, let $E$ denote the environment, which comprises the domain policy (e.g., business logic, guidelines) state space, the set of available tools $\mathcal{T}$, and an execution engine that operationalizes these tools. 
A \emph{tool call} is an action $a$ that invokes a tool $t\in\mathcal{T}$ with arguments: $a=t(\texttt{**kwargs})$. A task instance is defined as $x = (s_0, u, \sigma^\star)$, where $s_0$ is the initial state (e.g., a database), $u$ is the user instruction, and $\sigma^\star=(a_1^\star,\ldots,a_{n}^\star)$ is a gold (but not necessarily unique) tool-call sequence. The target final state is then derived as $s^\star = \texttt{apply}(\sigma^\star, s_0)$.
Let $\sigma = (a_1, \dots, a_k)$ denote the tool-call sequence produced during the interaction, i.e., the ordered list of all tool calls made by both agents.
The \emph{reward} is defined based on the state induced by $\sigma$ and $s^\star$: $r_x(\sigma) = \mathbf{1}[\texttt{apply}(\sigma, s_0) \equiv s^\star]$. Finally, we distinguish between a \emph{tool-call sequence} $\sigma$ and its corresponding \emph{tool sequence} $\bar\sigma=(t_1, \ldots, t_k)\in\mathcal{T}^k$, which is its projection onto the tool-identifiers space, that is, the ordered list of which tools were called, stripped of their arguments. In particular, $\bar\sigma^\star$ denotes the gold tool sequence. $\bar\sigma^\star$ is a central concept in our study, as it captures the structural requirements of a task and directly connects to the coverage and difficulty properties of a benchmark.

\subsection{Desiderata for Agent Benchmarks}
\label{sec:benchmark_quality}

We next propose three properties that characterize the quality of a benchmark (i.e., a set of tasks) for tool-using agents.

\paragraph{Validity.} A task is \emph{valid} if it is both \emph{verifiable} and \emph{correct}. Verifiability means that task success can be automatically checked (e.g., via final-state evaluation) rather than relying on subjective judgments. Correctness means that the gold final state $s^\star$ is consistent and faithful with the task specification $(s_0, u)$, i.e., it can be achieved given $s_0$ and a user instructed with $u$. Without a high proportion of valid tasks, agents may be penalized for benchmark defects rather than true capability limitations \citep{bestpractices2025Zhu}. Indeed, \citep{saber2025cuadron} identified dozens of invalid tasks in $\tau\text{-Bench}$, which were corrected in $\tau\text{-Bench Verified}$.

\paragraph{Coverage.}
A useful benchmark should cover the relevant task space, rather than sampling similar tasks for a smaller subspace. In our setting, each task $x$ is associated with a gold tool sequence $\bar\sigma^\star$ (i.e., the names of the tools leading to $s^\star$). We define coverage in terms of these gold tool sequences, thereby abstracting away from superficial linguistic variation in user instructions and initial states, and focusing instead on the procedural structure of the task. A benchmark has good coverage if its tasks represent structurally different tool sequences, rather than many near-duplicates. In addition, benchmark tasks should reflect common patterns of tool use, rather than arbitrary combinations or isolated outliers. Coverage can be quantified, for example, using the tool-edit distance between sequences or the number of unique $n$-grams of tools. This mirrors standard diversity/novelty quantification in NLP \citep{raven2023mccoy}, with the key distinction that our units are tool names rather than characters or words.

\paragraph{Difficulty.}
A benchmark should contain tasks that meaningfully differentiate between capability levels. If all tasks are too easy, the benchmark quickly saturates; if all tasks are too hard, it cannot provide a useful ranking of agents. In our setting, task difficulty arises from two sources. \emph{structural difficulty}, determined by properties of the gold tool sequence, such as its length and the types of tools involved (e.g., write operations versus simple reads), and \emph{interaction difficulty}, determined by the instruction ambiguity, initial-state complexity (e.g., distractors, dependencies, stale information), and user behavior (e.g., cooperative vs. adversarial).
