\section{Related Work}\label{sec:related}

% Failure-mode analysis is being refreshed against the leak-fix result tables.
% Keep the old figure out of the current Chinese-stage draft to avoid mixing result sources.

% We position Claw-SWE-Bench against three lines of work: foundational SWE benchmarks, agent-harness research, and other coding benchmarks.

\textbf{Foundational SWE benchmarks.}
SWE-bench~\cite{jimenez_swebench_2024} introduced the task formulation we adopt: resolving real GitHub issues against repository-level test suites. Multilingual coverage has since been pushed beyond Python by Multi-SWE-bench / SWE-bench-Multilingual~\cite{swe_smith}, which contributes 300 of our 350 instances, and Python coverage is supplied by the human-validated SWE-bench-Verified-Mini~\cite{swebench_verified_mini}, the source of our remaining 50 instances. Subset construction has precedent in SWE-bench Lite~\cite{swebench_lite} and in the more general anchor-set methodology of tinyBenchmarks~\cite{tinybenchmarks}; our 80-instance Lite subset extends this lineage with a rank-aware ILP. Orthogonal to evaluation, SWE-smith~\cite{swe_smith} scales SWE training data; its scope is data generation rather than harness comparison.

\textbf{Single-harness SWE evaluations.}
A growing line of work proposes individual harnesses and reports their per-system numbers on SWE-bench. SWE-agent~\cite{swe_agent} introduced the agent-computer-interface scaffold and is the most cited SWE harness. AutoCodeRover~\cite{zhang_autocoderover} adds code-aware retrieval to the agent loop. OpenHands~\cite{wang_openhands} is a generalist agent platform that ships a SWE-bench adapter. mini-SWE-agent~\cite{mini_swe_agent} represents the minimal-harness end of the design space. SWE-Bench Pro~\cite{deng_swebench_pro} extends the SWE-bench formulation to long-horizon tasks. Each of these reports per-harness resolved rates but does not vary the harness as an experimental axis: prompt, scaffold, runtime, and termination policy are bundled together with each release. Our five-claw $\times$ two-model claw sweep compares heterogeneous harnesses under a fixed outer protocol and reports accuracy together with cost, duration, and cache accounting.

% \paragraph{Holistic and resource-aware agent evaluation.}
% The two works closest to ours are HAL~\cite{kapoor_hal} and SWE-Effi~\cite{fan_swe_effi}. HAL advocates a holistic philosophy for agent leaderboards spanning accuracy, cost, and latency; methodologically, however, it ships only a single harness, leaving the harness $\times$ model interaction unidentifiable from its grid. SWE-Effi re-evaluates SWE-bench under resource constraints and concludes that scaffold and base model jointly determine effectiveness, so that no scaffold ranking is meaningful in isolation. SWE-Effi diagnoses the conflation but does not turn the harness into a controlled experimental variable: it varies scaffolds without holding prompt, max-turns, timeout, and concurrency identical, so the joint dependency it reports remains a caveat rather than a measurement. We operationalize their observation: by fixing the prompt template, \texttt{max-turns}=300, per-instance \texttt{timeout}=3600\,s, task set, and concurrency across all four harnesses and nine LLMs, we convert the joint scaffold $\times$ model dependency into a directly measurable interaction term, reported as finding F2 in Section~\ref{sec:results}. Unlike prior work, Claw-SWE-Bench treats the agent harness itself as a controlled variable: we hold the prompt template, max-turns (300), per-instance timeout (3600\,s), task set, and concurrency constant across four harnesses (openclaw, hermes-agent, zeroclaw, nanobot) and nine LLMs, yielding the first systematic harness $\times$ model evaluation grid for SWE-class coding tasks.

\textbf{Other coding benchmarks.}
A broader coding-evaluation literature complements SWE-bench-style issue resolution. HumanEval~\cite{humaneval}, MBPP~\cite{mbpp}, and APPS~\cite{apps} score function-level synthesis; CrossCodeEval~\cite{crosscodeeval} targets cross-file completion; CodeClash~\cite{codeclash} and PinchBench~\cite{pinchbench} probe further axes of coding capability. These benchmarks evaluate model code-generation skill rather than harness-mediated agentic behavior on real repositories; we treat them as orthogonal background.
