\section{Conclusion and Discussion}\label{sec:discussion}
\textbf{Conclusion.}
We introduced \textsc{Claw-SWE-Bench}, a 350-instance multilingual SWE-style benchmark, together with \textsc{Claw-SWE-Bench Lite}, an 80-instance subset. Through the adapter, a general-purpose agent such as OpenClaw can be constrained to the SWE-bench execution environment, patch contract, and scoring protocol, making it a repeatable, comparable, and scorable coding-agent evaluation target. Lite uses a 17-column cost-aware calibration spanning both model and harness variation, reproducing full-set aggregate Pass@1 within about $0.4$\,pp while reducing full-run cost to about $23\%$ of full-350. With prompt, budget, and task set fixed, the OpenClaw $\times$ nine-model sweep and the five-claw $\times$ two-model sweep jointly show that harness choice can reorder system rankings and reshape accuracy--cost trade-offs. SWE-style coding-agent evaluation should therefore report total API cost and cache hit rate alongside Pass@1; the harness should be treated as a first-class controlled variable rather than an implementation detail hidden behind a model score. We hope the full benchmark and Lite subset can serve as stable reference points for follow-on work, allowing new models, harnesses, and adapter changes to be compared under the same tasks, budgets, and cost-accounting conventions while making low-cost debugging and replication easier to incorporate into routine research workflows.

\textbf{Limitations and future directions.}
Several boundaries of the current results should be interpreted carefully. First, the main experiments report single-run aggregates; differences of only a few percentage points should therefore not be overinterpreted as stable system superiority, and future work should use multi-seed replication to estimate randomness and run-to-run variance. Second, the claw sweep covers five claws and two representative models, which is sufficient to show that the harness is a first-order variable in SWE-style coding-agent evaluation, but not sufficient to fully decompose harness $\times$ model interactions. A wider model axis would help determine which conclusions arise from general harness mechanisms and which depend on a specific backbone. Third, cost analysis depends on provider-side pricing and cache accounting. We therefore report total API cost, input/output tokens, and cache hit rate together; future releases should also retain raw token traces so that cost differences can be audited and re-priced. More broadly, whether the model--harness non-separability observed here generalizes to web agents or computer-use agents, and how harness components such as agent loop, tool surface, parser, and stopping rule drive accuracy--cost trade-offs, remain open questions.


