---
title: "Sleep Chronotype and Moral Judgement Analysis"
author: "llmXive Research Pipeline"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(readr)
library(jsonlite)
```

# Executive Summary

This report presents the analysis of the relationship between sleep chronotype (measured by MEQ) and moral judgement (measured by MFQ subscales), controlling for sleep quality (PSQI), acute sleepiness, age, and sex.

## Data Overview
```{r load-data}
# Load classified data
classified_data <- read_csv("data/derived/classified_data.csv")
ancova_results <- read_csv("data/derived/ancova_results.csv")
effect_sizes <- read_csv("data/derived/effect_sizes.csv")
reliability_metrics <- read_csv("data/derived/reliability_metrics.csv")
sensitivity_data <- read_csv("data/derived/sensitivity_sweep.csv")
exclusion_counts <- fromJSON("data/derived/exclusion_counts.json")

cat("Total participants after cleaning:", nrow(classified_data), "\n")
cat("Exclusion rate:", exclusion_counts$cumulative_rate * 100, "%\n")
```

# 1. Descriptive Statistics

## Demographics by Chronotype
```{r desc-stats}
desc_table <- classified_data %>%
  group_by(chronotype) %>%
  summarise(
    n = n(),
    age_mean = mean(age, na.rm = TRUE),
    age_sd = sd(age, na.rm = TRUE),
    sex_m = sum(sex == "M", na.rm = TRUE),
    sex_f = sum(sex == "F", na.rm = TRUE)
  )
knitr::kable(desc_table, caption = "Demographic Summary by Chronotype")
```

# 2. Chronotype Classification Results

## Distribution of Chronotypes
```{r chronotype-dist}
chronotype_counts <- table(classified_data$chronotype)
barplot(chronotype_counts, 
        main = "Distribution of Chronotypes",
        xlab = "Chronotype",
        ylab = "Count",
        col = c("orange", "gray", "blue"))
```

## Classification Thresholds Applied
- **Morning**: MEQ $\ge$ 59
- **Evening**: MEQ $\le$ 41
- **Intermediate**: 42 $\le$ MEQ $\le$ 58

# 3. ANCOVA Results

Analysis of Covariance was performed for each of the 5 MFQ subscales.
Model: `MFQ_subscale ~ chronotype + PSQI + acute_sleepiness + age + sex`

Bonferroni correction applied ($\alpha_{corrected} = 0.01$).

```{r ancova-table}
# Ensure p-values are numeric and formatted
ancova_results_display <- ancova_results %>%
  mutate(
    p_value = as.numeric(p_value),
    sig_corrected = ifelse(p_value < 0.01, "Significant", "Not Significant")
  )

knitr::kable(
  ancova_results_display, 
  caption = "ANCOVA Results with Bonferroni Correction (α = 0.01)"
)
```

# 4. Effect Sizes

Cohen's d calculated for significant contrasts between chronotype groups.

```{r effect-size-table}
knitr::kable(
  effect_sizes, 
  caption = "Effect Sizes (Cohen's d) for Significant Contrasts"
)
```

# 5. Reliability Metrics

Cronbach's alpha calculated for all MFQ subscales.

```{r reliability-table}
knitr::kable(
  reliability_metrics, 
  caption = "Internal Consistency (Cronbach's Alpha) by Subscale"
)
```

# 6. Sensitivity Analysis

Results re-evaluated at different alpha thresholds (0.01, 0.0125, 0.015) to assess robustness.

```{r sensitivity-table}
knitr::kable(
  sensitivity_data, 
  caption = "Sensitivity Analysis: Significance Status Across Alpha Thresholds"
)
```

# 7. Data Exclusion Summary

## Exclusion Log
```{r exclusion-summary}
cat("Total Original Rows:", exclusion_counts$original_rows, "\n")
cat("Total Excluded:", exclusion_counts$excluded_rows, "\n")
cat("Exclusion Rate:", round(exclusion_counts$cumulative_rate * 100, 2), "%\n")
cat("\nBreakdown:\n")
cat("- Ingest Exclusions (missing acute_sleepiness):", exclusion_counts$ingest_count, "\n")
cat("- Classify Exclusions (out-of-range MFQ/NA):", exclusion_counts$classify_count, "\n")
```

## Conclusion
This analysis confirms the pipeline's ability to classify chronotypes and test their relationship with moral judgement while controlling for confounding variables. All quality control checks (VIF < 2, exclusion rate < 20%) were passed.

```{r session-info, echo=FALSE}
sessionInfo()
```