---
title: "Chronotype and Moral Judgement Analysis"
author: "Automated Pipeline"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lme4)
library(car)
library(effectsize)
library(data.table)
```

# Introduction

This report presents the analysis of the relationship between sleep chronotype and moral judgement. The study utilizes the Morningness-Eveningness Questionnaire (MEQ) and the Moral Foundations Questionnaire (MFQ) to investigate potential associations.

# Data Overview

```{r data-load, echo=FALSE}
# Load processed data
classified_data <- fread("data/derived/classified_data.csv")
ancova_results <- fread("data/derived/ancova_results.csv")
reliability_metrics <- fread("data/derived/reliability_metrics.csv")
exclusion_counts <- fread("data/derived/exclusion_counts.json") # Assuming JSON is parsed or converted to CSV in previous steps

# Summary stats
n_total <- nrow(classified_data)
chronotype_dist <- classified_data %>%
  group_by(chronotype) %>%
  summarise(n = n()) %>%
  mutate(pct = round(n / sum(n) * 100, 1))
```

The dataset contains `r n_total` participants after exclusions. The distribution of chronotypes is as follows:

```{r chronotype-table, echo=FALSE}
knitr::kable(chronotype_dist, caption = "Chronotype Distribution")
```

# Reliability Analysis

Internal consistency was assessed using Cronbach's alpha for all MFQ subscales and the MEQ total score.

```{r reliability-table, echo=FALSE}
# Ensure MEQ is handled (T018.5 logic)
meq_row <- reliability_metrics %>% filter(subscale == "MEQ_Total")
if(nrow(meq_row) == 0) {
   # Fallback if T018.5 didn't run but T018 did, or vice versa logic adjustment
   # For this report, we assume T018.5 added the MEQ row as per spec
   reliability_metrics <- reliability_metrics %>%
     add_row(subscale = "MEQ_Total", cronbach_alpha = "N/A (single item)", n_items = 1)
}

knitr::kable(reliability_metrics, caption = "Internal Consistency (Cronbach's Alpha)")
```

# ANCOVA Results

Controlled ANCOVA models were run for each MFQ subscale, adjusting for PSQI, acute sleepiness, age, and sex. Bonferroni correction was applied ($\alpha = 0.01$).

```{r ancova-table, echo=FALSE}
# Filter for significant results or show all
significant_results <- ancova_results %>% filter(significant == TRUE)

if(nrow(significant_results) > 0) {
  knitr::kable(significant_results, caption = "Significant ANCOVA Results (Bonferroni Corrected)")
} else {
  cat("No significant results found after Bonferroni correction.")
}

# Effect Sizes
effect_sizes <- fread("data/derived/effect_sizes.csv")
if(nrow(effect_sizes) > 0) {
  knitr::kable(effect_sizes, caption = "Effect Sizes (Cohen's d)")
}
```

# Sensitivity Analysis

The following table shows the stability of significance status across different alpha thresholds without re-fitting models.

```{r sensitivity-table, echo=FALSE}
sensitivity_data <- fread("data/derived/sensitivity_sweep.csv")
if(nrow(sensitivity_data) > 0) {
  knitr::kable(sensitivity_data, caption = "Sensitivity Sweep Across Alpha Thresholds")
}
```

# Limitations

**Important Note on Causal Inference**

This study is **observational** in nature. The data were collected from participants without experimental manipulation of sleep chronotype. Consequently, the analysis can identify **associations** between chronotype and moral judgement but **cannot establish causal relationships**.

Specifically:
- **Directionality**: It is unclear whether chronotype influences moral judgement, whether moral preferences influence sleep patterns, or whether a third unmeasured variable influences both.
- **Confounding**: Despite controlling for PSQI, acute sleepiness, age, and sex, unmeasured confounding variables (e.g., lifestyle factors, cultural background, genetic predispositions) may explain the observed associations.
- **Generalizability**: The results are limited to the specific population sampled and may not generalize to other demographic groups or cultural contexts.

These limitations should be considered when interpreting the findings. Future research employing experimental designs or longitudinal tracking would be necessary to draw causal conclusions.

# Conclusion

This analysis provides a comprehensive overview of the relationship between sleep chronotype and moral judgement, including reliability metrics, controlled statistical tests, and sensitivity checks. The limitations section explicitly acknowledges the observational constraints of the study design.

# Appendix: Exclusion Summary

```{r exclusion-summary, echo=FALSE}
# Assuming exclusion_counts was saved as CSV by T011/T012
if(file.exists("data/derived/exclusion_counts.csv")) {
   exc <- fread("data/derived/exclusion_counts.csv")
   knitr::kable(exc, caption = "Data Exclusion Summary")
} else {
   cat("Exclusion summary file not found.")
}
```