\section{Related Work}\label{sec:related_work}
\paragraph{Reinforcement learning for agentic LLMs.}
Large language models are increasingly trained as interactive agents that operate over long horizons, invoke tools, and receive feedback from environments or verifiers~\citep{shridhar2020alfworld,yao2022webshop,jin2025searchr1,jimenez2023swebench,wu2026atlas}. Reinforcement learning has therefore become a natural post-training paradigm, with outcome-based methods such as GRPO providing a stable critic-free objective for on-policy rollouts~\citep{shao2024deepseekmath}. However, agentic environments typically provide sparse and delayed rewards. A terminal outcome can indicate whether a trajectory succeeds, but it does not identify which intermediate decisions caused success or failure. OPID targets this missing credit-assignment signal: it keeps outcome-based RL as the optimization backbone, but augments it with dense decision-level supervision extracted from the policy's own completed trajectories.

\paragraph{On-policy self-distillation.}
On-policy distillation trains a model from its own sampled outputs while using auxiliary teacher signals to provide token-level learning targets~\citep{agarwal2024onpolicy,gu2023minillm}. Recent self-distillation methods further remove the need for a separate teacher by comparing the same policy under different contexts or feedback conditions~\citep{zhao2026opsd,he2026sdzero}. For multi-turn agents, this suggests a useful decomposition: RL supplies task-level optimization, while self-distillation supplies dense shaping signals~\citep{lu2026sdar}. The key question is where the privileged signal should come from. Existing methods often rely on generic revision contexts, external hints, or task-level feedback transformations. OPID instead constructs the privileged branch from hindsight skills extracted from on-policy trajectories, making the distillation signal directly tied to the states, actions, and failures encountered by the current policy.

\paragraph{Skill-conditioned agent learning.}
Natural-language skills provide compact procedural knowledge for agents, including subgoal decompositions, action templates, and failure-avoidance rules~\citep{lu2026skill0,wang2026skillsd,lu2026sdar,wu2026maestro}. Existing skill-based methods commonly depend on external skill libraries, retrieved skill files, or persistent skill memories. These designs can improve agent behavior, but they introduce maintenance and retrieval costs, and retrieved skills may be mismatched with the state distribution induced by the current policy. This mismatch becomes more severe in long-horizon interaction, where small deviations can lead to substantial state drift. OPID makes a different design choice: it extracts hierarchical skills directly from completed on-policy trajectories, routes them according to decision criticality, and distills their behavioral effect into the policy during training. As a result, OPID provides distribution-matched hindsight supervision without requiring skill retrieval, analyzer calls, or privileged context at inference time.