% Theoretical appendix for OPID.
% Required packages: amsmath, amssymb, amsthm, natbib.
% Define the theorem environment if it is not already available:
% \newtheorem{proposition}{Proposition}

\section{Theoretical Analysis}
\label{app:theory}

This section provides three results that correspond to the main design choices of
OPID. We first place the proposed teacher advantage among representative
on-policy distillation objectives. We then show that it implements a sampled-token
reverse-KL update, characterize the benefit of collecting distillation contexts
on policy, and justify critical-first routing under a natural specialization
assumption.
\subsection{Notation and Representative On-Policy Distillation Objectives}
\label{app:theory_setup}

\subsubsection{Notation}

Let \(i=(\tau,t,\ell)\) index a valid token position in a response. We denote
the corresponding standard autoregressive context by
\(c_i=(h_{\tau,t},y_{\tau,t,<\ell})\), and its skill-augmented counterpart by
\(\widetilde c_i\). At each token position, define
\begin{equation*}
\begin{aligned}
 b_i(v) &\triangleq \pi_{\theta_{\mathrm{old}}}(v\mid c_i),
 & q_i(v) &\triangleq \pi_{\theta_{\mathrm{old}}}(v\mid \widetilde c_i),
 & p_{\theta,i}(v) &\triangleq \pi_{\theta}(v\mid c_i).
\end{aligned}
\end{equation*}
Here, \(b_i\) is the behavior distribution used to generate the response,
\(q_i\) is a detached skill-conditioned teacher distribution, and
\(p_{\theta,i}\) is the trainable policy evaluated under the standard context
available at inference time. The observed token
\(a_i\triangleq y_{\tau,t,\ell}\) is sampled from \(b_i\).

We further define the token-level log-likelihood gap and the policy importance
ratio as
\begin{equation}
\begin{aligned}
 \Delta_i(v) &\triangleq \log q_i(v)-\log b_i(v),
 & \rho_{\theta,i}(v) &\triangleq
 \frac{p_{\theta,i}(v)}{b_i(v)}.
\end{aligned}
\label{eq:gap_ratio}
\end{equation}
The quantity \(\Delta_i(v)\) measures the change in token log-probability
induced by the skill-augmented context. In particular,
\(\Delta_i(v)>0\) indicates that the skill-conditioned teacher assigns greater
probability to token \(v\) than the behavior policy does. The OPID skill
advantage associated with the observed token is therefore
\[
A_i^{\mathrm{skill}}=\Delta_i(a_i).
\]
Unless otherwise stated, all expectations below are taken over valid response
tokens; the response mask is consequently omitted for notational simplicity.

\subsubsection{Representative On-Policy Distillation Objectives}

On-policy distillation (OPD) applies teacher supervision at autoregressive
contexts generated by the student or a behavior policy, thereby reducing the
context-distribution mismatch between distillation training and free-running
inference \citep{agarwal2024onpolicy}. The context-generation policy and the
granularity of teacher supervision are orthogonal design choices. At each
on-policy context, output-space OPD objectives can be organized into three
common supervision granularities: full-vocabulary, Top-$K$, and sampled-token
distillation \citep{li2026rethinkingopd,fu2026revisitingopd}. OPID belongs to
the sampled-token category.

\paragraph{Full-vocabulary distribution matching.}

Let \(q_i\) and \(p_{\theta,i}\) denote the teacher and student next-token
distributions, respectively, at autoregressive context \(i\). When the complete
predictive distributions are available, OPD can minimize the forward KL,
reverse KL, or a generalized Jensen--Shannon divergence
\citep{hinton2015distilling,agarwal2024onpolicy,gu2024minillm}:
\begin{equation*}
\begin{aligned}
 \mathcal L_{\mathrm{FKL}}(\theta)
 &=
 \mathbb E_i\!\left[
 D_{\mathrm{KL}}(q_i\|p_{\theta,i})
 \right],\\
 \mathcal L_{\mathrm{RKL}}(\theta)
 &=
 \mathbb E_i\!\left[
 D_{\mathrm{KL}}(p_{\theta,i}\|q_i)
 \right],\\
 \mathcal L_{\mathrm{JSD}}^{(\alpha)}(\theta)
 &=
 \mathbb E_i\!\left[
 \alpha D_{\mathrm{KL}}(q_i\|m_i^{(\alpha)})
 +(1-\alpha)D_{\mathrm{KL}}(p_{\theta,i}\|m_i^{(\alpha)})
 \right],\\
 m_i^{(\alpha)}
 &=
 \alpha q_i+(1-\alpha)p_{\theta,i}.
\end{aligned}
\end{equation*}
Forward KL gives the conventional soft-target objective and emphasizes
coverage of teacher-supported probability mass. Reverse KL instead penalizes
student probability assigned to teacher-disfavored regions and therefore
typically exhibits more mode-seeking behavior. Generalized JSD compares both
models against a mixture distribution, with \(\alpha=\tfrac12\) recovering
the standard symmetric JSD \citep{kullback1951information,lin1991divergence}.

\paragraph{Top-\(K\) distribution matching.}

Top-\(K\) OPD retains distribution-level supervision over a restricted local
support. Common choices include a student-selected support
\citep{li2026rethinkingopd,ye2026opcd} and a teacher-selected support
\citep{fu2026revisitingopd}:
\begin{equation*}
\begin{aligned}
 S_{i,p}^{(K)}
 &\triangleq
 \operatorname{TopK}(p_{\theta,i},K),\\
 S_{i,q}^{(K)}
 &\triangleq
 \operatorname{TopK}(q_i,K),\\
 S_i^{(K)}
 &\in
 \left\{S_{i,p}^{(K)},S_{i,q}^{(K)}\right\}.
\end{aligned}
\end{equation*}
For the selected support \(S_i^{(K)}\), define the restricted and
renormalized distributions
\begin{equation*}
\begin{aligned}
 \bar p_{\theta,i}^{S_i^{(K)}}(v)
 &\triangleq
 \frac{
 p_{\theta,i}(v)\mathbf 1\{v\in S_i^{(K)}\}
 }{
 \sum_{u\in S_i^{(K)}}p_{\theta,i}(u)
 },\\
 \bar q_i^{S_i^{(K)}}(v)
 &\triangleq
 \frac{
 q_i(v)\mathbf 1\{v\in S_i^{(K)}\}
 }{
 \sum_{u\in S_i^{(K)}}q_i(u)
 }.
\end{aligned}
\end{equation*}
A representative truncated reverse-KL objective is
\begin{equation*}
\begin{aligned}
 \mathcal L_{\mathrm{TopK\text{-}RKL}}(\theta)
 &=
 \mathbb E_i\!\left[
 D_{\mathrm{KL}}\!\left(
 \bar p_{\theta,i}^{S_i^{(K)}}
 \,\middle\|\,
 \bar q_i^{S_i^{(K)}}
 \right)
 \right]\\
 &=
 \mathbb E_i\!\left[
 \sum_{v\in S_i^{(K)}}
 \bar p_{\theta,i}^{S_i^{(K)}}(v)
 \log
 \frac{
 \bar p_{\theta,i}^{S_i^{(K)}}(v)
 }{
 \bar q_i^{S_i^{(K)}}(v)
 }
 \right].
\end{aligned}
\end{equation*}
Top-\(K\) matching occupies an intermediate point between one-token and
full-vocabulary supervision. It preserves multi-token information at reduced
computational or communication cost, but discards probability mass outside the
selected support and is therefore a truncated, support-dependent approximation
to the full reverse KL.

\paragraph{Sampled-token distillation.}

At a fixed on-policy context, define the teacher--student log-ratio cost
\begin{equation*}
 \delta_i(v)
 \triangleq
 \log p_{\theta,i}(v)-\log q_i(v).
\end{equation*}
The token-level reverse KL can then be written exactly as an expectation over
student-sampled tokens:
\begin{equation*}
\begin{aligned}
 D_{\mathrm{KL}}(p_{\theta,i}\|q_i)
 &=
 \mathbb E_{a_i\sim p_{\theta,i}}
 \left[
 \delta_i(a_i)
 \right]\\
 &=
 \mathbb E_{a_i\sim b_i}
 \left[
 \rho_{\theta,i}(a_i)\delta_i(a_i)
 \right],
 \qquad
 \rho_{\theta,i}(a)
 \triangleq
 \frac{p_{\theta,i}(a)}{b_i(a)}.
\end{aligned}
\end{equation*}
The second equality requires \(p_{\theta,i}\ll b_i\)(support coverage condition). Consequently,
\(\rho_{\theta,i}(a_i)\delta_i(a_i)\) is an importance-weighted
single-sample estimator of the per-context reverse KL. Its score-function
gradient is
\begin{equation*}
\begin{aligned}
 \nabla_\theta
 D_{\mathrm{KL}}(p_{\theta,i}\|q_i)
 =
 \mathbb E_{a_i\sim b_i}
 \left[
 \rho_{\theta,i}(a_i)
 \operatorname{sg}\!\left[\delta_i(a_i)\right]
 \nabla_\theta\log p_{\theta,i}(a_i)
 \right],
\end{aligned}
\end{equation*}
where \(\operatorname{sg}\) denotes stop-gradient. This connection permits
sampled-token distillation to be implemented with policy-gradient or
importance-weighted policy-optimization machinery
\citep{gu2024minillm,lu2025onpolicydistillation,oh2026kl}.
Compared with full-vocabulary matching, sampled-token supervision requires
only the teacher probability of the realized token, but has higher
Monte Carlo variance and uses less information from the teacher distribution.


\paragraph{From the clipped OPID objective to its unclipped skill
surrogate.}
Let \(i=(\tau,t,\ell)\) index a valid rollout-token position, and let
\(\nu_b\) denote the distribution over valid token positions induced by
rollouts collected from the behavior policy. Given position \(i\), the
observed token \(a_i\) is sampled from \(b_i\). Recall from
Eq.~\ref{eq:gap_ratio} that
\[
 \Delta_i(v)
 =
 \log q_i(v)-\log b_i(v),
 \qquad
 \rho_{\theta,i}(v)
 =
 \frac{p_{\theta,i}(v)}{b_i(v)},
\]
where \(b_i\), \(q_i\), and the resulting advantages are detached during
the policy update. The skill advantage of a sampled token is
\begin{equation*}
 A_i^{\mathrm{skill}}(a_i)
 =
 \Delta_i(a_i).
\end{equation*}

The complete OPID advantage combines the outcome and skill signals:
\begin{equation*}
 A_i^{\mathrm{OPID}}(a_i)
 =
 A_i^{\mathrm{ep}}
 +
 \lambda_{\mathrm{skill}}\Delta_i(a_i).
\end{equation*}
Accordingly, the implemented clipped policy loss is
\begin{equation*}
\begin{aligned}
 \mathcal L_{\mathrm{policy}}(\theta)
 =
 -
 \mathbb E_{\substack{i\sim\nu_b\\a_i\sim b_i}}
 \Big[
 \min\Big(
   &\rho_{\theta,i}(a_i)
    A_i^{\mathrm{OPID}}(a_i),
 \\
   &\operatorname{clip}
   \bigl(
     \rho_{\theta,i}(a_i),
     1-\epsilon,
     1+\epsilon
   \bigr)
   A_i^{\mathrm{OPID}}(a_i)
 \Big)
 \Big].
\end{aligned}
\end{equation*}
In a realized rollout batch, this expectation is implemented as an
empirical average over the observed valid tokens \(a_i\).

Because PPO clipping is applied after the outcome and skill advantages
have been combined, the clipped objective does not in general decompose
into independently clipped outcome and skill losses. To isolate the
skill-distillation signal studied below, we therefore consider the
corresponding unclipped policy surrogate:
\begin{equation*}
 \mathcal L_{\mathrm{policy}}^{\mathrm{unclip}}(\theta)
 \triangleq
 -
 \mathbb E_{\substack{i\sim\nu_b\\a_i\sim b_i}}
 \left[
   \rho_{\theta,i}(a_i)
   A_i^{\mathrm{OPID}}(a_i)
 \right].
\end{equation*}
Unlike the clipped objective, this loss decomposes exactly as
\begin{equation*}
 \mathcal L_{\mathrm{policy}}^{\mathrm{unclip}}(\theta)
 =
 \mathcal L_{\mathrm{ep}}^{\mathrm{unclip}}(\theta)
 +
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta),
\end{equation*}
where
\begin{equation*}
 \mathcal L_{\mathrm{ep}}^{\mathrm{unclip}}(\theta)
 \triangleq
 -
 \mathbb E_{\substack{i\sim\nu_b\\a_i\sim b_i}}
 \left[
   \rho_{\theta,i}(a_i)A_i^{\mathrm{ep}}
 \right]
\end{equation*}
and
\begin{equation}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 \triangleq
 -
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\a_i\sim b_i}}
 \left[
   \rho_{\theta,i}(a_i)\Delta_i(a_i)
 \right].
\label{eq:opid_unclipped_skill_loss}
\end{equation}

Equation~\ref{eq:opid_unclipped_skill_loss} is the skill-distillation
loss analyzed in the next subsection. Although it is defined through the
unclipped surrogate, it characterizes the local skill-induced update of
the implemented PPO loss. In particular, let
\(\theta_0=\theta_{\mathrm{old}}\), so that
\(p_{\theta_0,i}=b_i\) and
\(\rho_{\theta_0,i}(a)=1\). Since \(1\) lies in the interior of the
clipping interval, the clipped and unclipped objectives have the same
value and gradient at the behavior policy:
\begin{equation*}
\begin{aligned}
 \mathcal L_{\mathrm{policy}}^{\mathrm{clip}}(\theta_0)
 &=
 \mathcal L_{\mathrm{policy}}^{\mathrm{unclip}}(\theta_0),
 \\
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{policy}}^{\mathrm{clip}}(\theta)
 \right|_{\theta=\theta_0}
 &=
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{policy}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}
 \\
 &=
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{ep}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}
 +
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}.
\end{aligned}
\end{equation*}
Thus,
\(\mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}\) is exactly the
skill-induced component of the first-order PPO update around the behavior
policy. Away from this local region, clipping couples the outcome and
skill signals through the sign of their combined advantage, and the
unclipped decomposition no longer describes the complete clipped
objective globally.

\subsection{The Unclipped OPID Skill Loss as a Relative-KL Surrogate}
\label{app:teacher_advantage}

We now analyze the unclipped skill-distillation loss introduced in
Eq.~\ref{eq:opid_unclipped_skill_loss}. Let \(\nu_b\) denote the
distribution over valid token positions induced by rollouts collected from
the behavior policy. Throughout this subsection, the rollout histories,
routed skills, and the corresponding distributions \(b_i\) and \(q_i\) are
detached and held fixed during the policy update.

We assume the common-support condition
\begin{equation*}
 p_{\theta,i}\ll b_i
 \qquad\text{and}\qquad
 p_{\theta,i}\ll q_i
\end{equation*}
for every \(i\) in the support of \(\nu_b\). This condition is satisfied by
standard softmax language models with finite logits.

Recall that
\begin{equation*}
\begin{aligned}
 \Delta_i(v)
 &\triangleq
 \log q_i(v)-\log b_i(v),
 &
 \rho_{\theta,i}(v)
 &\triangleq
 \frac{p_{\theta,i}(v)}{b_i(v)}.
\end{aligned}
\end{equation*}
The unclipped OPID skill loss is
\begin{equation}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 \triangleq
 -
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\a\sim b_i}}
 \left[
   \rho_{\theta,i}(a)\Delta_i(a)
 \right].
\label{eq:opid_skill_loss_recalled}
\end{equation}
In a realized rollout batch, this expectation is approximated by the
empirical average over the observed valid tokens. The expectation notation in
Eq.~\ref{eq:opid_skill_loss_recalled} makes the rollout-time token sampling
law explicit for the theoretical analysis.

Define the behavior-relative KL and the student--teacher reverse-KL loss as
\begin{equation*}
\begin{aligned}
 \mathcal D_b(\theta)
 &\triangleq
 \mathbb E_{i\sim\nu_b}
 \left[
   D_{\mathrm{KL}}
   \left(
     p_{\theta,i}\|b_i
   \right)
 \right],
 \\
 \mathcal L_{\mathrm{RKL}}(\theta)
 &\triangleq
 \mathbb E_{i\sim\nu_b}
 \left[
   D_{\mathrm{KL}}
   \left(
     p_{\theta,i}\|q_i
   \right)
 \right].
\end{aligned}
\end{equation*}


\begin{proposition}[Exact relative-KL decomposition]
\label{prop:relative_kl}
Under the assumptions above, for every admissible \(\theta\),
\begin{equation}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 =
 \lambda_{\mathrm{skill}}
 \left[
   \mathcal L_{\mathrm{RKL}}(\theta)
   -
   \mathcal D_b(\theta)
 \right].
\label{eq:relative_kl_decomposition}
\end{equation}
Let \(\theta_0=\theta_{\mathrm{old}}\) and suppose that
\(p_{\theta_0,i}=b_i\) for every \(i\). Then
\begin{align}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta_0)
 &=
 \lambda_{\mathrm{skill}}
 \mathcal L_{\mathrm{RKL}}(\theta_0),
\label{eq:value_equivalence}
\\
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}
 &=
 \lambda_{\mathrm{skill}}
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{RKL}}(\theta)
 \right|_{\theta=\theta_0}
\label{eq:gradient_equivalence}
\\
 &=
 -
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\a\sim b_i}}
 \left[
   \Delta_i(a)
   \left.
   \nabla_\theta
   \log p_{\theta,i}(a)
   \right|_{\theta=\theta_0}
 \right].
\label{eq:sampled_local_gradient}
\end{align}
\end{proposition}


\begin{proof}
Fix a valid token position \(i\). By the common-support assumption and a
change of measure from \(b_i\) to \(p_{\theta,i}\),
\begin{align*}
 &-
 \lambda_{\mathrm{skill}}
 \mathbb E_{a\sim b_i}
 \left[
   \rho_{\theta,i}(a)\Delta_i(a)
 \right]
 \\
 &\quad=
 -
 \lambda_{\mathrm{skill}}
 \sum_{v\in\mathcal V}
 b_i(v)
 \frac{p_{\theta,i}(v)}{b_i(v)}
 \bigl(
   \log q_i(v)-\log b_i(v)
 \bigr)
 \\
 &\quad=
 \lambda_{\mathrm{skill}}
 \sum_{v\in\mathcal V}
 p_{\theta,i}(v)
 \bigl(
   \log b_i(v)-\log q_i(v)
 \bigr).
\end{align*}
Adding and subtracting \(\log p_{\theta,i}(v)\) inside the summand gives
\begin{align*}
 &\lambda_{\mathrm{skill}}
 \sum_{v\in\mathcal V}
 p_{\theta,i}(v)
 \left[
   \log\frac{p_{\theta,i}(v)}{q_i(v)}
   -
   \log\frac{p_{\theta,i}(v)}{b_i(v)}
 \right]
 \\
 &\quad=
 \lambda_{\mathrm{skill}}
 \left[
   D_{\mathrm{KL}}(p_{\theta,i}\|q_i)
   -
   D_{\mathrm{KL}}(p_{\theta,i}\|b_i)
 \right].
\end{align*}
Averaging over \(i\sim\nu_b\) proves
Eq.~\ref{eq:relative_kl_decomposition}.

At \(\theta_0\), \(p_{\theta_0,i}=b_i\), and hence
\begin{equation*}
 \mathcal D_b(\theta_0)=0.
\end{equation*}
This proves Eq.~\ref{eq:value_equivalence}. Moreover,
\(\mathcal D_b\) is differentiable and attains its global minimum at
\(\theta_0\), so
\begin{equation*}
 \left.
 \nabla_\theta\mathcal D_b(\theta)
 \right|_{\theta=\theta_0}
 =0.
\end{equation*}
Differentiating Eq.~\ref{eq:relative_kl_decomposition} therefore proves
Eq.~\ref{eq:gradient_equivalence}.

Finally, because \(b_i\), \(q_i\), and \(\Delta_i\) are detached,
\begin{equation*}
\begin{aligned}
 \nabla_\theta
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 =
 -
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\a\sim b_i}}
 \left[
   \rho_{\theta,i}(a)\Delta_i(a)
   \nabla_\theta\log p_{\theta,i}(a)
 \right].
\end{aligned}
\end{equation*}
Substituting
\(\rho_{\theta_0,i}(a)=1\) proves
Eq.~\ref{eq:sampled_local_gradient}.
\end{proof}


\begin{remark}[Why the OPID skill loss is not the direct reverse-KL loss]
\label{rem:not_direct_kl}
The scaled direct reverse-KL loss is
\begin{equation*}
 \lambda_{\mathrm{skill}}
 \mathcal L_{\mathrm{RKL}}(\theta)
 =
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\v\sim p_{\theta,i}}}
 \left[
   \log p_{\theta,i}(v)-\log q_i(v)
 \right],
\end{equation*}
whereas the OPID skill loss can be written as
\begin{equation*}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 =
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\v\sim p_{\theta,i}}}
 \left[
   \log b_i(v)-\log q_i(v)
 \right].
\end{equation*}
The two expressions differ because the denominator in the detached
teacher advantage is the rollout policy \(b_i\), rather than the current
student \(p_{\theta,i}\). Importance weighting changes the sampling
distribution from \(b_i\) to \(p_{\theta,i}\), but it does not replace
\(\log b_i\) by \(\log p_{\theta,i}\). Consequently,
\begin{equation*}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 -
 \lambda_{\mathrm{skill}}
 \mathcal L_{\mathrm{RKL}}(\theta)
 =
 -
 \lambda_{\mathrm{skill}}\mathcal D_b(\theta).
\end{equation*}
Thus, the OPID skill loss is an exact relative-KL loss and only a local
surrogate for direct student--teacher reverse-KL matching.

This distinction also changes the global optimum. For example, consider
\[
 b=\left(\frac12,\frac12\right),
 \qquad
 q=\left(\frac34,\frac14\right).
\]
For a categorical distribution \(p=(p_1,p_2)\),
\begin{equation*}
 \frac{
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(p)
 }{
 \lambda_{\mathrm{skill}}
 }
 =
 p_1\log\frac{2}{3}
 +
 p_2\log 2,
\end{equation*}
which is linear in \(p\) and whose infimum is approached by concentrating
all probability mass on the first token. In contrast, the direct reverse-KL
loss is uniquely minimized at \(p=q\). Therefore, the two losses cannot be
identified globally.
\end{remark}


\begin{corollary}[First-order tightness around the behavior policy]
\label{cor:first_order_tightness}
Assume that \(p_{\theta,i}\) is twice continuously differentiable in a
neighborhood of \(\theta_0\). For \(\delta\to0\),
\begin{equation}
\begin{aligned}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}
 (\theta_0+\delta)
 &=
 \lambda_{\mathrm{skill}}
 \mathcal L_{\mathrm{RKL}}(\theta_0+\delta)
 \\
 &\quad-
 \frac{\lambda_{\mathrm{skill}}}{2}
 \delta^\top F_b\delta
 +
 o(\|\delta\|^2),
\end{aligned}
\label{eq:second_order_gap}
\end{equation}
where
\begin{equation*}
\begin{aligned}
 F_b
 &\triangleq
 \mathbb E_{\substack{i\sim\nu_b\\v\sim b_i}}
 \left[
   s_i(v)s_i(v)^\top
 \right],
 \\
 s_i(v)
 &\triangleq
 \left.
 \nabla_\theta\log p_{\theta,i}(v)
 \right|_{\theta=\theta_0}
\end{aligned}
\end{equation*}
is the behavior-policy Fisher information averaged over rollout contexts.
\end{corollary}


\begin{proof}
By Proposition~\ref{prop:relative_kl}, the discrepancy between the scaled
reverse-KL loss and the OPID skill loss is exactly
\(\lambda_{\mathrm{skill}}\mathcal D_b(\theta)\). The standard local
expansion of relative entropy around its reference distribution gives
\begin{equation*}
 \mathcal D_b(\theta_0+\delta)
 =
 \frac12
 \delta^\top F_b\delta
 +
 o(\|\delta\|^2).
\end{equation*}
Substituting this expansion into
Eq.~\ref{eq:relative_kl_decomposition} proves
Eq.~\ref{eq:second_order_gap}.
\end{proof}

Equation~\ref{eq:second_order_gap} gives the precise sense in which the
OPID skill loss is locally equivalent to reverse-KL distillation. At the
behavior policy, the two losses have the same value and gradient after
accounting for the factor \(\lambda_{\mathrm{skill}}\), while their
discrepancy is second order in the policy displacement.


\begin{corollary}[Exact recovery under a matching behavior-KL penalty]
\label{cor:matching_kl}
Consider the regularized auxiliary loss
\begin{equation}
 \mathcal L_{\mathrm{aux}}(\theta)
 \triangleq
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 +
 \beta\mathcal D_b(\theta).
\label{eq:regularized_auxiliary_loss}
\end{equation}
Then
\begin{equation}
 \mathcal L_{\mathrm{aux}}(\theta)
 =
 \lambda_{\mathrm{skill}}
 \mathcal L_{\mathrm{RKL}}(\theta)
 +
 \left(
   \beta-\lambda_{\mathrm{skill}}
 \right)
 \mathcal D_b(\theta).
\label{eq:regularized_decomposition}
\end{equation}
In particular, if
\(\beta=\lambda_{\mathrm{skill}}\), then
\begin{equation*}
 \mathcal L_{\mathrm{aux}}(\theta)
 =
 \lambda_{\mathrm{skill}}
 \mathcal L_{\mathrm{RKL}}(\theta)
\end{equation*}
for every admissible \(\theta\).
\end{corollary}


\begin{proof}
Substitute Eq.~\ref{eq:relative_kl_decomposition} into
Eq.~\ref{eq:regularized_auxiliary_loss} and collect the coefficients of
\(\mathcal D_b(\theta)\).
\end{proof}

The exact cancellation in Corollary~\ref{cor:matching_kl} requires both
(i) a KL penalty to the same behavior distribution \(b_i\), evaluated under
the ordinary context, and
(ii) the matching coefficient
\(\beta=\lambda_{\mathrm{skill}}\).
A KL penalty to a different reference distribution, or a different
coefficient, leaves the residual behavior-relative term in
Eq.~\ref{eq:regularized_decomposition} and is therefore not exactly
equivalent to direct student--teacher reverse-KL distillation.


\paragraph{Relation to the implemented PPO-clipped loss.}

The decomposition in Proposition~\ref{prop:relative_kl} applies exactly to
the unclipped skill loss
\(\mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}\).
In the implemented OPID objective, PPO clipping is applied to the combined
advantage
\[
 A_i^{\mathrm{OPID}}
 =
 A_i^{\mathrm{ep}}
 +
 \lambda_{\mathrm{skill}}\Delta_i(a_i),
\]
so the complete clipped loss does not globally decompose into independently
clipped outcome and skill losses.

Nevertheless, at
\(\theta_0=\theta_{\mathrm{old}}\),
\[
 \rho_{\theta_0,i}(a)=1.
\]
Since \(1\) lies in the interior of
\([1-\epsilon,1+\epsilon]\) for \(\epsilon>0\), the clipped and unclipped
policy losses have the same value and first derivative at the behavior
policy:
\begin{equation*}
\begin{aligned}
 \mathcal L_{\mathrm{policy}}^{\mathrm{clip}}(\theta_0)
 &=
 \mathcal L_{\mathrm{policy}}^{\mathrm{unclip}}(\theta_0),
 \\
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{policy}}^{\mathrm{clip}}(\theta)
 \right|_{\theta=\theta_0}
 &=
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{policy}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}
 \\
 &=
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{ep}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}
 +
 \left.
 \nabla_\theta
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 \right|_{\theta=\theta_0}.
\end{aligned}
\end{equation*}
Therefore,
Eq.~\ref{eq:gradient_equivalence} characterizes the skill-induced
component of the local PPO update. Once the policy ratio reaches a clipping
boundary, however, clipping couples the outcome and skill signals through
the sign of their combined advantage, and the exact relative-KL
decomposition no longer applies to the complete clipped objective.


\begin{corollary}[Non-degenerate token-level signal under reward ties]
\label{cor:dense_signal}
Fix one context \(i\), and parameterize
\(p_i=\operatorname{softmax}(z_i)\) using free categorical logits. Define
the corresponding full-action skill loss as
\begin{equation*}
 \mathcal L_{\mathrm{skill},i}^{\mathrm{unclip}}(z_i)
 \triangleq
 -
 \lambda_{\mathrm{skill}}
 \sum_{v\in\mathcal V}
 p_i(v)\Delta_i(v).
\end{equation*}
For \(\lambda_{\mathrm{skill}}>0\),
\begin{equation}
 \frac{
   \partial
   \mathcal L_{\mathrm{skill},i}^{\mathrm{unclip}}
 }{
   \partial z_i(v)
 }
 =
 -
 \lambda_{\mathrm{skill}}
 p_i(v)
 \left(
   \Delta_i(v)
   -
   \mathbb E_{u\sim p_i}[\Delta_i(u)]
 \right).
\label{eq:logit_gradient}
\end{equation}
At \(p_i=b_i\) with full support, the gradient in
Eq.~\ref{eq:logit_gradient} is zero for every \(v\) if and only if
\(q_i=b_i\).
\end{corollary}


\begin{proof}
Using
\[
 \frac{\partial p_i(u)}{\partial z_i(v)}
 =
 p_i(u)
 \left(
   \mathbf 1\{u=v\}-p_i(v)
 \right),
\]
we obtain
\begin{align*}
 \frac{
   \partial
   \mathcal L_{\mathrm{skill},i}^{\mathrm{unclip}}
 }{
   \partial z_i(v)
 }
 &=
 -
 \lambda_{\mathrm{skill}}
 \sum_u
 \Delta_i(u)
 p_i(u)
 \left(
   \mathbf 1\{u=v\}-p_i(v)
 \right)
 \\
 &=
 -
 \lambda_{\mathrm{skill}}
 p_i(v)\Delta_i(v)
 +
 \lambda_{\mathrm{skill}}
 p_i(v)
 \sum_u p_i(u)\Delta_i(u),
\end{align*}
which proves Eq.~\ref{eq:logit_gradient}.

Suppose that \(p_i=b_i\), \(b_i(v)>0\) for every \(v\), and the derivative
is zero for every \(v\). Since
\(\lambda_{\mathrm{skill}}>0\), it follows that
\(\Delta_i(v)\) is constant over the vocabulary. Hence
\[
 q_i(v)=e^c b_i(v)
\]
for some constant \(c\). Normalization of \(q_i\) and \(b_i\) implies
\(e^c=1\), and therefore \(q_i=b_i\). The converse is immediate.
\end{proof}

Corollary~\ref{cor:dense_signal} is a per-context logit statement. It shows
that even when group-relative outcome advantages vanish because all sampled
trajectories receive tied rewards, a nontrivial skill-conditioned teacher
still supplies a token-level learning signal whenever \(q_i\neq b_i\).
With shared neural parameters, gradients from different contexts may still
cancel; the result does not claim that the aggregate parameter gradient must
be nonzero.

\subsection{On-Policy Occupancy Matching for Distillation}
\label{app:occupancy_matching}

Recall that \(\nu_b\) denotes the distribution over valid token positions
induced by rollouts collected from the behavior policy. Let \(d_b\) denote
the corresponding distribution over ordinary autoregressive contexts
\(c_i\), i.e., the context marginal induced by \(i\sim\nu_b\). For an
arbitrary data-collection policy \(\mu\), let \(d_\mu\) denote the analogous
context distribution.

We define total variation as
\begin{equation*}
 \operatorname{TV}(P,Q)
 \triangleq
 \sup_A |P(A)-Q(A)|
 =
 \frac{1}{2}
 \int
 \left|
   \mathrm dP-\mathrm dQ
 \right|.
\end{equation*}

The following result isolates the effect of changing only the distribution
of ordinary autoregressive contexts. It applies to both nonnegative
distillation losses and signed surrogate losses.

\begin{proposition}[On-policy occupancy matching]
\label{prop:occupancy_matching}
Let
\(\ell_\theta:\mathcal C\rightarrow[m_\ell,M_\ell]\)
be a measurable per-context loss, where
\(-\infty<m_\ell<M_\ell<+\infty\). Then
\begin{equation}
\begin{aligned}
 &
 \left|
   \mathbb E_{c\sim d_b}
   \left[
     \ell_\theta(c)
   \right]
   -
   \mathbb E_{c\sim d_\mu}
   \left[
     \ell_\theta(c)
   \right]
 \right|
 \\
 &\qquad\le
 \left(
   M_\ell-m_\ell
 \right)
 \operatorname{TV}(d_b,d_\mu)
 \\
 &\qquad\le
 \left(
   M_\ell-m_\ell
 \right)
 \sqrt{
   \frac{1}{2}
   D_{\mathrm{KL}}(d_b\|d_\mu)
 }.
\end{aligned}
\label{eq:occupancy_bound}
\end{equation}
In particular, if \(d_\mu=d_b\), then the context-occupancy mismatch is
exactly zero.
\end{proposition}

\begin{proof}
Define
\begin{equation*}
 f_\theta(c)
 \triangleq
 \frac{
   \ell_\theta(c)-m_\ell
 }{
   M_\ell-m_\ell
 }.
\end{equation*}
Then \(0\le f_\theta(c)\le 1\). By the variational characterization of
total variation over measurable functions with range in \([0,1]\),
\begin{equation*}
 \left|
   \mathbb E_{d_b}[f_\theta]
   -
   \mathbb E_{d_\mu}[f_\theta]
 \right|
 \le
 \operatorname{TV}(d_b,d_\mu).
\end{equation*}
Multiplying both sides by \(M_\ell-m_\ell\) proves the first inequality in
Eq.~\ref{eq:occupancy_bound}. The second inequality follows from
Pinsker's inequality. If \(d_b\) is not absolutely continuous with respect
to \(d_\mu\), then
\(D_{\mathrm{KL}}(d_b\|d_\mu)=+\infty\), and the inequality remains valid
in the extended-real sense. Setting \(d_\mu=d_b\) proves the final
statement.
\end{proof}

For example, Proposition~\ref{prop:occupancy_matching} can be applied to
the per-context reverse-KL loss
\begin{equation*}
 \ell_{\mathrm{RKL},\theta}(c_i)
 \triangleq
 D_{\mathrm{KL}}
 \left(
   p_{\theta,i}\|q_i
 \right),
\end{equation*}
which is the distribution-matching loss locally approximated by the OPID
skill update. It can also be applied to a bounded version of the signed
per-context OPID skill loss
\begin{equation}
\begin{aligned}
 \ell_{\mathrm{skill},\theta}^{\mathrm{unclip}}(c_i)
 &\triangleq
 -
 \lambda_{\mathrm{skill}}
 \mathbb E_{a\sim b_i}
 \left[
   \rho_{\theta,i}(a)\Delta_i(a)
 \right]
 \\
 &=
 \lambda_{\mathrm{skill}}
 \left[
   D_{\mathrm{KL}}
   \left(
     p_{\theta,i}\|q_i
   \right)
   -
   D_{\mathrm{KL}}
   \left(
     p_{\theta,i}\|b_i
   \right)
 \right].
\end{aligned}
\label{eq:per_context_opid_skill_loss}
\end{equation}
Because the loss in Eq.~\ref{eq:per_context_opid_skill_loss} is signed
and need not be uniformly bounded for arbitrary probability distributions,
applying Proposition~\ref{prop:occupancy_matching} to it requires an
explicit bounded-range condition, such as probability flooring, log-ratio
clipping, or restriction to a compact parameter neighborhood. More general
versions can instead be obtained under appropriate moment or tail
conditions.

Proposition~\ref{prop:occupancy_matching} controls only the mismatch in the
outer distribution of ordinary autoregressive contexts. It assumes that
the same per-context loss map is evaluated under \(d_b\) and \(d_\mu\).
It does not by itself control changes in the hindsight skill, the routed
teacher \(q_i\), or other trajectory-dependent quantities that may also
change with the data-collection policy.
\subsection{Critical-First Hierarchical Routing}
\label{app:routing}

We next formalize how the episode-level and step-level skills determine the
detached teacher \(q_i\) used in
\(\mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}\).

Let \(q_i^\star\) denote an ideal privileged teacher at token position
\(i\). Let \(q_i^{\mathrm{ep}}\) and \(q_i^{\mathrm{step}}\) denote the
teachers induced by the episode-level and step-level skills, respectively.
Let
\begin{equation*}
 z_i^\star\in\{0,1\}
\end{equation*}
be an oracle criticality indicator, where \(z_i^\star=1\) means that the
step-level teacher is the appropriate specialized teacher. The analyzer
prediction is
\begin{equation*}
 \widehat z_i
 \triangleq
 \mathbf 1\{t\in C_\tau\}.
\end{equation*}

The critical-first routing rule defines
\begin{equation}
 q_i^{\mathrm{route}}
 \triangleq
 \widehat z_i q_i^{\mathrm{step}}
 +
 \left(
   1-\widehat z_i
 \right)
 q_i^{\mathrm{ep}},
 \qquad
 q_i\equiv q_i^{\mathrm{route}}.
\label{eq:routed_teacher}
\end{equation}
Thus, the \(q_i\) appearing in the OPID skill advantage
\(\Delta_i(v)=\log q_i(v)-\log b_i(v)\) is precisely the routed teacher in
Eq.~\ref{eq:routed_teacher}.

Measure the approximation errors of the two candidate teachers by
\begin{equation}
\begin{aligned}
 \mathcal E_i^{\mathrm{ep}}
 &\triangleq
 D_{\mathrm{KL}}
 \left(
   q_i^\star\|q_i^{\mathrm{ep}}
 \right),
 \\
 \mathcal E_i^{\mathrm{step}}
 &\triangleq
 D_{\mathrm{KL}}
 \left(
   q_i^\star\|q_i^{\mathrm{step}}
 \right),
 \\
 \mathcal E_i^{\mathrm{route}}
 &\triangleq
 D_{\mathrm{KL}}
 \left(
   q_i^\star\|q_i^{\mathrm{route}}
 \right).
\end{aligned}
\label{eq:routing_errors}
\end{equation}
Because the routing decision is hard,
\begin{equation*}
 \mathcal E_i^{\mathrm{route}}
 =
 \widehat z_i
 \mathcal E_i^{\mathrm{step}}
 +
 \left(
   1-\widehat z_i
 \right)
 \mathcal E_i^{\mathrm{ep}}.
\end{equation*}

\begin{proposition}[Routing optimality and detector-error regret]
\label{prop:routing_regret}
Assume that the episode-level and step-level teachers specialize according
to the oracle criticality label:
\begin{equation}
\begin{aligned}
 z_i^\star=1
 &\implies
 \mathcal E_i^{\mathrm{step}}
 \le
 \mathcal E_i^{\mathrm{ep}},
 \\
 z_i^\star=0
 &\implies
 \mathcal E_i^{\mathrm{ep}}
 \le
 \mathcal E_i^{\mathrm{step}}.
\end{aligned}
\label{eq:routing_specialization}
\end{equation}
Then, pointwise,
\begin{equation}
\begin{aligned}
 \mathcal E_i^{\mathrm{route}}
 &=
 \min
 \left\{
   \mathcal E_i^{\mathrm{ep}},
   \mathcal E_i^{\mathrm{step}}
 \right\}
 \\
 &\quad+
 \mathbf 1
 \left\{
   \widehat z_i\ne z_i^\star
 \right\}
 \left|
   \mathcal E_i^{\mathrm{ep}}
   -
   \mathcal E_i^{\mathrm{step}}
 \right|.
\end{aligned}
\label{eq:routing_regret_identity}
\end{equation}
Consequently, if
\begin{equation}
 \left|
   \mathcal E_i^{\mathrm{ep}}
   -
   \mathcal E_i^{\mathrm{step}}
 \right|
 \le
 \Gamma
\label{eq:routing_gap_bound}
\end{equation}
almost surely under \(i\sim\nu_b\), then
\begin{equation}
\begin{aligned}
 \mathbb E_{i\sim\nu_b}
 \left[
   \mathcal E_i^{\mathrm{route}}
 \right]
 &\le
 \min
 \left\{
   \mathbb E_{i\sim\nu_b}
   \left[
     \mathcal E_i^{\mathrm{ep}}
   \right],
   \mathbb E_{i\sim\nu_b}
   \left[
     \mathcal E_i^{\mathrm{step}}
   \right]
 \right\}
 \\
 &\quad+
 \Gamma
 \Pr_{i\sim\nu_b}
 \left(
   \widehat z_i\ne z_i^\star
 \right).
\end{aligned}
\label{eq:routing_expected_bound}
\end{equation}
Under perfect criticality detection,
\begin{equation*}
 \widehat z_i=z_i^\star
 \qquad
 \text{almost surely},
\end{equation*}
and therefore
\begin{equation*}
 \mathcal E_i^{\mathrm{route}}
 =
 \min
 \left\{
   \mathcal E_i^{\mathrm{ep}},
   \mathcal E_i^{\mathrm{step}}
 \right\}
\end{equation*}
pointwise, with
\begin{equation*}
\begin{aligned}
 \mathbb E_{i\sim\nu_b}
 \left[
   \mathcal E_i^{\mathrm{route}}
 \right]
 \le
 \min
 \left\{
   \mathbb E_{i\sim\nu_b}
   \left[
     \mathcal E_i^{\mathrm{ep}}
   \right],
   \mathbb E_{i\sim\nu_b}
   \left[
     \mathcal E_i^{\mathrm{step}}
   \right]
 \right\}.
\end{aligned}
\end{equation*}
\end{proposition}

\begin{proof}
Consider first the event
\(\widehat z_i=z_i^\star\). Under
Eq.~\ref{eq:routing_specialization}, the routing rule selects a teacher
with the smaller approximation error. Hence
\begin{equation*}
 \mathcal E_i^{\mathrm{route}}
 =
 \min
 \left\{
   \mathcal E_i^{\mathrm{ep}},
   \mathcal E_i^{\mathrm{step}}
 \right\}.
\end{equation*}
The second term in Eq.~\ref{eq:routing_regret_identity} is zero on this
event.

On the event
\(\widehat z_i\ne z_i^\star\), the routing rule selects the nonspecialized
teacher. Its excess error over the oracle choice is exactly
\begin{equation*}
 \left|
   \mathcal E_i^{\mathrm{ep}}
   -
   \mathcal E_i^{\mathrm{step}}
 \right|.
\end{equation*}
This proves Eq.~\ref{eq:routing_regret_identity}.

Taking expectations yields
\begin{equation*}
\begin{aligned}
 \mathbb E_{i\sim\nu_b}
 \left[
   \mathcal E_i^{\mathrm{route}}
 \right]
 &=
 \mathbb E_{i\sim\nu_b}
 \left[
   \min
   \left\{
     \mathcal E_i^{\mathrm{ep}},
     \mathcal E_i^{\mathrm{step}}
   \right\}
 \right]
 \\
 &\quad+
 \mathbb E_{i\sim\nu_b}
 \left[
   \mathbf 1
   \left\{
     \widehat z_i\ne z_i^\star
   \right\}
   \left|
     \mathcal E_i^{\mathrm{ep}}
     -
     \mathcal E_i^{\mathrm{step}}
   \right|
 \right].
\end{aligned}
\end{equation*}
Using
\begin{equation*}
 \mathbb E[\min\{X,Y\}]
 \le
 \min\{\mathbb E[X],\mathbb E[Y]\}
\end{equation*}
and Eq.~\ref{eq:routing_gap_bound} proves
Eq.~\ref{eq:routing_expected_bound}. The perfect-detection statements
follow by setting
\(\Pr_{i\sim\nu_b}(\widehat z_i\ne z_i^\star)=0\).
\end{proof}

Proposition~\ref{prop:routing_regret} separates the two requirements behind
critical-first routing: teacher specialization and criticality-detection
accuracy. Under specialization, perfect detection recovers the oracle
pointwise choice between the two candidate teachers. With imperfect
detection, the excess teacher-approximation error is controlled jointly by
the detector error probability and the difference between the two candidate
teacher errors.

The criterion in Eq.~\ref{eq:routing_errors} measures the quality of a
candidate teacher relative to \(q_i^\star\). It is distinct from the
student--teacher reverse-KL loss
\(D_{\mathrm{KL}}(p_{\theta,i}\|q_i)\) appearing in
\(\mathcal L_{\mathrm{RKL}}\). Therefore, without additional assumptions
relating the candidate teachers' likelihood ratios, the routing result
should not be interpreted as a direct upper bound on
\(\mathcal L_{\mathrm{RKL}}\).


\subsection{Summary}
\label{app:theory_scope}

Proposition~\ref{prop:relative_kl} analyzes the unclipped skill component
of the OPID policy loss:
\begin{equation*}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 =
 -
 \lambda_{\mathrm{skill}}
 \mathbb E_{\substack{i\sim\nu_b\\a\sim b_i}}
 \left[
   \rho_{\theta,i}(a)\Delta_i(a)
 \right].
\end{equation*}
Conditioned on fixed rollout histories, routed skills, and detached
distributions \(b_i\) and \(q_i\), this loss has the exact decomposition
\begin{equation*}
 \mathcal L_{\mathrm{skill}}^{\mathrm{unclip}}(\theta)
 =
 \lambda_{\mathrm{skill}}
 \left[
   \mathcal L_{\mathrm{RKL}}(\theta)
   -
   \mathcal D_b(\theta)
 \right].
\end{equation*}

Proposition~\ref{prop:occupancy_matching}  shows that collecting the ordinary autoregressive
contexts on policy eliminates the outer context-distribution mismatch:
when the collection distribution equals the behavior-policy distribution,
\(d_\mu=d_b\), the occupancy term in
Eq.~\ref{eq:occupancy_bound} is zero. 

Proposition~\ref{prop:routing_regret} analyzes how the teacher \(q_i\) is
selected from episode-level and step-level candidates. Under the stated
specialization assumption, critical-first routing recovers the lower-error
candidate under perfect detection, while the degradation under imperfect
detection is controlled by
\begin{equation*}
 \Gamma
 \Pr_{i\sim\nu_b}
 \left(
   \widehat z_i\ne z_i^\star
 \right).
\end{equation*}

Taken together, the three results establish that:

\begin{enumerate}
 \item The unclipped OPID skill loss is an exact relative-KL loss and is
 first-order equivalent to scaled reverse-KL distillation at the behavior
 policy;

 \item On-policy collection removes the mismatch in the outer distribution
 of ordinary autoregressive contexts; and

 \item Critical-first routing approaches the oracle candidate-teacher
 selection when the candidate teachers specialize and the criticality
 detector is accurate.
\end{enumerate}