\section{Introduction}
\label{sec:intro}
% (Align to https://arxiv.org/pdf/2601.20540)


High-fidelity three-dimensional geospatial reconstruction of the Earth's surface has emerged as a foundational pillar for modern digital twin infrastructures, smart city logistics, and virtual simulation ecosystems. In particular, aerial-view 3D representations provide critical geospatial priors for rapid disaster response, urban planning, and robotic exploration. Despite its immense utility, traditional large-scale 3D reconstruction pipelines~\cite{schoenberger2016sfm,schoenberger2016mvs,nex2014uav,WEHR199968,shan2018topographic}, primarily built on dense oblique photogrammetry and LiDAR scanning, are fundamentally constrained by extreme data acquisition costs, prolonged processing latencies, and high computational barriers, rendering real-time or on-demand planetary-scale modeling a long-standing challenge.

As a compelling alternative, generative 3D modeling offers a way to bypass these physical constraints by shifting the technical burden from exhaustive multi-view acquisition to learned structural priors. By leveraging implicit geometric knowledge, generative paradigms can synthesize complete 3D structures from highly sparse inputs, drastically reducing both data collection overhead and optimization latency. Under this paradigm, 3D generative modeling has reached remarkable maturity at the individual object scale, enabling the rapid, high-fidelity synthesis of diverse single assets~\cite{xiang2025native,xiang2025structured,CLAY,hunyuan3d22025tencent,lai2025hunyuan3d25highfidelity3d,Seed3D,gao2022get3d,jun2023shapegeneratingconditional3d}. Inspired by these object-level successes, the generative AI community has actively embarked on scaling these capabilities to the far more challenging domain of unbounded, large-scale outdoor scenes~\cite{earthcrafter,CityDreamer,sat2city,lin2023infinicity,sat2scene,Yang_2023_ICCV,Qian_2023_Sat2Density,sat2densitypp,qian2026sat3dgen,infinite_nature_2020}. 

However, directly migrating object-centric formulations~\cite{xiang2025native,xiang2025structured,CLAY,hunyuan3d22025tencent,lai2025hunyuan3d25highfidelity3d,Seed3D} to large-scale scene synthesis remains highly non-trivial. A major limitation of existing outdoor generators is their heavy reliance on synthetic virtual assets or unconstrained imaginary scene hallucination~\cite{Xie_2025_CVPR,CityDreamer,lin2023infinicity,sat2city,sat2scene,earthcrafter}. Because these generated environments are inherently artificial and lack real-world physical and geospatial authenticity, they fail to bridge the severe sim-to-real domain gap~\cite{8202133}, making them impractical for rigorous downstream simulation and real-world transfer. To address this issue, we advocate for generating native 3D scenes trained directly on high-quality real-world reconstructions. Among various 3D representations, 3DGS~\cite{3DGS} has revolutionized photometric reconstruction of real-world environments with its unmatched rendering fidelity, natively capturing complex non-manifold topologies like dense foliage, building facades, and specular water surfaces~\cite{liu2024citygaussian}. 

To scale this real-world-trained generative paradigm to unbounded, planetary-wide extents, establishing a scalable and globally consistent conditioning signal is paramount. In this context, globally ubiquitous and geospatially referenced remote sensing imagery serves as the ideal conditioning blueprint. By establishing a direct geospatial correspondence to the underlying terrain at virtually any coordinates on Earth, leveraging this pervasive modality unlocks the potential for infinite-scale, progressive 3D generation. In this work, we present \method{}, a generative 3D framework designed to realize this vision, enabling the synthesis of vast, near-seamless 3D aerial scenes conditioned on standard satellite imagery, without requiring knowledge of its precise acquisition angles or multi-view overlaps. By framing 3D generation directly in a native 3DGS generative space, \method{} preserves exceptional visual quality while achieving a scalable generation rate of under 10 minutes per square kilometer. Extensive quantitative and qualitative evaluations demonstrate that \method{} outperforms state-of-the-art baselines by a significant margin, exhibiting a clear superiority in rendering realism and geometric fidelity. To showcase its performance and versatility, the key capabilities of \method{} are highlighted as follows:

\begin{itemize}
    \item \textbf{Generation of Real-World Geospatial Complexity.} \method{} achieves remarkable improvements in capturing real-world complexity without relying on artificial heuristics or synthetic assets. Trained directly on diverse real-world urban reconstructions, the model is capable of robustly synthesizing highly detailed structures, such as intricate building facades, dense vegetation canopy, and coherent road networks, with natural textures. We successfully validate the model's performance across the vast majority of global metropolitan areas as well as numerous non-urban natural terrains, demonstrating its robust Earth-scale generalizability. Crucially, as a native 3DGS-based generative framework, our synthesized environments can be seamlessly integrated and co-edited with precisely reconstructed, high-fidelity 3DGS landmark models. This composability allows users to insert meticulously scanned real-world landmarks into generated contextual backdrops, yielding an exceptionally immersive, hybrid-reality experience.
    
    \item \textbf{Planetary-Scale Online Exploration via Native Multi-LOD.} To support seamless online exploration across vast geographical extents, \method{} natively generates hierarchical level-of-detail (multi-LOD) outputs. When paired with our customized, YunJing-based 3DGS visualization engine, this capability enables dynamic, viewport-dependent tile scheduling and streaming of trillion-scale Gaussian primitives. Users can seamlessly zoom in from a planetary overview to fine-grained, street-level details with fluid, interactive frame rates. This natively integrated multi-LOD pipeline avoids the need for expensive post-reconstruction downsampling, providing a smooth and continuous interactive experience for global-scale digital earth and geographic information systems.
    
    \item \textbf{Simulation-Ready Capabilities for Downstream Applications.} Grounded in physical and photometric realism, the 3D environments synthesized by \method{} are not merely static visual models, but interaction-ready virtual sandboxes. By generating authentic 3D geometry and multi-view consistent textures, the model establishes a high-fidelity closed-loop simulation and training platform for Embodied AI, particularly for unmanned aerial vehicle (UAV) navigation, obstacle avoidance, and control~\cite{AirSim,FlightGoggles}. Ultimately, by providing an ultra-low-cost, high-efficiency generation pipeline, \method{} lowers the technical and financial barriers to large-scale 3D reconstruction, empowering global researchers and enterprises to realize scalable, cost-effective geospatial applications in smart city planning, environmental monitoring, and rapid disaster response.
\end{itemize}

Through this unified generative framework, \method{} effectively mitigates the synthetic-to-real domain gap and delivers robust generalization across diverse environments worldwide. Ultimately, we hope this ultra-low-cost, highly efficient solution will break down technological and financial barriers, bridging the ``3D digital divide'' and fostering global geospatial technological equity, thereby empowering scalable 3D digital earth visualization and simulation on a global scale.
