\section{Conclusion}
\label{sec:conclusion}


The core value of ABot-Earth is to democratize 3D content production, transforming it from a high-cost, specialized process into a low-barrier, generative workflow. It rapidly generates foundational 3D layers for urban digital twins, significantly accelerating project timelines, and fills the void in unmapped regions globally, enabling the low-cost launch of 3D map services. Furthermore, ABot-Earth provides a crucial spatial prior for intelligent systems such as drones, enabling advanced navigation and simulation in complex environments. It also turns physical locations into versatile creative assets, allowing for the easy generation of a scenic area's seasonal variations, a city's future developments, or narrative-driven spaces for individual creators. These applications all converge on a single trend: the future of the digital Earth will transcend being a passive viewing tool to become a dynamic spatial intelligence platform—one that is generative, simulatable, and editable. ABot-Earth marks the beginning of this evolution.

Looking ahead, our next step is to bring this technology from the sky down to the ground~\cite{qian2026sat3dgen}. We are working to transition from our current aerial-level 3D to street-view level detail, while simultaneously exploring greater scene diversity and achieving reconstruction-grade fidelity in our outputs. Concurrently, we aim to systematically validate the scaling laws that govern outdoor 3D scene generation. Our vision is for ABot-Earth to become the foundational layer for the future 3D world, ensuring that a new generation of applications, from digital twins to robotics simulation, can all benefit and build upon it.