\documentclass[]{bytedance_seed}

% single-column: \documentclass[]{bytedance_seed}, 
%Please prioritize using single-column。

% twocolumn: \documentclass[twocolumn]{bytedance_seed}

\usepackage[toc,page,header]{appendix}

%%%%%%%%%%%%%%%%patches for this bytedance template%%%%%%%%%%%%%%%%%%%%
\titleformat*{\paragraph}{\normalsize\sffamily\bfseries}
\let\oldemph\emph
\renewcommand{\emph}[1]{\textit{#1}}
\usepackage{wrapfig}
\captionsetup[subtable]{
  format=plain,
  labelformat=parens,
  labelsep=space,
  font=footnotesize,
  labelfont=normalfont,
  textfont=normalfont,
  justification=centering,
  skip=1pt
}
%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%copy this for a new template%%%%%%%%%%%%%%%%%%%%
\usepackage{url}
\usepackage{nicefrac}
\usepackage{minitoc}
\usepackage{enumitem}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{paralist}
\usepackage{array}
\usepackage{makecell}
\usepackage{wrapfig}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{tabularx}
\usepackage{adjustbox}
\setcitestyle{square,numbers,comma}
\usepackage{float}


%%%%%%%%%%%%%%%%%%%%
\newcommand{\eg}{e.g.\@\xspace}
\newcommand{\ie}{i.e.\@\xspace}

\newcolumntype{C}[1]{>{\centering\arraybackslash}m{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}m{#1}}
\definecolor{posgreen}{RGB}{0,128,64}
\definecolor{negred}{RGB}{180,50,50}
\newcommand{\posdiff}[1]{{\color{posgreen}{+#1}}}
\newcommand{\negdiff}[1]{{\color{negred}{-#1}}}
\definecolor{MidnightBlue}{RGB}{25,25,112}

\newcommand{\modelname}{\textsc{\textbf{MMProLong}}\xspace}
\newcommand{\extractsingle}{\texttt{extract-single}\xspace}
\newcommand{\extractmulti}{\texttt{extract-multi}\xspace}
\newcommand{\reasoningtask}{\texttt{reasoning}\xspace}
\newcommand{\ocrfull}{\texttt{OCR-full}\xspace}
\newcommand{\ocrneedle}{\texttt{OCR-needle}\xspace}
\newcommand{\promptplaceholder}[1]{\colorbox{gray!12}{\textcolor{MidnightBlue}{\ttfamily #1}}}


\title{Training Long-Context Vision-Language Models Effectively with Generalization Beyond 128K Context}

\author[1,2, *]{Zhaowei~Wang}
\author[2]{Lishu~Luo}
\author[2]{Haodong~Duan}
\author[2]{Weiwei~Liu}
\author[2]{Sijin~Wu}
\author[2]{Ji~Luo}
\author[2]{Shen~Yan}
\author[2]{Shuai~Peng}
\author[2]{Sihang~Yuan}
\author[2]{Chaoyi~Huang}
\author[2, \dagger]{Yi~Lin}
\author[1, \dagger]{Yangqiu~Song}


%论文单位请使用ByteDance Seed
\affiliation[1]{CSE Department, HKUST}
\affiliation[2]{ByteDance Seed}


\contribution[*]{Work done at ByteDance Seed}
\contribution[\dagger]{Corresponding authors}

\abstract{
Long-context modeling is becoming a core capability of modern large vision-language models (LVLMs), enabling sustained context management across long-document understanding, video analysis, and multi-turn tool use in agentic workflows. 
Yet practical training recipes remain insufficiently explored, 
particularly for designing and balancing long-context data mixtures.
In this work, we present a systematic study of \emph{long-context continued pre-training} for LVLMs, extending a 7B model from 32K to 128K context with extensive ablations on long-document data. 
We first show that long-document VQA is substantially more effective than OCR transcription.
Building on this observation, our ablations further yield three key findings:
\begin{inparaenum}[i)]
\item for sequence-length distribution, balanced data outperforms target-length-focused data (e.g., 128K), suggesting that long-context ability requires generalizable key-information retrieval across various lengths and positions;
\item retrieval remains the primary bottleneck, favoring retrieval-heavy mixtures with modest reasoning data for task diversity;
\item pure long-document VQA largely preserves short-context capabilities, suggesting that instruction-formatted long data reduces the need for short-data mixing.
\end{inparaenum}
Based on these findings, we introduce \modelname, obtained by long-context continued pre-training from Qwen2.5-VL-7B with only a 5B-token budget. 
\modelname improves long-document VQA scores by 7.1\% and maintains strong performance at 256K and 512K contexts beyond its 128K training window, without additional training.
It further generalizes to webpage-based multimodal needle retrieval, long-context vision-text compression, and long-video understanding without task-specific supervision.
Overall, our study establishes a practical LongPT recipe and an empirical foundation for advancing long-context vision-language models.
}

%\date{\today}
\checkdata[Email]{\texttt{\{zwanggy, yqsong\}@cse.ust.hk};\ \texttt{linyi.james@bytedance.com}}

% You can add additional info fields as follows 
\checkdata[Project Page]{coming soon} %\url{xxx}

\begin{document}
\maketitle

%不需要目录就注释掉 注意目录不要和第一页放在一块 要有\newpage
%\newpage
%\tableofcontents
%\newpage

\input{sections/1_introduction}
\input{sections/2_relatedwork}
\input{sections/3_setup}
\input{sections/4_curation}
\input{sections/5_data_mixture_and_training_design}
\input{sections/6_final_results_and_generalization}

\section{Conclusion}
In this work, we presented a systematic study of long-context continued pre-training for LVLMs, focusing on how to construct and mix effective multimodal long-context data.
Our experiments show that long-document VQA is a strong and practical training task, as it provides diverse retrieval and reasoning supervision signals, preserves short-context capability, and supports context extension from 32K to 128K under a modest token budget.
Instantiated as \modelname, our recipe improves long-document VQA performance and generalizes beyond the training context window to 256K and 512K context lengths, as well as broader multimodal long-context tasks, such as long-video understanding.
We hope our study provides a practical foundation for building future LVLMs with reliable long-context capability.
\clearpage

\clearpage

\bibliographystyle{unsrtnat}
\bibliography{reference}

\clearpage
\beginappendix
\crefname{appendix}{Appendix}{Appendices}
\Crefname{appendix}{Appendix}{Appendices}
\crefalias{section}{appendix}
\crefalias{subsection}{appendix}
\crefalias{subsubsection}{appendix}
\input{sections/7_appendix}

\end{document}