\section{Introduction}
\label{sec:intro}

\section{Introduction}

The ultimate goal of robotics is to build general embodied agents that operate across diverse robotic bodies, realizing the vision of ``one brain, many forms''~\cite{rt2,liu2026rdt2,cheang2025gr}. While Large Language Models (LLMs) have achieved universality in the digital realm, embodied intelligence remains fragmented—policies trained on one robot rarely transfer to another, and high-level semantic reasoning struggles to translate into low-level physical precision. We ask a critical question: \textit{Why has the ``ChatGPT Moment'' for robotics not yet arrived?}
% 机器人学的终极目标是构建能够跨多种机器人本体运行的通用具身智能体，实现“一脑多形”（one brain, many forms）的愿景~\cite{rt2,liu2026rdt2,cheang2025gr}。尽管大语言模型（LLMs）已在数字领域实现普适性，具身智能却仍处于碎片化状态——针对单一机器人训练的策略难以迁移到其他平台，高层语义推理也难以转化为底层物理操作的精确控制。我们提出一个关键问题：为何机器人领域尚未迎来属于自己的“ChatGPT 时刻”？

We argue that the field is currently blocked by two invisible walls: the \textit{Curse of Data Heterogeneity} and the \textit{Manifold Misalignment}. Unlike the standardized text data that fueled the LLM revolution, robotic data is trapped in silos. High costs and diverse hardware create a ``Tower of Babel'', where the open-source community possesses vast but fragmented data, while proprietary models build moats around standardized but closed ecosystems. Furthermore, existing Vision-Language-Action (VLA) models often treat actions as unstructured noise (e.g., diffusion noise prediction~\cite{GR00T}), ignoring the physical reality that valid motions follow strict kinematic laws rather than random distributions.
% 我们认为，当前领域被两堵“隐形高墙”所阻隔：数据异构性诅咒（Curse of Data Heterogeneity）与流形错位（Manifold Misalignment）。与推动LLM革命的标准化文本数据不同，机器人数据深陷孤岛困境。高昂成本与硬件多样性构筑起一座“巴别塔”：开源社区虽拥有海量数据，却高度碎片化；而私有模型则围绕标准化但封闭的生态系统筑起护城河。此外，现有视觉-语言-动作（VLA）模型常将动作视为无结构的噪声（如基于扩散的噪声预测~\cite{GR00T}），忽视了物理现实——有效运动遵循严格的运动学规律，而非随机分布。

To shatter these barriers and democratize generalist robotics, we introduce \VLAFM{}, a unified foundation model built on the philosophy of ``United We Scale.'' Our primary contribution is not merely a model, but a reproducible \textbf{Open-Source Data Curation Pipeline} designed to break the data monopoly. We demonstrate that the community does not need to wait for a single massive dataset release; we already possess the critical mass—if properly refined. We systematically integrate six major disparate open-source datasets—including OXE~\cite{o2024oxe}, OXE-AugE~\cite{ji2025oxeauge}, Agibot-Beta~\cite{agibotbeta}, and RoboCoin~\cite{wu2025robocoin}—into a coherent training set of over \textbf{6 million trajectories}, the largest non-proprietary collection to date. Crucially, we establish a rigorous standardization protocol: unifying action spaces into delta-end-effector poses, encoding rotations via rotation vectors to avoid singularities, and applying a pad-to-dual strategy to support heterogeneous morphologies. This pipeline transforms scattered ``data fragments'' into high-octane fuel, proving that with rigorous curation, community-aggregated data can rival proprietary data moats.
% 为打破这些壁垒、推动通用机器人技术的民主化，我们提出 \VLAFM{}，一个基于“聚力方能规模化”（United We Scale）理念构建的统一基础模型。我们的核心贡献不仅是一个模型，更是一套可复现的开源数据整理管道，旨在瓦解数据垄断。我们证明：社区无需等待单一超大规模数据集的发布——我们已具备关键数据体量，只需加以精炼。我们系统性整合六大异构开源数据集（包括 OXE~\cite{o2024oxe}、OXE-AugE~\cite{ji2025oxeauge}、Agibot-Beta~\cite{agibotbeta} 与 RoboCoin~\cite{wu2025robocoin}），构建出包含逾600万条轨迹的统一训练集，成为迄今规模最大的非私有数据集合。尤为关键的是，我们建立了一套严格标准化协议：将动作空间统一为末端执行器位姿增量（delta-end-effector poses），采用旋转矢量（rotation vectors）编码旋转以避免奇异性，并引入“填充至双臂”（pad-to-dual）策略以支持异构形态。该管道将零散的“数据碎片”转化为高能燃料，证明经严谨整理的社区聚合数据足以媲美私有数据护城河。

Building on this unified foundation, we propose the \textbf{Action Manifold Hypothesis} to resolve the modeling misalignment. We posit that successful robotic actions do not scatter randomly in high-dimensional space but reside on a low-dimensional, smooth manifold shaped by physics and task constraints. Instead of the prevailing practice of fitting unstructured noise, we introduce \textbf{Action Manifold Learning (AML)}. This mechanism guides our DiT-based policy to project visual intent directly onto this feasible action manifold, filtering out physical jitter and significantly accelerating inference by reducing output uncertainty. To further bridge the gap between 2D semantics and 3D execution, we incorporate plug-and-play spatial experts, such as VGGT~\cite{wang2025vggt} and Qwen-Image-Edit~\cite{wu2025qwenimagetechnicalreport}, which inject scene-level structural priors without altering the core architecture.
% 基于这一统一基础，我们提出动作流形假设（Action Manifold Hypothesis）以解决建模错位问题。我们主张：成功的机器人动作并非在高维空间中随机散布，而是栖居于一个由物理规律与任务约束塑造的低维、平滑流形之上。我们摒弃拟合无结构噪声的主流做法，引入动作流形学习（Action Manifold Learning, AML）。该机制引导基于扩散Transformer（DiT）的策略将视觉意图直接投影至可行动作流形，滤除物理抖动，并通过降低输出不确定性显著加速推理。为弥合2D语义与3D执行间的鸿沟，我们进一步集成即插即用的空间专家模块（如 VGGT~\cite{wang2025vggt} 与 Qwen-Image-Edit~\cite{wu2025qwenimagetechnicalreport}），在不改动核心架构的前提下注入场景级结构先验。

Our results confirm that scale and structure must go hand in hand. Under a consistent training setup, \VLAFM{} achieves state-of-the-art performance, recording average success rates of \textbf{98.6\%} on LIBERO~\cite{liu2023libero}, 80.5\% on LIBERO-Plus~\cite{fei25libero-plus}, and 58.3\% on RoboCasa GR1 tasks~\cite{GR00T}, significantly outperforming strong baselines like $\pi_{0.5}$~\cite{pi_0.5} and OpenVLA~\cite{kim2024openvla}. These gains remain consistent across platforms, suggesting that our standardized representations generalize beyond dataset-specific artifacts. We will fully release our data curation pipeline and training code, providing a blueprint for the community to collaboratively build a shared, evolving, and truly open general-purpose robotic brain.
% 实验结果证实：规模与结构必须协同演进。\VLAFM{} 在统一训练设置下取得当前最优性能，在 LIBERO~\cite{liu2023libero}、LIBERO-Plus~\cite{fei25libero-plus} 与 RoboCasa GR1 任务~\cite{GR00T} 上分别达到 98.6%、80.5% 与 58.3% 的平均成功率，显著超越 
 % ~\cite{pi_0.5} 与 OpenVLA~\cite{kim2024openvla} 等强基线模型。性能增益在不同平台间保持一致，表明我们的标准化表征可泛化至数据集特有伪影之外。我们将完整开源数据整理管道与训练代码，为社区协作构建一个共享、演进且真正开放的通用机器人“大脑”提供蓝图。

