
% \\
% \\
% \section{Limitations \& Conclusion}
\section{Conclusion}
We presented ABot-AgentOS, a general-purpose robotic agent operating
system that connects foundation VLM/VLA models with physical robot
execution through a deliberative agent layer. Its modular Agent Harness separates
reasoning, context management, skill execution, and verification, while
its multi-modal memory system stores embodied experience as a typed,
source-grounded graph for traceable retrieval. We also introduced
EmbodiedWorldBench, an executable benchmark for long-horizon embodied
tasks across indoor, outdoor, and hybrid scenarios with dynamic events,
NPC interaction, and deterministic scoring.

The results indicate that ABot-AgentOS is useful both as an online
reasoning-execution layer and as a persistent learning substrate. The
graph memory performs strongly across long-term conversational recall,
embodied QA, multi-modal memory, video reasoning, and egocentric
daily-life QA, while failure-driven self-evolution improves later splits
through gated, trace-diagnosed evo-assets. Initial agent evaluations on
an EmbodiedWorldBench subset further suggest that hierarchical control,
task memory, skill feedback, and finish-time verification improve
long-horizon execution. We also outline a deployment-oriented path for
transferring ABot-AgentOS-style tool use to smaller agent models through
sandboxed trajectory distillation, SFT/RL, and judge-guided
optimization.

Several limitations remain. Large-scale real-world validation is still
needed under noisy perception, imperfect actuation, network latency,
safety constraints, and heterogeneous robot embodiments. Although
EmbodiedWorldBench covers diverse scenario types, its scene diversity,
social interaction depth, and current agent evaluation coverage remain
limited; future work will report the complete benchmark evaluation and
release EmbodiedWorldBench for open research use. The memory and
self-evolution pipeline also depends on structured traces and post-hoc
feedback, and privacy-aware edge-cloud memory sharing requires stronger
auditing, user control, access management, and failure analysis. Finally,
the small-model training pipeline should move beyond text-state semantic
sandboxes by incorporating visual observations, multi-modal feedback, and
additional simulation platforms.

Overall, ABot-AgentOS provides a unified foundation for physically
grounded, memory-augmented, and continuously improving embodied agents
that can perceive, remember, reason, act, verify, and adapt over
long-term interaction with the physical world.
\label{sec:conclusion}
