\section{Case Studies for Skill Composition}
\label{app:case_study}

This appendix provides detailed per-episode analysis of the three merging strategies discussed in \S4.5. Each case study examines a specific episode where the merging strategies diverge in outcome, revealing the underlying mechanism of success or failure.

\paragraph{Case 1: Complementary Capability Transfer.}
The task in ep4 is to examine a CD under a desklamp. The room contains 5 shelves (shelf 1--5), 3 drawers, and several sidetables; the desklamp is on sidetable 2 and the CD is on shelf 3. Look-Only fails within 50 steps, never visiting any shelf; its search is confined to the desk, drawers 1--3, sidetables 1--2, safe, and bed, repeatedly cycling through previously checked locations (desk visited 8 times, drawer 2 opened 3 times). The root cause is that the Look skill lacks a systematic object search strategy: the model successfully locates and activates the desklamp (Step 12) but cannot find the CD. Component Merging completes the task in 17 steps: after exhausting drawers 1--3, the model begins scanning shelf 1$\rightarrow$2$\rightarrow$3 at Step 12, discovers the CD on shelf 3 at Step 14, picks it up at Step 15, returns to sidetable 2 at Step 16, and activates the desklamp at Step 17. The behavioral difference is concentrated in a single dimension: the pick-specific component contributes a systematic search strategy that includes shelves in the search space, a strategy absent from the Look skill's task-specific component. Because Component Merging mounts the general and mistakes components only once and superposes look-specific and pick-specific independently, Pick's search capability is introduced cleanly without interfering with Look's lamp interaction behavior.

\paragraph{Case 2: Weight Redundancy and Decision Disruption.}
The task in ep6 is to examine a keychain under a desklamp. Dresser 1 holds both desklamp 1 and keychain 3; drawer 7 holds keychain 4; both keychains are valid targets. Look-Only completes the task in 5 steps by simultaneously discovering the desklamp and keychain on the dresser. Component Merging completes it in 18 steps after systematically scanning drawers 1--7. Direct Merging fails within 50 steps: the model identifies valid targets twice (keychain 4 at Step 27, keychain 3 at Step 35) yet executes no pick-up action after either identification, instead wandering to other locations; the desklamp is activated 3 times, but the keychain is never in hand when the lamp is on. This pattern reveals the core problem of Direct Merging: the Look and Pick skill texts share identical general and mistakes components, so mounting their complete LoRAs superimposes these shared components twice, over-amplifying general behavioral patterns that persistently suppress task-specific pick-up actions. Direct Merging does not destroy the model's ability to perceive targets but disrupts the decision threshold for triggering a pick-up upon target recognition. Component Merging avoids this by mounting general and mistakes components once each, with pick-specific supplementing capability only at the task-specific level, keeping the weight strength of general behavioral patterns unchanged and stabilizing the perception-to-action decision chain.

\paragraph{Case 3: OOD Encoding and Behavioral Incoherence.}
The task in ep36 is to look at an alarmclock under a desklamp. Desk 1 holds both alarmclock 2 and desklamp 1, making this a simple co-located task. Look-Only completes it in 4 steps; Component Merging completes it in 3 steps (go to desk $\rightarrow$ use desklamp $\rightarrow$ take alarmclock), one of the shortest successful paths across all episodes. Text Merging fails within 50 steps with a failure mode fundamentally different from Case 2: the model correctly identifies and picks up alarmclock 2 at Step 2 but puts it back at Step 3; it correctly activates desklamp 1 at Step 4 but immediately leaves the desk for sidetable 2 at Step 5. The model then cycles among the desk, sidetable 1, sidetable 2, and bed, activating desklamp 1 four times and desklamp 2 five times while the alarmclock remains on the desk. Unlike Direct Merging's failure mode of perceiving targets but being unable to trigger pick-up, Text Merging exhibits a different pathology: it can execute individual correct actions but cannot chain them into a coherent sequence. Pick-up actions are interrupted by lamp activation, which is in turn interrupted by exploration, with the two skills' behavioral patterns alternately activating yet never completing a handoff. The root cause is that Text Merging concatenates two skill texts as a single input to the hypernetwork, but the hypernetwork has only seen single skill texts during training; the concatenated input constitutes an out-of-distribution (OOD) input, and the resulting LoRA fails to separate the two skills' knowledge, causing their behavioral patterns to interfere at the parameter level without any dispatch mechanism. Component Merging, by contrast, decomposes skills at the text level into independent components, encodes each through the hypernetwork separately, and maintains each behavioral pattern as an independent module in parameter space, thereby supporting coherent sequential decision execution.
