\documentclass[11pt]{article}

% Pass [table] to xcolor (loaded by acl.sty) so \rowcolors is available.
\PassOptionsToPackage{table}{xcolor}

% Change "review" to "final" to generate the camera-ready version.
% Change to "preprint" to generate a non-anonymous version with page numbers.
\usepackage[preprint]{acl}

% In preprint/final mode the lineno package is not loaded, so \linenumbers is
% undefined; the document still toggles it around the single-column appendix.
% Provide a no-op so those calls compile (review mode keeps lineno's real one).
\providecommand{\linenumbers}{}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
% \usepackage{inconsolata}  % optional; install texlive-fonts-extra to enable

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{wasysym}  % \CIRCLE \LEFTcircle \Circle for the positioning table
\newcommand{\cfull}{\CIRCLE}
\newcommand{\chalf}{\LEFTcircle}
\newcommand{\cnone}{\Circle}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{algorithm}
\usepackage{algpseudocode}

\definecolor{rhoAccent}{HTML}{2F4F6F}
\definecolor{rhoComment}{HTML}{6F7480}
\definecolor{rowAlt}{HTML}{ECECEC}
\definecolor{rhoBlue}{HTML}{D6E6F2}

% \newcommand{\joy}[1]{{\color{purple}{joy: #1}}}
\newcommand{\joy}[1]{}

% Prompt listings: a rounded, outlined card with a dark filled title bar
% (tcolorbox) wrapping a line-wrapping listings block. Long prompt lines (up
% to 343 chars) wrap; the card breaks across pages (floats cannot). tcolorbox
% breakable conflicts with the [review] lineno package, so line numbers are
% disabled across single-column Appendix A (\onecolumn + \nolinenumbers).
\usepackage{listings}
\usepackage{enumitem}

\setlist[itemize]{leftmargin=*,topsep=0pt,itemsep=1pt}
\setlist[enumerate]{leftmargin=*,topsep=0pt,itemsep=1pt}
\usepackage[most]{tcolorbox}
\tcbuselibrary{listings,breakable,skins}

\definecolor{promptBg}{HTML}{F8FAFC}     % near-white body fill
\definecolor{promptStroke}{HTML}{2F4F6F} % = rhoAccent: border + title bar

\lstdefinestyle{prompt}{%
  basicstyle=\small\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  breaklines=true,
  breakatwhitespace=true,
  breakindent=1.5em,
  postbreak=\mbox{\textcolor{rhoComment}{$\hookrightarrow$}\space},
  showstringspaces=false,
  aboveskip=0pt, belowskip=0pt,
}

% \begin{promptbox}[lst:label]{Title}  ...  \end{promptbox}
\newtcblisting[auto counter]{promptbox}[2][]{%
  breakable, enhanced,
  listing only,
  listing options={style=prompt},
  colback=promptBg,
  colframe=promptStroke,
  boxrule=0.6pt,
  arc=4pt, outer arc=4pt,
  left=8pt, right=7pt, top=5pt, bottom=5pt,
  boxsep=1pt,
  before skip=1.2em, after skip=0.8em,
  colbacktitle=promptStroke,
  coltitle=white,
  fonttitle=\bfseries\footnotesize,
  toptitle=2.5pt, bottomtitle=2.5pt,
  title={Listing~\thetcbcounter\quad #2},
  label={#1},
}

\algrenewcommand\algorithmiccomment[1]{\unskip\hfill\textcolor{rhoComment}{\,$\triangleright$\,\textit{\footnotesize #1}}}
\algrenewcommand\algorithmicrequire{\textcolor{rhoAccent}{\textbf{Input:}}}
\algrenewcommand\algorithmicensure{\textcolor{rhoAccent}{\textbf{Output:}}}
\algrenewcommand\algorithmicreturn{\textcolor{rhoAccent}{\textbf{return}}}
\newcommand{\rhostage}[1]{\Statex \textcolor{rhoAccent}{\rule[1pt]{6pt}{6pt}}\,\textcolor{rhoAccent}{\textsc{\textbf{#1}}}}

\title{Evolving Agents in the Dark: \\ \textsc{Retrospective Harness Optimization} via Self-Preference}

\author{
  \textbf{Wenbo Pan\textsuperscript{1} \quad Shujie Liu\textsuperscript{2} \quad Chin-Yew Lin\textsuperscript{2} \quad Jingying Zeng\textsuperscript{2} \quad Xianfeng Tang\textsuperscript{2}} \\
  \textbf{Xiangyang Zhou\textsuperscript{2} \quad Yan Lu\textsuperscript{2} \quad Xiaohua Jia\textsuperscript{1}} \\[4pt]
  \textsuperscript{1}City University of Hong Kong \qquad \textsuperscript{2}Microsoft Research Asia
}

\newcommand{\method}{\textsc{RHO}}
\newcommand{\harness}{harness}
\newcommand{\tbd}{\textcolor{red}{\textsc{tbd}}}

\begin{document}
\maketitle

\begin{abstract}
AI agents rely on a \textit{harness} of skills, tools, and workflows to solve complex problems.
Continually improving this harness is essential for adapting to new tasks.
However, existing optimization methods typically require ground-truth validation sets, yet such labeled data is difficult to acquire in practical deployment settings.
To address this problem, we introduce \textit{Retrospective Harness Optimization} (\method{}), a self-supervised method that optimizes the agent harness using only past trajectories.
Specifically, \method{} selects a diverse coreset of challenging tasks from past trajectories and re-solves them in parallel.
The agent analyzes these rollouts using self-validation and self-consistency, then generates candidate harness updates and selects the most effective one by its own pairwise self-preference.
We evaluate \method{} across three diverse domains, spanning software engineering, technical work, and knowledge work.
Notably, a single optimization round improves the pass rate on SWE-Bench Pro from 59\% to 78\% without any external grading.
Furthermore, our analysis demonstrates that \method{} effectively targets prior failure modes. As a result, the optimized harness alters the agent's behavior patterns and sustains higher accuracy during long-horizon sessions.
Code is available at \url{https://github.com/wbopan/retro-harness} and the project website at \url{https://paper-rho.wenbo.io}.
\end{abstract}

\input{sec_introduction}
\input{sec_related_work}
\input{sec_problem_setting}
\input{sec_method}
\input{sec_experiments}
\input{sec_discussion}
\input{sec_conclusion}

\section*{Limitations}

In this paper we introduce \method{}, a self-supervised method that improves an agent's harness from its own past trajectories without any external grading.
However, operating without ground-truth feedback carries several limitations.
First, group rollout replays each coreset task several times, which assumes environments that reset cleanly and tolerate repeated attempts, leaving one-shot or irreversible tasks outside the setting \method{} targets.
Second, \method{} presumes that a meaningful portion of the agent's competence is mediated by an editable harness of prompts, skills, and tools; our experiments span software engineering, technical work, and knowledge work, and extending \method{} to domains with different harness surfaces, task fingerprints, and rollout budgets remains future work.
Third, \method{} trusts past trajectories as its only optimization input; in open environments these trajectories can embed adversarial content injected mid-task \citep{liu2026webtrap}, and harness updates distilled from compromised trajectories could entrench such behavior, reinforcing the need for the audit practices in the Ethics Statement.

\section*{Ethics Statement}

\method{} modifies persistent agent behavior from model-generated judgments.
This can amplify mistaken preferences, unsafe procedures, or biased behavioral rules if the evaluator prefers them.
Deployments should keep full audit logs, require human approval for sensitive harness edits, and use domain-specific safety checks before applying accepted harnesses to high-impact tasks.

\section*{Reproducibility Statement}

Every run persists prompts, completions, trajectories, diagnoses, candidate harnesses, harness diffs, configs, scores, run metadata, and held-out reports.
The numbers in this draft are direct reads from recorded run reports.
A detailed comparison with related work is given in Appendix~\ref{app:positioning}, prompts are listed in Appendix~\ref{app:prompts}, hyperparameters in Appendix~\ref{app:hyperparameters}, full pipeline details in Appendix~\ref{app:pipeline}, per-dataset specifications in Appendix~\ref{app:datasets}, baseline implementations in Appendix~\ref{app:baselines}, and per-method optimization-phase compute cost in Appendix~\ref{app:cost}.

\bibliography{custom}

\appendix

% Appendix A (positioning) and the prompts appendix both render single-column:
% the positioning table spans the full page width, and the verbatim prompts are
% wide artifacts whose long lines read far better at full page width
% (permitted by the ACL April-2025 single-column appendix exception).
\onecolumn
\nolinenumbers   % breakable tcolorbox prompt cards are incompatible with lineno
\input{appendix_g_positioning}
\input{appendix_a_prompts}
\linenumbers
\twocolumn
\input{appendix_b_hyperparameters}
\input{appendix_c_pipeline}
\input{appendix_d_datasets}
\input{appendix_e_baselines}
\input{appendix_f_cost}

% Concrete computed harness artifacts last: single-column, wide verbatim dumps.
\onecolumn
\nolinenumbers
\input{appendix_h_artifacts}

\end{document}
