\section{Related Work}
\label{sec:related}

\textbf{Unified Multimodal Models.}
Existing UMMs broadly fall into two families. The first generates within a \emph{single backbone}, either by modeling images as discrete tokens (Chameleon~\cite{chameleon}, Emu3~\cite{emu3}) or by attaching diffusion in a VAE latent space (Transfusion~\cite{transfusion}, Show-o~\cite{showo}, JanusFlow~\cite{janusflow}). These approaches further address the interference between understanding and generation through decoupled visual encoders in Janus series~\cite{janus,janus-pro,janusflow} or modality-specific experts in BAGEL~\cite{bagel}, but all depend on a separately pretrained visual tokenizer---a VQVAE or a continuous VAE---to define the latent space for generation. The second family \emph{stitches an LLM with an external diffusion model}: the LLM predicts visual representations such as CLIP features, which then condition a separately trained diffusion decoder to render images, as in Emu2~\cite{emu2}, SEED-X~\cite{seedx}, BLIP3-o~\cite{blip3o}, and MetaQueries~\cite{metaquery}.
Representation Forcing builds on the same MoT backbone design~\cite{bagel} but eliminates separately pretrained VAE. The decoder learns to predict intermediate visual representations from its own jointly trained understanding encoder, and uses them as in-context conditioning for pixel-space diffusion within the same transformer, yielding a fully end-to-end UMM.

\textbf{Pixel-Space Generation.}
State-of-the-art diffusion models~\cite{stable_diff,flux} typically operate in the latent space of a pretrained VAE~\cite{vae}, which reduces compute and enables high-resolution synthesis but prevents end-to-end training. A recent line of work has explored generating directly in pixel space~\cite{DDPM_paper,dhariwal2021diffusion}, enabling end-to-end learning from raw data: JiT~\cite{jit} demonstrates that plain Vision Transformers with $x$-prediction can generate on raw pixels, and other approaches explore alternative pixel-space architectures~\cite{sid2,pixelflow,pixnerd}. These methods mostly focus on standalone generation on the ImageNet~\cite{imagenet} dataset. In UMMs, the broader image distribution and richer text conditioning leave naive pixel-space diffusion with a clear quality gap; Representation Forcing closes this gap by providing a structural scaffold from the model's own understanding encoder.



\textbf{Representation Learning for Generation.}
Several works explore richer visual representations to improve generation. REPA~\cite{repa} aligns intermediate diffusion features with frozen pretrained representations to accelerate training convergence; RAE~\cite{rae,scale-rae-2026} goes further by replacing the VAE entirely with frozen pretrained encoders such as DINOv2~\cite{dinov2} and SigLIP~\cite{siglip}, providing a semantically richer latent space. However, all these methods still operate within a frozen, externally defined representation space. Representation Forcing differs by training the decoder to directly predict visual representations from its own jointly trained understanding encoder, making high-level visual structure a native output of the generation process.
