\begin{tcolorbox}[
    colback=black!2,
    colframe=nvidiagreen!75!white,
    boxrule=0.4pt,
    arc=1pt,
    left=4pt,
    right=4pt,
    top=4pt,
    bottom=4pt,
    title=Top-level Fields in the Image Caption Schema
]
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{\linewidth}{@{}>{\ttfamily}lX@{}}
\toprule
Field & Description \\
\midrule
subjects[] & Per-subject records capturing identity, appearance, spatial placement, pose, clothing, expression, and count-sensitive anatomy fields when applicable. Human or human-like subjects additionally include demographic and facial-attribute fields when visible. \\
subject\_details & Open-ended attributes or clarifications that do not fit cleanly into a single subject slot, including free-form fine-grained facial details when useful. \\
background\_setting & Global scene context, environment, and background elements. \\
lighting & Illumination conditions, directionality, shadows, and notable lighting effects. \\
aesthetics & Composition, color palette, mood, and repeated visual patterns. \\
cinematography & Framing, camera angle, depth of field, focus behavior, and lens characteristics. \\
style\_medium / artistic\_style & Medium, rendering style, and artistic treatment. \\
context & High-level narrative or situational context for the scene. \\
text\_and\_signage\_elements[] & Any visible text, its appearance, placement, category, and scene relevance. \\
quadrant\_scan & Region-wise scan of the top-left, top-right, bottom-left, bottom-right, and absolute center. \\
comprehensive\_t2i\_caption & A natural-language summary distilled from the structured fields. \\
resolution / aspect\_ratio & Image size metadata used to preserve scale and layout cues. \\
\bottomrule
\end{tabularx}
\end{tcolorbox}
