\section{Introduction}

Recent advances in diffusion-based video generation have produced powerful text-to-video (T2V) and text-and-image-to-video (TI2V) foundation models capable of synthesizing high-quality and temporally coherent videos~\cite{videoworldsimulators2024,bao2024vidu,yang2024cogvideox,lin2024open,zheng2024open,wan2025wan,kong2024hunyuanvideo}. These models provide strong generative priors for visual appearance, motion, and scene evolution, and therefore offer a promising starting point for building video world models. However, a high-quality offline video generator is not yet an interactive world model. An interactive video world model should support causal rollout, respond to user actions such as camera trajectories, and generate future frames with sufficiently low latency for real-time interaction~\cite{sun2025worldplay,genie3,tang2025hunyuan,mao2025yume,feng2025vidarc,huang2025live,sun2025streamavatar,hong2025relic,ye2025yan,xiang2025pan,he2025matrix,shin2025motionstream}.

Although recent works have explored autoregressive (AR) diffusion distillation to convert existing video foundation models into real-time interactive world models~\cite{yin2025slow,lin2025diffusion,huang2025self,zhu2026causal,zhao2026causal,yang2025towards}, these techniques remain scattered across separate pipelines. As a result, building an interactive video world model still requires substantial effort in data construction, controllable fine-tuning, AR training, few-step distillation, post-training alignment, and inference. \emph{A unified, reproducible, and extensible framework for this full pipeline is still missing}.

\input{Figures_tex/pipeline}

To this end, we present \textbf{minWM}, a full-stack open-source framework for building real-time interactive video world models. Instead of releasing a single trained checkpoint, minWM provides a reproducible end-to-end pipeline that converts existing T2V or TI2V video foundation models into camera-controllable few-step autoregressive video world models. The framework covers the complete workflow, including data construction, camera-controllable fine-tuning, autoregressive diffusion training, few-step distillation, and low-latency inference. Its modular design allows researchers to plug in different video backbones, control signals, training recipes, and inference configurations, making minWM easy to reproduce, adapt, and extend.

Concretely, minWM follows a two-phase recipe. First, it fine-tunes a bidirectional video diffusion backbone on camera-annotated or camera-generated video data, enabling the model to follow prescribed camera trajectories while preserving the visual quality of the original foundation model~\cite{li2026cameras}. Second, it applies Causal Forcing~\cite{zhu2026causal} or Causal Forcing++~\cite{zhao2026causal} to transform the camera-controllable multi-step bidirectional model into a few-step autoregressive generator. This stage consists of teacher-forcing AR diffusion training~\cite{teng2025magi}, causal ODE~\cite{zhu2026causal} or causal consistency distillation~\cite{zhao2026causal} initialization, and asymmetric DMD~\cite{wang2023prolificdreamer,luo2023diff,yin2024one,yin2025slow} post-training with self-rollout~\cite{huang2025self}. The resulting model supports camera-controllable autoregressive video generation with few-step inference, making it suitable for low-latency interactive applications.

We instantiate minWM on representative open video backbones, including Wan2.1-T2V-1.3B~\cite{wan2025wan} and HY1.5-TI2V-8B~\cite{kong2024hunyuanvideo}. These instantiations demonstrate two practical usages of the framework. First, minWM provides a complete conversion pipeline that starts from a bidirectional T2V or TI2V foundation model and progressively turns it into a real-time, camera-controllable autoregressive video world model. By releasing intermediate checkpoints for each training stage, minWM allows researchers to resume, modify, or extend the pipeline from any stage. Second, minWM supports adapting existing video world models, such as HY-WorldPlay~\cite{sun2025worldplay}, to new data distributions, training recipes, or latency targets through fine-tuning and distillation. Beyond final generation results, we further report practical ablations on camera trajectory quality of the dataset, controllability training steps, and minimal batch-size requirements, providing actionable guidance for reproducible interactive world-model training. The overall pipeline is illustrated in Fig. \ref{fig:pipeline}.

Our contributions are summarized as follows:

\begin{itemize}
    \item We release \textbf{minWM}, a fully open-source end-to-end pipeline for building real-time interactive video world models. The pipeline covers camera-conditioned data construction, controllable fine-tuning of bidirectional video diffusion models, and Causal Forcing / Causal Forcing++ distillation, including AR diffusion training, causal ODE or causal consistency distillation initialization, asymmetric DMD post-training, and low-latency inference. 

    \item We show that minWM is architecture-general and can convert multiple types of video foundation models into camera-controllable few-step autoregressive world models. We instantiate the framework on representative open backbones, including Wan2.1-T2V-1.3B with cross-attention-based condition injection and HY1.5-TI2V-8B with an MMDiT-style architecture~\cite{esser2024scaling}.

    \item We further support the adaptation of existing video world models, such as HY-WorldPlay, to new data distributions, training recipes, and latency targets. Together with practical ablations on camera trajectory quality, controllability training steps, and minimal batch-size requirements, minWM provides a reproducible and extensible recipe for building and adapting interactive video world models.
\end{itemize}