\section{Introduction}

Video generation models have shown strong potential as world simulators for embodied intelligence~\cite{openai2024sora,GoogleDeepMind2025Veo3,wan2025wan,kong2024hunyuanvideo,yang2024cogvideox}, providing scalable visual futures that can be used for data generation, world simulation, and downstream applications such as policy learning~\cite{bruce2024genie,o2024open,nvidia2026worldsimulationvideofoundation,ye2026worldactionmodelszeroshot,li2025unified,liang2025video, lingbot-va2026}. However, embodied world simulation requires more than photorealistic videos. The generated dynamics must also be physically plausible, especially in contact-rich manipulation.

This requirement remains difficult for current video generators. In contact-rich manipulation, physical violations often appear as local dynamic errors, such as discontinuous gripper trajectories, object penetration, or anti-gravity motion, and as global relational errors, such as a pushed object remaining static or a grasped object drifting away. These errors directly undermine the reliability of the generated video as a visual world simulator, since the predicted frames no longer represent a valid consequence of the robot action.

Existing approaches provide only partial solutions. General video models rely on large-scale visual pre-training but lack sufficient exposure to embodied contact dynamics~\cite{GoogleDeepMind2025Veo3,wan2025wan,kong2024hunyuanvideo,yang2024cogvideox}. Robot-oriented video or world models improve task relevance with manipulation data~\cite{jang2025dreamgen,feng2025vidar,cheang2024gr,du2023learning, chen2025lvp}, yet are usually trained with reconstruction objectives that treat physically critical regions and background pixels uniformly. Recent physics-aware methods introduce geometric cues such as depth~\cite{yang2024depth, video_depth_anything}, tracking~\cite{karaev2023cotracker,karaev24cotracker3,harley2025alltracker}, and 3D structure~\cite{zhen2025tesseract,shang2025roboscape,zhou2024robodreamer,guo2025ctrl}, or apply preference and reward alignment to suppress implausible generations~\cite{ouyang2022training,rafailov2024direct,ali2025abot,zhang2026mind}. However, geometric constraints mainly capture local motion consistency, while preference feedback is often sparse and weakly localized. As a result, existing methods still lack a unified training-time mechanism for aligning both local dynamics and global interaction outcomes.

This raises a key question: \textit{how can we inject physical supervision into video generation in a way that is both hierarchical and region-focused?} We observe that physical plausibility in manipulation videos is naturally hierarchical. At the pixel level, local motion should satisfy trajectory continuity, depth consistency, and contact-compatible displacement. At the semantic level, object relations should evolve according to the interaction semantics: a pushed object should move away, a grasped object should remain coupled with the gripper, and a placed object should rest on the support surface. Meanwhile, physical evidence is highly localized around manipulators, objects, contacts, and moving regions. Applying supervision uniformly over all pixels dilutes these signals and weakens alignment.

Based on this observation, we propose \textbf{PhysisForcing}, a region-focused hierarchical physics alignment framework for video generation. Instead of supervising all pixels uniformly, PhysisForcing first identifies physics-informative regions, including manipulators, manipulated objects, contact areas, and moving parts, and then applies two complementary alignment losses to these regions during backbone training. Specifically, the \textit{pixel-level physics alignment} module uses point tracking~\cite{karaev24cotracker3} to supervise per-point trajectories of the DiT feature, so that local motion stays continuous and contact-compatible. The \textit{semantic-level physics alignment} module instead aligns the pairwise token-similarity matrix of the DiT feature with that of a frozen video understanding encoder~\cite{bardes2024revisiting} on the same tokens, encouraging globally consistent robot-object interactions. In this way, PhysisForcing improves physical plausibility through hierarchical supervision on regions most relevant to robot-object interactions.

Extensive experiments on R-Bench, PAI-Bench, and EZS-Bench demonstrate the effectiveness of PhysisForcing for embodied video generation. Across the Wan2.2-I2V-A14B and Cosmos3-Nano backbones, our method improves physical plausibility scores by 22.3\% and 9.2\% over the base models, and by 7.1\% and 3.7\% over vanilla finetuning, with the Cosmos3-Nano variant attaining the best overall R-Bench score among all evaluated models. Beyond generation, under the WorldArena action-planner protocol~\cite{shang2026worldarena} PhysisForcing lifts the closed-loop success rate from 16.0\% to 24.0\%, surpassing strong world-model planners, and it also improves downstream policy success when used as the video backbone of a world action model~\cite{yuan2026fastwam}, indicating that physically aligned video models provide stronger representations for embodied intelligence.

In summary, our contributions are:
\begin{itemize}
    \item We formulate physical plausibility in robotic video generation as a hierarchical and region-focused alignment problem.
    \item We propose \textbf{PhysisForcing}, a training-time framework that focuses physical supervision on interaction-critical regions and aligns generation at both the pixel-level motion and the semantic-level inter-region relations.
    
    \item We demonstrate consistent gains on R-Bench, PAI-Bench, and EZS-Bench across diffusion video backbones of different scales and families, including Wan2.2-I2V-A14B and Cosmos3-Nano, with ablations validating the importance of hierarchical physical supervision. We further show that PhysisForcing benefits embodied decision making, improving world-model action planning on WorldArena and downstream policy success when used as the video backbone of a world action model.
\end{itemize}