\section{Related Work}

\subsection{Embodied Video Generation and World Models}

Video generation models have emerged as a promising paradigm for embodied intelligence, offering scalable solutions to the data scarcity challenge in robotics~\cite{openai2024sora,wan2025wan,kong2024hunyuanvideo,yang2024cogvideox,GoogleDeepMind2025Veo3}. General-domain models like Sora~\cite{openai2024sora} and Wan~\cite{wan2025wan} demonstrate impressive visual fidelity but often struggle with embodied manipulation tasks due to limited exposure to robot-object interaction dynamics. To address this, recent work has developed robotics-specific world models. Cosmos~\cite{nvidia2026worldsimulationvideofoundation}, DreamGen~\cite{jang2025dreamgen}, and GR-2~\cite{cheang2024gr} focus on action-conditioned video generation for robotic scenarios, while Vidar~\cite{feng2025vidar}, UnifoLM-WMA-0~\cite{unifolm-wma-0}, and Unified Video Action Model~\cite{li2025unified} emphasize world action modeling. Video generators have also been explored as robot policies~\cite{liang2025video,hu2024video}. Gen2Act~\cite{bharadhwaj2024gen2act} generates human videos in novel scenarios to enable generalizable manipulation. Genie~\cite{bruce2024genie} and Genie Envisioner~\cite{liao2025genie} explore interactive world models. However, these approaches primarily optimize visual quality or action controllability, without explicitly enforcing physical constraints during training. Consequently, they remain susceptible to physically implausible manipulation dynamics, including object penetration, trajectory discontinuities, and interaction outcomes that violate basic physical causality, which limits their utility as world simulators.

\subsection{Physics-aware World Simulator}

Recognizing the limitations of purely visual objectives, recent work has begun incorporating physical constraints into video generation. These approaches can be broadly categorized into three paradigms. \textit{Geometry-based methods} enforce pixel-level physical consistency through depth prediction~\cite{yang2024depth,video_depth_anything}, keypoint tracking~\cite{karaev2023cotracker,karaev24cotracker3,harley2025alltracker}, or 3D scene reconstruction~\cite{zhen2025tesseract}. RoboScape~\cite{shang2025roboscape} jointly learns temporal depth prediction and keypoint dynamics to improve geometric consistency and motion modeling. RoboDreamer~\cite{zhou2024robodreamer} learns compositional world models for robot imagination. CTRL-World~\cite{guo2025ctrl} proposes a controllable generative world model. While effective at capturing local spatial structure, these methods do not explicitly model semantic-level interactions or relational outcomes. \textit{Preference-based methods} apply post-training alignment~\cite{ouyang2022training,rafailov2024direct} to suppress physically implausible generations. ABot-PhysWorld~\cite{ali2025abot} employs DPO-based training with physics-aware discriminators to reduce violations like object penetration and anti-gravity motion. MIND-V~\cite{zhang2026mind} uses GRPO with a Physical Foresight Coherence reward to align generated dynamics with physical laws. While these approaches successfully reduce certain failure modes, they operate post-hoc—correcting rather than preventing physical violations—and may sacrifice visual quality in the alignment process. \textit{Simulator-based methods} leverage physics engines~\cite{wang2023robogen,yang2025physics} or test-time selection to ensure physical validity, but incur significant computational overhead and limit scalability.

Existing physics-aware video models often provide supervision at a single level or over the entire frame. However, contact-rich manipulation calls for two design principles. First, physical consistency should be hierarchical: pixel-level motion constraints govern per-point trajectories and contacts, while semantic-level relational cues capture inter-region interactions. Second, physical supervision should be localized: the most informative evidence lies around manipulators, contact interfaces, and moving parts, rather than static background regions. PhysisForcing follows these principles by introducing a region-focused hierarchical physics objective for training video generators.