\section{Related Work}

\textbf{Advancements in GRPO and Reasoning Models}\quad The shift from PPO \citep{ppo} to GRPO \citep{grpo} has significantly streamlined the post-training pipeline for Large Language Models by eliminating the need for a heavily parameterized value model. This efficiency has been instrumental in the development of state-of-the-art reasoning models like DeepSeek-R1 \citep{deepseek-r1}. Recently, several variants have been proposed to further enhance GRPO’s stability and efficiency. GSPO \citep{gspo} shifts the importance ratio calculation from the token level to the sequence level to mitigate variance. DAPO \citep{dapo} introduces dynamic sampling and token-level policy gradients to accelerate convergence. To address the issue of length explosion during reasoning, methods like GFPO \citep{gfpo} and DLER \citep{dler} incorporate length-based heuristics, such as filtering responses by reward-per-token ratios or applying simple truncation penalties. While these extensions improve specific aspects of the GRPO framework, they primarily focus on single-reward maximization or rely on rigid heuristics, which struggle to generalize to complex, multi-dimensional alignment tasks.

\textbf{Multi-Reward Reinforcement Learning in LLMs}\quad Integrating multiple, often conflicting, reward signals is increasingly vital for practical LLM deployments, ranging from balancing diverse human preferences \citep{alarm, jang2023personalized} to enforcing length efficiency \citep{gfpo, dler, luo2025o1} and strict formatting constraints in agentic tool-use \citep{toolrl, xlam}. To simultaneously optimize these diverse objectives, standard practices typically rely on Reward Combination (RC) or Advantage Combination (AC). While RC directly scalarizes raw rewards but frequently suffers from magnitude explosion, AC-based methods like GDPO \citep{gdpo} independently normalize each reward into an advantage before applying a static convex combination. Although GDPO mitigates extreme gradients, its reliance on fixed hyperparameters completely isolates the objectives during normalization. Our proposed DVAO diverges fundamentally from these approaches: by introducing a dynamic, variance-adaptive weighting scheme, DVAO strictly bounds advantage magnitudes while explicitly modeling cross-objective correlations, enabling a self-adaptive regularization mechanism that seamlessly scales to multiple objectives without manual tuning.