% arXiv / tech-report build of the paper.
% Mirrors iclr2026_conference.tex but swaps in arxiv_techreport.sty
% for a DeepMind-skeleton / OpenAI-restraint visual style.
% Shares sections/, figs/, .bib, and math_commands.tex with the ICLR build.


\documentclass{article}
\usepackage{arxiv_techreport}

\input{math_commands.tex}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\usepackage[section]{placeins}
\usepackage{fontawesome5}
\usepackage{subcaption}


\NewDocumentCommand{\runzhe} { mO{} }{\textcolor{blue}{\textsuperscript{\textit{runzhe:}}\textsf{\textbf{[#1]}}}}

\definecolor{hanGreen}{RGB}{0,100,60}
\NewDocumentCommand{\han} { mO{} }{\textcolor{hanGreen}{\textsuperscript{\textit{han:}}\textsf{\textbf{[#1]}}}}


\title{EvoPolicyGym: Evaluating Autonomous Policy Evolution in Interactive Environments}



\author{%
\begin{minipage}[t]{0.88\textwidth}
\centering
Zhilin Wang\textsuperscript{1,*}, Han Song\textsuperscript{2,*},
Runzhe Zhan\textsuperscript{3,*}\\
Jusen Du\textsuperscript{4}, Jiacheng Chen\textsuperscript{2},
Tianle Li\textsuperscript{2}, Qingyu Yin\textsuperscript{5},
Yulun Wu\textsuperscript{5}\\
Zhennan Shen\textsuperscript{9}, Tong Zhu\textsuperscript{6},
Yanshu Li\textsuperscript{7}, Guanjie Chen\textsuperscript{9} \\
% Junjie Wu\textsuperscript{8}\\
Derek F. Wong\textsuperscript{3}, Yafu Li\textsuperscript{2,\textdagger},
Yu Cheng\textsuperscript{2,\textdagger}, Yang Yang\textsuperscript{9,\textdagger}\\[0.35em]
{\normalfont\small 
\textsuperscript{1}University of Science and Technology of China \quad
\textsuperscript{2}The Chinese University of Hong Kong \\
\textsuperscript{3}University of Macau \quad
\textsuperscript{4}Tsinghua University \quad
\textsuperscript{5}Zhejiang University \\
\textsuperscript{6}Soochow University \quad
\textsuperscript{7}Brown University \\
% \quad
% \textsuperscript{8}The Hong Kong University of Science and Technology 
\textsuperscript{9}Shanghai Jiao Tong University \\
} 
{\normalfont\small \textsuperscript{*}Equal contribution \quad
\textsuperscript{\textdagger}Corresponding authors}
\end{minipage}
}



\newcommand{\evopolicygym}{\textsc{EvoPolicyGym}}
\newcommand{\policy}{\texttt{policy.py}}
\newcommand{\workspace}{\texttt{workspace}}

\begin{document}

\maketitle

\begin{abstract}
Autonomous agents are increasingly expected to improve executable policies through feedback, yet existing evaluations often collapse this process into a final score or confound it with open-ended software-engineering progress. 
We introduce Autonomous Policy Evolution, a controlled evaluation setting in which a harness–model agent repeatedly edits an executable policy system under a fixed interaction budget. 
We instantiate this setting in EvoPolicyGym, a benchmark built from compact interactive RL environments that evaluates how agents iteratively improve explored policies.
On the EvoPolicyGym suite, GPT-5.5 achieves the strongest aggregate rank score and top-two performance on all 16 environments. 
Beyond leaderboard results, EvoPolicyGym also provides trajectory-level diagnostics that distinguish how agents allocate budget, convert feedback into parametric tuning. 
These analyses show that strong autonomous policy evolution depends not only on isolated task wins, but on discovering task-appropriate mechanisms and refining policies under bounded feedback.

\begin{center}
\small
\faGithub~\href{https://github.com/Linzwcs/EvoPolicyGym}{\texttt{EvoPolicyGym}}
\qquad
\faDatabase~\href{https://huggingface.co/datasets/linzw/EvoPolicyGym-Exp-data}{\texttt{EvoPolicyGym-Exp-data}}\qquad
\faGlobe~\href{https://linzwcs.github.io/EvoPolicyGym/}{\texttt{EvoPolicyGym.io}}
\end{center}
\end{abstract}

\input{sections/intro}
\input{sections/related_work}
\input{sections/framework}

\input{sections/experiment}
\input{sections/analysis}

\section{Conclusion}
\evopolicygym{} casts autonomous policy improvement as a controlled evaluation
of the systems agents build over time. Each harness--model agent edits an
executable policy under a fixed interaction budget, learns only from visible
train feedback, and is judged by hidden validation-selected heldout
performance. The Core16 results show that high scores require more than
isolated task wins: strong agents infer task-appropriate abstractions,
translate feedback into mechanism-level code changes, and preserve useful
candidates under budget pressure. By pairing leaderboard scores with
trajectory-level diagnostics, \evopolicygym{} provides a concrete protocol for
measuring stable, feedback-driven autonomous policy evolution.

\section*{Acknowledgements}
We thank Jiayi Weng for the public blog post \emph{Learning Beyond Gradients}
\citep{learningbeyondgradients}. Its central insight, that coding agents can
continually maintain and improve heuristic systems rather than merely produce
one-off policy files, directly shaped the starting point of this work. During
this project, we found that the word ``heuristic''
was difficult to make operational. Traditional hand-written rules are often
called heuristics, but the boundary becomes unclear once the policy includes
tuned numeric parameters, learned components, or baselines optimized by methods
such as PPO. This difficulty pushed us to make the policy system the benchmark
object: an executable policy together with the state, code structure, feedback
traces, and revision history that an agent can maintain. Fixed
environment-interaction budgets then create optimization pressure while hidden
validation and heldout splits prevent leakage. This setting lets us compare
how well stronger models extract insight from feedback and turn that insight
into policy improvements.

\bibliography{iclr2026_conference}
\bibliographystyle{iclr2026_conference}

\input{sections/appendix}

\end{document}
