
\section{What Controls Subspace Locking?}
\label{sec:controls}
Section~\ref{sec:subspace-lock} shows that on-policy distillation (OPD) learns through an
early-emerging update channel sufficient for training. We now ask what
maintains this channel by perturbing three candidate factors: token-level
teacher supervision, rollout policy, and objective composition. We use stable rank as the primary diagnostic, with details in Appendix~\ref{app:control-details}.



\subsection{Runtime Perturbations Preserve the Lock}
\label{sec:runtime-controls}

We first perturb runtime sources while keeping the objective fixed. Token
sparsification retains either top-KL or randomly selected tokens at
$25\%$ and $50\%$ density. Figure~\ref{fig:controls}(a) shows that all
sparsified variants closely track the OPD stable-rank trajectory. Even
random $25\%$ retention changes update scale more than spectral shape,
indicating that the low-rank profile is not localized to a small set of
high-KL tokens.

% We next perturb rollout generation. 
% Off-policy rollout perturbations leave the stable-rank trajectory nearly
% unchanged, as shown in Figure~\ref{fig:controls}(b). Off-policy OPD produces
% a modestly larger update norm, but its stable rank remains
% matched to on-policy OPD. Thus, the lock is robust to both token-level
% thinning and rollout-policy changes.

Off-policy rollout perturbations also leave the stable-rank trajectory
nearly unchanged (Figure~\ref{fig:controls}(b)). Although off-policy OPD
produces a modestly larger update norm, its stable rank remains matched to
on-policy OPD, showing robustness to rollout-policy changes.

\subsection{Objective Mixing Changes the Rank Dynamics}
\label{sec:objective-mixing}
We perturb objective composition by interpolating the OPD and RLVR
advantage signals in our GRPO-style policy-gradient implementation:
\begin{equation}
    A_i^{(\alpha)}
    =
    \alpha A_{i,\mathrm{OPD}}
    +
    (1-\alpha)A_{i,\mathrm{RLVR}} .
    \label{eq:objective-mixing}
\end{equation}
Unlike token sparsification or off-policy generation, this changes the
gradient source rather than only the sampled data.

Figure~\ref{fig:controls}(c) shows a clear split. OPD-dominant mixtures
retain the OPD-like stable-rank trajectory. When the OPD component becomes
weak, the trajectory no longer follows the baseline and enters a distinct
spectral regime. Objective composition therefore changes the rank dynamics
in a way that runtime perturbations do not.

\paragraph{Mechanistic view.}
The controls separate sample perturbations from objective perturbations.
OPD gradients can be written as token-level sums,
\begin{equation}
    g_{\mathrm{OPD}}=\sum_t J_t^\top \delta_t , \
    \tilde g_{\mathrm{OPD}}=\sum_t m_t J_t^\top \delta_t ,
\end{equation}
where $J_t$ is the local Jacobian, $\delta_t$ is the teacher--student
token discrepancy, and $m_t$ is a token mask. If token gradients share a
dominant update subspace, masking primarily rescales the second moment,
\begin{equation}
    \mathbb{E}[\tilde g \tilde g^\top]
    \approx
    c\,\mathbb{E}[g_{\mathrm{OPD}}g_{\mathrm{OPD}}^\top]+\mathrm{noise},
\end{equation}
so the leading spectral directions are preserved. Off-policy rollouts
change the sampling distribution, but retain the same teacher-token
gradient source. Objective mixing changes the source itself:
\begin{equation}
\begin{split}
    g_\alpha
    &=
    \alpha g_{\mathrm{OPD}}
    +(1-\alpha)g_{\mathrm{RLVR}}, \\
    \Sigma_\alpha
    &\approx
    \alpha^2\Sigma_{\mathrm{OPD}}
    +(1-\alpha)^2\Sigma_{\mathrm{RLVR}} \\
    &\quad
    +\alpha(1-\alpha)\Sigma_{\mathrm{cross}} .
\end{split}
\end{equation}
Thus, weakening the OPD term can change the dominant covariance geometry,
explaining why objective mixing, unlike token thinning or rollout changes,
breaks the OPD-like rank trajectory.

\begin{takeawaybox}
\textbf{Implication.}
Subspace locking is objective-sensitive rather than runtime-induced.
Effective OPD recipes should regulate objective-induced update geometry,
not only token coverage or rollout generation.
\end{takeawaybox}