\section{Related Work}
\label{sec:related_work}

\subsection{Customized Image Generation}
Customized image generation has emerged as a pivotal task within the broader landscape of image synthesis, focusing on enabling pretrained diffusion models to understand specific concepts from limited data and re-render them in diverse contexts. Early optimization-based methods, such as Textual Inversion\cite{gal2022image} and DreamBooth\cite{ruiz2023dreambooth}, paved the way by learning specific tokens or fine-tuning the model for a single subject. 
% To bypass the time-consuming optimization process at inference, recent research has shifted toward zero-shot personalization using image encoders or in-context learning.
Methods like ELITE\cite{wei2023elite}, IP-Adapter\cite{ye2023ip}, InstantID\cite{wang2024instantid}, and MoMA\cite{song2024moma} treat personalization as a vision-conditioned generation task by training specialized adapters. 
% Such a technical paradigm also extends to the multi-concept personalized generation task\cite{kumari2023multi,wang2024ms,mou2025dreamo}. 
With the emergence of Diffusion Transformers (DiT)\cite{peebles2023scalable} like FLUX\cite{flux2024} and SD3\cite{esser2024scaling}, the paradigm has further evolved toward leveraging strong in-context capabilities. For instance, OmniControl\cite{xie2023omnicontrol} and EasyControl\cite{zhang2025easycontrol} adapt text-to-image models for precise personalized control, while unified models like Bagel\cite{deng2025emerging} attempt to harmonize understanding and generation. Furthermore, large-scale models such as FLUX Kontext\cite{labs2025flux}, Qwen-Image-Edit\cite{wu2025qwen}, and FLUX2\cite{flux-2-2025} also leverage in-context learning for image generation and editing. 
However, zero-shot adapters often fail on out-of-distribution effects, making custom LoRA training the reliable industrial standard. Yet, directly composing these LoRAs with acceleration modules (e.g., lightx2v\cite{lightx2v}) triggers severe feature interference and semantic drift. To resolve this, we distill multiple customized effects into a single, few-step unified LoRA, completely avoiding the conflicts inherent in multi-module composition.
% Despite the generalization of zero-shot adapters, they often struggle with highly specific image transformation effects like unique artistic styles or physical simulations that are absent from the base model's distribution. Consequently, the most reliable practice in industrial pipelines remains training a lightweight effect LoRA on several high-quality data examples. However, to speed up inference, these effect LoRAs must be sequentially applied  with acceleration modules (e.g., lightx2v\cite{lightx2v}) during deployment. This naive combination inevitably triggers severe feature interference, semantic drift, and fidelity degradation. To address these bottlenecks, we propose a novel paradigm that jointly distills multiple customized effects into a single, unified LoRA with few-step acceleration, eliminating the conflicts inherent in multi-module composition.

\vspace{-0.3cm}
\subsection{Few-Step Generation}
To address the inference inefficiency of diffusion models, Consistency Models (CMs) \cite{song2023consistency} and their derivatives \cite{luo2023latent,wang2024phased,zhai2024motion} enable few-step generation by enforcing trajectory self-consistency. Recently, Distribution Matching Distillation (DMD)~\cite{dmd1,dmd2} established a superior paradigm by directly minimizing the divergence between the teacher and student distributions. Recent advances further elevate DMD: Decoupled-DMD~\cite{decoupledmd} enhances fine details via independent noise schedules, while DMDR~\cite{jiang2025distribution} and Flash-DMD\cite{chen2025flash} integrate reinforcement learning to incorporate external preference rewards safely, surpassing the teacher's performance ceiling.
Despite these advancements, existing DMD-based methods are largely confined to single-teacher, homogeneous settings. They suffer from severe training instability and feature collapse when bridging large student-teacher gaps or simultaneously fitting multiple target distributions. To overcome these bottlenecks, we propose CollectionLoRA, a multi-teacher distillation framework designed to stabilize multi-source matching and prevent distribution collapse.
\vspace{-0.3cm}
\subsection{On-Policy Distillation}
Standard offline distillation often suffers from exposure bias and compounding errors~\cite{llmopd}. OPD~\cite{llmopd,minillm,rethinkingopd} mitigates these issues by applying teacher feedback directly to states visited by the student’s own rollouts. While initially validated in Large Language Models for superior stability over scalar-reward reinforcement learning~\cite{rlopd}, OPD has recently been adapted to continuous visual generation. In diffusion and flow matching (e.g., Flow-OPD~\cite{fang2026flowopdonpolicydistillationflow}, D-OPSD~\cite{jiang2026dopsdonpolicyselfdistillationcontinuously}), OPD matches the teacher’s dense velocity fields along student-sampled trajectories.
Technically, while traditional OPD typically aligns the conditional transition distribution step-wise along trajectories, DMD~\cite{dmd2} focuses on optimizing the marginal data distribution of generated samples. Following recent literature~\cite{gu2026anyflowanystepvideodiffusion,chern2025livetalkrealtimemultimodalinteractive}, we conceptually unify DMD under the OPD taxonomy, as both frameworks fundamentally rely on correcting the student's on-policy explored states with teacher signals.

Building upon this paradigm, CollectionLoRA pioneers large-scale multi-teacher distillation to efficiently consolidate 50 to 180 diverse visual effects alongside few-step generation capabilities into a single, unified module.
% While building upon the on-policy distillation paradigm, CollectionLoRA distinguishes itself by pioneering large-scale multi-teacher distillation, which efficiently consolidates 50 to 180 diverse visual effects alongside few-step generation capabilities into a single, unified module.
% % 最近LLM的一些工作将多教师蒸馏作为一种有效的后训练策略。他们将不同领域专家，如数学、....等，的知识蒸馏到一个统一的学生模型中以提高学生模型在各个领域的表现。目前扩散模型领域大多应用单教师蒸馏来使模型获得少步生成能力，对多教师蒸馏策略缺乏探索。CollectionLoRA不仅解决了当前定制化编辑大规模部署的挑战，还是一次对扩散模型多教师蒸馏策略的有效探索。
% Recently, multi-teacher distillation has emerged as an effective post-training strategy in the realm of Large Language Models (LLMs), where the capabilities of diverse domain-specific experts (e.g., mathematics, coding, and law) are distilled into a single unified model to significantly enhance comprehensive performance~\cite{coreteam2026mimov2flashtechnicalreport,yang2025qwen3technicalreport,deepseek_v4_pro}. However, existing DMD-style distillation methods~\cite{decoupledmd,dmd1,dmd2} for diffusion models typically adopt a single-teacher setup where both the student and teacher share the same initialization. Their primary goal is to compress a multi-step teacher into a few-step student. In this process, the student only learns to align with the teacher's distribution using fewer inference steps, rather than acquiring novel distributions. This fundamentally differs from CollectionLoRA, where effect teachers are initialized differently from the student.