\section{Introduction}
Autonomous agents are expected to improve through feedback rather than produce a single fixed output. Modern coding agents are able to call tools, observe failures, and revise executable artifacts over long horizons \citep{sweagent,openhands,codeact}, while self-improvement systems show that language models can use reflection to refine answers across attempts \citep{reflexion,selfrefine,alphaevolve}. However, this broader capability is hard to evaluate because improvement is both an outcome and a process: final scores can hide blind retries, overfitting to visible feedback, brittle special cases, missing verification, and other trajectory-level failure modes \citep{agentlens,codeagentbehavior,howcodingagentsfail}. 
Fully open-ended engineering tasks add further confounders, including evolving specifications and software-maintenance quality \citep{sweci,roadmapbench,specbench,slopcodebench}. 
% We therefore need an evaluation setting that is controlled enough to isolate primitive self-evolution abilities, while still forcing agents to improve an executable behavior through real environment feedback.
We therefore need a controlled setting that isolates an agent's ability to convert bounded environment feedback into generalizable improvements of an executable policy, while retaining the iterative decisions that make autonomous improvement difficult.

We address this gap by formalizing \emph{Autonomous Policy Evolution}, a problem in which an agent repeatedly revises an executable decision policy using feedback from prior deployments. 
Formally, the observable object is the sequence of submitted policy systems and train-feedback records, whereas the outcome is the held-out return of the checkpoint selected on hidden validation.
The goal is not only to maximize observed performance alone, but to produce a policy that generalizes to environment instances after limited interaction. 
The bounded budget is therefore part of the capability being measured: it requires each agent to choose what information to acquire, when to explore or exploit, and how efficiently to convert sparse behavioral evidence into robust policy improvement.

To evaluate this problem, we instantiate \emph{Autonomous Policy Evolution} in \textsc{EvoPolicyGym}, a controlled benchmark built from compact interactive environments. 
Unlike open-ended engineering benchmarks \citep{frontiereng}, 
\textsc{EvoPolicyGym} makes policy evolution itself the evaluated object: an agent repeatedly edits an executable policy system, submits it under a fixed interaction budget, and receives server-generated feedback from sandboxed rollouts. 
Train submissions return aggregate and trajectory-level feedback, whereas validation and held-out cases remain server-side.
This protocol makes policy evolution itself the evaluated object, rather than direct task execution or open-ended engineering progress. Across environments derived from standard reinforcement-learning substrates \citep{mujoco,minigrid,gymnasium}, \textsc{EvoPolicyGym} records the full execution--feedback--revise trajectory, enabling comparisons of not only final held-out performance but also how agents diagnose failures, allocate budget, and balance exploration with exploitation.


We conduct preliminary experiments on the classical RL gymnasium environment.
We evaluate four harness--model agents on Core16, a 16-environment suite spanning Gym/Box2D, MuJoCo, MiniGrid, and robotics/driving tasks, under a common 128-episode interaction budget. The results show that GPT-5.5 obtains the highest aggregate rank score and top-two performance on all 16 environments, whereas Claude Opus 4.7 leads the MiniGrid family. The remaining agents achieve isolated task wins but substantially lower cross-environment coverage.
% The results show that policy-evolution performance is highly environment-dependent: 

% Beyond these scores, our trajectory analysis shows that agents adopt different feedback-use strategies, including patching observed failures, using visual perception for control refinement, and trading off early exploration against late exploitation.
Beyond final scores, our analysis on aggregate edit statistics and selected-policy structure reveal systematic differences between structural synthesis and parameter tuning, while audited CarRacing and BipedalWalker traces illustrate how individual agents translate visible feedback into revisions.
Thus, \textsc{EvoPolicyGym} serves both as a leaderboard and as a diagnostic substrate for studying how self-evolving agents interact with environment feedback under bounded budgets.

To summarize, our contributions are threefold:
\begin{itemize}
    \item We formulate \emph{Autonomous Policy Evolution} as a benchmarkable setting for evaluating agents in policy searching tasks.
    \item We instantiate this setting in \textsc{EvoPolicyGym}, a controlled benchmark with strict visibility boundaries, bounded interaction, trajectory-level feedback, and hidden held-out generalization. 
    \item We introduce trajectory-level diagnostics that relate policy improvement to budget-conditioned policy evolution and audited trace revisions.
\end{itemize}

