\section{Introduction}

\input{Figures/fig_concept}

Large language model (LLM) agents are now routinely deployed as digital assistants that read documents, invoke external tools, and operate over complex environments~\cite{ReAct, SWE-agent, wu2024autogen}. Yet despite their growing capability, these agents remain reactive: they act only after a user issues an explicit request, whether to summarize a file, schedule a meeting, fix a failing test, or invoke a particular tool. This interaction model presumes that the user already knows what is wrong and what to ask.

In practice, however, the most consequential issues are often the ones a user has not yet noticed: a budget approval given verbally but not yet recorded in writing, holding up a vendor order on a hard deadline; two copies of the same report with conflicting numbers, both heading into an upcoming review; a recurring meeting the team has tacitly stopped attending, still blocking the only window for an urgent kickoff. Such issues sit, often in plain sight, inside the very documents, emails, and calendar entries that the agent could in principle inspect.

These otherwise different issues share a common structure that extends beyond the workspace setting above. Across the digital environments where agents operate (a personal workspace, a software repository, or another rich working context), evidence accumulates in which multiple problems coexist; none is articulated as a request; the number of coexisting problems is not known in advance; and resolving only the most salient one leaves the rest untouched. We therefore argue that proactive assistance is best framed not as anticipating a single user intent, but as the broader task of \textit{discovering multiple hidden problems from context}. Existing work on proactive agents has studied when to intervene~\cite{DBLP:conf/chi/LiuFSWIC25, DBLP:conf/emnlp/ZhangDLMKDCM25} or how to anticipate a single localized need~\cite{ProactiveAgent, ContextAgent, probe}, but largely sidesteps this multi-problem, context-wide setting that real workflows demand.

Meeting this task requires two complementary capabilities: broad coverage over coexisting problems that compete for attention with more salient ones, and enough precision per candidate to be actionable rather than speculative. To address these challenges, we present \textsc{TIDE} (\textbf{T}emplate-guided \textbf{I}terative \textbf{D}iscovery and r\textbf{E}solution; Figure~\ref{fig:concept}), a framework that combines two mechanisms operating along distinct axes. Specifically, motivated by the observation that single-pass prediction anchors on the most salient cases and yields generic claims, we propose \textit{iterative discovery}: rather than producing all predictions in one pass, the agent surfaces a small batch of candidates per round while conditioning on what has already been found, so that subsequent rounds are pushed beyond the cumulative discovery state. Additionally, for every surfaced candidate, the agent retrieves the supporting artifacts that serve as evidence and commits to a concrete action that proposes a resolution, so that the per-case output forms a multi-element plan that simultaneously identifies, grounds, and addresses each of the discovered problems. We then complement iteration with \textit{thought templates}: reusable schemas distilled from previously solved cases, each capturing a recognizable pattern of hidden problem and laying out the chain of contextual signals from which that pattern can be inferred, anchoring each prediction in a known problem class instead of leaving the agent to infer it from scratch. Iteration thus expands the set of problems the agent will consider, while templates provide a reusable prior on how problems manifest in evidence, sharpening each prediction.

We instantiate this \textsc{TIDE} framework in two different real-world settings that share the underlying multi-problem structure: personal workspaces, where the agent surfaces multiple unresolved bottlenecks from workspace documents and proposes how to address each, and software repositories, where the agent identifies multiple hidden bugs directly from source code and produces patches that resolve them. Across both settings and four LLM backbones, \textsc{TIDE} consistently outperforms single-shot and parallel multi-agent baselines on task coverage, identification, and resolution; analyses further show that iterative discovery and thought templates contribute largely complementary gains and that templates transfer across backbones. Taken together, these results suggest that proactive assistance is better cast as an explicit, multi-step discovery process over context than as single-shot prediction from a user request, and they offer a general recipe for building agents that can both surface what users would not have thought to ask and point toward how to address it.
