In recent years, Large Language Models (LLMs) are rapidly evolving, integrating deeper reasoning and extending into complex multimodal domains. Recent advances, such as OpenAI GPT-5.5 (\cite{openai2026gpt55}), Claude Opus 4.8 (\cite{anthropic2026claude}), Gemini 3.5 Flash (\cite{deepmind2026gemini}), and Qwen3.7 (\cite{qwenteam2026qwen}), demonstrate substantial progress in multimodal reasoning, long-context understanding, and specialized tool execution. These models increasingly exhibit the ability to decompose complex problems and process dynamic, information-dense visual streams. 

Building upon the strong foundation of our previous works, Keye-VL (\cite{kwaikeye2025vl}) and Keye-VL-1.5 (\cite{yang2025kwaikeyevl15}), which established robust vision-language alignment and state-of-the-art short-video comprehension, we introduce Kwai Keye-VL-2.0-30B-A3B. As we push the model's frontier from short-form visual perception to long-horizon agentic reasoning, two critical roadblocks naturally emerge. First is the infrastructural bottleneck: scaling to extreme video contexts incurs prohibitive computational and memory costs. Second is the algorithmic dilemma: integrating complex, heterogeneous agent tasks often induces catastrophic forgetting of foundational reasoning capabilities. In this report, we detail how Keye-VL-2.0 elegantly resolves these dual challenges by introducing paradigm-shifting improvements in architecture and alignment, focusing on the following two key aspects.

\textbf{Extreme Context Scaling via Multimodal DSA.} A key insight from extending multimodal models to 256K contexts is that standard dense attention mechanisms inevitably lead to catastrophic KV cache expansion and computational walls, forcing models to sacrifice temporal continuousness through aggressive frame subsampling. To overcome this latency and scalability limit, Keye-VL-2.0 introduces a 30-billion-parameter Mixture-of-Experts (MoE)(\cite{shazeer2017outrageously}) foundation with only 3 billion active parameters, ensuring exceptional deployment efficiency. Crucially, we pioneer the application of Multimodal DeepSeek Sparse Attention (DSA)(\cite{deepseek2025v32}) in the visual domain. By compressing and sparsifying video feature aggregation, DSA effectively constrains the linear growth of the KV cache. This design allows Keye-VL-2.0 to process 256K extreme long-video contexts losslessly, transforming video understanding from frame-limited perception to global-context reasoning while maintaining high inference efficiency.

\textbf{Resolving Modality Conflict via Cross-Modal MOPD.} Existing models often suffer from the "Multimodal Alignment Dilemma": directly injecting complex video understanding and tool-use capabilities often triggers catastrophic forgetting, degrading the model's foundational STEM, mathematical, and linguistic reasoning abilities. To circumvent the ambiguity and instability of end-to-end co-optimization, we introduce Cross-Modal Multi-Teacher On-Policy Distillation (MOPD)(\cite{mimoteam2026v2flash}). Through an innovative dynamic routing mechanism, MOPD leverages specialized teacher models to provide dense token-level feedback on student-generated trajectories across varying modalities and tasks during post-training. By supervising on-policy rollouts, MOPD effectively isolates task-specific expertise, which is subsequently distilled and seamlessly merged back into the unified MoE backbone. This bidirectional alignment ensures that Keye-VL-2.0 achieves substantial leaps in native agent capabilities (e.g., Code, Tool Use, Web Search) while robustly preserving its general-purpose reasoning baselines. During post-training, we further apply Context-RL(\cite{lu2026contextrl}) and Video-RL, equipped with bucket advantage scaling, to stabilize long-sequence decision trees and systematically reduce visual hallucinations.

As illustrated in Figure \ref{fig:teaser}, Keye-VL-2.0-30B-A3B demonstrates highly competitive performance against both open-source and closed-source state-of-the-art models. It achieves top-tier results in fine-grained temporal localization across TimeLens benchmarks (\cite{zhang2025timelens}) (ActivityNet, QVHighlights, Charades), substantially outperforming leading models such as Gemini-3-Flash~(\cite{deepmind2025gemini3flash}) on several temporal grounding settings. Furthermore, it exhibits a non-linear capability scale-up on extreme long-context evaluations, including Video-MME-v2 (\cite{fu2026videomme2}) and LongVideoBench (\cite{wu2024longvideobench}), proving its robustness in processing extensive temporal information. By resolving the dual challenges of context scaling and multi-task capability conflict, we offer a powerful, efficient, and versatile multimodal foundation for the open-source community, enabling researchers and developers to explore, refine, and deploy scalable multimodal applications.
