% Auto-extracted exemplar: 1 pattern + 1 sub-pattern.
% Sources: methodology_cards.tex (Audit and Pivot), tactical_cards_appendix.tex (C01).

\begin{tcolorbox}[colback=bgskill, colframe=cblue!70, title={\textbf{Audit and Pivot an Assumption} \textcolor{gray}{\small\texttt{assumption\_audit\_and\_pivot}} \hfill \textit{Audit the load-bearing assumption and pivot}}, breakable, before skip=4pt, after skip=4pt]
\textcolor{gray}{\small Cluster-level primary: Oral $n=94$ | HC $n=11$ | Reject $n=79$ | Total $n=181$}

\textbf{Definition.} Locate the load-bearing implicit assumption a result, guarantee, or defense rests on, then pivot on it: relax it to a weaker condition and re-prove (extending the guarantee), or violate it with a constructed counterexample/exploit (breaking the system or unlocking new behavior).

\textbf{Operational signature.} \textit{identify the implicit assumption a result or defense rests on $\to$ relax it (weaker condition) or violate it (counterexample/exploit) $\to$ re-derive the guarantee or demonstrate the new behavior}

\textbf{When to apply.} When a result's strength or a system's safety hinges on an assumption that real settings can weaken or that an adversary can violate.

\textbf{Step-by-Step:}
\begin{enumerate}[leftmargin=14pt,topsep=2pt,itemsep=1pt]
  \item Surface the single implicit assumption the target result, guarantee, or defense silently rests on --- the precondition the whole subfield has inherited as fixed (a quantity treated as immutable, a property treated as protective, a structure treated as required). Name it explicitly and show it is load-bearing: the existing result holds *because* of it. Do not pivot on an assumption whose removal leaves the achievable outcome unchanged --- relaxing a non-binding assumption yields only an incremental generalization, the cluster's most common rejection.
  \item Choose the pivot direction deliberately. Either relax the assumption to a strictly weaker condition and plan to re-derive the guarantee (extending it to settings previously thought impossible), or violate it with a constructed counterexample or exploit to break the system or unlock new behavior. The direction must change what is achievable --- reach an outcome the assumption was believed to forbid --- not merely restate the prior result under a cosmetic reframing.
  \item Establish that the pivoted condition is real, not a substitution of one strong assumption for another. For a relaxation, verify empirically --- in natural data or the actual deployment regime --- that the weaker condition genuinely holds, and confirm you have not smuggled in an equally restrictive or unverifiable replacement; the recurring failure is trading a known assumption for a distributional form or abstract regularity condition no real instance is shown to satisfy. For a violation, confirm the assumption is breakable by a realistic, resource-bounded actor rather than an idealized one.
  \item Re-derive rigorously and certify tightness. For a relaxation, prove the guarantee under the weaker condition and supply a matching lower bound, an exact equivalence, or an impossibility result, so the pivot is certified as the true barrier rather than an artifact of loose analysis. For a violation, demonstrate the exploit and trace its success to a structural mechanism of the pipeline, turning an isolated trick into a systematic, reusable template --- empirical success with no identified mechanism is read as anecdote.
  \item Prove the pivot was necessary and differentiate it. Show the prior approach fails precisely because of the audited assumption via an explicit counterexample or failure-case characterization, and sharply distinguish your move from any prior work that pivoted similarly. Avoid bundling off-the-shelf components and crediting the bundle, and never name a tool or framework as the contribution without exercising it in the actual derivation --- both are recurring rejection shapes in this cluster.
\end{enumerate}
\textbf{Success conditions} (from Oral):
\begin{itemize}[leftmargin=12pt,topsep=2pt,itemsep=2pt]
  \item \textbf{The relaxed or violated condition is shown to hold in the real setting, not merely assumed --- measured in natural data, or demonstrated reachable by a bounded adversary.} \\ A pivot is only valuable if the weaker precondition is actually satisfied where the method will be used; grounding the condition empirically converts a theoretical relaxation into a usable one and pre-empts the reviewer suspicion that one strong assumption was swapped for another.
  \item \textbf{A relaxed upper bound is paired with a matching lower bound, exact equivalence, or impossibility result that certifies the pivot as the genuine barrier.} \\ Tightness certificates separate a loose extension from a sharp one: they prove the audited assumption --- not the analysis technique --- was what limited prior results, which is the structural claim the contribution actually makes.
  \item \textbf{The audited assumption is replaced by a strictly weaker or structurally different one, often shifting the burden from the prior distribution to the geometry of the map or from pointwise to expectation-level conditions.} \\ Moving the constraint to a different and milder level of the problem is what unlocks new achievable results; a same-strength substitution re-imports the original barrier and leaves the frontier unchanged.
  \item \textbf{For exploits and violations, the success is traced to a structural property of the pipeline and packaged as a systematic design template rather than a one-off attack.} \\ A root-cause-level explanation generalizes across instances and gives the community a reusable lens, whereas an isolated demonstration is fragile and reads as engineering rather than insight.
  \item \textbf{Prior methods are shown to fail precisely because of the audited assumption, via explicit failure-case characterization presented before the positive result.} \\ Demonstrating that the assumption is the binding constraint --- not an incidental simplification --- is what makes the pivot necessary rather than optional, and reviewers consistently reward this clarity.
\end{itemize}
\textbf{Failure modes} (from Reject):
\begin{itemize}[leftmargin=12pt,topsep=2pt,itemsep=2pt]
  \item \textbf{Pivoting on an assumption whose relaxation produces no new achievable result, yielding a generalization that reviewers read as incremental.} \\ When the weaker condition leaves the attainable guarantee or rate unchanged, the contribution collapses to 'we also handle case X' with no unlocking; the absence of a qualitatively new capability is the most cited reason these attempts fall short.
  \item \textbf{Replacing the audited assumption with an equally strong or unverifiable substitute --- a restrictive distributional form, or an abstract regularity condition no known instance is shown to satisfy.} \\ The substitution re-imports the very barrier the paper claimed to remove, and reviewers flag that the new condition is untestable or unmotivated, so the claimed relaxation is not actually a relaxation.
  \item \textbf{Demonstrating an exploit or violation without validating it against an adaptive or informed defender.} \\ The central claim of a violation is that the property is hard to counter; showing only that an attack succeeds against a static target leaves that claim unproven, and the 'commonness implies safety' or 'stealth' arguments remain intuitive rather than substantiated.
  \item \textbf{Stacking several off-the-shelf components and attributing the gain to the bundle without isolating which pivot is load-bearing.} \\ Reviewers cannot identify a unifying principle and read the work as an engineering combination; without an ablation pinning the benefit to a specific structural move, the novelty above the constituent parts is unconvincing.
  \item \textbf{Naming a theoretical tool or framework as the contribution while never exercising it in the proofs, or shipping an error in the stated result.} \\ If the advertised machinery (a continuous-time analysis, a new perspective) is absent from the actual derivation, or a headline rate is wrong, confidence in the entire pivot evaporates regardless of the idea's appeal.
  \item \textbf{Failing to sharply differentiate from prior work that already performed a near-identical prior pivot.} \\ In crowded areas a correct but under-differentiated move reads as redundant; reviewers point to the closest precedent and demand a decisive reason to prefer the new approach over that pivot, which an unfocused positioning cannot supply.
\end{itemize}
\textbf{Oral vs Reject gap.} Accepted executions perform three observable moves that rejected ones skip. First, after naming the assumption they verify the pivoted condition concretely --- measuring that the weaker condition holds in real or natural data, or that the violated property is reachable by a bounded adversary --- whereas rejected papers relax an assumption and immediately substitute an equally strong or unverifiable one (a linear-Gaussian form, an abstract regularity condition with no exhibited satisfying function class). Second, accepted relaxations ship a tightness certificate --- a matching lower bound, an exact algebraic equivalence, or an impossibility result --- while rejected ones offer only a one-directional upper bound or a bare empirical demonstration, leaving open whether the pivot was the real barrier or just loose analysis. Third, accepted papers establish that prior methods fail precisely because of the audited assumption via an explicit counterexample or failure-case characterization, and for exploits they trace success to a structural pipeline mechanism that becomes a reusable template; rejected papers instead present the work as a generalization or an isolated trick, bundle off-the-shelf components without isolating the load-bearing pivot, or never exercise the theoretical tool they claim as their contribution.

\textbf{Oral vs HC gap.} With a moderate HC sample (\textasciitilde{}10 papers, several of them attacks, benchmarks, or scaling studies), the pattern is that HC papers pivot on an assumption and characterize the consequence at scale or package it as urgently needed reusable infrastructure --- a power-law fit for an attack, a measured safety/helpfulness trade-off, a large boundary-targeting benchmark, a broadly transferable defense --- but typically stop at empirical characterization or a single bound. Oral papers add a tightness certificate on top of the same pivot: a matching lower bound, an exact equivalence, an identifiability proof, or an impossibility/optimality result that converts 'this works and here is how well' into 'this is provably the right lever and here is the exact boundary.' Oral papers also more often formalize the failure modes of prior approaches before presenting the positive result, so the pivot reads as a principled diagnosis rather than a strong empirical finding. Given the sample size this is a tendency rather than a strict law.

\textbf{Reviewer expectations:}
\begin{itemize}[leftmargin=12pt,topsep=2pt,itemsep=1pt]
  \item \textit{[both]} When a result rests on a relaxed assumption, supply a matching lower bound or tightness certificate confirming the analysis is not loose; the absence of a matching bound is repeatedly cited as a reason to hold back.
  \item \textit{[reject\_reviews]} Characterize the new or weakened assumption --- show which function classes, distributions, or real settings actually satisfy it --- rather than introducing it as a convenient condition that merely lets the proof go through.
  \item \textit{[reject\_reviews]} For attacks and violations, validate against adaptive or informed defenders, not just a static target, before claiming the violated property is hard to counter.
  \item \textit{[both]} Sharply differentiate the pivot from prior work that made a similar move, giving a decisive reason to prefer it; overlap with a close precedent (a prior drift-correction, a prior attack surface) must be addressed head-on.
  \item \textit{[reject\_reviews]} Do not advertise a theoretical framework or tool as the contribution unless it is actually used in the derivation, and ensure headline rates and claims are error-free.
  \item \textit{[oral\_reviews]} Establish and cleanly characterize the failure cases of prior approaches before presenting the positive result, so the pivot is seen as a principled diagnosis of the binding constraint.
\end{itemize}
\textbf{Cognitive barriers:}
\begin{itemize}[leftmargin=12pt,topsep=2pt,itemsep=1pt]
  \item The audited assumption is invisible because the entire subfield inherited a framing that treats it as a fixed precondition rather than a controllable choice --- a 'fixed' local objective, a 'required' supervision signal, a trigger that 'must' live in the input. You cannot relax what you do not first see as a variable, so the hard part is denaturalizing a convention everyone reasons within.
  \item A stated safety or design property is mistaken for an inherent guarantee: practitioners read 'reduced probability mass,' 'norm-bounded imperceptibility,' or 'scale dilutes adversaries' as protection, which masks that the very property is the exploitable lever. Reframing a feature as a vulnerability requires inverting the intent the property was introduced to serve.
  \item The pivot often hinges on importing a tool from an adjacent domain and recognizing it transfers --- concentration tools from sequential testing, blocking from dependent-sequence analysis, scalarization from multi-objective optimization. Because the tool was built for a different purpose, its applicability to the audited assumption is not visible from within either field's standard practice.
  \item The relaxation frequently lives at a different level of the problem than the natural formulation suggests --- reparameterizing the latent basis instead of the observed variables, constraining the geometry of the generative map instead of the prior distribution, measuring increments or contrasts instead of raw values. The obvious formulation hides the lever, so the move feels unreachable until one deliberately changes the level of analysis.
\end{itemize}
\textbf{Examples} (paper-agnostic lessons):
\begin{itemize}[leftmargin=12pt,topsep=2pt,itemsep=1pt]
  \item \textbf{[Oral lesson]} When a guarantee is blocked by a quantity treated as fixed, ask whether that quantity is actually a controllable variable you can reshape cheaply each step --- turning a hard constraint into a degree of freedom often dissolves the barrier.
  \item \textbf{[Oral lesson]} Measure that the weaker condition you rely on actually holds in real data before claiming the relaxation is practical; a precondition verified in natural statistics is worth far more than one merely assumed.
  \item \textbf{[Oral lesson]} Pair every relaxed upper bound with a matching lower bound, exact equivalence, or impossibility result, so the pivot is certified as the true barrier rather than an artifact of loose analysis.
  \item \textbf{[Oral lesson]} Replace a distributional assumption with a structural or architectural one when the problem's geometry already encodes the constraint --- identifiability can rest on the shape of the generative map rather than on the prior over latents.
  \item \textbf{[Oral lesson]} When building an exploit, trace its success to a structural property of the pipeline and convert it into a reusable design template rather than presenting an isolated trick.
  \item \textbf{[Oral lesson]} Read every 'bounded,' 'small,' or 'imperceptible' guarantee as a budget an adversary can neutralize --- a stated protective property is frequently the exact lever to violate.
  \item \textbf{[Reject lesson]} Relaxing an assumption that leaves the achievable result unchanged reduces the contribution to a generalization reviewers read as incremental.
  \item \textbf{[Reject lesson]} Swapping the audited assumption for an equally strong or unverifiable one --- a restrictive distributional form, or a regularity condition no known instance is shown to satisfy --- re-imports the barrier you claimed to remove.
  \item \textbf{[Reject lesson]} Demonstrating an exploit only against a static defender leaves the central claim, that the violation is hard to counter, unproven.
  \item \textbf{[Reject lesson]} Bundling several off-the-shelf components and crediting the bundle, without isolating which pivot is load-bearing, invites a 'no unifying principle' rejection.
  \item \textbf{[Reject lesson]} Naming a theoretical framework as the contribution while never exercising it in the proofs --- or shipping an error in the stated rate --- collapses confidence in the whole result.
  \item \textbf{[Reject lesson]} Failing to sharply distinguish the work from a prior near-identical pivot makes even a correct result read as redundant in a crowded area.
\end{itemize}
\end{tcolorbox}

\noindent\textit{Representative sub-pattern under this pattern (the tactical recipe a Phase~2.2 candidate instantiates):}

\begin{tcolorbox}[colback=bgskill!50, colframe=cgreen!70, title={\textbf{C01} \textcolor{gray}{\small (O20/H9/R21)}}, breakable, before skip=4pt, after skip=4pt]
\textbf{Tactical pattern.} Across this cluster the audited assumption is always about the *representational locus* at which a security, detection, provenance, or safety property is presumed to be enforced --- and the tactical move is to relocate the attack or defense to a different locus the field treated as irrelevant or invariant. Concretely: triggers are moved from the input image to the supervision/label co-occurrence space; the comparison signal for malicious updates is moved from parameter/gradient divergence to scalar-score distribution geometry; the detection feature is moved from a scalar confidence to local curvature, positional-depth profile, or the full sampled output distribution; provenance is moved from raw parameter comparison to transformation-invariant kernel alignment over intermediate activations; defense interception is moved from the output boundary to internal state trajectories. For attacks, the papers show the overlooked locus is exploitable at far lower cost or access than the assumed locus implied; for defenses, they show the overlooked locus carries an attack-agnostic, transformation-invariant signal the assumed locus lacks. The unifying verbs are "relocate the signal," "re-anchor enforcement," and "demonstrate exploitability/robustness at the new locus."

\textbf{Step-by-Step:}
\begin{enumerate}[leftmargin=14pt,topsep=2pt,itemsep=1pt]
  \item Name the single representational locus that current attacks or defenses in the target security/detection setting all operate on (input perturbation, parameter/gradient divergence, scalar confidence score, deterministic mode output, externally-injected artifact), and state the tacit sufficiency-or-invariance claim bolted to it as a falsifiable proposition about where the load-bearing signal lives. The sharper the claim ('the trigger must be in the input', 'scale dilutes contamination', 'the scalar score is the signal'), the more leverage the pivot has.
  \item Identify an alternative locus --- supervision records, intermediate-activation geometry, score-distribution shape, the full stochastic output distribution, behavioral input-output traces, internal-state trajectories --- and argue structurally why the property of interest must manifest there too, ideally MORE robustly: invariant to the very transformations or access restrictions that defeat the assumed locus. If the alternative is only a cosmetic variant of the assumed locus, reviewers fold the contribution back into prior work, so the distinction must be load-bearing and explicit.
  \item Build the mechanism at the new locus: for an attack, the lowest-access, lowest-budget operation that exploits it; for a defense, a signal demonstrably agnostic to the attacker's configuration or insertion point. Where possible, couple detection and action through the same locus signal so the move is one unified primitive rather than two bolted-on stages.
  \item Validate against the strongest adversary that actually knows the new locus --- an adaptive attacker for a defense, or the realistic access/budget constraints for an attack --- not merely a fixed, non-adaptive baseline. Quietly demonstrating the move only against the bypassed locus, or against outdated baselines, is the single most common reason this tactic fails to convince.
  \item Supply a mechanistic account of WHY the new locus carries the signal --- an invariance proof, an impossibility-style bound, or a root-cause decomposition of the pipeline --- not merely evidence THAT it does. A relocation shown to work but never explained reads as a stacked combination of existing components or an anecdote, and forfeits the conceptual-lens credit that elevates these papers.
\end{enumerate}
\textbf{Differentiation within parent.} \textit{All siblings share the parent's audit-and-pivot signature, but they pivot mathematical objects inside a fixed locus, whereas this cluster pivots the locus itself in adversarial/empirical security and safety settings and validates with attack-success rates and detection performance rather than tighter analysis. Cluster 11 re-derives privacy-accounting bounds by reframing repeated selection as a resource-bounded mechanism --- the assumption is relaxed but the quantity (a cumulative bound) and its locus (the accounting calculus) are unchanged; this cluster instead relocates the enforcement point entirely (e.g., from parameter-divergence to score-distribution geometry) and shows the old locus was the wrong place to look. Likewise cluster 29 proves a new property of a projection operator on a constraint set and cluster 5 derives an exact algebraic equivalence between model families --- both are theorem-driven reformulations of analytic structure, while this cluster's deliverable is a demonstrated attack surface or a transformation-invariant detection signal. The contrast is empirical-locus-shift-under-adversary, not closed-form quantity tightening.}

\textbf{When to pick this one.} Pick this instantiation when the gap sits in a security, safety, detection, or provenance setting and you can observe that every existing attack or defense is anchored at one representational locus (raw input, parameters/gradients, a scalar score, a deterministic output, or an injected artifact). The trigger to instantiate here is a suspicion that the real load-bearing signal lives at an overlooked locus --- labels, activations, distribution shape, or behavior --- that the field has tacitly assumed irrelevant or invariant.

\textbf{Tactical failure mode.} The recurring failure is a successful relocation that is demonstrated but not adversarially closed: the paper shows its new-locus attack or defense beats a fixed, non-adaptive baseline, yet never stress-tests against an informed adversary who also operates at the new locus, so stealthiness, evasion, or robustness claims are asserted rather than proven. A second shape is a novelty-boundary collapse --- the "new" locus turns out to be one a prior paper already used (normalization statistics, activation-similarity alignment, pre-encoded semantic concepts), and the work fails to quantify what is genuinely different. A third is reintroducing the very assumption the relocation was meant to dissolve, by depending on a clean reference model, a known anomaly distribution, or other access the realistic threat model denies. A fourth is offering only a demonstration with no mechanistic account of why the locus carries the signal, which reads as a stacked pipeline of off-the-shelf parts. Self-check: can an adversary who knows your new locus defeat it as cheaply as the old one, is your locus provably distinct from published loci, and can you state structurally WHY the signal lives there?

\textbf{Examples} (paper-agnostic lessons):
\begin{itemize}[leftmargin=12pt,topsep=2pt,itemsep=1pt]
  \item \textbf{[Oral lesson]} Moving a trigger or detection signal from the input space to the supervision/label space opens an attack surface the entire defense literature was built around the wrong locus to catch.
  \item \textbf{[Oral lesson]} A detection or provenance signal anchored in intermediate-representation geometry survives exactly the transformations --- fine-tuning, pruning, unit reordering, rescaling --- that destroy parameter- and output-level signals.
  \item \textbf{[Oral lesson]} Treating the shape of a learned distribution (its local curvature, positional-depth profile, or full sampled spread) as the signal, rather than a scalar summary, surfaces structure that point estimates provably discard.
  \item \textbf{[Oral lesson]} Replacing a randomized, probabilistic guarantee with a deterministic combinatorial construction at the same structural locus can upgrade a soft bound into an exact certificate.
  \item \textbf{[Oral lesson]} Showing that a stated security property itself --- an imperceptibility norm bound, scale-induced dilution, or deterministic evaluation --- is the exploitable handle converts a design parameter into a provable vulnerability.
  \item \textbf{[Oral lesson]} Decomposing an empirically observed failure into structural root causes of the training pipeline turns ad hoc attack discovery into a systematic, predictive design process.
  \item \textbf{[Reject lesson]} Claiming a relocated signal is 'stealthier' or 'harder to remove' without defeating an adaptive defender who searches the same new locus leaves the central claim unverified.
  \item \textbf{[Reject lesson]} Relocating to a locus already exploited by published work --- normalization statistics, activation-similarity losses, semantic concept triggers --- collapses the novelty boundary unless the difference is made explicit and measured.
  \item \textbf{[Reject lesson]} A defense whose relocation silently depends on a clean reference model or a known anomaly distribution reintroduces precisely the strong assumption the move was supposed to eliminate.
  \item \textbf{[Reject lesson]} Demonstrating the new locus only on fixed, non-adaptive or outdated baselines while omitting any account of why it works reads as a stacked pipeline of existing components rather than a principled shift.
  \item \textbf{[Reject lesson]} Wiring off-the-shelf generators or tools into a new attack setting without a structural reason for the locus change gets judged as engineering integration, not insight.
\end{itemize}
\end{tcolorbox}
