\section{Introduction}
Interactive world models~\cite{genie3,Gamegen-x,Yume,matrix2,matrix3,Worldplay,team2026dreamx}, generative systems that synthesize an environment frame by frame in response to a stream of user or agent actions, have recently emerged as a promising substrate for game generation ~\cite{Gamegen-x,solaris,matrix2,matrix3,worlddirector} and
embodied simulation ~\cite{cosmos,Abot-physworld,giga}. Driven by advances in causal (autoregressive) video generation ~\cite{CasualVid,MAGI-1,SkyReels,yang2025longlive,longlive_2.0,minWM},
such models can in principle render an explorable world that unfolds indefinitely and reacts to its inhabitants in real time. Yet turning this principle into a usable system has remained difficult, for two reasons.

The first concerns \textbf{long-horizon stability}. Because each frame is conditioned on the frames generated before it, errors are fed back into the model and accumulate; over time, textures smear, geometry warps, and the scene drifts away from any plausible world. Most existing systems therefore remain visually stable for only seconds to a few minutes before degrading, far short of the persistence one expects from a world meant to be lived in. The second concerns \textbf{interactivity at high fidelity}, which is computationally expensive: rendering detailed video while reacting to live input strains compute
budgets, and prior work has typically bought interactivity by sacrificing resolution, smoothness, or control, leaving the user with little more than coarse camera movement through an otherwise inert scene.

We argue that these two limitations are what stand between current models and genuinely open-ended worlds, and we tackle them directly. Our starting point is a causal video generation model trained to resist the error accumulation described above; the resulting backbone holds its quality far longer than prior open models and, in turn, makes a clean teacher for distillation. The distilled student is what makes the system practical: it sustains an explorable world that never has to end, at a resolution and frame rate high enough to feel responsive rather than merely playable. We stress-test this claim with an hour-long uninterrupted session in which quality shows no visible decay, evidence that the model's stability is structural rather than a short-lived artifact of a favorable clip. And rather than confining the user to wandering a frozen landscape, the model responds to an expressive vocabulary of actions such as combat, archery, spell-casting, and ranged fire, and even reshapes the environment itself, summoning weather such as snow or rain on demand.

Even a capable backbone, however, does not \emph{play itself}~\cite{Hu2024StoryAgentCS,Lian2023LLMgroundedVD,Huang2023FreeBloomZT}. Here we take a cue from how today's coding assistants are built: a strong language model becomes a working software engineer only once it is placed inside a harness, a scaffold such as Codex that lets it inspect state, act, and chase a goal across many turns. We adopt the same philosophy for world modeling. Around our generative core we wrap an \textbf{agentic harness} in which a \emph{pilot} agent reads the current scene and decides what the controllable character should do next, whether stepping forward, swinging at a foe, or reaching for a nearby object, while a \emph{director} agent keeps the world from running dry, seeding fresh content, props, and events as exploration proceeds. The two together turn a frame predictor into something that behaves like a world: self-sustaining, goal-directed, and open-ended, with no author scripting each moment in advance.

In summary, this paper makes the following contributions:

\begin{itemize}
    \item \textbf{An open, state-of-the-art causal world model} whose key advance is durability: leading visual quality paired with strong resistance to drift, giving the community both a usable backbone and a capable distillation teacher.

    \item \textbf{A real-time distilled model} that renders an unbounded, drift-free interactive world at 720p and 60\,fps, verified by over an hour of continuous generation without quality loss.

    \item \textbf{A rich, controllable action space} that goes well beyond navigation to include character actions such as combat, archery, spell-casting, and shooting, alongside on-the-fly environmental changes.

    \item \textbf{A pilot and director agentic harness} that orchestrates the model into a living interactive experience.
\end{itemize}

\begin{table}[t]
    \small
    \centering
    \caption{
        \textbf{Comparison with recent interactive world models.} \method stands out as the only model to achieve hour-level (infinite) generation duration within a general domain. Furthermore, it uniquely combines this extended continuous generation with high dynamic degree, semantic interaction, and real-time performance, while being fully open-sourced.
    }
    \label{tab:comparison}
    \SetTblrInner{rowsep=1.2pt}      
    \SetTblrInner{colsep=4.6pt}     
    \definecolor{linegray}{HTML}{BDBDBD} 
    \definecolor{bg_gray1}{HTML}{FAFAFA}
    \definecolor{bg_gray2}{HTML}{F2F2F2}
    \definecolor{bg_gray3}{HTML}{EAEAEA}
    \definecolor{bg_gray4}{HTML}{E2E2E2}
    \definecolor{bg_gray5}{HTML}{DADADA}
    \definecolor{bg_purple}{HTML}{6A67F3}
    \begin{tblr}{
        cells={halign=l,valign=m},
        column{1}={bg=white},
        column{2}={bg=bg_gray1},
        column{3}={bg=bg_gray2},
        column{4}={bg=bg_gray3},
        column{5}={bg=bg_gray4},
        column{6}={bg=bg_gray5},
        column{7}={bg=bg_purple, fg=white},
        hline{2}={0.5pt, fg=linegray},
    }
    \ & \textbf{M-G 3.0}~\cite{matrix3} & \textbf{D-W}~\cite{team2026dreamx} & \textbf{LingBot-World}~\cite{lingbot-world} & \textbf{HappyOyster}~\cite{happyoyster} & \textbf{Genie 3}~\cite{genie3} & \textbf{Ours} \\
    Generation Duration & Minutes & Minutes & Minutes & Minutes & Minutes & Hours (Infinite) \\ 
    Semantic Interaction & None & None & None & Few & Few & Infinite \\
    Domain & Game & General & General & General & General & General \\ 
    Dynamic Degree & Medium & Medium & High & Medium & Medium & High \\ 
    Real-time & \ding{51} & \ding{51} & \ding{51} & \ding{51} & \ding{51} & \ding{51} \\ 
    Open-source & \ding{51} & \ding{51} & \ding{51} & \ding{55} & \ding{55} & \ding{51} \\ 
    \end{tblr}
\end{table}