\section{Conclusion and Discussion}
\paragraph{Conclusion.}
We present \method, an open causal video generation
model for interactive world modeling. It pairs state-of-the-art visual quality with strong resistance to drift, and from it we distill a real-time model that sustains an unbounded, drift-free world at 720p and 60\,fps. Beyond navigation, the model supports a rich action space including combat, archery, spell-casting,
and shooting, along with on-the-fly environmental changes such as snow and rain. Finally, a pilot and director agentic harness orchestrates the model into a self-sustaining, goal-directed, and open-ended experience. We release our models and harness to the community.

\paragraph{Limitations and discussion.}
Several limitations remain. The most fundamental is \textbf{long-term memory}.
While the model stays visually stable over long horizons, it does not truly remember the world it has generated: a region that leaves the context window and is later revisited tends to be regenerated rather than recalled, so the world is persistent in appearance but not in identity. We are exploring mechanisms such as dynamic KV-cache scheduling to retain a larger and more relevant history, but this faces an inherent tension, since memory cost grows with the horizon while the available budget is bounded. Giving interactive world models genuine, compact long-term memory remains an open problem. A second limitation concerns \textbf{identity and style consistency}. Over very long rollouts, specific characters can subtly change in appearance and the overall art style may gradually drift, even when the scene as a whole stays coherent. A third limitation lies in \textbf{physical understanding}. The model learns dynamics purely from pixels, without an explicit notion of geometry or collision, so its grasp of the underlying physics is imperfect. As a result, physically implausible artifacts occasionally arise, such as characters or objects intersecting or passing through one another rather than colliding as they would in a real world. Endowing the model with a more faithful sense of physical interaction is an important direction for future work. Finally, although our distilled model runs in real time, it still requires substantial computational resources. Reaching real-time, high-fidelity world modeling on commodity hardware will need further gains in efficiency, which we see as an important direction going forward.