% Related Work
\section{Related Work}

\subsection{Agentic Image Generation}

Recent work extends image generation and editing with agent capabilities such as planning, reasoning, memory, search, and self feedback. Planning-based methods decompose complex intents into intermediate steps~\citep{Yao2026PhotoAgentAP}; Reasoning-based methods handle implicit user intent for more intelligent generation and editing~\citep{He2026MindBrushIA}; Search-based methods incorporate web search and image search to improve grounding in open-world scenarios~\citep{Feng2026GenSearcherRA,He2026MindBrushIA}; Memory-based methods support long-horizon interactions through persistent memory~\citep{He2026GEMSAM}; and Feedback-based methods study test-time scaling for image generation~\citep{Jiang2026GenAgentST,Wang2025ImAgentAU}. However, from the perspective of generation context, existing methods remain fragmented in how they identify, acquire, and use the context required for real-world image generation. In contrast, Qwen-Image-Agent unifies plan, reason, memory, search, and feedback within a single framework, bridging the context gap in real-world image generation.

\subsection{Benchmarks for Image Generation}

Early image generation benchmarks mainly evaluate instruction following and text--image alignment, such as GenEval~\citep{Ghosh2023GenEvalAO} for compositional attribute binding and DPGBench~\citep{Hu2024ELLAED} for dense prompt following. More recent benchmarks target harder settings that are either knowledge-driven or reasoning-driven. Knowledge-driven benchmarks, such as WISE~\citep{Niu2025WISEAW} and PhyBench~\citep{Meng2024PhyBenchAP}, evaluate grounding in domain knowledge and physical commonsense. Reasoning-driven benchmarks, such as RISEBench~\citep{Zhao2025EnvisioningBT}, test whether models can translate logical, causal, and spatio-temporal reasoning into visual outputs. Mind-Bench~\citep{He2026MindBrushIA} covers both aspects. However, existing benchmarks mainly evaluate partial agent abilities, especially reasoning or search, while largely overlooking planning and memory. To support holistic evaluation of agentic image generation, we introduce IA-Bench, which covers the full spectrum of agent capabilities with fine-grained, checklist-based evaluation.
