\textbf{Hy-Embodied-0.5-VLA.}
We fine-tune Hy-Embodied-0.5-VLA from the foundation checkpoint hosted at \url{https://huggingface.co/tencent/Hy-Embodied-0.5-VLA-UMI}.
For simulation experiments, we follow its official open-source evaluation setting on RoboTwin 2.0, training the model with a batch size of 128 for 200K steps and using 6 historical images sampled every 20 steps for memory encoding.

\textbf{Spatial Forcing.}
We fine-tune Spatial Forcing from the \texttt{gs://openpi-assets/checkpoints/pi05\_base} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 60K steps.

\textbf{$\pi_{0.5}$.}
We fine-tune $\pi_{0.5}$ from the \texttt{gs://openpi-assets/checkpoints/pi05\_base} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 60K steps.
For real-world experiments, we use the same batch size and train the model for 30K steps.

\textbf{X-VLA.}
We fine-tune X-VLA from the \texttt{2toInf/X-VLA-Pt} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 100K steps.
For real-world experiments, we use the same batch size and train the model for 50K steps.

\textbf{Xiaomi-Robotics-0.}
We fine-tune Xiaomi-Robotics-0 from the \texttt{Xiaomi-Robotics-0-Pretrain} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 100K steps.
For real-world experiments, we use the same batch size and train the model for 50K steps.

\textbf{X-WAM.}
We fine-tune X-WAM from the \texttt{Wan2.2-TI2V-5B} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 32 for 40K steps.

\textbf{GigaWorld-Policy.}
We fine-tune GigaWorld-Policy from the \texttt{Wan2.2-TI2V-5B} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 32.
The video optimization stage is trained for 50K steps (1 epoch), followed by the action optimization stage for 250K steps (2 epochs), resulting in 300K total optimization steps (3 epochs).

\textbf{StarVLA-$\alpha$.}
We fine-tune StarVLA-$\alpha$ from the \texttt{Qwen3-VL-4B-Instruct} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 128 for 100K steps.
For real-world experiments, we use the same batch size and train the model for 30K steps.

\textbf{GalaxeaVLA.}
We fine-tune GalaxeaVLA from the \texttt{G0Plus\_3B-base} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 32 for 4 epochs.
For real-world experiments, we use the same batch size and train the model for 4 epochs.

\textbf{LingBot-VLA.}
We fine-tune LingBot-VLA from the \texttt{lingbot-vla-4b} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 15K steps.

\textbf{EventVLA.}
We fine-tune EventVLA from the \texttt{Qwen3-VL-4B-Instruct} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 128 for 150K steps.

\textbf{Fast-WAM.}
We fine-tune Fast-WAM from the \texttt{Wan2.2-TI2V-5B} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 20K steps.

\textbf{AHA-WAM.}
We fine-tune AHA-WAM from the \texttt{Wan2.2-TI2V-5B} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 768 for 25K steps.

\textbf{$\pi_0$.}
We fine-tune $\pi_0$ from the \texttt{gs://openpi-assets/checkpoints/pi0\_base} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 60K steps.
For real-world experiments, we use the same batch size and train the model for 30K steps.

\textbf{GR00T-N1.7.}
We fine-tune GR00T-N1.7 from the \texttt{nvidias/GR00T-N1.7-3B} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 640 for 100K steps.
For real-world experiments, we use the same batch size and train the model for 50K steps.

\textbf{InternVLA-A1.}
We fine-tune InternVLA-A1 from the \texttt{InternVLA-A1-3B} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 64 for 60K steps.
For real-world experiments, we use the same batch size and train the model for 30K steps.

\textbf{SmolVLA (Single Task).}
We fine-tune SmolVLA (Single Task) from the \texttt{smolvla\_base} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 512 for 100K steps.

\textbf{GO-1.}
We fine-tune GO-1 from the \texttt{GO-1} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 96 for 4 epochs, corresponding to 77,484 optimization steps.

\textbf{LDA-1B.}
We fine-tune LDA-1B from the \texttt{LDA-pretrain} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 128 for 300K steps.

\textbf{MolmoAct2.}
We fine-tune MolmoAct2 from the \texttt{MolmoAct2} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 128 for 100K steps.

\textbf{ACT (Single Task).}
We train ACT (Single Task) from scratch.
For simulation experiments, the model is trained with a batch size of 128 for 6K steps.

\textbf{H-RDT.}
We fine-tune H-RDT from the \texttt{pretrain-0618/checkpoint-500000} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 90K steps.

\textbf{Spirit v1.5.}
We fine-tune Spirit v1.5 from the \texttt{Spirit-v1.5} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 50K steps.
For real-world experiments, we use the same batch size and train the model for 30K steps.

\textbf{RDT-1B.}
We fine-tune RDT-1B from the \texttt{rdt-1b} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 256 for 200K steps.

\textbf{DM0.}
We fine-tune DM0 from the \texttt{DM0-base} foundation checkpoint.
For simulation experiments, the model is trained with a batch size of 32 for 100K steps.
For real-world experiments, we use the same batch size and train the model for 50K steps.
