\documentclass{mindlab}

\usepackage[toc,header]{appendix}
\usepackage{calc}
\usepackage{float}
\usepackage{pifont}
\usepackage{makecell}
\usepackage{enumitem}
\usepackage{mathrsfs}
\usepackage{adjustbox}
\usepackage{multicol}
\usepackage{changepage}
\usepackage{wrapfig}
\usepackage{amssymb}
\usepackage{array}
\usepackage{colortbl}
\usepackage{tikz}
\usepackage{svg}

\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\newcommand{\apphead}{\rowcolor{mindlabbluepale!45}}
\newcommand{\appgroup}[2]{\addlinespace[2pt]\rowcolor{mindlabbluepale!45}\multicolumn{#1}{@{}l}{\sffamily\bfseries\color{mindlabfg}#2}\\}
\newcommand{\appkey}[1]{{\sffamily\bfseries\color{mindlabblue}#1}}

\setlength{\cftbeforesubsecskip}{1.5pt}
\setlength{\headheight}{27pt}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\fittowidth}[1]{%
  \sbox0{#1}%
  \ifdim\wd0>\textwidth
    \resizebox{\textwidth}{!}{\usebox0}%
  \else
    \usebox0%
  \fi
}

\title{MinT: Managed Infrastructure for Training\\ and Serving Millions of LLMs}
\author{Mind Lab}

\correspondence{\email{contact@mindlab.ltd}}
\date{May 2026}

\abstract{
We present MindLab Toolkit (MinT), a managed infrastructure system for Low-Rank Adaptation (LoRA) post-training and online serving. MinT targets a setting where many trained policies are produced over a small number of expensive base-model deployments. Instead of materializing each policy as a merged full checkpoint, MinT keeps the base model resident and moves exported LoRA adapter revisions through rollout, update, export, evaluation, serving, and rollback. This adapter-revision path hides the complexity of distributed training, serving, scheduling, and data movement behind a service interface, making large-scale LoRA RL easier to run and reproduce.

MinT scales this adapter-revision path along three axes. \textbf{Scale Up} extends LoRA RL to frontier-scale dense and Mixture-of-Experts (MoE) architectures, including MLA and DSA attention paths via LoRA target mapping and rollout correction, while model-parallel training and serving paths are validated beyond 1T total parameters. \textbf{Scale Down} minimizes the training-serving handoff by moving only the exported LoRA adapter, which can be less than 1\% of the base-model size in compact rank-1 settings, eliminating full-checkpoint materialization entirely. In our measurements, adapter-only handoff reduces the measured handoff step by $18.3\times$ on a 4B dense model and by $2.85\times$ on a 30B MoE model; under the same resident-base allocation, concurrent multi-policy Group Relative Policy Optimization (GRPO) shortens wall time by $1.77\times$ and $1.45\times$, respectively, without increasing peak memory. \textbf{Scale Out} expands the policy namespace while keeping engine-local execution bounded. MinT separates durable policy addressability from CPU/GPU hot working sets: a tensor-parallel serving deployment supports $10^6$-scale addressable policy catalogs, with measured single-engine sweeps through 100K entries and thousand-adapter active waves at cluster scale. To address the main bottleneck that first-touch adapters serialize through the engine load path, MinT treats cold loading as scheduled service work and uses packed MoE LoRA tensors to improve live engine loading by $8.5$--$8.7\times$.Together, MinT makes multi-tenant training services a practical reality. Experimental validation demonstrates the infrastructure's ability to manage million-scale LoRA policy catalogs while training and serving selected adapter revisions over shared 1T-class base models.
}

% \abstract{Post-training has become increasingly important for advancing large-scale LLM performance, where tightly coupling training, serving, scheduling, and data transfer introduces significant infrastructure complexity. Low-Rank Adaptation (LoRA), originally a memory-efficient adaptation method, has recently shown surprising potential to address this complexity at scale.

% We present MindLab Toolkit (MinT), a Tinker-compatible managed system for LoRA-based reinforcement learning (RL). MinT treats a LoRA adapter as the operational unit of a policy: the base model remains resident, while adapters carry trainable behavior across rollout, update, serving, export, and rollback. This design hides the complexity of distributed training and serving behind a simple service interface, making large-scale LoRA RL easier to run and reproduce.

% MinT is organized around three scaling axes. \textbf{Scale Up} enables LoRA RL on large dense and MoE architectures, including Qwen-family and DeepSeek-style models, with frontier training paths specifically exercised on GLM-5.1 and Kimi K2. \textbf{Scale Down} minimizes the training-serving handoff to transferring only the LoRA adapter, which can be as small as 1\% of the base-model size (rank-1), eliminating full-checkpoint materialization entirely. \textbf{Scale Out} separates a $10^6$-entry addressable policy catalog from bounded host-resident and GPU-active working sets, with measured packed cache-miss live-load under 0.2 seconds.

% Together, MinT makes multi-tenant training services a practical reality. Experimental validation demonstrates the infrastructure's ability to manage million-scale LoRA policy catalogs while training and serving selected adapter revisions over shared trillion-parameter base models.}

\begin{document}
\maketitle

\input{sections/1_introduction}
\input{sections/2_system_overview}
\input{sections/3_architecture}
\input{sections/4_scaling}
\input{sections/5_capabilities}
\input{sections/6_related_work}
\input{sections/7_conclusion}

\bibliographystyle{plainnat}
\bibliography{paper}

\newpage
\begin{appendices}
\input{sections/appendix}
\end{appendices}

\end{document}
