\input{figures/figure_appendix_quality}
\input{figures/figure_appendix_valscore}

\section{Additional Evaluation Results}

\subsection{Per-Step Rollout Generation Time}
\label{app:gen_time_over_time}

\Cref{fig:app_gen_time_breakdown} shows rollout-generation time over training steps across models and SD baselines for RL rollouts.
\methodtitle{} achieves the lowest generation time throughout training on Qwen2.5-7B and Llama3.1-8B, and for most Qwen2.5-14B steps.
The small late-stage slowdown on Qwen2.5-14B coincides with longer sampled responses under \methodtitle{}, whose maximum response length reaches about 4,108 tokens, compared with 3,383 for \learned{} and 2,916 for \history{}.
We view this fluctuation as response-length stochasticity rather than a systematic toggle-policy failure.

\Alwayssd{} can improve over \nosd{} on Qwen2.5-7B and Llama3.1-8B, but can become slower when SD is enabled throughout rollout, especially on Qwen2.5-14B where the larger quantized drafter increases drafting cost.
\Learned{} achieves reasonable rollout speedup on Qwen2.5-14B because its auxiliary drafter combines low proposal cost with moderate block efficiency, as shown in~\cref{tab:eagle3_timing_decomp}.
These per-step trends are consistent with the aggregate timing results in~\cref{tab:e2e_speedup}.


\subsection{Training Dynamics and Quality Preservation}
\label{app:quality-preservation}

\Cref{fig:app_fidelity_reward,fig:app_fidelity_valscore} report training reward and validation accuracy trajectories over RL training.
Across Qwen2.5-7B, Qwen2.5-14B, and Llama3.1-8B-Instruct, \methodtitle{} closely follows the \nosd{} baseline in both metrics.
These results support that the rollout speedups do not materially alter training dynamics or degrade downstream model quality in our evaluated settings.

\input{figures/figure_appendix_adaptive_gamma}

\subsection{Adaptive Draft-Length Schedules Across Models}
\label{app:adaptive_gamma_schedule}

\Cref{fig:app_gamma_elevation} shows the $\gamma$ schedule selected by the adaptive controller across models.
Qwen2.5-7B increases from $5 \rightarrow 7 \rightarrow 9 \rightarrow 11$ as its \MAL{} repeatedly approaches the effective ceiling of the current draft length.
Before each elevation, its \MAL{} reaches $5.73$, $7.60$, and $9.48$, and after increasing $\gamma$ the \MAL{} further rises to $7.37$, $9.35$, and $11.25$, respectively.
Qwen2.5-14B follows the same pattern but later in training, increasing from $5 \rightarrow 7 \rightarrow 9$ when its \MAL{} reaches $5.73$ and $7.61$.
In contrast, Llama3.1-8B-Instruct stays at $\gamma=5$: although its \MAL{} improves from $5.07$ to around $5.5$, it never crosses the elevation threshold.
No downward adjustment is triggered in these runs.

\input{tables/rollout_history_breakdown}

\subsection{Why History-Based Drafting Does Not Accelerate}
\label{app:why-history-slow}

\Cref{tab:history_based_slow} analyzes the rollout-history-based baseline after rollout history becomes available, \textit{i.e.}, from the second epoch onward.
This isolates the active window where history-based drafting can actually be applied.
Unlike fixed-$\gamma$ chain drafting, the \history{} method reuses variable-length prefixes from previous rollouts, so block efficiency $\mal$ and effective acceptance rate $\alpha$ are not directly comparable.
We therefore analyze prefix reuse, verification overhead, and rollout-generation time in this window.
Even in this favorable window, the method does not reduce rollout-generation time.

The main reason is that history reuse does not provide enough accepted tokens to amortize verification cost.
For Qwen2.5-7B, the Spec-RL implementation reuses only 30 tokens on average out of a 660-token response, corresponding to a 4.5\% reuse rate.
For Llama3.1-8B-Instruct, the reused prefix is longer, but verification is also more expensive, adding 15.9\,s of overhead per active step.
As a result, even after excluding the first epoch and using the original lossy lenience factor to increase acceptance, rollout-history-based drafting increases rollout generation time by 19.4\% on Qwen2.5-7B and 8.8\% on Llama3.1-8B-Instruct.
These results show that the history-based approach is limited not only by its cold start, but also by low effective reuse relative to verification cost.

\input{figures/figure_appendix_eagle3_mal}
\input{tables/eagle3_analysis}

\subsection{Why Learned Auxiliary Drafting Remains Challenging}
\label{app:why-learned-auxiliary-challenging}

\paragraph{Block efficiency of \learned{}.}

EAGLE3-style auxiliary drafters typically require pretraining and online adaptation to match the target model and rollout data distribution~\citep{zhang2025fastgrpo, hu2026taming, iso2026accelerating}.
In our setting, a randomly initialized drafter yields \MAL{} close to 1.0, meaning that it rarely predicts even the first token correctly.
As shown in \cref{fig:app_eagle3_mal}, starting from a generally pretrained drafter improves early-stage behavior, but the initial \MAL{} remains limited: with a ShareGPT-pretrained drafter evaluated on SimpleRL-math rollouts, \MAL{} starts around 1.4--2.0.
A drafter pretrained directly on the target RL post-training distribution would likely provide higher initial \MAL{}~\citep{iso2026accelerating}; however, such in-distribution drafter pretraining is not always available in practice, so we use a drafter pretrained on general chat/text data~\citep{li2025eagle}.

\paragraph{Limited acceleration in Qwen-7B/14B.}

For the Qwen2.5-series models, learned auxiliary drafting provides limited generation-time acceleration mainly because its \MAL{} remains insufficient for our RL rollout workload.
Under high-temperature sampling ($T=1.0$) and long reasoning generation, the pretrained drafters do not provide sufficient proposal quality for our math RL rollout distribution.
In addition, system-aware activation remains important: as indicated by the gap between \alwayssd{} and \methodtitle{}, enabling SD during early large-batch phases can reduce or eliminate its benefit.
Higher \MAL{} may be achievable with in-distribution drafter initialization or more aggressive online adaptation, such as training the auxiliary drafter for multiple steps per RL iteration or running adaptation asynchronously.
However, these choices introduce additional drafter-training, scheduling, and configuration search burden; we further analyze the workload-matching requirement in~\cref{app:workload-matched-auxiliary-drafters}.

\paragraph{Unexpected slowdown on Llama3.1-8B.}

\citet{li2025eagle} report strong inference speedups for EAGLE3 in standard serving settings across Qwen and Llama models.
However, these results are not directly comparable to our RL rollout setting, which combines vLLM execution, high-temperature sampling ($T=1.0$), hard reasoning prompts, and long response budgets.
To diagnose the Llama3.1-8B-Instruct slowdown, we run short cold-start profiling experiments for the first five RL steps under the same rollout setting as in \cref{sec:eval-setting}, using vLLM EAGLE3 timing (\texttt{VLLM\_SD\_TIMING=1}), position-binned \MAL{} logging, and a custom proposal timer.

\Cref{tab:eagle3_timing_decomp} shows that the slowdown is not primarily due to drafter cost: the one-token drafter time $\DraftTime$ is similar across models, around 0.8--0.9\,ms.
Llama3.1-8B-Instruct has the largest verification-to-target ratio ($\VerifyTime/\TargetTime=1.45$), but this is only modestly higher than the Qwen models.
Thus, timing overhead alone does not explain the much larger slowdown.

The larger difference appears in depth-dependent block efficiency.
As shown in~\cref{tab:eagle3_depth_tau_step1}, before online adaptation, Llama3.1-8B-Instruct already shows block efficiency close to one after 512 generated tokens, indicating weak long-depth proposal quality from the pretrained drafter.
After the first five online-adaptation steps, this weakness persists: Llama3.1-8B-Instruct remains near 1.2--1.3 in the 1k--8k bins, whereas the Qwen models quickly recover block efficiency relative to their pre-adaptation values, especially at deeper positions (\cref{tab:eagle3_depth_tau_step1,tab:eagle3_depth_mal}).
This is precisely the regime that matters for rollout latency.
Long-tail responses dominate the rollout makespan, and the Llama drafter pays draft and verification overhead while most proposals are rejected, making the long tail even more expensive.
% The stronger depth robustness of the Qwen models is also consistent with prior reports of their instruction-following, mathematical reasoning, and self-reflection capabilities~\citep{zeng2025simplerl}.
This suggests a model- and workload-dependent failure mode of learned auxiliary drafting rather than a stable speedup across RL rollouts.

\input{figures/figure_appendix_eagle3_dapo_mal_combined}
\input{figures/figure_appendix_eagle3_dapo_gentime}


\subsection{Auxiliary Drafters Aligned with RL Rollout Distributions Are Difficult to Obtain}
\label{app:workload-matched-auxiliary-drafters}

\paragraph{NeMo-RL-native validation setup.}

To check whether the moderate \MAL{} of \learned{} stems from our veRL~\citep{sheng2025hybridflow}-based implementation or from the drafter itself, we additionally evaluate EAGLE3 drafters in NVIDIA's native NeMo RL SD stack~\citep{iso2026accelerating}.
This stack uses Megatron~\citep{shoeybi2019megatron}-based online drafter training with vLLM rollout serving, and we leave the core RL, SD, and reward paths unchanged.
We train on the DAPO-Math-17K dataset~\citep{yu2025dapo} using one node with 8 H100-SXM GPUs, $\DraftLength=3$ with chain drafting and verification, 30 RL steps, a maximum generation length of 8k, temperature $T=1.0$, and a global batch size of 1k rollouts.
For comparison, we also measure the \MAL{} of quantized self-drafters under the same DAPO-Math experimental setting in our self-SD stack.

\paragraph{Evaluated learned auxiliary drafters.}

For Qwen3-8B, we evaluate three public EAGLE3 drafters: RedHatAI's thinking drafter~\citep{redhatai2025qwen3eagle3thinking}, RedHatAI's non-thinking drafter~\citep{redhatai2025qwen3eagle3nothink}, and AngelSlim's drafter~\citep{angelslim2025qwen3eagle3}.
For Qwen2.5-7B and Qwen2.5-14B, we did not find compatible public EAGLE3 checkpoints, so we use our ShareGPT-pretrained drafters introduced in~\cref{app:learned_auxiliary_drafting}.
For Llama3.1-8B-Instruct, we evaluate both the RedHatAI public drafter~\citep{redhatai2025llama31eagle3} and the official checkpoint from the EAGLE3 authors released by yuhuili~\citep{yuhuili2024llama31eagle3}.
We also evaluate the public Qwen2.5-7B EAGLE-style drafter~\citep{mithanlab2026qwen25eagle} released with FastRL~\citep{hu2026taming} as an RL-targeted auxiliary-drafter checkpoint.
The checkpoint is reported to be trained on the OpenThoughts2-1M dataset~\citep{openthoughts2026openthoughts2}, a synthetic reasoning dataset spanning math, science, code, and puzzles.
Because this checkpoint is not directly compatible with NeMo RL's EAGLE3-specific online update path, we use it as a fixed drafter over the 30 RL steps without online adaptation.


\paragraph{Block efficiency gap.}

\Cref{fig:app_eagle3_dapo_mal} shows that evaluated learned auxiliary drafters achieve substantially lower block efficiency than target-induced quantized drafters.
Across the first 30 RL steps, quantized self-drafters stay near $\mal{}=3.6$--$3.9$ under $\DraftLength=3$, close to the ceiling of 4.
By contrast, EAGLE3 drafters remain mostly in the $\mal{}=1.2$--$2.4$ range, with the best public configuration, Qwen3-8B with the RedHatAI thinking drafter, reaching roughly $\mal{}=2.2$--$2.4$.
The FastRL Qwen2.5-7B EAGLE-style drafter reaches higher \MAL{} ($\mal{}=2.2$--$2.6$) than our ShareGPT-pretrained EAGLE3 drafter, but this is still insufficient for consistent rollout-generation speedup.
Within our 30-step window, online adaptation yields only modest \MAL{} improvement from the evaluated initializations.
These results suggest that the limitation is not the EAGLE/EAGLE3 architecture itself, but the availability of an auxiliary drafter sufficiently aligned with long, high-temperature RL rollout distributions and effective from the beginning of RL training.


\paragraph{Generation-time consequence.}

\Cref{fig:app_eagle3_dapo_gentime} shows the practical consequence of low block efficiency within the NeMo RL stack.
All timing curves compare \learned{} against the corresponding NeMo RL No-SD baseline.
Among the evaluated EAGLE3 drafters, Qwen3-8B with the RedHatAI thinking drafter is the only configuration that consistently reduces rollout-generation time, improving it by about 7.7\%.
The Qwen2.5-7B EAGLE-style FastRL drafter becomes consistently beneficial only late in training, reducing rollout-generation time by 17.6\% over steps 22--30.
The other configurations are slower than No-SD for much of training because their \MAL{} is too low to amortize draft, verification, and online-update overhead.
This confirms that learned auxiliary drafting can reduce rollout-generation time when the drafter reaches sufficiently high \MAL{}.


\paragraph{Output length and training distribution.}

One possible explanation is an output-sequence-length mismatch between drafter training and rollout inference.
However, when we run ablation probes with shorter output caps (e.g., 1k, 2k, and 4k), most public checkpoints or drafters pretrained on general chat/text data remain weak even under shorter caps, and several configurations are roughly length-flat.
This suggests that output length alone does not explain their low block efficiency.
A more central factor appears to be the drafter-training distribution: public drafters and our Qwen2.5 drafters are trained on general chat/text corpora such as ShareGPT or UltraChat, rather than target-generated rollouts from the RL workload.
\citet{hu2026taming} train their drafter on OpenThoughts2-1M~\citep{openthoughts2026openthoughts2}, a synthetic reasoning dataset closer to long reasoning workloads than general chat corpora.
More generally, reaching the high-\MAL{} regime reported by~\citet{iso2026accelerating} may require a more specialized auxiliary-drafter training pipeline, such as collecting target-generated long rollouts from the RL workload and then training the auxiliary drafter on the resulting data.


\paragraph{Takeaway.}

Our results show that reaching this regime depends on a well-aligned auxiliary drafter, which is not generally available from public checkpoints or generic drafter-training recipes.
Obtaining such a drafter can require a separate pipeline for target-generated long-rollout collection, offline auxiliary-drafter training, and possibly more aggressive online adaptation.
This raises the barrier to using \learned{} as a drop-in rollout accelerator.
Target-induced self-drafting provides a simpler point in this trade-off: it requires no external drafter checkpoint or offline auxiliary-drafter pretraining, yet achieves high \MAL{} from the beginning of RL training and translates it into practical rollout speedup.