\documentclass[11pt,a4paper]{article}

% Required base packages
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{float}
\usepackage{setspace}
\usepackage{microtype}

% hyperref - MUST load last with full config for clickable links
\usepackage[
    colorlinks=true,
    linkcolor=blue,
    citecolor=darkgray,
    urlcolor=blue,
    bookmarks=true,
    bookmarksnumbered=true,
    unicode=true
]{hyperref}

% Geometry config
\geometry{a4paper, top=2.5cm, bottom=2.5cm, left=3cm, right=2.5cm}

% Line spacing
\setstretch{1.15}

\begin{document}

\title{SparkVLA-M1: Cross-Embodiment Generalization via Unified Data Foundation and Action Manifold Learning}

\author{}
\date{}

\maketitle

\begin{abstract}
We introduce SparkVLA-M1, a co-optimized framework for cross-embodiment generalization in vision-language-action (VLA) modeling. To address data fragmentation, representation heterogeneity, and paradigm misalignment in embodied intelligence, we construct the largest open-source mixed-action dataset comprising 6M+ trajectories from six sources, with unified delta action representation in end-effector frames and rotation vectors for orientation encoding. We propose the \textbf{Action Manifold Hypothesis}, postulating that valid actions cluster on low-dimensional manifolds constrained by physics and task objectives, rather than scattering in high-dimensional space. Accordingly, we design \textbf{JAT (Joint Action Transformer)}, enabling direct prediction of clean action sequences instead of fitting unstructured noise, thereby improving decoding efficiency and policy robustness. Additionally, we introduce pluggable spatial enhancement modules---VGGT for single-view 3D priors and LongCat for multi-view fusion---to strengthen geometric understanding. Comprehensive experiments demonstrate that these techniques yield additive gains, achieving state-of-the-art zero-shot performance on Libero, Libero-Plus, RoboCasa, and Simpler-WidowX, significantly outperforming pi0 and uniVLA. We open-source all code and pipelines to advance community-driven general-purpose robotics.
\end{abstract}

\section{Introduction}

The ultimate goal of robotics should not be to develop specialized skills tailored for specific hardware, but rather to construct a general-purpose embodied agent akin to the human brain, one that can operate on arbitrary body configurations and achieve true one-brain-multiple-bodies capability. Such an agent should perceive the environment, understand instructions, reason about action consequences, and continuously learn through interaction. Despite decades of technological evolution, this grand vision remains unrealized. Embodied intelligence has yet to witness its GPT moment.

Upon reflection, we identify three critical bottlenecks in this process. First, data scale has not yet crossed the critical threshold. Embodied intelligence requires data that extends beyond images and text to include precise action annotations, which incur high collection costs and long cycles. Current approaches are predominantly constrained by downstream task requirements, adopting data from unified robot types or specific platforms during pretraining. While this facilitates local deployment, it severely limits the formation of generalizable strategies. Second, data quality varies significantly with missing standards. Different datasets employ inconsistent action representations, chaotic coordinate systems, and varying control frequencies, making it difficult for models to extract cross-embodiment commonalities from such fragmented inputs. The existing data volume and consistency are insufficient to support the emergence of intelligence. Third, there exists a fundamental mismatch in pretraining paradigms. Current VLA models are predominantly initialized from vision-language models, whose visual encoders focus on semantic recognition and lack modeling capabilities for three-dimensional spatial structure and physical laws. However, embodied decision-making relies on fine-grained spatial understanding and causal reasoning, capabilities that are difficult to acquire solely through 2D visual supervision and cannot be fully compensated through chain-of-thought mechanisms.

The path toward general embodied intelligence demands systematic optimization from data foundations and representation unification to training paradigms. We focus on a practical yet profound question: in the context of high self-collection costs for robot data, strong hardware dependencies, and the difficulty for individual teams to accumulate sufficient scale and diversity of interaction data, can we integrate global open-source robotics data from fragmented data silos into a unified data substrate and build general-purpose VLA models independent of specific hardware? This points to a new paradigm where embodied intelligence no longer emerges from closed-source systems but grows from the systematic reconstruction and capability accumulation of public knowledge.

To address this challenge, we propose SparkVLA-M1, a co-optimized technical pathway spanning from data to architecture. At the data level, we systematically integrate six open-source data sources including OXE, OXE-AugE, Agibot-Beta, and RoboCoin, constructing the largest and most comprehensive mixed training set within the non-proprietary ecosystem, comprising over 6 million trajectories. Faced with severe fragmentation across action formats, coordinate systems, and sampling frequencies in multi-source data, we establish a unified preprocessing pipeline: all actions are remapped to delta action representation in the end-effector coordinate frame, rotation components adopt rotation vectors to avoid singularities, and a pad-to-dual strategy enables unified modeling of single and dual-arm tasks. This standardized substrate breaks down data silos and provides a solid foundation for cross-configuration generalization.

At the model level, we propose two core innovations based on the VLM plus Action Expert architecture. First, current mainstream VLA action experts primarily predict noise through epsilon-prediction or v-prediction. These random noises are unstructured and meaningless, mixed with unreasonable actions such as jitter and discontinuities, scattered throughout the high-dimensional action space. Models require high information capacity to memorize these noises, leading to difficult and inefficient learning. Inspired by JIT, we propose the action manifold hypothesis: effective and smooth successful actions are not scattered throughout high-dimensional space but are constrained by physical laws, task objectives, and environmental conditions, clustering on a low-dimensional smooth manifold. Accordingly, we design the JAT mechanism, enabling the DiT backbone to directly predict clean action sequences themselves, transforming the prediction paradigm from fitting unstructured noise to projecting noisy actions back onto the low-dimensional action manifold, which significantly improves decoding efficiency and policy robustness. Second, to enhance spatial understanding capabilities, we introduce pluggable modules into the backbone: injecting scene-level structural priors through VGGT or fusing multi-view observation features via LongCat, achieving plug-and-play capability expansion.

We complete all ablation experiments under a unified training framework, and the results demonstrate that the three key techniques, data unification, architectural enhancement, and training paradigm innovation, are mutually orthogonal and yield additive gains. The final model undergoes comprehensive evaluation across multiple mainstream manipulation benchmarks including Libero, Libero-Plus, RoboCasa, and Simpler-WidowX, achieving average task success rates that significantly surpass advanced methods such as pi0, uniVLA, and openVLA-OFT. We thereby validate a complete technical pathway from data unification to architectural innovation to capability enhancement, demonstrating that high-performance, strongly generalizable embodied intelligence models can be built through systematic engineering even without private data. To this end, we will fully open-source the data processing pipeline and training code, dedicated to promoting a community-driven, shared, and progressive general-purpose robotics ecosystem.

\section{Data Foundation}

To support large-scale pretraining of general embodied intelligence, we complete two systematic efforts at the data level: establishing unified data representation to achieve cross-platform and cross-configuration operational semantic alignment, and conducting deep data cleaning and quality governance to address format inconsistencies, missing instructions, and semantic noise in multi-source data. Together, these constitute a reproducible, scalable, and high-purity embodied manipulation data foundation that provides a solid basis for subsequent high-performance VLA model training.

\subsection{Open-Source Dataset Analysis and Selection Strategy}

As embodied intelligence advances, an increasing number of high-quality open-source datasets have emerged. We systematically integrate the most representative VLA data resources currently available and conduct comprehensive evaluations of their characteristics to provide selection guidance for pretraining data construction.

Current open-source VLA datasets exhibit two development trends. One category prioritizes breadth, represented by OXE, which pursues data scale and embodiment coverage, making it suitable for pretraining foundation models. The other category prioritizes quality, such as Agibot, Galaxea, and RoboCOIN, which emphasize task structuring, annotation granularity, and physical realism, making them suitable for capability enhancement.

We believe that when building general-purpose VLA models, the principle of scale ensures the lower bound while quality raises the upper bound should be adhered to. Data scale determines the generalization boundary, data quality determines the performance ceiling, and unified formats such as LeRobot or RLDS are key to lowering access barriers and promoting community collaboration. Future high-quality datasets should ideally provide multimodal action representations including joint and end-effector configurations, dynamics information such as velocity and force or torque, high-definition multi-view synchronized video, fine-grained semantic annotations including task decomposition and keyframes and object relationships, and camera intrinsics and extrinsics with 2D and 3D geometric annotations. These metadata are not only sources of supervision signals but also core foundations for achieving visual-language-action alignment across temporal and spatial dimensions.

\subsection{Data Specification Unification}

We use the LeRobot format as the foundation to normalize and reconstruct all source data, ensuring complete consistency in action space, observation input, and training interface. This unified system comprises three key designs.

Action Representation: Delta Action in End-Effector Coordinate Frame with Rotation Vectors. We unify all task action outputs to delta action in the end-effector coordinate frame, where each step outputs the six-dimensional relative pose change and gripper state. This representation possesses inherent controller robustness and cross-platform compatibility. Furthermore, we compare multiple rotation representation methods including Euler angles, quaternions, and axis-angle, and through ablation experiments find that rotation vectors demonstrate optimal stability in action prediction, particularly in fine rotation tasks.

Unified Single and Dual-Arm Training Paradigm: Shared Policy Head via Zero Padding. To support single and dual-arm tasks sharing the same policy network, we adopt the pad-to-dual-arm strategy: in single-arm tasks, the unused other arm action dimensions are padded with zeros, and single-arm data is uniformly treated as right-arm behavior in dual-arm configurations. The model always outputs dual-arm action sequences, but only the corresponding valid arm channels are activated during execution, thereby achieving unified modeling and parameter sharing for single and dual-arm tasks. This approach enables the model to share most parameters while learning when to use only one arm and when dual-arm collaboration is required. Guided by task instructions, the model can automatically determine interaction modes during inference, achieving true unified modeling for single and dual-arm scenarios.

Data Distribution Statistics: Coverage of Diverse Configurations and Task Scenarios. The final unified dataset encompasses over 6 million trajectory fragments from multiple sources including OXE, RoboCasa, Agibot, and self-collected simulation data. The distribution covers various robot configurations including Widow, Panda, UR5, Franka dual-arm, Agibot custom dual-arm, and Stretch, with single-arm tasks comprising approximately 64 percent and dual-arm tasks approximately 36 percent, balanced to near 1:1 during training through sampling. The top task types include pick, place, open, close, grasp, move, push, insert, fold, and lift.

\subsection{Data Cleaning}

Although multiple open-source datasets provide rich manipulation trajectories, their original versions universally suffer from severe format inconsistencies, missing instructions, and semantic contamination. Direct usage would cause models to learn incorrect language-action mappings or invalid behavior strategies. We therefore conduct systematic rule-driven cleaning work. Through in-depth analysis of mainstream datasets, we identify the following typical quality issues.

OXE contains task prompts with non-English languages including French, Spanish, and Malagasy, as well as meaningless garbled text. Some episodes have empty task fields, and different subsets exhibit large frame rate variations with some as low as 5 FPS. OXE-AugE contains instructions in ASCII encoding format in episodes.jsonl, conflicting with dataloader parsing logic, and some subset tasks.jsonl files are completely empty despite having episode records. AgiBot provides scene descriptions rather than specific subtask instructions in tasks.jsonl, while actual action text exists in episodes.jsonl with frame range annotations but is not used by current dataloaders, causing models to degenerate into vision-only models. Galaxea subtask instructions mix in success judgment statements, empty instructions, and Chinese text stored in Unicode escape form, making true intent difficult to parse. RoboMind exhibits the same issues as AgiBot.

These problems severely affect the validity of language supervision signals and may even cause models to mistake failed attempts for correct strategies. We design a multi-stage rule-based cleaning pipeline focusing on the following sample categories. For invalid instructions, we delete episodes with empty or none task fields, filter garbled instructions containing obvious non-English characters such as Unicode encoding or special symbol strings, and attempt translation normalization for mixed-language instructions, removing those that cannot be repaired. For data mismatches, we fix the index inconsistency between episodes.jsonl and tasks.jsonl in OXE-Aug and supplement missing subtask instructions in AgiBot by extracting from action\_text and injecting according to frame ranges. For visual anomalies, we remove segments with all-black, blurry, or severely occluded images and exclude records with incorrect camera viewpoints such as wrist camera frames not aligned with operation areas. For action anomalies, we filter trajectories with excessively large consecutive action changes exceeding thresholds to prevent jitter interference and remove samples with severely mismatched action frequencies and video frame rates.

After the above cleaning, approximately 15 percent of low-quality trajectories are removed, retaining approximately 6 million high-confidence data points forming the CleanCore subset as the main pretraining source.

\section{Model Architecture}

To achieve end-to-end mapping from multimodal perception to robot action generation, we adopt a dual-component architecture of VLM plus Action Expert, and complete a two-stage training paradigm based on this: the first stage conducts unified pretraining on large-scale heterogeneous data to learn general operational strategies across tasks and configurations, establishing strongly generalizable action priors; the second stage introduces fine-grained spatial priors through supervised fine-tuning without destroying generalization capabilities, improving execution precision for complex tasks.

\subsection{VLM plus Action Expert Architecture}

Our model separates vision-language understanding and action generation into two independent yet closely cooperating modules. The VLM serves as the perception engine, responsible for fusing images and natural language instructions to output semantically rich multimodal representations and leveraging VLM capabilities to align actions into a unified space. The Action Expert serves as the action decision-maker, focusing on transforming the above representations into concrete robot control instructions.

We select Qwen3-VL as the base vision-language model, which possesses powerful image-text alignment capabilities, long-sequence modeling advantages, and good Chinese support, making it one of the few open-source VLMs that simultaneously satisfies high performance and broad applicability. Inputs include multi-view image sequences such as main view, wrist camera, and top view, input in frame-stacking form, and natural language instructions encoded through the tokenizer and concatenated with image tokens. Qwen3-VL outputs a set of spatially aligned visual features and text token representations through understanding of observations and instructions. These high-level semantic features are passed to the Action Expert as contextual basis for action generation.

The Action Expert is a lightweight DiT specifically designed to predict robot action sequences from multimodal features provided by the VLM. Unlike previous diffusion and flow matching action experts that train models to predict unstructured and meaningless noise through epsilon-prediction or noisy velocities through v-prediction, we advocate based on the action manifold hypothesis and JIT that DiT action experts should directly predict actions themselves. This addresses the problem of insufficient network capacity in high-dimensional space and provides better loss weighting directly corresponding to the generation process. Specifically, given a ground-truth action chunk and a flow matching timestep, sampled noise, the noised action chunk is computed as the linear interpolation between the ground-truth action and noise. The model directly predicts the action, which is then converted to velocity, and the flow matching loss is minimized.

\subsection{Two-Stage Paradigm}

Stage One: The core objective of this stage is to leverage massive heterogeneous data to train the model to master general operational patterns across tasks and embodiments, establishing strongly generalizable action priors. We train on the CleanCore dataset comprising approximately 6.2 million trajectories, covering single-arm, dual-arm, and multiple robot arm configurations, encompassing diverse task scenarios from basic grasping to complex interaction. All actions are represented as delta action in the end-effector coordinate frame, specifically six-dimensional relative pose change and gripper state, with rotation components encoded using rotation vectors to avoid Euler angle singularities and quaternion normalization optimization difficulties. Model inputs are multi-view image sequences and natural language instructions, with outputs being action sequences. The overall learning objective maps perceptual information to executable action space.

In architectural design, we achieve unified modeling for single and dual-arm tasks. By introducing the pad-to-dual-arm strategy, unused other arm action dimensions are padded with zeros in single-arm tasks, and single-arm operations are uniformly treated as right-arm behavior. The model always outputs complete dual-arm action sequences, but only corresponding valid channels are activated during actual execution. This strategy enables single and dual-arm tasks to share the same Action Expert parameters, significantly improving parameter utilization efficiency while enabling the model to autonomously determine whether dual-arm collaboration is required based on instructions, achieving flexible interaction mode switching. Furthermore, to address task and configuration distribution imbalances in the data, we adopt dual weighted sampling strategies for task categories and robot arm configurations: on one hand, sampling weights for rare tasks are increased according to inverse frequency, and on the other hand, balanced sampling is performed across different configurations to prevent the model from overfitting to specific hardware or high-frequency actions. Experimental results demonstrate that this strategy effectively alleviates long-tail bias, significantly improving zero-shot transfer performance on complex benchmarks such as RoboCasa and Libero-Plus, laying a solid foundation for subsequent fine-grained capability injection.

Stage Two: Although the pretrained model already possesses good generalization capabilities and performs well on most common tasks, it still exhibits action error accumulation and spatial alignment instability in high-precision operation scenarios such as fine insertion, cloth folding, or dual-arm collaborative operations. These tasks depend not only on language instruction understanding but also require precise modeling of three-dimensional spatial relationships, such as relative poses between holes and shafts, cloth fold directions, or coordinated timing between hands during operations. We therefore enter the second stage: spatial capability injection supervised fine-tuning, aiming to inject key spatial prior knowledge through selective fine-grained training without destroying the existing general strategy foundation.

In specific implementation, we update parameters with a lower learning rate, jointly fine-tuning the VLM and Action Expert modules. Dropout and action noise perturbation are introduced to enhance model robustness in real environments. Experimental results demonstrate that after this stage of training, model success rates significantly improve on challenging tasks such as insertion, folding, and dual-arm door opening, while performance on original simple tasks is not negatively affected, validating the effectiveness of capability stacking.

\section{Unified Specification Pretraining}

On the path toward general embodied intelligence, a core challenge is how to extract cross-embodiment and cross-task commonalities from massive heterogeneous manipulation data. Multi-source embodied data exhibits high heterogeneity in trajectory scale, task granularity, and embodiment diversity. If naive trajectory-level equal-probability sampling is adopted, the training distribution easily falls into volume-dominated pathological bias: on one hand, long-tail but critical high-value domains such as dual-arm or multi-configurations suffer from insufficient exposure and difficult convergence; on the other hand, limited training budgets are largely consumed by redundant sampling of high-frequency skills, suppressing model coverage efficiency for long-tail skills and consequently weakening cross-task and cross-configuration generalization capabilities. We therefore systematically investigate sampling ratio strategies in dual-arm learning, aiming to retain single-arm training priors while balancing embodiment coverage and skill learning efficiency through multi-granularity uniform sampling, seeking the optimal sampling balance point to achieve a data-driven generalization pathway.

\subsection{Data Ratio Strategy and Analysis}

Under the premise of fixing single-arm data ratios to OpenVLA default hyperparameters, we only compare sampling ratio strategies within dual-arm data to attribute performance differences as much as possible to variations in dual-arm ratios and sampling granularity. Based on the hierarchical data structure of embodiment, task, trajectory, and skill, we design three uniform sampling strategies: Trajectory-Uniform where each trajectory is sampled with equal probability to maintain original volume distribution, Task-Uniform where each task is sampled with equal probability to emphasize task diversity, and Embodiment-Uniform where each robot configuration is sampled with equal probability to emphasize embodiment coverage, without additional weighting within single trajectories.

To fairly evaluate different strategies on multi-source data, we construct a hierarchical validation set: randomly sampling 1000 trajectories from OXE for validation, stratified by task in AgiBotWorld with one random trajectory per task totaling 183, and stratified by embodiment in RoboCoin with 30 random trajectories per configuration totaling 240.

Faced with diverse data sources from different robot arm configurations, single and dual-arm settings, task types, and collection platforms, simple full mixing plus random sampling easily causes model bias toward high-frequency devices and common actions. We therefore design a refined data sampling strategy aiming to achieve balanced learning across embodiments and tasks.

We compare four typical data scheduling schemes. The first is single and dual-arm 1:1 balanced sampling, which enforces 50 percent single-arm and 50 percent dual-arm data in each training batch to prevent single-arm data from overwhelming dual-arm learning signals due to larger volume. The second sub-strategy is Task-Random, which still samples randomly by trajectory slice counts within single and dual-arm categories, preserving original task frequency characteristics. The third sub-strategy is Task-Diverse, which weights task categories to make sampling probabilities for each semantic verb such as pick, open, and fold tend toward uniformity, actively increasing long-tail task exposure rates. The fourth is Robot-Config 1:1:1 configuration balanced sampling, which divides robot arms into multiple typical configuration groups such as Widow, Panda, UR5, and dual-arm systems, with each group contributing equal sample sizes per batch, while still sampling by slice counts within groups to balance diversity and efficiency.

Through reasonable ratio design, we successfully construct a large-scale training set that is highly balanced across three dimensions: configuration, interaction complexity, and task semantics, laying a solid foundation for subsequent model learning of general representations.

\subsection{Multi-Dimensional Evaluation Validation}

To comprehensively evaluate pretrained effects under different ratio strategies, we establish a two-level evaluation system: open-loop L1 evaluation for rapid pretraining configuration screening and downstream benchmarks for final performance confirmation.

L1 evaluation is a lightweight, zero-shot open-loop test set specifically designed for the pretraining stage. All L1 tests are completed under zero fine-tuning and fixed inference head settings, truly reflecting pretraining knowledge absorption levels. Beyond L1 evaluation, we transfer models to the mainstream benchmark Libero-Plus for post-SFT evaluation.

\section{Dual-Stream Feature Collaboration for Action Generation}

On the path toward general embodied intelligence, a core challenge is enabling models to simultaneously understand task intent and master environmental structure. Vision-language models excel at parsing instructions, recognizing objects, and understanding context, but their perception of three-dimensional space often remains at the semantic relative position level, knowing that a cup is to the left of a box without knowing by how much or whether it is reachable. Robot action execution, however, relies on millimeter-level spatial judgment and dynamic coordination. Therefore, relying solely on VLM is insufficient to support fine-grained manipulation.

We therefore propose a dual-stream feature collaboration architecture, treating semantic streams provided by VLM and 3D spatial streams from diverse foundation models as two complementary pillars for action generation. The two streams are not input separately but interact hierarchically and fuse dynamically within a unified representation space, jointly guiding the Action Expert in making reasonable decisions.

Our VLM model undergoes large-scale VLA data pretraining, acquiring generalization capabilities for embodied scenarios and the ability to directly predict discrete actions in text form. However, differences remain between VLM and Action Expert policy model design in terms of model structure, training methods, and output format. To bridge this gap and better inject pretrained VLM model information into the Action Expert, we investigate which types of VLM features are more important for action generation after pretraining.

In experiments, we use Qwen3VL-4B and 16-layer DiT for comparison experiments, focusing on two main questions. First, after pretraining, which layer or layers of features in the VLM are more effective for the policy network, and are high-level or low-level features more helpful? We conduct three experiments: using the last layer of VLM features, using an intermediate layer at 70 percent depth, and using all hidden layers. Second, is the Action Query functionality better than original VLM features? We test two schemes: VLM hidden layers plus action query for the first 16 layers, and VLM hidden layers plus action query for the last 16 layers.

Experimental conclusions indicate that directly using VLM features performs better than concatenating action queries. Deep VLM features are more effective than shallow features. Adding more VLM feature layers does not bring additional benefits. Our model, after large-scale VLA data pretraining, possesses generalization capabilities for embodied scenarios and can directly predict discrete actions in text form. This indicates that the pretrained VLM can already extract certain action space features, with intermediate layers extracting multimodal features for embodied scenarios and deeper layers approaching action space semantic features, no longer requiring additional action queries to help VLM adapt to action space. Experimental results prove that using the last deep layer of pretrained VLM features performs better than shallow features and better than using all layers, equivalent to one deepest layer learning the most effective action space features. Furthermore, action queries do not show effectiveness in pretrained VLA and may even perform worse, as the model no longer needs additional queries to assist VLM rapid migration to VLA tasks. These experiments indirectly prove the effectiveness of our pretraining and provide an effective solution for feature selection after VLM pretraining.

\subsection{3D Information Injection}

Vision-language-action tasks are essentially embodied manipulation tasks executed in three-dimensional physical space: models must combine visual observations and natural language instructions to control robot arms or robots to complete specified actions in real three-dimensional environments. However, current mainstream VLA methods generally rely solely on two-dimensional images as visual input. This design has a fundamental limitation: when three-dimensional scenes are projected onto two-dimensional image planes, depth information, that is, the front-back and near-far relationships between objects, is inevitably lost. This information compression severely constrains model capabilities for reconstructing environmental geometric structure and consequently weakens spatial reasoning and geometric understanding capabilities.

Specifically, without explicit depth cues, models struggle to accurately judge occlusion relationships, parse relative spatial positions between objects such as above, below, front, and back, or precisely estimate distances and poses of target objects. These capabilities are crucial for fine manipulation such as grasping, insertion, and obstacle avoidance. Therefore, in scenarios with complex spatial layouts, ambiguous visual cues such as dramatic lighting changes, missing textures, limited viewpoints, or significant occlusions, existing VLA model performance often degrades significantly. This problem is particularly prominent in more challenging benchmark tests.

Taking the LIBERO-Plus evaluation set as an example, this dataset introduces diverse camera viewpoints, dynamic lighting conditions, complex scene layouts, and different robot arm initial poses. These factors collectively increase the difficulty of recovering three-dimensional semantics and geometric information from single two-dimensional images. Experiments show that current mainstream methods generally experience significant performance degradation on LIBERO-Plus, fully exposing the shortcomings of existing VLA frameworks in three-dimensional spatial perception.

To address the above three-dimensional understanding bottleneck caused by two-dimensional visual input, we propose a diverse three-dimensional feature injection mechanism based on existing vision-language models, aiming to enhance model perception capabilities for environmental geometric structure and spatial relationships. Specifically, we design and integrate two complementary three-dimensional feature representations.

The first is feedforward single-image three-dimensional features. We adopt the VGGT network pretrained on large-scale three-dimensional datasets to extract three-dimensional perception-capable features from single-view RGB images. VGGT effectively infers scene depth distribution and three-dimensional structure even from single images through joint learning of image appearance and underlying geometric priors.

The second is implicit multi-view image features. To further enhance spatial understanding robustness, we introduce a multi-view generation mechanism based on LongCat-Image-Edit. This model can synthesize multiple new viewpoint images from sparse single-view inputs, thereby implicitly modeling three-dimensional scene structure. Such features significantly enhance model generalization capabilities for object occlusion and viewpoint changes.

At the implementation level, to avoid destroying existing knowledge in pretrained VLM, we fuse three-dimensional features with VLM outputs before the Action Expert module. We systematically evaluate three feature interaction strategies: feature concatenation where VLM features and three-dimensional features are directly concatenated and fed into the Action Expert, cross-attention where VLM features serve as Query and three-dimensional features serve as Key and Value for fusion through attention mechanisms, and Q-Former where a set of learnable queries are introduced, combining multi-layer VLM and three-dimensional features as Key and Value through multi-layer cross-attention fusion.

Ablation experiments validate that single-layer cross-attention achieves the optimal balance between performance and computational efficiency, becoming the default fusion scheme. Ultimately, we fuse VGGT-extracted single-image three-dimensional features with the last layer of VLM visual-language representations through cross-attention as conditional input for action prediction.

For LongCat-generated implicit multi-view features, we conduct small-scale fine-tuning of LongCat-Image-Edit on Bridge and LIBERO datasets, using 50 pairs of paired data each including original images, synthesized viewpoint images, and corresponding edit prompts. For training and inference efficiency considerations, we adopt a two-step sampling strategy. Experiments show that introducing two synthesized viewpoints rather than a single viewpoint can more effectively enhance model generalization capabilities. Therefore, the final configuration adopts two synthesized viewpoints, two-step inference, and fusion with backbone VLM features through cross-attention.

The designed three-dimensional modules are pluggable and can be flexibly enabled for single-image depth features, multi-view enhancement, or both combinations according to task requirements, providing good adaptability and scalability for practical deployment.

\section{Experiments}

We conduct comprehensive evaluations across multiple mainstream benchmarks to validate the effectiveness of our proposed methods.

\subsection{Main Results}

Libero Benchmark: Our method achieves 98.6 percent average success rate across Libero-Spatial, Libero-Object, Libero-Goal, and Libero-Long tasks, significantly outperforming the previous state-of-the-art method starVLA at 95.4 percent.

Libero-Plus Benchmark: This more challenging benchmark introduces diverse camera viewpoints, dynamic lighting, complex backgrounds, and noise. Our method achieves 77.5 percent average success rate across all sub-tasks, substantially surpassing pi0-Fast at 69.5 percent and Uniform-VLA-baseline at 45.7 percent.

Simpler-WidowX Benchmark: Our method achieves 58.3 percent average accuracy, outperforming RDT at 47.8 percent, Pi0 at 54.6 percent, and other baseline methods.

RoboCasa Benchmark: Our method demonstrates strong performance on this benchmark as well, validating the effectiveness of our pretraining and data mixing strategies.

\subsection{Ablation Studies}

Rotation Vector Validation: To emphasize unified specifications, we validate the impact of rotation vectors based on models without pretraining on downstream tasks. Results show that rotation vectors achieve 66.7 percent average accuracy compared to 53.1 percent for Euler angle subtraction, confirming that rotation vectors provide more stable rotation predictions.

JAT Mechanism Validation: All validations based on pretrained models demonstrate JAT benefits. Models using JAT consistently outperform those using traditional noise prediction across all Libero-Plus sub-tasks. The best JAT configuration achieves 72.4 percent total success rate compared to 67.2 percent for the baseline without JAT.

VLM Feature Interaction: Experimental results indicate that directly using pretrained VLM features performs better than concatenating action queries. Deep VLM features are more effective than shallow features, and adding more VLM feature layers does not bring additional benefits. Using the last layer of VLM features with 16-layer DiT achieves 71.0 percent total success rate, while using all hidden layers drops to 53.2 percent.

3D Feature Interaction Effects: To systematically validate the effectiveness of the proposed three-dimensional feature injection mechanism, we conduct ablation experiments on LIBERO and its more challenging extension LIBERO-Plus. All experiments are based on the same pretrained VLM, introducing components to be validated only through supervised fine-tuning while strictly excluding interference from other enhancement strategies such as JAT, ensuring performance changes can be clearly attributed to the proposed three-dimensional perception mechanism.

First, we evaluate the introduction effect of VGGT single-image three-dimensional features and compare three fusion methods: feature concatenation, cross-attention, and Q-Former. Results show that regardless of fusion strategy, introducing VGGT features consistently improves model performance on LIBERO and LIBERO-Plus, validating the universal gain effect of single-view three-dimensional priors for VLA tasks. Among them, single-layer cross-attention significantly outperforms other methods, becoming the standard configuration for subsequent experiments.

Furthermore, we validate the effectiveness of implicit multi-view features. After introducing additional viewpoints generated by LongCat-Image-Edit, model generalization capabilities significantly improve: overall success rate increases by 2.8 percentage points on LIBERO and by 2 to 4 percentage points on average on LIBERO-Plus. Critically, using two synthesized viewpoints rather than a single viewpoint can further enhance three-dimensional spatial perception capabilities. This improvement is most prominent in the camera viewpoint perturbation sub-item, with accuracy improving by up to 14 percentage points, fully demonstrating that multi-view information effectively alleviates model dependence on fixed observation viewpoints and significantly enhances robustness in real dynamic environments.

\section{Conclusion}

We present SparkVLA-M1, a comprehensive pathway from data unification to architectural innovation for cross-embodiment generalizable VLA models. Through systematic integration of open-source data, unified action representation, novel action manifold learning via JAT, and pluggable 3D spatial enhancement, we demonstrate that high-performance embodied intelligence can be achieved without proprietary data or specialized hardware. Our extensive experiments across multiple benchmarks validate the effectiveness of each component and their additive contributions. We hope this work contributes to the open robotics ecosystem and inspires further research toward general-purpose embodied agents.

\end{document}