\clearpage
%% ============================================================
\section{Data Workflows}
\label{app:workflows}
%% ============================================================

\subsection{Airline CSM Workflows}
This domain covers 50 scenarios across seven workflow categories — IRROPS rebooking, voluntary changes, missed connections, same-day standby, cancellations and refunds, escalation and availability constraints, and adversarial compensation claims — backed by 15 tools. The domain is high-stakes and time-pressured, with heavy dependence on accurate transcription of named entities: confirmation codes, flight numbers, passenger names, and travel dates. Table~\ref{tab:workflows-csm} provides a description of each workflow, the expected number of tool calls, and the tools invoked by the agent.

\input{tables/08_appendix/i_data_workflow/01_workflow_airline}

\subsection{Healthcare HRSD Workflows}

This domain covers HR service delivery at a hospital, with callers drawn from clinical and administrative staff. It comprises 83 scenarios across 12 single-intent workflows backed by 47 tools, extended with dual-intent, triple-intent, and adversarial variants. It has the highest per-workflow complexity of any domain in \framework, with an average of 8.7 expected tool calls across all scenarios (5.0 for single-intent workflows, up to 18 for triple-intent). Its defining challenge is the density and complexity of named entities the caller must communicate over voice — NPI numbers, DEA registration numbers, state license numbers, and OTP codes — where a single transcription error can cascade into authentication or policy failures. Table~\ref{tab:workflows-hrsd} provides a description of each workflow, the expected number of tool calls, and the tools invoked by the agent.

\input{tables/08_appendix/i_data_workflow/03_workflow_medical_hr}

\subsection{Enterprise ITSM Workflows}

This domain covers an enterprise IT service desk spanning 21 workflows across six categories, backed by 59 tools. It comprises 80 scenarios: 29 single-intent, 14 double-intent, 14 triple-intent, 14 quadruple-intent, and 9 adversarial. Its defining characteristic is a branching flow structure: incident flows have both a troubleshooting-resolved path and an escalation-to-ticket path, testing whether the agent correctly gates escalation on failed resolution attempts. Authentication is tiered across three levels — standard, OTP-elevated, and manager-level — reflecting the sensitivity of different workflows. Table~\ref{tab:workflows-itsm} provides a description of each workflow, the expected number of tool calls, and the tools invoked by the agent.

\input{tables/08_appendix/i_data_workflow/02_workflow_itsm}
