\section{Conclusion}

Post-training now couples rollout, update, evaluation, serving, scheduling, and data movement around many policy variants over a small number of expensive base-model deployments. Full-checkpoint workflows make every task branch, benchmark candidate, product version, tenant variant, rollback point, or personal policy look like another complete model deployment. This abstraction does not scale when the target workload is a large and continuously changing population of trained behaviors over shared frontier bases.

MinT makes exported LoRA adapter revisions the managed unit for this setting. Base-model deployments remain resident, while adapter revisions carry trained behavior through rollout, update, export, evaluation, serving, and rollback. The adapter revision is the behavior-carrying payload; the policy record is the service state that makes that payload reproducible, schedulable, and durable across worker changes, training checkpoints, rollout records, and serving-cache movement.

The measurements validate the same adapter-revision path along three scaling axes. \textit{Scale Up} supports LoRA RL on dense and MoE architectures with model-parallel training and serving paths exercised beyond 1T total parameters. \textit{Scale Down} removes full-checkpoint materialization from the training-serving handoff: adapter-only handoff reduces the measured handoff step by $18.3\times$ on a 4B dense model and by $2.85\times$ on a 30B MoE model, while concurrent multi-policy GRPO shortens wall time by $1.77\times$ and $1.45\times$ under the same resident-base allocation. \textit{Scale Out} separates durable policy addressability from CPU/GPU hot working sets: the serving path supports $10^6$-scale addressable policy catalogs, with measured single-engine sweeps through 100K entries and thousand-adapter active waves at cluster scale; first-touch cold loading becomes scheduled service work, and packed MoE LoRA tensors improve live engine loading by $8.5$--$8.7\times$.

Together, these results show that multi-tenant LoRA training services can be made practical without turning every trained policy into a full-model server. MinT lets policy count grow through adapter revisions, policy records, cache tiers, and controlled cold-load admission, while selected revisions train and serve over bounded resident working sets. This makes LoRA a service-level unit for large-scale post-training today and a practical path toward larger populations of organizational and personal policies over shared 1T-class base models.