\section{Curvilinear Demonstration Selection}
\label{sec:CDS}

Motivated by the curvature--performance correlation in Section~\ref{sec:curvature_corr}, we introduce \emph{Curvilinear Demonstration Selection} (CDS), a practical method for constructing an ordering of many-shot CoT demonstrations.
CDS aims to produce a smooth trajectory in embedding space, avoiding abrupt transitions between successive demonstrations.

\subsection{Experiment Settings}
We evaluate CDS on three reasoning tasks spanning diverse domains, including geometry, number theory, and DetectiveQA.
Our primary experiments use reasoning LLMs from the Qwen3 family (8B and 14B) across multiple demonstration budgets, with the motivation and experimental details provided in Appendix~\ref{sec:exp_setting}.


\paragraph{Objective.}
Given a set of $n$ demonstrations with projected embeddings $\{\mathbf{e}_i\}_{i=1}^n$, CDS seeks a permutation
$O=[\mathbf{d}_{\pi(1)},\ldots,\mathbf{d}_{\pi(n)}]$
that minimizes the total curvature
\begin{equation}
\Theta(O)=\sum_{t=2}^{n-1}\arccos\!\left(
\frac{\mathbf{v}_t \cdot \mathbf{v}_{t+1}}{\|\mathbf{v}_t\|\;\|\mathbf{v}_{t+1}\|}
\right)
\label{eq:CDS_curvature}
\end{equation}

\begin{equation}
    \mathbf{v}_t = \mathbf{e'}_{\pi(t)}-\mathbf{e'}_{\pi(t-1)}.
\end{equation}

\paragraph{TSP-based approximation.}
Directly optimizing Eq.~\eqref{eq:CDS_curvature} is combinatorial for large $n$: exact minimization requires evaluating $n!$ permutations, which is infeasible for the longest prompts studied in this paper ($n\le128$).
Moreover, optimizing angles alone can produce trajectories that are geometrically straight but make very large jumps across the embedding space.
We therefore use a practical TSP-based heuristic with a combined transition cost that balances spatial proximity and local curvature:
\vspace{-0.5mm}
\begin{equation*}
D_{\mathrm{CDS}} = D_{\mathrm{euclidean}} + D_{\mathrm{curvature}}
\end{equation*}
\vspace{-0.5mm}
The Euclidean component keeps adjacent demonstrations in related local neighborhoods, while the curvature component discourages sharp conceptual turns.
We build a complete graph under this combined cost and compute a short tour/path using a nearest-neighbor heuristic followed by 2-opt local search~\cite{tsp}.
We then linearize the resulting tour to obtain the final demonstration order.
Our theoretical claim only requires a sufficiently smooth pedagogical progression, not the global minimum of Eq.~\eqref{eq:CDS_curvature}; empirically, this approximation is effective and remains inexpensive, taking under one minute on a standard CPU for $n\le128$.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\resizebox{\linewidth}{!}{%
\begin{tabular}{lllcccc}
\toprule
Task & Model & Method & \multicolumn{4}{c}{Number of demonstrations} \\ 
\cmidrule(lr){4-7}
 & & & 16 & 32 & 64 & 124 \\ \midrule
number\_theory & gpt-5.2 & origin & 89.63 & 91.11 & 88.56 & 91.48 \\
 &  & CDS & 89.26 & \textbf{92.59} & \textbf{92.04} & \textbf{91.85} \\
 &  & CDS$_{bge}$ & 88.15 & \textbf{91.48} & \textbf{88.70} & 91.30 \\
 & Qwen3 & origin & 86.67 & 87.96 & 86.30 & 90.93 \\
 &  & CDS & 85.56 & 87.85 & \textbf{87.78} & 90.74 \\
 &  & CDS$_{bge}$ & 85.37 & 87.78 & \textbf{87.22} & \textbf{91.30} \\ \midrule
geometry & gpt-5.2 & origin & 75.99 & 74.74 & 75.37 & 75.78 \\
 &  & CDS & \textbf{81.21} & \textbf{78.08}
  & \textbf{80.79} & \textbf{75.99}\\
 &  & CDS$_{bge}$ & \textbf{80.37} & \textbf{78.29} & \textbf{76.83} & 75.11 \\
 & Qwen3 & origin & 66.18 & 65.76 & 65.14 & 73.07 \\
 &  & CDS & 65.55 & \textbf{68.27} & \textbf{68.89} & \textbf{73.90} \\
 &  & CDS$_{bge}$ & \textbf{66.60} & \textbf{67.85} & \textbf{70.36} & \textbf{74.32} \\ \midrule
DetectiveQA & gpt-5.2 & origin & 80.52 & 82.47 & 83.77 & 85.71 \\
 &  & CDS & 80.52 & \textbf{83.12} & \textbf{85.06} & \textbf{88.31} \\
 &  & CDS$_{bge}$ & \textbf{81.17} & \textbf{85.71} & \textbf{86.36} & \textbf{88.31} \\
 & Qwen3 & origin & 75.97 & 74.03 & 70.78 & 72.73 \\
 &  & CDS & \textbf{76.62} & \textbf{75.32} & \textbf{73.38} & \textbf{75.32} \\
 &  & CDS$_{bge}$ & 75.32 & \textbf{77.29} & \textbf{75.32} & \textbf{76.63} \\ \bottomrule
\end{tabular}%
}
\caption{CDS robustness across tasks, embedding models, and target LLMs. CDS uses the original embedding model, while  CDS$_{bge}$ replaces it with bge-m3.}
\label{tab:CDS_robustness}
\vspace{-5mm}
\end{table}













\paragraph{High-curvature control.}
To isolate curvature from local clustering, we compare CDS with a high-curvature ordering constructed from the same demonstrations.
Both use Euclidean proximity, but the high-curvature variant inverts the curvature objective:
$$D_{\mathrm{high\ curv}} = D_{\mathrm{euclidean}} + \left(\max(D_{\mathrm{curvature}})-D_{\mathrm{curvature}}\right)$$
Thus, it still groups semantically related demonstrations while forcing abrupt turns between consecutive examples.



\paragraph{Result.} We evaluate CDS on three reasoning tasks: geometry proof generation, number theory problem solving, and DetectiveQA logical reasoning.
We further test robustness by replacing the original embedding model with bge-m3 and by evaluating an additional closed-source model, gpt-5.2.
Table~\ref{tab:CDS_robustness} shows that the gains persist across embedding models and target LLMs, especially on geometry and DetectiveQA; number\_theory shows smaller margins because baseline accuracies are already high, leaving less room for ordering to improve performance.
The performance gains also depend on the curvature of the original ordering. Our ablation in Table~\ref{tab:high_curvature_ablation} against a high-curvature baseline strengthens our curvature claim.



\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\resizebox{\linewidth}{!}{%
\begin{tabular}{lllcccc}
\toprule
Task & Model & Method & \multicolumn{4}{c}{Number of demonstrations} \\
\cmidrule(lr){4-7}
 & & & 16 & 32 & 64 & 128 \\ \midrule
number\_theory & gpt-5.2 & CDS & \textbf{88.15} & \textbf{91.48} & \textbf{88.70} & \textbf{91.30} \\
 &  & high curv & 85.74 & 89.63 & 86.48 & 88.15 \\
 & Qwen3-14B & CDS & \textbf{85.37} & \textbf{87.78} & \textbf{87.22} & \textbf{91.30} \\
 &  & high curv & 79.26 & 84.44 & 84.26 & 90.37 \\ \midrule
geometry & gpt-5.2 & CDS & \textbf{80.37} & \textbf{78.29} & \textbf{76.83} & \textbf{75.11} \\
 &  & high curv & 72.65 & 73.90 & 76.33 & 74.53 \\
 & Qwen3-14B & CDS & \textbf{66.60} & \textbf{67.85} & \textbf{70.36} & \textbf{74.32} \\
 &  & high curv & 66.38 & 66.81 & 66.60 & 71.80 \\ \bottomrule
\end{tabular}%
}
\caption{Controlled smoothness ablation with bge-m3 embeddings. The same demonstrations are used for both orderings; only the transition curvature objective is inverted.}
\label{tab:high_curvature_ablation}
\vspace{-6mm}
\end{table}



