\section{Settings}
We establish an experimental framework for studying many-shot In-Context Learning (ICL), with and without Chain-of-Thought (CoT), under long-context constraints.
Our design spans three dimensions: \emph{task type} (non-reasoning vs.\ reasoning), \emph{model type} (standard instruction-tuned vs.\ explicitly ``reasoning'' models), and \emph{ICL configuration} (prompt format and number of demonstrations).



\subsection{Tasks Studied} 
Prior many-shot work has largely emphasized non-reasoning classification benchmarks~\cite{struggle, bertsch-etal-2025-context}.
We extend evaluation to include both classification-style tasks and multi-step reasoning tasks, while using a unified \emph{open-ended generation} evaluation for all datasets.


\paragraph{Evaluation protocol.}
For each test instance, the model generates a free-form text completion.
We map the completion to a predicted answer using task-specific extraction and normalization, and score it by \emph{exact match} against the reference. Prompt templates for evaluation are provided in Appendix~\ref{sc:apx_tasks}.
For numerical datasets (e.g., GSM8K/MATH), we extract the final numeric value or mathematical expression from the completion and compare it to the ground truth under the same exact-match criterion.


\paragraph{Non-reasoning tasks.}
These tasks require little intermediate reasoning and primarily test semantic understanding and label mapping.
We include benchmarks with different label-space sizes: SuperGLUE~\cite{superglue} (small label space), NLU~\cite{nlu}, TREC~\cite{trec}, and BANKING77~\cite{banking} (large label space).

\paragraph{Reasoning tasks.}
These tasks require deduction and/or mathematical derivation.
We focus on mathematical reasoning with GSM8K~\cite{cobbe2021gsm8k} and MATH~\cite{hendrycksmath2021}, and include DetectiveQA~\cite{xu2025detectiveqaevaluatinglongcontextreasoning} for narrative reasoning over long contexts.
For tasks that provide gold rationales, we use the dataset-provided reasoning chains $C_i$ as the CoT component in demonstrations (Section~\ref{sec:icl_config}).

\subsection{LLMs Studied}
\label{sec:models}
We evaluate a range of LLMs and group them by whether they explicitly contain extended reasoning at inference time.

\paragraph{Non-reasoning LLMs.}
These models are primarily tuned to produce direct answers given instructions, without an explicit ``thinking'' token.
We evaluate LLaMA 3.1 (\texttt{\small Llama-3.1-8B-Instruct}), LLaMA 3.3 (\texttt{\small Llama-3.3-70B-Instruct})~\cite{llama3}, Qwen 2.5 (7B) (\texttt{\small Qwen2.5-7B-Instruct}), and Qwen 2.5 (14B) (\texttt{\small Qwen2.5-14B-Instruct}).

\paragraph{Reasoning-oriented LLMs.}
These models expose an explicit reasoning segment (e.g., a \texttt{<think>} token)
We evaluate Qwen 3 (8B) (\texttt{\small Qwen3-8B}) and Qwen 3 (14B) (\texttt{\small Qwen3-14B})~\cite{yang2025qwen3technicalreport}, QwQ (32B) (\texttt{\small QwQ-32B})~\cite{qwen2025qwen25technicalreport}, and DeepSeek-R1 (685B) (\texttt{\small DeepSeek-R1})~\cite{deepseekai2025deepseekr1incentivizingreasoningcapability}.
For reasoning-oriented models, we enable the model's reasoning mode during inference to allow the generation of intermediate reasoning tokens.

\paragraph{Long-context configuration.}
To support many-shot prompts (up to 131K tokens for Qwen-family models), we apply the official RoPE scaling configurations provided by each model release.
All other decoding and system-prompt settings follow the model providers' recommended defaults unless stated otherwise.

\subsection{ICL Configuration}
\label{sec:icl_config}
We study scaling from few-shot to many-shot under two prompting paradigms.

\paragraph{Traditional ICL.}
Prompts consist of $k$ input--output pairs $(x_i, y_i)$ followed by a query $x'$.
The model produces an output $y'$ conditioned on the demonstration set:
\begin{equation}
y' = \mathrm{LLM}\!\left(x' \mid \{(x_i, y_i)\}_{i=1}^n\right).
\end{equation}

\paragraph{CoT-ICL.}
Prompts consist of $k$ triples $(x_i, C_i, y_i)$, where $C_i$ is a reasoning chain.
Given a query $x'$, the model generates both an intermediate chain $C'$ and a final answer:
\begin{equation}
(C', y') = \mathrm{LLM}\!\left(x' \mid \{(x_i, C_i, y_i)\}_{i=1}^n\right).
\end{equation}

\paragraph{Context scaling.}
CoT demonstrations are substantially longer than standard ICL examples (e.g., in our geometry setting, a single CoT demonstration can be $\sim$30$\times$ longer than a BANKING77 example).
As a result, while hundreds to thousands of demonstrations may fit for traditional ICL, CoT-ICL is typically limited to at most a few hundred demonstrations by context length.
We therefore focus our scaling analysis on $n \le 128$, which captures the most informative trade-offs between model type, task type, and demonstration count in our long-context regime.









