\section{Related Work}
\label{sec:related}

% \noindent\textbf{Game Benchmarks.} Game benchmarks evaluate models in environments that require reasoning, planning, and multimodal action. AgentBench covers diverse interactive tasks \citep{liu2023agentbench}, while GameBench, BALROG, and GameWorld use games to probe strategic reasoning and multimodal game play \citep{costarelli2024gamebench,paglieri2024balrog,ouyang2026gameworld}. Their breadth makes memory-specific failure hard to isolate: an error can arise from perception, rule understanding, exploration, planning, or action formatting. Our environments make hidden state explicit and controllable instead, so that Matching Pairs and 3D Maze both probe the same question: whether a model can reconstruct information no longer in the current observation and use it to act.

\noindent\textbf{Game Benchmarks.} Game benchmarks evaluate reasoning, planning, and multimodal action in interactive settings. AgentBench covers diverse agent tasks \citep{liu2023agentbench}, while GameBench, BALROG, and GameWorld use games to test strategic reasoning and multimodal game play \citep{costarelli2024gamebench,paglieri2024balrog,ouyang2026gameworld}. However, these broad benchmarks make memory-specific failures hard to isolate, as errors may come from perception, rule understanding, exploration, planning, or action formatting. In contrast, our environments make hidden state explicit and controllable, allowing Matching Pairs and 3D Maze to test whether models can recover information that is invisible and use it for later actions.

\noindent\textbf{Long-Context and Retrieval Benchmarks.} Long-context benchmarks study how well models use information distributed across extended inputs. LongBench, L-Eval, M4LE, and Ada-LEval provide broad long-context evaluation suites \citep{bai2023longbench,an2024leval,kwan2024m4le,wang2024adaleval}, while Lost in the Middle, RULER, NoLiMa, HELMET, and LongBench v2 probe position sensitivity, literal-match shortcuts, and realistic long-context reasoning \citep{liu2024lostmiddle,hsieh2024ruler,modarressi2025nolima,yen2025helmet,bai2025longbenchv2}. Retrieval and reading benchmarks such as MS MARCO, TriviaQA, KILT, and BEIR test whether models can locate relevant evidence from available sources \citep{nguyen2016msmarco,joshi2017triviaqa,kwiatkowski2019naturalquestions,yang2018hotpotqa,petroni2021kilt,thakur2021beir}, and MMNeedle extends this to long visual contexts \citep{wang2024mmneedle}. These show that placing evidence in a context window does not guarantee robust use of it. Our setting differs because long context is produced by interaction rather than given as a fixed input, and key evidence may be seen only once before it becomes useful later.