\section{Related Work} \label{sec:related}

\textbf{Visual Detection and Grounding in VLMs}. Visual grounding/detection tasks traditionally rely on task-specific heads~\citep{carion2020end, liu2024grounding, ren2016faster, jiang2024far3d}, but recent VLMs like Qwen-VL series~\citep{bai2025qwen2.5vl, bai2025qwen3vltechnicalreport}, InternVL~\citep{chen2024internvl} and Shikra~\citep{chen2023shikra} formulate it as an autoregressive token generation problem. This generative paradigm, however, often suffers from structural hallucinations and high latency~\citep{li2023evaluating}. To mitigate these issues, Rex-Omni~\citep{jiang2025rexomni} employs point-based prediction, while Patch-as-Decodable-Token (PaDT)~\citep{su2026patch} and Groma~\citep{ma2024groma} utilize visual reference tokens to point directly to image patches. Complementary innovations such as Pink~\citep{xuan2024pink}, ViP-LLaVA~\citep{cai2024vipllava}, Griffon~\citep{zhan2024griffon},  DnU~\citep{lin2024draw} and PAM~\citep{lin2025perceive} focus on enhancing 2D referential comprehension through visual prompt engineering and multi-granularity feature scaling.
LLMDet~\citep{fu2025llmdet} boosts detection recall by data distribution tuning. To bypass serial decoding bottlenecks, WeDetect~\citep{fu2025wedetect} treats detection as a parallel retrieval task. Advanced perception logic is further integrated via Chain-of-Thought (CoT)~\citep{qi2025cot4det}, while post-training strategies such as Vision-R1~\citep{zhan2025visionr1}, UniVG-R1~\citep{bai2025univg} and GW-VLM~\citep{jiang2026gwvlm} utilize reinforcement learning to align model outputs with visual feedback and reduce grounding errors~\citep{zhang2024ferretv2}.

\noindent\textbf{Parallel Decoding via MTP and Diffusion LLMs}.
To mitigate autoregressive latency, parallel generation techniques such as MTP~\citep{gloeckle2024better, cai2024medusa, samragh2025your} predict multiple future tokens simultaneously, often coupled with speculative decoding to accelerate inference. Recent extensions such as Future Summary Prediction~\citep{mahajan2025beyond} capture long-term dependencies via auxiliary heads. Concurrently, Diffusion Language Models (DLMs) such as LLaDA~\citep{nie2025large}, Dream~\citep{ye2025dream}, and DiffuCoder~\citep{gong2025diffucoder} frame sequence generation as a discrete denoising process, enabling bidirectional context modeling and non-autoregressive decoding. Hybrid semi-autoregressive paradigms, including Block Diffusion~\citep{arriola2025block}, SDLM~\citep{liu2025sequential} and Fast-dLLM v2~\citep{wu2025fast}, decode fixed-size token blocks in parallel while maintaining causal dependencies to preserve KV-caching compatibility. More advanced frameworks~\citep{wang2025diffusion, lu2025adablock} unlock inter-block parallelism and adaptive block scheduling. These paradigms have been extended to the multimodal domain via DiffusionVL~\citep{li2025diffusionvl}, translating autoregressive LMMs into high-performance diffusion-based models.


\textbf{LocateAnything} differs from existing works in two key aspects. First, instead of generating bounding boxes via slow NTP, we output the complete box in a single parallel step.
Second, recent MTP paradigms group tokens into arbitrary chunks.
Instead, our PBD treats the entire coordinate set as a single atomic block, resolving both the fragmentation of NTP and the arbitrary chunking of MTP, seamlessly unifying high throughput with structural coherence.
