\section{Conclusion}\label{sec:conclusion}
We presented LocateAnything, a unified framework that reformulates visual grounding and detection in VLMs via \emph{Parallel Box Decoding}. By elevating geometric elements to atomic units rather than 1D streams, LocateAnything aligned the training supervision with the inherently coupled nature of spatial coordinates. With massive 138M text-image training queries and a flexible on-demand inference mechanism, LocateAnything not only delivered SOTA accuracy across diverse tasks, but also achieved up to a $2.5\times$ speedup over competitive methods. Our method provided a practical and scalable route for real-time visual perception, opening the door to deploying general-purpose VLMs in latency-sensitive embodied robotics and interactive agents.

\noindent\textbf{Limitation.} Currently, our model is primarily trained with supervised fine-tuning. Reinforcement learning is an important next step to further optimize the block-level decoding policy, reduce fallback frequency, and encourage effective exploration in hard dense/long-tail cases, which could improve both robustness and worst-case decoding speed. We leave it for future work.

\noindent\textbf{Acknowledgement.} The authors would like to thank the valuable discussions and input from Qing Jiang, Amala Sanjay Deshmukh, Karan Sapra, Mingjie Liu, Yi Dong, Pavlo Molchanov, Yonggan Fu, Collin McCarthy, Mike Ranzinger, Greg Heinrich, Wonmin Byeon, Yexuan Li, Chi-Pin Huang, Fu-En Yang, Frank Wang, Jin Huang, Le An, Jaehun Jung, Shaokun Zhang, Hao Zhang, Johan Bjoerck, Jim Fan, Patrick Langechuan Liu, Sifei Liu, Xiaolong Li, Paris Zhang, Yilin Zhao, Subhashree Radhakrishnan, Shiyi Lan, Jose Alvarez, Sanja Fidler, Yan Wang, Xiaodong Yang, Yin Cui, Tsung-Yi Lin, Padmavathy Subramanian and more. We would also like to thank the NVIDIA infra, legal and data teams, including Xinyou Ma, Katherine Cheung, Timo Roman, and Yao Xu for their prompt and helpful support. Finally, the authors would like to additionally acknowledge the following teams, including Nemotron-Diffusion, Nemotron VLM, Cosmos, GR00T, Alpamayo, Gigas and Metropolis, for the engagement and downstream applications.
