% !TEX root = ../main.tex
\section{Introduction} \label{sec:intro}

Beyond their success in content creation, diffusion-based \cite{ho2022video,ho2022imagen,singer2022make,bar2024lumiere,blattmann2023stable,wan2025wan,seedance2026seedance} and autoregressive \cite{yan2021videogpt,villegas2022phenaki,gupta2024photorealistic} video models have demonstrated remarkable ability to synthesize temporally coherent and photorealistic sequences conditioned on text, images, and other control signals \cite{esser2023structure,guo2023animatediff,bruce2024genie,team2026advancing,sun2025worldplay}. This capability has motivated a growing body of work that interprets video models as implicit simulators of the physical world, enabling their use in robotics \cite{agarwal2026cosmos,liao2025genie,li2026causal}, autonomous driving \cite{russell2025gaia,ren2025cosmos}, and interactive environments \cite{bruce2024genie}. In this paradigm, video models serve not only as generative systems but also as predictive world models \cite{assran2025v} that support planning, policy learning, and imagination-based control. However, translating these models from passive video generation to active embodied reasoning and intelligence remains an open challenge.

Despite their promise, a fundamental gap persists between video generation models and embodied intelligence requirements. Most video foundation models are optimized for perceptual quality—such as realism, aesthetics, and text alignment—rather than physical correctness or controllability. While these objectives yield visually compelling results, they do not explicitly enforce consistency with physical interaction constraints, such as contact stability, rigid-body dynamics, or long-horizon state consistency under intervention. This highlights a key tension: while internet-scale video provides rich visual diversity, it does not guarantee fidelity to the constraints of embodied interaction.

Existing efforts to bridge video generation and embodied intelligence typically fall short across three tightly coupled dimensions: architecture, data, and training objectives. Architecturally, most diffusion-based video transformers rely on dense computation, where all parameters are activated uniformly across tokens and timesteps, resulting in prohibitive inference costs and limited scalability. While recent sparse Mixture-of-Experts (MoE) formulations~\cite{deepseek2024v3,lepikhin2020gshard} demonstrate promising efficiency gains in large language models (LLM) \cite{achiam2023gpt,touvron2023llama,team2023gemini,bai2023qwen,bi2024deepseek,zeng2022glm}, their adoption in video generation remains limited. 
Data-wise, training corpora are dominated by internet videos lacking robot embodiment priors or precise interaction dynamics, leading to weak grounding in physical simulation.
From a training perspective, current alignment strategies primarily optimize aesthetic quality and text-video correspondence, without incorporating explicit physical feasibility, task completion, or long-horizon reward signals. Consequently, existing systems struggle to simultaneously achieve scalability, physical consistency, and embodiment grounding.


In this work, we propose \method, a DiT-based video pretraining paradigm specifically designed for embodied intelligence. Our approach addresses the aforementioned limitations through three integrated components. First, we introduce a \textbf{Mixture-of-Experts (MoE) video frameword}, which enables sparse conditional computation, improving inference efficiency while scaling model capacity for complex spatiotemporal dynamics. Second, we construct \textbf{a robot-augmented pretraining corpus} that unifies internet-scale videos with robot manipulation, navigation, and egocentric datasets, thereby injecting explicit embodiment and interaction priors into the model. Third, we develop \textbf{a multi-dimensional reward system} that extends beyond aesthetic objectives to incorporate physical rationality and task-oriented success signals, encouraging the model to learn dynamics and interactions consistent with embodied environments. Together, these components enable a more physically grounded and computationally efficient video foundation model.

Overall, we present \method, to the best of our knowledge the \textbf{first} large-scale open-source Mixture-of-Experts video foundation model for embodied intelligence, bridging the gap between digital video generation and physical actuation. Our contributions are threefold:
\begin{itemize}
    \item We introduce a sparse Mixture-of-Experts (MoE) video diffusion framework with a scalable training paradigm, enabling scalability–efficiency trade-off for spatiotemporal modeling.
    \item We develop a dedicated data profiling engine that systematically analyzes, filters, and rebalances heterogeneous video sources. This enables effective integration of large-scale internet videos with embodied datasets, resulting in improved grounding in physical interactions, action semantics, and embodiment-specific dynamics.
    \item We propose a multi-dimensional reward system that incorporates physical plausibility and task-level success signals, extending beyond conventional perceptual and text-alignment objectives.
\end{itemize}
