\section{Related Work}
\label{sec:related}

\paragraph{Spatial Intelligence in VLMs.}
Recent work has sought to improve the spatial intelligence of VLMs by scaling spatial supervision, introducing geometry-aware architectures, or designing spatially focused training objectives. 
Cambrian-S~\citep{cambrian_s} and SenseNova-SI~\citep{sensenova_si} construct large-scale spatial instruction data, while Spatial-MLLM~\citep{spatial_mllm} and VST~\citep{vst} inject explicit spatial modeling or visual spatial tuning into multimodal backbones. 
Other works, such as SpaceR~\citep{spacer}, ViLaSR~\citep{vilasr}, MindCube~\citep{mindcube}, and SpatialLadder~\citep{spatialladder}, further improve spatial reasoning through reinforcement learning, verifiable rewards, or curriculum design. 
These efforts have advanced performance on spatial benchmarks such as BLINK~\citep{blink}, 3DSR~\citep{threedsr}, EmbSpatial~\citep{embspatial}, MMSI-Bench~\citep{mmsi_bench}, and VSI-Bench~\citep{vsi_bench}. 
However, most of them remain \textit{training-driven and single-shot}: the model is expected to encode spatial capability into its parameters and produce an answer in one forward pass, relying on the model's internalized spatial knowledge rather than explicit, scene-specific evidence acquisition at inference time.


\paragraph{Agentic Spatial Reasoning.}
Tool-use agents extend language and vision-language models by interleaving reasoning with calls to external tools, as shown in general agent frameworks such as ReAct~\citep{react} and visual tool-use systems such as ViperGPT~\citep{vipergpt}, Visual ChatGPT~\citep{visual_chatgpt}, and MM-ReAct~\citep{mm_react}. 
More recent work brings this paradigm to spatial reasoning by equipping VLMs with explicit geometric tools or structured computation. 
VADAR~\citep{vadar} synthesizes Python programs over dynamically constructed 3D APIs, SpaceTools~\citep{spacetools} trains VLMs to coordinate vision and robotic tools through reinforcement learning, and GCA~\citep{gca} constrains the reasoning process with formal reference-frame and objective constraints before deterministic geometric computation. 
Concurrent to these efforts, Think3D~\citep{zhang2026think3d} equips VLM agents with 3D reconstruction and camera-manipulation tools, enabling active exploration through ego/global-view switching and novel-view rendering. 
These methods demonstrate the promise of agentic spatial reasoning, but they are still limited in capturing the continuous nature of human spatial understanding, where partial observations are integrated over time, object states are maintained across viewpoints, and spatial judgments are made from an evolving scene representation.

\paragraph{Long-video and Multi-view Understanding.}
Methods commonly handle continuous observations through frame compression or reconstruction-first pipelines. 
Frame-compression methods sample, retrieve, or summarize a limited set of frames before feeding them to long-context VLMs~\citep{videollama,longva,tstar}, improving efficiency but risking the loss of question-relevant spatial evidence. 
Reconstruction-first methods instead build an explicit 3D representation using multi-view geometry or feed-forward reconstruction models~\citep{vggt,da3}, providing stronger geometric grounding but often incurring unnecessary computation when the query only requires sparse or localized evidence. 
However, the selected frames or reconstructed geometry are typically consumed as fixed context, leaving spatial grounding, cross-view association, and metric comparison largely to implicit reasoning or a separate downstream step. 
Thus, they improve access to visual or geometric information, but do not fully close the loop between evidence acquisition, spatial computation, and persistent scene-level reasoning.





% \paragraph{Spatial Intelligence in Vision-Language Models.}
% A first line of work attacks the ``spatial gap'' of generic VLMs by scaling spatial supervision and reshaping the training recipe. Cambrian-S~\cite{cambrian_s} and SenseNova-SI~\cite{sensenova_si} curate massive corpora of metric 3D scenes and spatial QA, while Spatial-MLLM~\cite{spatial_mllm} and VST~\cite{vst} inject geometry-aware encoders or visual spatial tuning into the backbone. A second cluster -- SpaceR~\cite{spacer}, ViLaSR~\cite{vilasr}, MindCube~\cite{mindcube}, and SpatialLadder~\cite{spatialladder} -- pushes reinforcement learning with verifiable rewards or curricula tailored to multi-view mental rotation, distance, and layout. These models are evaluated on increasingly hard spatial benchmarks: BLINK~\cite{blink}, 3DSR~\cite{threedsr}, and EmbSpatial~\cite{embspatial} for single-image probing, and MMSI-Bench~\cite{mmsi_bench} and VSI-Bench~\cite{vsi_bench} for multi-image and long-video regimes. All of the above are \emph{single-shot, training-driven} VLMs: the model emits an answer in one forward pass and never invokes external perception. S-Agent is orthogonal to this axis -- we add no new training data and instead introduce an explicit agent loop that dispatches spatial tools at inference time.

% \paragraph{Tool-Use Agents and Visual Reasoning.}
% The agent paradigm equips an LLM or VLM with a controller that interleaves reasoning with calls to external tools. ReAct~\cite{react} and Toolformer~\cite{toolformer} established the template at the language side, while ViperGPT~\cite{vipergpt}, Visual ChatGPT~\cite{visual_chatgpt}, and MM-ReAct~\cite{mm_react} extended it to visual question answering by composing detectors, captioners, and code executors. These systems have driven strong gains in generic VQA, web navigation, and software engineering, demonstrating that explicit tool dispatch can outperform monolithic models on tasks that demand structured perception. Yet the \emph{spatial-3D} domain has remained largely beyond their reach: the available tools target 2D recognition, and the controllers are not designed to reason over 3D geometry, viewpoint, or temporal grounding. Bringing the agent paradigm to spatial intelligence therefore requires a new set of tools and a new orchestration policy, which a small concurrent literature has begun to explore.

% \paragraph{Agentic Spatial Reasoning.}
% Three concurrent works are the closest prior art to ours. VADAR~\cite{vadar} (CVPR 2025) proposes a Pythonic dynamic API in which an LLM synthesizes programs that call 3D primitives such as depth and pose; it is evaluated on Omni3D-Bench, where every question is grounded in a \emph{single image}, and it does not handle multi-view sets or video. SpaceTools~\cite{spacetools} (Dec.\ 2025) trains a VLM via dual-loop reinforcement learning to coordinate depth, segmentation, and pose tools, reporting results on RoboSpatial-Home, BLINK, and BOP-ASK; all of these benchmarks again expose the agent to a \emph{single image at a time}, with no multi-view fusion or long-video reasoning. GCA~\cite{gca} (Nov.\ 2025) is a training-free, geometrically-constrained agentic VLM that invokes 3D tools through structured prompts, but its evaluation is likewise restricted to \emph{single-image} 3D QA and does not address multi-view or temporal inputs. In contrast, S-Agent is, to our knowledge, the first tool-use agent that operates on \emph{multi-view image sets and long-form ego videos} -- the regimes captured by MMSI-Bench~\cite{mmsi_bench} and VSI-Bench~\cite{vsi_bench} -- and that schedules its 3D operations across views and time rather than over a fixed single frame.

% \paragraph{Long-Video and Multi-View Understanding.}
% Existing pipelines for long videos and multi-view inputs broadly fall into two camps. Frame-compression methods sparsely sample a fixed budget of frames and run a single forward pass through a long-context VLM (e.g., VideoLLaMA~\cite{videollama}, LongVA~\cite{longva}); they trade temporal fidelity for tractable context length and commit to one viewing pattern regardless of the question. Reconstruction-first methods instead pre-build an offline 3D representation -- via VGGT~\cite{vggt}, Depth-Anything-3~\cite{da3}, or a COLMAP-style pipeline -- and then query that frozen geometry, paying a heavy compute cost per scene whether or not the downstream question actually needs dense 3D. S-Agent takes a third route: the agent dynamically selects which views to attend to via T$^\star$ keyframe search~\cite{tstar} and decides which 3D operation to invoke per question, so perception is \emph{on-demand and question-conditional}. This couples the strengths of both lines -- multi-view geometric grounding and long-video temporal selection -- without paying their full cost on every query, and is what enables S-Agent to scale tool use to the multi-view and long-video regimes that prior agentic spatial systems do not address.
