% Claim:
%   The report contributes a session-centered runtime dataflow, a compact
%   object vocabulary, backend-aware boundaries, and an evidence-oriented
%   release protocol.
% Evidence:
%   initial-tech-report/references/OpenRath-README.md lines 22-84 and 170-241;
%   audited local OpenRath source snapshot and generated evidence packets.
% Open questions:
%   Implementation Milestones and evidence packets should anchor each
%   contribution to concrete files, commands, and artifacts.

This report makes four technical contributions to the runtime-state boundary
for agent systems.

\begin{enumerate}[itemsep=0.1em, topsep=0.2em]
  \item \textbf{A session-centered runtime dataflow.}
  OpenRath treats \texttt{Session} as the value that moves through the agent
  runtime, so conversation chunks, placement, lineage, usage, pending work,
  tool evidence, and memory-boundary records are represented as one inspectable
  flow rather than separate controller bookkeeping.

  \item \textbf{A PyTorch-like object vocabulary for agent programs.}
  The framework organizes agent programs around
  \texttt{Session}, \texttt{Sandbox}, \texttt{Tool}, \texttt{Agent},
  \texttt{Memory}, \texttt{Workflow}, and \texttt{Selector}. Each object has a
  narrow runtime boundary while preserving the same \texttt{Session -> Session}
  shape, including runtime-routed control flow.

  \item \textbf{Backend-aware boundaries for tools and memory.}
  OpenRath separates runtime state from the execution backend that runs tools
  and the memory backend that persists recallable state. This lets local
  execution, optional OpenSandbox placement, MCP-style tools, and memory
  services participate in one session-centered model as their evidence packets
  are verified.

  \item \textbf{An audit-first release protocol.}
  The report maps claims to packets: lineage export, local sandbox execution,
  workflow transcript, focused tests, visual QA, claim ledger, and a
  memory source audit. Broad benchmark superiority, human preference
  results, and cross-system leaderboard claims are reserved for follow-on
  quantitative evaluation.
\end{enumerate}
