% Claim:
%   OpenRath's programming model is a compact Session-centered vocabulary, not
%   a large taxonomy of agent framework objects.
% Evidence:
%   initial-tech-report/references/OpenRath-README.md lines 22-164 and 170-245.
% Open questions:
%   Memory remains an intended runtime plane until local source/example/test
%   evidence passes the release smoke suite.

OpenRath keeps the programming model small on purpose. The core rule is that
runtime components transform or annotate \texttt{Session}; they should not each
invent a private transcript, placement record, tool log, memory format, or
workflow state. This rule is what makes the PyTorch analogy useful: the analogy
is not about tensor math, but about one value flowing through reusable
transformations with explicit placement and persistent state boundaries.

\begin{table}[H]
\vspace{0.15em}
\centering
\small
\caption{The compact OpenRath object vocabulary.}
\label{tab:programming-model-vocabulary}
\begin{tabularx}{\linewidth}{@{}lX@{}}
\toprule
Object & Runtime boundary \\
\midrule
\texttt{Session} & Flowing runtime value for chunks, placement, lineage, usage, pending work, tool evidence, and memory evidence when enabled. \\
\texttt{Agent} & Reusable \texttt{Session -> Session} transformation with local prompt, provider, tools, and memory policy. \\
\texttt{Tool} & Model-visible callable operation backed by schema validation, session context, sandbox dispatch, and returned evidence. \\
\texttt{Sandbox} & Placement boundary for file, command, code, and external tool execution. \\
\texttt{Memory} & Intended persistent-state plane for recall and commit across runs, kept separate from prompt text. \\
\texttt{Workflow} & Composition surface for agents, tools, branches, compression, memory, and child workflows. \\
\texttt{Selector} & Runtime router over self-describing workflows: it reads the current session and picks the next workflow, so dynamic control flow stays explicit instead of hard-coded. \\
\bottomrule
\end{tabularx}
\vspace{0.15em}
\end{table}

\begin{figure}[!htbp]
\centering
\begin{tikzpicture}[
  font=\footnotesize\sffamily,
  leftbox/.style={draw=RathBlueMorandi, fill=RathBlueMorandiSoft, minimum width=2.05cm, minimum height=0.58cm, align=center, inner sep=2pt},
  rightbox/.style={draw=RathPinkMorandi, fill=RathPinkMorandiSoft, minimum width=2.05cm, minimum height=0.58cm, align=center, inner sep=2pt},
  pt/.style={text=RathMuted},
  ra/.style={->, draw=RathPinkMorandi, line width=0.5pt}
]
\node[draw=RathBlueMorandi, fill=RathBlueMorandiSoft, line width=0.55pt, minimum width=3.15cm, minimum height=5.95cm] at (0, -2.55) {};
\node[draw=RathPinkMorandi, fill=RathPinkMorandiSoft, line width=0.55pt, minimum width=3.15cm, minimum height=5.95cm] at (4.1, -2.55) {};
\foreach \i/\ptlabel/\orlabel in {
  0/{\texttt{Tensor}}/{\texttt{Session}},
  1/{\texttt{Device}}/{\texttt{Sandbox}},
  2/{\texttt{Parameter}}/{\texttt{Memory}},
  3/{Function}/{\texttt{Tool}},
  4/{\texttt{nn.Linear}}/{\texttt{Agent}},
  5/{\texttt{nn.Module}}/{\texttt{Workflow}},
  6/{control flow}/{\texttt{Selector}}
}{
  \node[leftbox] (p\i) at (0, -\i*0.78) {\ptlabel};
  \node[rightbox] (o\i) at (4.1, -\i*0.78) {\orlabel};
  \draw[ra] (p\i.east) -- (o\i.west);
}
\node[pt, anchor=south, font=\footnotesize\sffamily\bfseries] at (0, 0.72) {PyTorch};
\node[anchor=south, text=RathInk, font=\footnotesize\sffamily\bfseries] at (4.1, 0.72) {OpenRath};
\end{tikzpicture}
\caption{The PyTorch lens. Each agent-runtime concern maps onto one OpenRath
object, with \texttt{Session} as the flowing value (the tensor of the runtime)
and \texttt{Selector} routing control flow at run time. The mapping is a
teaching device, not a claim that agent systems are neural networks.}
\label{fig:pytorch-lens}
\end{figure}

The most important design choice is what each object does not own. An
\texttt{Agent} does not own the entire conversation graph; lineage belongs to
\texttt{Session}. A \texttt{Tool} does not own placement; it executes through
the active sandbox. A \texttt{Workflow} does not create a separate orchestration
state; it composes transformations over sessions. \texttt{Memory} does not
become hidden prompt text; recall and commit should remain visible runtime
events. These separations keep the system inspectable when a run becomes
multi-agent, multi-branch, and multi-backend.

The tool boundary illustrates the pattern. A flow-level tool exposes a name,
description, and JSON schema to the model, while its Python call receives the
active \texttt{Session} and validated arguments. Built-in tools can then create
backend payloads for file, command, code, or MCP-like execution without changing
the model-visible contract. The same principle applies to workflows: a workflow
may fork a session, call an agent, validate in a sandbox, compress context, and
return a new session, but the evidence remains attached to the shared runtime
value.

Control flow follows the same discipline through \texttt{Selector}. Rather than
hard-coding which agent or workflow runs next, a \texttt{Selector} reads the
current \texttt{Session} and routes to one of several self-describing
workflows, returning an empty workflow when the task is done. This keeps
branching and looping over agents as ordinary, inspectable runtime decisions:
the routing choice becomes part of the session record instead of vanishing into
controller code. It is also where OpenRath departs from a static workflow
graph---the next step is decided at runtime from session state, yet every
decision still flows through one value.

Memory is described with a deliberately bounded claim. OpenRath provides local
memory with lexical recall, optional embeddings, and an optional external
backend, exposed through agent-level recall and commit operations so that
remembering and recalling stay visible runtime events rather than hidden prompt
text. What this report does not claim is retrieval \emph{quality}: how well a
given corpus, embedding choice, and commit policy serve a task is an empirical
question left to a follow-on evaluation. The programming model reserves the
correct boundary---memory as a session-visible persistent plane---without
asserting that every quality and backend trade-off has been measured.
