% Claim:
%   OpenRath is motivated by the gap between loop-shaped agents and the
%   branchable, replayable runtime state required by multi-agent systems.
% Evidence:
%   initial-tech-report/references/OpenRath-README.md lines 22-84 and 170-245.
% Open questions:
%   Later release evidence should keep distinguishing implemented runtime
%   claims from future benchmark comparisons.

Section~1 framed the gap with a single run, and the related work above situated
OpenRath among adjacent agent systems---tool-using and acting agents, multi-agent
frameworks, runtime/observability layers, memory and retrieval, evaluation
environments, and provenance standards. This section states the gap that runs
underneath all of them as a general property of multi-agent work. The loop
boundary that suffices for one assistant becomes a weak state boundary the moment
an application branches across roles, tools, memory, files, sandboxes, and
resumed runs. The transcript still shows a final answer, but the runtime path
that produced it is spread across controller code, tool logs, memory stores,
workspace state, and provider traces.

This pressure is becoming more visible as agent products move from demos to
longer-running workflows. Once an agent edits a repository, calls external
tools, resumes after interruption, or routes work through multiple roles, the
state boundary becomes an engineering contract rather than an implementation
detail. Users and reviewers need to ask ordinary runtime questions: what was
the input state, what changed, which backend performed the change, which memory
or artifact influenced the decision, and how can the run be resumed or
replayed? A framework that cannot answer those questions may still produce a
plausible response, but it cannot easily support release review, debugging,
audit, or systematic evaluation.

This hidden state is the central motivation for OpenRath. Multi-agent work is
not merely one conversation with more roles. It naturally creates multiple
runtime paths: one branch gathers context, another edits or tests an artifact,
another validates evidence, and another compresses the result. If every
intermediate step is placed into one shared transcript, later agents inherit too
much noise. If intermediate work is hidden in controller state, reviewers cannot
reconstruct which branch produced a claim, what memory was recalled, which
sandbox touched a file, or what evidence was discarded during compression.

OpenRath treats branchability as a property of runtime state rather than as a controller-side convention. A branch should inherit the portion of parent context required for independent work, accumulate local evidence during execution, and merge useful results back without erasing provenance. The object being branched is therefore \texttt{Session}: the runtime value that flows through agents, tools, memory-boundary operations, sandbox placement, compressors, and workflows.

This boundary is the foundation for the remainder of the report. The next section makes it concrete through a compact object vocabulary centered on \texttt{Session}, the value that every other component reads, transforms, annotates, or passes forward.

