% Claim:
%   OpenRath is a milestone implementation with a precise, bounded claim
%   status, not a roadmap or a broad benchmark result.
% Evidence:
%   Audited local OpenRath source snapshot inspected May 29, 2026; generated
%   packets verify the session, backend, tool, lineage, and workflow surfaces
%   and mark memory and optional OpenSandbox as not yet substantiated.
% Open questions:
%   A public archival release still needs a clean source snapshot and
%   quantitative benchmark runs before broader empirical claims.

OpenRath is a working implementation, not an architecture sketch. It is
distributed as a Python package whose modules realize the objects of the
preceding sections: a session core; an execution-backend layer; the flow layer
of tools, agents, workflows, and compressors; an LLM-provider layer; and
persistence with lineage export. This report is written against an audited
snapshot of that codebase, which we treat as adequate for technical review
rather than as a tagged archival release.
Table~\ref{tab:implementation-claim-status} records which surfaces the current
implementation substantiates and where its claims are deliberately bounded.

\begin{table}[!htbp]
\vspace{0.15em}
\centering
\small
\caption{Implementation surface and claim status.}
\label{tab:implementation-claim-status}
\begin{tabularx}{\linewidth}{@{}>{\raggedright\arraybackslash}p{0.22\linewidth}>{\raggedright\arraybackslash}X@{}}
\toprule
Surface & Status \\
\midrule
Session core & Implemented runtime value: ordered chunks, branching via \texttt{fork}, \texttt{detach}, and \texttt{merge}, usage accounting, and JSONL lineage export, exercised by focused tests. \\
Backend placement & Local execution implemented and verified; OpenSandbox available as an optional backend, unconfigured in this environment. \\
Tool layer & Implemented boundary: model-visible schemas with backend-dispatched side effects returned to the session, with custom-tool and MCP examples. \\
Agent and workflow & Implemented composition surface over the uniform \texttt{Session -> Session} contract, including a scripted multi-stage workflow. \\
Provider layer & Live-inference prerequisites in place; model quality and provider runs are out of scope for this report. \\
Memory plane & Intended runtime plane, not yet substantiated by a local module with examples and tests (see below). \\
Examples & A worked example set spanning lineage, backends, tools, streaming, usage, and multi-agent workflows. \\
\bottomrule
\end{tabularx}
\vspace{0.15em}
\end{table}

The substantiated claims are deterministic and local. The session core---its
ordered chunks, branching operations, usage accounting, and JSONL lineage
export---is exercised by focused tests, as are local sandbox placement and the
tool-dispatch path; tool and workflow behavior is further demonstrated by
custom-tool, MCP, and scripted-workflow examples. The optional OpenSandbox
backend and the LLM-provider layer are present but environment- or
provider-dependent, and live model quality lies outside the scope of this report.

What the milestone demonstrates is not any individual surface but that they
share a single object model. \texttt{Session} is the value that flows; backends
determine where code and tools execute; tool calls enter the session as
structured events rather than disappearing into an executor log; agents and
workflows transform sessions without maintaining private transcript formats; and
persistence with lineage export renders the resulting state inspectable outside
the running process. This is the minimum structure an agent runtime requires to
support branching, audit, and replay without rebuilding a separate observability
system for each application.

