% Claim:
%   The report should close by restating the narrow, memorable thesis:
%   OpenRath is about making agent runtime state explicit as Session.
% Evidence:
%   Earlier sections define the programming model, implementation milestones,
%   ecosystem boundary, evidence protocol, and scoped limitations.
% Open questions:
%   The conclusion should stay stable as evidence packets mature; avoid adding
%   claims here that are not backed elsewhere in the report.

OpenRath's contribution is deliberately narrow: it makes the state that agents
operate on explicit. A multi-agent system is not only a prompt graph, a tool
registry, a trace stream, or a benchmark harness. It is a runtime in which
conversation chunks, branch lineage, sandbox placement, tool effects, memory
interactions, usage, artifacts, and replay evidence must remain connected.

\texttt{Session} is OpenRath's proposed boundary for that runtime state. The
programming model is compact because every major component either transforms a
\texttt{Session}, annotates it, dispatches work through its placement, or emits
evidence that can be inspected after the run. Because that evidence lives in the
value the program already passes around rather than in a side channel
reconstructed afterward, it stays available exactly when a reviewer needs it.
This makes OpenRath complementary to graph runtimes, tracing SDKs, tool
protocols, sandbox providers, and real-environment benchmarks rather than a
replacement for them.

The current technical report therefore makes a scoped claim: deterministic
runtime behavior can be reviewed through release packets today, while broader
quality comparisons, memory-quality evaluation, and live-provider results
belong in follow-on benchmark artifacts. The durable thesis is that reliable
agent systems need a first-class runtime value, and OpenRath makes
\texttt{Session} that value.

That thesis is also the practical standard for the next iteration of the
project. New capabilities should enter the report only when they preserve the
same boundary: they should transform a \texttt{Session}, attach evidence to a
\texttt{Session}, or expose a backend effect through a \texttt{Session}. This
keeps the system from becoming a collection of hidden side channels. It also
keeps the report honest: implementation milestones, case studies, and
evaluation claims can expand without changing the core argument that agent
systems become easier to compose, debug, review, and evaluate when their
runtime state is explicit.

If the last decade of deep learning made the tensor the value a network is
built around, the next generation of agent systems needs the same move: a
single runtime value that everything reads, transforms, and explains. OpenRath
proposes that value is the \texttt{Session}.
