\section{Related Works}
\label{sec:rw}

\noindent\textbf{Evaluation of VLMs' spatial reasoning.}
A growing body of benchmarks has established that spatial reasoning remains a persistent weakness of modern vision-language models. Early datasets target brittleness in basic spatial predicates: SpatialSense~\cite{yang2019spatialsense} reduces language priors through adversarial crowdsourcing, while VSR~\cite{liu2022vsr} scales relation types in a caption-verification format, and What'sUp~\cite{kamath2023whatsup} uses minimal-pair testing to reveal systematic failures on left/right and above/below distinctions. More recent work shifts from 2D relations to viewpoint and 3D structure. 3DSRBench~\cite{ma20253dsrbench} shows that models fail under modest changes in perspective, depth, and occlusion, and ViewSpatial-Bench~\cite{li2025viewspatialbench} identifies a ``perspective gap'': models often succeed in camera-centered views but break when asked to adopt human-centered viewpoints.

Benchmarks have also expanded to multi-image and video settings where maintaining a consistent spatial state is essential. VSI-Bench~\cite{yang2025thinkinginspace} tests whether models can build a persistent mental map from videos, while MMSI-Bench~\cite{yang2025mmsibench} reports large human--model gaps on cross-view scene reconstruction. MindCube~\cite{yin2025mindcube} is closely aligned with our motivation, targeting spatial mental modeling from limited views, including perspective-taking and ``what-if'' scene dynamics. Counting Stacked Objects~\cite{dumery2025countingstackedobjects} studies 3D object counting under heavy occlusion across multiple views, directly analogous to our Multiview Counting setting. Finally, benchmark design work emphasizes that shortcuts remain pervasive: Brown \etal~\cite{brown2025trainontestset} construct VSI-Bench-Debiased by iteratively pruning samples solvable via priors, reinforcing the need for evaluations where success requires genuine spatial computation.

Collectively, these benchmarks diagnose \emph{where} VLMs fail spatially, but they typically evaluate \emph{discriminative} understanding---reading off a relation from an observed view---rather than \emph{constructive} spatial imagination. Our work is complementary: we isolate \emph{imaginative perception} as a standardized intermediate substrate, and pair each task with a ground-truth intermediate spatial imagination rather than only a final answer label.


\noindent\textbf{Intermediate representations for spatial reasoning.}
Chain-of-thought prompting~\cite{wei2022cot} can improve multi-step reasoning, but serializing viewpoint transformations, occlusions, and geometric constraints into language is often awkward and error-prone. This motivates intermediate representations in modalities better aligned with spatial computation. One direction externalizes reasoning into explicit visual buffers: Visual Sketchpad~\cite{hu2024visualsketchpad} equips models with drawing actions for iterative refinement, and MVoT~\cite{li2025mvot} introduces visualization-of-thought traces that help on dynamic spatial tasks where text CoT struggles. ThinkMorph~\cite{gu2025thinkmorph} studies interleaved text--image reasoning traces, and OpenAI describes o3/o4-mini as using chains-of-thought that include simple image transformations during reasoning~\cite{openai2025thinkingimages}. A complementary line introduces latent visual scratchpads: Mirage~\cite{yang2025mirage} frames latent tokens as ``machine mental imagery,'' and Mull-Tokens~\cite{ray2025mulltokens} generalizes to modality-agnostic latent thinking tokens.

Our work differs in what the intermediate is meant to represent. Many prior approaches treat intermediate images or latents as optional visualizations of \emph{visible} structure. We instead target \emph{imaginative perception}: predicting what would be perceived under an unobserved spatial configuration (e.g., a rotated viewpoint or a top-down path state), a representation constrained by the input but not present in it. This framing provides a principled criterion for when intermediate visual thoughts are necessary and a controlled way to supervise them.


\noindent\textbf{Unified multimodal models for interleaved understanding and generation.}
Producing imaginative perceptual intermediates within a single model requires the ability to both understand and generate images. Unified decoder-only architectures treat image tokens as first-class sequence elements, enabling arbitrary text--image interleaving. Chameleon~\cite{chameleon2024chameleon} is an early example, while Show-o2~\cite{xie2025showo2} and Janus~\cite{chen2025januspro} offer alternative unified designs that balance understanding and generation. We build on BAGEL~\cite{deng2025bagel}, a unified model pretrained on large interleaved corpora that exhibits strong spatial capabilities, making it a natural substrate for producing intermediate spatial imaginations. Crucially, however, a unified architecture alone does not guarantee that intermediate images are \emph{used} in a way that supports reasoning; our work provides task constructions and supervision that make imaginative perception the relevant computational substrate.