\section{Related Work}
\label{sec:related}

\paragraph{Post-training workload.}
MinT targets a workload where post-training leaves many policy variants over a smaller set of base models. Yao argues that future progress shifts more weight toward problem definition, evaluation, and agent-environment interaction~\citep{yao_second_half_2025}. Silver and Sutton frame future agents as learning from streams of experience in addition to static human data~\citep{silver_sutton_experience_2025}. Recent frontier-model reports expose the same systems pressure through reasoning, coding, tool use, executable environments, asynchronous agent RL, and long-horizon evaluation~\citep{deepseek_v4_release_2026,deepseekv4_2026,glm5_2026,kimi_k25_2026,minimax_m27_2026,qwen35_2026,openai_gpt55_2026,anthropic_opus45_2025,anthropic_agent_autonomy_2026}. These workloads produce task variants, evaluation candidates, product branches, tenant adapters, rollback points, and personal adapters that share base models while retaining separate histories.

\paragraph{Service interfaces.}
Remote post-training interfaces make training loops programmable. Tinker exposes low-level post-training primitives through a remote service~\citep{tinker2025,tinker_cookbook}. SkyRL tx implements a Tinker-style backend and documents model and benchmark coverage across Qwen3 dense and MoE models, Llama 3, DeepSeek V3, GSM8K, and DAPO/AIME recipes~\citep{skyrl_tx}. OpenTinker frames a broader RL-as-a-Service stack around agents, environments, protocols, scheduling, training, and inference~\citep{opentinker2026}. Tinker specifies how users program remote post-training; MinT keeps a compatible programming surface and defines the LoRA-specific service path underneath it: policy record, training checkpoint, rollout record, exported adapter revision, and adapter cache state.

\paragraph{RL execution systems.}
HybridFlow/verl, AReaL, OpenRLHF, Relax, ROLL, StreamRL, AsyncFlow, Laminar, and NeMo-Aligner study rollout scheduling, colocated and disaggregated execution, asynchronous optimization, GPU utilization, failure isolation, and end-task quality~\citep{hybridflow2025,areal2025,openrlhf2024,relax2026,roll2025,streamrl2025,asyncflow2025,laminar2025,nemoaligner2024}. Their central systems objects are actors, rollout replicas, parameter services, queues, and placement groups. MinT uses the same execution concerns and adds LoRA-specific service state: adapter revisions, optimizer state, rollout records, MoE route records, DSA correction metadata, exported adapters, and adapter cache state stay tied to the policy that produced them.

\paragraph{Training-serving consistency.}
Modern RL pipelines often generate rollouts through an inference engine and score them through a training backend. Yao et al. study token-probability gaps in hybrid vLLM/FSDP pipelines and propose truncated importance sampling~\citep{rollout_training_mismatch2025}. Jet-RL shows that mixed BF16 training and FP8 rollout can destabilize on-policy RL under long generations, then uses a unified precision flow to reduce numerical mismatch~\citep{jetrl2026}. R3 studies MoE router disagreement between training and inference and reuses inference-time expert-route records during training~\citep{r3_moe_router2025}. MinT addresses the same problem by storing the adapter revision, rollout record, MoE route or DSA correction metadata, and served evaluation path that name which behavior generated and scored a token.

\paragraph{Parameter-efficient tuning and multi-LoRA training.}
LoRA freezes the base model and trains low-rank matrices attached to selected layers~\citep{lora2022}. AdaLoRA allocates rank budget across matrices according to importance~\citep{adalora2023}. QLoRA combines frozen quantized bases with LoRA updates for memory-efficient fine-tuning~\citep{qlora2023}. LoRA Without Regret argues that LoRA can reach strong post-training quality~\citep{lora_without_regret_2025}, supporting the premise that LoRA can be more than a memory-saving approximation. MinT turns that premise into managed infrastructure for RL service operation: each update restores one adapter and optimizer state over a resident base, each export produces a serving revision, and each rollout or serving request selects that revision through cache and residency state.

\paragraph{Multi-LoRA serving.}
Multi-LoRA serving systems optimize inference after adapters already exist. Punica, S-LoRA, dLoRA, dynamic operator optimization, and vLLM improve batching, memory management, scheduling, and kernels for many adapters over a shared base~\citep{punica2024,slora2023,dlora2024,zhou2025dynamic,vllm2023}. Compress then Serve reduces adapter storage and serving overhead through individual and joint LoRA compression~\citep{compress_then_serve2024}. FastLibra manages LoRA and KV cache dependencies in a unified HBM cache~\citep{fastlibra2025}. LoRAServe handles rank heterogeneity, placement, and routing across distributed LoRA serving clusters~\citep{loraserve2025}. MinT connects the serving catalog to the training lifecycle: an addressable adapter is an exported revision of a trained policy, and a cache miss reloads that revision from the policy's exported adapter file.

\paragraph{Large-model infrastructure.}
Ray provides the distributed execution framework used by many AI systems~\citep{ray2018}. Megatron-LM, Efficient Megatron-LM, ZeRO, and MoE Parallel Folding provide the model-parallel, memory-partitioning, and MoE-parallel techniques that make large-model and sparse-MoE training practical~\citep{megatronlm2019,efficient_megatronlm2021,zero2019,moe_parallel_folding2026}. DeepSeek-V4 describes large-model infrastructure through concrete state units such as hybrid KV cache entries, state caches, on-disk cache storage, rollout WALs, and teacher-state scheduling~\citep{deepseekv4_2026}. MinT applies the same resource-accounting style to LoRA RL: base deployments stay resident, adapter revisions move across subsystems, and adapter cache state controls which exported policies are hot.
