\section{Introduction}
\begin{table}[t]
\caption{\textbf{Comparison between \bench\ and existing image editing benchmarks.} 
\texttt{AVR} denotes Algorithm Visual Reasoning, \texttt{MIA} denotes Multi-Image Awareness, \texttt{WKR} denotes World Knowledge Reasoning, \texttt{DM} denotes Dynamic Manipulation, and \texttt{HP} denotes Human Preference evaluation.}
\vspace{0.5em}
\label{tab:compare_image_editing_bench}
\centering
\resizebox{ 0.95\linewidth}{!}{%
\begin{tabular}{lcccccccc}
\toprule
\textbf{Benchmark} & \textbf{Size} & \textbf{Tasks} & \textbf{Lang.} & \textbf{AVR} & \textbf{MIA} & \textbf{WKR} & \textbf{DM}& \textbf{HP} \\
\midrule

Magicbrush~\cite{zhang2023magicbrush} & 1,053 & 5 & EN & \icono & \icono &\icono & \icono &  Low \\
Emu Edit~\cite{cui2025emu3} & 3,055 & 7 & EN & \icono & \icono & \icono & \icono  & Low \\
AnyEdit-Bench~\cite{yu2025anyedit} & 1,250 & 25 & EN & \icono & \icono & \icono & \icono & Low \\
ImgEdit-Bench~\cite{ye2025imgedit} & 811 & 11 & EN & \icono & \icono & \icono & \icono & Mid \\
GEdit-Bench~\cite{liu2025step1x} & 606 & 14 & EN/ZH & \icono & \icono & \icono & \icono & Mid \\
ICE-Bench~\cite{pan2025ice} & 6,538 & 31 & EN & \icono & \icono & \icono & \icono & Low \\
RISE-Bench\cite{zhao2025envisioning} & 360 & 4 & EN & \icono & \icono & \icohalf & \icono  & Mid \\
UniREditBench~\cite{han2025unireditbench} & 2,700 & 13 & EN & \icono & \icono & \icono & \icohalf  & Mid \\
WiseEdit~\cite{pan2025wiseedit} & 1,220 & 13 & EN/ZH & \icono & \icohalf  & \icoyes & \icono  & Mid \\
\midrule
\bench~(Ours) & 2,388 & 36 & EN/ZH & \icoyes & \icoyes & \icoyes & \icoyes  & High \\
\bottomrule
\end{tabular}%
}
% \vspace{-1.70em}
\end{table}

Recent image editing models~\cite{brooks2023instructpix2pix, chen2025opengpt,labs2025flux, wang2025scone,tong2026cof,zhu2026layermultilayerrepresentationfusion,zhao2024ultraedit,yu2025anyedit} have achieved remarkable progress, evolving from simple instruction-driven editing toward more advanced capabilities involving multimodal understanding, complex reasoning, and multi-image editing. 
As frontier models continue to improve, accurately evaluating their editing quality becomes increasingly challenging.
However, existing benchmarks~\cite{ye2025imgedit,liu2025step1x,pan2025ice} often exhibit a noticeable discrepancy between benchmark scores and human judgment, particularly for strong frontier models.
This limitation mainly stems from insufficient task difficulty and coarse-grained evaluation protocols, making it difficult to reliably distinguish subtle capability differences among advanced models.
Accurate evaluation is also crucial for reinforcement learning (RL) based image editing optimization.
Recent works such as EditScore~\cite{luo2025editscore} and EditReward~\cite{wu2025editreward} train reward models to support FlowGRPO-based~\cite{liu2025flow} image editing optimization.
However, existing reward model benchmarks often suffer from a distribution mismatch between evaluation samples and the edited images encountered during RL training, limiting their ability to faithfully assess reward model quality in realistic optimization settings.
Together, these limitations hinder a deeper understanding of frontier image editing models and their corresponding reward models, highlighting the need for a more comprehensive benchmark for faithful image editing evaluation.



To address these challenges, we introduce \textbf{{\bench}} and \textbf{{\rmbench}}, a unified evaluation suite for image editing models and reward models.
As illustrated in Figure~\ref{fig:gallery}, {\bench} contains $2,388$ carefully annotated instances spanning six progressively challenging task categories.
These tasks cover a diverse range of capabilities, including general editing, world perception, dynamic manipulation, visual reasoning, and multi-image understanding.
Beyond broad task coverage, {\bench} further adopts a fine-grained and multi-dimensional evaluation framework.
Each editing result is evaluated through chain-of-thought reasoning guided by well-defined scoring rubrics, enabling more reliable and interpretable assessment in complex editing scenarios.
This design better aligns benchmark evaluation with human judgment while improving evaluation consistency and sensitivity for frontier models.
In parallel, {\rmbench} contains $2,251$ preference pairs that simulate realistic decision-making scenarios encountered by reward models during RL optimization.
Together, this unified benchmark suite enables systematic evaluation of both frontier image editing models and reward models.
It further provides a realistic testbed for analyzing the effectiveness of reward-guided optimization in RL-based image editing.


To validate the effectiveness and difficulty of {\bench} and {\rmbench}, we conduct extensive evaluations on a broad range of frontier models, including $29$ image editing models and $21$ reward models.
For image editing, our evaluation covers state-of-the-art proprietary models, such as Nano-Banana Pro~\cite{nanobananapro}, Wan2.7-Image~\cite{wan}, and Seedream 4.5~\cite{seedream2025seedream}, as well as leading open-source models including Qwen-Image-Edit~\cite{wu2025qwen} and Joy-Image-Edit~\cite{joyaiimage2026}.
The results reveal a substantial performance gap between closed-source and open-source models.
The best proprietary model achieves an overall score of $3.99$, while the strongest open-source model, Qwen-Image-Edit~\cite{wu2025qwen}, reaches only $2.69$.
Beyond overall performance, fine-grained analysis further reveals clear weaknesses in multi-image understanding, world knowledge awareness, and visual reasoning, even for frontier models.
On the reward modeling side, native multimodal large language models~\cite{qwen3.5,qwen36_35b_a3b,zhang2025debiasing,wang2025monet,qwen3.6-27b,shi2025mavors} achieve stronger overall performance than existing open-source reward models, including models explicitly trained on preference data.
This finding suggests that current reward models remain limited in evaluating visual consistency and perceptual quality under complex editing scenarios.
Overall, our results reveal a fundamental limitation of current image editing systems: while existing models perform reasonably well on shallow perception-level editing tasks, they still struggle with deeper reasoning, world knowledge understanding, and complex multi-image editing.