\section{Introduction}
\label{sec:intro}

Reconstructing the 3D structure of a scene from 2D observations, commonly formulated as multi-view 3D reconstruction~\cite{hartley2003multiple, schonberger2016structure, furukawa2010towards, rosinol2020kimera, cramariuc2022maplab}, is a foundational problem in computer vision. It unlocks a wide range of real-world applications, including autonomous navigation~\cite{ceccarelli2022rgb, vargas2021overview, zhu2024llava}, robotics~\cite{kim2015real, lingbot-depth2026, 9506550}, and augmented and virtual reality~\cite{park2011handling, wloka1996interactive, okumura2006augmented}. Recent feed-forward reconstruction models~\cite{depthanything3, wang2025vggt, wang2025pi, wang2024dust3r, leroy2024grounding} have significantly advanced this task by replacing traditional multi-stage 3D reconstruction pipelines~\cite{hartley2003multiple, schonberger2016structure, 
furukawa2010towards} with end-to-end architectures that directly infer scene geometry from multi-view inputs. Built on transformer architectures~\cite{vaswani2017attention, dosovitskiy2020image}, their attention mechanism encode cross-view information to learn \textit{geometry-aware} representations, enabling accurate and scalable reconstruction under ideal imaging conditions.

% enabling accurate and scalable reconstruction in a single forward pass under ideal imaging conditions.

However, real-world multi-view observations often deviate from this ideal setting. In practice, captured images and video sequences frequently suffer from degradations such as motion blur induced by camera motion~\cite{nah2017deep, rim2020real, malyugina2025unsupervised, liu2024boosting, han2025emergent}. These effects obscure fine textures and structural cues essential for reliable feature extraction and cross-view matching. As a result, the learned representations become less discriminative, disrupting geometric consistency across views. Since feed-forward models~\cite{depthanything3, wang2025vggt, wang2025pi, wang2024dust3r, leroy2024grounding} directly infer scene geometry from these features in a single forward pass, they lack mechanisms to explicitly correct such errors, allowing them to propagate through the network and accumulate in the final reconstruction. Improving robustness to such imperfect inputs therefore remains a key challenge for achieving reliable and consistent performance in multi-view 3D reconstruction.


A key design question is where restoration should be performed in the 3D reconstruction pipeline. A straightforward approach would be to adopt a \textit{restore-then-reconstruct} paradigm (Fig.~\ref{fig:comparison} (a)), where degraded inputs are first restored in image space using existing image restoration models~\cite{zamir2022restormer, conde2024high, chen2023hierarchical, liang2024vrt, Youk_2024_CVPR, mao2025sir, zamfir2024complexityexperts} before being passed to the feed-forward reconstructor. However, current image restoration models are predominantly designed for single-view restoration and thus fail to leverage multi-view information and cannot enforce cross-view geometry consistency during restoration. While a recent multi-view restoration approach~\cite{mao2025sir} partially addresses this issue, it operates in a heavily compressed VAE-based latent space~\cite {kingma2013auto}, where information bottlenecks hinder the preservation of fine-grained details and geometric fidelity. Consequently, existing approaches remain suboptimal for multi-view image restoration and 3D reconstruction.

On the other hand, recent advances in Representation Autoencoders (RAEs)~\cite{zheng2025diffusion, tong2026scaling, kumar2026learning} further highlight the inherent shortcomings of conventional latent spaces. In particular, the compressed representations learned by VAEs~\cite{kingma2013auto, yao2025reconstruction} introduce information bottlenecks that hinder the preservation of fine-grained details and structural fidelity, which are essential for accurate multi-view 3D reconstruction. RAEs address this limitation by adopting high dimensional, semantically rich latent representations that better retain both global structure and local details, while maintaining a dedicated decoder for reconstructing high fidelity images. Motivated by this insight, we move beyond conventional image space~\cite{zamir2022restormer, chen2023hierarchical, conde2024high, liang2024vrt, Youk_2024_CVPR} and VAE-based formulations~\cite{mao2025sir} and instead exploit the geometry-aware feature space inherently encoded by feed-forward reconstruction models as a more suitable domain for denoising, while enabling image recovery through an auxiliary decoder.

To this end, we propose \textbf{Geometry-Aware Representation Denoising (GARD)}, a novel framework that learns a diffusion-based multi-view restoration denoiser model operating directly within the \textit{geometry-aware feature space} of a feed-forward reconstructor (Fig.~\ref{fig:comparison} (b)). By conducting denoising in this feature space, the proposed approach exploits high-dimensional representations inherently structured for scene geometry estimation, as well as the cross-view consistency encoded by feed-forward reconstruction models. This design preserves geometric fidelity while mitigating the information bottlenecks associated with VAE-based latent spaces and the inconsistencies introduced by image-space restoration. Furthermore, we adopt a dedicated image decoder~\cite{jang2026gld} to reconstruct high-quality RGB images from the refined representations, thereby enabling the joint recovery of high-quality imagery and accurate 3D scene geometry within a unified framework.

We validate our approach through extensive experiments on the Depth Anything 3 benchmark~\cite{depthanything3}, where controlled degradations are introduced to establish a rigorous evaluation protocol for restoration and reconstruction under motion blur degradation. To enable a fair comparison, we train the multi-view diffusion restoration model in both the geometry-aware feature space of the feed-forward reconstructor~\cite{depthanything3} and a conventional VAE-based latent space~\cite{kingma2013auto}, thereby isolating the impact of representation choice. We further compare our approach against restore-then-reconstruct pipelines for image restoration and scene geometry recovery, as well as dedicated image restoration methods evaluated using standard image quality metrics. Experimental results demonstrate that operating in the geometry-aware feature space yields improved geometric fidelity and visual quality, resulting in strong performance across pose estimation, 3D reconstruction, and image restoration benchmarks.





% Multi-view 3D reconstruction~\cite{hartley2003multiple, schonberger2016structure, furukawa2010towards, rosinol2020kimera, cramariuc2022maplab} aims to recover accurate scene geometry from a set of multi-view input images. Significant progress has been made with the advent of feed-forward reconstruction models~\cite{depthanything3, wang2025vggt, wang2025pi, wang2024dust3r, leroy2024grounding}, which jointly encode the multi-view inputs using transformer attention mechanism~\cite{vaswani2017attention, dosovitskiy2020image} to produce multi-view representations that are vital for capturing the underlying 3D geometry. Although effective, these models are typically trained and evaluated on high-quality inputs and therefore implicitly assume sharp, degradation-free multi-view imagery.

% In real-world scenarios, however, multi-view observations and video sequences often suffer from degradations, most notably motion blur induced by camera motion. Such degradations impair edge and local texture cues, impairing 3D reconstruction performance. Therefore, developing robust 3D reconstruction methods that can handle such imperfect inputs remains a important challenge for practical applications, including autonomous navigation~\cite{ceccarelli2022rgb, vargas2021overview, zhu2024llava}, drone manipulation~\cite{truong2020slimdeblurgan,Kolbeinsson_2024_CVPR, Fonder2019MidAir}, AR/VR~\cite{park2011handling, wloka1996interactive, okumura2006augmented}, and robotic perception~\cite{kim2015real, lingbot-depth2026, 9506550}.

% A straightforward solution is to employ a restore-and-reconstruct pipeline, where degraded inputs are first restored using off-the-shelf image restoration models~\cite{zamir2022restormer} before being passed to reconstruction models. However, most restoration approaches operate on single-view inputs, making it difficult to enforce multi-view constraints such as cross-view geometric consistency. While recent work~\cite{mao2025sir} extends restoration to multi-view settings, it performs denoising in a VAE latent space~\cite{kingma2013auto}, where heavily compressed, low-dimensional representations introduce an information bottleneck that limits the preservation of geometric structures and fine-grained visual details.

% Notably, we observe that feed-forward reconstruction models inherently encode multi-view representations that remain structurally informative even under degraded inputs. In particular, these representations continue to encode consistent geometric relationships across degraded views, suggesting that they provide a more suitable space for restoration than image space or compressed latent spaces. This observation motivates a shift in perspective: instead of restoring images independently, restoration should be performed directly in a geometry-aware multi-view feature space.

% Building on this insight, we identify two key requirements for effective multi-view restoration: (1) jointly modeling all views to enforce cross-view consistency and leverage additional view information, and (2) operating in a representation space that preserves geometric fidelity. To this end, we propose \ourmodel, a diffusion-based multi-view restoration framework that performs denoising directly in the latent feature space of feed-forward 3D reconstruction models. Our approach is inspired by Representation Autoencoders (RAEs)~\cite{zheng2025diffusion}, which demonstrate that high-dimensional, semantically rich latent spaces can preserve both global structure and fine-grained details more effectively than conventional VAE representations. By operating on multi-view features produced by reconstruction networks, \ourmodel naturally retains geometric cues and enables consistent restoration across views.


% This design offers several key advantages. First, \textbf{geometric consistency}: by denoising directly in a multi-view feature space, our method preserves cross-view geometric relationships that are often degraded in image-space or VAE-based restoration. Second, \textbf{robust recovery}: the learned representations jointly encode low-level textures and high-level structure, enabling reliable reconstruction under severe degradations such as motion blur. Third, \textbf{computational efficiency}: operating in a compact latent multi-view feature space avoids redundant high-resolution image processing, allowing scalable restoration for large-scale multi-view inputs.

% We validate our approach on two representative feed-forward reconstructors, VGGT~\cite{wang2025vggt} and DA3~\cite{depthanything3}, across tasks including pose estimation, 3D reconstruction, and monocular depth estimation. Experiments on a motion-blur-augmented benchmark demonstrate that \ourmodel consistently outperforms prior image-space and VAE-based restoration methods, highlighting the effectiveness of feature-space denoising for preserving geometric consistency and enabling high-fidelity 3D reconstruction under challenging real-world degradations.

