\section{Conclusion}\label{sec:conclusion}

This work has argued that visual generation is moving beyond the goal of producing visually plausible images toward a broader problem of \emph{visual intelligence}. The central question is no longer only how realistic an output looks, but what capabilities a model acquires as it becomes controllable, persistent, interactive, and eventually world-aware. To make this transition explicit, we proposed a five-level taxonomy---\emph{Atomic Generation}, \emph{Conditional Generation}, \emph{In-Context Generation}, \emph{Agentic Generation}, and \emph{World-Modeling Generation}---that organizes progress from one-shot rendering to causal simulation under intervention.

We used this taxonomy as the organizing thread for the paper. First, we reviewed the modeling foundations and architectural components of modern visual generators, from GANs, diffusion, flow matching, and autoregressive models to hybrid and unified multimodal systems. Second, we analyzed the full training lifecycle, emphasizing that recent progress is shaped not only by model scale but also by data density, VLM-driven relabeling, continued training, SFT, preference-based alignment, reward modeling, and inference acceleration. Third, we examined the data and infrastructure layer, showing how the field has shifted from passive web scraping toward actively engineered and synthetic data pipelines. Fourth, we organized applications from conditional generation and domain adaptation to reasoning-driven editing and embodied visual prediction.

Our main contribution is therefore not a catalogue of methods, but a capability-centered framework for reading a rapidly expanding field. The five-level taxonomy clarifies what is being improved; the method, training, data, and application sections explain how those improvements are being produced; and the stress tests expose where frontier systems still fall short. Across spatial puzzles, physical reasoning, multi-turn editing, structured diagrams, coding interfaces, and embodied scenarios, we observed the same pattern: current models are increasingly strong at semantic and visual plausibility, but remain fragile when correctness depends on exact structure, persistent state, verification, or causal grounding.

This positioning also defines the next agenda. Visual generation will need stronger intermediate reasoning, closed-loop agents, tool-augmented rendering, synthetic-data feedback loops, and evaluation protocols that test structural and causal correctness rather than perceptual quality alone. In this sense, the path from atomic mapping to playable world models is not simply a scaling path; it is a shift from image synthesis to verifiable visual reasoning. We hope this roadmap provides a common language for that shift and a practical framework for locating future work along the trajectory from powerful generators to genuinely grounded visual intelligence.