\subsubsection{Robotics Manipulation Trajectories}

Real and simulated robot manipulation trajectories form the core of our
pretraining corpus, accounting for approximately 74.2\% of the total
pretraining mixture. The data spans tabletop manipulation, mobile manipulation,
bimanual tasks, and dexterous hand control across a diverse set of robot
embodiments.

\paragraph{Public datasets.}

To improve data scale, diversity, and cross-embodiment generalization, we train
on a broad mixture of publicly available real-robot datasets, including
RobotSet~\citep{RoboSet}, Galaxea~\citep{galaxea_g0},
AgiBot World~\citep{agibot_world_colosseo}, RoboCOIN~\citep{robocoin},
RoboMIND V1/V2~\citep{robomind_v1,robomind_v2}, RDT-1B~\citep{RDT-1B},
DROID~\citep{droid}, BridgeData V2~\citep{bridgev2},
RH20T~\citep{RH20T}, RT-1~\citep{rt1}, and BC-Z~\citep{bc_z}.
These datasets cover diverse robotic settings, including tabletop
manipulation, mobile manipulation, bimanual manipulation, dexterous hand
control, and in-the-wild task execution, amounting to over 10{,}000 hours of
interaction data across heterogeneous embodiments and task scenarios. This
large-scale mixture provides complementary supervision over robot morphologies,
camera viewpoints, object categories, backgrounds, language instructions, and
action distributions, thereby reducing overfitting to any single embodiment or
environment and improving robustness under distribution shifts.

In addition to real-robot data, we include simulation-based manipulation
trajectories from InternData-A1~\citep{interndata-a1} and
GR00T-X-Embodiment-Sim~\citep{nvidia2025gr00t}, which are generated through motion
planning in diverse virtual environments. These simulated trajectories augment
scene and object diversity, especially for long-tail object configurations and
task layouts. Before training, all real and simulated datasets are curated and
converted into a unified observation-action format while preserving their
original task semantics, enabling scalable training over heterogeneous robot
experience.
\paragraph{Proprietary datasets.}
We supplement with over 1{,}000 hours of in-house collected real-robot trajectories
covering a variety of manipulation tasks and robot platforms, together
contributing approximately 20\% of the total pretraining mixture.
In addition, we generate over 8M synthetic simulation trajectories using our
scalable simulation pipeline illustrated in \cref{subsubsection:SyntheticSimulationData}, contributing a further 3.7\%
of the mixture and substantially expanding the diversity of scenes, objects,
and task configurations beyond what real-robot teleoperation alone can cover.

\paragraph{Embodiment-aware prompt conditioning.}
As described in \cref{sec:embodiment}, each training example is prepended with
a robot-specific prompt specifying the current platform, arm configuration, and
control convention.
Table~\ref{tab:robots} lists the representative embodiments in our pretraining
corpus with their arm configurations and action types.

\begin{table}[h]
\centering
\caption{Representative robot embodiments in the pretraining corpus.
``EEF'' = end-effector pose; ``Joint'' = joint angles;
``$\Delta$'' = delta (relative) commands; ``Abs'' = absolute commands;
``G'' = gripper; ``DH'' = dexterous hand.}
\label{tab:robots}
\setlength{\tabcolsep}{4pt}
\begin{tabular}{llll}
\toprule
\textbf{Robot} & \textbf{Arms} & \textbf{Action type} \\
\midrule
WidowX            & Single & $\Delta$EEF + G \\
Google Robot      & Single & $\Delta$EEF + G \\
Franka Panda      & Single / Dual & $\Delta$EEF + G;\; Abs Joint + G \\
ARX5              & Dual   & $\Delta$EEF + G \\
Fourier GR-1      & Dual   & $\Delta$EEF + G \\
Mobile ALOHA      & Dual   & $\Delta$EEF + G;\; Abs Joint + G \\
AgiBot A2-D       & Dual   & Abs Joint + G;\; Abs Joint + DH \\
Galaxea R1        & Dual   & Abs Joint + G \\
AIRBOT MMK2       & Dual   & Abs Joint + DH \\
TienKung          & Dual   & Abs Joint + G;\; Abs Joint + DH \\
Real Human        & Dual   & $\Delta$EEF (from MANO) \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Action representation.}
Different datasets adopt different action conventions: some provide absolute
end-effector poses in Cartesian space~\citep{xie2026unifyrobotactionscamera}, others provide delta end-effector
commands, and others use absolute or relative joint-space control signals.
We preserve each dataset's original action format rather than converting to a
shared representation, relying on embodiment-aware prompt conditioning
to inform the model of the current control convention.

All action dimensions are normalized per dataset using quantile statistics.
For each action dimension $d$ in dataset $k$, we compute the 1st and 99th
percentiles $q^k_{01}$ and $q^k_{99}$ over all trajectories and apply a
linear mapping:
\begin{equation}
    \tilde{a}_d = 2 \cdot \frac{a_d - q^k_{01}}{q^k_{99} - q^k_{01}} - 1,
\end{equation}
clipping the result to $[-1, 1]$. This per-dataset quantile normalization
removes scale differences across embodiments and action spaces while preserving
relative motion structure within each source.

\paragraph{Language instructions.}
Language instructions are sourced from a combination of original dataset
annotations and model-generated captions provided alongside the data.
All instructions undergo a quality filtering and consistency checking pipeline:
trajectories whose language annotations are inconsistent with the observed
motion are discarded, ensuring reliable language--action alignment across
diverse sources.
Beyond action-labelled data, we incorporate fine-grained embodied video--caption
pairs (described in \cref{sec:vl-data}) as auxiliary vision-language supervision.
This dense language signal bridges the gap between coarse task labels and rich
semantic descriptions of robot behaviour, enabling the model to generalise to a
broader range of natural-language instructions at inference time.

\paragraph{Camera view representation.}
Many robot datasets provide observations from multiple camera viewpoints
simultaneously, \eg a head-mounted ego-centric camera alongside one or two
wrist-mounted cameras.
To make the view provenance explicit to the model, we wrap each image in the
token stream with a pair of view-specific boundary tokens:
\begin{center}
\texttt{<|}\textit{tag}\texttt{\_start|>}\ \textlangle image\textrangle\ \texttt{<|}\textit{tag}\texttt{\_end|>}
\end{center}
where \textit{tag} identifies the camera source, \eg \texttt{ego},
\texttt{cam\_left\_wrist}, or \texttt{cam\_right\_wrist}.
The explicit view labels allow the VLM backbone to form view-aware
representations and let the action expert attend selectively to each
viewpoint when producing actions, without requiring any architectural
changes or additional input channels.

\paragraph{Data cleaning.}
We apply a multi-stage filtering pipeline to remove trajectories with corrupted
or missing frames, near-zero-variance action sequences indicative of static
recordings, and episodes with anomalous lengths. For datasets without explicit
action labels, pseudo-actions are recovered via finite differences on
proprioceptive state sequences.
