\section{Related Work}
\label{sec:related-work}


\paragraph{Benchmarks for Agent Memory.} 
Recent benchmarks have advanced the evaluation of long-term memory in LLM agents. LoCoMo and LongMemEval evaluate long-term conversational recall, multi-session and temporal reasoning, and knowledge updates \citep{maharana-etal-2024-evaluating,wu2025longmemevalbenchmarkingchatassistants}. MemoryAgentBench further studies retrieval, test-time learning, long-range understanding, and selective forgetting in incremental multi-turn interactions \citep{hu2025evaluating}. Beyond post hoc recall, another line evaluates whether memory improves realistic agent behavior. RealMem studies long-term project-oriented interaction \citep{bian2026realmembenchmarkingllmsrealworld}, EverMemBench extends memory evaluation to multi-party and multi-group collaboration \citep{hu2026evaluatinglonghorizonmemorymultiparty}, and MemoryArena and Mem2ActBench examine memory use in multi-session tasks and action execution \citep{he2026memoryarenabenchmarkingagentmemory,shen2026mem2actbenchbenchmarkevaluatinglongterm}. These benchmarks primarily assess memory capability or utility. \gatemem{} targets a complementary dimension, memory governance, asking whether stored information is accessed, withheld, and deleted according to requester-specific boundaries in shared-memory environments.

\paragraph{Benchmarks for Agent Security.} 
Agent-security benchmarks examine risks at multiple levels. Model-level benchmarks such as AdvBench and HarmBench evaluate harmful generation, jailbreak robustness, and refusal behavior \citep{chen2022should,mazeika2024harmbenchstandardizedevaluationframework}. Tool-level benchmarks such as InjecAgent, AgentDojo, and Agent Security Bench study prompt injection, unsafe tool use, and compromised action execution in tool-using agents \citep{zhan2024injecagent,debenedetti2024agentdojo,zhang2025agent}. Persistent memory introduces a distinct attack surface, where sensitive information may be retained, disclosed to the wrong requester, or reconstructed after deletion. Recent work studies contextual integrity, over-persistent memory, and privacy leakage in memory-augmented or multi-agent systems \citep{mireshghallah2025cimemories,pulipaka2026persistbenchlongtermmemoriesforgotten,yagoubi2026agentleakfullstackbenchmarkprivacy}. However, these efforts do not jointly evaluate utility, access control, and active forgetting in multi-principal shared-memory agents. \gatemem{} fills this gap with a unified benchmark protocol for memory and security-oriented governance evaluation.