\section{Baseline Details}
\label{app:baselines-details}

\begin{table}[t]
\centering
\caption{
Comparison of baseline memory mechanisms evaluated in GateMem.
Baselines are grouped by family and compared by memory representation, update mechanism, and query-time access pattern.
}
\label{tab:baseline_methods}
\vspace{-3pt}

\small
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.12}
\begin{tabularx}{\linewidth}{@{}p{2.25cm}YYY@{}}
\toprule
\textbf{Method} & \textbf{Memory Representation} & \textbf{Update Mechanism} & \textbf{Query-time Access} \\
\midrule

\rowcolor{ctxbg}
\multicolumn{4}{c}{\textcolor{gray}{\textsc{Context-based baseline}}} \\

\textbf{Long-Context}
& Raw dialogue history
& Append by turn
& Direct context read \\

\addlinespace[2pt]

\rowcolor{retbg}
\multicolumn{4}{c}{\textcolor{gray}{\textsc{Retrieval-based baselines}}} \\

\textbf{RAG-Naive}
& Chunked text memory
& Append by chunk
& Similarity-based top-$k$ retrieval \\

\textbf{RAG-Policy}
& Chunked text with access metadata
& Append with record and scope metadata
& Metadata-filtered retrieval \\

\addlinespace[2pt]

\rowcolor{agentbg}
\multicolumn{4}{c}{\textcolor{gray}{\textsc{Agentic memory baselines}}} \\

\textbf{A-Mem}~\citep{xu2026mem}
& Linked memory notes
& Note construction with link evolution
& Graph retrieval and reranking \\

\textbf{Mem0}~\citep{chhikara2025mem0}
& Persistent memory entries
& Extract and consolidate memories
& Persistent memory search \\

\textbf{REMem-I}~\citep{shu2026remem}
& Hybrid episodic graph of gists and facts
& Index episodes into memory graph
& Iterative tool-based retrieval and reasoning \\

\textbf{REMem-S}~\citep{shu2026remem}
& Hybrid episodic graph of gists and facts
& Index episodes into memory graph
& Single-step retrieval and generation \\

\bottomrule
\end{tabularx}
\end{table}

The implementation supports seven baseline configurations. These include three custom baselines---\textsc{Long-Context}, \textsc{RAG-Naive}, and \textsc{RAG-Policy}---and four configurations derived from prior agentic memory systems: \textsc{A-Mem}~\citep{xu2026mem}, \textsc{Mem0}~\citep{chhikara2025mem0}, \textsc{ReMem-I}, and \textsc{ReMem-S}~\citep{shu2026remem}. The main result tables report the subset of baseline runs completed for the experimental configurations evaluated in this paper.

\paragraph{\textsc{Long-Context}.}
\textsc{Long-Context} places the full available interaction history in the model prompt and does not construct an explicit external memory store. This baseline provides a strong utility-oriented reference when prompt length is not tightly constrained, but it is not governance-aware and may expose information that should no longer be accessible.

\paragraph{\textsc{RAG-Naive}.}
\textsc{RAG-Naive} retrieves the top-$k$ relevant snippets from prior turns and provides them to the model as context. It follows the standard retrieval-augmented pattern of reducing prompt length through retrieved external context \citep{lewis2020retrieval}. It does not apply a dedicated access-control or deletion-aware policy layer, and therefore relies on the base model to decide whether retrieved information is safe to use.

\paragraph{\textsc{RAG-Policy}.}
\textsc{RAG-Policy} augments retrieval with policy-oriented metadata, including requester identity, domain rules, and access constraints. At query time, these metadata are used to make access-relevant information salient to the model. This baseline tests whether lightweight policy-aware retrieval and prompting are sufficient for shared-memory governance, without introducing a dedicated governance mechanism.

\paragraph{\textsc{A-Mem}.}
\textsc{A-Mem} is a Zettelkasten-inspired agentic memory system that stores interactions as structured notes with LLM-generated context, keywords, tags, embeddings, and links \citep{xu2026mem}. New memories can trigger dynamic link generation and memory evolution, making it a useful baseline for testing whether adaptive memory organization improves governance-sensitive long-horizon tasks.

\paragraph{\textsc{Mem0}.}
\textsc{Mem0} is a production-oriented long-term memory system that incrementally extracts salient facts from recent interactions and updates an external memory store through add, update, delete, or no-op operations \citep{chhikara2025mem0}. We include it as a practical persistent-memory baseline designed to balance long-term conversational continuity with retrieval efficiency.

\paragraph{\textsc{ReMem}.}
\textsc{ReMem} is an episodic memory framework that indexes interaction histories as a hybrid graph of time-aware gists and time-scoped facts, then reasons over this graph using retrieval and graph-exploration tools \citep{shu2026remem}. We evaluate two variants: \textsc{ReMem-I}, which performs iterative tool-based retrieval and reasoning, and \textsc{ReMem-S}, which uses a single-step embedding-retrieval-and-generation procedure with lower inference overhead.