
\section{Discussion}\label{sec12}
 
 

 
 

 

 
 
 

 
 
 

 
 
 
 
 

 
 
 

In this work, we introduced \projName{}, a multimodal scientific foundation model for native structural reasoning across proteins, small molecules and inorganic crystals.
The central premise is that structure--property relationships cannot be fully addressed by treating scientific entities as text strings, low-dimensional descriptors or black-box inputs to property predictors.
\projName{} instead represents structures as primary objects of inference through a unified structure-aware vocabulary that discretizes three-dimensional coordinates, molecular topologies and crystallographic lattices into domain-native tokens, and integrates them with language instructions in a single autoregressive model.
This design allows residues, molecular fragments, conformational cues and crystal descriptors to function as addressable evidence units within generated reasoning trajectories, rather than as peripheral annotations appended to text.


Our results support native structural reasoning as a model-level capability for scientific AI.
Rather than using structure only as an input format or auxiliary descriptor, \projName{} makes residues, molecular fragments, conformational cues and crystal descriptors available as evidence within generated reasoning trajectories.
This shift is important because structure--property analysis requires not only accurate outputs, but intermediate claims that can be inspected against the physical object being studied.

The strongest evidence for this capability comes from regimes in which superficial similarity is insufficient.
 
For Cellular Component annotation of low-homology and orphan-like proteins, \projName{} improved $F_{\max}$ from 0.42 to 0.55, with its attention enriched at contact-defined DNA-binding residues and protein--DNA interfaces. 
In retrosynthesis, it increased single-step accuracy from 0.63 to 0.72 while identifying reaction-relevant fragments, strategic bond disconnections and chemically valid precursors.
In crystalline materials, its reasoning trajectories invoked symmetry-distinct sites, coordination environments and periodic bonding motifs when linking structure to electronic and thermodynamic behaviour.
These behaviours indicate that structural organization functions as the working substrate of inference, not merely as a representation used to produce a final score.

This distinction clarifies the contribution of \projName{} relative to existing scientific AI systems.
A Foldseek-derived structural fragment can ground a fold- or function-level hypothesis; a molecular substructure can justify a retrosynthetic disconnection; a conformational cue can support a pharmacophore-level similarity judgement; and a crystal-connectivity token can support an inference about stability or electronic response.
Because such intermediate claims are expressed in the reasoning trajectory, they can be checked against the input structure rather than accepted as opaque post hoc rationales.
This helps bridge two previously separated capabilities: domain-native structural representation and explicit scientific reasoning.
The double-blind expert evaluation, in which experts preferred or tied \projName{} over a frontier large language model in 98\% of cases, further suggests that this inspectability is practically useful for scientific assessment.


 
 
 
 
 
 

Overall, \projName{} demonstrates that a single autoregressive foundation model can unify sequence, structure, and natural-language reasoning across major scientific modalities while achieving specialist-level accuracy and interpretable scientific explanation. 
This provides a route toward scientific AI systems that move beyond prediction toward reasoning over the physical structures from which functions, properties, and mechanisms emerge.