\abstract{
Optimizer selection for large-scale model training has become a system-level
design decision constrained jointly by compute, memory, tuning budget, and task
diversity, yet the landscape of over one hundred methods remains fragmented.
We therefore present \textbf{OmniOpt},
a unified survey and benchmark cookbook of optimizers for the research community.
OmniOpt rests on four coupled components. First, we treat every optimizer update
as a structured transformation through a five-stage meta-pipeline, and show that
most methods engage only one or two of these stages. Second, we use
norm-constrained linear minimization oracles (LMOs) to unify different
optimizers. Third, these two views ground a dual-dimension taxonomy, one
dimension assigning each method to a mechanism family and the other recording the
measurable training objectives it aims to improve. Fourth, and at the core of
this paper, we instantiate the full taxonomy in a unified cross-domain benchmark
spanning representative optimizers, model scales, and training regimes from
language model pretraining to image classification, systematically analyzing each
method family across multiple effect objectives and laying out their trade-offs.
OmniOpt thus supplies the research community with an operational
coordinate system for selecting optimizers under explicit mechanism and
objective assumptions, and charts a direction for the future development of the
optimizer community.
%
% Sentences condensed out of the abstract (kept for reference, not deleted):
% 1. Meta-pipeline stage list: "...through a five-stage meta-pipeline covering
%    parameter routing, gradient transformation, state evolution, update
%    reconstruction, and finalization..."
% 2. LMO direction examples: "Norm-constrained linear minimization oracles then
%    provide a geometric primitive that places sign, spectral-orthogonalized,
%    Kronecker-factored preconditioned, and projected directions within one
%    four-axis decomposition."
% 3. Taxonomy detail: "a methodological dimension assigning each of five mechanism
%    families a unique pipeline locus, and an effect-objective dimension recording
%    six measurable training properties."
% 4. Benchmark scale numbers: "spanning language model pretraining at 60M to 1B
%    parameters across four architectures and context lengths from 256 to 32k
%    tokens, together with image classification on CIFAR, evaluating
%    representative optimizers on convergence, cost, memory, stability,
%    robustness, and generalization."
% 5. Detailed findings (moved to Introduction contribution 4): "Key findings
%    include that no single optimizer dominates the multi-objective frontier
%    across domains, that aggressive state compression excelling under short
%    context degrades sharply when input complexity grows, that structured-matrix
%    methods transfer most stably across architectures and tasks but at
%    substantial per-step cost, and that optimizer rankings exhibit systematic
%    domain-dependent crossings unpredictable from any single-setting comparison."
}
