\subsection{Experimental Setup}
\label{sec:benchmark-setup}

\paragraph{Optimizer selection and taxonomy coverage.}
We instantiate the taxonomy by selecting 24 representative optimizers spanning
all five methodological families (T1--T5) and the major regions of the
four-axis coordinate space. Their four-axis coordinates are listed in
Table~\ref{tab:four-axis-instantiation} (Section~\ref{sec:four-axis-decomposition}),
which places each optimizer in its family and records its update domain, state
estimator, geometry-and-precondition operator, and finalization. This makes
explicit that our optimizer set is a systematic coverage of the classification
space.

% \input{tables/tab_opt_coverage} % merged into Table~\ref{tab:four-axis-instantiation}

\paragraph{Training protocol.}
Our evaluation comprises two complementary stages.
Stage~1 sweeps all optimizers on C4~\cite{raffel2020exploring} under the
LLaMA architecture at sequence length 256, across four model scales---60M,
130M, 350M, and 1B (trained for 10k, 20k, 60k, and 100k steps,
respectively)---using final C4 validation perplexity as the quality metric.
Stage~2 transfers the stronger Stage-1
optimizers to the higher-quality FineWeb-Edu
corpus~\cite{penedo2024finewebdatasetsdecantingweb,karpathy2024finewebedu100b}
with long 32k sequences,
at 340M and 1B ($\sim$30k steps), and across four
architectures---Transformer++~\cite{yang2023gated}
(standard attention) and three linear-attention variants (Gated
DeltaNet~\cite{yang2025gated},
DeltaNet~\cite{yang2024parallelizing}, GLA~\cite{yang2023gated})---to probe portability across dataset, sequence length, and
architecture. We weight the 350M and 1B results most heavily, as small-scale
rankings are unreliable predictors of behavior at scale. To ensure that
observed differences genuinely stem from the optimizers themselves, we adopt a
strict \emph{controlled-variable} principle: within each
stage, only optimizer-related hyperparameters are tuned (\texttt{betas},
\texttt{eps}, \texttt{lr}, and method-specific knobs such as APOLLO's projection
rank and projection interval~$T$), while all architectural, data, and schedule
settings are held fixed across optimizers.
 
\paragraph{Stage-dependent regularization.}
The two S5 (Update Finalization) operations of the meta-pipeline---decoupled
weight decay and gradient clipping---are treated differently across the two
stages, by design. In Stage~1 both are \emph{disabled}, so that the comparison
isolates the S2/S3 machinery that distinguishes one family from another, and
separates it from a generic finalization-stage regularizer. In Stage~2
both are \emph{enabled} identically for all optimizers, to measure
generalization under a realistic, production-style recipe. This also lets us
distinguish whether an optimizer's advantage is intrinsic to its S2/S3 design
(surviving into the regularized regime) or merely an artifact of the
unregularized Stage-1 setting.
 



\paragraph{Evaluation metrics (mapped to O1--O6).}
\textbf{O1 (optimization quality):} C4 validation PPL in Stage~1, while
Stage~2 uses WikiText~\cite{merity2016pointer} test PPL and downstream Commonsense Reasoning average
accuracy (CS Avg.). CS Avg. is evaluated with lm-eval-harness~\cite{eval-harness} and
computed over ten tasks/splits: ARC-Easy, ARC-Challenge, HellaSwag,
WinoGrande, PIQA, OpenBookQA, BoolQ, COPA, LAMBADA-OpenAI, and
SciQ~\cite{clark2018think,zellers2019hellaswag,sakaguchi2021winogrande,bisk2020piqa,mihaylov2018can,clark2019boolq,roemmele2011choice,paperno-EtAl:2016:P16-1,welbl2017crowdsourcing}. \textbf{O2 (per-step compute):} isolated wall-clock
optimizer runtime per step ($T$, ms). \textbf{O3 (memory):} isolated
optimizer-state memory (Mem, GB), excluding model parameters, activations, and
gradients. \textbf{O4 (training stability):} long-context gradient-norm
dynamics, mainly post-warmup GNormCV, with spike and NaN/Inf diagnostics.
\textbf{O5 (hyperparameter robustness):} learning-rate perturbation around the
tuned learning rate, using $1/5$ and $5\times$ perturbations. \textbf{O6
(generalization):} cross-scenario transfer across dataset, context length,
scale, and architecture, summarized by rank stability and sequence-length
sensitivity. These measurements support the optimizer-level and family-level
O1--O6 assessment below.