\section{Benchmark Study}
\label{sec:benchmark}

This section presents a mechanism-aware benchmark of representative optimizers
for LLM pretraining. Rather than treating the benchmark as a standalone
leaderboard, we use the framework developed in the preceding sections---the
universal meta-pipeline, the LMO-driven four-axis decomposition, and the
dual-dimension taxonomy---to organize and interpret the empirical results.
Each experiment is linked to a concrete framework question: which optimizer
families perform well under controlled short-context pretraining, which
quality--efficiency trade-offs emerge across optimization quality, runtime,
and memory cost (O1--O3), whether short-context advantages transfer to
long-context and cross-architecture settings (O6), and how training dynamics
and hyperparameter perturbations reveal additional stability and robustness
properties (O4--O5). We further use Muon as a representative
matrix-structured optimizer to study how its sub-operations compose across
scale and architecture.


Compared with existing LLM-pretraining optimizer benchmarks
(e.g., \emph{Fantastic Pretraining Optimizers}~\cite{wen2025fantastic},
\emph{Benchmarking Optimizers for LLM Pretraining}~\cite{semenov2025benchmarking}),
our benchmark emphasizes three aspects. First, results are analyzed not only at
the optimizer level, but also at the family and axis levels, so that empirical
patterns can be mapped back to the methodological taxonomy. Second, we evaluate
optimizers under a multi-objective protocol: final quality is measured by
perplexity, Stage~2 additionally reports downstream Commonsense Reasoning
average accuracy (CS Avg.), efficiency is measured by per-step runtime and
optimizer-state memory, stability is analyzed through gradient-norm dynamics,
robustness is probed by learning-rate perturbation, and generalization is
evaluated by cross-scenario transfer. Third, we conduct a mechanistic ablation
of Muon to examine how matrix-structured optimizer
components---orthogonalization, scaling, momentum placement, and
finalization---interact. In this way, the benchmark both evaluates practical
optimizer choices and provides empirical feedback on the proposed framework.

\input{Chapters/6_1_setups}
\input{Chapters/6_2_results_and_analysis}
\input{Chapters/6_3_mechanistic_ablation_of_muon}