\section{Discussion}
\label{sec:discussion}

This paper was organized around three questions about where an optimizer acts,
why its update takes a particular geometric form, and what training objective it
is meant to improve. We answered them through three layers. The first is the
meta-pipeline of Section~\ref{sec:meta-pipeline}, the second is the four-axis
decomposition of Section~\ref{sec:four-axis-decomposition}, and the third is
the dual-dimension taxonomy of Section~\ref{sec:taxonomy}. All three are
grounded in the benchmark of Section~\ref{sec:benchmark}. This section steps back from those instruments and asks what they collectively do and
do not establish. We first delimit the scope
of the present study (Section~\ref{sec:discussion-limitations}), then use the
framework to extract technique-level lessons about benefit, composition, and
conflict (Section~\ref{sec:discussion-compositions}), and finally
distill the open directions that follow (Section~\ref{sec:discussion-open-problems}).

% =====================================================================
\subsection{Limitations}
\label{sec:discussion-limitations}

The introduction framed optimizer selection as a jointly empirical,
mathematical, and engineering problem. A modern training run is constrained by more than whether an update reduces the
loss. Compute, optimizer-state memory, tuning budget, and task diversity all
bound it. The literature is also hard to navigate because mechanisms are named
locally, empirical claims are protocol-sensitive, and mathematical accounts are
fragmented
(Section~\ref{sec:intro-background}). Our framework targets exactly these gaps,
and the benchmark supplies controlled evidence for them. The same framing,
however, makes the boundaries of the current study explicit.

\paragraph{Empirical scope.}
The empirical study substantiates the framework primarily on language-model
pretraining, covering four model scales from 60M to 1B parameters, four
architectures spanning standard and linear attention, and context lengths from
256 to 32k tokens. In addition, the benchmark includes a CIFAR100 vision-backbone study across
ResNet50, DeiT-S, and CAFormer-S12. This study tests whether optimizer behavior
transfers across architecture families within image classification.
Thus, claims about \emph{domain-dependent crossings} are supported by the
language-modeling regime and by a targeted CIFAR100 cross-architecture check.
In future work we will add validation on ImageNet-scale vision and additional
modalities, extending these into a full cross-modality conclusion.

\paragraph{Mechanism attribution.}
Every effect-target assessment in Section~\ref{sec:benchmark} is deliberately a
mechanism-informed prior. The mechanistic explanations we give are also \emph{qualitative}. They cover
APOLLO's rank-bounded compression, Muon's interaction with architecture topology,
and MARS's variance reduction on the state estimator while preserving base
geometry. We can say \emph{why} an optimizer plausibly behaves as observed. In future work
we will develop metrics that quantify these explanations, measuring quantities
such as the effective
rank of the gradient as context grows, the staleness of a preconditioning basis,
and the fraction of a technique's benefit that is intrinsic to the algorithm
versus
induced by the tuning protocol. Such interpretability metrics must be designed
and validated carefully, and we will provide a quantitative interpretability
evaluation in a subsequent version of this work. For future work, we plan to add
heavy-tailed self-regularization indices as training-quality metrics and to
evaluate optimizers on downstream commonsense reasoning benchmarks.

\paragraph{Protocol dependence.}
Our conclusions are tied to a strict controlled-variable protocol, in which only
optimizer-related hyperparameters are tuned while architecture, data, and
schedule are held fixed. This isolates optimizer mechanism, but it also makes
several findings protocol-relative. Many apparent
improvements over AdamW shrink or disappear once the baseline is retuned. The
learning-rate robustness objective (O5) is reported as a \emph{local} diagnostic.
The long-context comparison is
matched on dataset and architecture family but not on token budget, so it
certifies a degradation \emph{trend}.
A few memory-efficient methods are drawn from recent preprints whose broader
LLM-scale behavior is still being established. Their results should therefore
be interpreted under the present protocol and remain preliminary for the
method class.

\paragraph{Method coverage.}
We instantiate 24 representative optimizers chosen to cover the major regions of
the four-axis space (Table~\ref{tab:four-axis-instantiation}), but the surveyed literature
exceeds one hundred methods, and not every axis combination is realized by an
existing optimizer or exercised by our benchmark. Consequently, family-level statements should be read as conclusions about the
\emph{tested} instances and protocol. This caution is especially important for the claim that
curvature-aware and geometry-regularized methods (T5) are competitive only in
limited regimes under the current benchmark. These boundaries motivate the
technique-level reading and the open directions that follow.

% =====================================================================
\subsection{Technique-Level Lessons}
\label{sec:discussion-compositions}

A central payoff of the framework is that it shifts attention from
\emph{which optimizer is the best} to a more tractable set of technique-level
questions. Which mechanisms carry the gains? Which ones can be stacked? Which
ones interfere?
The governing principle, established by the meta-pipeline and the cross-dimension
analysis (Section~\ref{sec:taxonomy-cross}), is one of \emph{locality}.
Mechanisms that act on different pipeline stages or different axes tend to
compose, whereas mechanisms that contend for the same slot require an explicit
ordering and may conflict.

\paragraph{Benefit carriers.}
Under the tested protocol, the strongest gains are associated with
\emph{geometry-sensitive direction maps} (Axis~III) and with the
\emph{structured state} that feeds them (Axis~II). Scalar control tweaks alone
contribute little. The Muon ablation makes this concrete. Removing the
diagonal second moment alone collapses quality (PPL $17.78\!\rightarrow\!70.74$),
whereas replacing it with Newton--Schulz orthogonalization recovers and then
surpasses the AdamW baseline ($\rightarrow\!16.86$). This identifies spectral
orthogonalization as the indispensable core.
Kronecker-factored preconditioning is another strong carrier. SOAP attains
the strongest transferable quality profile in Stage~2, with a mean rank near the
top across architectures. This comes at the cost of the highest per-step runtime
and the highest memory use.
Variance reduction in the style of MARS is a genuine, reusable improvement to the
state estimator. It improves the gradient signal while preserving the base update
geometry, which makes MARS-AdamW the most stable AdamW-style enhancement in
Stage~2. The gain comes with extra state and computation, so MARS is better read
as a quality-oriented extension. Finally, simple
memory techniques that \emph{preserve} geometry, such as row/column
factorization (AdaFactor) and low-bit state (8-bit Adam), deliver
reliable memory savings without changing the descent direction, which is why they
are safe defaults when memory is the binding constraint.

\paragraph{Limited returns.}
Conversely, several techniques are marginal in the tested regime. Small
element-wise refinements of the AdamW template land in the weakest tier for
tuned perplexity. These include early-variance rectification (RAdam),
Nesterov lookahead (NAdam), and prediction-error variance (AdaBelief). Automatic
step-size tuning (Prodigy) lands in the same tier. This is consistent with
their stated targets of robustness or convenience.
Diagonal curvature proxies and geometric wrappers (Sophia, LAMB) are likewise
weak or only situationally competitive here. They add cost through an extra
Hessian-vector product, clipping or trust-ratio parameters, or a second gradient
evaluation, yet they do not deliver a consistent quality return under matched
tuning. These mechanisms still have value. Their advantage is simply
conditional, and the benchmark is meant to expose that condition.

\paragraph{Compatible compositions.}
Locality predicts, and the survey confirms, several orthogonal compositions.
(i)~A variance-reduced signal (Axis~II) can feed essentially any base direction,
because it is largely independent of basis and curvature choice. This is exactly
the pattern behind the MARS-AdamW, MARS-Lion, and MARS-Shampoo constructions.
(ii)~Low-rank projection (T2.3, acting on the signal subspace across S2 and
S4) and state quantization (T4.2, acting on the state representation across S3
and S4) modify disjoint objects and combine cleanly, as in Q-GaLore.
(iii)~Post-update filters (T5.3) wrap a base optimizer because they act
\emph{after} the core direction is chosen. This yields the cautious variants of
AdamW and Lion from a single mechanism. (iv)~Layer-wise trust ratios (T5.4)
act at finalization (S5) and can therefore wrap an element-wise adaptive update
without touching its direction. (v)~Even within a single method, the Muon
ablation shows that smaller gain operations can act through partially separable
mechanisms. In Muon, symmetric
two-way learning-rate scaling and post-orthogonalization Nesterov momentum stack
across model scales on the standard Transformer.

\paragraph{Conflicts and limits.}
The same locality principle predicts where stacking fails. (i)~Two strong S2
matrix constraints, spectral orthogonalization and low-rank projection, contend
for the same slot. Combining them is therefore not free. The
result depends on whether projection precedes orthogonalization or vice versa,
and the two orders yield different directions and state footprints.
(ii)~Fused backprop-update streaming (LOMO-style, T4.4) shortens the gradient's
lifetime and thereby conflicts with
any mechanism that needs complete gradient statistics, including global
clipping, delayed low-rank basis refresh, SAM-style second gradients, or matrix
preconditioners.
(iii)~Some compositions are valid but impractical. Stacking SAM's double
forward-backward on top of expensive Kronecker-factored preconditioning multiplies cost
for little expected return. (iv)~Most importantly, two conflicts are
\emph{empirical and quantitative}. Aggressive state compression is
\emph{rank-bounded}. APOLLO is the single best optimizer at 256 tokens but the
worst at 32k. Its degradation is
$+21.87$ PPL, which is roughly three times the AdamW baseline degradation. The
reason is that a fixed low-dimensional projection discards proportionally more
information as the gradient's effective rank rises
with context (Section~\ref{sec:benchmark-mem-eff-sensitivity},
Table~\ref{tab:seqlen_numbers}). Muon's spectral geometry is also
\emph{architecture-conditional}. The gain
operations that stack on the standard Transformer no longer improve beyond the best single gain on Gated DeltaNet.
The operator order matters as well. Accumulating momentum in the
orthogonalized space hurts, and so does scaling the
learning rate before orthogonalization
(Section~\ref{sec:benchmark-muon-ablation},
Table~\ref{tab:muon_cross_validation}).

\paragraph{Design takeaway.}
Read together, these results recommend treating the four-axis coordinate of a
method as a practical compatibility test. Techniques that occupy different axes
or pipeline stages are good candidates for composition. Techniques that share a
slot demand an explicit ordering. Benefits that depend on gradient rank or
architecture topology, especially compression and spectral geometry, must be
validated in the target regime, and their transfer must be confirmed.
The implication is that optimizer geometry must be matched to model structure
and training dynamics. This is the sense in which no single optimizer dominates
the multi-objective frontier. The unit of design is a \emph{composition} chosen
against a specific budget, context length, and architecture.

% =====================================================================
\subsection{Open Problems and Future Directions}
\label{sec:discussion-open-problems}

The limitations and the technique-level reading point to a coherent agenda for
LLM optimizer design and evaluation.

\paragraph{Diagnostics and adaptive compression.}
The most direct gap is the absence of measures that turn our qualitative
mechanism attributions into testable quantities. Useful examples include an
effective-rank estimate that predicts when compression will become lossy, a
basis-staleness measure that predicts when a Kronecker or low-rank preconditioner
should refresh, and an intrinsic-versus-protocol decomposition that separates
genuine algorithmic gains from tuning artifacts. Such metrics would convert the
present coordinate system into a predictive diagnostic, and they point to a
concrete algorithmic target. The APOLLO collapse shows the payoff of an
effective-rank estimate, namely compression schemes whose retained rank or projection
adapts to the rising effective rank of the gradient as context length grows, so
that short-context memory wins do not become long-context quality failures.

\paragraph{Architecture-aware geometry and transfer.}
Rankings cross across dataset, context length, and architecture in ways that no
single-setting comparison anticipates, and the two clearest mechanisms behind
those crossings, compression versus gradient rank and spectral geometry versus
attention topology, are understood only post hoc. This motivates a predictive
account of \emph{when} a given geometry transfers, tied to measurable properties
of the model and data, which would let practitioners select an optimizer without
an exhaustive sweep. The same handle suggests a design principle, namely to choose
optimizer geometry from measurable properties of the architecture and the
evolving training dynamics. RMNP~\citep{deng2026rmnp} is one example. Under row-wise
block-diagonal dominance in the layer-wise Hessian and diagonal dominance in the
momentum Gram matrix, its row normalization is a structure-aware specialization
of Muon-style preconditioning,
and future optimizers should co-design the preconditioner with the architecture
and the statistics that emerge during training.

\paragraph{Multi-objective selection.}
Because the O1--O6 objectives genuinely trade off and no method is uniformly
best, optimizer choice is a multi-objective problem that the field still treats
as a single-number leaderboard. Cost-aware comparison (token efficiency versus
wall-clock and memory), Pareto-aware selection, and optimizer designs that target
a specified region of the frontier are all open.

\paragraph{Compositional search.}
The framework identifies compatible and conflicting compositions by hand.
Automating this would mean searching axis-compatible combinations, resolving
same-slot ordering, and co-designing variance reduction, direction geometry, and
state compression. It also includes the under-explored question of how to combine
two S2 constraints without sacrificing either.

\paragraph{Cost-effective curvature.}
Methods that use curvature across T2 and T5 show the same cost tension.
Sophia follows a T5 diagonal-curvature route, while Shampoo and SOAP follow T2
Kronecker-basis routes. They deliver either inconsistent gains or strong gains
at prohibitive cost. Cheaper and more stable curvature estimation would help
determine when these mechanisms become default tools and when they remain
specialist choices. The same is true for a clear characterization of the regimes
in which curvature information is worth its overhead.

\paragraph{Matched protocols.}
Finally, the protocol sensitivity documented throughout argues for standardized,
matched-budget evaluation. Useful examples include full learning-rate-transfer
studies, matched-token long-context
comparisons, and reporting conventions that make the controlled-variable
assumptions explicit. This would make it easier to separate intrinsic
algorithmic progress from tuning and protocol effects.
