\begin{abstract}
Reinforcement learning (RL) is becoming increasingly important for post-training large language models (LLMs). 
Previous RL pipelines for LLMs were mostly synchronous and batch-interleaved, which is inefficient for long-horizon agentic tasks. 
Recently, asynchronous RL has emerged as a more efficient alternative by updating the model as rollouts arrive. 
However, existing asynchronous RL systems often emphasize throughput, while leaving training stability and task effectiveness largely underexplored. 
For example, a key challenge is that group-wise sampling in the widely-used GRPO framework does not naturally fit asynchronous agentic training. 
In this paper, we present Single-rollout Asynchronous Optimization (\model{}) to address the stability and off-policy challenges in asynchronous RL. 
To reduce off-policy effects and improve generalization, we replace group-wise sampling with single-rollout sampling, that is, using one rollout per prompt. 
We further improve this single-rollout strategy with practical value-model training designs.
To improve optimization stability, we introduce a strict double-side token-level clipping strategy. 
\model is able to train stably for one thousand steps and consistently outperform GRPO and its variants on agentic coding and reasoning benchmarks, such as SWE-Bench Verified, BeyondAIME, and IMOAnswerBench.  
We also demonstrate that single-rollout RL is particularly effective in a simulated online learning setting, where the model must adapt to changing evolving environments. 
To this end, \model{} is successfully deployed in the agentic RL pipeline for training the open GLM-5.2 model (750B-A40B). 
\end{abstract}


%Reinforcement learning (RL) is becoming increasingly important for large language models (LLMs), yet most LLM-RL pipelines remain synchronous and batch-interleaved, which can be inefficient for long-horizon tasks. Asynchronous RL offers a more efficient alternative by updating on rollouts as they arrive; however, existing systems often emphasize throughput while leaving training stability and task effectiveness underexplored.
%In this paper, we present \model to address the stability and off-policy challenges in asynchronous RL. We introduce a strict token-level clipping strategy to improve optimization stability. To further reduce off-policy effects and improve generalization, we replace group-wise sampling with single-rollout sampling (one rollout per prompt) and boost performance through improved value-model training.
%\model trains stably for a thousand steps and consistently outperforms GRPO and variants in agentic reasoning and coding benchmarks.  We also demonstrate that RL with single-rollout can be uniquely effective in a simulated online learning setting to dynamically adapt to evolving environments.