\section{Related Work}

\textbf{Self-Improvement for LLM and Agent.} A growing line of work studies how language models and agents can improve using their own outputs to self-evolve~\citep{selfimprovement, selftraining}. STaR~\citep{star} bootstraps reasoning by filtering correct rationales for fine-tuning, while Self-Refine~\citep{madaan2023self} revises outputs at inference time through self-generated feedback. Self-Rewarding Language Models~\citep{selfreward} use the model itself as a judge for preference optimization. In agentic settings, Reflexion~\citep{shinn2023reflexion} converts environmental feedback into verbal reflections, and Voyager~\citep{wang2023voyager} accumulates reusable skills through continual exploration. These methods typically refine individual trajectories or rely on the model to judge its own outputs. In contrast, \ours\ treats self-improvement as a structured search problem, systematically discovering high-quality solutions that facilitate model self-improvement.

\textbf{Search in LLM and Agent.} Recent work applies search to both training and inference of LLMs and agents~\citep{efficientevolution, funsearch}. On the training side, tree-structured exploration provides richer supervision and finer-grained credit assignment than standard rollout-based methods. Tree-GRPO~\citep{treegrpo} and TreeRL~\citep{treerl} integrate tree search directly into reinforcement learning, while ReST-MCTS*~\citep{restmcts}, MCTS-DPO~\citep{mctsdpo}, and rStar-Math~\citep{rstarmath} use search to bootstrap higher-quality training data for self-improvement. On the inference side, Tree of Thoughts~\citep{tot}, Graph of Thoughts~\citep{got}, and RAP~\citep{rap} established reasoning as an explicit search problem by expanding chain-of-thought into tree-structured exploration~\citep{abmcts}. More recently, evolutionary approaches such as AlphaEvolve~\citep{alphaevolve}, ShinkaEvolve~\citep{shinkaevolve}, and ThetaEvolve~\citep{thetaevolve} tackle hard and even open problems~\citep{languageevolutionary} by maintaining candidate populations with LLM-driven mutations and external evaluation. However, these methods predominantly rely on tree search and thus struggle to explore beyond the policy's own distribution.

\textbf{Classical Search Methods.} The recent surge of search-based methods in LLMs and agents draws on a long history of classical search. In symbolic planning and graph search, algorithms such as A*~\citep{astar} and bidirectional search introduced core principles including heuristic guidance, frontier expansion, and search-space reduction. Branch-and-bound methods~\citep{bbsurvey}, which prune subtrees whose bounds certify suboptimality, offer a natural analogy for modern methods that use verifiers to discard unpromising candidates early. Evolutionary methods, including genetic search~\citep{genesearch,geneprogram} and differential evolution~\citep{differentialevolution}, optimize by maintaining and iteratively refining candidate populations. 