\section{Implementation Details}

\subsection{Evaluation Metrics}
We follow the DA3~\cite{depthanything3} evaluation protocol for pose estimation and 3D reconstruction. For pose estimation, we report the area under the curve (AUC) of Relative Rotation Accuracy (RRA) and Relative Translation Accuracy (RTA) at thresholds of 5 and 30, reflecting strict and relaxed evaluation criteria. For 3D reconstruction, we report accuracy and completeness, and compute the Chamfer Distance (CD) as their average. We additionally report the F1-score under a distance threshold to assess overall reconstruction quality. For image restoration, we use PSNR and LPIPS to evaluate pixel-level fidelity and perceptual quality, respectively. PSNR measures reconstruction accuracy against the ground-truth image, while LPIPS assesses perceptual similarity in a deep feature space. Together, they evaluate both low-level fidelity and perceptual quality of the restored images. Higher PSNR and lower LPIPS indicate better restoration performance.



\subsection{RGB Image Decoder}
Our GARD framework jointly restores 3D scene geometry and high-quality RGB images by propagating the restored representations to their corresponding decoders. The underlying feed-forward 3D reconstruction model already incorporates a 3D geometry decoder $\mathcal{D}$ for scene geometry prediction. To reconstruct restored RGB images, we adopt the ViT-based RGB decoder $\mathcal{D}_{\text{rgb}}$ introduced in GLD~\cite{jang2026gld}. As the original RGB decoder was designed exclusively for the DA3-BASE backbone, we further adapt and fine-tune it for the DA3-GIANT backbone, which serves as the primary feed-forward 3D reconstruction model in our framework. Concretely, we initialize the RGB decoder using the released DA3-BASE checkpoint and introduce an additional DA3 linear projection adapter to bridge the feature dimensional discrepancy between the DA3-BASE and DA3-GIANT representations. The restored multi-scale features extracted from the four selected feature levels $\mathcal{M}=\{20, 28, 34, 40\}$ are concatenated and subsequently fed into the RGB decoder to reconstruct the restored RGB images. We refer readers to the original GLD work~\cite{jang2026gld} for additional architectural and implementation details.



\subsection{Training Details}

\paragraph{Training dataset.}
We train our model on two photo-realistic synthetic datasets: Hypersim~\cite{roberts2021hypersim} and TartanAir~\cite{wang2020tartanair}, which together cover diverse 3D environments. Hypersim provides high-fidelity indoor scenes with realistic lighting and materials, making it well-suited for learning appearance-aware geometric representations. In contrast, TartanAir includes both indoor and outdoor environments with dynamic camera trajectories, emphasizing temporal consistency and large viewpoint changes for multi-view geometry learning. By combining them, we leverage complementary strengths in visual realism and motion diversity, enabling more robust and generalizable geometric representations. Degraded images are generated by applying motion blur kernels using the codebase\footnote{https://github.com/LeviBorodenko/motionblur}. During training, we resize the input frames so that the longest side is 504 pixels and randomly select up to 4 views per iteration to train the model in various multi-view settings. 

\paragraph{Training hyperparameters.}
We train the GARD denoiser for 10 epochs with a global batch size of 8 on NVIDIA H100 GPUs, using the AdamW optimizer with a learning rate of $2 \times 10^{-4}$, $(\beta_1, \beta_2) = (0.9, 0.95)$, and no weight decay. All parameters are updated in full precision (fp32). We apply exponential moving average (EMA) with a decay rate of 0.9995 to stabilize training. The flow-matching objective uses a linear path with velocity prediction and logit-normal time sampling. Gradients are clipped to a maximum norm of 1.0.


\paragraph{Attention alignment.}
Fig.~\ref{fig:suppl_attn_target} visualizes the target correspondence map $\mathbf{A}^*$ under different temperature values $T$. To construct $\mathbf{A}^*$, we first estimate depth and camera pose from the clean high-quality (HQ) input views and unproject each pixel into a shared 3D world coordinate system. The resulting point cloud is spatially pooled to the ViT patch resolution (patch size 14) to align with the spatial token grid of the GARD denoiser $\mathcal{S}_{\theta}(\cdot)$. We then compute pairwise negative $\ell_2$ distances across all $VN$ patch tokens and apply a temperature-scaled softmax to obtain $\mathbf{A}^*$. Lower temperatures produce sharper and more spatially localized correspondence distributions, providing a stronger supervisory signal for the cross-entropy-based attention alignment loss $\mathcal{L}_{\text{attn}}$, whereas higher temperatures yield softer distributions across neighboring patches. For training the GARD denoiser, we used $T=0.01$ to construct the target correspondence map.
% Fig.~\ref{fig:suppl_attn_target} illustrates visualizations of the target correspondence map $\mathbf{A}^*$ under different temperature values $T$. To construct $\mathbf{A}^*$, we first estimate depth and camera pose from the clean high-quality (HQ) input views and unproject each pixel into a shared 3D world coordinate system. The resulting point cloud is subsequently spatially pooled to the ViT patch resolution (patch size 14) in order to align with the spatial token grid of the GARD denoiser $\mathcal{S}_{\theta}(\cdot)$. We then compute pairwise negative $\ell_2$ distances across all $VN$ patch tokens and apply a temperature-scaled softmax to obtain the entries of $\mathbf{A}^*$. Lower temperature values produce sharper correspondence distributions that concentrate probability mass on the geometrically nearest patches, whereas higher temperatures generate softer distributions that spread probability more broadly across neighboring patches. As illustrated in Fig.~\ref{fig:suppl_attn_target}, lower temperatures yield sharper and more spatially localized correspondence maps, thereby providing a stronger supervisory signal for the cross-entropy-based attention alignment loss $\mathcal{L}_{\text{attn}}$. For training the GARD denoiser, we used $T=0.01$ to construct the target correspondence map.

Fig.~\ref{fig:suppl_attn} presents visualizations of the GARD denoiser attention maps before and after applying the attention alignment loss. Specifically, we first train the GARD denoiser using only the flow-matching objective and visualize the global attention maps from the wide decoder layers within the denoiser. While the 11th and 13th decoder layers exhibit relatively accurate attention toward the green query point, the 9th decoder layer fails to establish reliable geometric correspondences. Motivated by both this observation and the significance of wide decoder layers demonstrated in RAE~\cite{zheng2025diffusion}, we explicitly regularize the global attention weights of the $J$-th decoder layer ($J=9$) to encourage attention toward geometrically corresponding regions rather than spurious artifacts. As reported in Tab.~\ref{tab:abl_gard_training}, although attention alignment does not consistently improve performance under standard flow matching, combining it with interpolated flow matching yields additional gains. We attribute this improvement to the initialization from an LQ structural prior rather than pure noise, which enables more effective learning of geometrically consistent correspondences.
% As reported in Tab.~\ref{tab:abl_gard_training}, although the attention alignment objective does not consistently yield performance improvements under the standard flow-matching formulation, its integration with interpolated flow matching leads to further performance gains. We attribute this improvement to the ability of interpolated flow matching to initialize the denoising process from a low-quality (LQ) structural prior rather than pure noise, thereby enabling the attention alignment objective to more effectively guide the learning of geometrically consistent correspondences.
\input{suppl/suppl_fig/suppl_attn_target}
\input{suppl/suppl_fig/suppl_attn}


\paragraph{View selection strategy.}
To construct multi-view training samples, we adopt a view selection strategy based on the ground-truth camera poses provided in each dataset. Given a target number of views $V$, we retrieve neighboring frames using an expansion ratio that defines a larger candidate window to select the required number of views. For Hypersim, we use a \textit{near-camera} strategy since adjacent views often exhibit relatively large viewpoint changes. For every anchor frame, we first precompute the ranking between frame pairs based on their geometric overlap, and then select additional candidates to form the multi-view input. We use an expansion ratio of 2, which encourages stronger geometric consistency across selected views. In contrast, TartanAir consists of long video sequences with strong temporal continuity, where consecutive frames are often highly redundant. To increase viewpoint diversity during training, we employ a \textit{near-random} strategy with an expansion ratio of 4. Using this strategy, we first collect a larger set of nearby frames around the anchor frame and then randomly sample the remaining views from this candidate window. For training, we set the maximum number of training input views to $V=4$.

\subsection{Inference Details}
During inference, degraded multi-view images are first processed using the DA3-GIANT-1.1 multi-view encoder~\cite{depthanything3}. We extract the intermediate geometry-aware representation from the $K=18$-th encoder layer and use it as input to the proposed GARD denoiser. The denoiser restores the degraded representation by modeling the flow field in the latent feature space. To perform denoising, we solve the corresponding probability flow ordinary differential equation (ODE) using a first-order Euler solver with 50 discretized sampling steps. The restored representation is subsequently propagated through the remaining layers of the multi-view encoder to obtain refined multi-scale features. From these features, four feature levels $\mathcal{Z}_{\text{res}} = {\mathbf{z}_{\text{res}}^{l}}_{l \in \mathcal{M}}$ are selected, where $\mathcal{M}=\{20, 28, 34, 40\}$ for DA3-GIANT-1.1 and $\mathcal{M}=\{6, 8, 10, 12\}$ for DA3-BASE. These features are provided to the geometry decoder and RGB decoder~\cite{jang2026gld} for downstream 3D reconstruction and image restoration, respectively. Overall, the proposed GARD framework jointly reconstructs high-quality multi-view images and accurate 3D geometry in a single forward pass, eliminating the need for separate restoration and reconstruction stages.

        