% \clearpage
\section{Introduction}
\label{sec:intro}
\vspace{-0.3em}

Recent MLLM-based mobile GUI agents can understand screenshots, reason over user goals, and control devices~\cite{gao2026ui,zhou2025mai,hong2024cogagent,xu2026mobile,liu2025llm}. However, they remain unreliable in long-horizon settings that require retaining intermediate facts, tracking multi-step progress, and carrying information across app boundaries~\cite{liu2026memgui,kong2025mobileworld}. As task horizons grow, even the strongest 8B end-to-end model, \llmname{GUI-Owl-1.5-8B}~\cite{xu2026mobile}, drops from 71.6\% on AndroidWorld~\cite{rawles2025androidworld} (avg.\ 8.4 steps) to 38.2\% on MobileWorld~\cite{kong2025mobileworld} (avg.\ 27.8 steps), and further to 11.7\% on MemGUI-Bench~\cite{liu2026memgui} (avg.\ 36.2 steps).

\input{figures/fig-teaser.tex}

We attribute this degradation to context management. \emph{Agentic frameworks} couple a backbone MLLM with planning, memory, and grounding modules~\cite{agashe2025agent,rawles2025androidworld,wang2024mobile,wang2025mobileagentselfevolving}, but rely on complex pipelines and often proprietary backbones. \emph{End-to-end models} are simpler and more deployable~\cite{gao2026ui,zhou2025mai,xu2026mobile}, yet manage context through passive strategies (e.g., Action-Thought, Multi-turn, Rule-based, No History; see Figure~\ref{fig:teaser} and Table~\ref{tab:main-results}) that append, truncate, or discard records mechanically rather than according to task needs. Consequently, prompts grow with task horizon (\emph{prompt explosion}), while critical cross-app facts such as prices, identifiers, or copied text may be diluted, paraphrased, truncated, or forgotten (\emph{information loss})~\cite{liu2026memgui}. We review related mobile GUI agents and context-management methods in Appendix~\ref{sec:related-work}.

Long-horizon mobile GUI control requires both compact working context and persistent UI-derived facts across screens, steps, and app transitions. Existing strategies fail to provide both: external memory moves context curation outside the end-to-end policy, while prompt-based or rule-based methods either accumulate passive logs or discard old facts mechanically. This motivates a policy-level mechanism that decides what to compress, what to remember, and what to keep available while acting.

\obsbox{\textit{Can a single end-to-end mobile GUI agent jointly choose UI actions and manage its working context, so that long-horizon interaction stays compact while critical UI facts persist across steps and app transitions?}}

We answer this question with \textbf{\ourmethod}, an end-to-end long-horizon mobile GUI agent with proactive context management. \ourmethod is built on \textbf{\underline{Con}}text-as-\textbf{\underline{Act}}ion (\textbf{\conact}), which casts context management as first-class actions emitted alongside UI actions. Instead of passively appending history, \conact maintains three structured context fields: \textcolor{histcolor}{\emph{Folded Action History}} \textcolor{histcolor}{$H_t$} for compressed trajectory summaries, \textcolor{memstatecolor}{\emph{Folded UI State}} \textcolor{memstatecolor}{$M_t$} for persistent UI-derived facts, and \textcolor{recentcolor}{\emph{Recent Step Record}} \textcolor{recentcolor}{$L_t$} for recent interaction details. All three are predicted in a single forward pass, keeping context curation inside the agent policy.

As shown in Figure~\ref{fig:main-performance}(a), \conact keeps prompt growth flatter than ReAct-style prompting. Figure~\ref{fig:main-performance}(b) shows that this lower context cost translates into stronger long-horizon task success: applied zero-shot to \llmname{Qwen3-VL-235B-Thinking}, \ourmethod reaches 62.5\% Pass@3 on MemGUI-Bench, surpassing agentic workflows built on \llmname{Gemini-2.5-Pro}. To make proactive context management learnable across model scales, we construct \textbf{MemGUI-3K}, a 2{,}956-trajectory dataset with full \conact annotations. Training an 8B model on MemGUI-3K produces \textbf{MemGUI-8B-SFT}, which achieves the best open-data 8B performance on MemGUI-Bench and generalizes to the out-of-distribution MobileWorld benchmark.

\paragraph{Contributions.}
\textbf{(1)}~We introduce \textbf{\ourmethod}, an end-to-end long-horizon mobile GUI agent built on \conact, a Context-as-Action paradigm that unifies history folding, UI memory, and self-describing step outputs within a single policy~(\S\ref{sec:method}).
\textbf{(2)}~We construct \textbf{MemGUI-3K}, a 2{,}956-trajectory dataset with full \conact annotations, supporting supervised training and offline analysis of proactive context management across model scales~(\S\ref{sec:dataset}).
\textbf{(3)}~We train \textbf{MemGUI-8B-SFT}, an 8B \ourmethod supervised on MemGUI-3K, achieving the best open-data 8B performance on MemGUI-Bench and generalizing to the out-of-distribution MobileWorld benchmark~(\S\ref{sec:experiments}).
