\section{Related Work}
\label{sec:related-work}
\vspace{-0.25em}

\paragraph{Mobile GUI agents.}
Mobile GUI agents built on multimodal models generally follow two paradigms. 
\emph{Agentic frameworks} combine a backbone MLLM with planning, memory, and grounding modules~\cite{rawles2025androidworld,wang2024mobile,wang2025mobileagentselfevolving,agashe2025agent}, often around proprietary backbones such as \llmname{Gemini-2.5-Pro}~\cite{comanici2025gemini}; they can perform strongly but are complex and difficult to reproduce. 
\emph{End-to-end models}~\cite{gao2026ui,zhou2025mai,hong2024cogagent,xu2026mobile} directly map screenshots to actions, offering simplicity and deployability, but typically rely on passive context mechanisms: Action-Thought traces~\cite{zhang2025appagent,gu2025ui-venus,xu2026mobile,bai2025qwen3}, multi-turn prompts~\cite{qin2025ui}, rule-based aggregation~\cite{zheng2024seeact}, or no-history prompting~\cite{hong2024cogagent}. 
These mechanisms either let prompts grow with task horizon or lose UI-derived facts needed across screens and apps~\cite{liu2026memgui,kong2025mobileworld}. 
\ourmethod brings proactive context management into the end-to-end paradigm itself.

\paragraph{Context management in LLM agents.}
Long-context LLM agents manage histories through external memory or prompt-internal curation. 
External memory maintains a separate store, such as retrieval-augmented databases~\cite{lewis2020retrieval,chhikara2025mem0,xu2026mem}, hierarchical memory~\cite{zhang2026g}, or experiential replay~\cite{zhao2024expel}. 
Prompt-internal curation edits the in-task context through truncation, summarization, or model-controlled compression, as in recent web-agent methods such as MEM1~\cite{zhou2025mem1}, MemAgent~\cite{yu2025memagent}, and AgentFold~\cite{ye2025agentfold}. 
Unlike text-only web settings, mobile GUI control must preserve UI-derived facts, such as prices, codes, and copied text, verbatim across screen and app transitions. 
\conact addresses this by treating context curation as joint action over structured history, UI state, and recent-step fields within the same end-to-end policy.