\section{Related Work}
\label{sec:realted_word}
\vspace{-2mm}
\paragraph{Foundation Models for Robotic Manipulation.}
Large vision-language and vision-language-action models have recently become a central paradigm for building generalizable robotic manipulation systems. 
One line of work turns multimodal foundation models into robot policies by adding action-generation modules and training on large-scale robot trajectories, often using diffusion or flow-matching objectives for continuous action prediction~\citep{li2024vision,kim2024openvla,pi0.5}. 
These models acquire broad manipulation skills, but their adaptation often depends on substantial robot data, and their internal action representations make explicit constraint checking and plan repair difficult. 
Another line of work uses VLMs as high-level reasoning modules and grounds their outputs through structured spatial representations, such as affordance map and keypoints relations~\citep{huang2023voxposer,fangandliu2024moka,huang2024rekep,yuan2024robopoint}. 
These methods improve spatial grounding and compositionality, but often rely on carefully designed perception-action interfaces or hand-specified planning primitives.

Recent work further explores explicit action reasoning before execution. MolmoAct and MolmoAct2 reason about actions through interpretable spatial or language-conditioned representations~\citep{lee2025molmoact,fang2026molmoact2actionreasoningmodels}, while ThinkAct compresses embodied reasoning into a visual plan latent that conditions a downstream action model~\citep{huang2026thinkact}. Hierarchical methods such as HAMSTER~\citep{li2025hamster} separate high-level reasoning from low-level control policies. 
These approaches improve the reasoning ability of robot policies, but the execution still typically depends on learned action heads. 
This makes it difficult to explicitly inspect a plan and iteratively repair it according to the task specification.

\vspace{-2mm}
\paragraph{Harnessing Agents for Robotic Manipulation.}
Pioneering works such as Code-as-Policies~\citep{liang2022codeap} and ProgPrompt~\citep{singh2023progprompt} show that language models can compose perception outputs, control primitives, and task-specific APIs into executable policies or situated task plans. 
These harnessing frameworks inherit the modularity of classical robotic systems: perception, planning, and control can be represented as callable tools, while the language model composes them into task-specific behavior. 
Recent work has extended this paradigm to multimodal and multi-agent settings. 
RoboCodeX~\citep{mu2024robocodex} translates high-level instructions and scene understanding into executable robotic programs through multimodal code generation, while RoCo and related works~\citep{mandi2024roco,chen2025emos} integrate generated code with multi-robot coordination and motion planning. 
Recent work Maestro~\citep{shi2025maestro} and concurrent work Cap-X~\citep{fu2026capxframeworkbenchmarkingimproving} further scale this paradigm by enabling robots to call diverse perception and control tools through writing programs.
However, most existing harness frameworks rely on one-shot program generation and execution, providing limited opportunities for agents to react to execution outcomes or recover from failures.
In contrast, we study how harness design can enable effective embodied manipulation through a ReAct-style workflow~\citep{yao2023reactsynergizingreasoningacting} that continuously interleaves perception, reasoning, and action execution. 
We further investigate whether such a harness can act as a general interface for transferring embodied capabilities to small open models.
