% \section{Related Work}

% \paragraph{DeepResearch Tasks.}
% Recent evaluation paradigms for \emph{deep research agents} have shifted beyond short-form QA toward report-style synthesis and groundedness\citep{shao2024assistingwritingwikipedialikearticles,Semnani_2023, wei2024longformfactualitylargelanguage}. Notably, FINDER \citep{zhang2025fargenuinelyusefuldeep} introduces checklist-style rubrics and the DEFT taxonomy to categorize failures in retrieval and reasoning. Similarly, general agent benchmarks focusing on long-horizon browsing \citep{Zhou2023WebArenaAR,mialon2023gaiabenchmarkgeneralai,xie2024osworldbenchmarkingmultimodalagents} have driven progress but remain predominantly outcome-centric. These metrics often treat the agent as a black box, making it difficult to localize the provenance of errors or explain why trajectories incur high computational costs. Our work complements these report-level evaluations by shifting focus to trajectory dynamics diagnostics—such as identifying earliest failures and redundancy—to provide actionable signals for improving research behavior.

% \paragraph{DeepResearch systems.}
% Agent frameworks for deep research typically integrate planning, tool orchestration, memory management, and termination heuristics \citep{xi2023risepotentiallargelanguage}. While these architectural choices act as critical determinants of robustness and efficiency, their impact is frequently entangled with the capabilities of the underlying LLM in empirical comparisons. Beyond deep research, systematic failure analyses have surfaced in multi-agent settings; for instance, MAST \citep{cemri2025multiagentllmsystemsfail} establishes an empirically grounded taxonomy of failure modes across various multi-agent frameworks, highlighting challenges in specification and verification. Building on this systematic approach, we employ controlled comparisons to decouple \emph{framework-induced instability} from \emph{model-induced unreliability}. Distinctively, we focus on single-agent deep research trajectories, introducing granular diagnostic anchors—such as EFS, propagation spans, recovery, and redundancy—specifically tailored to long-horizon, tool-mediated workflows.

% \paragraph{Process reward models.}
% Process Reward Models (PRMs) \citep{lightman2023letsverifystepstep, wang2024mathshepherdverifyreinforcellms, zhang2025lessonsdevelopingprocessreward,uesato2022solvingmathwordproblems} offer granular oversight for intermediate reasoning steps, alleviating the ambiguity inherent in sparse terminal supervision. However, evaluating PRMs remains a non-trivial challenge, as binary metrics of ``step correctness'' often fail to capture latent reasoning flaws or subtle inefficiencies. While recent benchmarks like PRMBench \citep{song2025prmbenchfinegrainedchallengingbenchmark} have introduced multi-dimensional labeling to expose deficiencies in current PRMs and LLM-based critics \citep{mcaleese2024llmcriticshelpcatch,gao2024llmcriticshelpcatch, zhang2025criticvvlmcriticshelp}, and others explore specific paradigms such as Socratic reasoning \citep{li2025socraticprmbenchbenchmarkingprocessreward}, our work targets a distinct frontier. Diverging from synthetic or purely logic-centric domains, our benchmark evaluates diagnostic capabilities within \emph{deep research trajectories} that involve complex tool interactions. We prioritize fine-grained error localization and the detection of redundancy or stagnation, utilizing these diagnostics to advance PRM training through supervised learning and preference optimization.




\section{Related Work}


\paragraph{Deep-research systems and outcome-level evaluation.}
Recent agent benchmarks, such as GAIA, BrowseComp, WebArena, and OSWorld, have shifted focus from static QA to long-horizon tasks including web navigation and tool use \citep{mialon2023gaiabenchmarkgeneralai, wei2025browsecomp, zhou2024webarenarealisticwebenvironment, xie2024osworldbenchmarkingmultimodalagents}. Furthermore, newer benchmarks like DeepResearch Bench, DeepResearchGym, LiveResearchBench, and DRBench emphasize rubric-based assessment of citation-grounded reports \citep{du2025deepresearchbenchcomprehensivebenchmark, coelho2025deepresearchgymfreetransparentreproducible, wang2026liveresearchbenchlivebenchmarkusercentric, abaskohi2026drbenchrealisticbenchmarkenterprise}. Despite improving realism, these evaluations remain primarily outcome-centered: they assess final task completion but fail to localize where a research trajectory first becomes unreliable. In contrast, TELBench evaluates the process itself by segmenting trajectories into semantic spans, requiring models to pinpoint harmful errors within ordered evidence.

\paragraph{Process-level evaluation and trajectory diagnosis.}
Moving beyond outcome-level metrics, recent works evaluate intermediate reasoning and agent traces. Frameworks such as ProcessBench, PRMBench, Delta-Bench, VisualProcessBench, AgentProcessBench, and TRACE focus on step-level or tool-use errors~\citep{zheng-etal-2025-processbench, song2025prmbenchfinegrainedchallengingbenchmark, he2025largelanguagemodelsdetect, wang2025visualprm, fan2026agentprocessbenchdiagnosingsteplevelprocess, Kim2025BeyondTF}, while MAST, TRAIL, AgentRx, and CodeTracer address failure localization and trace debugging~\citep{cemri2025multiagentllmsystemsfail, deshpande2025trailtracereasoningagentic, barke2026agentrxdiagnosingaiagent, li2026codetracertraceableagentstates}. While valuable (Table~\ref{tab:related-error-localization}), prior signals are mostly built for shorter or more structured settings such as math reasoning, VLM reasoning~\citep{wei2026agentic}, coding traces, and API workflows. Deep-research trajectories are longer and noisier, mixing useful exploration, weak evidence, failed searches, and harmful mistakes. \textsc{TELBench} focuses on this setting by evaluating whether models can localize harmful error spans from ordered semantic spans, rather than only judging final answers or overall trajectory quality.
