\begin{mdframed}[style=mdpurplebox,frametitle={USAMO 2026 Problem 6}]
Let $a$ and $b$ be positive integers such that $\varphi(ab+1)$ divides $a^2 + b^2 + 1$. Prove that $a$ and $b$ are Fibonacci numbers.
\end{mdframed}
\solutionheading{Solution}
\begin{remark*}
  7/7.
\end{remark*}
Let \(a\) and \(b\) be positive integers such that\par
\[
\varphi(ab+1) \mid a^{2}+b^{2}+1.
\]
Set \(N = ab+1\) and define the positive integer\par
\[
k = \frac{a^{2}+b^{2}+1}{\varphi(N)}.
\]
We will prove that both \(a\) and \(b\) must be Fibonacci numbers.\par
\smallskip
\solutionheading{1. Preliminary bounds}
Since \(N > 1\), we have \(\varphi(N) \le N-1 = ab\) (with equality exactly when \(N\) is prime). Hence\par
\[
a^{2}+b^{2}+1 = k\,\varphi(N) \le k\,ab. \quad\text{(1)}
\]
By the quadratic-mean-arithmetic-mean inequality, \(a^{2}+b^{2} \ge 2ab\), so\par
\[
a^{2}+b^{2}+1 \ge 2ab+1. \quad\text{(2)}
\]
Combining (1) and (2) yields\par
\[
2ab+1 \le k\,ab \quad\Longrightarrow\quad k \ge 2 + \frac{1}{ab}.
\]
Because \(k\) is an integer, if \(ab > 1\) we obtain \(k \ge 3\). If \(ab = 1\) then necessarily \(a=b=1\), \(N=2\), \(\varphi(2)=1\), and then \(k = 3\). Thus in every case\par
\[
\boxed{k \ge 3}. \quad\text{(3)}
\]
\smallskip
\solutionheading{2. Parity and the even case}
\textbf{Lemma 2.1.} If \(N\) is even and \(N>2\), then the divisibility condition cannot hold.\par
\emph{Proof.} For \(N>2\), \(\varphi(N)\) is even (a standard fact: \(\varphi(n)\) is even for every \(n>2\)). Because \(N\) even implies \(ab = N-1\) odd, both \(a\) and \(b\) are odd. Then \(a^{2}+b^{2}+1\) is odd (odd+odd+1). An even number cannot divide an odd number. The only even possibility is \(N=2\), which gives \(a=b=1\) and indeed \(\varphi(2)=1\) divides \(3\).\par
Consequently, the only even \(N\) that can occur is \(N=2\), producing \((a,b)=(1,1)\) - both Fibonacci numbers. From now on we assume\par
\[
N\ \text{is odd and}\ N>2.
\]
Then \(ab = N-1\) is even, so at least one of \(a,b\) is even. If both were even, \(a^{2}+b^{2}+1\) would be odd, while for odd \(N>2\) we still have \(\varphi(N)\) even. Hence exactly one of \(a,b\) is even and the other is odd. In particular,\par
\[
a^{2}+b^{2}+1 \equiv 0+1+1 \equiv 2 \pmod{4}. \quad\text{(4)}
\]
\smallskip
\solutionheading{3. Case I: \(N\) is prime}
Assume \(N = ab+1\) is prime. Then \(\varphi(N) = N-1 = ab\), and (1) becomes\par
\[
a^{2}+b^{2}+1 = k\,ab. \quad\text{(5)}
\]
We first determine \(k\).\par
\solutionheading{3.1. The value of \(k\)}
Without loss of generality, order the variables so that \(a \le b\). Write (5) as a quadratic in \(b\):\par
\[
b^{2} - (k a)\,b + (a^{2}+1) = 0.
\]
If \(b\) is a root, the other root is\par
\[
b' = \frac{a^{2}+1}{b} = k a - b,
\]
which is a positive integer.\par
\textbf{Claim.} If \(a \ge 2\), then \(b' < b\).\par
\emph{Proof.} Suppose \(b' \ge b\). Then \(\frac{a^{2}+1}{b} \ge b\), i.e. \(a^{2}+1 \ge b^{2}\). Because \(b \ge a\), we have \(a^{2} \le b^{2} \le a^{2}+1\). Hence either \(b^{2}=a^{2}\) or \(b^{2}=a^{2}+1\).\par
\compactbullet{If \(b^{2}=a^{2}\), then \(b=a\). Substituting \(b=a\) into (5) gives \(2a^{2}+1 = k a^{2}\) \(\Rightarrow\) \(k = 2 + \frac{1}{a^{2}}\), which is not an integer for \(a \ge 2\). Contradiction.}
\compactbullet{If \(b^{2}=a^{2}+1\), then \(a^{2}+1\) is a perfect square. For \(a \ge 2\), this is impossible because between \(a^{2}\) and \((a+1)^{2}=a^{2}+2a+1\) there is no other square.}
Thus whenever \(a \ge 2\), we can replace \((a,b)\) by \((a,b')\) (which is also a solution of (5)) with \(a+b' < a+b\). Repeating this descent (the sum \(a+b\) strictly decreases each step) we eventually reach a solution with \(a=1\) (the minimal sum cannot have \(a \ge 2\)).\par
Now consider \(a=1\). Equation (5) becomes\par
\[
1 + b^{2} + 1 = k \cdot 1 \cdot b \quad\Longrightarrow\quad b^{2} - k b + 2 = 0. \quad\text{(6)}
\]
For integer \(b\), the discriminant \(\Delta = k^{2}-8\) must be a perfect square, say \(d^{2}\). Then\par
\[
k^{2} - d^{2} = 8 \quad\Longrightarrow\quad (k-d)(k+d)=8.
\]
Both factors are positive integers of the same parity. The factor pairs of \(8\) are \((1,8)\) and \((2,4)\).\par
\compactbullet{\((1,8)\) gives \(k-d=1\), \(k+d=8\) \(\Rightarrow\) \(2k=9\) \(\Rightarrow\) \(k=9/2\), not integer.}
\compactbullet{\((2,4)\) gives \(k-d=2\), \(k+d=4\) \(\Rightarrow\) \(2k=6\) \(\Rightarrow\) \(k=3\), \(d=1\).}
Hence \(k=3\) is forced. With \(k=3\), equation (6) is \(b^{2}-3b+2=0\), whose roots are \(b=1\) and \(b=2\). Thus the only solutions with \(a=1\) are \((1,1)\) and \((1,2)\) (and by symmetry, \((2,1)\) if we had not assumed \(a\le b\)). Since the reduction process preserves \(k\), every solution of (5) must have \(k=3\).\par
Therefore, when \(N\) is prime,\par
\[
a^{2}+b^{2}+1 = 3ab \quad\Longleftrightarrow\quad a^{2} - 3ab + b^{2} = -1. \quad\text{(7)}
\]
\solutionheading{3.2. Solving \(a^{2} - 3ab + b^{2} = -1\)}
Equation (7) is symmetric; we may assume \(a \le b\).\par
Define a sequence \((A_n)_{n\ge 0}\) by\par
\[
A_0 = 1,\qquad A_1 = 1,\qquad A_{n+2} = 3A_{n+1} - A_n\ \ (n\ge 0).
\]
One verifies by induction that \((A_n, A_{n+1})\) satisfies (7): For \(n=0\), \(1^2-3\cdot1\cdot1+1^2=-1\). Assuming it holds for \(n\), using the recurrence one checks that it also holds for \(n+1\). (The computation is straightforward and can be filled in similarly to the proof in Lemma 4.1 of the original draft.)\par
Now let \((a,b)\) be any positive integer solution of (7) with \(a \le b\). View (7) as a quadratic in \(b\):\par
\[
b^{2} - 3a\, b + (a^{2}+1) = 0.
\]
Its two roots are \(b\) and\par
\[
b' = \frac{a^{2}+1}{b} = 3a - b,
\]
which is an integer. A similar argument as in 3.1 shows that if \(a \ge 2\), then \(b' < b\). Consequently, by infinite descent on \(a+b\), we can reduce any solution to one with \(a=1\). The only solutions with \(a=1\) are \((1,1)\) and \((1,2)\) (and their symmetric versions). Therefore every solution can be obtained from these minimal ones by the inverse transformation: if \((x,y)\) is a solution with \(x \le y\), then \((y,\, 3y - x)\) is also a solution. Starting from \((1,1)\) and repeatedly applying this transformation yields the increasing sequence\par
\[
(1,1),\ (1,2),\ (2,5),\ (5,13),\ (13,34),\ \dots
\]
which exactly corresponds to the pairs \((A_n, A_{n+1})\) for \(n\ge 0\).\par
It remains to identify these \(A_n\) with Fibonacci numbers. Define the Fibonacci numbers by\par
\[
F_1 = 1,\quad F_2 = 1,\quad F_{n+2} = F_{n+1} + F_n.
\]
We claim that for \(n \ge 1\),\par
\[
A_n = F_{2n-1}.
\]
\emph{Proof by induction.}\par
\compactbullet{Base \(n=1\): \(A_1 = 1 = F_1\).}
\compactbullet{\(n=2\): \(A_2 = 3A_1 - A_0 = 3\cdot1 - 1 = 2 = F_3\).}
\compactbullet{Inductive step: Assume \(A_n = F_{2n-1}\) and \(A_{n+1} = F_{2n+1}\). Then}
\[
A_{n+2} = 3A_{n+1} - A_n = 3F_{2n+1} - F_{2n-1}.
\]
Using the identities
\[
F_{2n+1} - F_{2n-1} = F_{2n},\quad \text{and}\quad F_{2n+1} + F_{2n} = F_{2n+2},
\]
we compute\par
\[
\begin{aligned}
3F_{2n+1} - F_{2n-1} &= 2F_{2n+1} + (F_{2n+1} - F_{2n-1}) \\
&= 2F_{2n+1} + F_{2n} \\
&= F_{2n+1} + (F_{2n+1} + F_{2n}) \\
&= F_{2n+1} + F_{2n+2} \\
&= F_{2n+3}.
\end{aligned}
\]
Thus \(A_{n+2} = F_{2n+3} = F_{2(n+2)-1}\), completing the induction.\par
Hence all solutions of (7) are pairs of Fibonacci numbers (specifically, consecutive odd-index Fibonacci numbers). This completes \textbf{Case I}.\par
\smallskip
\solutionheading{4. Case II: \(N\) is composite and odd (\(N>2\))}
Now \(N = ab+1\) is composite, odd, and \(N>2\). From (4) we have\par
\[
a^{2}+b^{2}+1 \equiv 2 \pmod{4}.
\]
\solutionheading{4.1. \(N\) cannot have two distinct odd prime factors}
Suppose \(N\) has at least two distinct odd prime divisors \(p\) and \(q\). Then \(\varphi(N)\) is divisible by \((p-1)(q-1)\). Since \(p\) and \(q\) are odd, both \(p-1\) and \(q-1\) are even, hence \(4 \mid (p-1)(q-1) \mid \varphi(N)\). Thus \(4 \mid \varphi(N)\). Since \(\varphi(N) \mid a^{2}+b^{2}+1\), we would have \(4 \mid a^{2}+b^{2}+1\), contradicting (4). Therefore \(N\) cannot possess two distinct odd prime factors.\par
It follows that \(N\) must be a prime power:\par
\[
N = p^{e}, \qquad e \ge 2\ (\text{since } N \text{ is composite}).
\]
\solutionheading{4.2. The prime \(p\) cannot be \(\equiv 1 \pmod{4}\)}
If \(p \equiv 1 \pmod{4}\), then \(p-1\) is divisible by \(4\), so \(4 \mid \varphi(N)\). Again this forces \(4 \mid a^{2}+b^{2}+1\), which is impossible by (4). Hence\par
\[
p \equiv 3 \pmod{4}. \quad\text{(8)}
\]
\solutionheading{4.3. Analysis modulo \(3\)}
We now examine the residue of \(p\) modulo \(3\).\par
\textbf{Subcase }\(p \equiv 1 \pmod{3}\).\textbf{ Then }\(3 \mid (p-1)\), so \(3 \mid \varphi(N)\). Moreover, \(p \equiv 1 \pmod{3}\) implies \(p^{e} \equiv 1 \pmod{3}\), so \(ab = p^{e} - 1 \equiv 0 \pmod{3}\). Thus \(3 \mid ab\).\par
Now evaluate \(S = a^{2}+b^{2}+1\) modulo \(3\). Since \(3 \mid ab\), there are two possibilities:\par
\compactbullet{\(3 \mid a\) and \(3 \mid b\): then \(a^{2}\equiv0\), \(b^{2}\equiv0\), so \(S \equiv 1 \pmod{3}\).}
\compactbullet{Exactly one of \(a,b\) is divisible by \(3\): then \(S \equiv 0+1+1 = 2 \pmod{3}\).}
In neither case is \(S\) divisible by \(3\). However, \(\varphi(N) \mid S\) and \(3 \mid \varphi(N)\) would force \(3 \mid S\). Contradiction. Therefore \(p \not\equiv 1 \pmod{3}\).\par
\textbf{Subcase }\(p \equiv 2 \pmod{3}\).\textbf{ Then }\(p-1 \equiv 1 \pmod{3}\), so \(3 \nmid (p-1)\). But because \(e \ge 2\), \(\varphi(N) = p^{e-1}(p-1)\) contains the factor \(p^{e-1}\), so in particular \(p \mid \varphi(N)\). Consequently,\par
\[
p \mid S. \quad\text{(9)}
\]
Also, from \(ab = p^{e} - 1\) we have\par
\[
ab \equiv -1 \pmod{p}. \quad\text{(10)}
\]
In particular, \(p\) does not divide \(a\) or \(b\) (otherwise (10) would give \(0 \equiv -1 \pmod{p}\), impossible). Hence \(a\) and \(b\) are invertible modulo \(p\). Write \(b \equiv -a^{-1} \pmod{p}\). Substituting into (9):\par
\[
a^{2} + (-a^{-1})^{2} + 1 \equiv 0 \pmod{p},
\]
i.e.,
\[
a^{2} + a^{-2} + 1 \equiv 0 \pmod{p}.
\]
Multiply by \(a^{2}\) (invertible modulo \(p\)):\par
\[
a^{4} + a^{2} + 1 \equiv 0 \pmod{p}. \quad\text{(11)}
\]
Set \(y = a^{2} \pmod{p}\); then (11) becomes\par
\[
y^{2} + y + 1 \equiv 0 \pmod{p}. \quad\text{(12)}
\]
Equation (12) has a solution in \(\mathbb{F}_p\) iff \(-3\) is a quadratic residue modulo \(p\). Indeed, multiplying (12) by \(4\) gives \((2y+1)^{2} \equiv -3 \pmod{p}\).\par
We need the following standard lemma:\par
\textbf{Lemma.} For an odd prime \(p \neq 3\), \(-3\) is a quadratic residue modulo \(p\) if and only if \(p \equiv 1 \pmod{3}\).\par
\emph{Proof.} The multiplicative group \(\mathbb{F}_p^{\times}\) is cyclic of order \(p-1\). The equation \(y^{2}+y+1=0\) is equivalent to \(y^{3}=1\) with \(y \neq 1\) (since \((y-1)(y^{2}+y+1)=y^{3}-1\). Hence a solution exists iff the group contains an element of order \(3\), i.e., iff \(3 \mid p-1\), which is \(p \equiv 1 \pmod{3}\). Conversely, if \(p \equiv 1 \pmod{3}\), then \(3 \mid p-1\), so such an element exists.\par
Since \(p \equiv 2 \pmod{3}\), we have \(p-1\) not divisible by \(3\), so there is no element of order \(3\) in \(\mathbb{F}_p^{\times}\); therefore (12) has no solution modulo \(p\). This contradicts the existence of \(a\). Thus \(p \not\equiv 2 \pmod{3}\).\par
The only remaining possibility for the odd prime \(p\) is \(p = 3\).\par
\solutionheading{4.4. The exponent \(e\)}
Now \(N = 3^{e}\) with \(e \ge 2\). Then\par
\[
\varphi(N) = 3^{e-1} \cdot 2.
\]
\textbf{If }\(e \ge 3\):\textbf{ Then }\(3^{e-1}\) is at least \(3^{2}=9\), so \(9 \mid \varphi(N)\). Since \(\varphi(N) \mid S\), we must have \(9 \mid S\).\par
But we will show that \(S \equiv 3 \pmod{9}\), which is not divisible by \(9\), a contradiction. Let us verify this congruence.\par
Because \(e \ge 3\), \(N = 3^{e}\) is divisible by \(27\), in particular \(N \equiv 0 \pmod{9}\). Hence\par
\[
ab = N-1 \equiv -1 \pmod{9}. \quad\text{(13)}
\]
Moreover, \(N \equiv 0 \pmod{3}\) gives \(ab \equiv -1 \equiv 2 \pmod{3}\). Hence \(3 \nmid a\) and \(3 \nmid b\) (otherwise \(ab \equiv 0 \pmod{3}\)). Thus \(a\) and \(b\) are each \(\equiv 1\) or \(2 \pmod{3}\), and their product is \(2 \pmod{3}\); the only possibilities are \(\{a,b\} \equiv \{1,2\} \pmod{3}\). Consequently,\par
\[
a + b \equiv 1 + 2 \equiv 0 \pmod{3}. \quad\text{(14)}
\]
Now compute\par
\[
S = a^{2}+b^{2}+1 = (a+b)^{2} - 2ab + 1.
\]
From (14), \(a+b = 3t\) for some integer \(t\), so \((a+b)^{2} = 9t^{2} \equiv 0 \pmod{9}\). Using (13), \(ab \equiv -1 \pmod{9}\), so \(-2ab \equiv -2(-1) = 2 \pmod{9}\). Thus\par
\[
S \equiv 0 + 2 + 1 = 3 \pmod{9}.
\]
Thus \(9 \nmid S\), contradicting \(9 \mid \varphi(N) \mid S\). Therefore \(e\) cannot be \(\ge 3\).\par
The only remaining possibility is \(e = 2\), i.e., \(N = 9\). Then \(\varphi(9) = 6\) and \(ab = 8\). The positive factor pairs of \(8\) are\par
\[
(1,8),\ (2,4),\ (4,2),\ (8,1).
\]
Compute \(S\) for each:\par
\compactbullet{\((1,8)\): \(S = 1 + 64 + 1 = 66\), and \(66 / 6 = 11 \in \mathbb{Z}\).}
\compactbullet{\((2,4)\): \(S = 4 + 16 + 1 = 21\), \(21/6 = 3.5 \notin \mathbb{Z}\).}
Thus only \((1,8)\) and \((8,1)\) satisfy the divisibility condition. Both \(1\) and \(8\) are Fibonacci numbers (\(1 = F_1\), \(8 = F_6\)).\par
\smallskip
\solutionheading{5. Conclusion}
We have examined all possible values of \(N = ab+1\):\par
\compactbullet{\textbf{Even case:} \(N=2\) gives \((a,b)=(1,1)\) - Fibonacci numbers.}
\compactbullet{\textbf{Prime case:} \(N\) prime leads to the equation \(a^{2} - 3ab + b^{2} = -1\), whose complete set of positive integer solutions consists of pairs of Fibonacci numbers (as shown in 3.2).}
\compactbullet{\textbf{Composite case:} \(N\) composite forces \(N=9\) and yields \((a,b) = (1,8)\) or \((8,1)\), again Fibonacci numbers.}
Hence, under the hypothesis \(\varphi(ab+1) \mid a^{2}+b^{2}+1\), the integers \(a\) and \(b\) are necessarily Fibonacci numbers.\par
\smallskip
\emph{Remark.} The Fibonacci numbers are defined by \(F_1=F_2=1\) and \(F_{n+2}=F_{n+1}+F_n\). The solutions from the prime case are precisely \((F_{2n-1},F_{2n+1})\) (and symmetric) for \(n \ge 1\), together with the degenerate pair \((1,1)\) which can be seen as \((F_1,F_2)\) or \((F_2,F_1)\). The pair \((1,8)\) from the composite case also fits the Fibonacci sequence (\(F_1=1\), \(F_6=8\)). This completes the proof.\par
\par
