\documentclass[10pt, logo, onecolumn]{shengshu}

\input{preamble}
\input{math_commands}

\title{Causal Forcing++: Scalable Few-Step Autoregressive Diffusion Distillation for Real-Time Interactive Video Generation}

\author{%
  \textbf{Min Zhao}$^{1,2~*}$, \textbf{Hongzhou Zhu}$^{1,2~*}$, \textbf{Kaiwen Zheng}$^{1}$, \textbf{Zihan Zhou}$^{3}$,
  \textbf{Bokai Yan}$^{3}$, \textbf{Xinyuan Li}$^{1}$, \textbf{Xiao Yang}$^{1}$, 
  \textbf{Chongxuan Li}$^{3}$, \textbf{Jun Zhu}$^{1,2~\dagger}$\\
  $^1$Tsinghua University\,
  $^2$ShengShu\,
  $^3$Renmin University of China\\
  {\small $^*$Equal contribution. \quad $^\dagger$Correspondence to: Jun Zhu.}\\
\texttt{\{gracezhao1997, suinibian74\}@gmail.com;}
\texttt{dcszj@tsinghua.edu.cn}
}

\begin{abstract}
Real-time interactive video generation requires low-latency, streaming, and controllable rollout. Existing autoregressive (AR) diffusion distillation methods have achieved strong results in the chunk-wise 4-step regime by distilling bidirectional base models into few-step AR students, but they remain limited by coarse response granularity and non-negligible sampling latency. In this paper, we study a more aggressive setting: frame-wise autoregression with only 1--2 sampling steps. In this regime, we identify the initialization of a few-step AR student as the key bottleneck: existing strategies are either target-misaligned, incapable of few-step generation, or too costly to scale. We propose \textbf{Causal Forcing++}, a principled and scalable pipeline that uses \emph{causal consistency distillation} (causal CD) for few-step AR initialization. The core idea is that causal CD learns the same AR-conditional flow map as causal ODE distillation, but obtains supervision from a single online teacher ODE step between adjacent timesteps, avoiding the need to precompute and store full PF-ODE trajectories. This makes the initialization both more efficient and easier to optimize. The resulting pipeline, \ours, surpasses the SOTA 4-step chunk-wise Causal Forcing under the \textit{\textbf{frame-wise 2-step setting}} by 0.1 in VBench Total, 0.3 in VBench Quality, and 0.335 in VisionReward, while reducing first-frame latency by 50\% and Stage 2 training cost by $\sim$$4\times$. We further extend the pipeline to action-conditioned world model generation in the spirit of Genie3.
Project Page: \textbf{\href{https://github.com/thu-ml/Causal-Forcing}{\textcolor{shengshublue}{https://github.com/thu-ml/Causal-Forcing}}} \& \textbf{\href{https://github.com/shengshu-ai/minWM}{\textcolor{shengshublue}{https://github.com/shengshu-ai/minWM}}}.
\end{abstract}

\begin{document}
\maketitle

\input{src/1-Introduction.tex}
\input{src/2-Background.tex}
\input{src/3-Method.tex}
\input{src/4-Experiment.tex}
\input{src/5-Conclusion.tex}

\clearpage

{
  \small
  \bibliographystyle{unsrt}
  \bibliography{main}
}

\end{document}
