% World models aim to transform video generation from passive visual synthesis
% into interactive simulation, where users can navigate a generated environment,
% revisit previously observed regions, and trigger events that persist over time.
% Achieving this goal in a general-purpose setting requires jointly addressing
% data diversity across visual domains, geometrically precise camera control,
% long-horizon consistency, and low-latency generation.
% We present \method{}, a general-purpose interactive text/image-to-video world
% model built as an integrated data, training, and deployment system.
% Our data engine combines camera-accurate Unreal Engine rendering, action-rich
% gameplay recordings, and real-world videos with recovered camera geometry,
% enabling a single model to generate photorealistic, game-style, and stylized
% worlds.
% For camera control, we introduce \eprope{}, a lightweight variant of projective
% positional encoding that achieves comparable trajectory-following precision to
% the full formulation with substantially lower computational cost.
% We progressively transform a bidirectional video generator into a few-step
% autoregressive world model through a training recipe that combines
% autoregressive adaptation, DMD2-style distillation, and forcing-based rollout
% correction. By training on self-generated long-horizon contexts, this recipe
% improves long-horizon generation over hundreds of frames while substantially
% mitigating the progressive drift in visual style and color that commonly
% emerges during autoregressive rollouts.
% Geometry-guided memory retrieval with error recycling further improves scene
% persistence over long-horizon camera revisits.
% Event Instruction Tuning introduces composable event control, and reinforcement
% learning improves visual quality and motion liveliness.
% An optimized asynchronous deployment pipeline reaches up to 16\,FPS on eight
% RTX\,5090 GPUs.
% On interaction-centric evaluation, \method{} achieves a camera-control score
% of \textbf{73.75} and an overall score of \textbf{84.76}, outperforming HY-WorldPlay (\textbf{80.79})
% and LingBot-World (\textbf{80.45}).

\method{} is a general-purpose interactive text/image-to-video world model for
controllable long-horizon generation.
It supports camera navigation, revisits to previously observed regions, and
promptable events across photorealistic, game-style, and stylized domains.
Our data engine combines camera-accurate Unreal Engine rendering, action-rich
gameplay recordings, and real-world videos with recovered camera geometry.
For camera control, we introduce \eprope{}, a lightweight variant of projective
positional encoding that retains PRoPE's projective camera geometry while
applying camera-aware attention to spatially reduced tokens.
We convert a bidirectional video generator into a few-step autoregressive world
model using causal forcing, DMD-style distillation, and long-rollout training.
Training on self-generated long-horizon contexts exposes the model to its own
generated history and reduces the style and color drift that accumulates across
autoregressive chunks.
Memory-Conditioned Scene Persistence retrieves earlier views through
camera-geometry-based retrieval, while residual recycling makes the conditioning
path less sensitive to imperfect memory latents.
Event Instruction Tuning adds composable event control, and reinforcement
learning alignment recovers camera control and visual quality after distillation.
With mixed-precision DiT execution, residual reuse, 75\%-pruned VAE decoding,
and asynchronous pipeline parallelism, \method{} reaches up to 16\,FPS on eight
RTX\,5090 GPUs.
On our 5-second basic evaluation, \method{} achieves a camera-control score
of \textbf{73.75} and an overall score of \textbf{84.76}, outperforming
HY-WorldPlay 1.5 and LingBot-World in overall score, which achieve \textbf{80.79}
and \textbf{80.45}, respectively.