\section{Introduction}
\label{sec:intro}

AI coding agents are rapidly moving toward autonomous scientific research~\citep{karpathy2026autoresearch,lu2026towards,gottweis2026accelerating}, from reproducing published implementations to conducting end-to-end research workflows.
As these systems begin to target real scientific problems, rigorous evaluation becomes critical: without reliable benchmarks, it is impossible to tell whether an agent is genuinely advancing the state of the art or merely fitting familiar patterns to new data.

However, existing benchmarks for evaluating agent capabilities on scientific research have several limitations.
Paper-based benchmarks~\citep{starace2025paperbench,siegel2024corebench,wang2026firebench} measure whether an agent can re-implement a published method, but stop short of the more consequential question: can an agent \emph{discover} a competitive method on its own?
Engineering-optimization benchmarks~\citep{shern2024mlebench,rank2026posttrainbench,nathani2025mlgym,qiang2026mle} target Kaggle competitions or post-training tasks, which do not require the domain reasoning, specialized tooling, or cross-discipline knowledge that characterize research in the natural sciences, and suffer from environment fragmentation that makes independent re-running fragile.
Credibly evaluating whether autonomous research agents can advance the frontier of AI-for-Science requires a benchmark that is both challenging and bidirectional. It must test \emph{discovery}, whether an agent can devise methods that surpass the published state of the art, on \emph{genuine scientific problems} drawn from the natural sciences rather than on engineering proxies.

We present \textbf{\benchmark}, a cross-discipline benchmark of 90 tasks distilled from peer-reviewed Nature-family publications, designed to evaluate whether AI coding agents can move beyond reproduction toward discovery.
\benchmark simultaneously extends both axes: the \textbf{PaperBench axis} from Understanding $\to$ Coding to Discovery, and the \textbf{PostTrainBench axis} from Engineering Optimization to Science.
It is built on \textbf{NatureGym}, an automated pipeline that converts a published paper into a containerized task package comprising a task brief, the paper's dataset, a held-out test set with hidden ground truth, and an automated evaluator, addressing the environment-fragmentation problem in prior benchmarks.
We collect approximately $5{,}500$ papers from ten Nature-family journals published between 2022 and 2025 and apply a three-stage build-then-verify pipeline to yield the final $90$ task packages (Figure~\ref{fig:gym-pipeline}).
An information firewall removes the source method from each package, so agents must discover solutions rather than reproduce them.
The benchmark spans six scientific task domains (cellular omics, protein biology, biomedical modeling, physical modeling, molecular design, and relational reasoning) and uses a SOTA-normalized relative gap $g$ as the primary metric, supplemented by a post-hoc validity judge that detects shortcut behaviors such as output fabrication and feedback gaming.

We evaluate ten agents spanning three coding-agent harnesses (Claude Code, Codex CLI, Gemini CLI) and ten frontier models under a strict web-search-disabled protocol as shown in Figure~\ref{fig:intro-main-results}.
The strongest agent, Claude Opus 4.7, surpasses the published SOTA ($g > 0.1$) on only $17.8\%$ of tasks and matches it on $47.8\%$.
Analysis of the $900$ task--agent runs reveals that success is driven primarily by methodological translation, where agents convert scientific tasks into familiar supervised-prediction problems, accounting for $45.5\%$ of validated successes, rather than by scientific invention.
Failures are dominated by wrong method choice ($45.1\%$) and insufficient compute budget ($24.4\%$), not by task misunderstanding.
Our contributions are as follows:
\begin{itemize}[leftmargin=1em, itemsep=2pt, topsep=4pt]
    \item \textbf{NatureGym}, an automated pipeline that constructs reproducible, containerized per-task environments from Nature-family papers, addressing the environment-fragmentation problem that has limited the credibility of prior agent-on-research benchmarks.
    \item \textbf{NatureBench}, a benchmark of $90$ Nature-sourced tasks across six scientific task domains with a Discovery-oriented evaluation protocol (Surpass-SOTA, Match-SOTA, validity judge) that separates genuine algorithmic progress from engineering optimization and shortcut-taking.
\end{itemize}
