\section{Conclusion}
In this paper, we introduce Qwen-Image-VAE-2.0, a suite of high-compression image VAEs ($f16$ and $f32$) designed to overcome the long-standing bottlenecks in native high-resolution image synthesis. Our work demonstrates a clear technical path to resolving the fundamental tripartite trade-off between high compression ratio, reconstruction fidelity, and downstream diffusability: by leveraging expanded latent channel dimensions to compensate for spatial information loss caused by high compression, while simultaneously applying advanced semantic alignment techniques to ensure the latent space remains suitable for diffusion modeling. To achieve state-of-the-art reconstruction fidelity in high-compression regimes, we adopt an improved VAE architecture featuring Global Skip Connections (GSC), establishing a direct path for fine-grained detail recovery. This is bolstered by a comprehensive data strategy where we scale the training corpus to billions of images and utilize a specialized synthetic document-rendering pipeline to ensure the legibility of dense text—a traditional failure point for high-compression models. To address the challenge of diffusability inherent in high-dimensional latent space, we introduce an improved semantic alignment strategy. By aligning latent representations with middle-layer feature of DINOv2, we significantly accelerate the convergence of downstream DiT. Finally, to ensure practical utility, we implement a light-encoder paradigm and an attention-free backbone, maintaining high throughput and minimal encoding overhead even at ultra-high resolutions. Extensive evaluations on public benchmarks and OmniDoc-TokenBench demonstrate that Qwen-Image-VAE-2.0 not only preserves exceptional structural and textual integrity but also facilitates efficient generative modeling. These models provide a robust foundation for the next generation of visual synthesis systems, marking a significant milestone in efficient image generation.