\section{Data}
\subsection{Data Collection}
\paragraph{Scaling Data to Billion Scale.} To ensure robust generalization across diverse domains, we scale the VAE training corpus to encompass billions of images. This large-scale dataset covers a wide spectrum of visual content, spanning various categories, resolutions and aspect ratios. However, data at this scale inevitably contains noise like edge blur and compression artifacts, which impedes model's ability to learn high-frequency detail. To mitigate this, we employ clarity and blur filters to prune low-quality samples, ensuring that the VAE is supervised by high-fidelity signals.

\paragraph{Text-Rich Image Collection.} To address the reconstruction bottleneck in text-rich scenarios, we adopt a two-fold strategy. First, we leverage an OCR filter to identify and prioritize samples with high character density from extensive real-world datasets. Second, we curate a specialized document corpus, which includes screenshots of academic papers, presentation slides, posters, and complex web pages. By training on these real-world text-rich images, our models learn to prioritize the preservation of sharp edges of characters and semantic structures, enabling legible text reconstruction that is challenging for high compression VAEs.

\subsection{Data Synthesis}
To further enhance character-level reconstruction, we develop a synthetic pipeline that renders text documents into images. Our pipeline supports both alphabetic (English) and logographic (Chinese) languages, accounting for their distinct stroke densities and complexities. Crucially, we observed models trained on background-free synthetic data (e.g., black text on white backgrounds) generalize poorly to real-world images where text is often overlaid on complex textures. To bridge this gap, we implement background-contained synthesis, where text is rendered onto backgrounds randomly sampled from general-domain images. Moreover, to adapt different compression settings, we construct synthetic datasets of varying difficulty by rendering characters ranging from 5 to 20 pixels. This multi-granularity supervision forces the VAE to capture fine detail, ensuring legibility even at $f32$ compression.
