\section*{Limitations and Future Work}
\label{app:limitations}

Although OpenComputer is designed around executable, hard-coded verification, not every realistic desktop task can be fully reduced to reliable programmatic checks. Some generated tasks require visual or geometric judgments that are difficult to express using application state alone. For example, in Draw.io, a verifier can often inspect the existence of shapes, labels, and connector objects, but it may be difficult to determine with high confidence whether an arrow visually and semantically connects two specific boxes in the intended way without inspecting the rendered screenshot. Similar cases arise when the desired outcome depends on spatial layout, visual alignment, or other presentation-level properties that are only partially exposed through file formats or application APIs.

When a generated task contains criteria that cannot be reliably checked by a hard-coded verifier, we mark those criteria as requiring LLM-based visual judgment rather than treating them as fully programmatic rewards. However, to keep the official benchmark auditable and reproducible, we exclude such tasks from the main benchmark and from all reported evaluation results. In the current task-generation process, we found 17 generated tasks with at least one success criterion that could not be fully verified by hard-coded checkers; these tasks were retained only for diagnostic analysis and were not included in the finalized OpenComputer benchmark. We will release these tasks in the repository, together with the procedure used to identify visually grounded criteria and the LLM-as-judge pipeline used for analysis. This provides a controlled starting point for future work on hybrid verification, where executable state checks can be combined with visual judgments for desktop tasks whose success depends on layout, geometry, or rendered appearance.