\section{Methodology}

\textit{DomainShuttle} is designed to advance flexible and high-fidelity cross-domain personalization, which consists of three modules: Domain-MoT, Video-Reference DualRoPE, and Cross-Pair Consistent Loss, as illustrated in Fig. \ref{fig:method_overview}. During training,  video latents and reference image features are processed through separate branches in \textbf{DomainMoT} to achieve feature disentanglement, facilitating the incorporation of domain-aware AdaLN modulation in the reference branch. We then introduce \textbf{Video-Reference DualRoPE}, which assigns reference image tokens to a RoPE space independent of video tokens, enabling precise subject-level control: different subjects are explicitly separated, while representations of the same subject are pulled closer in the latent space. Finally, the \textbf{Cross-Pair Consistent Loss} (CCL) aligns two sets of reference images corresponding to the same video to precisely capture subject-specific features. In this section, we first introduce the preliminaries of video generation models. We then present the design of DomainMoT, Video-Reference  DualRoPE, and the Cross-Pair Consistent Loss in detail. Finally, we describe the construction of our training dataset.






\subsection{Preliminaries} 

Our method is trained on the DiT-based video generation model \cite{wan2025wan}. During training, the text encoder  $\boldsymbol{\mathcal{E}}_t(\cdot)$ \cite{chung2023unimax} encodes the text into text feature $\boldsymbol{c}_t$, and the 3D VAE encoder  $\boldsymbol{\mathcal{E}}_v(\cdot)$ encodes the video and $N$ reference images $\mathcal{R} = \{ I_0, I_1, \cdots, I_{N-1} \}$ into video latents $\boldsymbol{z}_1$ and reference image features $\boldsymbol{c}_r$. The model is optimized by flow-matching loss \cite{lipman2023flow}, which is defined as follows:

\begin{equation}
\mathcal{L}_{\mathrm{FM}}=\mathbb{E}_{t,\boldsymbol{z}_0,\boldsymbol{z}_1}\|{G}_\theta(\boldsymbol{z}_t,t,\boldsymbol{c}_t,\boldsymbol{c}_r)-(\boldsymbol{z}_1-\boldsymbol{z}_0)\|_2^2,
\label{eq:flow_matching_loss}
\end{equation}
where continuous time $t\in[0,1]$. $\boldsymbol{z}_0$ denotes a sample drawn from the prior distribution, and $\boldsymbol{z}_1$ denotes video latent encoded by 3D VAE. ${G}_\theta$ represents a learnable vector field parameterized by $\theta$.

\subsection{Model Architecture} 
\subsubsection{Domain-MoT}


A key issue in cross-domain S2V is the entanglement between intrinsic subject features and domain-specific attributes, which makes it difficult to preserve subject features while enabling flexible transitions across domains. To address this issue, we propose Domain-MoT, as shown in Fig. \ref{fig:method_overview} (a). Specifically, Domain-MoT decouples video latents and reference image features into two independent processing paths, and explicitly injects domain attributes into the reference image branch via Domain-aware AdaLN to distinguish different domains in the feature space, achieving more precise cross-domain S2V.

As shown in Fig. \ref{fig:method_overview} (a), the video latents and reference image features after 3D VAE encoding are separately patchified to extract their patch embeddings $\boldsymbol{f}_v$ and $\boldsymbol{f}_r$, which facilitates the subsequent integration of the in-context self-attention and Domain-aware AdaLN, achieving more stable reference injection.

Next, Domain-MoT employs in-context self-attention with independent QKV projections and independent RoPE for video latents and reference features. This decoupled design preserves the inherent capability of the video branch as the base model, while allowing the reference image branch to focus on extracting more precise subject features for flexible and high-fidelity personalization. The in-context self-attention is shown below:


\begin{equation}
\label{eq:attention}
 \text{Softmax}\left( \frac{ [ R_{v}(  \boldsymbol{Q}_{v}); R_{r}(\boldsymbol{Q}_{r})] \cdot [ R_{v}(  \boldsymbol{K}_{v}); R_{r}(\boldsymbol{K}_{r})]}{\sqrt{d}} \right)  [\boldsymbol{V}_{v},\boldsymbol{V}_{r}], 
\end{equation}
Where $ R_{v}$ and $ R_{r}$ denote the RoPE applied to the video branch and reference image branch, respectively, which will be discussed in the next subsection. The queries $\boldsymbol{Q}_v=\boldsymbol{W}^q_v\cdot \boldsymbol{f}_v$ and $\boldsymbol{Q}_r=\boldsymbol{W}^q_r \cdot \boldsymbol{f}_r$,  keys  $\boldsymbol{K}_v=\boldsymbol{W}^k_v \cdot \boldsymbol{f}_v$ and $\boldsymbol{K}_r=\boldsymbol{W}^k_r \cdot \boldsymbol{f}_r$, values $\boldsymbol{V}_v=\boldsymbol{W}_v^v \cdot \boldsymbol{f}_v$ and $\boldsymbol{V}_r=\boldsymbol{W}^v_r \cdot f_r$. $\boldsymbol{W}^q$,  $\boldsymbol{W}^k$ and $\boldsymbol{W}^v$ are weight parameters. \( [; ] \) denotes feature concatenation. %rope的描述将在Sec. 2.3.2节中描述。


To preserve intrinsic text guidance capability of the base model across complex cross-domain scenarios, we freeze the textual cross-attention during training. Textual cross attention enables the interaction between textual features $\boldsymbol{f}_t$  and concatenated visual features $\boldsymbol{f}_c = [\boldsymbol{f}_v; \boldsymbol{f}_r]$, as follows:
\begin{equation}
\label{eq:attention2}
 \text{Softmax}\left( \frac{   \boldsymbol{Q}_{c} \cdot \boldsymbol{K}_{t}}{\sqrt{d}} \right)  \boldsymbol{V}_{t}, 
\end{equation}
where the query $\boldsymbol{Q}_c=\boldsymbol{W}^q\cdot \boldsymbol{f}_c$,  key  $\boldsymbol{K}_t=\boldsymbol{W}^k \cdot \boldsymbol{f}_t$ and value $\boldsymbol{V}_t=\boldsymbol{W}^v \cdot \boldsymbol{f}_t$. 


\paragraph{\textbf{Domain-aware AdaLN.}} We then introduce Domain-aware AdaLN, a novel mechanism that aims to achieve flexible cross-domain generation by decoupling domain attributes. Existing S2V methods typically modulate reference tokens and video latents indiscriminately, causing domain attributes from the reference image to be entangled with video latents. Domain-aware AdaLN addresses this by structurally decoupling the noise AdaLN and reference AdaLN. Notably, the reference AdaLN is modulated by both the reference domain attributes and time features, while the noise AdaLN is modulated only by time features.

The explicit injection mechanism of reference domain attributes decouples features of content and domain, enabling cross-domain generation without disturbing the content and temporal structure by simply exchanging domain attributes. On the one hand, in in-domain generation scenarios, this mechanism can utilize in-domain prior knowledge to improve generation quality; on the other hand, in cross-domain scenarios, replacing the injected domain attributes achieves better cross-domain generation results. Specifically, the AdaLN mechanisms for video latent and reference image features are as follows:


\begin{equation}
\left\{
\begin{aligned}
& \hat{\boldsymbol{f}}_v \, =  \, g_{v}(t) \,\odot\, \big[ \mathcal{F}\big( \mathrm{LN}( \boldsymbol{f}_v) \,\odot\, (1+\gamma_{v}(t))  \;+\; \beta_{v}(t) \big)\big] + \boldsymbol{f}_v , \\
& \hat{\boldsymbol{f}}_r \, = \, g_{r}(t,a) \,\odot\, \big[ \mathcal{F} \big( \mathrm{LN}( \boldsymbol{f}_r) \,\odot\,  (1+\gamma_{r}(t,a))  \;+\; \beta_{r}(t,a)\big) \big] +\boldsymbol{f}_r  ,
\end{aligned}
\right.
\end{equation}
where $t$ denotes time features and $a \in  \{A_1, A_2, A_3, \ldots, A_K\}$ denotes one of the $K$ domain attributes. $\mathrm{LN}$ denotes layer normalization and  $\odot$ denotes Hadamard product.  $\hat{\boldsymbol{f}}_v$ and $\hat{\boldsymbol{f}}_r$ denote the modulated video features and reference features, respectively. The modulation coefficients are given by the scale $\gamma \in \mathbf{R}^{d}$, shift $\beta \in\mathbf{R}^{d}$, and residual gate $g \in \mathbf{R}^{d}$ conditioned on $t$ and $a$. $\mathcal{F}(\cdot)$ denotes general residual functions (\eg, attention and FFN).




\subsubsection{ Video-Reference DualRoPE}
\label{3.2.2}

Currently, DiT-based video generation models \cite{wan2025wan,yang2024cogvideox} commonly adopt Rotary Positional Encoding (RoPE) to distinguish positional information among video tokens. RoPE assigns a positional index to each token, modulating the interaction strength between different tokens, which means that tokens with closer indices correspond to shorter latent distances. Specifically, each video token is assigned a positional index $(i,j,k)$, where $i\in[0,f-1]$, $j\in[0,h-1]$, and $k\in[0,w-1]$. Here, $f$, $h$, and $w$ denote the number of frames, height, and width in the video latent space, respectively.



Existing methods mainly inherit the RoPE mechanism of the base model, applying the reference image RoPE by treating each reference image as an additional video frame. This mechanism distinguishes different reference images solely through temporal indices, ignoring that multiple reference subjects lack temporal continuity and that multiple reference images may jointly describe a single subject. Notably, a reference subject is not equivalent to a single reference image, as one subject can correspond to multiple reference images. To explicitly disentangle video tokens from different reference subjects, we propose \textbf{Video-Reference (VR) DualRoPE}, which allocates reference image tokens into a RoPE space fully decoupled from the video token space, enabling precise RoPE spatial distance relationships among different reference subjects, as shown in Fig. \ref{fig:method_overview} (b). Through this design, the model achieves joint optimization of subject consistency and text controllability in open-domain scenarios. The video noise RoPE $R_v(i,j,k)$ and the reference image RoPE $R_r(i,j,k)$ are defined as follows:
\begin{equation}
\left\{
\begin{aligned}
& R_v(i,j,k) = \theta (i+1,j,k)  , \\
&R_r(i,j,k) = \theta(0,j+h\times(m+1),k+w\times(n+1)),
\end{aligned}
\right.
\end{equation}
where $m \in [0, M-1] $ denotes the $m$-th reference subject, $n \in [0, N-1]$ denotes the $n$-th reference image, and $\theta$ represents the rotation function. For reference images, the temporal index is set to 0 while the temporal index for video starts from 1, explicitly separating the reference RoPE space from the video RoPE space. For two adjacent reference images representing different subjects, the RoPE offset is $\Delta=(0,h,w)$. When two reference images represent different parts of the same subject, they are treated as sub-images of a large reference image, whose offset is set to $\Delta=(0,0,w)$. This design distinguishes semantic differences between different reference images and keeps images of the same subject closer in the latent space, thus explicitly establishing their identity associations.






\subsubsection{ Cross-Pair Consistent Loss}

To further enhance the flexibility of our model in various complex cross-domain scenarios, we propose a \textbf{Cross-pair Consistent Loss} (CCL) to precisely extract key features from references.

We construct multiple sets of reference images for each video, forming a reference pool (see Fig. \ref{data_pipeline} for dataset details). During training, two sets of reference images are randomly sampled from the pool, and the generated video latent noises are aligned at the same timestep as shown in Fig. \ref{fig:method_overview} (c). This encourages the model to extract shared features (\eg, shape, texture style, and subject identity) from different references and suppresses overfitting to redundant features in a single frame. Different reference sets vary in viewpoint, occlusion, motion blur, and illumination. By forcing a learnable branch to match a frozen reference branch, the model learns representations that are insensitive to these perturbations, thereby improving flexibility in complex cross-domain scenarios. Compared to randomly sampling a single reference set at different timesteps during training, our strategy achieves more precise consistency alignment at the same noise level, leading to more effective extraction of precise subject features in the reference images. The Cross-Pair Consistent Loss $\mathcal{L}_{\mathrm{C}}$ is defined as:

\begin{equation}
\mathcal{L}_{\mathrm{C}}=\|{G}_\theta(\boldsymbol{z}_t,t,\boldsymbol{c}_t,\boldsymbol{c}_r)-{G}^*_\theta(\boldsymbol{z}_t,t,\boldsymbol{c}_t,\boldsymbol{c}^*_r)\|_2^2,
\label{eq:loss_2}
\end{equation}
where $\boldsymbol{c}_r$ and $\boldsymbol{c}^*_r$ represent two different sets of reference image features. The ${G}^*_\theta$ branch is frozen, while the ${G}_\theta$ branch is trainable.




\subsection{Training Data Pipeline}
\label{data_pipeline}
Our training data comprises two components: image and video personalization datasets. We build a 200K image dataset containing multiple subjects based on open-source datasets \cite{UNO,echo-4o} to endow the model with basic personalization capability. For video personalization, we use three datasets: Phantom-Data \cite{phantom-data}, OpenS2V \cite{yuan2025opens2v}, and Ditto-1M \cite{zhang2025region}. We first filter low-quality videos using aesthetics and motion metrics. Phantom-Data naturally provides numerous cross-pairs after retrieval, enabling direct application of Cross-Pair Consistent Loss on its cross-pair set. For OpenS2V and Ditto-1M, we utilize Grounding-DINO \cite{groundingdino} for object detection and SAM2 \cite{sam2} for multi-frame segmentation to build a reference image set for Cross-Pair Consistent Loss. We then use visual-semantic alignment via MLLM \cite{qwen25} to remove low-quality references (\eg, incomplete, blurred, or semantically mismatched subjects). Notably, Ditto-1M is a video editing dataset in which segmented reference images can align with multiple edited and source videos, facilitating ``single reference set → multiple videos'' pair construction, which further promotes the extraction of precise subject features. 

In total, we obtain a 750K high-quality, open-domain video personalization dataset covering various scenarios (\eg, humans, objects, fantasy subjects, and backgrounds) and supporting cross-pair configurations of both ``multiple reference set → single video'' and ``single reference set → multiple videos''.  In the training dataset, only 50K from the Ditto-1M dataset. These 50K samples include 25K reference-image/original-video pairs and \textbf{25K} reference-image/edited-video pairs, so edited videos only account for \textbf{3.3\%} of the total data. This 3.3\% video editing data is not used for main supervision, but only as data augmentation for cross-domain scenarios. Original-video/edited-video pairs are not used for training. We conduct an ablation study on whether to use Ditto-1M, as shown in Tab. \ref{video_editing} of the supplementary material B.

