\section{Related Work}

\subsection{Turn-based Models and Products}
\label{sec:product}
Almost all of today's models are \textit{turn-based}: they act only when a user takes a turn, answering once addressed and otherwise idle. This is a structural property, not a matter of speed. A turn-based system has no mechanism for reacting to an event in the world that arrives with no accompanying user utterance, however quickly it can reply once spoken to. Two families of recent systems push hard on responsiveness while keeping this turn-based core.

\paragraph{\textbf{Realtime and omni models.}} OpenAI's GPT-Realtime-2 reasons inside a single speech-to-speech loop \cite{openai2026gptrealtime2}, and Alibaba's Qwen3.5-Omni \cite{qwen2026omni} is a natively pretrained omni-modal model (text, audio, image, video) with real-time streaming and built-in turn-taking and interruption handling, the latter openly released. These remove the latency of cascaded stacks, but what they optimize is \textit{conversational turn-taking}: how quickly and naturally they answer once the user has spoken. Their interaction is organized around dialogue and waits for the user's turn; they are not built to watch a visual stream and decide, unprompted, that this is the moment to speak. Being open, real-time, and multimodal (as Qwen3.5-Omni already is) is therefore necessary but not sufficient for the setting we target.

\paragraph{\textbf{Consumer video-call products.}} In-app ``video call'' features are the most familiar real-time multimodal assistants and our most recognizable baseline, yet their cores remain turn-based. Doubao's runs on ByteDance's Volcano Engine conversational-AI stack, a configurable ASR/VLM/TTS pipeline with stop/turn detection, and its documentation makes the mechanism plain: while no one is speaking, the server only caches incoming video frames and does not send them to the model, so the application must periodically fire an \texttt{ExternalTextToLLM} trigger to obtain any analysis \cite{volcengine2026videounderstanding}; this characterization is based on the publicly available version of the stack, which may differ from the deployed product. ``Monitoring'' is thus an external clock bolted onto a turn-based model: reaction to an on-screen event waits for the next trigger and can never beat the polling interval. Gemini's video call goes less far still; in our use it answers only about the frame visible the instant a question is asked. In neither case does the decision of when to respond live in the model.

% \subsection{Turn-based Models and Products}
% Almost all of today's models are \textit{turn-based}: they act only when a user takes a turn---answering once addressed and otherwise idle. This is a structural property, not a matter of speed; a turn-based system has no mechanism for reacting to an event in the world that arrives with no accompanying user utterance, however quickly it can reply once spoken to. Two families of recent systems push hard on responsiveness while keeping this turn-based core.

% \paragraph{\textbf{ realtime and omni models.}} OpenAI's GPT-Realtime-2 reasons inside a single speech-to-speech loop \cite{openai2026gptrealtime2} , and Alibaba's Qwen3.5-Omni \cite{qwen2026omni} is a natively pretrained omni-modal model (text, audio, image, video) with real-time streaming and built-in turn-taking and interruption handling, the latter openly released. These remove the latency of cascaded stacks, but what they optimize is \textit{conversational turn-taking}---how quickly and naturally they answer once the user has spoken. Their interaction is organized around dialogue and waits for the user's turn; they are not built to watch a visual stream and decide, unprompted, that this is the moment to speak. Being open, real-time, and multimodal---as Qwen3.5-Omni already is---is therefore necessary but not sufficient for the setting we target.

% \paragraph{\textbf{Consumer video-call products.}} In-app ``video call'' features are the most familiar real-time multimodal assistants and our most recognizable baseline, yet their cores remain turn-based. Doubao's runs on ByteDance's Volcano Engine conversational-AI stack---a configurable ASR/LLM/TTS pipeline with stop/turn detection---and its documentation makes the mechanism plain: while no one is speaking, the server only caches incoming video frames and does not send them to the model, so the application must periodically fire an \texttt{ExternalTextToLLM} trigger to obtain any analysis \cite{volcengine2026videounderstanding}. ``Monitoring'' is thus an external clock bolted onto a turn-based model: reaction to an on-screen event waits for the next trigger and can never beat the polling interval. Gemini's video call goes less far still---in our use it answers only about the frame visible the instant a question is asked. In neither case does the decision of when to response live in the model.

% \subsection{Interaction Models}
% A distinct, newer line breaks with the turn-based assumption by moving interactivity \textit{into the model}.

% \paragraph{TML interaction models.} Thinking Machines Lab recently named this an \textit{interaction model} and argued---as we independently and concurrently concluded---that scalable interactivity should come from the model itself rather than an external harness of turn detectors and activity heuristics. Their model handles audio, video, and text in real time and can speak unprompted, acting on the ongoing stream rather than waiting for a user's turn; it was released as a research preview \cite{thinkingmachines2026interactionmodels}. That two groups arrived at this conviction at nearly the same moment we read as a sign that the move from turn-based to interactive models is a direction whose time has come. 

% \paragraph{\textbf{TML interaction models.}} Thinking Machines Lab recently named this an \textit{interaction model} and argued---as we independently and concurrently concluded---that scalable interactivity should come from the model itself rather than an external harness of turn detectors and activity heuristics. Their model handles audio, video, and text in real time and can speak unprompted, acting on the ongoing stream rather than waiting for a user's turn, and---as in our system---it is paired with an asynchronous background model for slower, harder reasoning, the two sharing context; it was released as a research preview \cite{thinkingmachines2026interactionmodels}. That two groups arrived at this approach at nearly the same moment we read as a sign that the move from turn-based to interactive models is a direction whose time has come. Within this shared interaction--background design, our aesthetic differs in two deliberate ways. On modality, rather than fusing audio and vision into the model alike, we keep speech (ASR/TTS) a pluggable module and make vision the model's intrinsic, primary modality for proactive interaction---fitting a watch-and-react setting in which voice is interchangeable I/O. On scale, TML-Interaction-Small is a 276B-parameter mixture-of-experts model (12B active) that its authors describe as the smallest they can serve at the required latency, whereas we deliberately choose a compact $\sim$8B model---balancing interaction-grade responsiveness, local and low-cost deployability, and capability---so that it can be run, fine-tuned, and reproduced widely rather than only on large infrastructure.

% \paragraph{\textbf{Full-duplex speech.}} For speech, Kyutai's Moshi \cite{kyutai2024moshi} is the prominent open instance: a real-time, full-duplex speech--text model in which both sides can speak and listen at once rather than strictly alternating, and the closest open precedent for putting interaction in the model. Its modality and scope---spoken conversation---set it apart from the visual, event-driven setting we address.

% \subsection{Streaming Video Understanding}
% A research line lets models process video as it arrives rather than after the fact, spanning streaming video LLMs, proactive or anticipatory response, and long-horizon video memory [cite, cite, cite]. It is closest to our model contribution, but three gaps separate it from a deployable interaction model. It typically advances one property at a time---responsiveness, \textit{or} proactivity, \textit{or} memory---rather than the three together. It is usually measured on offline benchmarks, at times even in the manner of offline video understanding, which never tests reaction to live events under real-time constraints. And it stops at the model, without the surrounding system---serving, memory, transduction, delegation---that hours of sustained real-time presence demand.

% \subsection{Our Position.} Read along two axes---\textit{how} a model interacts and \textit{what} is released---JoyAI-VL-Interaction occupies a cell no prior work does. Against turn-based systems, native or product, the decision of when to act lives inside our model and is taken every second, event-driven, so reaction is bounded by inference rather than by a user's turn or a trigger clock. Against TML's interaction model, we make \textbf{vision} the first-class driver---the watch-and-react setting rather than audio--video conversation---and we open-source not a research preview but the model, its training recipe, and a complete, composable, deployable system. Against Moshi, we are vision-driven rather than speech-centric, reacting to events that carry no conversational turn at all. And against streaming-video research, we bring responsiveness, proactivity, and memory together in one model and run it inside a system built for sustained presence. To our knowledge, JoyAI-VL-Interaction is the first open, vision-driven interaction model released together with a complete deployable system---the point where all of these properties meet, rather than any one alone.

\subsection{Interaction Models}
A distinct, newer line breaks with the turn-based assumption by moving interactivity into the model.
\paragraph{\textbf{TML interaction models.}} Thinking Machines Lab recently named this an \textit{interaction model} and argued, as we independently and concurrently concluded, that scalable interactivity should come from the model itself rather than an external harness of turn detectors and activity heuristics. Their model handles audio, video, and text in real time and can speak unprompted, acting on the ongoing stream rather than waiting for a user's turn, and, as in our system, it is paired with an asynchronous background model for slower, harder reasoning, the two sharing context; it was released as a research preview \cite{thinkingmachines2026interactionmodels}. That two groups arrived at this approach at nearly the same moment we read as a sign that the move from turn-based to interactive models is a direction whose time has come. Within this shared interaction--background design, our aesthetic differs in two deliberate ways. On modality, rather than fusing audio and vision into the model alike, we keep speech (ASR/TTS) a pluggable module and make vision the model's intrinsic, primary modality for proactive interaction, fitting a watch-and-interact setting in which voice is interchangeable I/O. On scale, TML-Interaction-Small is a 276B-parameter mixture-of-experts model (12B active) that its authors describe as the smallest they can serve at the required latency, whereas we deliberately choose a compact $\sim$8B model (balancing interaction-grade responsiveness, local and low-cost deployability, and capability) so that it can be run, fine-tuned, and reproduced widely rather than only on large infrastructure.
\paragraph{\textbf{Full-duplex speech.}} For speech, Kyutai's MoshiRAG \cite{chien2026moshiragasynchronousknowledgeretrieval} is the prominent open instance: a real-time, full-duplex speech--text model in which both sides can speak and listen at once rather than strictly alternating, and the closest open precedent for putting interaction in the model. Its modality and scope (spoken conversation) set it apart from the visual, event-driven setting we address.


\subsection{Streaming Video Understanding}
A research line lets models process video as it arrives rather than after the fact, spanning streaming video LLMs, proactive response \cite{xia2025streamingvideoinstructiontuning,qian2025dispider,yan2026proactvl}, real-time inference \cite{xu2025streamingvlmrealtimeunderstandinginfinite,wang2026acceleratingstreamingvideolarge}, and long-horizon video memory \cite{yao2025timechatonline,zhang2026hermeskvcachehierarchical,zeng2026streamforest}. It is closest to our model contribution, but three gaps separate it from a deployable interaction model. It typically advances one property at a time (responsiveness, \textit{or} proactivity, \textit{or} memory) rather than the three together. It is usually measured on offline benchmarks, at times even in the manner of offline video understanding, which never tests reaction to live events under real-time constraints. And it stops at the model, without the surrounding system (serving, memory, transduction, delegation) that hours of sustained real-time presence demand. 
% 丁钰师妹辛苦了！

\subsection{Our Position.} Read along two axes, \textit{how} a model interacts and \textit{what} is released, JoyAI-VL-Interaction occupies a cell no prior work does. Against turn-based systems, native or product, the decision of when to act lives inside our model and is taken every second, event-driven, so reaction is bounded by inference rather than by a user's turn or a trigger clock. Against TML's interaction model, we make vision the first-class driver, the watch-and-interact setting rather than audio--video conversation, and we open-source not a research preview but the model, its training recipe, and a complete, composable, deployable system. Against MoshiRAG, we are vision-driven rather than speech-centric, reacting to events that carry no conversational turn at all. And against streaming-video research, we bring responsiveness, proactivity, and memory together in one model and run it inside a system built for sustained presence. To our knowledge, JoyAI-VL-Interaction is the first open, vision-driven interaction model released together with a complete deployable system: the point where all of these properties meet, rather than any one alone.