% \section{Conclusion}

\section{Conclusion}\label{sec:conclusion}
We set out to move large models past the turn-based paradigm, in which a model opens its eyes only when it is addressed, toward genuine streaming interaction, in which the model is present in the world and decides for itself when to act. To that end we release, to our knowledge, the first open vision-driven interaction model, together with everything needed to build on it: the time-aligned data, the training recipe, the 8B model, and a complete, deployable system.

At the center is a single conviction, that interactivity should scale as a capability of the model itself rather than be simulated by a harness around it. Scaling interactivity means scaling three things at once: the model's freedom to speak on its own when a moment is worth a word, its ability to interact in real time, and its sense of elapsed time. As these grow, a model behaves less like a tool that waits to be queried and more like a participant that is simply present, watching what is happening and responding as a person would. We see this as a direction of scaling in its own right, alongside the scaling of intelligence, and one that brings models closer to being genuinely present in human settings.

Across the streaming scenarios we study, we already see early signs that an interaction model carries natural advantages for real deployment, from monitoring and live narration to step-by-step guidance and companionship. The moment we are reaching for is an everyday one: you come home worn out after a long day, and before you have said anything, a quiet voice notices and offers, ``I can see you're tired; today must have been hard on you.'' Presence like that, given unasked, is what an interaction model makes possible and a turn-based one, waiting to be addressed, never can. Much remains open in both the model and the system, and we have released the whole stack openly to lower the barrier and accelerate this shift. Turn-based models, products, and optimizations have been built up over many years, while vision-driven interaction models are only beginning; if this release opens even a first foothold for them among those long-settled peaks, and draws others into building it with us, it will have done its job. We invite the community to explore, with us, what a model that is truly present in the world can become. 