\section{Conclusion}

In this paper, we study A2UI-based Generative UI as a unified problem in which the assistant produces both natural language and executable UI actions under a fixed declarative protocol. To support this setting, we construct an A2UI-grounded corpus from heterogeneous dialogue sources, build A2UI-Bench for controlled evaluation, and develop a two-stage training pipeline with schema-light SFT and reward-driven RL. Experiments on both 30B and 235B models show that this training recipe improves protocol correctness, interaction quality, and user experience by a large margin. The best 235B model performs strongly in the minimal-prompt setting and slightly surpasses the strongest full-prompt frontier baseline in overall score. These results show that Generative UI capability does not have to depend on heavy schema prompting at inference time, and can instead be learned and internalized through training.

This project is a key step toward bringing Generative UI into real production environments, but it still has clear limitations. The underlying A2UI protocol is still evolving and currently remains at version 0.8. Model capability is also a bottleneck, especially for complex multi-turn interaction and user experience. In addition, latency can become an issue when the model must generate, validate, and render structured UI in real time. In future work, we will continue to explore more general, flexible, and token-efficient ways to build Generative UI systems. We release our models, benchmark, and evaluation protocol to support broader research and development in this direction.