\beginappendix

\lstdefinestyle{svpromptcode}{
  basicstyle=\ttfamily\scriptsize,
  breaklines=true,
  columns=fullflexible,
  keepspaces=true,
  frame=single,
  framerule=0.25pt,
  rulecolor=\color{memblue!45},
  backgroundcolor=\color{white},
  xleftmargin=0mm,
  xrightmargin=0mm,
  aboveskip=0.5em,
  belowskip=0.5em,
  literate={─}{{\textSFx}}1 {│}{{\textSFxi}}1 {└}{{\textSFii}}1 {├}{{\textSFviii}}1
}

\newtcolorbox{svpromptbox}[1]{
  colback=memblue2,
  colframe=memblue,
  colbacktitle=memblue,
  coltitle=white,
  arc=1pt,
  boxrule=0.8pt,
  fonttitle=\bfseries\sffamily,
  fontupper=\small,
  title={#1},
  left=6pt,
  right=6pt,
  top=6pt,
  bottom=6pt,
  before skip=0.9em,
  after skip=1.0em,
  breakable,
  width=\linewidth
}

\newcommand{\svpromptheading}[1]{\par\smallskip\noindent{\sffamily\bfseries\color{memblue}#1}\par}
\newcommand{\svpromptsubheading}[1]{\par\smallskip\noindent{\sffamily\bfseries #1}\par}
\newcommand{\svinlinecode}[1]{{\ttfamily #1}}
\newcommand{\svpartsep}{\par\medskip\noindent{\color{memblue!45}\leaders\hbox{\rule[0.45ex]{5pt}{0.35pt}\hspace{3pt}}\hfill\kern0pt}\par\medskip}
\newlength{\svtracelabelwidth}
\setlength{\svtracelabelwidth}{6.8em}
\newcommand{\svtrace}[2]{%
  \par\smallskip\noindent
  \begin{minipage}[t]{\svtracelabelwidth}#1\end{minipage}%
  \hspace{0.6em}%
  \begin{minipage}[t]{\dimexpr\linewidth-\svtracelabelwidth-0.6em\relax}#2\end{minipage}%
  \par
}
\newcommand{\svcasebenchmark}[1]{\par\noindent{\sffamily\scriptsize\bfseries\color{memblue}Benchmark: #1}\par\smallskip}

\setlist[itemize]{leftmargin=*, itemsep=2pt, topsep=2pt}
\setlist[enumerate]{leftmargin=*, itemsep=2pt, topsep=2pt}

\section{Implementation Details}
\label{app:implementation-details}

\ours implements benchmark evaluation based on Harbor. The implementation integrates Skill Recommendation, Task Attribution, and Skill Evolution into the Harbor workflow.

\subsection{Lifecycle of Harbor Evaluation Framework}
A Harbor job first parses the task collection and expands each task instance into a trial. The main lifecycle of a trial includes creating the trial working directory, starting the task container, running agent setup, passing the instruction to the solver agent, downloading agent logs and sessions, running the verifier, stopping the container, recording the final trial result, and triggering the trial-end hook.

This order determines the following implementation logic of \ours:
\begin{itemize}
    \item The recommendation stage must run before the solver agent starts, because it controls which skills are installed into the agent-visible directory and appends compressed skill-use guidance to the task instruction.
    \item Attribution and skill evolution must run after the trial ends, because they depend on the complete agent session and verifier result. At that point, the task container has already stopped, so both stages run on the host side.
\end{itemize}

\ours implements the above mechanism using the trial lifecycle hook exposed by Harbor. When the trial-end hook is triggered, it can already access the trial result, agent artifacts, and verifier artifacts, which match the inputs required by attribution and evolution. If a failed trial will still be retried by Harbor itself, \ours skips attribution and evolution for that attempt and waits until the final attempt finishes, avoiding duplicate writes of intermediate failures into the experience library.

The task-solving process of the solver agent keeps Harbor's original logic, including execution and verifying. \ours only changes the preparation stage before task execution and the attribution stage after task execution, without modifying the solver agent's task-solving workflow.

\subsection{Dataset Preparation and Environment Setup}
Harbor's official dataset images do not guarantee a fixed preinstalled agent CLI. If each trial downloads the agent CLI during setup, it creates two experimental issues: concurrent runs generate heavy download traffic, and CLI version changes may affect experimental results. \ours therefore prebuilds experiment images with the agent CLI on top of the original task images and skips repeated installation at runtime.

The prebuild process first downloads the Harbor dataset, reads the image definition for each task, builds a new image with a preinstalled agent from the original task image as the base image, and uses the built image in the task configuration. The prebuilt image fixes \texttt{nvm}~\texttt{0.40.4}, Node.js~\texttt{22}, and Codex CLI~\texttt{0.125.0}.

This approach reduces repeated downloads in concurrent experiments, fixes the agent CLI version, and shortens the trial preparation time.

\subsection{Experiment Configuration and Orchestration}
\ours provides a lightweight launcher outside Harbor. The launcher uses YAML to manage both native Harbor configuration and \ours-specific extended configuration, registers different modules, and makes experimental setup convenient.

This design allows baselines, recommendation, online evolution, and offline evolution to share the same launcher. Different experiments only need to change YAML, without modifying Harbor code or copying multiple execution scripts.

\subsection{Integration of Solver Agent}
We implement our solver agent based on the Codex integration provided by Harbor, but do not modify its task execution logic. We only modify the preparation work before trial execution. Because the image already preinstalls Codex~\texttt{0.125.0}, agent setup no longer installs the CLI and only creates the necessary agent directories. During formal execution, Codex still runs the task instruction in execution mode, with JSON logging and the unified terminal execution tool enabled so that Harbor can download the session and execution status.

Codex's built-in system skills and plugins may inject extra prompt content, interfering with the measurement of the \ours skill library. The system first initializes the agent home so that Codex discovers system skills and plugins; it then generates the agent configuration according to the experiment setting and marks system skills and plugins as disabled.

\subsection{Integration of Skill Recommendation}
The goal of the recommendation stage is to select a small set of skills that are most relevant to the task and least redundant before the solver agent executes. The agent itself does not solve the task; it only searches the candidate skill library, reads candidate skill documents and resources, and generates skill usage guidance for the solver agent.

\begin{enumerate}
    \item The candidate skill library is mounted into the task container through Harbor as a read-only directory. The agent's \texttt{cwd} is set to the candidate skill root, not the benchmark task workspace.
    \item An isolated recommendation environment is created. The recommendation stage uses a temporary agent home and a temporary output directory. System skills and plugins are disabled in the same controlled way as for the solver, and the candidate skill root is used only as the trusted directory for recommendation.
    \item The recommendation prompt is rendered and executed. The recommendation stage reuses the solver agent's model and CLI parameters and runs with bypass permissions.
    \item The recommendation stage writes the JSON schema, structured output file, command log, and recommendation session. The structured output is parsed and validated; if the output is missing or malformed, it is retried up to three times. Logs, intermediate outputs, final outputs, and the recommendation session are downloaded to the host trial directory.
    \item If recommendation succeeds and returns a nonempty skill-name list, the system copies only those skill directories into the solver agent's skill directory, and the concise usage guidance generated by the agent is appended to the end of the task instruction as the solver agent's skill usage context. If the recommendation stage repeatedly fails to produce valid output, or if installing the selected skills fails, the system records the error and falls back to copying all candidate skills, allowing the benchmark trial to continue.
    \item After recommendation ends, the temporary recommendation directory, temporary agent home, and temporary credential files inside the container are cleaned up.
\end{enumerate}

\subsection{Integration of Task Attribution}
This stage turns a complete agent trajectory and verifier result into structured subtasks. It does not simply compress the original session into text; instead, it resumes the original Codex session so that the agent continues reasoning inside the native agent harness context. This prevents loss of contextual details, and because commercial models often encrypt chain-of-thought, the resume mechanism is the only way to avoid discarding that context.

\begin{enumerate}
    \item When the trial-end hook is triggered, Harbor has already downloaded agent artifacts, run the verifier, and stopped the task container. The host-side session and verifier artifacts can therefore be obtained.
    \item An isolated working path and a new agent home are created for each trial. The original solver agent session, visible skills, and related artifacts are copied.
    \item Because each trial is expected to produce exactly one agent session file, the system reads the session id needed for resume from that session file and runs Codex resume with the isolated working path and agent home to restore the context at that time. The prompt is appended as a new user prompt through standard input; resume does not replace the original system prompt, preserving native session context management.
    \item Verifier evidence is provided in a controlled way, mainly in two modes:
    \begin{itemize}
        \item Online mode provides only task-level test counts, including the total, passed, and failed counts. These counts can come from CTRF reports, pytest summaries, benchmark-specific JSON, or Harbor reward.
        \item Offline oracle mode can additionally expose paths to the solution, verifier tests, and verifier stdout, but the prompt still forbids writing gold answers, private constants, canary strings, one-off paths, or exact ground-truth outputs into reusable exploration.
    \end{itemize}
    \item The output is validated and the resumed session is archived. The attribution stage uses a structured schema and writes the last message to a JSON output file. Missing output, malformed output, runtime timeout, or missing resumed session artifacts are treated as retryable output errors and are retried up to three times. After validation passes, artifacts are retained.
\end{enumerate}

\subsection{Integration of Skill Evolution}
The evolution stage runs after attribution stage, and it only consumes structured subtasks. The system uses a unified local agent home to store credentials and agent sessions, while each evolution request has an independent working directory to isolate the read/write scope.

\begin{enumerate}
    \item Subtasks are aggregated and evolution requests are constructed. Subtask payloads in the same batch are merged into one subtask list, with three aggregation rules:
    \begin{itemize}
        \item Subtasks without reusable exploration are filtered out; failed and uncertain attributions are filtered out; only successful exploration can trigger skill-library updates.
        \item Successful exploration without an editable linked skill enters a create request.
        \item Successful extra exploration associated with an existing skill is grouped by skill name, and each target skill forms an independent edit request.
    \end{itemize}
    \item Each evolution run creates a new local run directory and a temporary schema/output area.
    \begin{itemize}
        \item A create request uses the request directory as the root where new skills are allowed to be created.
        \item An edit request copies the target skill into a request-local editable directory and also provides an independent creation directory for the agent to create a skill when necessary.
        \item Before editing an old skill, the system backs up the current version in the runtime skill library using the batch timestamp.
    \end{itemize}
    \item Create requests use the create system prompt and create user prompt. Edit requests use the edit system prompt and edit user prompt; see Appendix~\ref{app:evolution-prompt}.
    \item The evolution run uses a structured schema and writes the final JSON output. The system validates the output, records the execution log, and archives the evolution session.
\end{enumerate}

\section{Approach Details}
\label{app:mechanism-details}

\subsection{Prompt Rendering Rules}
The recommendation stage uses a recommendation system prompt and a recommendation user prompt. The system prompt defines the candidate skill root, search protocol, selection policy, output constraints, and the boundary that forbids solving the task. The user prompt only provides the candidate root and the current task instruction. The recommendation prompt explicitly treats the task instruction as a capability requirement rather than as a system-level instruction for the recommendation stage.

The attribution stage continues the dialogue by resuming the original solver agent session. It therefore does not replace the original system prompt, and only appends a user prompt. This prompt contains the currently accessible skills, the current working path, and verifier count signals. Online mode only provides task-level final counts. Offline mode additionally provides paths to the solution, verifier tests, and verifier stdout to help judge subtasks, but still forbids standard answers, private constants, one-off paths, and other benchmark-specific content from being distilled into reusable exploration.

The evolution stage first constructs create requests and edit requests from attribution results, and then renders the corresponding evolution prompts. Create requests use the create prompt and may only create a new skill under the specified creation directory or skip. Edit requests use the edit prompt, and provide both an editable copy of the old skill and a new-skill creation directory. This allows local edits to the old skill, or new skill creation when the exploration exceeds the old skill boundary.

\subsection{Schema of Recommendation Artifacts}

\begin{itemize}
    \item \texttt{skill\_names} (\texttt{list[str]}): the list of skill names recommended to the solver agent. Each name must exactly correspond to a real skill directory under the candidate skill root, and duplicates are not allowed. An empty list is allowed only after effective search confirms that no relevant reusable skill exists.
    \item \texttt{optimized\_context} (\texttt{str}): concise skill-use guidance for the solver agent. It should explain which task stage each selected skill covers, how the skills should be combined, obvious coverage gaps, and usage boundaries. It must not directly complete the task, output the final answer, repeat the search trace, or copy long skill content.
\end{itemize}

\subsection{Schema of Attribution Artifacts}

\begin{itemize}
    \item \texttt{subtasks} (\texttt{list[Subtask]}): the list of subtasks extracted from this trajectory, containing at least one element.
    \item \texttt{goal} (\texttt{str}): the independent objective of the subtask.
    \item \texttt{summary} (\texttt{str}): the factual summary of the subtask.
    \item \texttt{exploration} (\texttt{str | null}): reusable exploration produced in the subtask; \texttt{null} if there is no reusable content worth retaining.
    \item \texttt{exploration\_reason} (\texttt{str}): the explanation for \texttt{exploration}.
    \item \texttt{judge} (\texttt{enum}): the primary judgment signal type used by the subtask, with values shown in Table~\ref{tab:judge-enum}.
    \item \texttt{judge\_reason} (\texttt{str}): the evidence explanation for selecting this judge type.
    \item \texttt{attribution} (\texttt{enum}): the final result and main-cause category of the subtask, with values shown in Table~\ref{tab:attribution-enum}.
    \item \texttt{attribution\_reason} (\texttt{str}): the evidence explanation for selecting this attribution.
    \item \texttt{skill\_linked} (\texttt{str | null}): the single skill name associated with the subtask.
    \item \texttt{skill\_refs} (\texttt{list[SkillRef]}): the skill text spans actually relied on by the subtask.
    \item \texttt{SkillRef.file\_path} (\texttt{str}): the relative path of the referenced file inside the skill directory.
    \item \texttt{SkillRef.start\_line} (\texttt{int | null}): the 1-based starting line number of the referenced knowledge span.
    \item \texttt{SkillRef.end\_line} (\texttt{int | null}): the 1-based ending line number of the referenced knowledge span.
    \item \texttt{SkillRef.capability} (\texttt{str}): the capability, instruction, or knowledge summary expressed by the referenced span.
    \item \texttt{SkillRef.used\_for} (\texttt{str}): how the knowledge span was actually used in the current subtask.
    \item \texttt{ground\_truth\_path} (\texttt{str | null}): the oracle directory path attached by the program in offline oracle mode. It is not directly output by the agent.
\end{itemize}

\begin{table}[!p]
    \centering
    \small
    \caption{Judge signal types used by task attribution. Each type identifies the primary evidence source for judging a subtask outcome.}
    \label{tab:judge-enum}
    \renewcommand{\arraystretch}{1.35}
    \begin{tabularx}{\linewidth}{>{\raggedright\arraybackslash}p{0.22\linewidth} @{\hspace{0.12\linewidth}} X}
        \toprule
        Judge Type & Meaning \\
        \midrule
        \texttt{environment} & Primarily judged by observable environment feedback. \\
        \texttt{human} & The result depends on human preference or manual review. \\
        \texttt{unknown} & No clear judgment signal exists. \\
        \bottomrule
    \end{tabularx}
\end{table}

\begin{table}[!p]
    \centering
    \small
    \caption{Attribution categories produced for each subtask. Successful categories can trigger skill creation or update, while failed and uncertain categories are skipped during skill evolution.}
    \label{tab:attribution-enum}
    \renewcommand{\arraystretch}{1.35}
    \begin{tabularx}{\linewidth}{>{\raggedright\arraybackslash}p{0.34\linewidth} X @{\hspace{0.025\linewidth}}>{\centering\arraybackslash}p{0.15\linewidth}}
        \toprule
        Attribution Type & Meaning & Evolution Type \\
        \midrule
        \path{success_viewed_skill_but_not_used} & The agent viewed a skill, but the skill did not materially shape the successful path. & Create \\
        \path{success_no_skill_seen} & The agent did not view a skill and still completed the subtask through independent exploration. & Create \\
        \path{success_skill_used_with_extra_exploration} & The agent genuinely relied on a skill, performed extra exploration, and succeeded. & Edit or Create \\
        \path{fail_skill_issue} & The main failure cause lies in the skill itself. & Skip \\
        \path{fail_agent_limit} & The main failure cause lies in the agent. & Skip \\
        \path{fail_client_env} & The main failure cause lies in the client-side environment. & Skip \\
        \path{fail_external_env} & The main failure cause lies in external systems or services. & Skip \\
        \path{fail_unknown_env} & The subtask clearly failed because of the environment, but the evidence cannot distinguish the client environment from the external environment. & Skip \\
        \path{uncertain_human_judge_required} & Human judgment is required but currently unavailable. & Skip \\
        \path{uncertain_environment_judge_inconclusive} & Environment signals exist but are insufficient for complete judgment. & Skip \\
        \path{uncertain_no_judge} & No clear judgment signal exists and the goal is not self-evident enough. & Skip \\
        \bottomrule
    \end{tabularx}
\end{table}

\subsection{Schema of Evolution Artifacts}

\begin{itemize}
    \item \texttt{request\_dir\_name} (\texttt{str}): the working directory name of the evolution request.
    \item \texttt{target\_skill\_name} (\texttt{str | null}): the old skill name corresponding to an edit request; \texttt{null} for a create request.
    \item \texttt{subtasks} (\texttt{list[Subtask]}): the subtasks supporting this evolution request.
    \item \texttt{actions} (\texttt{list[Action]}): the list of actions returned by the agent.
    \item \texttt{action\_type} (\texttt{enum}): the evolution action category.
    \item \texttt{rationale} (\texttt{str}): the reason for executing the action.
    \item \texttt{summary} (\texttt{str | null}): the edit summary when editing an old skill; \texttt{null} for creation or skip.
    \item \texttt{skill\_dir\_path} (\texttt{str | null}): the absolute directory path of a newly created skill; \texttt{null} when editing an old skill or skipping.
\end{itemize}

\begin{table}[!p]
    \centering
    \small
    \caption{Evolution action types and corresponding operations. Each action describes how reusable exploration is evolved, ranging from editing an existing skill to creating a new skill or skip.}
    \label{tab:action-enum}
    \renewcommand{\arraystretch}{1.45}
    \begin{tabularx}{\linewidth}{>{\raggedright\arraybackslash}p{0.27\linewidth} @{\hspace{0.035\linewidth}} X @{\hspace{0.04\linewidth}}>{\centering\arraybackslash}p{0.17\linewidth}}
        \toprule
        Evolution Action Type & Meaning & Skill Operation \\
        \midrule
        \texttt{error\_fix} & Correct clearly wrong or misleading guidance in an old skill. & Edit \\
        \texttt{knowledge\_addition} & Add missing reusable steps, branches, or fallback guidance to an old skill. & Edit \\
        \texttt{prerequisite\_addition} & Add prerequisites, applicability boundaries, warnings, or guardrails. & Edit \\
        \texttt{create\_skill} & Create a new independent skill. & Create \\
        \texttt{skip} & Do not update. & Skip \\
        \bottomrule
    \end{tabularx}
\end{table}

\subsection{Aggregation of Subtasks}
Before evolution, the system first merges all subtasks payloads in a batch, and then checks the subtasks one by one. A subtask enters evolution only when it satisfies two conditions:
\begin{itemize}
    \item \texttt{exploration} is nonempty, because we only evolve explorations performed by the agent into skills.
    \item \texttt{attribution} belongs to the three successful exploration categories. Failed and uncertain subtasks do not trigger library updates.
\end{itemize}

After aggregation, successful exploration without an editable linked skill is placed into a create request. Successful exploration with \texttt{success\_skill\_used\_with\_extra\_exploration} and a nonempty \texttt{skill\_linked} is grouped by linked skill into multiple edit requests. Each edit request corresponds to exactly one old skill.

\FloatBarrier

\clearpage
\section{Case Study}
\label{app:case-study}

\subsection{Configure Web Server}
\label{app:case-apache-git}
\svcasebenchmark{Terminal-Bench 2.0}

\noindent This case shows that an existing skill transfers a reusable success pattern to an unseen composite task, rather than copying a source solution.\par\medskip

\begin{svpromptbox}{Baseline}

\svpromptheading{Task Instruction}
Configure a Git server so that a user can run:
\begin{itemize}
    \item \texttt{git clone user@server:/git/server}
    \item \texttt{echo "hello world" > hello.html}
    \item \texttt{git add hello.html}
    \item \texttt{git commit -m "add hello.html"}
    \item \texttt{git push origin master}
\end{itemize}
The pushed data should be served by a webserver on port~\texttt{8080}; \texttt{curl http://server:8080/hello.html} should return \texttt{hello world}. Login setup is assumed to be handled externally.

\svpartsep
\svpromptheading{Agent Trajectory}
\svtrace{[Thought]}{Inspect the workspace and wire a Git receive/update flow to an HTTP serving path so pushed files materialize on port~\texttt{8080}.}
\svtrace{[Command]}{Probe available runtimes and service binaries, including Python, Node.js, Ruby, Go, \texttt{sshd}, and \texttt{git}.}
\svtrace{[Thought]}{Use a bare repository with a \texttt{post-receive} hook that checks out the pushed branch into a web root, plus a small static server on \texttt{8080}.}
\svtrace{[Command]}{Attempt to install Git with \texttt{apt-get update} and \texttt{apt-get install -y git}.}
\svtrace{[Observation]}{\texttt{apt-get} cannot proceed because \path{/var/lib/apt/lists/lock} is held by another process.}
\svtrace{[Command]}{Kill stuck \texttt{apt-get} processes and continue with generated setup files.}
\svtrace{[Final]}{The agent writes setup scripts and documentation, but cannot fully exercise the Git hook because Git installation failed in the environment.}

\end{svpromptbox}

\begin{svpromptbox}{Evolution}

\svpromptheading{Task Instruction}
Configure a Git server so that a user can clone \path{user@server:/git/server}, commit \path{hello.html}, push to \path{master}, and then retrieve the pushed file from \texttt{http://server:8080/hello.html}. The solver also receives the following skill usage:

\begin{quote}
Use \path{ubuntu-apache-vhost} for the HTTP side: bind the service on \texttt{*:8080}, set a stable \path{DocumentRoot}, and validate the vhost after reload so pushed files are actually served from the expected path. There is no dedicated Git-server skill in \path{/skills}, so the downstream agent needs to handle the bare repository and \path{post-receive} deployment hook directly while using this skill for persistent webserver configuration and verification.
\end{quote}

\svpartsep
\svpromptheading{Recommendation Usage of Skills}
\begin{itemize}
    \item \textbf{skills:} \path{ubuntu-apache-vhost}
    \item \textbf{usage of skills:} Use \path{ubuntu-apache-vhost} for the HTTP side: bind the service on \texttt{*:8080}, set a stable \path{DocumentRoot}, and validate the vhost after reload so pushed files are actually served from the expected path. The Git repository and \path{post-receive} deployment hook remain outside the recommended skill and must be handled directly by the solver.
\end{itemize}

\svpartsep
\svpromptheading{Skill Text Used}
\svpromptsubheading{[1] \path{ubuntu-apache-vhost}, line 14--20:}
\begin{enumerate}
    \item Inspect live bindings with \texttt{apache2ctl -S} and confirm whether Apache is already listening on the target port.
    \item Check service state; in minimal Ubuntu/Debian environments, \texttt{systemctl} may be unavailable, so use \texttt{service apache2 start} or \texttt{service apache2 restart}.
    \item Put global listeners in \path{ports.conf}.
    \item Put site-specific settings in \path{sites-available}.
    \item Enable the site with a symlink or \texttt{a2ensite}.
    \item Add required modules before relying on their directives.
    \item Validate with \texttt{apache2ctl -t}, then restart or reload Apache.
\end{enumerate}
\svpromptsubheading{[2] \path{ubuntu-apache-vhost}, line 24--26:}
\begin{itemize}
    \item Use \texttt{Listen} in \path{ports.conf} for persistent port bindings.
    \item Keep the virtual host on the matching \texttt{*:PORT} stanza.
    \item Define a clear \path{DocumentRoot} and explicit log paths.
\end{itemize}
\svpromptsubheading{[3] \path{ubuntu-apache-vhost}, line 36--40:}
\begin{itemize}
    \item Confirm the configured port responds with \texttt{200 OK}.
    \item Confirm the access log records the expected fields.
    \item Confirm a denied path produces \texttt{403 Forbidden} and the custom 403 body.
    \item Check the error log for the denied access entry.
    \item Use \texttt{apache2ctl -S} after enabling the site to verify the new vhost is bound to the intended \texttt{*:PORT} listener.
\end{itemize}

\svpartsep
\svpromptheading{Agent Trajectory}
\svtrace{[Thought]}{Use \path{ubuntu-apache-vhost} for the HTTP side, inspect the Apache layout, and then wire the bare Git repository plus deployment hook so pushes land in Apache's document root.}
\svtrace{[Command]}{Read \path{/logs/agent/skills/ubuntu-apache-vhost/SKILL.md}.}
\svtrace{[Command]}{Run \texttt{apache2ctl -S} and check Apache service state with \texttt{service apache2 status} or \texttt{systemctl status apache2}.}
\svtrace{[Command]}{Install the required packages: Apache, Git, and OpenSSH server.}
\svtrace{[Command]}{Inspect \path{/etc/apache2/ports.conf} and existing site configurations.}
\svtrace{[Command]}{Create the bare repository, add an executable \path{post-receive} hook, enable the Apache site, and run \texttt{apache2ctl -t}.}
\svtrace{[Observation]}{\texttt{apache2ctl -S} reports a \texttt{*:8080} vhost from \path{/etc/apache2/sites-enabled/git-server.conf}.}
\svtrace{[Observation]}{A local commit and push to \path{/git/server}, followed by \texttt{curl http://127.0.0.1:8080/hello.html}, returns \texttt{hello world}.}

\svpartsep
\svpromptheading{Distilled Subtasks}
\begin{enumerate}[label=\textnormal{[\arabic*]}, leftmargin=*]
    \item \textbf{goal:} Configure Apache to serve a persistent virtual host on port~\texttt{8080} from a stable document root. \textbf{summary:} The agent installed packages, configured \path{/etc/apache2/ports.conf}, added a \texttt{*:8080} vhost with \path{DocumentRoot /srv/git-server/www}, enabled the site, and verified it with \texttt{apache2ctl -t}, \texttt{apache2ctl -S}, and \texttt{curl}. \textbf{exploration:} On Ubuntu/Debian, persistent Apache port binding should be placed in \path{ports.conf}; site config should live in \path{sites-available}; \texttt{apache2ctl -S} plus a live HTTP request are fast checks that the vhost is bound and serving from the intended document root. \textbf{attribution:} \texttt{success\_skill\_used\_with\_extra\_exploration}. \textbf{skill\_linked:} \path{ubuntu-apache-vhost}. \textbf{skill\_refs:} \path{SKILL.md} line 14--20, line 24--26, and line 36--40.
    \item \textbf{goal:} Create a bare Git repository with a deploy hook that publishes pushed \path{master} contents into the Apache document root. \textbf{summary:} The agent created \path{/git/server} as a bare repository and \path{/srv/git-server/www} as the deployment target, added an executable \path{post-receive} hook, pushed \path{hello.html}, and verified that \texttt{curl} returned \texttt{hello world}. \textbf{exploration:} A bare repository can serve as a lightweight deployment endpoint when \path{post-receive} checks out the pushed branch into the web root and cleans stray files so the web root mirrors the branch. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
\end{enumerate}

\svpartsep
\svpromptheading{Skill Evolution}
\svpromptsubheading{[1] [edit] \path{ubuntu-apache-vhost}, line 40:}
Use \texttt{apache2ctl -t} for syntax, then \texttt{apache2ctl -S} plus a live \texttt{curl} request after enabling the site to verify the new vhost is bound to the intended \texttt{*:PORT} listener and serving the intended \path{DocumentRoot}.

\svpromptsubheading{[2] [create] \path{git-web-publish}:}
\begin{itemize}
    \item Configure a bare Git repository with a \path{post-receive} hook that publishes pushed branches into a web document root.
    \item Use the push-to-bare-repository pattern: receive a push, run \path{post-receive}, and mirror the deployed branch into the web root.
    \item Check out the deployed branch into the target directory, clean stray files so the web root mirrors the branch, then verify by pushing a commit and requesting the served file.
\end{itemize}

\end{svpromptbox}

\subsection{Configure Apache Logging And Rat}
\label{app:case-apache-logging-rat}
\svcasebenchmark{Terminal-Bench Pro}

\noindent This case shows how \ours creates a new skill by distilling reusable procedures from successful exploration.\par\medskip

\begin{svpromptbox}{Evolution}
\svpromptheading{Task Instruction}
Configure an Apache web server environment to serve content with comprehensive logging and rate limiting.
\begin{itemize}
    \item \textbf{Apache2:} installed and available via \texttt{apache2ctl}.
    \item \textbf{Port:} listen on port~\texttt{9090}.
    \item \textbf{Document Root:} serve static content from \path{/srv/www/content}.
    \item \textbf{Log Format:} include timestamps, HTTP methods, status codes, response sizes, and referrers.
\end{itemize}
The task requires a \texttt{*:9090} virtual host with \texttt{ServerName localhost}, detailed access and error logs under \path{/var/log/apache2}, a maximum of 20 concurrent connections per IP, a custom \path{/403.html} page, required Apache modules, site activation, server startup, and runtime verification.

\svpartsep
\svpromptheading{Agent Trajectory}
\svtrace{[Thought]}{Inspect the existing Apache layout and repository-specific configuration so the changes can be made persistently.}
\svtrace{[Command]}{Run \texttt{apache2ctl -v}, list loaded modules, and search files under \path{/etc/apache2}.}
\svtrace{[Command]}{Read the benchmark site configuration and \path{/etc/apache2/ports.conf}.}
\svtrace{[Command]}{Search Apache packages for connection-limiting modules such as \texttt{qos}, \texttt{evasive}, and \texttt{ratelimit}.}
\svtrace{[Command]}{Install \texttt{libapache2-mod-qos} and inspect \texttt{QS\_} directives.}
\svtrace{[Command]}{Validate the Apache configuration with \texttt{apache2ctl -t} and inspect vhost bindings with \texttt{apache2ctl -S}.}
\svtrace{[Observation]}{Live requests to \texttt{http://127.0.0.1:9090/} and \path{/403.html} verify serving and custom error behavior.}
\svtrace{[Observation]}{A denied real path under the document root produces the expected access and error log entries.}

\svpartsep
\svpromptheading{Distilled Subtasks}
\begin{enumerate}[label=\textnormal{[\arabic*]}, leftmargin=*]
    \item \textbf{goal:} Inspect the Apache environment and locate the existing benchmark site configuration. \textbf{exploration:} On Ubuntu Apache layouts, persistent port bindings belong in \path{ports.conf}, while site-specific behavior should live in \path{sites-available} and be activated via symlinks. Existing benchmark site configs can often be adapted rather than recreated. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
    \item \textbf{goal:} Implement persistent Apache configuration for port~\texttt{9090}, custom logging, and per-IP connection limiting. \textbf{exploration:} For Apache concurrency limits, \texttt{mod\_qos} provides \texttt{QS\_SrvMaxConnPerIP}, which is practical when the stock module set only includes \texttt{mod\_ratelimit} for bandwidth throttling rather than true connection concurrency control. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
    \item \textbf{goal:} Start Apache and verify the server responds correctly on port~\texttt{9090} with the expected logging and 403 behavior. \textbf{exploration:} A reliable Apache 403 test is easier to validate by denying a real directory under the document root than by relying on a nonexistent path or rewrite rule. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
\end{enumerate}

\svpartsep
\svpromptheading{Skill Evolution}
\svpromptsubheading{[1] [create] \path{ubuntu-apache-vhost}:}
\begin{itemize}
    \item Configure persistent Apache vhosts on Ubuntu or Debian, including ports, logs, access control, and validation.
    \item Put global listeners in \path{ports.conf}, site-specific settings in \path{sites-available}, enable the site, and validate with \texttt{apache2ctl -t} before restart or reload.
    \item Use \texttt{Listen} for persistent port bindings, keep the virtual host on the matching \texttt{*:PORT} stanza, and define explicit \path{DocumentRoot} and log paths.
    \item For custom 403 behavior, serve \path{/403.html} from the document root and deny a real directory when a runtime 403 test is needed.
    \item For true per-IP connection limits, use \texttt{mod\_qos} and \texttt{QS\_SrvMaxConnPerIP} when the stock module set is insufficient.
\end{itemize}

\end{svpromptbox}

\subsection{Configure Apache Analytics Virtu}
\label{app:case-apache-analytics-virtu}
\svcasebenchmark{Terminal-Bench Pro}

\noindent This case shows how a previously created skill is incrementally refined with new operational guardrails from later successful use.\par\medskip

\begin{svpromptbox}{Evolution}
\svpromptheading{Task Instruction}
Configure Apache2 to implement custom port listening and advanced access log recording.
\begin{enumerate}
    \item Create \path{analytics-site.conf} under \path{/etc/apache2/sites-available/}, using \texttt{<VirtualHost *:9090>}, \texttt{ServerName localhost}, and \path{DocumentRoot /var/www/content}.
    \item Configure \path{/var/log/apache2/traffic-access.log} with the format \texttt{\%t \%m \%s \%\{Referer\}i [\%a]}.
    \item Configure \path{/var/www/content} with \texttt{Options Indexes FollowSymLinks}, \texttt{AllowOverride None}, and \texttt{Require all granted}.
    \item Add \texttt{Listen 9090} at the end of the configuration file.
    \item Enable the site with \texttt{a2ensite analytics-site.conf}.
\end{enumerate}

\svpartsep
\svpromptheading{Skill Text Used}
\svpromptsubheading{[1] \path{ubuntu-apache-vhost}, line 14--18:}
\begin{enumerate}
    \item Put global listeners in \path{ports.conf}.
    \item Put site-specific settings in \path{sites-available}.
    \item Enable the site with a symlink or \texttt{a2ensite}.
    \item Add required modules before relying on their directives.
    \item Validate with \texttt{apache2ctl -t}, then restart or reload Apache.
\end{enumerate}
\svpromptsubheading{[2] \path{ubuntu-apache-vhost}, line 24--26:}
\begin{itemize}
    \item Define a clear \path{DocumentRoot} and explicit log paths.
    \item Use \texttt{ErrorDocument 403 /403.html} only when the file is served from the document root.
    \item If a real 403 test is needed, deny a real directory under the document root rather than relying on a nonexistent path.
\end{itemize}

\svpartsep
\svpromptheading{Agent Trajectory}
\svtrace{[Thought]}{Use \path{ubuntu-apache-vhost} because the task is an Apache vhost and listener configuration task; inspect the current layout before editing.}
\svtrace{[Command]}{Read \path{/root/.agents/skills/ubuntu-apache-vhost/SKILL.md}.}
\svtrace{[Command]}{List Apache configuration files and inspect current bindings with \texttt{apache2ctl -S}.}
\svtrace{[Command]}{Check service state using \texttt{systemctl is-active apache2} or \texttt{service apache2 status}.}
\svtrace{[Command]}{Enable \path{analytics-site.conf}, validate syntax with \texttt{apache2ctl -t}, and start Apache with \texttt{service apache2 start}.}
\svtrace{[Observation]}{Requests to \texttt{http://localhost:9090/} with a referrer create the expected \path{/var/log/apache2/traffic-access.log} entries.}
\svtrace{[Command]}{Confirm the listener with \texttt{ss -ltnp} and inspect the site file for \path{traffic-access} and \texttt{Listen 9090}.}

\svpartsep
\svpromptheading{Distilled Subtasks}
\begin{enumerate}[label=\textnormal{[\arabic*]}, leftmargin=*]
    \item \textbf{goal:} Inspect the Apache environment and determine the correct configuration path for a \texttt{9090} analytics vhost. \textbf{exploration:} Use \texttt{apache2ctl -S} together with the Debian/Ubuntu Apache directory layout to verify existing listeners and vhost bindings before making changes. \textbf{attribution:} \texttt{success\_skill\_used\_with\_extra\_exploration}. \textbf{skill\_linked:} \path{ubuntu-apache-vhost}.
    \item \textbf{goal:} Implement and activate the analytics Apache site on port~\texttt{9090} with the required document root, log format, and homepage. \textbf{exploration:} Placing \texttt{Listen 9090} at the end of the site file worked in this environment, and \texttt{service apache2 start} was the viable startup path when \texttt{systemctl} was unavailable. \textbf{attribution:} \texttt{success\_skill\_used\_with\_extra\_exploration}. \textbf{skill\_linked:} \path{ubuntu-apache-vhost}.
    \item \textbf{goal:} Verify the deployed Apache site, response body, and custom access-log behavior against the benchmark expectations. \textbf{exploration:} Full validation required both runtime probing and file inspection because the verifier checks HTTP response, log creation, exact log formatting, enabled-site state, and static config content. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
\end{enumerate}

\svpartsep
\svpromptheading{Skill Evolution}
\svpromptsubheading{[1] [edit] \path{ubuntu-apache-vhost}, line 14--20:}
\begin{enumerate}
    \item Inspect live bindings with \texttt{apache2ctl -S} and confirm whether Apache is already listening on the target port.
    \item Check service state; in minimal Ubuntu/Debian environments, \texttt{systemctl} may be unavailable, so use \texttt{service apache2 start} or \texttt{service apache2 restart}.
    \item Put global listeners in \path{ports.conf}.
    \item Put site-specific settings in \path{sites-available}.
    \item Enable the site with a symlink or \texttt{a2ensite}.
    \item Add required modules before relying on their directives.
    \item Validate with \texttt{apache2ctl -t}, then restart or reload Apache.
\end{enumerate}
\svpromptsubheading{[2] [edit] \path{ubuntu-apache-vhost}, line 40:}
Use \texttt{apache2ctl -S} after enabling the site to verify the new vhost is bound to the intended \texttt{*:PORT} listener.

\end{svpromptbox}

\subsection{NodeBB Group Invite API}
\svcasebenchmark{SWE-Bench Pro}

\noindent This case shows how \ours performs attribution in a complex task and evolves the successful subtasks.\par\medskip

\begin{svpromptbox}{Evolution}
\svpromptheading{Task Instruction}
\svpromptsubheading{Lack of API Support for Managing Group Invitations Limits Extensibility}
Group invitation logic for issuing, accepting, and rejecting invitations is handled through socket events and the web layer. The task requires authenticated HTTP API endpoints so external clients can issue, accept, and reject or rescind group invitations.

\begin{itemize}
    \item \textbf{Issue invite:} \texttt{POST /groups/\{slug\}/invites/\{uid\}} lets a group owner or admin issue an invitation and log \texttt{group-invite}.
    \item \textbf{Accept invite:} \texttt{PUT /groups/\{slug\}/invites/\{uid\}} lets the invited user accept their own invite. If \texttt{uid} differs from the caller, return \texttt{[[error:not-allowed]]}; if no invite exists, return \texttt{[[error:not-invited]]}; log \texttt{group-invite-accept}.
    \item \textbf{Reject or rescind invite:} \texttt{DELETE /groups/\{slug\}/invites/\{uid\}} lets the invited user reject or an owner/admin rescind an invite, with \texttt{[[error:not-invited]]} or \texttt{[[error:not-allowed]]} on invalid cases. Rejection by the invited user logs \texttt{group-invite-reject}.
\end{itemize}
The implementation should add exported functions \texttt{issueInvite}, \texttt{acceptInvite}, and \texttt{rejectInvite} in \path{src/api/groups.js}, add controllers in \path{src/controllers/write/groups.js}, and update the web client and OpenAPI specification.

\svpartsep
\svpromptheading{Recommendation Usage of Skills}
\begin{itemize}
    \item \textbf{skills:} \path{nodebb-core-route-module}, \path{nodebb-v3-write-api-repro}, \path{nodebb-bootstrap-repro}, \path{debug-http-status-api-errors}, \path{nodebb-api-error-and-teaser-debug}
    \item \textbf{usage of skills:} Use \path{nodebb-core-route-module} to mount \texttt{POST}, \texttt{PUT}, and \texttt{DELETE} routes at \texttt{/groups/:slug/invites/:uid}. Use \path{nodebb-v3-write-api-repro} to keep controller success and error envelopes consistent with v3 write API conventions. Use \path{nodebb-bootstrap-repro} to build an authenticated Mocha repro under \path{scripts/} with \path{test/mocks/databasemock}. Use \path{debug-http-status-api-errors} when the new endpoints return embedded errors or wrong statuses. Use \path{nodebb-api-error-and-teaser-debug} to preserve raw error keys such as \texttt{[[error:not-invited]]} and \texttt{[[error:not-allowed]]} through the serializer.
\end{itemize}

\svpartsep
\svpromptheading{Skill Text Used}
\svpromptsubheading{[1] \path{nodebb-core-route-module}, line 14--27:}
\begin{itemize}
    \item Find the main route composition entry point.
    \item Create a small route module.
    \item Register and mount the route.
    \item Sanity check that new routes are not nested under unrelated routers unless explicitly intended.
    \item Add a minimal integration test or repro when needed.
\end{itemize}
\svpromptsubheading{[2] \path{nodebb-v3-write-api-repro}, line 13--28:}
\begin{itemize}
    \item Use \texttt{helpers.formatApiResponse(status, res, payload)} for successful v3 write responses.
    \item Prefer surfacing non-2xx outcomes by throwing or returning \texttt{Error} instances.
    \item For a minimal \path{/api/v3/**} Mocha repro, import \path{test/mocks/databasemock}.
    \item Use \path{test/helpers} to create users, log in, obtain CSRF tokens, and maintain an authenticated cookie jar.
    \item Assert the v3 envelope shape for success and error cases.
\end{itemize}
\svpromptsubheading{[3] \path{nodebb-bootstrap-repro}, line 24--42:}
\begin{itemize}
    \item Reuse existing helper utilities for authenticated HTTP requests instead of re-implementing login flows.
    \item Load helper modules after bootstrap.
    \item Use helper HTTP clients and login utilities to obtain an authenticated cookie jar or session.
    \item Prefer helper request wrappers for CSRF and cookies; if helpers are not used, fetch CSRF from \path{/api/config} and send it as \texttt{x-csrf-token}.
\end{itemize}
\svpromptsubheading{[4] \path{nodebb-api-error-and-teaser-debug}, line 17--27:}
\begin{itemize}
    \item Locate the route wrapper and confirm whether it uses an API wrapper.
    \item Confirm the error formatting path ends in a shared formatter.
    \item Prefer throwing invalid-input errors so the wrapper produces a consistent API error response.
    \item Avoid ad-hoc payloads that bypass canonical error serialization.
\end{itemize}

\svpartsep
\svpromptheading{Agent Trajectory}
\svtrace{[Thought]}{Scan the repository for group invitation logic, then add v3 write routes, controllers, API methods, client updates, OpenAPI entries, and an authenticated repro.}
\svtrace{[Command]}{Inspect \path{src/groups/invite.js}, \path{src/api/groups.js}, \path{src/controllers/write/groups.js}, and \path{src/routes/write/groups.js}.}
\svtrace{[Command]}{Inspect the existing OpenAPI invite spec and group details client code.}
\svtrace{[Command]}{Create \path{scripts/repro-group-invite-write-api.js} to exercise issue, accept, reject, and rescind flows through \texttt{/api/v3/groups/:slug/invites/:uid}.}
\svtrace{[Command]}{Run the Mocha repro script.}
\svtrace{[Observation]}{The first repro run fails because Redis is unavailable.}
\svtrace{[Command]}{Start Redis with \texttt{redis-server -{}-daemonize yes -{}-port 6379} and confirm \texttt{redis-cli -p 6379 ping} returns \texttt{PONG}.}
\svtrace{[Observation]}{The repro then reaches the application and confirms route absence before implementation.}
\svtrace{[Command]}{Patch \path{src/api/groups.js}, \path{src/controllers/write/groups.js}, \path{src/routes/write/groups.js}, the web client, and the OpenAPI spec.}
\svtrace{[Command]}{Search for \texttt{formatApiResponse} and existing \texttt{[[error:not-allowed]]} handling to align response envelopes and statuses.}
\svtrace{[Command]}{Patch status mapping and serializer behavior, then rerun the repro.}
\svtrace{[Final]}{The local repro passes for the new group invite Write API, but the verifier still reports six private failures.}

\svpartsep
\svpromptheading{Distilled Subtasks}
\begin{enumerate}[label=\textnormal{[\arabic*]}, leftmargin=*]
    \item \textbf{goal:} Locate the existing group invitation implementation and identify where to add equivalent authenticated HTTP endpoints. \textbf{summary:} The agent found group invite logic primarily in \path{src/socket.io/groups.js} and \path{src/groups/invite.js}, with v3 write structure under \path{src/routes/write/groups.js}, \path{src/controllers/write/groups.js}, and \path{src/api/groups.js}. It confirmed invite HTTP routes were missing or commented out and that OpenAPI only documented \texttt{GET /groups/\{slug\}/invites}. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
    \item \textbf{goal:} Create an executable end-to-end repro that boots NodeBB with the test database mock and exercises the new invite HTTP routes. \textbf{summary:} The agent created \path{scripts/repro-group-invite-write-api.js}, booted through \path{test/mocks/databasemock}, created users and a private group, logged in to obtain authenticated jars, and issued HTTP \texttt{POST}, \texttt{PUT}, and \texttt{DELETE} requests against \texttt{/api/v3/groups/:slug/invites/:uid}. The initial run failed with Redis \texttt{ECONNREFUSED}; after starting Redis, the repro confirmed the pre-implementation 404. \textbf{exploration:} For NodeBB end-to-end repros, use a Mocha script requiring \path{test/mocks/databasemock}, use \path{test/helpers} for login and CSRF handling, then exercise HTTP routes; if bootstrap fails with Redis \texttt{ECONNREFUSED}, start local Redis on that host and port before rerunning. \textbf{attribution:} \texttt{success\_skill\_used\_with\_extra\_exploration}. \textbf{skill\_refs:} \path{nodebb-bootstrap-repro} Mocha, databasemock, and authenticated HTTP pattern.
    \item \textbf{goal:} Implement authenticated v3 write API endpoints to issue, accept, and reject or rescind group invitations using \texttt{slug} and \texttt{uid} path parameters. \textbf{summary:} The agent implemented \texttt{groupsAPI.issueInvite}, \texttt{groupsAPI.acceptInvite}, and \texttt{groupsAPI.rejectInvite} in \path{src/api/groups.js} with permission checks and event logging, then added controllers and routes for \texttt{POST}, \texttt{PUT}, and \texttt{DELETE /api/v3/groups/:slug/invites/:uid}. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
    \item \textbf{goal:} Update the web client and OpenAPI write specification to reflect and use the new invite management routes. \textbf{summary:} The agent updated \path{public/src/client/groups/details.js} to use the v3 API module for invite issue, accept, reject, and rescind actions, and added OpenAPI path/spec files for \texttt{POST}, \texttt{PUT}, and \texttt{DELETE /groups/\{slug\}/invites/\{uid\}}. \textbf{attribution:} \texttt{uncertain\_no\_judge}.
    \item \textbf{goal:} Ensure invite endpoint errors return correct HTTP statuses and preserve canonical error keys in API responses. \textbf{summary:} The agent observed a repro failure where \texttt{[[error:not-allowed]]} surfaced with HTTP~\texttt{400} and a translated \texttt{not-allowed} message. It adjusted \path{src/controllers/helpers.js} to map \texttt{[[error:not-allowed]]} and \texttt{[[error:not-invited]]} to HTTP~\texttt{403} and preserve untranslated keys for \path{/api/v3}; the repro then passed. \textbf{exploration:} When write API tests require exact error keys, avoid automatic translation in the \path{/api/v3} serializer and map common permission-related keys to \texttt{403}; otherwise callers may see translated strings and inconsistent status codes. \textbf{attribution:} \texttt{success\_no\_skill\_seen}.
    \item \textbf{goal:} Achieve full private test suite compliance for the implemented invite API changes without regressing unrelated API behavior. \textbf{summary:} The local repro passed, but the verifier reported \texttt{421/427} private tests passing with six failures remaining, suggesting incomplete edge-case handling or regressions around global API status/translation behavior or permission semantics. \textbf{attribution:} \texttt{fail\_agent\_limit}.
\end{enumerate}

\svpartsep
\svpromptheading{Skill Evolution}
\svpromptsubheading{[1] [edit] \path{nodebb-bootstrap-repro}, line 22--29:}
\begin{itemize}
    \item When a repro needs a logged-in user and CSRF handling, reuse NodeBB test helper utilities instead of re-implementing login flows.
    \item Load the helpers module after bootstrap.
    \item Use helper HTTP clients, agents, and login utilities to obtain an authenticated cookie jar or session.
    \item Prefer helper request wrappers for subsequent HTTP calls so CSRF tokens and cookies are handled the same way core tests do.
\end{itemize}
\svpromptsubheading{[2] [edit] \path{nodebb-bootstrap-repro}, line 65--68:}
\begin{itemize}
    \item Start a temporary Redis locally on the expected host and port with \texttt{redis-server -{}-port 6379}.
    \item Optionally run it in the background with \texttt{redis-server -{}-daemonize yes -{}-port 6379}.
    \item Confirm connectivity before rerunning the repro.
\end{itemize}

\end{svpromptbox}

\clearpage
\section{Prompts}
\label{app:prompt}
\input{sections/appendix_prompts.tex}
