\section{Related Work}
\paragraph{Long-horizon Agentic Search.}
LLM agents increasingly perform information seeking through long-horizon, tool-augmented search rather than single-step retrieval.
Benchmarks such as BrowseComp~\citep{wei2025browsecomp}, BrowseComp-Plus~\citep{chen2025browsecompplus}, GAIA~\citep{mialon2024gaia}, xBench-DeepSearch~\citep{chen2025xbench}, HLE~\citep{phan2025hle}, and BrowseComp-ZH~\citep{zhou2025browsecompzh} require agents to maintain state across dozens of tool-use turns. Search tools, especially retrievers those tuned with agentic reasoning~\citep{chen2026agentir,tang2026largereasonembedding}, support exploration and evidence localization, but they also produce observation-heavy context growth. This setting amplifies long-context degradation and lost-in-the-middle effects~\citep{liu2024lost}, as backbone models must recover sparse evidence from accumulated retrieval noise.  

\paragraph{Context Management for Autonomous Agents.} 
To manage long histories without disrupting task logic, recent work builds on the concept of working memory inspired by MemGPT~\citep{packer2023memgpt} and studies context management heuristics, including static truncation~\citep{liu2025deepseekv32,zeng2026glm5,moonshot2025kimi_researcher}, heuristic eviction, and online compression. Adaptive methods such as AgentFold~\citep{ye2025agentfold} and ReSum~\citep{wu2025resum} add flexibility but also introduce costs: AgentFold uses non-native tool-calling loops, while ReSum requires rolling summarization that increases intermediate computation and can reduce KV-cache reuse. To study without heavy architectural overhead, our work leverages a minimal diagnostic setup to unearth how observation dynamics shape context management utility. Appendix~\S\ref{app:related-work} provides a more detailed discussion.