\section{Conclusion}
\label{sec:conclusion}

We have introduced \emph{latent spatial memory}, a 3D cache that stores the video diffusion model's own latent features rather than RGB colors at each world-space point, and built \method{} around this representation as a video world model that operates entirely within the VAE latent manifold.
By reading the cache through a single latent-resolution projection, \method{} removes the rasterise-and-encode round trip that dominates the per-step cost of RGB point cloud caches.
The decode-and-re-encode pair required to grow the cache is amortised over an entire chunk and never appears in the conditioning loop, so the per-step critical path is freed of pixel-space operations and the cache footprint shrinks by the squared VAE compression factor.
Across WorldScore and RealEstate10K, \method{} delivers state-of-the-art quality while reading from the cache an order of magnitude faster and with an order of magnitude less GPU memory than RGB-cache baselines, turning world-consistent generation into a process that scales with the rollout horizon.

\boldstart{Limitations and future work.}
The dynamic-region filter excludes moving entities from the persistent memory because their geometry is unreliable, so \method{} does not maintain the state of dynamic actors across chunks.
Scenes dominated by pervasive motion therefore benefit less from the cache than scenes dominated by rigid geometry, and persisting dynamic content across chunks is a natural direction for future work.
\vspace{1cm}

\section{Acknowledgments}
This work was supported by computing resources from Microsoft.