\section{Related Work}
\label{sec:related}


\subsection{GUI Agent: From Single-plantform to Multi-plantform}

Graphical user interface (GUI) agents interpret natural language
instructions and graphical user interfaces to automate humanz
interaction tasks in digital environments. Early GUI agent~\citep{xiaomigui0} research has largely focused on single-platform settings. Web environments provide online evaluation for browser-based navigation \citep{zhou2024webarena, koh2024visualwebarena}, where method such as WebVoyager improve web browsing. Mobile environments extend GUI agent evaluation to executable app control, cross-app workflows, personalization, and proactive assistance \citep{androidworld, mobileworld, Knowu-bench}, with methods such as UI-R1 \citep{UI-R1} and ClawGUI \citep{clawgui} exploring reinforcement learning. Desktop environments evaluate operating system level tasks \citep{osworld, bonatti2024windows, yang2026macosworld}, while recent computer-use agents such as EvoCUA \citep{evocua} and ComputerRL \citep{lai2025computerrl} study learning and decomposition for desktop automation. Recent efforts on generalist GUI agents, such as MobileAgent-v3.5 \citep{Mobile-agent-v3.5}, UI-Venus-1.5 \citep{UI-venus-1.5}, and UI-TARS-2 \citep{UI-TARS-2}, have explored unified modeling and control across heterogeneous GUI platforms. In contrast,
\METHODNAME focuses on continual learning for multi platform GUI agents, aiming
to train an end to end model for long horizon navigation while preserving
and transferring interaction capabilities across platforms.



\subsection{Multi-Teacher On-Policy Distillation}

During post training, different capabilities often exhibit a seesaw effect~\citep{scaling}: for example, mathematical RLVR may shorten reasoning traces and impair open ended writing, while each specialized training stage tends to improve one capability at the expense of others. To mitigate this issue, on policy distillation (OPD)~\citep{OPD, ExOPD, Uni-OPD} and on policy self distillation (OPSD)~\citep{OPSD, SDAR, SDFT} have emerged as effective solutions. OPD samples trajectories from the student model and matches the teacher distribution along these trajectories through reverse KL divergence, thereby providing dense token level supervision.
A natural extension of OPD is multi-teacher on-policy distillation (MOPD), which assigns the strongest checkpoint in each capability dimension as a teacher and enables the student model to absorb multiple capabilities in a single distillation stage. Recently, MOPD has been explored in foundation model post training. For example, MiMo-V2-Flash~\citep{Mimo-v2} and GLM-5~\citep{GLM-5} use MOPD as the final post training step to distill a unified model from multiple expert models. Nemotron-Cascade 2~\citep{Nemotron} uses MOPD as a forgetting recovery step between specialized RL stages, while DeepSeek-V4~\citep{DeepSeek-V4} further employs full vocabulary logits, more than ten teacher models, and dedicated infrastructure for teacher scheduling and fault tolerant trajectory generation.
However, MOPD remains largely unexplored in GUI agent. To our knowledge, \METHODNAME is the first to introduce MOPD into GUI agents. We further propose a platform conditioned distillation framework for multi platform GUI agents, enabling models to preserve and transfer interaction capabilities across different GUI platforms.
