\section{Related Work}
\subsection{LLM Agents and Skill Systems}
 
LLM agents solve complex tasks by interleaving reasoning and action \citep{Yao2023ReAct} and can improve from failures through self-reflection \citep{Shinn2023Reflexion}. To extend agents beyond the knowledge boundary of a single model, a growing body of work has explored injecting external experiential knowledge at decision time \citep{zhou2026externalization}. Early approaches store raw trajectories or reflective summaries in external memory banks and retrieve them into the context during inference \citep{Zhao2024ExpeL, Chhikara2025Mem0}. Subsequent work converged on skills as the core abstraction: \citet{Wang2023Voyager} introduced an ever-growing library of executable skills for embodied agents; \citet{Xia2026SkillRL} distill trajectories into a hierarchical SkillBank that co-evolves with the agent's policy through recursive reinforcement learning; and \citet{Wu2025EvolveR} refine interaction experience into reusable strategic principles.  On the industry side, Anthropic's Agent Skills specification has been adopted by mainstream harnesses including Claude Code, Cursor, and Gemini CLI \citep{SkillsBench2026, yang2025surveyaiagentprotocols}. All of the above deliver skills as natural-language text injected into the context window. \citet{Lu2026Skill0} take a different direction by internalizing skills into model parameters through a training-time curriculum that progressively withdraws skill context, enabling zero-shot execution at inference. Their skills, however, are irreversibly fused into the backbone, forfeiting modular flexibility. In contrast, \method{} encodes skills as LoRA modules that are independent of the backbone, simultaneously achieving zero context overhead and plug-and-play flexibility.
 
\subsection{Hypernetworks for LoRA Generation}
 
Hypernetworks generate the weights of a target network through a separate learned network \citep{Ha2017Hypernetwork} and have recently been applied to produce LoRA adapters \citep{Hu2022LoRA} for LLMs in a single forward pass, bypassing the iterative overhead of per-task fine-tuning. Text-to-LoRA \citep{charakorn2025texttolorainstanttransformeradaption} employs small MLPs to independently generate weight segments for each layer and concatenate them into a complete adapter, but this segment-wise strategy fails to capture global dependencies across layers. Generative Adapter \citep{Chen2025GenAdapter} utilizes the hidden states of the backbone LLM to produce LoRA weights, yet its parameter cost restricts coverage to only a subset of target modules. ICAE \citep{ge2024incontextautoencodercontextcompression} compresses contexts into a fixed set of compact tokens before generating LoRA, but is constrained by the resulting information bottleneck. SHINE \citep{Liu2026SHINE} extracts memory states at every layer and enables bidirectional cross-layer information flow through an attention mechanism, but does not analyze the intrinsic structure of the generated LoRA weight space. Doc-to-LoRA \citep{charakorn2026doctoloralearninginstantlyinternalize} generates adapters through a Perceiver-style encoder, but likewise focuses on context internalization. In contrast, \method{} targets the agent skill encoding setting and is the first to systematically reveal that the weight space of hypernetwork-generated LoRA exhibits structuredness, controllability, and composability.
