\subsection{Generator User Guide}

While benchmark results provide a quantitative assessment of generation quality, effectively using an omnimodal world model also requires understanding its inference interface, prompting methodology, and sampling configurations. Because Cosmos 3 supports a diverse set of modalities and generation modes—including image, video, audio-visual, transfer, forward-dynamics, inverse-dynamics, and policy generation—the quality of the outputs depends not only on the model weights but also on how generation requests are specified and conditioned. In this section, we provide practical guidance for using Cosmos 3 Generator, including recommended prompting strategies, structured caption formats, sampling hyperparameters, and the role of the Cosmos 3 Reasoner as a prompt upsampler. These guidelines reflect the configurations used throughout our evaluations and serve as a reference for obtaining high-quality, physically plausible generations across a wide range of Physical AI applications.

\subsubsection{Generation Guide}
\label{sec::gen_guide}

The Cosmos 3 Generator supports flexible visual (and audio-visual) generation across a broad inference envelope: frame rates from 10–30 FPS, 5 to 400 frames, resolutions spanning 256p, 480p, and 720p, and common aspect ratios (1:1, 3:4, 4:3, 9:16, 16:9). This allows a single model to serve use cases ranging from short preview clips to longer, higher-resolution landscape or portrait video without changing the sampling interface. The Cosmos 3 Generator is trained for forward dynamics, inverse dynamics and policy modes to support different action-related applications. For action generation, the base Cosmos3-Nano and Cosmos3-Super models support action prediction at native control frequencies ranging from 10–30 FPS, with prediction horizons spanning 16–400 frames across inverse dynamics and policy modes. Post-trained models specialize to a single mode and frequency---for example, Cosmos3-Nano-Policy-DROID operates at 15 FPS with a 32-step prediction horizon. Below, we detail the key components for successful generation using Cosmos 3 Generator. These details are also summarized in \cref{tab:sampling_configs}.

\input{tables/results/generation_settings}

\paragraph{Media specifications and prompting guide.}
The Cosmos 3 Generator is trained on structured JSON captions that provide fine-grained control over scene composition, covering subjects, background, lighting, aesthetics, cinematography, and for video, temporal fields such as actions, state changes, camera motion, and segment-level descriptions (full schema in Appendix~\ref{appendix:captioning_details}). At inference time, a prompt upsampler---served either by Claude Opus 4.6 or by the Cosmos 3 Reasoner---converts user requests into this same structured format, ensuring that generation prompts match the distribution seen during training (see Appendix~\ref{appendix:upsampler_prompt_template} for template instructions). The upsampler is instructed to first describe the scene layout and world state, then specify the temporal progression of events, and finally add any audio descriptions. The specific upsampler instruction varies slightly across generation modes---for instance, action and transfer generation impose additional task constraints tailored to their conditioning inputs (see Appendix~\ref{appendix:generator_transfer_prompt_prefix} for the transfer generation prompt prefix, and Appendix~\ref{appendix:action_upsampler_prompt} for the action prompt guide). The JSON specification additionally includes explicit media controls (duration, FPS, spatial height and width, and aspect ratio), keeping prompt interpretation and sampling configuration inspectable and reproducible.

\paragraph{Negative prompt.}
We tune negative prompts separately for each model and generation mode through automated benchmark iteration. For each configuration, we ablate over candidate templates spanning natural-language descriptions, keyword lists, instruction-style directives, compositional extensions targeting physical consistency and identity preservation, and the null string. The best-performing variant is selected based on automated benchmark scores. For the base Cosmos3-Nano and Cosmos3-Super generators, the explicit negative prompt can be found in Appendix~\ref{appendix:generator_negative_prompt}. For the post-trained variants, we found that the null string negative prompt works best for Cosmos3-Super-Text2Image, while using negative prompts automatically derived from the user prompts yields the best result for Cosmos3-Super-Image2Video (Appendix~\ref{appendix:sft_i2v_upsampler_prompt}). For action generation modes, we found the null string negative prompt to work best.

\paragraph{Generation sampling hyperparameters.} We adopt the following sampling parameters for different modalities:
\begin{enumerate}
    \item \textit{Audio-visual generation.} For audio-visual generation, we tune sampling hyperparameters on automated benchmarks for image and video generation (\cref{subsec::image_eval} and \cref{subsec::video_eval}). For the base Cosmos3-Nano and Cosmos3-Super generators, we use 50 denoising steps, a guidance scale of 6, a time shift of 10, and full-range classifier-free guidance. For the post-trained Cosmos3-Super-Text2Image model, we use a guidance scale of 4 and a time shift of 3. For the post-trained Cosmos3-Super-Image2Video model, we use a shift of 5.
    \item \textit{Action generation.} Action generation has three supported modes. For forward and inverse dynamics, we use 50 denoising steps, a guidance scale of 1, a time shift of 5, and full-range classifier-free guidance. For policy mode, we switch to 4 denoising steps and a guidance scale of 3.
    \item \textit{Transfer generation.} We tune sampling hyperparameters for video transfer generation on the automated PAIBench-C benchmark (\cref{subsec::transfer_eval}). We use 50 denoising steps, a text guidance scale of 3, a control guidance scale of 1.5, a time shift of 10, and full-range classifier-free guidance.
\end{enumerate}
