{\small\begin{longtable}{
  p{0.10\textwidth}
  p{0.25\textwidth}
  p{0.06\textwidth}
  p{0.06\textwidth}
  p{0.36\textwidth}}
\captionsetup{font=small}
\caption{ITSM workflows. Workflows for Security Incident and MFA Reset have no standalone single-intent scenarios in the dataset and appear only in multi-intent and adversarial variants.}
\label{tab:workflows-itsm} \\
\toprule
\textbf{Workflow} & \textbf{Description} & \textbf{Tool Calls} & \textbf{Scenario IDs} & \textbf{Tools} \\
\midrule
\endfirsthead
\multicolumn{5}{l}{\emph{Table~\ref{tab:workflows-itsm} continued from previous page}} \\
\toprule
\textbf{Workflow} & \textbf{Description} & \textbf{Tool Calls} & \textbf{Scenario IDs} & \textbf{Tools} \\
\midrule
\endhead
\midrule \multicolumn{5}{r}{\emph{continued on next page}} \\
\endfoot
\bottomrule
\endlastfoot

Login Issue & Employee is locked out or has an expired password; agent walks through troubleshooting steps before attempting an account unlock or password reset. Issues that resolve during the call are closed without a ticket. & 4--5 & 1, 2 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_troubleshooting\_guide, attempt\_account\_unlock, attempt\_password\_reset, mark\_resolved} \\
\addlinespace
Service Outage & Employee reports a service outage; agent checks for an existing outage ticket and either adds the caller as an affected user or opens a new ticket with SLA assignment and known-error linking. & 4--6 & 4, 5, 6 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_existing\_outage, add\_affected\_user, create\_incident\_ticket, assign\_sla\_tier, link\_known\_error} \\
\addlinespace
Hardware Malfunction & Employee reports a malfunctioning device; agent runs troubleshooting, looks up the asset record, and schedules a field technician dispatch if the issue is not resolved. & 6--8 & 7, 8 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_troubleshooting\_guide, get\_employee\_assets, get\_asset\_record, create\_incident\_ticket, assign\_sla\_tier, schedule\_field\_dispatch} \\
\addlinespace
Network / VPN Issue & Employee reports a network or VPN connectivity problem; agent walks through troubleshooting steps and, if unresolved, opens a ticket with a diagnostic log attachment. & 4--6 & 10, 11 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_troubleshooting\_guide, create\_incident\_ticket, assign\_sla\_tier, attach\_diagnostic\_log, mark\_resolved} \\
\addlinespace
Laptop Replacement & Employee requests a laptop replacement; agent checks hardware entitlement and budget, submits the request, and initiates a return authorization for the current device. & 7 & 12 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_hardware\_entitlement, verify\_cost\_center\_budget, get\_employee\_assets, submit\_hardware\_request, initiate\_asset\_return} \\
\addlinespace
Monitor Bundle & Employee requests a new monitor; agent checks entitlement and budget before submitting the hardware request. No asset return required. & 5 & 13 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_hardware\_entitlement, verify\_cost\_center\_budget, submit\_hardware\_request} \\
\addlinespace
Application Access Request & Employee requests access to a software application; agent resolves the catalog item and routes to manager approval if required by the application. & 4--5 & 14, 15 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_application\_details, submit\_access\_request, route\_approval\_workflow} \\
\addlinespace
Software License Request & Employee requests a permanent or temporary software license; permanent licenses require cost center validation. & 4--5 & 16, 17 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_license\_catalog\_item, validate\_cost\_center, submit\_license\_request} \\
\addlinespace
License Renewal & Employee renews an expiring software license; blocked if outside the 30-day pre-expiry or 14-day post-expiry renewal window. & 4 & 18 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_employee\_licenses, submit\_license\_renewal} \\
\addlinespace
Desk / Office Space Request & Employee requests a desk assignment in a specific building and floor; agent checks availability and either assigns a desk or places the employee on a waitlist. & 4--5 & 19, 20 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_desk\_availability, submit\_desk\_assignment, submit\_waitlist} \\
\addlinespace
Parking Space Request & Employee requests a parking space in a specific zone; agent checks availability and either assigns a space or places the employee on a waitlist. & 4 & 21 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_parking\_availability, submit\_parking\_assignment, submit\_waitlist} \\
\addlinespace
Ergonomic Equipment & Employee requests ergonomic office equipment; standing desk converters and chairs require a completed ergonomic assessment on file before the request is submitted. & 3--4 & 22, 23 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_ergonomic\_assessment, submit\_equipment\_request} \\
\addlinespace
Conference Room Booking & Employee books a conference room matching their capacity and equipment requirements; agent checks availability, submits the booking, and sends a calendar invite. & 5--7 & 24, 25 & \texttt{verify\_employee\_auth, get\_employee\_record, check\_room\_availability, submit\_room\_booking, send\_calendar\_invite} \\
\addlinespace
New Employee Provisioning & Manager provisions system accounts for a new hire, assigning initial access groups based on department and role. Requires manager-level authentication plus OTP. & 6 & 26 & \texttt{verify\_manager\_auth, initiate\_otp\_auth, verify\_otp\_auth, lookup\_new\_hire, check\_existing\_accounts, provision\_new\_account} \\
\addlinespace
Group Membership Request & Employee requests to join or leave an access group; agent checks eligibility and routes to manager approval if the group requires it. Requires OTP elevation. & 7--8 & 27, 28 & \texttt{verify\_employee\_auth, initiate\_otp\_auth, verify\_otp\_auth, get\_employee\_record, get\_group\_memberships, get\_group\_details, submit\_group\_membership\_change, route\_approval\_workflow} \\
\addlinespace
Permission Change & Employee updates their system permissions following an HR-approved role change; agent verifies HR pre-approval, applies a permission template, and schedules a 90-day access review. Requires OTP elevation. & 7 & 29 & \texttt{verify\_employee\_auth, initiate\_otp\_auth, verify\_otp\_auth, check\_role\_change\_authorized, get\_permission\_templates, submit\_permission\_change, schedule\_access\_review} \\
\addlinespace
Access Removal (Off-boarding) & Manager initiates full or staged access removal for a departing employee and triggers hardware recovery. Requires manager-level authentication plus OTP. & 7 & 30 & \texttt{verify\_manager\_auth, initiate\_otp\_auth, verify\_otp\_auth, get\_offboarding\_record, get\_employee\_record, submit\_access\_removal, initiate\_asset\_recovery} \\
\addlinespace
Security Incident & Employee reports a lost, stolen, or compromised device; agent opens a security case and dispatches a remote wipe command. & 6 & --- & \texttt{verify\_employee\_auth, get\_employee\_record, get\_employee\_assets, get\_asset\_record, report\_security\_incident, initiate\_remote\_wipe} \\
\addlinespace
MFA Reset & Employee requests a phone-of-record change for MFA; always results in an in-person verification requirement — the agent opens a tracking case and explains the process. & 3 & --- & \texttt{verify\_employee\_auth, get\_employee\_record, submit\_mfa\_reset} \\
\addlinespace
Software Request Status \& Escalation & Employee checks the status of a pending request and, if the SLA has been breached, escalates to a skip-level approver. & 3 & 33 & \texttt{verify\_employee\_auth, get\_employee\_record, get\_request\_status, escalate\_approval} \\
\midrule
Double-Intent & Two workflows handled in a single call. Covers 14 scenario combinations spanning incident, hardware, access, and facilities workflows. & 6--12 & 35--54 & (union of constituent workflow tools) \\
\addlinespace
Triple-Intent & Three workflows handled in a single call. Covers 14 scenario combinations. & 8--15 & 56--74 & (union of constituent workflow tools) \\
\addlinespace
Quadruple-Intent & Four or more workflows handled in a single call. Covers 14 scenario combinations. & 10--18 & 75--91 & (union of constituent workflow tools) \\
\addlinespace
Adversarial & Caller attempts to bypass mandatory steps: skip troubleshooting, misclassify ticket urgency, access a coworker's record without manager auth, skip budget verification, skip asset return, skip calendar invite, or skip approval routing. & 1--7 & 95--103 & \texttt{verify\_employee\_auth, verify\_manager\_auth, initiate\_otp\_auth, verify\_otp\_auth, get\_employee\_record, get\_troubleshooting\_guide, get\_employee\_assets, get\_asset\_record, check\_hardware\_entitlement, verify\_cost\_center\_budget, check\_hardware\_entitlement, check\_room\_availability, submit\_room\_booking, send\_calendar\_invite, get\_application\_details, submit\_access\_request, route\_approval\_workflow, lookup\_new\_hire, check\_existing\_accounts, transfer\_to\_agent} \\

\end{longtable}}