\section{Experiments}
\subsection{Experimental setups}
% \paragraph{Benchmark and Metrics.}



\paragraph{Benchmark and Metrics.}
For reward model evaluation, we curated a high-quality and diverse set of 5,000 reference images and instructions from the same public image-editing benchmark. 
We then utilized various models, including SeedEdit-3.0~\citep{wang2025seededit}, BAGEL~\citep{deng2025emerging}, and FLUX.Kontext~\citep{batifol2025flux}, to produce several edited outputs for each input. Finally, these generated outputs were manually annotated using pairwise preference comparisons. 
Annotators were also allowed to mark a pair as ``same'' when the two edits were of comparable quality or when no reliable preference could be established. These ambiguous pairs were excluded when evaluating verifier accuracy, yielding a cleaner pairwise benchmark.
The accuracy of the reward model in predicting these human-annotated preferences is used as our evaluation metric. We also report performance on the public EditRewardBench~\citep{luo2025editscore} for a comprehensive comparison.
For image editing model evaluation, we adopt GEdit-Bench-EN~\citep{liu2025step1x}, a standardized benchmark with multi-dimensional automatic metrics. Following the original protocol, we report scores across three key aspects, each assessed by GPT-4.1: semantic consistency (SC), which measures how well the edited image aligns with the given instruction; perceptual quality (PQ), which captures the visual fidelity of the edited image; and overall score (O), computed as the geometric mean of SC and PQ. In addition, we report SC scores for different categories, as presented in Tab.~\ref{tab:model_evaluation}.
% \vspace{-.8em}
% \paragraph{Implementation Details.}

% Our RRM is built on Qwen-VL-2.5~\citep{bai2025qwen2}. Cold-start SFT uses edits generated by SeedEdit3.0, BAGEL, and FLUX.Kontext~\citep{wang2025seededit,deng2025emerging,batifol2025flux}. GCPO uses 10k human preference pairs ($<1\%$ of the SFT-scale data), suggesting that its gains mainly come from better human alignment. For downstream optimization, we apply \editrone{} to FLUX.Kontext and Qwen-Image-Edit with Flow-GRPO~\citep{liu2025flow} ($G=24$, $\beta=0.04$). Training remains practical with packed inference.


\paragraph{Implementation Details.}
Our RRM is built on the open-source Qwen-VL-2.5~\citep{bai2025qwen2}.
In the cold-start SFT phase in Sec.~\ref{sec:cold_start_reward}, we constructed editing pairs using a mixture of models, including SeedEdit3.0~\citep{wang2025seededit}, BAGEL~\citep{deng2025emerging}, and FLUX.Kontext~\citep{batifol2025flux}.
% For the \gcpo{} phase in Sec.~\ref{sec:r3l}, we further collected 10k human-annotated preference pairs.
For the \gcpo{} phase in Sec.~\ref{sec:r3l}, we further collected 10k human-annotated preference pairs, which is less than 1\% of the SFT-scale training data. Therefore, the gains from GCPO are mainly attributable to better human alignment rather than increased data volume.
For editing model optimization, we apply our \editrone{} framework to two strong open-source models: FLUX.Kontext~\citep{batifol2025flux} and Qwen-Image-Edit~\citep{wu2025qwen}.
The models are optimized using the \grpo{} strategy described in Sec.~\ref{sec:rl_for_editing}, with our trained \rrm{} serving as the reward signal.
We adopt Flow-GRPO~\citep{liu2025flow} with a group size of $G=24$ and a KL penalty coefficient of $\beta=0.04$.
Although GCPO requires rollouts from both preferred and non-preferred groups, the overall training cost remains manageable in practice due to the small rollout group size and efficient packed inference.
\subsection{Reward Model Performance}
Our proposed \editrone{} framework yields a state-of-the-art reward model for predicting human preferences. As shown in Tab.~\ref{tab:full_rm_results}, our 7B Edit-RRM, trained via our full two-stage pipeline, achieves a top accuracy of \textbf{82.2\%}. 
% Tab.~\ref{tab:full_rm_results} shows that our 7B Edit-RRM achieves the best accuracy of 82.2\% on the internal benchmark.
This result significantly surpasses strong closed-source APIs like Seed-1.5-VL (79.3\%) and demonstrates the effectiveness of our training strategy.
\begin{table}[h] 
    \centering
    \caption{Accuracy on our internal benchmark. \textbf{T}, \textbf{V}, and \textbf{T+V} denote Think, Verify, and Think+Verify, respectively.}
    % \vspace{-.5em}
    \label{tab:full_rm_results}
    \renewcommand{\arraystretch}{1.2}
    \setlength{\tabcolsep}{3pt} 
    \small
    \begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}l ccc c}
        \toprule
        \textbf{Model} & \textbf{T} & \textbf{V} & \textbf{T+V} & \textbf{GCPO} \\
        \midrule
        \multicolumn{5}{l}{\textit{Inference w. API Baselines}} \\
        Seed-1.5-VL~\citep{guo2025seed1}  & 72.2\% & --- & 79.3\% & --- \\
        Seed-1.6-VL~\citep{guo2025seed1}  & 71.2\% & 69.4\% & 77.2\% & --- \\
        \midrule
        \multicolumn{5}{l}{\textit{Our Method (SFT \& RL Stages)}} \\
        Qwen-7B (VIESCORE) & \multicolumn{3}{c}{68.3\%} & --- \\
        \midrule
        Qwen-3B  & 64.1\% & 66.1\% & 69.3\% & 72.0\% \\
        Qwen-7B  & 68.9\% & 70.9\% & 75.4\% & \textbf{82.2\%} \\
        \bottomrule
    \end{tabular*}
\end{table}
% \vspace{-1.5em}
\paragraph{Verifier vs. Scorer on Public Benchmarks.}
% On the public EditRewardBench, our verifier-based RRM outperforms the holistic scorer baseline: 73.3\% vs. 65.9\% after SFT, and 78.2\% after GCPO (Tab.~\ref{tab:editscore_benchmark_comparison}). Since this benchmark is independent of our internal pipeline, the gain also indicates good generalization.
On the public EditRewardBench, our verifier-based RRM consistently outperforms the holistic scorer baseline. As shown in Tab.~\ref{tab:editscore_benchmark_comparison}, our SFT-RRM already surpasses EditScore-7B (73.3\% vs. 65.9\%), and GCPO further improves the final RL-RRM to 78.2\%. Since EditRewardBench is independently constructed from our internal pipeline, this gain indicates that our improvement is not due to internal benchmark bias. 
% \vspace{-1.2em}
\begin{table*}[t]
% \small
\centering
\caption{
Detailed performance comparison on the GEdit-Bench-EN (Full set). Higher scores are better. \textbf{Bold} scores highlight the best result within each model family. 
Columns 1–11 report SC scores for different editing categories (see Appendix for details).
%"O" stands for the Overall Score. 
}
\label{tab:model_evaluation}
\renewcommand{\arraystretch}{1.25} 
% \setlength{\tabcolsep}{4pt} 
\resizebox{\textwidth}{!}{
\begin{tabular}{l *{11}{S[table-format=1.2]} S[table-format=1.2] S[table-format=1.2] S[table-format=1.2]}
\toprule
\multirow{2}{*}{\textbf{Model}} & \multicolumn{11}{c}{\textbf{Category SC}} & \multicolumn{3}{c}{\textbf{Overall}} \\
\cmidrule(lr){2-12} \cmidrule(lr){13-15}
& {1} & {2} & {3} & {4} & {5} & {6} & {7} & {8} & {9} & {10} & {11} & {SC $\uparrow$} & {PQ $\uparrow$} & {O $\uparrow$} \\
\midrule
% ---------- Group 1: Edited Models ----------
\multicolumn{15}{l}{\textbf{Edited Models}} \\
Step-Edit~\citep{liu2025step1x}       & 8.77 & 8.90 & 7.52 & 4.35 & 4.10 & 7.73 & 8.56 & 7.81 & 8.26 & 2.82 & 7.30 & 6.53 & 6.72 & 5.90 \\
UniPic2~\citep{wei2025skywork}        & 8.07 & 8.70 & 6.75 & 3.57 & 4.78 & 7.13 & 8.36 & 8.36 & 7.87 & 5.39 & 7.97 & 6.84 & 7.24 & 6.41 \\
Bagel~\citep{deng2025emerging}        & 8.54 & 8.32 & 7.42 & 4.97 & 5.07 & 7.71 & 8.75 & 8.03 & 8.22 & 7.14 & 6.62 & 7.32 & 7.02 & 6.65 \\
GPT-4o                               &      &      &      &      &      &      &      &      &      &      &      & 7.74 & 8.13 & 7.49 \\
\midrule
% ---------- Group 2: FLUX.Kontext Family ----------
\multicolumn{15}{l}{\textbf{FLUX.Kontext Family}~\citep{batifol2025flux}} \\
FLUX.Kontext  & 7.65 & 8.17 & 6.90 & 3.02 & 3.54 & 6.86 & 7.43 & 7.54 & 6.95 & 5.14 & 7.85 & 6.27 & 7.25 & 5.77 \\
\rowcolor{gray!10}
RL w. SeedVLM-1.5                    & 8.23 & 8.41 & 7.00 & \textbf{5.00} & 4.05 & 7.17 & 7.93 & 7.60 & 7.11 & 5.51 & 8.60 & 6.74 & 6.44 & 6.03 \\
\rowcolor{gray!10}
RL w. SFT-RRM (3B)                   & 7.48 & 8.25 & 6.65 & 3.20 & 4.01 & 7.25 & 7.77 & 7.15 & 7.61 & 5.70 & 7.90 & 6.52 & 6.26 & 5.63 \\
\rowcolor{gray!20}
RL w. RL-RRM (3B)                    & 8.13 & 7.65 & \textbf{7.33} & 3.63 & 3.70 & 7.25 & 7.78 & 7.65 & \textbf{7.68} & 6.03 & \textbf{7.93} & 6.67 & 7.09 & 6.10 \\
\rowcolor{gray!10}
RL w. SFT-RRM (7B)                   & \textbf{8.25} & 8.52 & 7.10 & 3.85 & 3.68 & 7.27 & \textbf{8.26} & 7.68 & 7.38 & 6.44 & 7.80 & 6.81 & \textbf{7.25} & 6.20 \\
\rowcolor{gray!20}
RL w. RL-RRM (7B)                    & 8.15 & \textbf{8.62} & 6.82 & 4.42 & \textbf{4.22} & \textbf{7.30} & 8.21 & \textbf{7.99} & 7.41 & \textbf{6.44}  & 7.82 & \textbf{6.86} & 7.20 & \textbf{6.24} \\
\midrule
% ---------- Group 3: Qwen-Edit Family ----------
\multicolumn{15}{l}{\textbf{Qwen-Edit Family}~\cite{wu2025qwen}} \\
Qwen-Edit                            & \textbf{8.85} & 9.02 & \textbf{8.20} & 4.01 & 6.04 & 7.61 & \textbf{8.80} & 8.32 & 8.74 & \textbf{9.00} & \textbf{8.37} & 7.94 & \textbf{7.78} & 7.45 \\
\rowcolor{gray!20}
RL w. RL-RRM (7B)                & 8.75 & \textbf{9.05} & 8.10 & \textbf{4.62} & \textbf{6.17} & \textbf{7.76} & 8.72 & \textbf{8.43} & \textbf{8.79} & 8.69 & 8.32 & \textbf{7.99} & 7.76 & \textbf{7.50} \\
\bottomrule
\end{tabular}
}
\end{table*}
\paragraph{Principled Data Curation and SFT.}
% The foundation of our model's performance is a meticulously curated SFT dataset. Tab.~\ref{tab:full_rm_results} details the ablation studies on the components of our data generation process. For our Qwen-based models, the full SFT pipeline using both ``Think'' and ``Verify'' components (T+V) consistently yields the best performance before the GCPO stage. For instance, the 7B model's accuracy improves from 68.9\% (T only) to 75.4\% (T+V), highlighting the importance of rigorous filtering. As a baseline, we also trained with SFT data generated with prompts from VIESCORE~\cite{ku2024viescore}, achieved 68.3\% accuracy.
The foundation of our model's performance is a carefully curated SFT dataset. Tab.~\ref{tab:full_rm_results} shows that the full SFT pipeline with both ``Think'' and ``Verify'' (T+V) consistently performs best before GCPO. For Qwen-7B, accuracy improves from 68.9\% with ``Think'' only to 75.4\% with ``Think+Verify'', highlighting the importance of principled decomposition and rigorous filtering. As a reference baseline, SFT data generated with VIESCORE prompts~\cite{ku2024viescore} achieves 68.3\%. These results support our claim that each component in the cold-start pipeline is important for reward-model supervision.
% \vspace{-.8em}
\paragraph{Ablation on GCPO and Benchmarks.} 
% We conducted an ablation study on the public EditScore-RM benchmark (Tab.~\ref{tab:editscore_benchmark_comparison}). The \gcpo{} stage provides a significant performance boost, improving accuracy from 73.3\% to 78.2\%. This demonstrates that \gcpo{} is critical for refining the model's alignment with human preferences. Notably, our final RRM (78.2\%) substantially outperforms the original EditScore-7B model, even with its inference scaling enhancement.
We further analyze the impact of \gcpo{} on the public EditScore-RM benchmark (Tab.~\ref{tab:editscore_benchmark_comparison}). GCPO provides a clear performance boost, improving accuracy from 73.3\% to 78.2\%, which shows that it is critical for refining the model's alignment with human preferences. Notably, our final RRM also outperforms EditScore-7B with inference scaling. Since the GCPO stage uses only 10k human preference pairs---less than 1\% of the SFT-scale data, whose gain is mainly attributable to better human alignment.
\paragraph{Impact of Principled Data Curation.}
Our initial experiments (not shown in the main tables for brevity) confirmed that both the ``Think" (reasoning generation) and ``Verify" (external filtering) components are critical. Removing the ``Verify" step led to a significant accuracy degradation, highlighting the importance of rigorous data filtering. Meanwhile, the reasoning traces from the ``Think" step provide essential supervisory signals that enhance the model's evaluative capabilities.
% \vspace{-.6em}
\subsection{Image Editing Performance}
Fig.~\ref{fig:edit_dynamics} shows that our RRMs provide stable reward signals, while GCPO-refined RRMs yield higher evaluation rewards and act as stricter evaluators.
\vspace{-.5em}
\paragraph{Overall Performance.}
As shown in Tab.~\ref{tab:model_evaluation}, our framework demonstrates strong performance on both model families. Optimizing FLUX.Kontext with our RL-RRM (7B) boosts its Overall Score (O) from 5.77 to 6.24 and its Semantic Consistency (SC) score from 6.27 to 6.86. This result exceeds others, confirming that our RRM is a highly effective and competitive reward source for policy optimization.
Meanwhile, we carried out experiments on the SOTA open resource, Qwen-Edit, whose overall score showed a modest improvement from 7.45 to 7.50. This is largely because the baseline model already benefits from Best-of-N scaling~\cite{luo2025editscore}. However, the validity of Edit-R1 is highlighted in its ability to address the model's specific weaknesses.
Notably, as shown in the final row of Tab.~\ref{tab:model_evaluation}, our framework yields a significant 15.2\% relative gain (from 4.01 to 4.62) in the challenging Motion Change category, demonstrating its effectiveness in enhancing performance even on highly optimized models. These gains are further supported by human evaluation: FLUX.Kontext optimized with our RL-RRM (7B) achieves a GSB score of +23.2 against the original FLUX.Kontext baseline. The user-study protocol and full human-evaluation details are provided in Appendix~\ref{sec:human_eval}, and additional qualitative examples are shown in Fig.~\ref{fig:qualitative_results}.
% Notably, as shown in the final row of Tab.~\ref{tab:model_evaluation}, our framework yields a significant 15.2\% relative gain (from 4.01 to 4.62) in the challenging Motion Change category, demonstrating its effectiveness in enhancing performance even on highly optimized models. These gains are also supported by human evaluation, where the user-study protocol and GSB results are provided in Appendix~\ref{sec:human_eval}. Additional qualitative examples are shown in Fig.~\ref{fig:qualitative_results} in the Appendix.
% Notably, as shown in the final row of Tab.~\ref{tab:model_evaluation}, our framework yields a significant 15.2\% relative gain (from 4.01 to 4.62) in the challenging Motion Change category, demonstrating its effectiveness in enhancing performance even on highly optimized models.
\paragraph{Impact of GCPO and Scalability.} 
As shown in Tab.~\ref{tab:model_evaluation}, RRM with GCPO consistently outperforms its SFT counterpart. Training curves in Fig.~\ref{fig:edit_dynamics} further reveal the mechanism. Although RL-RRMs provide lower training rewards, they yield higher evaluation rewards, indicating that GCPO transforms the reward model into a stricter and more reliable evaluator. This stricter supervision pushes editing models to adhere more closely to human preferences and achieve higher-quality outputs.
\begin{table}[t!]
    \centering
    \caption{Comparison of our RRM against the baseline on the EditReward benchmark. All results are for 7B models.}
    \label{tab:editscore_benchmark_comparison}
    \renewcommand{\arraystretch}{1.15}
    \small
    \begin{tabular}{lc}
        \toprule
        \textbf{Method} & \textbf{Accuracy (\%)} \\
        \midrule
        \multicolumn{2}{l}{\textit{Baseline Model}} \\
        EditScore-7B & 65.9\% \\
        \quad + \textit{inference scaling} & 72.7\% \\
        \midrule
        \multicolumn{2}{l}{\textit{Our Method}} \\
        Our RRM (SFT only) & 73.3\% \\
        Our RRM (SFT + GCPO) & \textbf{78.2\%} \\
        \bottomrule
    \end{tabular}
\end{table}
% \vspace{-1.0em}
\paragraph{Qualitative Analysis.} 
Beyond quantitative metrics, the qualitative improvements are equally compelling. As shown in the Appendix (Fig.~\ref{fig:qualitative_results}), models optimized with our framework exhibit markedly better instruction adherence and visual fidelity. 
In challenging edits such as subject addition/removal and motion change, our method successfully follows the instruction, whereas the baseline fails. For localized edits such as Color Alter, our approach precisely modifies the target object without introducing global color shifts or artifacts. These qualitative results provide concrete evidence of the practical effectiveness of our framework.
% \vspace{-1.0em}

