\subsubsection{Video Generation Evaluation}
\label{subsec::video_eval}

We evaluate the video generation capabilities of Cosmos 3 through complementary automated and human benchmarks. Automated benchmarks offer scalable, reproducible comparisons and capture a broad range of quality and domain-specific signals, but their discriminative power diminishes as models improve (especially in domain-specific Physical-AI regimes, where such metrics are usually myopic to temporal and physics-based failures). Human evaluation addresses these gaps by catching long-tail physics, embodiment, and semantic failures that automated protocols systematically miss, while spreading scores across a meaningfully wider range so that small but real differences between state-of-the-art models remain detectable. We report results on three automated benchmarks---PAIBench-G, RBench, and Physics-IQ---followed by Cosmos HUE, a dedicated human evaluation protocol for Physical AI video generation, and Human World Bench (HWB), targeted human evaluation for realistic human motion from task-level instructions. For all benchmarks (automated or human-evaluation), we use Claude-Opus-4.6 as a prompt rewriter to convert text prompts into a structured format described in~\cref{sec::gen_guide} and~\cref{sec::prompt_upsampling}, ensuring they match the training prompt distribution.

\phantomsection\label{subsubsec::paibench_g}
\paragraph{PAIBench-G.} PAIBench-G is the video generation track of the Physical AI Benchmark~\citep{zhou2025paibench}, comprising 1,044 image–text prompt pairs across six Physical AI domains: Human (299), Autonomous Vehicle (239), Common Sense (174), Robotics (107), Physics (107), and Industry (107). We adopt it for its broad domain coverage and balanced scoring: each model receives a Quality Score (aggregating metrics for frame consistency, motion smoothness, aesthetic quality, and video–text alignment) and a Domain Score (VLM-as-Judge binary verification of physical and semantic accuracy across domains). The overall score weights both equally: $\text{Overall} = 0.5 \times \text{Quality} + 0.5 \times \text{Domain}$. With a Pearson correlation of
r=0.918 against human ELO rankings, PAIBench-G is well-suited for reliably measuring progress in early training phases, before models converge to the frontier where automated scores begin to saturate. For every prompt, we generate videos across 5 seeds and evaluate at 720p resolution, 16:9 aspect ratio for 189 frames. While PaiBench-G natively supports only Image-to-Video evaluation, we extend it to compute Text-to-Video scores as well\footnote{We found that the public PAIBench-G Image-to-Video leaderboard results (judged by Qwen3-VL-235B-A22B) were not reproducible. We therefore use Qwen2.5-VL-72B-Instruct as the domain-score judge for our internal evaluation (\cref{tab:paibench_rbench_results_combined}) and independently submit to the public leaderboard; on both, Cosmos3-Super and Cosmos3-Nano rank first and second overall.}. \cref{tab:paibench_rbench_results_combined} reports both text-to-video and image-to-video PAIBench-G results using Qwen2.5-VL-72B-Instruct as the VLM judge; Cosmos3-Super achieves state-of-the-art overall score on both tracks, emerging as the best open-source model, outperforming strong closed-source models such as Veo-3.1.

\input{tables/results/paibench_combined}

\paragraph{RBench.} RBench~\citep{deng2026rbench} evaluates video generation in embodied task scenarios, emphasizing task correctness and physical plausibility in robot–object interactions over purely perceptual realism. While PAIBench-G covers Physical AI domains broadly, RBench delves deeper into robotics as a critical Physical AI use case, stress-testing models on generating physically coherent manipulation sequences across diverse robot morphologies. The benchmark comprises 650 image–text evaluation cases from two complementary splits: 250 task-oriented pairs spanning five categories (Common Manipulation, Long-horizon Planning, Multi-entity Collaboration, Spatial Relationship, and Visual Reasoning) and 400 embodiment-specific pairs covering four robot morphologies (Dual-arm, Humanoid, Single-arm, and Quadruped). Prompts are sourced from RoVid-X~\citep{deng2026rbench}, a curated corpus of 4M robotics clips spanning over 1,300 skills. Each video is scored on Task Completion (TC, averaging physical-semantic plausibility and task-adherence consistency) and Visual Quality (VQ, a penalized combination of robot–subject stability and motion smoothness), and the final score is the mean of both across all 650 cases. With a Spearman correlation of $\rho = 0.96$ with human judgments across 25 evaluated models, RBench provides a reliable signal for embodied generation quality. For every prompt, we generate videos with a single seed and evaluate at 720p resolution, 16:9 aspect ratio for 121 frames. \cref{tab:paibench_rbench_results_combined} reports Cosmos 3 image-to-video results on RBench -- Cosmos 3 emerges as the best open-source model.

As shown in \cref{tab:paibench_rbench_results_combined}, Cosmos3-Super achieves the highest overall PAIBench-G scores on both T2V and I2V across all models, including closed-source, while Cosmos3-Nano matches the second best result on RBench. Both Cosmos 3 variants also significantly outperform their Cosmos-Predict2.5 predecessors, reflecting the gains from the omnimodal architecture. While both benchmarks incorporate physical plausibility into their scoring, they do not probe it in depth. We therefore turn to Physics-IQ for a targeted evaluation of physics adherence.

\paragraph{Physics-IQ.} Physics-IQ~\citep{motamed2025generative} evaluates whether video generation models capture physical principles by conditioning on a real-world starting context and scoring how closely the generated continuation matches an actual physical outcome. Unlike PAIBench-G and RBench, which incorporate physical plausibility as one scoring component among many, Physics-IQ isolates it as the sole evaluation axis, measuring whether generated motion matches ground-truth physical outcomes along precise spatial and temporal dimensions.
Physics-IQ covers five physics categories---solid mechanics, fluid dynamics, optics, thermodynamics, and magnetism---across 396 real-world scenes captured from three fixed viewpoints, with paired textual descriptions for text-conditioned models.
Two evaluation modes are supported.
In \emph{image-to-video} (I2V), the model is conditioned on a single \emph{switch frame} plus an optional text prompt and predicts the subsequent motion.
In \emph{video-to-video} (V2V) continuation, the model is conditioned on a 3\,s conditioning video plus an optional text prompt and predicts the next 5\,s of motion.
Generated videos are compared against ground-truth physical continuations along four complementary axes (spatial overlap, temporal alignment, magnitude-weighted spatial agreement, and pixel-level error), normalized by a real-vs-real upper bound into a single 0–100 score directly comparable across models and conditioning modes.

\input{tables/results/physics_iq}

We evaluate Cosmos3-Super in both I2V and V2V modes in \cref{tab:physics_iq_results} alongside leading open-source and commercial baselines in both modes.
For V2V, we use the full 3\,s conditioning video as input.
We also use a prompt upsampler, following a similar iterative-refinement strategy to PhyT2V~\citep{xue2025phyt2v}, to generate text prompts from the switching frame in the I2V setting and from the 3\,s conditioning video in the V2V setting.
Following the public Physics-IQ leaderboard and the WMReward inference-time alignment protocol~\citep{yuan2026wmreward}, we additionally run WMReward + best-of-$N$ (BoN) scores for Cosmos3-Super in both I2V and V2V.
From \cref{tab:physics_iq_results}, we find that Cosmos3-Super achieves state-of-the-art in I2V: its direct score of $43.8$ exceeds the direct I2V baselines, and WMReward+BoN further improves the score to $48.9$, above the strongest listed I2V baseline with WMReward+BoN.
In V2V, Cosmos3-Super also achieves the state-of-the-art, reaching $59.7$ directly and $63.4$ with WMReward + BoN, above the strongest listed V2V baseline with WMReward+BoN.

While automated evaluation metrics are valuable for measuring quality, prompt alignment, and physical plausibility, even the strongest VLM-based metrics miss certain artifacts and failure cases. Human evaluation complements automated metrics on two fronts: coverage of long-tail failures that automated protocols systematically miss, and discriminability across a wider scoring range. For instance, on the same prompt set, the evaluated T2V generators span $\sim$10 points on human evaluation versus $\sim$4 points on the comparable automated PAIBench-G overall score (see Appendix~\ref{sec::cosmos_hue_bench}). We therefore complement our automated results with two human evaluation protocols: Cosmos HUE, which targets broad Physical AI video generation, and Human World Bench (HWB), which focuses specifically on realistic human motion under task-level instructions.

\paragraph{Cosmos HUE.} Cosmos HUE (HUman Evaluation) is a human scoring protocol grounded in the same PAIBench-G prompt set (\cref{subsubsec::paibench_g}). HUE departs from prior
human-eval protocols in two ways. First, it replaces subjective Likert-scale grading with \emph{atomic binary} verification: each video is decomposed into a set of single-fact \emph{Yes / No / Unclear} questions phrased so that ``Yes'' always denotes the desirable outcome, shifting the annotator's task from holistic judgment to objective fact-verification. Each (video, question) pair is independently rated by two annotators, with disagreements escalated to a quality-control reviewer.   Second, the per-prompt question set is \emph{automatically generated by a three-layer VLM pipeline}: a Domain Strategist classifies the prompt into one of seven Physical AI domains, a Scene Parser produces a structured scene manifest from the prompt (T2V) or from sampled frames of the paired ground-truth reference video (I2V), and an Auditor emits the final atomic questions. This ensures that questions track the actual prompt and reference scene rather than a static rubric. All three layers run on GPT~5.2. The VLM-generated question set is then supplemented by a human review pass in which reviewers inspect generated videos from top-ranked models and propose additional questions targeting failure modes the automated pipeline does not yet cover; the proposed questions are folded back into the question bank. The evaluation set consists of 100 fixed prompts sampled from PAIBench-G preserving its native domain distribution, with each model producing 5 random-seed generations per prompt for 500 videos per model; the per-prompt question set (up to 20 binary questions) is applied to all 5 generated videos. Each question is classified into one of four dimensions: ``Semantic Alignment'', ``Physical Laws'', ``Geometric Reasoning'', and ``Visual Integrity''. Per-dimension and per domain T2V and I2V leaderboards, the formal scoring scheme, and reliability estimates are detailed in Appendix~\ref{sec::cosmos_hue_bench}.

\input{tables/results/human_eval}


\paragraph{Human World Bench (HWB).} While Cosmos HUE evaluates broad Physical AI video generation, Human World Bench (HWB) focuses on a more targeted and challenging setting: egocentric image-to-video generation for human manipulation tasks under task-level instructions.
The benchmark videos in HWB are sourced from EgoVerse~\citep{punamiya2026egoverse} and comprise 180 samples.
HWB uses an absolute failure-mode protocol: annotators judge each generated video independently for instruction following and physical plausibility. \label{subsec::hwb}
We report two top-level pass rates.
\emph{Instruction following} measures whether the video depicts the requested actions and objects.
\emph{Physical plausibility} measures temporal coherence and physical plausibility, including object dynamics, contact, and hand anatomy.
The HWB score is the average of the instruction-following and physics pass rates.


\cref{tab::human_eval_results} reports Cosmos-HUE scores on both T2V and I2V, alongside Human World Bench (HWB) results. On HUE T2V, Cosmos3-Super is the best open-source model at 89.3, behind the closed-source Veo-3.1 (91.3) and Seedance-1.5-Pro (90.0). On HUE I2V, Cosmos3-Super is again the best open-source model and is essentially tied with the leading closed-source generator, trailing Veo-3.1 by only 0.1 points (89.6 vs. 89.7). Cosmos3-Nano is also competitive, finishing second among open-source models on I2V (88.5) and third on T2V(87.6).

On HWB, Cosmos3-Super achieves 71.9, the state-of-the-art score among all evaluated models in \cref{tab::human_eval_results}, outperforming the strongest listed closed-source baseline, Veo-3.1 (67.8), by 4.1 points and the strongest non-Cosmos open-source baseline, Wan2.2-A14B (60.7), by 11.2 points.
Cosmos3-Nano also performs strongly at 66.9, ranking second among open-source models and outperforming every non-Cosmos open-source baseline.
Together, the two Cosmos 3 variants take the top two open-source positions on HWB, demonstrating strong egocentric human-motion generation across both model scales.

\input{figures/results/sft_i2v/sft_i2v}
\paragraph{Artificial Analysis Image-to-Video Leaderboard.} To evaluate Cosmos 3 under real-world scenarios, we submitted our specialized I2V model, Cosmos3-Super-Image2Video, to the Artificial Analysis Image-to-Video leaderboard (No Audio) for crowdsourced public voting. The model ranked the top among all open-weight models, and achieved the \#22 position globally including proprietary models. In particular, the model is on par with proprietary offerings such as Veo 3.1 \citep{veo31} and Wan 2.5 \citep{wan2pt2}, demonstrating its exceptional temporal dynamics and visual fidelity.
