\subsection{More Related Works}
\label{ap:related_work}


\subsubsection{Agentic AI Systems.} 
In recent years, LLMs and AI Agents are becoming more powerful due to comprehensive pre-training \cite{DBLP:journals/corr/abs-2601-03267}, advanced backbone architectures \cite{DBLP:journals/corr/abs-2505-06708, DBLP:journals/corr/abs-2510-26692}, training algorithms \cite{DBLP:journals/corr/abs-2501-12948}, and extended modalities supports \cite{DBLP:journals/corr/abs-2603-03276, DBLP:journals/corr/abs-2505-07062, DBLP:journals/corr/abs-2509-17765}. As a result, LLM and AI Agent applications are revolutionizing a variety of industries \cite{DBLP:journals/corr/abs-2503-21460}. More recently, we identify three trends which are related to our \textit{Eywa}.

\textbf{Agentic Reasoning.}
Real-world tasks are often too complex to be solved by AI agents in a single forward pass after receiving an instruction \cite{DBLP:journals/corr/abs-2602-07338}. Instead, agentic systems typically approach such tasks through structured reasoning procedures. A common paradigm is to first formulate an explicit plan or decompose the task into manageable subgoals \cite{DBLP:conf/nips/0001ST00Z23,ningmc}, and then solve these subgoals through iterative execution, reflection, clarification, and continuation across multiple interaction turns \cite{DBLP:journals/corr/abs-2509-23537, DBLP:journals/corr/abs-2601-11868, DBLP:journals/corr/abs-2603-00873}. This line of work highlights the importance of moving beyond one-shot language generation toward more deliberate and interactive reasoning processes.
Scientific tasks exhibit similar, and often more pronounced, complexity: they may require reasoning over specialized data modalities, domain-specific constraints, long-horizon dependencies, and expert-level predictive models \cite{jumper2021highly, merchant2023scaling, DBLP:journals/corr/abs-2507-01903, DBLP:journals/corr/abs-2508-14111, DBLP:journals/corr/abs-2511-02864}. However, most existing agentic reasoning frameworks remain primarily centered on the language capabilities of AI agents. Our \textit{Eywa} is motivated by this gap. Rather than relying solely on language agents to reason over scientific problems, \textit{Eywa} introduces domain-specific foundation models into agentic systems, enabling specialized scientific models to contribute domain representations, predictions, and feedback throughout the reasoning process.


\textbf{Orchestrated Intelligence.}
Building on the capabilities of individual agents, orchestrated intelligence has emerged as a central paradigm for tackling complex tasks that exceed the reach of any single component in the system. Early frameworks such as AutoGen~\cite{DBLP:journals/corr/abs-2308-08155} and MetaGPT~\cite{DBLP:conf/iclr/HongZCZCWZWYLZR24} demonstrate that conversational collaboration among role-playing agents can effectively decompose and solve intricate problems. Beyond flat conversational patterns, recent efforts have begun to explicitly investigate the orchestration layer itself. AgentOrchestra~\cite{zhang2025agentorchestra} adopts a hierarchical architecture in which a planning agent dispatches specialist sub-agents through a tool--environment--agent protocol, while evolving-orchestration approaches~\cite{DBLP:journals/corr/abs-2505-19591} train a centralized orchestrator via reinforcement learning to adaptively sequence and prioritize agents according to the evolving task state. Difficulty-aware orchestration~\cite{DBLP:journals/corr/abs-2509-11079} further tailors multi-agent workflows to query-level complexity through learned routing.
Despite this progress, existing orchestration frameworks remain largely language-centric: they primarily route, sequence, or coordinate agents that communicate through natural language, while non-linguistic expert models are often treated as passive callable tools. This design is insufficient for scientific domains, where reasoning may depend on heterogeneous representations such as molecular structures, crystal graphs, time series, spatial fields, and tabular measurements. Our \textit{EywaOrchestra} addresses this limitation by dynamically orchestrating heterogeneous experts, including LLM agents and domain-specific foundation models. By jointly selecting agent configurations, model specializations, and communication topologies, \textit{EywaOrchestra} extends orchestrated intelligence from language-agent workflow optimization toward heterogeneous scientific model collaboration.


\subsubsection{Modeling Structured Data with LLMs.}

Structured data modeling has long been a central problem in machine learning, spanning data types such as graphs, tables, time series, knowledge graphs, and other relational or compositional objects. Since the rise of deep learning, a wide range of neural architectures have been developed to capture the inductive biases of different structured domains, including graph neural networks, temporal models, tabular learning methods, and representation learning techniques for relational data \cite{wu2020comprehensive, lim2021time, borisov2022deep, antelmi2023survey, han2022data, fu2023investigating}. More recently, the rapid development of large language models and AI agents has opened new possibilities for leveraging language-based reasoning, instruction following, and tool use to support structured data modeling.

\textbf{LLMs directly handle structured data.}
One line of work investigates whether LLMs can directly process structured data by converting non-textual structures into textual or sequence-based representations \cite{DBLP:conf/sigir/Tang00SSCY024, DBLP:journals/corr/abs-2506-11040, DBLP:journals/corr/abs-2601-23204, DBLP:journals/corr/abs-2407-12522, DBLP:conf/emnlp/JiangZDYZW23}. These approaches benefit from the general reasoning and in-context learning capabilities of LLMs, enabling them to perform tasks such as question answering, classification, forecasting, and relational reasoning over structured inputs. However, direct textualization may lose important structural information, especially when the data contains long-range dependencies, high-order relations, precise numerical values, or domain-specific semantics that are difficult to faithfully express in natural language \cite{DBLP:conf/wsdm/SuiZZH024, DBLP:journals/corr/abs-2507-13646, DBLP:journals/bmcbi/SadeghiBFLN24, DBLP:journals/corr/abs-2510-01538, DBLP:conf/emnlp/YangTXZLH25}.

\textbf{LLMs to help existing models.}
Another line of work uses LLMs as assistants to improve existing structured data models rather than replacing them \cite{DBLP:journals/tkde/JinLHJJH24, DBLP:journals/sigkdd/ChenMLJWWWYFLT23, DBLP:journals/corr/abs-2502-08942}. For example, textual metadata or descriptions produced by LLMs can be combined with graph, tabular, or temporal encoders to enrich structured representations \cite{DBLP:journals/corr/abs-2510-21131, DBLP:conf/nips/YanLLY0ZYZHSDZ023, DBLP:conf/acl/0006ZJFJBH025, DBLP:journals/corr/abs-2509-00687, DBLP:journals/corr/abs-2506-00009}. LLMs can also act as controllers that select models, design prompts or queries, invoke external tools, and interpret the outputs of specialized predictors \cite{DBLP:conf/icml/TriratJH25, DBLP:journals/tmlr/TornedeDEGMRSTT24, DBLP:conf/mm/LuoFN024}. These methods preserve the strengths of domain-specific architectures while using LLMs to provide semantic knowledge and flexible reasoning.

\textbf{Re-programming LLMs for structured data.}
A third direction adapts LLMs themselves to structured data through re-programming, prompting, or lightweight adaptation. Instead of training a structured-data model from scratch, these methods map structured inputs into the embedding or token space of a pretrained language model, allowing the LLM to be reused for tasks beyond natural language \cite{DBLP:conf/iclr/0005WMCZSCLLPW24, DBLP:journals/corr/abs-2402-05862, DBLP:journals/corr/abs-2502-13449, DBLP:journals/corr/abs-2409-03444}. 
% Representative strategies include prefix or prompt tuning, modality projection, tokenization of numerical or relational objects, and alignment between structured representations and language embeddings. 
This paradigm leverages the broad prior knowledge and scaling properties of LLMs, but it also faces challenges in preserving structural fidelity, handling large-scale inputs, and ensuring that the language model's inductive biases align with the target scientific domain.

Despite these advances, most existing approaches still center on adapting structured data to the language-model interface. In contrast, many scientific problems are naturally handled by specialized foundation models that operate directly over structured representations. Our work therefore takes a complementary perspective: rather than forcing all structured data into language, \textit{Eywa} introduces domain-specific foundation models into agentic systems, allowing LLM agents to coordinate, invoke, and reason with specialized models that retain their native structured-data inductive biases.


\subsection{Future Directions}

\textbf{Scaling heterogeneous scientific model ecosystems.}
A natural future direction is to scale heterogeneous agentic systems beyond a small set of manually integrated domain-specific foundation models. Enabling heterogeneous components to collaborate within a unified agentic system remains challenging from the engineering perspective, because different models may operate over incompatible input formats and provide outputs with varying degrees of interpretability. As the number of available experts grows, the system must also determine which models are relevant to a task, whether their predictions are reliable, and how conflicting evidence from multiple experts should be reconciled.
Building such an ecosystem requires more than simply adding more APIs. It calls for standardized model interfaces, rich metadata descriptions, and capability profiling. Addressing these challenges would move heterogeneous agentic systems from small-scale proof-of-concept integrations toward scalable scientific model ecosystems.

\textbf{Learning better orchestration policies.}
In this work, orchestration is primarily driven by task-level reasoning over available agents, foundation models, and communication topologies. A promising direction is to learn orchestration policies from interaction data, execution traces, and task outcomes. Future systems may train orchestrators to predict which expert models to invoke, when to invoke them, and which communication topology is most suitable for a given task.

\textbf{More advanced integration between LLMs and scientific foundation models.}
The stronger "Tsaheylu" is, the better. Future work could explore tighter coupling mechanisms beyond model context protocol, including shared representation spaces, differentiable interfaces, bidirectional alignment between language and scientific embeddings, and memory/skill mechanisms that preserve domain-specific evidence across reasoning steps.


\textbf{Extending EywaBench for heterogeneous scientific reasoning.}
Another important direction is to further extend \textit{EywaBench} into a broader benchmark suite for heterogeneous scientific reasoning. While \textit{EywaBench} provides an initial testbed, scientific reasoning spans a much wider range of domains, data modalities, task formats, and expert models. Future extensions could incorporate more datasets as well as additional data types.. Beyond expanding coverage, \textit{EywaBench} can also be extended to evaluate richer capabilities such as model selection, cross-domain evidence integration, communication efficiency, and adaptive topology selection. Such extensions would make \textit{EywaBench} a more comprehensive testbed for understanding when heterogeneous orchestration is genuinely beneficial over language-only agents, single-domain models, or static multi-agent workflows.


\subsection{Limitation}

\textbf{Dependence on foundation model capabilities.}
\textit{Eywa} builds on existing language models and domain-specific foundation models, and therefore its performance can be influenced by the reasoning ability of the underlying LLMs, the predictive quality of scientific experts, and the reliability of their interfaces. As these base models continue to improve, we expect \textit{Eywa} to benefit from stronger general-purpose and domain-specific models.

\textbf{Coverage of domains and expert models.}
Although \textit{EywaBench} covers a diverse set of scientific domains and tasks, it cannot exhaustively represent the full space of scientific applications. Some domains may involve specialized data formats, assumptions, or expert models that are not included in the current evaluation. Extending \textit{EywaBench} to broader datasets and scientific workflows remains an important future direction.

\textbf{Computational cost.}
Compared with single-agent baselines, heterogeneous collaboration may introduce additional computation, latency, and communication overhead. This cost can be mitigated through more efficient orchestration, selective expert invocation, and adaptive stopping strategies, which we leave for future work.