\section{Introduction}
\label{sec:introduction}

World models are becoming a key interface for embodied simulation and interactive environments~\citep{world_models,genie3,gaia1,dreamdojo,aether,hy_worldplay,lingbot_world,infinite_world}.
We study camera-controlled world modeling: given a first frame, text, and a 6-DoF camera trajectory, the model synthesizes a one-minute 720p video that follows the input motion while preserving scene identity.
Recent open-source systems achieve minute-scale, action-conditioned rollouts~\citep{infinite_world,matrixgame3,lingbot_world,hy_worldplay}, but typically require large models, large-scale data, long training schedules, and multi-GPU inference.
A tempting lower-cost alternative is to distill long-rollout models from short-video generators, but such short-horizon teachers provide limited supervision for minute-scale scene persistence and trajectory following.
We therefore ask: \textit{can we natively train a high-fidelity, camera-controllable, one-minute world model while keeping data, training, and inference costs accessible?}

\noindent We introduce \modelname, a 2.6B-parameter open-source video world model designed around efficiency as a first-class objective.
\modelname is trained for one-minute generation using only ${\sim}213$K public video clips with metric-scale pose supervision and 15 days on 64 H100 GPUs.
At inference time, it supports three single-GPU variants: a bidirectional generator for high-quality offline synthesis, a chunk-causal autoregressive generator for sequential rollout, and a few-step distilled autoregressive generator for faster deployment.
Fig.~\ref{fig:teaser} shows representative generations, and Fig.~\ref{fig:pipeline_overview} summarizes the training and inference pipeline.
The improvements mainly lie in four key components.

\noindent\textbf{Efficient Native One-Minute Backbone.}
One-minute 720p generation stresses both token count and long-context modeling, so \modelname pairs a high-compression LTX2 tokenizer~\citep{ltx2} with a hybrid Linear DiT backbone.
The backbone combines frame-wise Gated DeltaNet~\citep{gated_deltanet} blocks for efficient recurrent context aggregation with periodic softmax attention for exact long-range recall.
This design keeps minute-scale context affordable while preserving the modeling capacity needed for scene persistence and camera-conditioned motion.

\noindent\textbf{Dual-Branch Camera Control.}
Precise action-conditioned world modeling requires generated videos to faithfully follow continuous action trajectories, rather than merely aligning with text prompts.
\modelname therefore uses a dual-rate camera conditioning design: a latent-rate UCPE branch~\citep{ucpe} captures global trajectory structure, while a raw-frame Pl\"ucker mixing branch restores fine camera motion inside each temporal VAE stride.
This lets the model preserve control accuracy despite aggressive video compression.

\noindent\textbf{Two-Stage Visual Refinement.}
To further improve visual quality, \modelname adopts a two-stage generation pipeline with a dedicated refinement stage.
We adapt an independent refiner
to operate on long \modelname outputs, correcting structural artifacts and sharpening details across the full minute.
This refinement stage is used as a quality-improvement pass after stage-1 generation.

\noindent\textbf{Robust Data Annotation and Evaluation Benchmark.}
To train camera-controlled videos without proprietary action labels, we build a robust annotation pipeline that recovers accurate metric-scale camera poses from public videos using pose and geometry estimators~\citep{vipe,pi3x,moge2}.
After filtering, this pipeline yields ${\sim}213$K video clips with precise metric-scale pose annotation.
Since existing benchmarks do not target minute-scale world modeling, we build a one-minute benchmark for action following, visual quality, and efficiency.
It contains 80 initial scenes generated by Nano Banana Pro~\citep{google2025nanobananapro} across four scene types, each paired with two revisit trajectories.

\noindent Experiments show that \modelname achieves higher accuracy in action-following than prior open-source baselines, with comparable visual quality, while delivering up to $36\times$ higher generation throughput.
Most importantly for accessibility, it reduces minute-scale generation to a single-GPU inference setting: the bidirectional and chunk-causal variants fit within one H100, and our distilled variant brings 1-minute video generation to 34s on a single RTX 5090 with NVFP4 quantization.

\noindent In summary, our contributions are: (i) a natively one-minute-trained, 720p, action-controllable world model with accessible training and inference cost; (ii) an efficiency-oriented architecture combining high-compression video latents, hybrid GDN/softmax long-context modeling, and dual-branch camera control; (iii) a long-video second-stage refiner that improves stage-1 visual quality; and (iv) a robust data annotation and evaluation pipeline for long-horizon world modeling.
















