\section{Method}
\label{sec:method}

\input{figures_tex/model}

\modelname is a world model designed to generate minute-long high-resolution videos with precise camera control under strict efficiency constraints. 
Scaling to this regime introduces three key challenges: (i) the prohibitive compute and memory cost of modeling 720p sequences; (ii) accurate, high-frequency action conditioning on continuous 6-DoF camera trajectories; and (iii) degraded visual quality when a base generator is trained under limited data and compute. 
To address these, \modelname builds on SANA-Video~\citep{chen2025sana} with three complementary designs: a hybrid GDN/softmax attention architecture for efficient long-context modeling, dual-rate camera conditioning for coarse-to-fine trajectory control, and a second-stage refiner for minute-length video to improve fidelity.


\subsection{Progressive Training Strategy}
\label{sec:training_strategy}

We train progressively from short clips to minute-scale videos, increasing sequence length and introducing architectural components in four stages:

\noindent\textbf{Stage 1: Efficient VAE Adaptation.}
To make minute-scale video modeling computationally feasible, we replace the baseline VAE~\citep{wan, dc-video-gen} with LTX2-VAE~\citep{ltx2} to leverage its superior spatiotemporal compression ratio. 
Given changed channel dimensions, we discard the original patchify layer and final output projection, and re-initialize them from scratch to align with the LTX2 space. 
Full-model fine-tuning adapts the model to this new latent distribution in 50k steps. 
The representation is $2.0\times$ smaller than ST-DC-AE and $8.0\times$ smaller than Wan2.1-VAE, improving training and inference efficiency.


\noindent\textbf{Stage 2: Hybrid Architecture Adaptation.} 
To improve the efficiency--quality trade-off of the backbone, we adapt the pre-trained SANA-Video model to the Hybrid GDN-Softmax architecture (Sec.~\ref{sec:long_context_modeling}). This architecture change is first optimized on short video clips, where training is cheaper and failure modes are easier to diagnose, before scaling to longer sequences.

\noindent\textbf{Stage 3: Minute-Scale Extension and Action Conditioning.} 
After stabilizing the architecture, we extend the sequence length to minute-scale videos to enable long-horizon temporal modeling. In parallel, we incorporate Dual-Branch Camera Control (Sec.~\ref{sec:camera_control}) to support metric 6-DoF trajectory conditioning, allowing explicit control over camera motion.

\noindent\textbf{Stage 4: Chunk-Causal Fine-Tuning and Few-Step Distillation.}
Starting from the one-minute bidirectional camera-control model, we further fine-tune a chunk-causal variant for autoregressive rollout.
We then use self-forcing distillation~\citep{huang2024selfforcing} to reduce sampling to four denoising steps.
For deployment, we add attention-sink tokens and local temporal windows to the softmax attention layers, keeping softmax memory and per-chunk latency constant with respect to rollout length.

\subsection{Memory-Efficient Long-Context Modeling}
\label{sec:long_context_modeling}

As background, SANA-Video~\citep{chen2025sana} uses ReLU-based cumulative linear attention in place of causal softmax attention.
For latent frame $t\ge0$ with $S=H_\ell W_\ell$ spatial tokens, let $\mathbf{Q}_t,\mathbf{K}_t,\mathbf{V}_t\in\mathbb{R}^{D\times S}$ collect the per-head queries, keys, and values, and let $\phi(\cdot)=\mathrm{ReLU}(\cdot)$.
Unlike softmax attention, which forms pairwise weights from $\mathbf{Q}_t^\top\mathbf{K}_{\le t}$, cumulative linear attention accumulates key--value outer products before applying the current queries:
{\small
\begin{equation}
\label{eq:cumulative_linear_attention}
\widetilde{\mathbf{O}}^{\mathrm{LA}}_t
=
\underbrace{\left(\sum_{\tau=0}^{t}\mathbf{V}_\tau\phi(\mathbf{K}_\tau)^\top\right)}_{\mathbf{A}^{\mathrm{LA}}_t}
\phi(\mathbf{Q}_t)
=
\left(\mathbf{A}^{\mathrm{LA}}_{t-1}+\mathbf{V}_t\phi(\mathbf{K}_t)^\top\right)\phi(\mathbf{Q}_t).
\end{equation}
}
Eq.~\ref{eq:cumulative_linear_attention} shows only the unnormalized numerator for brevity; the standard linear-attention denominator is omitted.
With $\mathbf{A}^{\mathrm{LA}}_{-1}=\mathbf{0}$, the cumulative numerator state $\mathbf{A}^{\mathrm{LA}}_t\in\mathbb{R}^{D\times D}$ is updated once per latent frame after aggregating all spatial-token outer products, so memory stays constant.

\noindent \textbf{Limitations of Cumulative Linear Attention.}
This compact state has no explicit decay or saliency mechanism: stale features accumulate with the same effective weight as more recent ones. At the minute scale, the unbounded growing state causes drift and degrades training stability.

\noindent \textbf{From Token-wise GDN to Frame-wise GDN.}
Gated DeltaNet (GDN)~\citep{gated_deltanet} augments the same recurrent state with a decay gate and a delta-rule correction:
\begin{equation}
\label{eq:token_gdn}
\mathbf{S}_i
=\gamma_i\mathbf{S}_{i-1}
+\big(\mathbf{v}_i-\gamma_i\mathbf{S}_{i-1}\hat{\mathbf{k}}_i\big)\beta_i\hat{\mathbf{k}}_i^\top,
\qquad
\mathbf{o}_i=\mathbf{S}_i\hat{\mathbf{q}}_i .
\end{equation}
Here $\mathbf{S}_i\in\mathbb{R}^{D\times D}$ is the token-wise recurrent state, $\hat{\mathbf{q}}_i,\hat{\mathbf{k}}_i,\mathbf{v}_i\in\mathbb{R}^{D}$ are the normalized query, normalized key, and value vectors, $\beta_i\in[0,1]$ is an update gate, and $\gamma_i\in(0,1]$ is a decay gate.
The correction term updates only the residual between the target value and the current state prediction, while $\gamma_i$ forgets obsolete content.
Standard GDN scans one token per recurrent step.
Our video model instead scans one latent frame per step.
For frame $t$, let $\hat{\mathbf{Q}}_t,\hat{\mathbf{K}}_t,\mathbf{V}_t\in\mathbb{R}^{D\times S}$ collect the query features, key features, and values used by frame-wise GDN; our additional key scaling is defined in the stabilization step below.
Let $\gamma_t\in(0,1]$ be a frame-level decay, and let $\boldsymbol{\beta}_t=\mathrm{diag}(\beta_{t,1},\ldots,\beta_{t,S})$ be per-token update gates.
The frame-wise state update becomes
\begin{equation}
\label{eq:frame_gdn}
\mathbf{S}_t
=\mathbf{S}_{t-1}\mathbf{M}_t+\mathbf{U}_t,\qquad
\mathbf{M}_t=\gamma_t\!\left(\mathbf{I}-\hat{\mathbf{K}}_t\boldsymbol{\beta}_t\hat{\mathbf{K}}_t^\top\right),\qquad
\mathbf{U}_t=\mathbf{V}_t\boldsymbol{\beta}_t\hat{\mathbf{K}}_t^\top,\qquad
\mathbf{O}_t=\mathbf{S}_t\hat{\mathbf{Q}}_t .
\end{equation}
Here $\mathbf{S}_t,\mathbf{M}_t,\mathbf{U}_t\in\mathbb{R}^{D\times D}$ are frame recurrent state, transition matrix, and additive update, respectively, and $\mathbf{O}_t\in\mathbb{R}^{D\times S}$ contains the output tokens for frame $t$.
Thus the recurrent state remains $D\times D$, while one recurrent step consumes all $S$ spatial tokens from a latent frame.

\noindent \textbf{Algebraic Stabilization for Spatial Explosion.}
Since $\mathbf{S}_t$ is repeatedly multiplied by $\mathbf{M}_t$, the transition should be non-expansive.
Let $\bar{\mathbf{K}}_t=[\bar{\mathbf{k}}_{t,s}]_{s=1}^{S}=\mathrm{ReLU}(\mathrm{RMSNorm}(\mathbf{K}_t))$ and $\mathbf{A}_t=\bar{\mathbf{K}}_t\boldsymbol{\beta}_t\bar{\mathbf{K}}_t^\top$.
The unscaled key energy is
\begin{equation}
\label{eq:gdn_key_mass}
\operatorname{tr}(\mathbf{A}_t)
=\sum_{s=1}^{S}\beta_{t,s}\|\bar{\mathbf{k}}_{t,s}\|_2^2 ,
\end{equation}
Since $\mathbf{A}_t$ is positive semidefinite, $\lambda_{\max}(\mathbf{A}_t)\le\operatorname{tr}(\mathbf{A}_t)$; an $O(S)$ trace can make $\mathbf{I}-\mathbf{A}_t$ expansive.
We therefore scale only the keys:
\begin{equation}
\label{eq:k_scale}
\hat{\mathbf{K}}_t = \bar{\mathbf{K}}_t \cdot \frac{1}{\sqrt{D \cdot S}} .
\end{equation}
With RMS-normalized keys and $\beta_{t,s}\in[0,1]$, $\operatorname{tr}(\hat{\mathbf{K}}_t\boldsymbol{\beta}_t\hat{\mathbf{K}}_t^\top)\le1$, hence $\|\mathbf{M}_t\|_2\le\gamma_t\le1$; $1/\sqrt{D}$ matches token-wise GDN L2 key normalization, and the extra $1/\sqrt{S}$ averages over spatial tokens.

\noindent \textbf{Bidirectional and Chunk-Causal GDN Variants.}
We use the same recurrence bidirectionally by summing forward and reversed-time scans. For chunk-causal inference, we partition latent frames into chunks, keep the forward scan global, and reset the reversed scan at chunk boundaries, giving each chunk local future context without leakage.

\noindent \textbf{Hybrid GDN/Softmax Attention.}
To enhance long-video generation performance, we further fine-tune the GDN model by replacing every fourth block with standard softmax attention~\citep{flash_attention2}, while retaining the original QKV and output projections.


\subsection{Dual-Branch Camera Control}
\label{sec:camera_control}

We use dual-rate geometric conditioning: latent-frame UCPE~\citep{ucpe} captures global 6-DoF pose, while raw-frame Pl\"ucker mixing compensates motion inside each VAE stride.

\noindent \textbf{Coarse Branch: Ray-Local UCPE.}
For each latent token at frame $t$ and spatial cell $s$, let $\mathbf{T}^{\mathrm{c2w}}_t=[\mathbf{R}_t\,|\,\mathbf{o}_t]$ be the camera-to-world pose and let $\mathbf{A}_t$ be the camera intrinsic matrix.
We unproject the corresponding pixel with $\mathbf{A}_t$ and transform it by $\mathbf{T}^{\mathrm{c2w}}_t$, obtaining a world-space ray with camera center $\mathbf{o}_{t}\in\mathbb{R}^{3}$ and unit direction $\mathbf{d}_{t,s}\in\mathbb{R}^{3}$.
We build a ray-local basis $\mathbf{z}=\operatorname{norm}(\mathbf{d}_{t,s})$, $\mathbf{x}=\operatorname{norm}(\mathbf{u}_{t}\times\mathbf{z})$, and $\mathbf{y}=\mathbf{z}\times\mathbf{x}$, where $\mathbf{u}_{t}$ is the camera vertical axis and $\operatorname{norm}$ denotes $L_2$ normalization.
This defines a homogeneous ray transform $\mathbf{D}_{t,s}\in\mathbb{R}^{4\times4}$ from world coordinates to the ray-local frame.
Following UCPE~\citep{ucpe}, we split each camera-branch attention-head vector into geometric channels and standard RoPE channels.
For token $i=(t,s)$, define $\mathbf{D}_i=\mathbf{D}_{t,s}$ and let $\mathrm{RoPE}_i$ be the standard spatiotemporal rotary operator for token $i$.
We apply the ray-local transform to the geometric channels and RoPE to the remaining channels:
\begin{equation}
\begin{aligned}
\widetilde{\mathbf{Q}}^{c}_{i} &=
  \big(\mathbf{D}_{i}^{\top}\oplus \mathrm{RoPE}_{i}\big)\mathbf{Q}^{c}_{i},\\
(\widetilde{\mathbf{K}}^{c}_{i},\widetilde{\mathbf{V}}^{c}_{i}) &=
  \big(\mathbf{D}_{i}^{-1}\oplus \mathrm{RoPE}_{i}\big)(\mathbf{K}^{c}_{i},\mathbf{V}^{c}_{i}),\\
\mathbf{O}^{c}_{i} &=
  \big(\mathbf{D}_{i}\oplus \mathrm{RoPE}_{i}^{-1}\big)
  \mathrm{GDN}_{\mathrm{cam}}(\widetilde{\mathbf{Q}}^{c},\widetilde{\mathbf{K}}^{c},\widetilde{\mathbf{V}}^{c})_{i},
\end{aligned}
\label{eq:ucpe_camera_branch}
\end{equation}
where superscript $c$ denotes the camera branch, $\mathbf{Q}^c_i,\mathbf{K}^c_i,\mathbf{V}^c_i\in\mathbb{R}^{D}$ are per-head camera-branch query/key/value vectors, and $\widetilde{\cdot}$ denotes the pose-transformed representation.
The operator $\oplus$ denotes a block-diagonal composition over the UCPE and RoPE channel groups; $\mathbf{D}_i$ is applied blockwise to 4-D homogeneous coordinate groups within the UCPE channels.
The camera branch uses its own QKV projections but shares the frame-wise GDN gates with the main branch; its zero-initialized projection is added to the main attention output.

\noindent \textbf{Fine Branch: Raw-Frame Pl\"ucker Mixing.}
The coarse UCPE branch operates at the latent-frame rate, whereas each latent token summarizes eight raw frames and their distinct camera poses.
For raw frame $r$ and pixel $p$, let $\mathbf{T}^{\mathrm{c2w}}_r=[\mathbf{R}_r\,|\,\mathbf{o}_r]$ and $\mathbf{A}_r$ denote the raw-frame camera pose and intrinsics, with camera center $\mathbf{o}_{r}\in\mathbb{R}^{3}$ and unit ray direction $\mathbf{d}_{r,p}\in\mathbb{R}^{3}$.
We compute pixel-wise Pl\"ucker raymaps $\boldsymbol{\rho}_{r,p}=(\mathbf{d}_{r,p},\mathbf{o}_{r}\times\mathbf{d}_{r,p})\in\mathbb{R}^{6}$ from $\mathbf{T}^{\mathrm{c2w}}_r$ and $\mathbf{A}_r$.
For each latent frame, we pack the eight raw-frame raymaps within one VAE temporal stride into a $48$-channel tensor and pass it through a zero-initialized 3D patch embedder.
A zero-initialized per-block projection then adds this embedding immediately after each self-attention output, preserving the pretrained model at initialization.


\subsection{Second Stage Refiner}
\label{sec:refiner}
Following LTX-Video~\citep{ltx_video}, we add a second-stage refiner to improve stage-1 \modelname visual fidelity. 
The refiner is trained on paired latents $(\mathbf{x}_{\ell},\mathbf{x}_{h})$, where $\mathbf{x}_{\ell}$ is a stage-1 or degraded latent and $\mathbf{x}_{h}$ is the high-fidelity target. 
We use truncated-$\sigma$ flow matching: $\mathbf{x}_{\ell}$ is perturbed with a large starting noise ($\sigma_{\mathrm{start}}=0.9$), and the model learns to map this noisy source toward $\mathbf{x}_{h}$, encouraging refinement over full reconstruction. 
We condition on text, camera, and a reference image, concatenate the reference to the sequence, and exclude it from the loss to preserve stage-1 appearance. 
For minute-long refinement, we initialize from the 17B LTX-2 model and train LoRA adapters on paired synthetic data plus real videos processed by downsampling and upsampling; details are in App.~\ref{sec:appendix_refiner}.
