\section{Experiments}
\label{sec:experiments}


\subsection{Implementation Details}
\label{sec:impl_details}

\noindent\textbf{Backbone configuration.}
\modelname utilizes a Diffusion Transformer (DiT) architecture with 20 transformer blocks ($d_\text{model}{=}2240$, 20 heads). To balance global dependency modeling and efficiency, we interleave 15 frame-wise GDN blocks with softmax attention blocks at layers $\{3,7,11,15,19\}$. Every block incorporates dual UCPE + Pl\"ucker mixing camera conditioning, and the LTX2 VAE uses $C{=}128$ latent channels for high spatio-temporal compression~\citep{dit,gated_deltanet,ucpe,plucker_lfn,ltx2}.

\noindent\textbf{Training scale.}
Our final model is trained on the 213K metric-pose clips described in Sec.~\ref{sec:data_pipeline}. 
The one-minute stage is trained on 961-frame clips with the standard flow-matching objective~\citep{flow_matching}. 
For 961-frame (60s) sequences, Context-Parallel (CP) training shards the latent sequence along time. 
For efficiency, we use custom fused Triton kernels for GDN scan and gate operations. 
The model is trained on 64 H100 GPUs for approximately 15 days; architectural details, hyperparameters, and the training schedule are in App.~\ref{sec:impl_details_appendix} and App.~\ref{sec:appendix_training_hyperparams}.



\subsection{Evaluation Setup}
\label{sec:eval_setup}



\noindent
\begin{minipage}[t]{0.50\textwidth}
\vspace{0pt}
\textbf{Benchmark construction.}
For evaluation, we build a 60-second world-model benchmark with 80 initial images generated by Nano Banana Pro~\citep{google2025nanobananapro} across four scene categories: game, indoor, outdoor-city, and outdoor-nature (20 per category).
Each image is paired with Simple and Hard revisit trajectories (Fig.~\ref{fig:benchmark_trajectories}); construction details and examples are in App.~\ref{sec:appendix_benchmark_construction} and App.~\ref{sec:appendix_benchmark_examples}.

\textbf{Protocol and metrics.}
The main table uses each model's multi-step, undistilled AR setting.
We report VBench~\citep{vbench} scores for visual quality and Umeyama~\citep{umeyama}-aligned Pi3X~\citep{pi3x} pose accuracy for action following; revisit memory (PSNR/SSIM/LPIPS~\citep{lpips}) and long-term degradation $\Delta$IQ are reported in App.~\ref{sec:appendix_revisit_temporal}, with metric details in App.~\ref{sec:appendix_eval_protocol}.
\end{minipage}%
\hfill%
\input{figures_tex/bench_trajectory}

\vspace{-0em}


\subsection{Main Results}
\label{sec:main_results}

\input{tables/main_table}

\noindent\textbf{Camera control accuracy.}
Tab.~\ref{tab:vbench} compares against recent world-model baselines~\citep{infinite_world,lingbot_world,hy_worldplay,matrixgame3} and shows that \modelname gives the strongest action following on both trajectory splits.
The refined output obtains the best RotErr, TransErr, and CamMC, with RotErr $4.50^\circ$/$8.34^\circ$ and CamMC $1.41$/$1.44$, improving over both large 480p and 720p baselines.

\noindent\textbf{Visual quality.}
With the second-stage refiner, \modelname attains $80.62$/$81.89$ VBench Overall on the Simple-/Hard-Trajectory splits, close to LingBot-World ($81.82$/$81.89$) while generating 720p videos on one GPU per clip.

\noindent\textbf{Inference efficiency.}
\modelname fits in $51.1$\,GB and reaches $24.1$ videos/hour; with the refiner, the full pipeline remains within the 80\,GB H100 budget ($74.7$\,GB) and reaches $22.0$ videos/hour, still $3.7\times$ faster than the fastest visible 480p baseline.
Although LingBot-World supports 720p, minute-long 720p inference is unaffordable under our 8-H100 evaluation budget; the closest-efficiency baseline, Infinite-World, runs at 480p.

\noindent\textbf{Revisit memory and temporal stability.}
After refinement, \modelname reaches $14.46$/$14.80$\,dB revisit PSNR, ranking second on Simple and first on Hard among the visible methods.
The refiner also reduces long-horizon visual drift: $\Delta$IQ drops from $3.79$/$3.09$ for stage-1 AR outputs to $1.17$/$0.31$, avoiding the severe late-window degradation observed for HY-WorldPlay ($23.59$/$25.88$); full per-method numbers are in App.~\ref{sec:appendix_revisit_temporal}.

\input{figures_tex/vis-main}

\noindent\modelname (Fig.~\ref{fig:vis-main}) preserves scene identity and viewpoint-consistent structure under hard actions, whereas baselines often blur, change layout, or collapse; additional examples are in App.~\ref{sec:appendix_qualitative}.



\subsection{Ablation Studies}
\label{sec:ablations}


\noindent\textbf{Deployment efficiency path.} %
Fig.~\ref{fig:efficiency-analysis}~(a) isolates the internal \modelname speed path under the 60-second 720p protocol, from the 60-step H100 autoregressive generator to 4-step distillation, attention-sink deployment, and NVFP4 quantization. Note that the attention-sink variant means that we use the first latent frame as the attention sink with local window attention on softmax attention layers only, so that the memory consumption remains constant.

\noindent\textbf{Progressive-training ablation.}
Tab.~\ref{tab:ablation_stages} isolates progressive upgrades at a 5\,s horizon on VBench-I2V~\citep{vbench} with identical data and inference settings. LTX2~\citep{ltx2} is quality-neutral ($+0.0012$ Total) but cuts peak memory from $8.9$ to $5.4$\,GB and latency $3.4{\times}$; the hybrid 15-GDN/5-softmax backbone raises Total to $0.853$ while keeping memory at $5.7$\,GB.

\input{tables/ablation_vbench}

\input{}

\noindent\textbf{GDN key scaling.}
We evaluate training stability under identical conditions: 81-frame sequences and an all-GDN~\citep{gated_deltanet} architecture initialized from a shared LTX2-VAE~\citep{ltx2} cumulative-linear checkpoint.
As shown in Fig.~\ref{fig:gdn-key-scaling}, our $1/\sqrt{D S}$ scaling is the only variant that ensures stable convergence.
In contrast, both the $L_2$ ($1/\sqrt{D}$) and no-scale baselines suffer from immediate gradient instability, triggering NaN events at steps 16 and 1, respectively.



\input{figures_tex/train-stability-camera-condition}

\noindent\textbf{Camera conditioning.}
Tab.~\ref{tab:ablation_camera} ablates camera conditioning on our held-out OmniWorld~\citep{omniworld} validation split with 5s clips after 10k fine-tuning steps. Input-level Pl\"ucker~\citep{plucker_lfn} gives only small gains, attention-level PRoPE~\citep{li2025cameras} and UCPE~\citep{ucpe} greatly improve control, and the dual UCPE + Pl\"ucker mixing branch gives the lowest Pi3X~\citep{pi3x} errors with competitive FVD~\citep{unterthiner2018towards}.

\input{figures_tex/efficiency-latency-gpu-combine}

\noindent\textbf{Efficiency scaling.}
Fig.~\ref{fig:efficiency-analysis}~(b) reports AR recurrent single-step H100 latency and memory with matched resolution, camera conditioning, and FFN settings. The recurrent/linear variants keep compact clean-context states, while all-softmax grows its KV cache and runs out of memory at 60s.




