\section{Quantitative Evaluation}

\subsection{Evaluation Setup}

We evaluate LightMem-Ego on three daily memory scenarios:
object finding, conversation recall, and life summarization. 
For each scenario, we construct queries with manually annotated gold
evidence. 
We report retrieval performance with Recall@1/3/5 and
MRR, QA performance with LLM- and human-judged accuracy, and
interactive latency with P50 / P90 retrieval, generation, and
end-to-end QA time on phone and glasses-style clients.

\input{tables/recall}
\subsection{Memory Retrieval Accuracy}
Table~\ref{tab:retrieval_accuracy} summarizes retrieval accuracy in the three demonstration scenarios. 
LightMem-Ego retrieves relevant memory evidence within the top-3 results for most queries, achieving 74.1 overall R@3 and 0.627 MRR. 
The results suggest that the hierarchical memory store provides sufficiently accurate evidence selection for downstream experience QA in everyday object finding, conversation recall, and life summarization settings.

\input{tables/accuracy}
\subsection{Experience Question Answering Accuracy}
Table~\ref{tab:qa_accuracy} reports QA accuracy against annotated
experience evidence. 
LightMem-Ego obtains 51.9 LLM-judged accuracy and 55.6 human-judged accuracy overall, with the best performance on life summarization.
These results indicate that, when relevant evidence is retrieved, the system can produce usable memory-grounded answers for the demonstrated everyday QA scenarios, while fine-grained object finding and conversation recall remain more challenging in the current prototype.

\input{tables/latency}
\subsection{Latency on Mobile and Wearable Clients}
Table~\ref{tab:latency} reports the runtime of LightMem-Ego on
phone and glasses-style clients. 
For short-term memory QA, the system achieves interactive response times, with P50 end-to-end latency of 5.86s on phone and 7.01s on the glasses-style client.
Long-term memory QA requires additional retrieval and evidence
aggregation, leading to higher P50 latency of 14.87s and 19.96s,
respectively. 
This latency profile is consistent with the intended usage of the demo: recent-memory queries support near-interactive assistance, while long-term queries remain practical for retrospective recall and life-summary interactions.

\input{tables/capability_comparison}

\subsection{Capability Comparison with Existing Assistants and Memory Systems}
We further compare LightMem-Ego with representative commercial assistants, text-based memory systems, and recent egocentric multimodal assistants in terms of publicly described capabilities. 
As shown in Table~\ref{tab:capability_comparison}, existing systems cover different parts of the design space: ChatGPT Memory and Mem0-style systems mainly focus on conversational or text-based agent memory, while Memories.ai explores long-term visual memory for searchable video archives~\cite{GPT-4o, GPT-5, Mem0, MemoriesAI}. Gemini Live and Ray-Ban Meta AI Glasses support real-time multimodal interaction, while recent egocentric assistants such as Vinci, VisualClaw, VisionClaw, Egocentric Co-Pilot, and EgoButler explore real-time perception, wearable assistance, tool use, or long-context egocentric QA~\cite{Vinci, VisualClaw, VisionClaw, Egocentric_Co-Pilot, EgoLife}.
However, these systems generally do not expose an explicit hierarchy that jointly supports current multimodal memory, short-term event memory, long-term multimodal episodic memory, semantic routine memory, and timestamped evidence retrieval.

In contrast, LightMem-Ego is designed specifically for everyday-life memory assistance over streaming visual-audio experience.
Rather than emphasizing only in-the-moment perception, text-based personalization, or agentic task execution, it organizes user experience into current, short-term, and long-term memory stores, enabling both immediate assistance and retrospective experience QA such as object finding, conversation recall, life summarization, and routine discovery.
