In this work, we present \bench, a dataset of 235 questions that remain challenging for modern models but straightforward for humans. We introduce a fine-grained taxonomy, curated reference solutions, and an automatic grading pipeline to evaluate diverse VLMs, LLMs, and image-generation models. Our analyses show that closed-source frontier models outperform open-weight models, while open-weight models are often more cost-effective. We further find that scaling model size or enabling tool use does not consistently improve performance. Finally, taxonomy-level analysis reveals several distinctive patterns in model failure modes.
% In this work, we present \bench, a comprehensive dataset presenting tasks that remain challenging for AI models but  intuitive for humans. Rather than relying on aggregate metrics, we introduce a fine-grained taxonomy and curated reference solution for automatic evaluation. Using a developed grading pipeline, we conduct different analyses over a diverse set of VLMs, LLMs, and image generation models to diagnose failure modes across different sub-tasks. Our analyses show that closed-source frontier models outperform open-weights counterparts. However, the latter tend to be the more cost-effective choice. The results also indicate that simply scaling model size or allowing for tool usage does not consistently improve the performance. Finally, taxonomy evaluation reveals that model performance varies across each sub-task, with models exhibiting expertise in certain areas and while lacking performance in others.


\textbf{Limitations.}
While \bench provides useful insights into model failure modes, several limitations remain. First, the dataset is relatively modest in size and imbalanced across sub-tasks. Although sufficient for our analyses, increasing the number of samples per sub-task would strengthen the statistical reliability of the results. Second, because the questions were primarily created by students to challenge two frontier models, they may be biased toward those models' specific weaknesses and may miss other types of reasoning failures. Finally, although \bench questions are intended to be easy for humans, adding a human baseline would more directly quantify the gap between model and human performance.
% While \bench presents key insights into failure reasoning, there are several limitations that should be considered. First, the dataset size is relatively modest and exhibits an imbalance in the number of questions across different sub-tasks.  Although the current size was enough to conduct our analysis, expanding the number of samples per sub-task would strengthen the statistical results observed. Second, since the dataset was created against two main frontier models, it could include bias. By focusing on these specific model's weaknesses, \bench might overlook other types of reasoning failures. Finally, although the questions in \bench are designed to be easy for humans, including a human baseline would  further highlight the reasoning gap between models and human performance.

\begin{ack}
This project was first proposed in the EPFL course Reasoning in Artificial Intelligence (MATH-700), taught by Emmanuel Abbé. We thank all students in the course for each proposing five questions to challenge frontier models at the time. We also thank Kevin Qiu, Amene Gafsi, Rafaila Galanopoulou, Szymon Sobczak, Julien David Laurendeau, and Zhipeng Xue for their early efforts to dataset cleaning. Finally, we thank the RCP team at EPFL for GPU and model support, and the Laboratory of Mathematical Data Science, which ran the course, for covering the associated costs.
\end{ack}