We evaluate a diverse set of state-of-the-art LLMs, VLMs and image generation models, covering a range of sizes and capabilities. The following model families are included: \texttt{Qwen3} and \texttt{Qwen3.5} \citep{qwen2025qwen3, qwen2026qwen35omni}, \texttt{GLM 4.7, 5, 5.1, 5.2} \citep{zeng2026glm5}, \texttt{Kimi K2.5} and \texttt{2.6} \citep{kimi2026k25}, \texttt{DeepSeek V4} \citep{deepseek2026v4}, \texttt{Gemma-4} \cite{google2026gemma4}, and \texttt{GPT-OSS} \citep{agarwal2025gptoss}. For frontier models, we consider Google's \texttt{Gemini-2.5, 3, 3.1} \citep{comanici2025gemini25, google2025gemini3, google2026gemini31} and OpenAI \texttt{GPT-5, 5.2, 5.4, 5.5} \citep{openai2025gpt5, openai2026gpt54, openai2026gpt55}. When possible, we include multiple sizes of the same model to analyze the impact of scaling model size. VLMs are tested on both multi-to-text and text-only questions, while LLMs are evaluated on text-only questions. The evaluation is repeated $k=4$ times for tasks with textual output, and only once for image generation tasks, due to their significantly higher cost.

All models are evaluated with \texttt{thinking} mode enabled, and \texttt{medium} effort when these options are available, with the maximum number of output tokens set to 32{,}768. Unless stated otherwise, no model has access to tools. Generation requests are retried at most five times in case of failure, and are counted as incorrect if the model fails to produce any response. All open-weight LLMs and VLMs were accessed through EPFL Research Computing Platform (RCP) AI-Inference-as-a-Service (AIaaS) which deploys and operates AI models, using using vLLM \citep{kwon2023efficient} as backbone and exposing OpenAI-comapatible APIs.

For image generation tasks we evaluate different versions of the \texttt{GPT-Image} and \texttt{Gemini-Image} models, which are specifically designed for generative image tasks. Due to technical constraints, no open-weight image generation models are included.

A detailed list of the models along with their configuration is provided in \autoref{appendix:models_info}, where the reported open models pricing refers to the offering by RCP AIaaS.