\section{Reproducibility Statement}\label{appendix:reproducibility}

This appendix consolidates the artefacts and protocol required to reproduce every cell of \S\ref{sec:experiments}.

\subsection{Code release}

\paragraph{Harness adapters.} Five claw-adapter packages -- \texttt{openclaw\_swebench}, \texttt{hermes\_swebench}, \texttt{zeroclaw\_swebench}, \texttt{nanobot\_swebench}, and the \texttt{generic} baseline adapter -- are released. Each ships \verb|run_infer.py| and \verb|run_eval.py| together with the per-harness orchestrator, agent adapter, and workspace modules. The same registry also hosts the minimal \emph{bare adapter} used for the diagnostic in \S\ref{sec:adapter_diagnostic}.

\paragraph{Lite construction scripts.} A Node.js toolkit implements the cost-aware rank-aware Lite selection, K-sweep, sensitivity checks, quartile-stratification logic, and final-report generators.

\paragraph{Figure-generation scripts.} \texttt{matplotlib} scripts that regenerate the cost--accuracy Pareto figure, the future-commit-cleanup comparison, and the Lite parity panels from the released result workbooks are in the \texttt{scripts/} directory of the release (\texttt{generate\_pareto\_figure.py}, \texttt{generate\_leak\_fix\_figure.py}, \texttt{generate\_lite\_figures.py}).

\subsection{Data release}

\textbf{Claw-SWE-Bench} (350 instance IDs plus metadata) and \textbf{Claw-SWE-Bench-Lite} (80 instance IDs with cost-aware selection metadata) are released as JSON files. The underlying issues and repositories are hosted on the upstream SWE-bench-Multilingual and SWE-bench-Verified-Mini sources \cite{swe_smith, swebench_verified_mini}; we redistribute only the curated ID sets and metadata.

\subsection{Reproduction protocol}

\paragraph{Per-instance run.}
\begin{verbatim}
python3 run_infer.py \
    --harness <openclaw|hermes|zeroclaw|nanobot|generic> \
    --dataset multilingual \
    --model <provider/model_name> \
    --run_id <run_label> \
    --timeout 3600 \
    --workers 3
\end{verbatim}

\paragraph{Evaluation.}
\begin{verbatim}
python3 run_eval.py \
    --predictions artifacts/<run_id>/predictions.jsonl \
    --dataset_name SWE-bench/SWE-bench_Multilingual \
    --run_id <run_id> \
    --max_workers 8
\end{verbatim}

The CLI surface is identical across all five claws; per-harness adapter flags are documented in Appendix~\ref{appendix:harness}.

\subsection{Random seeds and runs}

A single run per (instance, harness, model) cell is executed with 3-thread concurrency. The 1-repeat choice is a cost-driven trade-off; multi-seed validation on a 50-instance slice is left to future work.

\subsection{Compute requirements}

Full hardware and cost figures are in Appendix~\ref{appendix:compute}. In summary: a single 16-core / 61\,GiB-RAM Linux server is sufficient, with no GPU required since all model inference is routed through external APIs.
