% ============================================
% §2 Related Work
% ============================================
\section{Related Work}
\label{sec:related}


We build on four recent research threads---prompt-history agents,
externalized memory, skill libraries, and long-horizon game
testbeds---and target their joint problem: which slice of prior
experience enters each decision, and whether that route can itself
be ablated.

\paragraph{Loop and context engineering.}
In practice, agent design is increasingly framed as \emph{loop
engineering}: specifying a goal, tools, termination, and---centrally---the
memory and context policy of a control loop that runs over hundreds of
steps and multiple sessions~\citep{anthropic-context-2025}. The recurring
failure mode is context growth: appending full transcripts and tool logs
to every call overflows the window and dilutes attention, whereas
``token-poor'' loops keep only the last few messages or short summaries
and therefore depend on an explicit memory store~\citep{memory-survey-2025}.
This is exactly the axis our contract isolates. That literature is largely
qualitative and centered on coding agents; the academic threads below
supply the mechanisms, and we make the memory stage of the loop a typed,
bounded, ablatable contract on a hard long-horizon game.

\paragraph{Long-horizon LLM agents and prompt-visible history.}
ReAct and Reflexion made it natural for observations and
self-critiques to reappear in later LLM
calls \citep{react2023,reflexion2023}. Recent horizon analyses
\citep{sinha-horizon-2026,wang-planning-2026,horizon-mirage-2026,%
kwa-metr-2025} examine how small errors compound over many steps.

\paragraph{Typed and structured memory.}
MemGPT, Mem0, MemoryOS, GAM, hierarchical procedural memory, and
Agent Workflow Memory \citep{memgpt2024,mem0-2025,memoryos2025,%
gam-2026,macla-2025,awm-2025} move information out of raw message
history into external stores. Adjacent typed-memory frameworks
\citep{coala-2024,generative-agents-2023,a-mem-2025} group memory
by capacity; we instead role-type slots by mutability and retrieval
source. Most evaluation is in dialogue or QA; in our setting
retrieval feeds an action policy in a stochastic environment.

\paragraph{Self-evolving skill libraries.}
Voyager pioneered an external library of agent-written skills
\citep{voyager2023}; SkillsBench, SkillOS, Memento-Skills,
SAGE/SkillRL, SkillWeaver, ExpEL, and DyStIL extend the design
\citep{skillsbench2026,skillos2026,memento-skills-2026,%
sage-skillrl-2025,skillweaver2025,expel2024,dystil2025}. In the
SoK notation $S=(C,\pi,T,R)$ \citep{sok-agentic-skills-2026}, our
$L_5$ guides correspond to $(C,\pi)$: a trigger selects a prose
policy for the next decision.

\paragraph{LLM agents on games.}
Games such as Crafter, NetHack, BALROG, LMGame-Bench, DSGBench,
Gameverse, and RAGEN \citep{crafter2022,nethack2020,balrog2024,lmgame-bench-2025,%
dsgbench-2025,gameverse2026,ragen2025} provide stochastic testbeds
for agents. Card-game work includes end-to-end policy networks
\citep{xiao-hearthstone-2023}, LoRA-tuned draft models
\citep{urzagpt-2025}, cross-card-game LLM evaluation
\citep{wang-cardgames-2025}, and LLM play on the original Slay the
Spire with simplified rule sets
\citep{bateni-whitehead-2024,rethinking-agent-design-2025}.

\paragraph{Public \stspad{} LLM agents.}
The public \stspad{} ecosystem---STS2MCP, HermesBridge, AI-Spire,
CharTyr \citep{sts2mcp-2026,hermesbridge-2026,ai-spire-2026,%
chartyr-sts2-2026}---does not report a matched ablation over
prompt strictness, episodic memory, and triggered skills.
AGI-Eval~\cite{agi-eval-spire-2026} lists zero $A_0$ victories across
five frontier-model rows with mixed denominators.

\begin{table*}[t]
\centering\small
\caption{Positioning map for prior-experience interfaces. Green checkmarks mark a central axis; orange triangles mark a partial or optional axis. Citations appear in the surrounding paragraphs.}
\label{tab:memcompare}
\setlength{\tabcolsep}{6pt}
\renewcommand{\arraystretch}{1.08}
\begin{tabularx}{\textwidth}{@{}p{0.22\textwidth}cccccY@{}}
\toprule
Family & Transcript & Typed memory & Skills & Layer ablation & Game policy & Gap \\
\midrule
Prompt-history / replay & \yesmark & \nomark & \partmark & \nomark & \partmark & attribution \\
Structured memory       & \nomark & \yesmark & \nomark & \partmark & \nomark & policy \\
Skill-library agents    & \partmark & \partmark & \yesmark & \partmark & \partmark & contract \\
\textbf{Ours}            & \textbf{\nomark} & \textbf{\yesmark} & \textbf{\yesmark} & \textbf{\yesmark} & \textbf{\yesmark} & \textbf{joint test} \\
\bottomrule
\end{tabularx}
\end{table*}

We combine these threads in one bounded-memory contract, so future
work can compare alternative contracts under the same harness.
