% ============================================
% §3 STS2 Testbed (edited — non-game reader pass)
% ============================================
\section{The \stspad{} Testbed}
\label{sec:testbed}

\stspad{} is a turn-based deck-building roguelike: an agent builds
a deck during a run, fights stochastic battles, chooses routes and
rewards, and climbs an ordinal difficulty ladder. The game is
useful for evaluating LLM-agent memory because it is long-horizon
but not visually opaque. Rules, cards, relics, enemies, events,
legal actions, and state transitions can be represented as text
records, while success still requires sustained planning across
many contingent decisions. We release the resulting runs as a
reusable evaluation resource (\S\ref{sec:methodology:repro}).

\subsection{Four properties of \stspad{} as an LLM testbed}
\label{sec:testbed:props}

\textbf{(\emph{P1}) Closed, enumerable, LLM-readable rule space.}
Public database snapshots index hundreds of typed records (576
cards, 293 relics, 115 monsters, 87 encounters, and 66 events in
Spire Codex's May~2026 API; these are database counts, not unique
experimental-patch counts) \citep{spirecodex-2026}. Unlike
pixel-rendered Crafter \citep{crafter2022} or the inherited code
complexity of NetHack \citep{nethack2020}, \stspad{} has a compact
rule space that can be loaded into a typed knowledge layer
(\S\ref{sec:arch}). This makes $L_3$ part of the evaluation
substrate rather than a hidden source of game knowledge.

\textbf{(\emph{P2}) Empirically long horizon.} A typical run lasts
a median $\sim$80~min wall-clock (IQR 37--109\,min) and contains
67 LLM strategic calls (IQR 27--105). Roughly $500$ additional
per-run decisions, such as combat targets, treasure choices, map
nodes, and hand selection, are mechanically resolved or routed to a
fast tier with bounded combat context (\S\ref{sec:arch}). This is
precisely the regime where message-history accumulation becomes
costly relative to typed recomposition
\citep{sinha-horizon-2026,horizon-mirage-2026,kwa-metr-2025}.

\textbf{(\emph{P3}) Multi-axis stochasticity.} Random card draw,
shuffle order, reward offerings, map paths, relic effects, elite
and event placements, and Ascension modifiers prevent simple
trajectory replay. A strong policy must generalize over states, not
memorize a fixed route.

\textbf{(\emph{P4}) State-conditioned combat math.} Damage and
status pipelines combine hand contents, enemy intent, block timing,
and effects such as \emph{vulnerable, weak, strength}, and
\emph{dexterity}. The agent must compute from the current state;
web-like recall is much less useful than state-conditioned
calculation.

\subsection{Ascension ladder and scoring}
\label{sec:testbed:difficulty}

The 11-level Ascension ladder ($A_0$--$A_{10}$, with $A_{10}$ the
maximum) gives the testbed an ordinal difficulty scale. Higher
Ascensions stack modifiers that change strategic priorities, so
climbing the ladder is not just repeating $A_0$ with larger
numbers. We therefore use two complementary evidence streams: a
fixed-$A_0$ matrix that isolates components at one difficulty
(\S\ref{sec:methodology}) and an auto-mode ladder in which the
agent advances after victories and retries after defeats
(\S\ref{sec:methodology}).

Runs are scored with a derived analysis score:
\begin{equation}
s =
\begin{cases}
100 & \text{if victory,}\\
\mathrm{floor} + \tfrac{52}{3}\cdot\mathrm{bosses} & \text{otherwise,}
\end{cases}
\label{eq:score}
\end{equation}
where $\mathrm{bosses}$ counts cleared act bosses (0/1/2 by reached
floor for non-victory runs, 3 for victories; full mapping in
Appendix~\ref{app:configs}). The value is recomputed from outcome,
floor, and boss-count fields, not copied from the raw archive
\texttt{score} field. The $52/3$ coefficient calibrates three
cleared bosses to $52$ points, the approximate mid-Act-3 floor
reach. A $\pm 10\%$ perturbation of the coefficient checks
score-based qualitative comparisons; win-rate claims do not depend
on this score scale.

\subsection{Data corpus, release, and harness}
\label{sec:testbed:data}

The released archive contains $298$ completed independent game
trajectories spanning fixed-$A_0$ ablations, cross-backbone probes,
and auto-mode ascension runs. Each trajectory records target and
reached Ascension, outcome, wall-clock duration, LLM-call counts,
condition tag, and the active memory/scaffold setting. The headline
fixed-$A_0$ comparison uses a pre-specified balanced subset of 50
completed games, namely the first ten per condition under the
frozen configuration. Other completed trajectories support the
cross-backbone and ladder diagnostics rather than entering the
fixed-$A_0$ estimate. Frozen $L_4{+}L_5$ stores and per-condition tags
are released with the public artifact archive
\citep{deal-checklist-2025,liu-versioning-2025}.

The game alone is not the full benchmark: a game interface must
be paired with a decision protocol. Public \stspad{}
implementations \citep{sts2mcp-2026,hermesbridge-2026,%
ai-spire-2026,chartyr-sts2-2026} let LLMs act in the game, while
the independently evaluated AGI-Eval configurations
\citep{agi-eval-spire-2026} report no listed victory. Our
architecture (\S\ref{sec:arch}) supplies the bounded-memory
contract that makes the released trajectories a reusable evaluation
surface.
