% ============================================
% Limitations
% ============================================
% No \section here — main.tex provides "\section*{Limitations}".

\paragraph{Sample size and statistical inference.}
The headline fixed-$A_0$ result uses a balanced 50-game subset,
which is at or above the per-condition sample size typical for
recent LLM-agent game benchmarks
\citep{voyager2023,balrog2024,lmgame-bench-2025}. The
cross-backbone and ladder streams are smaller diagnostic streams,
reported separately rather than pooled into the headline table.
Wilson intervals at this sample size are read at the
scaffolded-versus-unscaffolded level; finer-grained equivalence
tests among scaffolded variants and smooth backbone-transfer
curves are natural extensions that the released archive supports.

\paragraph{Same-codebase accumulating-context variant.}
The release is organized so that an accumulating-context cell
sharing our codebase, condition tags, scoring scripts, and frozen
stores can be added as one further row in the same matrix. This is
the cleanest direct comparison to the bounded contract, and the
artifact provides the run protocol, prompt records, and analysis
scripts needed to run it.

\paragraph{Single character and game-version coverage.}
The headline runs target one playable character, Silent, chosen to
keep the typed substrate ($L_3$ enumerable game knowledge,
$L_4{+}L_5$ stores) self-consistent in the present submission.
Released trajectories carry game-version tags so that
version-stratified re-analysis is possible from the archive, and
cross-character runs follow the same harness once $L_3/L_4/L_5$
are repopulated for the new character.

\paragraph{External player and ecosystem references.}
Mega Crit, Spiracle, and AGI-Eval enter the paper only as
difficulty and ecosystem context, with cached snapshots released
alongside the artifact. Matched human inference would require a
separately designed user study with controlled denominators, which
is outside the present resource.

\paragraph{Architectural scope.}
The evaluation is training-free and single-game. The bounded
contract is tuned for \emph{turn-based decision} settings like
\stspad{}; continuous or streaming control loops, visual input,
multi-agent play, online human correction, model-internal
fine-tuning, and cross-game transfer are deliberate non-targets of
this release. Stub templates and expert seed skills are
author-curated; Mode~B measures within-interface template filling,
which we report as one operating point rather than as fully
autonomous skill invention.
