% ============================================
% Ethical Considerations
% ============================================
% No \section here — main.tex provides "\section*{Ethical Considerations}".

\paragraph{Subject of study.}
The agent acts only inside a single-player video game on a community
mod. No human subjects are involved, and the project collects no
personal data. We cite aggregate public community statistics and
public repository metadata, but do not redistribute user-level
community runs.

\paragraph{Game and mod licensing.}
\stspad{} is a commercial title, and the surrounding mod ecosystem
includes open-source projects such as STS2MCP, AI-Spire,
HermesBridge, and CharTyr. The artifact records community-mod
ancestry and license notes, including the CharTyr STS2 mod
\citep{chartyr-sts2-2026}, without redistributing game asset dumps.
Game names, screenshots, and textual references throughout this paper
and the released artifact are used for research documentation only;
all rights to the underlying game and its content remain with the
rights holders, and verbatim game-string tables are regenerated locally
rather than redistributed.

\paragraph{Reproducibility and data sharing.}
We release run logs, frozen $L_4{+}L_5$ artifacts, configuration
snapshots, and analysis scripts in a separate public archive with
license and provenance notes (\S\ref{sec:methodology:repro}). The
archive contains derived fields and local scripts, not a
redistributable copy of the game. It excludes scraped human dialogue,
model weights, copyrighted raw art, and game asset dumps.

\paragraph{Use of AI assistance.}
LLM systems were used for drafting and editing assistance. The
authors verified the experimental results, statistical analyses,
architecture choices, and interpretations against frozen logs and
scripts. LLM systems were not used to generate or alter raw logs or
to fabricate numerical results.

\paragraph{Risk of dual use.}
Bounded retrieval and skill write-back are general agent capabilities.
The single-player game setting has low direct harm. Deployment in
domains with real-world actions would require separate safety review,
logging, and write-gate controls.
