\section{Cosmos-HumanEval Benchmark (Cosmos-HUE)}
\label{sec::cosmos_hue_bench}

\textbf{Cosmos-HumanEval} (Cosmos-HUE or HUE) is an evaluation benchmark for video generation introduced in the main results section. HUE provides a human-reference signal via atomic binary questions generated per video by a Vision-Language Model (VLM) pipeline and scored along the four HUE dimensions (Semantic Alignment, Physical Laws, Geometric Reasoning, Visual Integrity). This appendix complements that overview with the formal binary scoring scheme, the annotation protocol and reliability estimates, and the per-dimension and per-domain Text-to-Video (T2V) and Image-to-Video (I2V) leaderboards.

\paragraph{Binary response schema and scoring.} Each Layer~3 question elicits \textbf{Yes} (criterion met), \textbf{No} (clear violation), or \textbf{Unclear} (e.g., the region is
occluded or the motion is too fast to assess, or the video fails to depict the event the question presupposes, e.g., the prompt calls for a right turn but the lead vehicle never
turns, so the follow-up question about a smooth curved arc has no clear Yes or No). \emph{Unclear is treated as No}: a model is not rewarded for failing to produce the prompted action,
and a non-confident annotator is conservatively counted against the model. Concretely, Unclear does not contribute to the Yes numerator but remains in the denominator, so a video that
is mostly Unclear cannot score above one with the same number of explicit Yes answers. Let $\mathcal{Q}_v$ denote the set of questions issued for video $v$; the per-video
HUE score is
\begin{equation}
  \mathrm{HUE}(v) \;=\; \frac{\sum_{q \in \mathcal{Q}_v} \mathbf{1}\!\left[\mathrm{ans}(v,q) = \textsc{Yes}\right]}{\lvert \mathcal{Q}_v \rvert} \times 100\%,
  \label{eq:hue_video}
\end{equation}
with $\mathrm{HUE}(v) \in [0, 100]\%$ since ``Yes'' is always the desirable outcome. Model-level scores aggregate over $V$ test videos:
\begin{equation}
  \mathrm{HUE}(m) \;=\; \frac{\sum_{v=1}^{V} \sum_{q \in \mathcal{Q}_v} \mathbf{1}\!\left[\mathrm{ans}(v,q) = \textsc{Yes}\right]}{\sum_{v=1}^{V} \lvert \mathcal{Q}_v \rvert} \times
100\%.
  \label{eq:hue_model}
\end{equation}
The grand mean weights every answered (video, question) observation equally, avoiding inflation from videos that happen to receive fewer applicable questions. Dimension-level scores
are computed identically by restricting the sums to questions belonging to each dimension.

\paragraph{Annotation protocol.} Each video receives up to 16 atomic binary questions from the three-layer pipeline above, and each (video, question) pair is independently rated by \emph{two} human annotators. If the two annotators agree, the consensus answer is recorded as the canonical response; if they disagree, the question is escalated to a third \emph{quality-control (QC) reviewer}, whose answer is final. This double-rating + QC-tiebreaker workflow yields a single canonical answer per (video, question) pair, bounds the contribution of any single annotator, and produces the auditable response stream that the scoring formulas in Equations~\ref{eq:hue_video}--\ref{eq:hue_model} operate on.

\paragraph{Question design and iteration.} Authoring and refining atomic questions targets two goals: \textbf{low variance} (re-running the same model on the same prompts produces a similar score), and \textbf{GT saturation} (real videos paired with the prompts score at or near $100\%$). We use these criteria to evaluate candidate questions and to iterate the question bank, revising or dropping items that fail any criterion. The process is ongoing: the GT score remains below $100\%$ (\cref{tab:hue_t2v_leaderboard,tab:hue_i2v_leaderboard}), so question-bank refinement continues.

\paragraph{Reliability and confidence intervals.} The current T2V evaluation pool samples 100 prompts from PAIBench-G, with 5 random-seed generations per prompt and up to 20 questions
per video, yielding up to 10{,}000 binary observations per model checkpoint. Treating each as an independent Bernoulli trial with underlying Yes-rate $p = \mathrm{HUE}(m)/100$, the
95\% confidence interval on the model-level HUE score (in percentage points) is $\mathrm{HUE}(m) \pm 100 \cdot 1.96\sqrt{p(1-p)/N}$ with $N$ the observation count; in practice this
gives CI$_{95}$ widths around $\pm 0.6$ points for top-tier scores on the scale of \cref{tab:hue_t2v_leaderboard}. Atomic binary judgments reduce within-annotator variance compared to
Likert-scale grading, where the annotator must integrate multiple observations, weigh them, and map onto an arbitrary scale. Each of these three operations introduces noise. The
\emph{Real video GT} row scores 93.6 on T2V prompts and 94.4 on I2V prompts, reflecting (i) minor residual prompt--video mismatches in the sampled PAIBench-G pairs despite manual
review, and (ii) imperfections in the auto-generated question set, where overly strict, ambiguous, or off-target questions can yield No or Unclear on a real video. We treat the gap
from GT to $100\%$ as a north-star for ongoing question-bank refinement.

\paragraph{T2V leaderboard.} \cref{tab:hue_t2v_leaderboard} reports current T2V results across our model panel and the strongest external T2V baselines. Veo-3.1 leads overall (91.3) and Seedance-1.5-Pro is second (90.0); Cosmos3-Super is the best open-source generator at 89.3, ahead of Wan2.2-A14B (88.2) and Cosmos3-Nano (87.6). The per-dimension picture is more favorable to Cosmos3-Super: it is the best open-source model on 9 of 12 axes, and beats every generator, including closed-source, on AV (87.7) and Physics (91.5). The gap from the strongest generator to real video (\emph{Real video GT}, 93.6) is $\sim$2.3 points overall, leaving meaningful headroom for the next round.

\paragraph{I2V leaderboard.} \cref{tab:hue_i2v_leaderboard} reports the parallel I2V evaluation. Veo-3.1 leads overall (89.7), with Cosmos3-Super trailing by only $0.1$ points (89.6);
Cosmos3-Nano (88.6) edges Wan2.2-A14B (88.4) and Seedance-1.5-Pro (87.6). Cosmos3-Super beats every generator outright on Visual Integrity (94.2), Robotics (91.1), and Miscellaneous
(94.8), and ties Veo-3.1 for the lead on Semantic Alignment (both 90.3); Cosmos3-Nano wins AV (87.6) outright; Wan2.2-A14B wins Physics (91.9). The full I2V slate spans $\sim$9 points
across generators.

\definecolor{rowours}{HTML}{E8F4E0}      % light green for current-gen Cosmos 3 (ours)
\definecolor{rowourslight}{HTML}{F2F8EC} % lighter green for prior-gen Cosmos2.5 (ours)
\begin{table}[H]
  \centering
  \caption{\textbf{Cosmos HUE T2V leaderboard.} Per-dimension and per-domain breakdowns (\%; higher is better). Left block: Overall HUE plus the four dimensions (Semantic
Alignment, Physical Laws, Geometric Reasoning, Visual Integrity). Right block (after the rule): Overall HUE restricted to each PAIBench-G prompt domain (\textbf{C.S.}, \textbf{AV},
\textbf{Rob.}, \textbf{Ind.}, \textbf{Hum.}, \textbf{Phy.}, \textbf{Misc.}\ abbreviate Common Sense, Autonomous Vehicle, Robotics, Industry, Human, Physics, Miscellaneous).
\textcolor{nvidiagreen!60!black}{\textbf{Cosmos 3 generators (ours)}} are shaded green; \emph{Real video GT} is the upper reference. \textbf{Bold} marks the best in the column,
\underline{underline} the second best. \textbf{Cosmos3-Super} leads open-source on 9 of 12 axes and beats every generator, including closed-source, on AV (87.7) and Physics (91.5).}
  \label{tab:hue_t2v_leaderboard}
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \renewcommand{\arraystretch}{1.15}
  \begin{tabular}{l l c cccc | ccccccc}
      \toprule
      \textbf{Model} & \textbf{Type} & \textbf{Overall}
      & \makecell{\textbf{Sem.}\\\textbf{Align.}} & \makecell{\textbf{Phys.}\\\textbf{Laws}} & \makecell{\textbf{Geo.}\\\textbf{Reas.}} & \makecell{\textbf{Vis.}\\\textbf{Integ.}}
      & \textbf{C.S.} & \textbf{AV} & \textbf{Rob.} & \textbf{Ind.} & \textbf{Hum.} & \textbf{Phy.} & \textbf{Misc.} \\
      \midrule
      \rowcolor{black!8}
      \emph{Real video GT (PAI-Bench)} & --- & 93.6 & 95.0 & 90.5 & 94.1 & 94.8 & 92.0 & 93.7 & 94.9 & 94.4 & 93.2 & 93.1 & 93.6 \\
      \midrule
      \rowcolor{rowours}
      \textbf{Cosmos3-Super} & Open-sourced & 89.3 & \underline{92.4} & 85.4 & 86.6 & \underline{93.7} & 89.5 & \textbf{87.7} & 88.4 & 89.4 & 88.6 & \textbf{91.5} & 93.0 \\
      \rowcolor{rowours}
      \textbf{Cosmos3-Nano} & Open-sourced & 87.6 & 91.5 & 83.6 & 84.1 & 92.5 & 89.5 & \underline{87.0} & 86.6 & 85.8 & 86.2 & 87.9 & \underline{94.0} \\
      \midrule
      Wan2.2-A14B & Open-sourced & 88.2 & 92.1 & 83.4 & 85.6 & 92.6 & 87.8 & 84.9 & 83.7 & 91.1 & 89.9 & 87.5 & 93.4 \\
      HunyuanVideo-1.5 & Open-sourced & 86.5 & 88.4 & 81.8 & 83.6 & 93.6 & 88.3 & 81.0 & 81.3 & 88.0 & 87.9 & 87.4 & 93.3 \\
      Wan2.1-14B & Open-sourced & 84.0 & 87.7 & 77.1 & 78.5 & 92.7 & 85.3 & 79.1 & 78.2 & 87.0 & 85.1 & 85.8 & 90.7 \\
      Wan2.2-5B & Open-sourced & 80.8 & 86.9 & 73.8 & 73.6 & 89.9 & 83.4 & 73.5 & 76.0 & 84.4 & 81.3 & 82.8 & 87.9 \\
      Cosmos-Predict2.5-14B & Open-sourced & 82.1 & 88.1 & 74.6 & 75.8 & 90.7 & 81.3 & 84.4 & 76.8 & 82.4 & 81.3 & 84.5 & 90.2 \\
      Cosmos-Predict2.5-2B & Open-sourced & 81.8 & 88.1 & 74.0 & 75.1 & 90.6 & 81.5 & 81.8 & 79.5 & 82.7 & 80.3 & 82.7 & 89.8 \\
      \midrule
      Veo-3.1 & Closed-sourced & \textbf{91.3} & \textbf{94.3} & \underline{87.7} & \textbf{91.0} & \textbf{93.9} & \textbf{92.7} & 85.6 & \textbf{91.5} & \textbf{94.7} &
\underline{90.3} & 90.4 & \textbf{95.8} \\
      Seedance-1.5-Pro & Closed-sourced & \underline{90.0} & 91.2 & \textbf{88.4} & \underline{89.8} & 92.8 & \underline{90.5} & 83.6 & \underline{89.0} & \underline{92.9} &
\textbf{90.7} & \underline{91.4} & 91.7 \\
      \bottomrule
  \end{tabular}
\end{table}

\begin{table}[H]
  \centering
  \caption{\textbf{Cosmos HUE I2V leaderboard.} Per-dimension and per-domain breakdowns (\%; same protocol as \cref{tab:hue_t2v_leaderboard}, with image conditioning). Layout,
domain abbreviations, color shading, and \textbf{bold}/\underline{underline} conventions match \cref{tab:hue_t2v_leaderboard}.}
  \label{tab:hue_i2v_leaderboard}
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \renewcommand{\arraystretch}{1.15}
  \begin{tabular}{l l c cccc | ccccccc}
      \toprule
      \textbf{Model} & \textbf{Type} & \textbf{Overall}
      & \makecell{\textbf{Sem.}\\\textbf{Align.}} & \makecell{\textbf{Phys.}\\\textbf{Laws}} & \makecell{\textbf{Geo.}\\\textbf{Reas.}} & \makecell{\textbf{Vis.}\\\textbf{Integ.}}
      & \textbf{C.S.} & \textbf{AV} & \textbf{Rob.} & \textbf{Ind.} & \textbf{Hum.} & \textbf{Phy.} & \textbf{Misc.} \\
      \midrule
      \rowcolor{black!8}
      \emph{Real video GT (PAI-Bench)} & --- & 94.4 & 94.8 & 93.2 & 95.1 & 95.4 & 94.2 & 96.0 & 94.9 & 93.4 & 92.1 & 96.3 & 98.1 \\
      \midrule
      \rowcolor{rowours}
      \textbf{Cosmos3-Super} & Open-sourced & \underline{89.6} & \textbf{90.3} & \underline{87.5} & \underline{87.0} & \textbf{94.2} & \underline{90.7} & \underline{86.2} &
\textbf{91.1} & 89.6 & \underline{87.1} & \underline{91.5} & \textbf{94.8} \\
      \rowcolor{rowours}
      \textbf{Cosmos3-Nano} & Open-sourced & 88.6 & \underline{89.2} & 86.4 & 86.3 & \underline{93.5} & 90.6 & \textbf{87.6} & \underline{90.6} & 88.0 & 84.7 & 91.0 & 93.9 \\
      \midrule
      Wan2.2-A14B & Open-sourced & 88.4 & 88.6 & 86.0 & 85.7 & \underline{93.5} & 90.1 & 84.7 & 84.8 & \underline{92.0} & 86.5 & \textbf{91.9} & 92.5 \\
      HunyuanVideo-1.5 & Open-sourced & 85.6 & 87.1 & 80.8 & 83.0 & 92.5 & 90.0 & 81.7 & 77.2 & 89.6 & 85.2 & 89.5 & 91.8 \\
      Wan2.1-14B & Open-sourced & 83.9 & 85.0 & 80.2 & 80.6 & 91.2 & 88.1 & 76.8 & 74.8 & 89.9 & 84.3 & 85.8 & 93.2 \\
      Wan2.2-5B & Open-sourced & 80.4 & 83.4 & 74.6 & 75.7 & 89.5 & 86.1 & 74.3 & 68.6 & 88.4 & 79.6 & 84.8 & 90.5 \\
      Cosmos-Predict2.5-14B & Open-sourced & 83.0 & 85.0 & 77.9 & 77.5 & 92.4 & 88.1 & 85.2 & 78.4 & 83.0 & 78.5 & 84.8 & 93.2 \\
      Cosmos-Predict2.5-2B & Open-sourced & 82.6 & 86.1 & 77.3 & 78.2 & 90.2 & 85.6 & 86.0 & 79.0 & 85.6 & 77.1 & 85.1 & 92.4 \\
      \midrule
      Veo-3.1 & Closed-sourced & \textbf{89.7} & \textbf{90.3} & \textbf{87.7} & \textbf{89.2} & 93.2 & \textbf{92.3} & 86.0 & 88.6 & \textbf{93.4} & \textbf{87.2} & 91.2 &
\underline{94.3} \\
      Seedance-1.5-Pro & Closed-sourced & 87.6 & 87.7 & 85.4 & 86.5 & 91.8 & 89.5 & 82.9 & 85.7 & 90.6 & 85.7 & 90.6 & 93.0 \\
      \bottomrule
  \end{tabular}
\end{table}
