\section{Training and Implementation Details}
\label{app:training_and_implementation}

\begin{table*}[t]
\centering
\setlength{\tabcolsep}{4.5pt}
\renewcommand{\arraystretch}{1.1}

\resizebox{\textwidth}{!}{%
\begin{tabular}{lccccccccc}
\toprule
\multirow{2}{*}{\textbf{Method}}
& \multicolumn{3}{c}{\textbf{Dragonball}} & \multicolumn{3}{c}{\textbf{HotpotQA}} & \multicolumn{3}{c}{\textbf{SQuAD}} \\
\cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
  & \textit{IE$@1\uparrow$} & \textit{Prec.$@1\uparrow$} & \textit{Rec.$@1\uparrow$}  & \textit{IE$@1\uparrow$} & \textit{Prec.$@1\uparrow$} & \textit{Rec.$@1\uparrow$}  & \textit{IE$@1\uparrow$} & \textit{Prec.$@1\uparrow$} & \textit{Rec.$@1\uparrow$} \\
\midrule
RAPTOR
  & 2.74 & 36.40 & 7.53
  & 5.40 & 55.63 & 9.70
  & 5.40 & 29.77 & 18.13 \\
Meta-Chunking-MSP
  & 3.21 & 37.20 & 8.63
  & 8.42 & \underline{60.30} & 13.97
  & 12.24 & 38.97 & 31.40 \\
Meta-Chunking-PPL
  & 5.07 & 39.80 & 12.73
  & 10.65 & \textbf{61.23} & 17.40
  & 11.78 & 38.37 & 30.70 \\
DenseXRetrieval
  & 0.00 & 1.08 & 0.32
  & 1.19 & 39.17 & 3.03
  & 4.74 & 28.17 & 16.83 \\
SAKI-RAG
  & 15.31 & 68.37 & 22.40
  & 13.43 & 51.60 & 26.03
  & 65.15 & 85.80 & 75.93 \\
\hline
LLM
  & 17.87 & 73.53 & 24.30
  & 15.29 & 51.83 & 29.50
  & 70.70 & 88.63 & 79.77 \\
LLM + 10 Topics
  & \textbf{26.32} & \textbf{84.43} & \textbf{31.17}
  & \textbf{21.41} & 55.33 & \textbf{38.70}
  & \textbf{80.30} & \textbf{92.83} & \textbf{86.50} \\
\hline
\rowcolor{oursblue}
MCompassRAG + 10 Topics
  & \underline{23.46} & \underline{79.80} & \underline{29.40}
  & \underline{19.19} & 52.40 & \underline{36.63}
  & \underline{79.35} & \underline{92.37} & \underline{85.90} \\
\bottomrule
\end{tabular}
}% end resizebox

\vspace{2pt}

\resizebox{\textwidth}{!}{%
\begin{tabular}{lccccccccc}
\toprule
\multirow{2}{*}{\textbf{Method}}
& \multicolumn{3}{c}{\textbf{DRBench}} & \multicolumn{3}{c}{\textbf{LegalBench-RAG}} & \multicolumn{3}{c}{\textbf{SCI-DOCS}} \\
\cmidrule(lr){2-4}\cmidrule(lr){5-7}\cmidrule(lr){8-10}
  & \textit{IE$@1\uparrow$} & \textit{Prec.$@1\uparrow$} & \textit{Rec.$@1\uparrow$}  & \textit{IE$@1\uparrow$} & \textit{Prec.$@1\uparrow$} & \textit{Rec.$@1\uparrow$}  & \textit{IE$@1\uparrow$} & \textit{Prec.$@1\uparrow$} & \textit{Rec.$@1\uparrow$} \\
\midrule
RAPTOR
  & 1.38 & 29.27 & 4.70
  & 1.52 & 29.23 & 5.20
  & 62.51 & 80.27 & 77.87 \\
Meta-Chunking-MSP
  & 2.87 & 32.63 & 8.80
  & 2.67 & 33.10 & 8.07
  & 64.50 & 81.03 & 79.60 \\
Meta-Chunking-PPL
  & 4.32 & 34.07 & 12.67
  & 3.65 & 34.53 & 10.57
  & 0.16 & 15.10 & 1.07 \\
DenseXRetrieval
  & 0.42 & 21.87 & 1.93
  & 0.47 & 21.93 & 2.13
  & 55.45 & 76.83 & 72.17 \\
SAKI-RAG
  & 14.54 & 58.80 & 24.73
  & 7.04 & 43.30 & 16.27
  & 73.43 & 89.77 & 81.80 \\
\hline
LLM
  & 18.68 & 64.93 & 28.77
  & 9.07 & 47.40 & 19.13
  & 78.68 & 92.60 & 84.97 \\
LLM + 10 Topics
  & \textbf{31.71} & \textbf{79.67} & \textbf{39.80}
  & \textbf{15.08} & \textbf{56.47} & \textbf{26.70}
  & \textbf{89.19} & \textbf{99.10} & \textbf{90.00} \\
\hline
\rowcolor{oursblue}
MCompassRAG + 10 Topics
  & \underline{28.30} & \underline{75.07} & \underline{37.70}
  & \underline{12.97} & \underline{52.10} & \underline{24.90}
  & \underline{88.03} & \underline{98.25} & \underline{89.60} \\
\bottomrule
\end{tabular}
}% end resizebox
\caption{Retrieval performance at depth $k{=}1$ across six benchmarks (IE\,$@1\uparrow$, Precision\,$@1\uparrow$, Recall\,$@1\uparrow$). \textbf{Bold} = best; \underline{underline} = second-best. \colorbox{oursblue}{\strut\,\name{}\,} rows are shaded. LLM and LLM\,+\,10\,Topics are oracle upper bounds that use an LLM at retrieval time.}
\label{tab:results_at1}
\vspace{-1em}
\end{table*}

\paragraph{Training details.}
For each benchmark, \name{} is trained separately using its corresponding training split. When a benchmark does not provide a sufficiently large training set, we use 10\% of the available data for synthetic training data construction. For DRBench~\cite{abaskohi2026drbench} and LongBenchV2~\cite{bai-etal-2025-longbench}, which are smaller and do not provide suitable retrieval training labels, we train using EDR-200~\cite{prabhakar2025enterprisedeepresearchsteerable} and LongBenchV1~\cite{bai-etal-2024-longbench}, respectively. For each dataset, we sample 2{,}000 training chunks and generate 10 synthetic queries per chunk, resulting in 20{,}000 query--chunk pairs before negative sampling. We train the metadata selector, abstraction module, and MLP relevance classifier while keeping the student encoder, topic centroids, and cached chunk-topic distributions fixed. Unless otherwise specified, all hyperparameters follow our default setting: AdamW~\cite{loshchilov2018decoupled} with learning rate $2\times10^{-5}$, batch size 16, weight decay 0.01, dropout 0.1, and 3 training epochs. The distillation temperature is set to $\tau=1.0$, and the loss interpolation coefficient is set to $\alpha=0.5$. For generation, we use temperature $\tau=0.7$ and top-$p=0.9$; for teacher relevance scoring, we use temperature $\tau=0.0$ to obtain deterministic judgments.

\paragraph{Evaluation.}
Because the compared methods use different chunk granularities, evaluating all systems with a fixed number of retrieved chunks can be unfair: the same top-$k$ may correspond to very different amounts of retrieved text. We therefore use two complementary evaluation protocols. For retrieval quality, we report Recall, Precision, and Information Efficiency (IE), where $\mathrm{IE@k}=\mathrm{Precision@k}\times\mathrm{Recall@k}$. These metrics are computed at $k \in \{1,3,5\}$ and averaged over three runs. For downstream evaluation, we use task-appropriate generation metrics, including Accuracy, F1, ROUGE-L~\cite{lin-2004-rouge}, METEOR~\cite{banerjee-lavie-2005-meteor}, and BERTScore~\cite{Zhang2020BERTScore}, depending on the benchmark. To ensure fairness in downstream comparisons, retrieved chunks are added in ranked order until a fixed token budget is reached (1K), so each method provides the generator with the same maximum amount of evidence regardless of its chunk size. This protocol evaluates retrieval methods under comparable evidence budgets while still allowing each method to use its own native chunking strategy. We use $L=50$ and $M=10$ in our experiments.