\begin{abstract}
  Few-step video generation has been significantly advanced by consistency distillation.
  However, the performance of consistency-distilled models often degrades as more sampling steps are allocated at test time, limiting their effectiveness for any-step video diffusion.
  We argue that this limitation arises because consistency distillation replaces the original probability-flow ODE trajectory with a consistency-sampling trajectory, weakening the desirable test-time scaling behavior of ODE sampling.
  To address this limitation, we introduce \textbf{AnyFlow}, the first any-step video diffusion distillation framework based on flow maps.
  Instead of distilling a model for only a few fixed sampling steps, AnyFlow optimizes the full ODE sampling trajectory.
  To this end, we shift the distillation target from endpoint consistency mapping ($\mathbf{z}_t\!\to\!\mathbf{z}_0$) to flow-map transition learning ($\mathbf{z}_t\!\to\!\mathbf{z}_r$) over arbitrary time intervals.
  We further propose \textbf{Flow Map Backward Simulation}, which decomposes a full Euler rollout into shortcut flow-map transitions, enabling efficient on-policy distillation that reduces test-time errors (\ie discretization error in few-step sampling and exposure bias in causal generation).
  Extensive experiments across both bidirectional and causal architectures, at scales ranging from 1.3B to 14B parameters, demonstrate that AnyFlow achieves performance comparable to or better than consistency-based counterparts in the few-step regime, while supporting flexible and scalable sampling under varying step budgets. In the causal text-to-video setting, AnyFlow-FAR reaches a VBench score of 84.05 at 4 NFEs and further improves to 84.41 at 32 NFEs, surpassing Krea-Realtime-14B (83.25 at 4 NFEs). For image-to-video generation, AnyFlow-FAR achieves a VBench-I2V score of 87.87 at 4 NFEs, comparable to Wan2.1-I2V-14B using 50$\times$2 NFEs (87.71). In the 14B bidirectional text-to-video setting, AnyFlow reaches 84.04 at 4 NFEs, outperforming rCM-14B (83.73 at 4 NFEs).
  Code is released at \url{https://github.com/NVLabs/AnyFlow}.
  \keywords{Flow Map Distillation \and Video Diffusion Distillation \and On-Policy Distillation}
\end{abstract}
