\section{Conclusion}
We present AnyFlow, the first any-step video diffusion distillation framework based on a two-time flow map formulation. By learning transitions between arbitrary time pairs, AnyFlow supports a wide range of sampling budgets within a single model. Built on large pretrained video diffusion backbones, it combines an improved forward flow map training recipe with flow map backward simulation for on-policy distillation, thereby reducing discretization error and exposure bias during sampling. Across both bidirectional and causal architectures, and across model scales from 1.3B to 14B, AnyFlow consistently matches or outperforms strong consistency-based counterparts in the few-step regime while continuing to improve as the sampling budget increases. These results highlight a practical and scalable path toward high-quality any-step video generation.

\myPara{Limitations.} The primary limitation of our method is its reliance on external datasets for flow map training. Even when synthetic data are used, the training distribution may still differ from that of the base model, which can introduce mild distribution shift, such as smoother textures. This issue could be mitigated by applying AnyFlow with the same data used to pretrain the base model.

\myPara{Future Work.} Although AnyFlow establishes the first any-step, flow-map-based video diffusion distillation framework, there is still substantial room for improving the training methodology. One important direction is to develop more effective and stable forward flow map training strategies to improve robustness across different NFE regimes. In addition, because AnyFlow provides a scalable recipe for learning any-step causal video diffusion from both data and a teacher model, a natural next step is to extend the framework to autoregressive long-video generation with dedicated long-video training.
