\begingroup
\small
\setlength{\tabcolsep}{2.4pt}
\renewcommand{\arraystretch}{1.05}
\setlength{\LTleft}{0pt}
\setlength{\LTright}{0pt}
\setlength{\LTcapwidth}{\textwidth}
\begin{longtable}{>{\scriptsize\raggedright\arraybackslash}p{0.22\linewidth}
                  >{\scriptsize\raggedright\arraybackslash}p{0.25\linewidth}
                  >{\scriptsize\raggedright\arraybackslash}p{0.43\linewidth}}
\caption{Complete 30-entry profile bank used for long-term
personalization. Each row corresponds to one persona-intent profile entry. The
third column reports English summaries of stored \texttt{profile\_json} fields;
long values are shortened for readability.}
\label{tab:appendix_profile_library_catalog}\\
\toprule
Persona & Role-intent bucket & Stored profile fields (English summaries) \\
\midrule
\endfirsthead
\toprule
Persona & Role-intent bucket & Stored profile fields (English summaries) \\
\midrule
\endhead
\bottomrule
\endfoot
Postsecondary teacher & Course unit teaching &
\texttt{layout.slide\_structure}: cover $\rightarrow$ learning objectives
$\rightarrow$ core concept breakdown $\rightarrow$ method workflow
$\rightarrow$ classroom example $\rightarrow$ recap/Q\&A;
\texttt{content.notes}: organize course units around objectives, concepts,
methods, and examples. \\
Postsecondary teacher & Research method explanation &
\texttt{layout.slide\_structure}: two-column layout with explanation on the
left and architecture/local diagrams on the right;
\texttt{content.notes}: for graduate students, clarify problem definition,
method steps, key experimental settings, and method rationale. \\
Postsecondary teacher & Academic progress review &
\texttt{layout.slide\_structure}: background/goals---stage progress---key
results---open issues---next plan;
\texttt{content.notes}: include core parameterized formulas and emphasize
precise academic definitions. \\
\midrule
Software developer & Architecture walkthrough &
\texttt{layout.slide\_structure}: overview/architecture diagram
$\rightarrow$ module structure $\rightarrow$ data flow $\rightarrow$ design
trade-offs $\rightarrow$ experimental evidence;
\texttt{visual.chart\_type\_priority}: architecture/topology diagrams for
overview pages and flowcharts for data flow. \\
Software developer & Sprint risk update &
\texttt{layout.slide\_structure}: current goal $\rightarrow$ completed
capabilities $\rightarrow$ main risks $\rightarrow$ key dependencies
$\rightarrow$ next steps;
\texttt{visual.chart\_type\_priority}: status-risk comparison diagrams and
milestone timelines. \\
Software developer & Technical value brief &
\texttt{content.notes}: organize into application scenarios, core technical
capabilities, performance evidence, and engineering value;
\texttt{visual.chart\_type\_priority}: comparison tables and evidence charts. \\
\midrule
Management analyst & Organizational diagnosis brief &
\texttt{layout.slide\_structure}: current issues---root-cause breakdown---capability
gaps---optimization directions;
\texttt{visual.chart\_type\_priority}: issue trees, $2\times2$ priority
matrices, capability heatmaps, and fishbone diagrams. \\
Management analyst & Framework training session &
\texttt{layout.slide\_structure}: use structured summary pages such as
$2\times2$ matrices or card frameworks with action orientation;
\texttt{visual.chart\_type\_priority}: process diagrams, decision trees, and
priority matrices. \\
Management analyst & Recommendation follow-up review &
\texttt{layout.slide\_structure}: multi-column card layout and priority
matrix for follow-up actions;
\texttt{content.notes}: prioritize conclusions and logic chains over raw data
stacking. \\
\midrule
Marketing manager & Go-to-market pitch &
\texttt{layout.slide\_structure}: open with a hook, use numeric anchors in the
middle, and close with a call to action;
\texttt{content.notes}: prefer short, forceful titles and remove redundant
labels such as ``objective.'' \\
Marketing manager & Brand campaign showcase &
\texttt{layout.slide\_structure}: put the core narrative on the cover; use big
numbers and conclusion cards on inner slides while reducing tables;
\texttt{content.notes}: avoid meta-labels on the cover. \\
Marketing manager & Campaign performance review &
\texttt{layout.slide\_structure}: highlight key metrics with KPI cards and
reduce table/screenshot clutter;
\texttt{visual.chart\_type\_priority}: KPI cards and metric cards, avoiding
table screenshots. \\
\midrule
Graphic designer & Visual direction showcase &
\texttt{layout.slide\_structure}: cover with theme/signature color
$\rightarrow$ mood board $\rightarrow$ color palette and usage ratio
$\rightarrow$ font pairing and type scale $\rightarrow$ grid/spacing rules;
\texttt{content.notes}: avoid information overload and keep the visual
direction sparse and clear. \\
Graphic designer & Design rationale pitch &
\texttt{layout.slide\_structure}: cover $\rightarrow$ design goals/audience
$\rightarrow$ constraints and brand tone $\rightarrow$ alternatives/trade-offs
$\rightarrow$ visual rationale;
\texttt{visual.chart\_type\_priority}: annotated design diagrams, trade-off
matrices, scorecards, and before/after cards. \\
Graphic designer & Design iteration review &
\texttt{layout.slide\_structure}: emphasize clear typography hierarchy and
avoid overlap between text, graphics, and containers;
\texttt{visual.chart\_type\_priority}: side-by-side before/after cards and
annotated mockups with numbered callouts. \\
\midrule
Training and development specialist & Capability training workshop &
\texttt{layout.slide\_structure}: organize by numbered learning-path steps;
\texttt{content.notes}: use action-oriented titles with step numbers and cover
objectives, key steps, examples, and recap reminders. \\
Training and development specialist & Learning solution pitch &
\texttt{layout.slide\_structure}: consolidate information into three or four
logical modules and reduce scattered bullets;
\texttt{content.notes}: prefer diagrammatic expression while keeping key terms
and conclusion prompts such as expert recommendations. \\
Training and development specialist & Rollout effectiveness review &
\texttt{layout.slide\_structure}: organize around training goals, execution
coverage, effectiveness observations, key issues, and improvement plans;
\texttt{visual.chart\_type\_priority}: process diagrams, highlighted tables,
and comparison tables. \\
\midrule
Financial manager & Budget and ROI assessment &
\texttt{layout.slide\_structure}: one table per slide for comparison/impact
tables, avoiding table walls; organize around inputs, expected returns, key
metrics, and constraints;
\texttt{visual.chart\_type\_priority}: key-variable impact tables and KPI
summary tables. \\
Financial manager & Finance control training &
\texttt{layout.slide\_structure}: organize around policy goals, key controls,
common risks, and execution requirements;
\texttt{visual.chart\_type\_priority}: comparison tables, checklists, process
diagrams, and KPI scorecards. \\
Financial manager & Financial risk review &
\texttt{layout.slide\_structure}: use conclusion/risk-stance titles and risk
indicator panels with changes in liquidity, leverage, and coverage;
\texttt{content.notes}: prefer quantitative indicators over vague wording. \\
\midrule
Operations manager & Execution alignment plan &
\texttt{layout.slide\_structure}: prioritize process flow, combine timelines
and kanban boards, and keep work-package boundaries clear;
\texttt{visual.chart\_type\_priority}: flowcharts, timelines, and kanban
boards. \\
Operations manager & Process adoption training &
\texttt{layout.slide\_structure}: convert lists into structured visuals:
horizontal process diagrams for steps, responsibility matrices for ownership,
and card grids for risks;
\texttt{visual.chart\_type\_priority}: horizontal flowcharts, card grids, and
responsibility matrices. \\
Operations manager & Delivery risk review &
\texttt{layout.slide\_structure}: use kanban/card layouts for baseline and
configuration points; use tables for risk-remedy mapping;
\texttt{content.notes}: cover delivery goal, current state, blocking risks,
impact scope, and remediation plan. \\
\midrule
Medical and health services manager & Clinical workflow evidence review &
\texttt{layout.slide\_structure}: structured ten-slide flow around workflow
status, key evidence, metric changes, and improvement directions;
\texttt{visual.chart\_type\_priority}: metric dashboards, horizontal clinical
pathway diagrams, alert cards, and risk panels. \\
Medical and health services manager & Compliance protocol training &
\texttt{layout.slide\_structure}: use tables for checklists and comparisons,
especially two-column tables in right-side modules;
\texttt{content.notes}: clearly distinguish investigational from approved
status for drugs or treatments. \\
Medical and health services manager & Quality and safety review &
\texttt{layout.slide\_structure}: visual-first layout with large metric cards,
comparison diagrams for risks, and flowcharts for processes;
\texttt{content.notes}: organize quality/safety reviews around key metrics,
incidents, cause analysis, corrective actions, and next steps. \\
\midrule
Legislator & Public issue evidence brief &
\texttt{layout.slide\_structure}: ten-slide evidence brief organized as issue
background---core evidence---policy options---public impact;
\texttt{visual.chart\_type\_priority}: comparison tables for policy trade-offs
and tabular evidence. \\
Legislator & Policy communication briefing &
\texttt{layout.slide\_structure}: policy goals $\rightarrow$ key content
$\rightarrow$ implementation points $\rightarrow$ public understanding path;
\texttt{visual.chart\_type\_priority}: evidence-linkage tables. \\
Legislator & Implementation accountability review &
\texttt{layout.slide\_structure}: cover---commitments/framing---implementation
progress and deviations---issue list and impact---accountability---remediation
and supervision---milestone timeline---decision points;
\texttt{visual.chart\_type\_priority}: responsibility matrices, issue-owner
tracking tables, and compact remediation Gantt charts. \\
\end{longtable}
\endgroup
