\gene{} reframes automated-research evaluation around scientific lineage. The central question is not only whether a system can write plausible research text, but whether it can identify, verify, and extend the inherited mechanisms that make a proposal a coherent descendant of prior work. The \ideagene{} framework---typed \genome{} objects, \genomediff{} records, and evolutionary dynamics---provides the representation layer for making that judgment precise and auditable.

The results make that distinction visible. Even the best system reaches only 27.3\% exact accuracy on \geneexam{}; tool scaffolds help T2 Inheritance Tracing but barely move T4 Lineage Verification; and structured lineage context separates systems rather than lifting them uniformly. In \genearena{}, PES decomposition shows the same plausibility--coherence gap: proposals can sound novel while failing to preserve the parent mechanism or limitation-delta relation needed for coherent descent. These findings point to a concrete design direction: auto-research systems need compositional verification modules, not only better retrieval. \gene{} provides an evaluation substrate for measuring whether future auto-research systems can move from retrieving papers and producing plausible text toward verifying and extending scientific lineages.

\section{Limitations}
The six evolutionary dynamics in \gene{} are operational categories for auditable evaluation, not an exhaustive theory of scientific development. Real lineages can mix multiple transition patterns within the same successor; \gene{} assigns primary drivers to keep annotation and evaluation consistent.
