%
\newenvironment{customabstract}{
    \begin{tcolorbox}[
        enhanced,
        colframe=taobaocolor,
        colback=white,
        boxrule=0.7pt,
        arc=12pt,
        auto outer arc,
        left=15pt,
        right=15pt,
        top=10pt,
        bottom=15pt,
        breakable
    ]
    \begin{center}
        \Large\bfseries\color{taobaocolor} Abstract
    \end{center}
    \vspace{0.2em}
}{
    \par\vspace{1em}
    \hfill \textit{Date: \today} % 这里添加时间
    \end{tcolorbox}
}

\begin{customabstract}
%
Human-centric video customization, particularly at the garment level, has shown significant commercial value. However, existing approaches cannot support low-latency and interactive garment control, which is crucial for applications such as e-commerce and content creation.
% 
This paper studies how to achieve interactive multi-garment video customization while preserving motion coherence using only single-garment video data.
% 
We present \textbf{\method}, a real-time and interactive framework for human-garment customization in autoregressive video generation, where users can interactively switch garment during generation.
\method consists of three key techniques:
% 
(i) Instead of training on multi-garment video data, we train a \textbf{Teacher Model with In-Context Learning} on a single reference–garment pair. By retaining the \textit{image-to-video training paradigm} while \textit{enforcing a mismatch between the reference and garment image}, the model is encouraged to implicitly preserve coherence during single-garment switching.
% 
(ii) To achieve consistency and efficiency during generation, we introduce \textbf{Streaming Distillation with In-Context Learning}, which fine-tunes the model with \textit{in-context teacher forcing} and improves extrapolation consistency via \textit{gradient-reweighted distribution matching distillation}.
% 
(iii) To extend the model for interactive multi-garment video customization, we propose \textbf{Training-Free KV Cache Rescheduling}, which includes \textit{garment KV refresh}, \textit{historical KV withdraw}, and \textit{reference KV disentangle} to achieve garment switching while preserving motion coherence.
% 
Our \method uniquely supports interactive customization and consistent long-video extrapolation, while achieving real-time generation at 23.8 FPS on a single GPU, 30-180$\times$ faster than existing baselines.
% 
\end{customabstract}
