\section{Introduction}
The striking advances of large foundation models have been driven by pretraining on 
text corpora at unprecedented scale \citep{brown2020language, kaplan2020scaling, hoffmann2022training}.
While effective, this paradigm rests on a strong implicit assumption: 
that all knowledge worth learning can be losslessly encoded as a linear sequence of text tokens.
However, cognitive science has long shown that humans routinely reason with
diagrams, spatial layouts, mathematical notation, and other representational
forms whose visual cues make certain relations directly available for
inference \citep{larkin1987diagram,zhang1997external,barsalou1999perceptual,barsalou2008grounded}.
Converting these visual cues into plain text before training is therefore inherently lossy.
Here we challenge this assumption and show that foundation models can 
learn directly from visual corpora without text extraction or image-text pairing supervision, 
yielding stronger language intelligence than text pretraining on the same underlying corpus.

Scientific documents are a particularly acute case of this loss.
Papers, textbooks, and technical reports communicate complex content through figures,
tables, formula layouts, and page-level spatial organization, all of which encode
geometric constraints, symbolic topologies, and structural correspondences essential
to scientific reasoning.
The recently proposed Platonic Representation Hypothesis \citep{huh2024platonic}
formalizes a related intuition for machine learning, arguing that representations learned
by different models and modalities converge toward shared abstractions of reality.
These observations imply that the linguistic information
extracted from a scientific document is only a projection of a richer underlying
structure, and that this structure could in principle be learned directly from raw
visual documents.

Yet existing approaches do not exploit this possibility along either of two dimensions.
At the data level, pretraining corpora reduce visual documents to plain text,
either by parsing HTML or LaTeX source \citep{paster2024openwebmath, gao2020pile}
or by applying neural document-parsing models to PDFs as a preprocessing step
\citep{blecher2024nougat, lee2023pix2struct, kim2021donut, wang2024mineru}, after
which the language model is trained exclusively on the resulting text.
At the training level, recent multimodal foundation models do incorporate visual modalities during training \citep{wang2024qwen2, chen2024internvl, liu2023visual},
but they treat visual inputs as conditioning context for text-token prediction, 
preventing visual content from being deeply integrated into the predictive process.
In both regimes the raw documents are consumed and then discarded leaving the visual modality outside the model's learning objective.

\input{tab/result_txt}

Here we present Visual Pretraining (VP), a framework in which a
foundation model learns visual information directly from raw documents without any text extraction or image-text pairing supervision.
Through a systematic study under carefully matched data sources, we
find that across multiple LLM backbones and scientific reasoning benchmarks, VP
consistently outperforms text-only pretraining on the same underlying corpus, is
substantially more efficient (using only 25\% of the token budget) with respect to both model size and data scale, 
and strengthens cross-modal alignment, establishing VP as a scalable pathway for
learning both language and visual intelligence.
This work makes three contributions. 
First, we show that reducing visual documents to plain text incurs substantial information loss, 
and that visual pretraining recovers this otherwise discarded information.
Second, we introduce an autoregressive visual pretraining framework that trains a foundation model to predict document patches in latent space, 
deeply integrating visual latent into the predictive process. 
Third, through a unified empirical study with matched corpora across multiple architectures and
benchmarks, we establish VP as an effective, efficient, and scalable alternative to text-only pretraining.

