\section{Introduction}
Olympiad competitions provide one of the clearest stress tests for long-horizon
reasoning. Unlike many standard benchmarks, these problems require a model to 
search over many possible solution paths, control assumptions precisely, verify
intermediate claims, and present a final argument that can survive strict
grading across mathematical and scientific settings. Recent systems have made
rapid progress in this direction: AlphaGeometry combined neural guidance with
symbolic search for olympiad geometry~\citep{trinh2024alphageometry}, while
AlphaProof, AlphaGeometry 2, and Gemini Deep Think reached silver- or
gold-medal standards on International Mathematical Olympiad problems with
larger search and verification budgets~\citep{deepmind2024alphaproof,deepmind2025geminiimo}.
At the same time, general reasoning models have improved through
chain-of-thought prompting, math-specialized post-training, and reinforcement
learning with verifiable rewards~\citep{wei2022chain,shao2024deepseekmath,yang2024qwen25math,r1,yan2025learning,exgrpo}, while scientific olympiad benchmarks test transfer to modeling, derivation, and competition-style justification~\citep{olympiadbench,chen2025p1,luo2026p1vl}.

% This report studies a practical question raised by this progress: 

% We use a 30B-A3B model as the main vehicle and keep
% the pipeline deliberately modular: supervised fine-tuning (SFT) first reshapes
% the model's reasoning behavior, reinforcement learning (RL) then scales solving
% capability, and test-time scaling (TTS) spends additional inference compute on
% the hardest proof-search instances.
% A central question is therefore \textit{whether a compact post-trained reasoning
% backbone can be pushed to olympiad-level performance with a simple and
% unified
% recipe}.  
A central question is therefore whether \textit{a reasoning backbone can be pushed to olympiad-level performance with a compact, domain-unified recipe} that
applies the same reasoning-centric pipeline across mathematical and scientific problems. 
Using
a 30B-A3B model, we build a modular pipeline: SFT reshapes reasoning behavior, RL scales solving capability, and TTS allocates additional inference compute to the
hardest proof-search problems. Together, these stages align behavior shaping, reward design, experience replay, and self-verification into a compact recipe for rigorous
mathematical and scientific reasoning. The desgin follows a specializable-generalist view: rather than building a narrow olympiad solver, we specialize a broadly capable post-trained model toward expert-level proof reasoning while preserving transfer across scientific domains.
  
\textbf{The first stage aims to instill a more disciplined proof-search pattern.} 
Starting from a post-trained model that is already competitive on scientific reasoning tasks,
we curate long-form solution, self-verification, and self-refinement
trajectories from mathematical, scientific, coding, and instruction-following
sources. 
After filtering, the SFT mixture contains 338K
trajectories with responses shorter than 8K tokens. SFT on this rigorous proof
data instills reasoning behaviors centered on proof search, self-checking, and
repair. We then order the examples by reverse perplexity so
that each pass starts with trajectories most mismatched to the initial policy
before consolidating on more familiar examples. This curriculum helps preserve
and recover the capability of the post-trained model with its reasoning
behavior reshaped.
% \haoran{Why does a high-perplexity-first ordering help the model learn the form of rigorous reasoning?}

\textbf{The second stage scales this behavior through two levels of RL.}
% \haoran{and TTS?}
Coarse RL uses
verifiable prompts and efficient outcome checking to scale the reasoning
behaviors introduced by SFT under reliable binary rewards, following the broader RLVR paradigm for efficient reasoning improvement~\citep{r1,shao2024deepseekmath}. 
Refined RL then shifts the target from answer correctness to proof quality.  It combines a proof-level generative reward model for scoring complete proofs,
self-refinement prompts for training critique-and-repair behavior, and
experience replay for preserving rare successful trajectories on hard
problems.
Finally, we apply test-time scaling through a self-verification-and-refinement loop to elevate the trained
model to olympiad-level reasoning~\citep{huang2025winning}.


On answer-verifiable benchmarks, the resulting model,
\textbf{\textcolor{iclrdeepblue}{SU-01}}, nearly matches the strongest
similar-size baseline, Qwen3.6-35B-A3B, across AnswerBench, AMO-Bench, AIME
2025/2026, and FrontierScience-Olympiad. On proof-oriented evaluation, SU-01
reaches 57.6\% on IMO-ProofBench with direct generation and 70.2\% with TTS,
substantially outperforming similar-size models and approaching competitive
commercial systems such as Gemini 3.1 Pro Thinking. Beyond solving
competition problems, SU-01 obtains the best similar-size overall score on
FrontierScience-Research, suggesting that the recipe generalizes scientific
reasoning toward research-style problems beyond olympiad benchmarks.

On official competition problems, SU-01 shows strong end-to-end reasoning beyond benchmark-style evaluation. 
Direct SU-01 already exceeds the IPhO gold lines for
both 2024 and 2025, and clears the bronze-medal lines on IMO 2025 and USAMO 2026. 
With test-time scaling, it reaches \textbf{35 points} on both mathematical olympiads, meeting the \textbf{IMO
2025 gold} line and exceeding the \textbf{USAMO 2026 gold} line by 10 points. 
Notably, on USAMO 2026, this matches the highest reported human total among 340 competitors,
indicating that the overall recipe can elicit top-level human-like olympiad reasoning from a compact 30B-A3B model. 
The TTS traces further show how this
capability emerges at inference time: SU-01 can sustain reasoning trajectories beyond 100K tokens, condition on its own drafts and error analyses, and repeatedly
verify and repair candidate proofs. 
Overall, these results support a specializable-generalist view of compact reasoning
models: with the right training and inference recipe, a broadly capable backbone
can be driven toward expert-level proof reasoning while retaining meaningful
scientific transfer. 
% than single-pass evaluation reveals.

 