\section{Conclusion}

% This report presents a simple and unified recipe for moving a compact
% post-trained reasoning model toward olympiad-level mathematical and scientific
% reasoning. By combining reverse-perplexity curriculum SFT, efficient coarse RL
% with outcome verification, refined RL with proof-level rewards, self-refinement,
% and experience replay, and finally test-time scaling, the recipe decomposes hard
% reasoning improvement into a small number of aligned stages. The resulting
% \textbf{SU-01} shows that a 30B-A3B backbone can reach
% gold-medal-level olympiad performance while remaining efficient, and suggests a
% broader path for scaling rigorous reasoning through better behavior shaping,
% reward design, verifier reliability, and training--inference co-design.
This report presents a simple and unified recipe for turning a compact
post-trained reasoning model into a stronger mathematical and scientific
reasoner. Starting from a broadly capable 30B-A3B backbone, SU-01 combines
reverse-perplexity curriculum SFT, efficient coarse RL with outcome verification,
refined RL with proof-level rewards, self-refinement and experience replay, and
test-time scaling through self-verification and refinement. Together, these
stages decompose rigorous reasoning improvement into behavior shaping, scalable
reward feedback, proof-level specialization, and inference-time repair. 
The resulting model reaches gold-medal-level performance on mathematical and
physical olympiad competitions, sustains reasoning trajectories beyond 100K
tokens during inference, and shows transfer to scientific domains beyond the
main math and physics training signals. 
In summary, SU-01 supports a specializable-generalist view of compact reasoning models: with the right training and inference recipe, a broadly capable backbone can be driven toward expert-level proof reasoning while retaining meaningful scientific transfer.


\section*{Acknowledgments}
This work was supported by the Shanghai Artificial Intelligence Laboratory. 
We thank the authors and maintainers of prior open research and infrastructure
that made this work possible. 
In particular, we are
grateful to DeepSeek for open-sourcing strong reasoning policies and generative reward models, which provided an important
reference point for our work~\citep{deepseekai2025deepseekv32,deepseekmathv2}.
IMO-Bench, AMO-Bench, and FrontierScience
helped guide the overall system optimization by offering challenging
mathematical and scientific reasoning benchmarks and evaluation
protocols~\citep{luong2025imobench,an2025amobench,wang2026frontierscience}. We
also thank prior data efforts that supported our SFT and RL data curation,
including DeepMath, NaturalReasoning, Eurus, OpenCodeReasoning, P1, and
OPC~\citep{he2025deepmath,yuan2025naturalreasoning,cui2025process,ahmad2025opencodereasoning,chen2025p1,dekoninck2025opc},
as well as the many public problem sources and communities that cannot all be
listed here. We further acknowledge the broader open-source infrastructure
ecosystem, including slime for training and SGLang for efficient inference and
serving~\citep{thudm2026slime,sglang2026repo}. 

