\section{Achieving Gold-Medal-Level Reasoning via Test-time Scaling}
\label{sec:tts}

% Training alone is not the full story for olympiad-level reasoning. 
Even with a
strong reasoning policy, the hardest problems often require substantial search
and revision at inference time. This is not merely a matter of sampling more
answers. IMO-style tasks demand complete and rigorous proofs, and a solution
with the right final conclusion can still fail if it contains a hidden gap or a
logical fallacy. Recent work on IMO 2025 makes this point explicit: strong
frontier models already contain significant mathematical capability, but their
single-pass outputs and even best-of-many selection can remain far below the
level obtained by a structured verification-and-refinement pipeline
\citep{huang2025winning}. 
% Test-time scaling is therefore a core part of the
% SU-01 recipe rather than a minor evaluation trick.

The need for test-time scaling is also tied to reasoning budget. A single generation has a
finite context and thinking budget, while an olympiad proof may require several
rounds of exploration, lemma checking, counterexample search, and exposition
repair. A model can spend most of its budget discovering a promising approach
and still fail to fully close the proof. Breaking inference into repeated
solve--verify--refine stages effectively allocates additional computation to the
same problem while keeping each step focused and auditable. This extra budget is
useful only when the model can remain coherent across repeated drafts,
critiques, and repairs. 
After the full training pipeline, SU-01 is able to use
this budget productively on difficult problems, sustaining coherent reasoning
trajectories longer than 100K tokens during inference.

Our TTS procedure follows the verification-and-refinement paradigm of
\citet{huang2025winning} as a self-verification and refinement loop. The model
first produces an initial solution under a solver prompt that prioritizes proof
rigor rather than merely reaching a final answer. It then enters refinement,
where it revisits the draft, repairs weak steps, and tries to turn a promising
argument into a complete proof. The refined candidate is next checked through a
verification prompt: the model inspects the full solution and writes a structured
bug report, identifying issues such as critical errors, unjustified claims, or
missing cases. A verdict step interprets this report and decides whether the
candidate should be accepted, rejected, or sent back for another refinement
round. This loop is repeated until the solution consistently passes
self-verification or the refinement budget is exhausted. Multiple independent
runs can be executed in parallel or serial, and accepted candidates are selected
only after the proof is stable under repeated verification. The corresponding
inference setting is summarized in \cref{app:inference-serving-details}.

% This inference-time procedure matches the behavior encouraged during training.
% SFT installs a proof-search grammar, refined RL rewards complete and rigorous
% solutions, self-refinement teaches the model to repair flawed arguments, and
% experience replay preserves rare successful proof trajectories. Test-time
% scaling then converts these learned behaviors into stronger end performance by
% allowing the model to search more broadly, verify more carefully, and refine
% candidate solutions more effectively.
