\section{Boosting Reasoning Capability with RL}
\label{sec:rl}

Once the model has acquired a stronger long-form reasoning pattern,
reinforcement learning provides the scalable feedback mechanism for turning this
pattern into stronger expert behavior. We split this stage into two levels.
\textbf{Coarse RL} converts the SFT reasoning pattern into stronger
answer-seeking behavior under reliable, mostly verifiable reward signals,
improving search, coverage, and direct solving performance on hard tasks.
\textbf{Refined RL} then specializes the policy toward complete, auditable proof
construction, using more fine-grained feedback to encourage proof rigor and
self-refinement.
% \haoran{\textbf{Refined RL} then targets the quality of the reasoning process itself, using more fine-grained feedback to encourage proof rigor and self-refinement.}
The shared
RL launch configuration and stage-specific hyperparameters are summarized in
\cref{app:rl-training-details}.

\subsection{RL Data Curation}

RL training uses a separate prompt pool from SFT, curated to support both
answer-verifiable optimization and proof-quality refinement. The physics subset
is derived from olympiad-level physics data associated with P1
\citep{chen2025p1}. The mathematical subset follows the same source families as
our SFT data, including AoPS, online competition training books, Evan Chen's
olympiad materials, and the Shuzhimi Forum. We refer readers to
\S\ref{sec:sft-data-curation} for source attribution. We additionally include
OPC\footnote{OPC dataset card:
\href{https://huggingface.co/datasets/INSAIT-Institute/OPC}{link}.}, a
human-evaluated corpus of advanced mathematical proofs \citep{dekoninck2025opc},
to increase coverage of proof-oriented prompts.

We split the resulting RL pool into a verifiable set and a non-verifiable set.
The verifiable set contains prompts whose final answers or structured outputs
can be checked reliably, while the non-verifiable set includes proof-oriented
or open-ended reasoning prompts that require softer judgment, e.g., generative reward. 
% \haoran{softer judgment? One more concise sentence of explanation would help here}. 
Before training,
we first deduplicate and decontaminate the prompt pool. We then apply rejection
sampling to remove examples that are already too easy or too hard for the
current policy, and further filter noisy prompts that are poorly formatted or otherwise unreliable. The final RL pool
contains 8,967 verifiable prompts and 16,287 non-verifiable prompts.


\subsection{Coarse RL}
\label{sec:coarse-rl}

Coarse RL trains the SFT model on the 8,967 verifiable prompts described above.
We formulate this stage as reinforcement learning with verifiable rewards
(RLVR; \citealt{lambert2024tulu,r1}), using Group Sequence Policy Optimization
(GSPO; \citealt{zheng2025gspo}). GSPO is better aligned with outcome-reward
training than token-level GRPO because both reward assignment and policy
clipping operate at the complete-response level. For each prompt
$q \in \mathcal{D}_{\mathrm{ver}}$ (the verifiable prompt set), the rollout policy
$\pi_{\theta_{\mathrm{old}}}$ samples a group of $K$ candidate solutions
$\mathcal{G}_q=\{o_i\}_{i=1}^K$. The verifier converts each final answer into a
binary outcome reward $r(q,o)=1$ if the extracted final answer is verified as
correct, and $r(q,o)=0$ otherwise.
The group-relative advantage is computed from the within-prompt reward baseline.
We use the unnormalized form without group standard-deviation normalization,
$\widehat{A}_i=r(q,o_i)-\mu_{\mathcal{G}_q}$, where
$\mu_{\mathcal{G}_q}=\frac{1}{K}\sum_{j=1}^{K}r(q,o_j)$.
The key GSPO quantity is the length-normalized sequence-level importance ratio
$s_i(\theta)=\exp\{\frac{1}{|o_i|}\sum_{t=1}^{|o_i|}\log
\frac{\pi_{\theta}(o_{i,t}\mid q,o_{i,<t})}
{\pi_{\theta_{\mathrm{old}}}(o_{i,t}\mid q,o_{i,<t})}\}$. The policy is
updated with the clipped sequence-level surrogate
\begin{equation}
\mathcal{J}_{\mathrm{GSPO}}(\theta)=
\mathbb{E}_{q,\{o_i\}}
\left[
\frac{1}{K}\sum_{i=1}^{K}
\min\left(
s_i(\theta)\widehat{A}_i,
\operatorname{clip}(s_i(\theta),1-\epsilon,1+\epsilon)\widehat{A}_i
\right)
\right].
\label{eq:coarse-gspo-objective}
\end{equation}
% Compared with token-level clipping, this objective treats each generated
% solution as the unit of correction, matching the binary verifier reward above.
These definitions are also the interface used by the experience replay variants in the subsequent subsection: 
replayed trajectories can reuse the same reward, advantage, and sequence-ratio
notation while changing the source policy in the denominator of $s_i(\theta)$.
Following the routing-replay motivation in GSPO \citep{zheng2025gspo}, we freeze
the MoE router during RL so replayed trajectories are evaluated under stable
expert-routing decisions, which reduces replay-induced instability.

The reward system is intentionally layered to keep high-precision automatic
checks before more expensive model-based judgments. We first extract the final
answer and apply canonicalized text matching. Unresolved cases are then checked by Math-Verify\footnote{Math-Verify
repository: \href{https://github.com/huggingface/Math-Verify}{link}.}, a
rule-based mathematical-expression evaluation pipeline for LLM outputs. Samples
that still fail these rule-based checks are sent to
gpt-oss-120b\footnote{gpt-oss-120b model card:
\href{https://huggingface.co/openai/gpt-oss-120b}{link}.} \citep{openai2025gptoss}
for generative verification. This ordering makes the reward conservative by
default, while still recovering correct solutions whose final answers are
equivalent but difficult to normalize with rule-based parsers alone.

\subsection{Refined RL}

After coarse RL has established strong search behavior, refined RL shifts the
optimization target from answer correctness to proof quality. The central issue
is that many olympiad solutions can reach a correct final answer while still
containing hidden gaps, unjustified transformations, or incomplete case
analysis. Refined RL therefore uses a stronger process-level reward and adds two
memory mechanisms: self-refinement, which turns recent failures into repair
tasks, and experience replay, which preserves rare successful proofs long
enough for the policy to learn from them.

\paragraph{Generative proof reward.}
We use DeepSeekMath-V2 as a generative reward model for refined RL
\citep{deepseekmathv2}, except for physics prompts. 
For every rollout from both the verifiable and
non-verifiable subsets, the reward model reads the problem and the complete
solution or proof, then outputs a binary score $r_{\mathrm{proof}}(q,o)\in
\{0,1\}$. 
Unlike the coarse verifier in \S\ref{sec:coarse-rl}, this score is not restricted to
checking whether the final answer matches a reference answer. 
It evaluates
whether the full reasoning path is mathematically valid, sufficiently rigorous,
and complete. This makes the reward more aligned with the final goal of
olympiad reasoning, but also more expensive and more vulnerable to judge
artifacts. We therefore apply anti-hack preprocessing before sending a response
to the reward model: malformed generations with leaked chat-template tokens,
unbalanced thinking delimiters, or severe repetition are replaced by a safe
fallback answer. This prevents the policy from receiving reward by exploiting
formatting or verifier-input pathologies rather than by improving the proof. The
reward-model serving configuration is summarized in
\cref{app:inference-serving-details}.

\paragraph{Self-refinement.}
Self-refinement exposes the policy to the same repair pattern that we use at
test time: propose a solution, inspect it, locate gaps, and produce a corrected
proof. After each rollout, responses are grouped by query. If a query group has
average proof reward below a threshold $\tau_{\mathrm{ref}}=0.5$, failed
responses from that group are converted into refinement prompts. Each prompt
contains the original problem, the previous incorrect solution, and an
instruction to critique the argument, fix proof 
% \runzhe{the corrections were not restricted just to math and physical queries} 
errors, fill
missing justifications, and output a complete final solution. These prompts are
stored in a self-refinement buffer and mixed into subsequent batches with target
ratio $\eta_{\mathrm{ref}}=0.2$. Normal samples displaced by refinement queries
are returned to a buffer, so refinement does not silently discard fresh
training data. We also do not recursively enqueue failed refinement attempts,
which avoids spending repeated updates on examples that remain outside the
current policy's learnable region.

\paragraph{Experience replay.}
% The second memory mechanism is experience replay.
On difficult proof problems,
the policy may occasionally discover a valid solution trajectory even though it
usually fails on the same query. Immediately discarding such a trajectory wastes
a high-value training signal. Following ExGRPO
\citep{exgrpo}, we keep a replay buffer $\mathcal{E}$ keyed by query, but our
implementation is simpler: it uses the same GSPO-style update and does not apply
the policy-shaping transform introduced in ExGRPO. After each rollout, a query
is admitted to the replay buffer only when it is hard but solvable, operationalized as
$0<n_+(q)<2$, where $n_+(q)$ is the number of successful trajectories in the
current group. In answer-only RLVR, such a unique success can be a lucky hit: a
trajectory may end with the correct final answer while still containing brittle
or invalid reasoning\citep{exgrpo}. 
In our refined RL setting, however, success is assigned by the
DeepSeekMath-V2 proof reward, which inspects the full solution rather than only
the final answer. This does not eliminate reward-model noise, but it makes a
rare successful rollout substantially more likely to encode a reusable proof
pattern and therefore a safer replay target. Stored trajectories are
deduplicated, and a query is retired once fresh on-policy rollouts solve it
often enough, with threshold $n_+(q)\ge 4$.

Replay is mixed with fresh proof-reward training rather than run as a separate
mode. A replay rollout injects one stored successful trajectory for the selected
query, and the replay ratio is controlled by $\rho=0.25$ over the non-refinement
portion of the batch. When multiple successful trajectories are stored for a
query, we select the lowest-entropy one, $o^*=\arg\min_{o\in\mathcal{E}(q)}
H(o;\pi_\theta)$, using rollout-side top-$k$ log probabilities as an efficient
entropy estimate, following the trajectory-selection principle in ExGRPO
\citep{exgrpo}. The resulting refined objective is
\begin{equation}
\mathcal{J}_{\mathrm{refined}}(\theta)=
(1-\rho)\,\mathbb{E}_{\mathcal{B}_{\mathrm{fresh}}}
\left[\mathcal{J}_{\mathrm{GSPO}}(q,\mathcal{G}_q;\theta,\pi_{\theta_{\mathrm{old}}})\right]
+\rho\,\mathbb{E}_{\mathcal{B}_{\mathrm{exp}}}
\left[\mathcal{J}_{\mathrm{GSPO}}(q^*,\{o^*\}\cup\mathcal{G}_{q^*};\theta,\pi_{\theta_{\mathrm{src}}})\right],
\label{eq:refined-replay-objective}
\end{equation}
where $\pi_{\theta_{\mathrm{src}}}=\pi_{\theta_{\mathrm{past}}}$ for the replayed
trajectory and $\pi_{\theta_{\mathrm{src}}}=\pi_{\theta_{\mathrm{old}}}$ for fresh
rollouts. This replay design is targeted rather than exhaustive: it stores rare
valid proofs, prefers the most stable stored trajectory, replays it at a
controlled ratio, and removes it once the current policy can reliably reproduce
the behavior on-policy.
