\section{Related Work}
\label{sec:related_work}

\subsection{Consistency Models}
Consistency Models (CMs) \cite{song2023consistency} accelerate diffusion sampling by directly learning a mapping from $\mathbf{z}_t$ to $\mathbf{z}_0$ along the probability-flow ODE (PF-ODE) defined by a teacher model. For multi-step generation, CMs repeatedly inject noise into intermediate states and perform iterative denoising. Subsequent studies improve CM training stability through annealed time schedules \cite{song2024improved, geng2024consistency} and segmented consistency objectives \cite{wang2024phased, ren2024hyper, lee2024truncated}. sCM \cite{lu2024simplifying} further simplifies these design choices and provides a strong practical baseline. Building on sCM, rCM \cite{zheng2025large} introduces score distillation as a regularizer, achieving strong performance in both image and video diffusion distillation. In contrast to prior CM-based approaches, we explore flow map distillation to enable any-step video generation.


\subsection{Flow Map Models}
Flow Map Models~\cite{boffi2024flow, sabour2025align, kim2024consistency} offer a unified perspective on diffusion modeling and diffusion distillation by learning a generalized transition operator $\mathbf{f}_\theta(\mathbf{z}_t, t, r)$ between arbitrary points along PF-ODE trajectories. From this perspective, the formulation recovers consistency modeling~\cite{song2023consistency} when $r=0$ and reduces to standard flow matching~\cite{liu2022flow, lipman2022flow} when $t=r$. However, training flow map models is challenging because it requires accurate transition learning across arbitrary time pairs. MeanFlow~\cite{geng2025mean} is a representative approach that connects instantaneous and average velocities, but it relies on Jacobian-vector products (JVPs), which are difficult to scale under Fully Sharded Data Parallel (FSDP). To overcome this limitation, subsequent studies either approximate JVP terms numerically~\cite{wang2025transition} or derive JVP-free algebraic formulations~\cite{luo2025soflow, guo2025splitmeanflow}. Building on these advances, AnyFlow leverages flow map distillation for scalable any-step video diffusion, enabling efficient and flexible inference for high-fidelity generation.

\subsection{Video Diffusion Distillation}
As shown in \cref{tab:paradigm_comp}, recent video diffusion distillation methods typically follow a two-stage pipeline: (1) forward training for initialization and (2) on-policy distillation for refinement. The first stage improves optimization stability, while the second reduces rollout drift during few-step inference. For example, rCM~\cite{zheng2025large} initializes from sCM~\cite{lu2024simplifying} and then applies consistency backward simulation~\cite{yin2024improved} with distribution matching~\cite{yin2024one}. In causal video generation, Self-Forcing~\cite{huang2025selfforcing} adopts data-free consistency ODE initialization~\cite{yin2025causvid} before performing a similar on-policy refinement stage. Our method follows the same high-level pipeline but replaces consistency modeling with
a flow map formulation, enabling stronger any-step performance.

A concurrent work, TMD~\cite{nie2026transition}, also studies a flow map formulation for bidirectional video diffusion distillation. The main difference lies in how efficient rollout is achieved. TMD improves rollout efficiency from an architectural perspective by sharing the backbone and introducing an additional flow head. In contrast, our method improves simulation efficiency from the flow-trajectory perspective by decomposing a full trajectory into shortcut transitions based on the composition property of flow maps. As a result, our approach naturally supports arbitrary step budgets and generalizes to both bidirectional and causal architectures.

\myPara{Causal Video Diffusion.} A growing line of work studies causal video diffusion models~\cite{yang2025longlive, jinpyramidal,zhang2025framepack,songhistory}. These models often suffer from exposure bias, which leads to error accumulation during autoregressive generation. Self-Forcing~\cite{huang2025selfforcing} mitigates this issue through on-policy distillation, but its consistency-based formulation is mainly tailored to few-step settings for rollout efficiency. Another line of work~\cite{guo2025end} focuses more on multi-step models by explicitly modeling test-time errors during pretraining, but it is not directly designed for efficient few-step sampling. By contrast, our method starts from a flow map formulation and uses flow map backward simulation to support both few-step and multi-step causal video generation within a unified framework.


\subsection{On-Policy Distillation}
On-policy distillation mitigates test-time errors  by training the student on its own generated trajectories and supervising them with a stronger teacher. This approach has demonstrated benefits in both large language models \cite{lu2025onpolicy, zhao2026self, ye2026policy} and diffusion models \cite{yin2024one, huang2025selfforcing, yin2024improved}.

For diffusion models, on-policy distillation is typically implemented via Distribution Matching Distillation (DMD)~\cite{yin2024one, yin2024improved} or adversarial distillation~\cite{lindiffusion, lin2025autoregressive}. Self-Forcing~\cite{huang2025selfforcing} combines self-rollouts with distribution matching against a bidirectional teacher, while APT2~\cite{lin2025autoregressive} evaluates one-step video rollouts using a discriminator. Our method follows the same on-policy distillation paradigm but introduces flow map backward simulation, specifically designed for the flow map formulation, enabling any-step video generation.