\section{Definitions \& Key Terms}
\label{app:definitions}

\begin{description}[leftmargin=2.2cm, labelwidth=2cm, labelsep=0.2cm, style=nextline]

  \item[\textbf{STT}] \textit{Speech-to-Text.} A model or service that transcribes spoken audio into text. In voice agent pipelines, STT serves as the input stage, converting user utterances into a form consumable by a downstream language model. \label{def:stt}

  \item[\textbf{TTS}] \textit{Text-to-Speech.} A model or service that synthesizes natural-sounding speech from text. In voice agent pipelines, TTS serves as the output stage, rendering language model responses as audio delivered to the user. \label{def:tts}


\item[\textbf{S2S}] \textit{Speech-to-Speech.} An end-to-end model architecture that accepts raw audio as input and produces audio as output, bypassing discrete STT and TTS stages entirely. S2S systems may also emit text output, used to invoke tool calls, alongside or prior to synthesizing a spoken response. S2S systems minimize modality-crossing latency and preserve paralinguistic signals throughout the pipeline. \label{def:s2s}

  \item[\textbf{LALM}] A large language model that accepts audio as a direct input modality---without prior transcription---and produces text output. AudioLLMs internalize acoustic features (prosody, speaker characteristics, noise) that would otherwise be lost in an STT transcription step. Also referred to as an \textbf{AudioLLM} or a \textbf{SpeechLLM}. \label{def:audio-llm}

  \item[\textbf{Hybrid}] A voice agent architecture that combines an AudioLLM (for audio-aware language understanding) with a discrete TTS module (for speech synthesis). Hybrid systems are Audio-Native on the input side while retaining a text-to-speech output stage, and occupy a middle ground between fully cascaded (STT$+$LLM$+$TTS) and fully end-to-end (S2S) pipelines.  \label{def:hybrid}
    
  \item[\textbf{Audio-Native}] An umbrella term for voice agent architectures that process audio directly at one or more stages, rather than relying solely on text-based STT$\to$LLM$\to$TTS cascades. Audio-Native systems include both \textit{S2S} and \textit{Hybrid} configurations.  \label{def:audio-native}

  \item[\textbf{Multi-turn}] A conversational interaction comprising more than one exchange between user and agent---i.e., at least one agent response followed by a subsequent user utterance within the same session. Multi-turn evaluation tests an agent's ability to maintain task state, handle clarifications, and recover from errors across dialogue turns. \label{def:multi-turn}

  \item[\textbf{Bot-to-Bot}] An automated evaluation protocol in which a \textit{user simulator} (a separate agent or model) plays the role of the human caller, conducting full spoken conversations with the system under evaluation.  \label{def:bot-to-bot}


\item[\textbf{pass@1}] The fraction of all trials that pass, measuring average performance: $\text{pass@1} = \frac{1}{T}\sum_{t=1}^{T} \mathbf{1}[\text{pass}^{(d)}_t]$, where $T = Nk$ is the total number of trials across $N$ scenarios and $k$ trials each.  \label{def:passatone}

  \item[\textbf{pass@k}] The fraction of scenarios where at least one of $k$ trials passes, measuring ceiling performance: $\text{pass@k} = \frac{1}{N}\sum_{i=1}^{N} \mathbf{1}\!\left[\sum_{j=1}^{k} \mathbf{1}[\text{pass}^{(d)}_{i,j}] \geq 1\right]$. See also: \textit{Peak}. \label{def:passatk}

  \item[\textbf{pass\^{}k}] The expected probability that a system passes all $k$ independent future trials on a given scenario, measuring reliable performance: $\text{pass\^{}k} = \frac{1}{N}\sum_{i=1}^{N} \hat{p}_i^{\,k}$, where $\hat{p}_i = \frac{1}{k}\sum_{j=1}^{k} \mathbf{1}[\text{pass}^{(d)}_{i,j}]$ is the per-scenario pass rate across $k$ trials. See also: \textit{Reliable}. \label{def:passpowerk}


    \item[\textbf{Peak}] Generally refers to \textit{\passatk} scores. A peak score characterizes the best-case performance of a system over $k$ independent attempts---i.e., the probability that at least one of $k$ runs produces a correct outcome. Peak scores are informative about a system's ceiling capability but do not reflect consistency.  \label{def:peak}

  \item[\textbf{Reliable}] Generally refers to \textit{\passpowerk} scores. A reliability score characterizes the probability that a system produces a correct outcome on \textit{every one} of $k$ independent attempts. Reliability scores penalize variance and are informative about how consistently a system can be expected to succeed across repeated deployments.  \label{def:reliable}


\end{description}
