\clearpage
\section{User Simulator Details}
\label{app:user-sim}
We use ElevenLabs ElevenAgents \cite{elevenlabs2024agents} as the user simulator with the following cascade system: \texttt{Scribe v2.2 Realtime + GPT-5.1 + Eleven V3 Conversational}. We select these models for their high transcription accuracy, \metricbehavior, user realism for GPT-5.1 \cite{zhou2026mind}, and for their naturalness and realism for Eleven v3 Conversational. ElevenLabs also provides a large voice library, enabling testing of a wide variety of user accents, languages, speaking styles, etc.

We created four ElevenLabs agents for the user simulator, covering two accents (English and French) and two genders each. When creating a new agent, select \textit{Blank Agent} as the starting template, then apply the configuration described in Table~\ref{tab:elevenlabs-config}. The four voice names are listed in Table~\ref{tab:elevenlabs-voices}. All parameters not listed are set to their default values provided by ElevenLabs at agent creation.

\input{tables/08_appendix/d_user_simulator/01_el_simulator_config}

\input{tables/08_appendix/d_user_simulator/02_voice_name}

When enabling the "End Conversation" system tool, the name must be \texttt{end\_call}, and the description to provide is shown below. This allows the simulator to hang up programmatically.

\begin{tcolorbox}[promptbox, title={\small "End Conversation" Tool Description}]
Use this to end the phone call and hang up.

Call this function when its time to end the call and one of the following is true:
\begin{enumerate}
    \item The agent has confirmed your request is resolved (all steps are completed) and you have said goodbye
    \item The agent has initiated a transfer to a live agent
    \item The agent has been unable to make progress for at least 5 consecutive turns
    \item The agent says goodbye or indicates the conversation is over
    \item The agent indicates that the remainder of your request cannot be fulfilled.
    \item If the assistant says something along the lines of "I'm sorry I encountered an error processing your request."
\end{enumerate}

IMPORTANT: never call this tool in the same turn that you provide the agent with data, an identifier, a request to transfer to a live agent, an approval to proceed, or any kind of additional information. 

Before calling this tool, always say a brief goodbye first.
\end{tcolorbox}

Once the agent is configured, click \texttt{Publish} in the top-right corner. The \texttt{agent-id} can be retrieved from the \texttt{Widget} tab of the agent dashboard, under \texttt{Embed code}.

The simulator is prompted in ~\framework~with a specific user goal and is instructed to stay on task, communicate all required named entities clearly, and terminate the conversation when the goal is accomplished, or the task is clearly unlikely to succeed. The system prompts are provided in Appendix \ref{app:simulator-prompts}.

\input{tables/08_appendix/d_user_simulator/03_validation_metrics}

\subsection{User Simulation Validation}

Table~\ref{tab:validation_metrics} lists the three validation metrics used to verify user-simulator behavior and conversation integrity before any \framework~evaluation is performed.

\subsubsection{\metricvalidend.}
Before invoking any LLM judges, we run a deterministic check on the conversation logs to verify that each simulation terminated correctly. A valid end state is one in which either the agent failed to respond to the user, or the user invoked the end-call tool. Conversations meeting this criterion proceed to the User Speech and Behavioral Fidelity judges; all others are rerun.

This gate primarily catches infrastructure-level failures: WebSocket connections that closed unexpectedly, conversations that failed to start, or user simulator timeouts. These are pipeline errors rather than agent errors, and filtering them deterministically before any judge invocation avoids wasting LLM calls on malformed simulations.

\subsubsection{\metricbehavior.}

\textbf{Motivation.}
A key assumption underlying our evaluation is that the user simulator faithfully follows its assigned goal and decision logic, since the ground truth end database state is derived from this assumption. If the user deviates, the agent may fail to reach the ground truth state through no fault of its own. We therefore define a validation gate that detects user behavior errors capable of corrupting the evaluation, organized into five \emph{corruption types}:

\begin{enumerate}
    \item \textbf{Extra modifications.} The user makes requests beyond its stated goal that invoke modification tools writing to the scenario database. The user simulator prompt explicitly instructs the user to decline any such offers from the agent, but we check for violations regardless.

    \item \textbf{Premature ending.} In our simulations, the user is responsible for ending the call once its goal is complete. If the user hangs up prematurely—for example, providing actionable information and ending the call in the same turn—the agent has no opportunity to execute the required tool calls. We therefore verify that the user does not terminate the conversation in the same turn it provides critical information to the agent.

    \item \textbf{Missing information.} If the user fails to provide information the agent needs to complete the task, the evaluation is corrupted since task success cannot reasonably be expected.

    \item \textbf{Duplicate modifications.} Occasionally, the user simulator (particularly when using non-primary models) enters a loop and repeats requests the agent has already fulfilled. The agent then acts on the duplicate request, causing redundant writes to the scenario database that cause the final state comparison to fail.

    \item \textbf{Decision tree violations.} Each user is given a structured decision tree governing how to navigate choices during the interaction (e.g., \textit{``accept the earlier flight if the price difference is under \$200, otherwise decline''}). We verify that the user adheres to this logic, since deviations would cause the agent to reach a final state inconsistent with the ground truth.
\end{enumerate}

\textbf{Implementation.}
We prompt a judge model with a description of the five corruption types, the full conversation trace (including agent tool calls), and the list of available agent tools. The judge first produces a \emph{corruption analysis} assessing each corruption type in turn, then outputs a binary flag per corruption type, and finally an overall binary rating for the conversation. Conversations receiving a rating of 0 are rerun. See~\autoref{app:prompt-user-behavior} for the full prompt.

\textbf{Validation.}
To validate the judge, we constructed a human-annotated dataset of real user simulator failures from earlier evaluation artifacts alongside correct behavior examples, labeled with both overall ratings and per-corruption-type annotations. GPT-5.2 with medium reasoning achieved 100\% accuracy across three independent runs, and was therefore selected as the judge model.

\textbf{Error Categorization.}
Across 714 trials flagged by the \metricbehavior~judge—spanning four evaluated systems across all domains—premature ending was the most frequent corruption type (63.9\%), followed by decision tree violations (52.9\%), missing information (28.6\%), extra modifications (1.3\%), and duplicate modifications (0\%). Notably, 42.4\% of flagged trials exhibited two or more corruption types simultaneously.

\subsubsection{\metricuserspeechfidelity.}

\textbf{Motivation.}
As noted in \autoref{sec:conversation-simulation}, speech fidelity flags are rare in practice, which may invite questions about why this validation gate was included at all. The answer lies in one of the primary failure modes we observe in voice agents: failure to transcribe and understand key entities. To attribute such failures confidently to the agent rather than the user simulator, we must verify that the user's synthesized speech correctly conveys all critical entities. This is particularly consequential because most flows begin with an authentication step requiring the agent to correctly capture entities such as names, confirmation codes, and account IDs—if the user's speech corrupts these, the agent cannot proceed regardless of its own capability.

\textbf{Implementation.}
We adapt the same prompt used to evaluate agent speech fidelity (Section~\ref{app:prompt-speech-fidelity}), with one key modification to the rating scale. Unlike for the agent, we do not require the user's speech to precisely mirror every word of the user-side LLM output; only key entities and major informational content must be conveyed accurately. The judge therefore rates each turn on a 3-point scale: 3 indicates full fidelity, 2 indicates minor errors that do not affect the agent's ability to progress (e.g., slight disfluencies, but all entities intact), and 1 indicates an entity error or significant omission or addition that would prevent the conversation from proceeding sensibly. Any conversation in which any turn receives a rating of 1 is rerun. The full prompt can be found in \autoref{app:prompt-user-speech-fidelity}.

\textbf{Validation.}
Because this prompt is closely derived from the agent speech fidelity judge, we inherit its validation. That judge achieved high inter-annotator agreement with human linguists ($\kappa =  0.777$, 95\% CI [0.704, 0.835]), and the core capability it requires—accurately parsing audio and detecting entity-level errors—is shared. The rating scale and its interpretation are sufficiently well-defined that additional annotation studies were not deemed necessary. See table \ref{tab:judge-agreement} for more details on human-judge agreement.


\input{tables/08_appendix/d_user_simulator/04_distribution_corruption_types}