
\section{Additional Real-World Experimental Results}
\label{sec:appendix_additional_realworld}

\subsection{Additional Quantitative Results for Real-World Experiments}

We further evaluate whether skills learned from MolmoSpaces play can transfer to additional
real-world manipulation tasks. We add two real-world tasks that require object rearrangement and
articulated-object interaction. In \textit{Swap Cubes}, the robot must pick up the cube on the
platform, move it off the platform, and exchange it with the cube initially placed below the platform.
In \textit{Close Drawer}, the robot must close an initially open drawer. For both tasks, we compare
the standard \textsc{CaP-Agent0} system against \textsc{CaP-Agent0} augmented with skills learned
by \textsc{RATs} during MolmoSpaces play.

Table~\ref{tab:additional-real-world} shows that adding MolmoSpaces play-learned skills improves
performance on both tasks. On \textit{Swap Cubes}, the standard \textsc{CaP-Agent0} baseline fails
to solve the task, while the skill-augmented agent succeeds in 7 out of 30 trials. On
\textit{Close Drawer}, the success rate improves from 2/30 to 8/30. Averaged over the two tasks,
MolmoSpaces skills improve real-world success from 3.3\% to 25.0\%, corresponding to a
+21.7 percentage-point gain. These results suggest that play-learned skills from MolmoSpaces can
provide useful reusable behaviors for real-world manipulation beyond the original benchmark tasks.

\begin{table}[ht]
    \centering
    \setlength{\tabcolsep}{13pt}
    \caption{\textbf{Additional real-world evaluation with MolmoSpaces skills.}
    Skills are learned by \textsc{RATs} during MolmoSpaces play and reused with the
    \textsc{CaP-Agent0} system. }
    \label{tab:additional-real-world}
    \resizebox{\robosuiteTableWidth}{!}{%
    \begin{tabular}{lccc}
        \toprule
        Task
        & \textsc{CaP-Agent0}
        & \textsc{CaP-Agent0} + \textsc{RATs} Skills (MolmoSpaces)
        & $\Delta$ \\
        \midrule
        Swap Cubes
            & 0/30 (0.0\%)
            & \textbf{7/30 (23.3\%)}
            & +23.3 pp \\
        Close Drawer
            & 2/30 (6.7\%)
            & \textbf{8/30 (26.7\%)}
            & +20.0 pp \\
        \midrule
        \textbf{Average (Real World)}
            & 2/60 (3.3\%)
            & \textbf{15/60 (25.0\%)}
            & +21.7 pp \\
        \bottomrule
    \end{tabular}
    \vspace{-2em}
    }
\end{table}



\subsection{Additional Qualitative Results}
\label{sec:appendix_additional_qualitative}

We also provide additional qualitative examples of successful real-world executions using
\textsc{CaP-Agent0} augmented with \textsc{RATs} skills learned from MolmoSpaces play. As shown
in Figure~\ref{fig:real}, the skill-augmented agent successfully solves
the newly added \textit{Swap Cubes} and \textit{Close Drawer} tasks, as well as an \textit{Open Drawer}
task. Specifically, in \textit{Swap Cubes}, the robot exchanges a cube on the platform with a cube 
below the platform; in \textit{Close Drawer}, the robot closes an initially open drawer; and in 
\textit{Open Drawer}, the robot opens a closed drawer. These examples illustrate that the MolmoSpaces 
skill library can support both object rearrangement and articulated-object manipulation in the real world. 
We include video results of these real-world rollouts, along with additional simulation rollouts, 
on the project webpage.




