\section{Details about the Evaluation Benchmark}
\label{sec:appendix_evaluation_benchmarks}

During evaluation, the play-time skill library is frozen and the system solves
externally specified tasks. This section records the benchmark composition and
replay protocol for the two simulated evaluation domains.

\subsection{MolmoSpaces Evaluation Benchmark}
\label{sec:appendix_molmospaces_benchmark}

For MolmoSpaces, we evaluate on a compact held-out subset of the
benchmark. The subset contains
40 test-split episodes: ten each for opening, closing, picking, and
pick-and-place. It was constructed by randomly selecting ten
episodes per task type. It covers four components of the upstream
benchmark, namely Open-v1, Close-v1, Pick-v2-classic, and PnP-v2.

\begin{table}[t]
\centering
\caption{
\textbf{MolmoSpaces held-out core evaluation subset.}
The main benchmark sweep repeats each selected task ten times.
}
\label{tab:molmospaces_eval_benchmark}
\begin{tabular}{llrr}
\toprule
\textbf{Task type} & \textbf{Scene dataset} & \textbf{Tasks} & \textbf{Main rollouts} \\
\midrule
Open & iTHOR & 10 & 100 \\
Close & iTHOR & 10 & 100 \\
Pick & ProcTHOR-Objaverse & 10 & 100 \\
pick-and-place & ProcTHOR-Objaverse & 10 & 100 \\
\midrule
Total & -- & 40 & 400 \\
\bottomrule
\end{tabular}

\end{table}

The selected episodes span 37 houses. Each episode's JSON stores the scene
dataset and house index, the Franka initialization, object poses, task state,
natural-language referral expressions, and camera specification. Success is evaluated by the simulator task's \texttt{judge\_success()} predicate: articulated tasks check the requested
joint state, pick tasks, check the object target state, and pick-and-place tasks
check placement support on the receptacle.

The benchmark definition is independent of the number of repeated trials. The
main sweep uses ten trials per task, for $40 \times 10 = 400$ rollouts.

\subsection{LIBERO-PRO Evaluation Benchmark}
\label{sec:appendix_libero_pro_benchmark}

For LIBERO-PRO, we evaluate the object, goal, and spatial
categories under two perturbation types. The \emph{Pos} setting corresponds to
the \texttt{swap} suite, which perturbs object positions. The \emph{Task}
setting corresponds to the \texttt{task} suite, which perturbs the task
specification. Each suite contains ten language-conditioned manipulation
tasks.

\begin{table}[t]
\centering
\caption{
\textbf{LIBERO-PRO evaluation.}
Reported comparisons use ten initial-state trials per task.
}
\label{tab:libero_pro_eval_benchmark}
\begin{tabular}{llrr}
\toprule
\textbf{Setting} & \textbf{LIBERO-PRO suite} & \textbf{Tasks} & \textbf{Reported rollouts} \\
\midrule
Object Pos & \texttt{libero\_object\_swap} & 10 & 100 \\
Object Task & \texttt{libero\_object\_task} & 10 & 100 \\
Goal Pos & \texttt{libero\_goal\_swap} & 10 & 100 \\
Goal Task & \texttt{libero\_goal\_task} & 10 & 100 \\
Spatial Pos & \texttt{libero\_spatial\_swap} & 10 & 100 \\
Spatial Task & \texttt{libero\_spatial\_task} & 10 & 100 \\
\midrule
Total & -- & 60 & 600 \\
\bottomrule
\end{tabular}

\end{table}

LIBERO-PRO provides 50 initial states for each task. A complete sweep over all
available states therefore contains $6 \times 10 \times 50 = 3{,}000$
rollouts. For the reported comparisons, we cap evaluation at ten
initial-state trials per task, yielding $6 \times 10 \times 10 = 600$ rollouts
for each evaluated setup. The baseline and learned-library conditions use the
same task and trial budget.
