


    










\section{Additional Studies}
\label{sec:appendix_additional_studies}

\subsection{Ablation Results on MolmoSpaces}

We further ablate the effect of play-learned skills and the test-time execution system on
MolmoSpaces. This setting complements the LIBERO-PRO ablation in the main paper by evaluating
whether the same play-time skill acquisition mechanism remains useful in a scene-grounded
benchmark with open, close, pick, and pick-and-place task categories. As in the main ablation, all
play-based variants use 50 play iterations, and all play-time skills are learned by our proposed
\textsc{RATs} system.

Table~\ref{tab:molmospaces-ablation} shows that play-learned skills improve performance even when
they are only plugged into the standard \textsc{CaP-Agent0} test-time agent. Under \textsc{CaP-Agent0},
Curious Play increases the average success rate from 21.0\% to 25.8\%, with the largest gain on
closing tasks. When the same learned skills are used by the full \textsc{RATs} execution system,
performance improves more substantially, from 32.8\% without play to 38.0\% with Curious Play.
These results suggest that play-time skill learning and the structured \textsc{RATs} test-time
execution system provide complementary benefits: the learned skill library alone can improve a
standard Code-as-Policy agent, while the full execution system is better able to retrieve, verify, and
compose those skills for downstream tasks.

\begin{table}[t]
    \centering
    \setlength{\tabcolsep}{3.5pt}
    \caption{\textbf{MolmoSpaces ablation over play strategy and test-time system.} All
    play-based variants use 50 play iterations. All play-time skills are learned by our proposed \textsc{RATs} system.}
    \label{tab:molmospaces-ablation}
    \resizebox{\ablationTableWidth}{!}{%
    \begin{tabular}{llccccc}
        \toprule
        Test-Time System
        & Play-Time Skills
        & Open
        & Close
        & Pick
        & pick-and-place
        & Avg. \\
        \midrule
        \multirow{2}{*}{\textsc{CaP-Agent0}}
            & No Play
             & 14.0 & 36.0 & 23.0 & 11.0 & 21.0 \\
            & Curious Play
             & 17.0 & 62.0 & 14.0 & 10.0 & 25.8 \\
        \midrule
        \multirow{2}{*}{\textsc{RATs Exec.}}
            & No Play
             & 11.0 & 65.0 & \textbf{45.0} & 10.0 & 32.8 \\
            & Curious Play
             & \textbf{20.0} & \textbf{73.0} & 37.0 & \textbf{22.0} & \textbf{38.0} \\
        \bottomrule
    \end{tabular}
    }
\end{table}


\subsection{Token Cost Analysis}
\label{sec:token_cost_analysis}

Because \textsc{RATs} relies heavily on Large Language Models (LLMs) for task proposal, planning,
code generation, and failure diagnosis, we provide a detailed analysis of the token consumption
during both the autonomous play phase and the test-time evaluation phase. All LLM calls documented
in this section utilized \texttt{gpt-5.5}.

\textbf{Play-Time Token Consumption.} We analyzed a play-mode run in the
\texttt{libero\_spatial} environment. Over the play iterations, where each iteration is budgeted with
a maximum of five attempts, token consumption is heavily skewed toward failed iterations that
exhaust the retry budget. A component-wise breakdown in Table~\ref{tab:play_time_tokens} shows
that the Write-Execute-Verify-Diagnose loop is the primary cost driver. This indicates that while
intrinsic task proposal is relatively token-efficient, extracting diagnostics and revising policies from
repeated physical failures remains the dominant computational expense during play.

\begin{table}[htbp]
    \centering
    \caption{\textbf{Play-time token consumption by each agent component (10 iterations).}}
    \label{tab:play_time_tokens}
    \begin{tabular}{lrr}
        \toprule
        \textbf{Agent Role} & \textbf{Total Tokens} & \textbf{Share (\%)} \\
        \midrule
        Failure Diagnoser & 2,071,881 & 40.5\% \\
        Policy Writer & 1,472,949 & 28.8\% \\
        Failure Memory Distillation & 993,770 & 19.4\% \\
        Verifier & 194,318 & 3.8\% \\
        Memory Curator & 111,563 & 2.2\% \\
        Task Proposer & 98,505 & 1.9\% \\
        Planner & 78,359 & 1.5\% \\
        Environment Creator & 31,616 & 0.6\% \\
        Skill Library (Duplicate Check) & 31,377 & 0.6\% \\
        Skill Proposer & 30,969 & 0.6\% \\
        Feedback Generator & 5,875 & 0.1\% \\
        \midrule
        \textbf{Total} & \textbf{5,121,182} & \textbf{100.0\%} \\
        \bottomrule
    \end{tabular}
\end{table}

\textbf{Compute-Matched Test-Time Comparison.} To ensure a fair comparison, we account for the
upfront token cost incurred by \textsc{RATs} during autonomous play. A full 50-iteration play phase
consumes approximately 30M tokens. We amortize this play-time cost across the 60 held-out tasks in
LIBERO-PRO.

The baseline \textsc{CaP-Agent0} consumes approximately 1.6M tokens per task under a standard
10-turn retry budget. Adding the amortized play-time token cost of \textsc{RATs} to this baseline
grants \textsc{CaP-Agent0} a larger test-time compute budget, allowing it to run for roughly 15 turns
per task. We therefore evaluate a compute-matched baseline, \textsc{CaP-Agent0} with a 15-turn
budget, and compare it against the standard 10-turn \textsc{CaP-Agent0} agent augmented with the
skill library learned by \textsc{RATs} through Curious Play. This comparison isolates whether the
same additional token budget is more useful when spent reactively on extra test-time retries or
proactively on play-time skill acquisition.

\begin{table}[htbp]
    \centering
    \caption{\textbf{Compute-matched performance comparison on LIBERO-PRO.} The
    \textsc{CaP-Agent0} (15 turns) baseline is granted additional retry budget at test time to match
    the amortized token cost that \textsc{RATs} spent during play-time. The \textsc{CaP-Agent0} +
    \textsc{RATs} Skills row uses the standard 10-turn \textsc{CaP-Agent0} test-time system, with
    skills learned from 50 iterations of Curious Play.}
    \label{tab:compute_matched_comparison}
    \resizebox{\textwidth}{!}{%
    \begin{tabular}{lccccccr}
        \toprule
        \multirow{2}{*}{\textbf{Method}} & \multicolumn{2}{c}{\textbf{Object}} & \multicolumn{2}{c}{\textbf{Goal}} & \multicolumn{2}{c}{\textbf{Spatial}} & \multirow{2}{*}{\textbf{Avg.}} \\
        \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
        & Pos. & Task & Pos. & Task & Pos. & Task & \\
        \midrule
        \textsc{CaP-Agent0} (10 turns) & 27.0 & 31.0 & 29.0 & 16.0 & 13.0 & 23.0 & 23.2 \\
        \textsc{CaP-Agent0} (15 turns, compute-matched) & 28.0 & 32.0 & 30.0 & \textbf{24.0} & \textbf{22.0} & 20.0 & 26.0 \\
        \textsc{CaP-Agent0} (10 turns) + \textsc{RATs} Skills from Play
        & \textbf{51.0} & \textbf{47.0} & \textbf{34.0} & 20.0 & 19.0 & \textbf{23.0} & \textbf{32.3} \\
        \bottomrule
    \end{tabular}%
    }
\end{table}

As shown in Table~\ref{tab:compute_matched_comparison}, simply allocating more test-time compute
to the baseline yields only modest improvements, increasing average success from 23.2\% to 26.0\%.
In contrast, spending the comparable compute budget proactively during play-time and then reusing
the learned skills with the same 10-turn \textsc{CaP-Agent0} test-time system improves average
success to 32.3\%. This result indicates that the gain does not come merely from giving the agent
more inference-time retries. Instead, play-time computation is more effective when it is distilled into
a reusable skill library that can be retrieved and composed for held-out tasks.
