% ========================================================
% -------------------- Contributors ----------------------
% ========================================================
\section{Author List}\label{sec:contributors}
\setnavsection{sec:contributors}


\subsection{Core Contributors} 
Yihao~Wang\textsuperscript{*}, \
Yuheng~Ji\textsuperscript{*,$\dagger$}, \ 
Mingyu~Cao\textsuperscript{*}, \
Yanqing~Shen\textsuperscript{*}, \
Runze~Xiao\textsuperscript{*}


\paragraph{Model Pre-Training.} 
Huaihai~Lyu, \
Senwei~Xie, \
Mingyu~Cao\textsuperscript{*}


\paragraph{Data Infra.}
Euan~Liu, \
Klara~Tian, \
Tianfeng~Long, \
Yichi~Zhang, \
Zhengliang~Cai,\
Ruike~Chen,  \
Jifan~Zhao, Yanqing~Shen\textsuperscript{*} 


\paragraph{Evaluation.} 
Ruochuan~Shi, \
Zihan~Tang, \
Jing~Lyu, \
Runze~Xiao\textsuperscript{*}

\paragraph{Real Robot.}
Jing~Lyu, \
Wenxing~Tan, \
Ningbo~Zhang,\ 
Yangtao~Hu,\
Euan~Liu,\
Yuming~Gao,\
Xiansheng~Chen, \
Junkai~Zhao,\
Runze~Xiao\textsuperscript{*}

\paragraph{Downstream Post-Training.}
Senwei~Xie, \
Huaihai~Lyu, \
Congsheng~Xu,\
Boan~Zhu,\
Ziqi~Wang

\paragraph{Infrastructure.} 
Yupu~Feng, \
Qiongqiong~Zhang


\subsection{Contributors} 

\paragraph{Infrastructure.}
Yingli~Zhao,\
Yulong~Ao

\paragraph{Real-Robot Data.}
Shaoxuan~Xie, \
You~Liu, \
Guocai~Yao

\paragraph{Product \& Operations.}
Leiduo~Zhang, \
Xiaodan~Liu, \
Yunyan~Zhang, \
Yance~Jiao

\paragraph{Brand Management.}
Xinyan~Yang,\
Jiaxing~Wei

\paragraph{Platform Management.}
Xu~Liu,\
Tengfei~Pan

\paragraph{System Management.}
Shaokai~Nie,\
Chunlei~Men


\subsection{Expert Consultant (Ordered by English last name alphabetically)}
Sen~Cui, \
Xiaojie~Jin, \
Hongyang~Li, \
Jianlan~Luo, \
Yao~Mu, \
Yunchao~Wei, \
Jun~Yan, \
Hang~Zhao, \
Xiaolong~Zheng


\subsection{Research Leads}
Jiaming~Li, \ 
Yonghua~Lin, \ 
Tiejun~Huang, \ 
Zhongyuan~Wang\textsuperscript{\Letter}, \ 
Pengwei~Wang\textsuperscript{\Letter}

\vspace{5ex}


\makeatletter
\begingroup
  \renewcommand{\thefootnote}{}
  \renewcommand{\@makefntext}[1]{\noindent #1}
  \footnotetext{%
    \small
    \textsuperscript{*} Equal Contribution.
    \par
    \textsuperscript{$\dagger$} Project Lead.\par
    \textsuperscript{\Letter} Corresponding Authors: zhongyuan@baai.ac.cn, pwwang@baai.ac.cn}
\endgroup
\makeatother
