\section{Introduction}
\label{sec:introd}

Recent progress in LLMs~\citep{kimi2.6, GPT5.5, claude-4.6, gemini-3-pro, deepseekai2026deepseekv4, Qwen3.5} is rapidly pushing AI from passive language models toward autonomous agents that can plan, use tools, interact with environments, and improve through feedback. In real-world scenarios such as software engineering~\citep{mlebench}, scientific research~\citep{frontierscience, phan2025humanity}, and complicated decision making~\citep{browsecomp}, agents must operate over long horizons: they need to acquire information, decompose tasks, call tools, verify intermediate results, and continuously adjust their strategies. Such long-horizon settings are especially challenging since early mistakes can accumulate and strategies often need to be revised as new external information becomes available.

Existing efforts to improve long-horizon agents \cite{zeng2026glm,du2026towards,GLM5.2,team2026kimi,sun2025scaling,copet2025cwm} broadly follow two scaling routes. One route \cite{GLM5.2,team2026kimi,claude-4.6} scales parameters: frontier models commonly rely on scaling model parameters to internalize a wide range of reasoning patterns, tool-use behavior patterns, and domain knowledge. This route is effective, but it largely makes it difficult to reproduce the same agentic competence without comparable model scale, data, and training resources. The other route scales the horizon rather than enlarging parameters, which makes the intermediate decision process explicit, turning knowledge acquisition, action execution, observation interpretation, and verification into trainable supervision. However, this route also exposes two key bottlenecks.

A key bottleneck lies in the knowledge infrastructure required to support the scaling of long-horizon trajectories. Long-horizon trajectory training requires a unified environment that connects external knowledge, agentic actions, observations, and verification signals, enabling models to learn from grounded feedback rather than isolated text supervision. Without such a knowledge-tool interaction infrastructure, agents can hardly acquire the ability to plan over extended trajectories, invoke tools appropriately, verify evidence, incorporate feedback, and recover from failures in realistic settings.

Beyond infrastructure, scaling the horizon also requires integrating a broad set of heterogeneous and compositional abilities across domains. These abilities include multi-step information retrieval, tool use, executable iteration, constraint tracking, and result reflection. They often emerge unevenly across domains and interact with each other in complex ways. As a result, it is challenging to effectively combine highly different specialized abilities into one unified agentic model.

In this work, we introduce \ProjectName, a 35B MoE agentic model designed to address the key challenges mentioned above. To support agentic model training, we build a knowledge-action infrastructure that connects external knowledge, intermediate actions, observations, execution results, and verification signals, producing agentic trajectories with an average length of 45K tokens. Based on this infrastructure, we train \ProjectName~using a three-stage training recipe. First, we perform full-domain supervised fine-tuning to obtain a general agentic model with broad long-horizon abilities. Next, we train domain-level teacher models to achieve improvements in specialized domains. Finally, we propose a domain-routed on-policy distillation (OPD) with salient vocabulary alignment to unify capabilities from six heterogeneous domains into a single deployable student model.

It also aims to provide the community with a scalable technical path from model-parameter scaling to agent-horizon scaling. As shown in Fig.~\ref{fig:benchmark_performance}, \ProjectName~outperforms 1T-parameter models (Kimi-K2.6~\citep{kimi2.6} and DeepSeek-V4~\citep{deepseekai2026deepseekv4}) on SEAL-0~\citep{sealqa}, IFBench~\citep{IFBench}, HiPhO~\citep{hipho}, FrontierScience-Olympiad~\citep{frontierscience}, and MolBench-Bind~\citep{zhang2026molclaw}. It also achieves strong results on SciCode~\citep{scicode}, HLE~\citep{phan2025humanity}, and BrowseComp~\citep{browsecomp}. Our contributions include the following three aspects:
\begin{itemize}
    \item We present \ProjectName, 35B MoE model designed to scale heterogeneous agentic abilities across multiple domains. \ProjectName~can match or even outperform 1T-parameter models in long-horizon interactive agent capabilities for science and research.
    \item We build a Long-Horizon Knowledge-Action Infrastructure for multi-turn interaction with external knowledge and tools in long-horizon tasks. This infrastructure improves the agent’s ability to obtain useful information, summarize key issues, call the right tools, and execute and verify tasks.
    \item We propose a Domain-Routed On-Policy Distillation with Salient Vocabulary Alignmen to reduce conflicts caused by different reasoning patterns across domains when scaling the horizon.
\end{itemize}
