\section{Full Schema of {\ours}}
\label{app:schema}
%entities
\begin{longtable}{p{2.8cm}p{5.5cm}p{3cm}}
\caption{Node types and attributes in the Neo4j schema.}
\label{tab:node-attributes} \\

\toprule
Node Type & Attribute & Type \\
\midrule
\endfirsthead

\toprule
Node Type & Attribute & Type \\
\midrule
\endhead

\bottomrule
\endfoot

% ===== Author =====
Author & id & ID \\
Author & label & string \\
Author & display\_name & string \\
Author & orcid & string \\
Author & works\_count & int \\
Author & cited\_by\_count & int \\
Author & h\_index & int \\
Author & i10\_index & int \\
Author & mean\_citedness\_2y & float \\
Author & created\_date & string \\
Author & updated\_date & string \\

% ===== Domain =====
Domain & id & ID \\
Domain & label & string \\
Domain & display\_name & string \\
Domain & description & string \\
Domain & works\_count & int \\
Domain & cited\_by\_count & int \\
Domain & created\_date & string \\
Domain & updated\_date & string \\

% ===== Field =====
Field & id & ID \\
Field & label & string \\
Field & display\_name & string \\
Field & description & string \\
Field & works\_count & int \\
Field & cited\_by\_count & int \\
Field & created\_date & string \\
Field & updated\_date & string \\

% ===== Institution =====
Institution & id & ID \\
Institution & label & string \\
Institution & display\_name & string \\
Institution & ror & string \\
Institution & country\_code & string \\
Institution & country & string \\
Institution & city & string \\
Institution & type & string \\
Institution & works\_count & int \\
Institution & cited\_by\_count & int \\
Institution & h\_index & int \\
Institution & homepage\_url & string \\
Institution & created\_date & string \\
Institution & updated\_date & string \\

% ===== Keyword =====
Keyword & id & ID \\
Keyword & label & string \\
Keyword & text & string \\
Keyword & text\_normalized & string \\
Keyword & frequency & int \\
Keyword & text\_embedding & float[] \\

% ===== Paper（续）=====
Paper & created\_date & string \\
Paper & updated\_date & string \\
Paper & pdf\_url & string \\
Paper & pdf\_source\_id & string \\
Paper & pdf\_source\_display\_name & string \\
Paper & pdf\_source\_type & string \\
Paper & pdf\_is\_oa & boolean \\
Paper & pdf\_is\_published & boolean \\
Paper & pdf\_version & string \\
Paper & venue\_source\_id & string \\
Paper & venue\_source\_display\_name & string \\
Paper & venue\_source\_type & string \\
Paper & venue\_raw\_source\_name & string \\

% ===== Source =====
Source & id & ID \\
Source & label & string \\
Source & display\_name & string \\
Source & type & string \\
Source & issn\_l & string \\
Source & is\_oa & boolean \\
Source & is\_core & boolean \\
Source & works\_count & int \\
Source & cited\_by\_count & int \\
Source & created\_date & string \\
Source & updated\_date & string \\

% ===== Subfield =====
Subfield & id & ID \\
Subfield & label & string \\
Subfield & display\_name & string \\
Subfield & description & string \\
Subfield & works\_count & int \\
Subfield & cited\_by\_count & int \\
Subfield & created\_date & string \\
Subfield & updated\_date & string \\

% ===== Topic =====
Topic & id & ID \\
Topic & label & string \\
Topic & display\_name & string \\
Topic & description & string \\
Topic & keywords & string[] \\
Topic & works\_count & int \\
Topic & cited\_by\_count & int \\
Topic & domain\_id & string \\
Topic & field\_id & string \\
Topic & subfield\_id & string \\
Topic & created\_date & string \\
Topic & updated\_date & string \\

\end{longtable}

% relations
\begin{longtable}{p{3cm}p{2.5cm}p{2.5cm}p{5cm}}
\caption{Relationship types in the Neo4j schema.}
\label{tab:relationships} \\

\toprule
Type & Source & Target & {\raggedright Properties\par} \\
\midrule
\endfirsthead

\toprule
Type & Source & Target & {\raggedright Properties\par} \\
\midrule
\endhead

\bottomrule
\endfoot

AFFILIATED\_WITH & Author & Institution & {\raggedright is\_current (boolean)\par} \\
AUTHORED & Author & Paper & {\raggedright position (int), is\_corresponding (boolean), raw\_name (string)\par} \\
CITES & Paper & Paper & none \\
COAUTHOR & Author & Author & count (int) \\
COOCCUR & Keyword & Keyword & count (int) \\
DOMAIN\_OF & Field & Domain & none \\
FIELD\_OF & Subfield & Field & none \\
HAS\_KEYWORD & Paper & Keyword & {\raggedright relevance\_score (float)\par} \\
HAS\_TOPIC & Paper & Topic & {\raggedright score (float), is\_primary (boolean)\par} \\
RELATED\_TO & Paper & Paper & none \\
SUBFIELD\_OF & Topic & Subfield & none \\

\end{longtable}


{
\begin{longtable}{@{}p{5.6cm}p{1.7cm}p{5.8cm}@{}}
\caption{Indexes in the Neo4j schema.}
\label{tab:indexes-general} \\

\toprule
{\raggedright Index Name\par} & {\raggedright Type\par} & {\raggedright Definition\par} \\
\midrule
\endfirsthead

\toprule
{\raggedright Index Name\par} & {\raggedright Type\par} & {\raggedright Definition\par} \\
\midrule
\endhead

\bottomrule
\endfoot

{\raggedright paper\_\allowbreak title\_\allowbreak normalized\_\allowbreak idx\par} & RANGE & {\raggedright :Paper(title\_normalized)\par} \\
{\raggedright paper\_\allowbreak text\_\allowbreak ft\par} & FULLTEXT & {\raggedright :Paper(title, abstract)\par} \\
{\raggedright paper\_\allowbreak title\_\allowbreak ft\par} & FULLTEXT & {\raggedright :Paper(title)\par} \\
{\raggedright paper\_\allowbreak abstract\_\allowbreak ft\par} & FULLTEXT & {\raggedright :Paper(abstract)\par} \\
{\raggedright keyword\_\allowbreak text\_\allowbreak ft\par} & FULLTEXT & {\raggedright :Keyword(text, text\_normalized)\par} \\

\end{longtable}
}

% vector index
{
\begin{longtable}{@{}p{5.5cm}p{1.5cm}p{4.8cm}@{}}
\caption{Vector indexes in the Neo4j schema.}
\label{tab:indexes-vector} \\

\toprule
{\raggedright Index Name\par} & Node & {\raggedright Configuration\par} \\
\midrule
\endfirsthead

\toprule
{\raggedright Index Name\par} & Node & {\raggedright Configuration\par} \\
\midrule
\endhead

\bottomrule
\endfoot

{\raggedright paper\_\allowbreak title\_\allowbreak embedding\_\allowbreak idx\par} & Paper &
{\raggedright dimensions=1024, similarity=COSINE\par} \\

{\raggedright paper\_\allowbreak abstract\_\allowbreak embedding\_\allowbreak idx\par} & Paper &
{\raggedright dimensions=1024, similarity=COSINE\par} \\

{\raggedright keyword\_\allowbreak text\_\allowbreak embedding\_\allowbreak idx\par} & Keyword &
{\raggedright dimensions=1024, similarity=cosine\par} \\

\end{longtable}
}

\section{Prompts Used in this Report}

\subsection{Keyword Extraction}
\label{app:keyword_prompt}
\begin{tcolorbox}[breakable, colback=black!5!white, colframe=black!60!black,
  left=1ex, right=1ex, top=0.5ex, bottom=0.5ex,
  fontupper=\small\ttfamily, fontlower=\small\ttfamily,
  title=\textbf{Keyword Extraction}]
\textbf{System Prompt:}
You are an expert assistant that extracts high-level academic keywords for knowledge graph construction.

\medskip\textbf{User Prompt:}
\textbf{Goal:}
From an academic paper abstract, extract a small set of canonical, high-level keywords that represent the main research topics, tasks, or method categories of the paper.

These keywords will be used as entities in a knowledge graph to connect related papers across many scientific and engineering fields.

\textbf{Important:}
The goal is NOT to capture detailed phrases from the abstract.
The goal is to identify general, reusable concepts that are likely to appear in many different papers.

\textbf{Requirements:}
\begin{itemize}\small
\item Extract only 3--8 keywords.
\item Prefer high-level research areas, problem types, method families, or evaluation paradigms.
\item Normalize detailed wording into broader concepts when possible.
\item Keywords should be reusable across many papers and suitable as knowledge graph entities.
\item Use concise noun phrases (typically 1--4 words).
\item If the abstract does not contain many strong high-level concepts, return fewer keywords.
\item Also score each keyword's relevance to the abstract on a 1-10 integer scale.
\end{itemize}

\textbf{Avoid:}
\begin{itemize}\small
\item Long descriptive phrases copied from the abstract
\item Paper-specific terminology or system names
\item Highly customized or marketing-style expressions
\item Implementation details or narrow technical descriptions
\end{itemize}

\textbf{Good keyword examples} (general, reusable concepts):
\begin{itemize}\small
\item machine learning
\item wireless communication
\item computer vision
\item protein structure prediction
\item finite element method
\item Monte Carlo simulation
\item graph neural networks
\item energy optimization
\item fault detection
\item numerical simulation
\end{itemize}

\textbf{Bad keyword examples} (paper-specific phrasing):
\begin{itemize}\small
\item hierarchical dual-path adaptive learning framework
\item multi-stage cross-modal feature fusion architecture
\item lightweight energy-aware dynamic routing mechanism
\item novel high-performance prototype system
\item end-to-end task-specific optimization pipeline
\end{itemize}

\textbf{Abstract:}
\{abstract\}

\textbf{Output format:}
Return ONLY a JSON object:
\begin{verbatim}
{"keywords": ["keyword1", "keyword2"], "scores": [8, 7]}.
\end{verbatim}
\texttt{scores} must be integers from 1 to 10, aligned with \texttt{keywords}, with no extra text.
\end{tcolorbox}

\subsection{Downstream Tasks}
\label{app:prompt_downstream}

\subsubsection{Idea Grounding -- Query Generation}
\label{prompt:task1-query-gen}

\begin{tcolorbox}[breakable, colback=blue!5!white, colframe=blue!60!black,
  left=1ex, right=1ex, top=0.5ex, bottom=0.5ex,
  fontupper=\small\ttfamily, fontlower=\small\ttfamily,
  title=\textbf{Idea Grounding -- Query Generation:} Generate dense-retrieval queries from structured scientific extraction results.]
\textbf{System Prompt:}
You are a scientific grounding agent.
Generate retrieval-oriented paragraph search queries from structured scientific extraction results.
Return strict JSON only.

\medskip\textbf{User Prompt:}
Generate dense-retrieval queries from the structured extraction below.

Requirements:
\begin{itemize}\small
\item Use only the \texttt{motivation} and \texttt{method} sections as query sources.
\item Consider all provided sentences from those two sections before deciding which queries to emit.
\item Select the most retrieval-useful items yourself. Do not mirror every input sentence if some are redundant.
\item Produce at most \texttt{max\_queries} total queries across both sections combined.
\item You may allocate the total freely across \texttt{motivation} and \texttt{method}.
\item Each output item must contain:
  \begin{itemize}\small
  \item \texttt{section}: either \texttt{motivation} or \texttt{method}
  \item \texttt{sentence}: the source sentence you selected
  \item \texttt{query}: the final retrieval-oriented rewrite
  \end{itemize}
\item Keep the selected sentence meaning exactly.
\item Write concise academic retrieval phrases or sentences likely to match paper paragraphs.
\item Preserve task, object, method, mechanism, training objective, dataset, baseline, metric, or analysis anchors when present.
\item Avoid vague wording such as ``the framework'', ``this method'', ``evaluation'', ``performance improvement'', or ``how it works''.
\item Do not introduce unsupported facts.
\end{itemize}

Return JSON with this schema only:
\begin{verbatim}
{
  "items": [
    {
      "section": "motivation | method",
      "sentence": "selected source sentence",
      "query": "retrieval-oriented rewrite optimized for paragraph retrieval"
    }
  ]
}
\end{verbatim}

Structured extraction:
\{items\_json\}
\end{tcolorbox}


\subsubsection{Idea Grounding -- Grounding}
\label{prompt:task1-refinement}

\begin{tcolorbox}[breakable, colback=green!5!white, colframe=green!60!black,
  left=1ex, right=1ex, top=0.5ex, bottom=0.5ex,
  fontupper=\small\ttfamily, fontlower=\small\ttfamily,
  title=\textbf{Idea Grounding -- Grounding:} Analyze how retrieved evidence aligns with a research idea unit.]
\textbf{System Prompt:}
You are a scientific grounding alignment agent.
Given a research idea unit, a retrieved paragraph, and local paper context, analyze how the retrieved evidence aligns with the target idea.
Use only the provided evidence. Return strict JSON only.

\medskip\textbf{User Prompt:}
Analyze how the retrieved evidence aligns with a specific research idea unit.

Research idea:
\{idea\_text\}

Target query section:
\{query\_section\}

Target query sentence:
\{query\_sentence\}

Retrieval query:
\{query\_text\}

Paper title:
\{paper\_title\}

Paper abstract:
\{paper\_abstract\}

Section path:
\{section\_path\_text\}

Retrieved paragraph:
\{paragraph\_text\}

Previous paragraphs:
\{previous\_paragraphs\}

Next paragraphs:
\{next\_paragraphs\}

Requirements:
\begin{itemize}\small
\item Judge whether the retrieved evidence supports the target query unit.
\item Use the paragraph as the primary evidence and use the surrounding context only to clarify scope, subject, or omitted details.
\item Do not invent claims that are not grounded in the provided paragraph or context.
\item If the paragraph is weakly related, say so explicitly.
\item \texttt{focus\_aspect} should state which specific aspect of the target query is actually addressed by the evidence.
\item \texttt{grounded\_passage} should be a concise evidence-focused passage, usually 1--3 sentences, that is more context-aware and better aligned to the target query than the raw paragraph alone.
\item \texttt{evidence\_span} should quote or closely paraphrase the most directly relevant span from the retrieved paragraph or immediate context.
\item \texttt{shared\_points} should list concrete aspects that are aligned between the target idea unit and the evidence.
\item \texttt{different\_points} should list concrete mismatches, missing parts, narrower scope, or different emphasis between the evidence and the target idea unit.
\item \texttt{coverage\_label} must be one of: \texttt{high}, \texttt{partial}, \texttt{limited}, \texttt{none}.
  \begin{itemize}\small
  \item Use \texttt{high} only when the evidence covers most of the target idea unit.
  \item Use \texttt{partial} when there is clear overlap but also important missing coverage.
  \item Use \texttt{limited} when the evidence is only weakly related or covers a small sub-aspect.
  \item Use \texttt{none} when it is essentially irrelevant.
  \end{itemize}
\item \texttt{why\_this\_matches} should briefly explain the overall judgment.
\end{itemize}

Return JSON with this schema only:
\begin{verbatim}
{
  "status": "supported | partially_supported | weak_match | irrelevant",
  "focus_aspect": "which specific aspect of the target query is grounded",
  "grounded_passage": "context-aware grounding passage",
  "evidence_span": "most relevant evidence span",
  "shared_points": ["..."],
  "different_points": ["..."],
  "coverage_label": "high | partial | limited | none",
  "why_this_matches": "brief explanation of the match quality"
}
\end{verbatim}
\end{tcolorbox}


\subsubsection{Idea Generation}
\label{prompt:task2-idea-gen}

\begin{tcolorbox}[breakable, colback=magenta!5!white, colframe=magenta!60!black,
  left=1ex, right=1ex, top=0.5ex, bottom=0.5ex,
  fontupper=\small\ttfamily, fontlower=\small\ttfamily,
  title=\textbf{Idea Generation:} Propose novel research ideas from a collection of papers.]
\textbf{System Prompt:}
You are an expert research idea generator and only return valid JSON.

\medskip\textbf{User Prompt:}
You are an expert research idea generator.
Given the papers below, propose novel research ideas that extend, combine, or contrast this work.
Return strict JSON only with this schema:
\begin{verbatim}
{
  "ideas": [
    {
      "title": "concise idea title",
      "description": "2-3 sentence core description of the idea",
      "novelty": "why this is novel compared to existing work",
      "significance": "potential impact and importance",
      "key_references": ["paper title 1", "paper title 2"]
    }
  ]
}
\end{verbatim}

Generate exactly \{idea\_count\} ideas.

Papers:

\{paper\_context\}
\end{tcolorbox}

\subsubsection{Research Trend Predicting}
\label{prompt:task3-trend}

\begin{tcolorbox}[breakable, colback=orange!5!white, colframe=orange!60!black,
  left=1ex, right=1ex, top=0.5ex, bottom=0.5ex,
  fontupper=\small\ttfamily, fontlower=\small\ttfamily,
  title=\textbf{Research Trend Predicting:} Summarize research trends from chronologically ordered papers.]
\textbf{System Prompt:}
You analyze academic topic evolution and only return valid JSON objects.

\medskip\textbf{User Prompt:}
You are an expert research trend analyst.
Given chronologically ordered papers from one topic, summarize the research trend.
Return strict JSON only with this schema:
\begin{verbatim}
{
  "one_sentence_summary": "...",
  "trend_summary": "...",
  "stage_summary": [
    {"period": "...", "theme": "...", "description": "..."}
  ],
  "methodological_shifts": ["..."],
  "emerging_topics": ["..."],
  "open_gaps": ["..."],
  "future_directions": ["..."],
  "representative_papers": [
    {"title": "...", "year": 2024, "why_representative": "..."}
  ]
}
\end{verbatim}

Papers:
\{papers\_by\_year\}
\end{tcolorbox}


\subsubsection{Author Research Profile}
\label{prompt:task5-author-profile}

\begin{tcolorbox}[breakable, colback=purple!5!white, colframe=purple!60!black,
  left=1ex, right=1ex, top=0.5ex, bottom=0.5ex,
  fontupper=\small\ttfamily, fontlower=\small\ttfamily,
  title=\textbf{Author Research Profile:} Summarize a researcher's major directions from their publication list.]
\textbf{System Prompt:}
You summarize an author's research trajectory and only return valid JSON objects.

\medskip\textbf{User Prompt:}
You are an expert academic intelligence analyst.
Summarize one researcher's major directions from their publication list.
Return strict JSON only with this schema:
\begin{verbatim}
{
  "author_name": "\{author\_name\}",
  "overall_academic_profile": "...",
  "main_research_directions": [
    {"theme": "...", "active_years": "...", "description": "..."}
  ],
  "technical_arsenal": ["..."],
  "representative_papers": [
    {"title": "...", "year": 2024, "why_representative": "..."}
  ]
}
\end{verbatim}

Papers:
\{paper\_lines\}
\end{tcolorbox}


