\section{Limitations and Future Work}
Our {\ours} is under continuous maintenance and updates.
To further facilitate automated scientific discovery, we enumerate several important directions for future work.

\paragraph{CLI and Skills.}
Currently, our KG is primarily accessed through the Neo4j interface.
Although we provide usage guidelines, users are still required to write Neo4j queries if conducting secondary development.
To facilitate user adoption, particularly for integration with AI agents, we will encapsulate various KG retrieval and invocation functionalities into Command Line Interfaces (CLI).
 Additionally, for downstream tasks, we will distill the best practices identified during our experimental process into agentic skills, enabling one-click loading when utilizing agents for automated scientific discovery.

\paragraph{Integrating More Knowledge Forms.}
Currently, the scientific knowledge in our KG primarily encompasses papers, keywords, authors, and other paper-centric entities.
However, the complete research workflow extends beyond these elements to include atomic knowledge, theorems and standards, experimental experiences, datasets and code, among others.
How to acquire such knowledge and establish its associations with papers to form a more extensive and well-organized knowledge network that facilitates agentic utilization and reasoning constitutes a crucial research direction for our future work.
\textbf{We argue that KG is an indispensable knowledge organization form for scientific discovery} because, although LLMs have achieved remarkable advancements in semantic understanding, they still exhibit substantial deficiencies in capturing logical relationships among knowledge entities, a capability of paramount importance for scientific research that transcends mere semantic associations.

\paragraph{Benchmark and Evaluation.}
Benchmarks serve as a critical engine driving scientific progress.
Although automated scientific research has gained considerable popularity, numerous stages within this domain still lack high-quality benchmarks and evaluation metrics that faithfully simulate real-world research scenarios.
Furthermore, many scientific tasks involve long-form responses, and the evaluation of such outputs is often ambiguous, making it difficult to establish definitive verifiers.
KGs, as symbolic knowledge repositories, can provide essential reference points for such verification processes.
Additionally, the knowledge stored within KGs can serve as valuable data sources for benchmark construction.
In this paper, we merely present running examples of downstream tasks, remaining at the qualitative analysis level.
In future work, we will develop dedicated benchmarks based on {\ours} to quantitatively assess the downstream application capabilities of agent scientists.

\paragraph{Dynamic Update.}
Currently, our KG updates primarily rely on periodic manual execution of fixed scripts.
Although we support user-initiated updates, automated real-time updates are essential to keep pace with the rapidly evolving knowledge landscape.
In future work, we will systematize the real-time update strategies mentioned in \S\ref{sec:real_time_update} to support daily KG update mechanisms.