\section{Training}
\label{sec:training}

We train {SenseNova-Vision} by adapting Bagel-7B-MoT~\cite{bagel} to the unified vision-task corpus constructed in Sec.~\ref{sec:dataset_construction}.
Instead of training a multimodal model from scratch, we aim to endow an existing UMM with benchmark-compatible computer vision abilities while mitigating the degradation of its open-ended capabilities, including image understanding, instruction following, and image generation.
This section describes the mixed-task fine-tuning strategy, the high-resolution and multi-view training setup, and the training hyperparameters.

\textbf{Mixed-task fine-tuning.}
We perform supervised fine-tuning from the Bagel checkpoint on the SenseNova-Vision Corpus together with general-purpose multimodal data spanning visual question answering (VQA), text-to-image, and image-to-image tasks.
In a unified understanding and generation framework, this data mixture allows our model to learn representations of benchmark-readable outputs for structured visual understanding, dense geometric prediction, segmentation, and multi-view visual geometry, while mitigating degradation of the broad capabilities of the base UMM.

We adopt a joint sampling strategy that draws mini-batches from a weighted mixture of converted computer vision samples and auxiliary multimodal samples; the realized training mixture is shown in Fig.~\ref{fig:training-mixture-composition}.
Each mini-batch may contain samples drawn from multiple task categories and therefore produces a mixture of text and visual supervision targets.
During training, text and visual targets are interleaved in the same optimization process despite their different objectives.
Text-form outputs, including detection boxes, OCR strings, keypoints, and camera parameters, are tokenized and optimized with the standard cross-entropy (CE) loss under the next-token-prediction paradigm.
Visual outputs, including masks, depth maps, normal maps, and point maps, are encoded into a VAE latent space and optimized with the rectified-flow training objective inherited from Bagel.
In this way, heterogeneous computer vision targets are learned through the native text and image decoders of the base model, without introducing task-specific prediction heads.

The mixed-task joint training strategy enables the model to learn shared parameters across interleaved task formats and output modalities, potentially improving its generalization to zero-shot task variants.
Sec.~\ref{sec:text-prompted-interseg} gives further analysis.

\textbf{High-resolution and multi-view training.}
Compared with the VAE pathway, SigLIP2~\cite{siglip2} provides stronger semantic conditioning over input images, which benefits language-guided and region-level perception.
For image-input tasks requiring fine spatial conditioning, especially segmentation, we keep the SigLIP2 input resolution up to 980 pixels for both understanding and image-conditioned generation.
This preserves high-resolution conditioning whenever generation depends on an input image, whereas Bagel uses a lower SigLIP2 conditioning limit on the generation side.

For multi-view visual geometry, each data sample corresponds to a multi-view scene.
Due to memory constraints, each training sample is formed by randomly selecting at most 10 views from the corresponding scene.
For point map reconstruction, the selected views are aligned to the first view, center-normalized, and invalid pixels are mapped to a distant sky box.
For camera pose estimation, we reserve the final 2,009 vocabulary entries of the base model and repurpose them as a dedicated set of special tokens.
Among these tokens, 2,001 encode quantized pose parameters, including quaternion-based rotations, unit translation vectors, and scales, while the remaining 8 act as structural placeholders, as detailed in Appendix Table~\ref{tab:pose_token_markers}.
These special tokens are used exclusively for camera pose estimation, while other structured outputs remain serialized using ordinary text tokens.

\textbf{Training setup.}
We perform SFT with the VAE visual encoder frozen, while allowing all other modules and connectors to be updated.
We use the AdamW optimizer with a learning rate of $2.5\times10^{-5}$ and no weight decay.
We follow the method used in Bagel to pack training mini-batches, with 32K--36K tokens per rank and a maximum context window of 32K per sample.
The dropout rates for text, ViT and VAE input tokens are set to 0.05, 0.1 and 0.1, respectively.
The model is trained for 50K steps including 500 warm-up steps, and an EMA ratio of 0.995; the EMA checkpoint is used for evaluation.
All other configurations remain the same as in the SFT stage of Bagel.
