\section{Conclusion}
This paper explores a core question: how can we better absorb the capabilities of multiple experts into a single model? We identify that both conventional approaches, mixed RLVR (i.e., directly mixing data relevant to each expert) and the traditional static OPD pipeline (first training experts, then distilling them into a policy model), suffer from non-negligible loss of expert capabilities. We propose a central idea: Co-evolving Policy Distillation, where distillation among experts should occur \textit{during} training rather than \textit{after} it, and multiple expert models should serve as teachers and students to one another, co-evolving in synergy. Each expert's corresponding branch performs RLVR on its own data to explore the capability frontier, interleaved with mutual OPD from other experts to bring their behavioral patterns closer, making them more learnable from one another. Experiments validate that CoPD achieves an all-in-one consolidation of multiple expert capabilities, even outperforming the respective expert models. The strong performance of CoPD suggests an intriguing perspective: \textit{model parallel training} may serve as a promising scaling paradigm for further broadening the boundaries of model capabilities.


This paper is the third installment of our \textit{Self-Taught RLVR} research series, which aims to investigate how LLMs can better learn from themselves and self-evolve. We explore three complementary dimensions: the first installment, RLSD~\cite{yang2026selfdistilledrlvr}, investigates the \textit{informed self}---a self augmented by privileged information that teaches the  base self; the second installment, NPO~\cite{qin2026nearfuturepolicyoptimization}, focuses on the \textit{temporal self}---a near-future self teaching its past self; this paper explores the \textit{parallel self}---parallel selves mutually teaching each other. Each of these three papers is self-contained and complete; together, they instantiate \textit{Self-Taught RLVR}.


