\section{Introduction}
Reinforcement learning with verifiable rewards (RLVR)~\cite{grpo,liu2025understandingr1zeroliketrainingcritical,yu2025dapoopensourcellmreinforcement,zheng2025groupsequencepolicyoptimization} has emerged as the dominant post-training paradigm for large models, driving rapid progress on visual reasoning~\cite{huang2026visionr1incentivizingreasoningcapability,yao2025r1sharevlincentivizingreasoningcapability}, text reasoning~\cite{deepseekr1,yang2025qwen3technicalreport}, video understanding~\cite{feng2025videor1,wang2025timer1posttraininglargevision}, and other domains~\cite{kimiteam2026kimik25visualagentic,glm5team2026glm5vibecodingagentic}. However, training a single model on mixed-capability data with RLVR often exhibits a capability trade-off~\cite{shukor2025scalinglawsoptimaldata}, where gains in one capability come at the expense of another, as illustrated in Figure~\ref{fig:teaser}(a). We refer to this phenomenon as \textit{capability divergence}: distinct capabilities favor different optimization directions, making it difficult for a single training run to advance all of them simultaneously. Since no single RLVR run can master all capabilities at once, the field has converged on a two-stage pipeline: separate copies of the base model are first trained with RLVR on each capability to obtain specialized experts, and these experts are then consolidated into a unified policy model through on-policy distillation (OPD)~\cite{lu2025onpolicy-opd}, also known as Multi-teacher OPD (MOPD) when multiple experts serve as teachers simultaneously~\cite{mimo2025flash-mopd,deepseekai2026deepseekv4}. By separating capability-specific training from cross-capability consolidation, OPD avoids the gradient conflicts caused by capability divergence
% . Combined with its dense token-level supervision on the student's own trajectories, this design 
, which has made OPD a widely adopted approach for capability consolidation.


However, the way OPD is currently applied leaves an important issue unaddressed. In current practice, each expert is first trained until convergence on its own capability data, and then used as a fixed teacher for distillation into a separate student model, as illustrated in Figure~\ref{fig:teaser}(b). We argue that by the time distillation begins, the teacher has drifted too far from the student for its supervision to be effectively absorbed. To examine this, we measure the behavioral distance between teacher and student via \textit{top-$k$ token overlap} along on-policy trajectories. Our pilot study (\S\ref{sec:motivation}) shows that effective distillation requires teacher and student to remain behaviorally close. When they are too far apart, the teacher's supervision becomes difficult for the student to absorb, which is precisely where the static OPD pipeline falls short. We therefore argue that teacher and student should \textit{co-evolve}, with their behavioral distance kept within the absorbable range throughout training, rather than being set up as a static expert and a passive learner only at the moment of distillation.

Building on this insight, we propose Co-Evolving Policy Distillation (CoPD), which unifies capability exploration and consolidation into a single co-evolving process. CoPD maintains parallel training branches, each associated with a different capability. Instead of freezing each branch as a static teacher, CoPD interleaves branch-specific RLVR with cross-branch mutual distillation throughout training, so that each branch continuously deepens its own expertise while serving as an evolving teacher for the other.
% CoPD maintains parallel learning branches, each associated with a different capability. First, the branches take turns serving as teacher and student across capability domains, with each branch acting as the teacher on its own domain and as the student on the other's, so that both sides remain actively training rather than being cast into fixed teacher and separate student roles. Second, distillation is interleaved with capability-specific RLVR rather than postponed until convergence, so that teaching occurs before teacher and student drift too far apart. 
Concretely, training proceeds in alternating phases, as illustrated in Figure~\ref{fig:teaser}(c). In the \textit{RLVR phase}, each branch independently optimizes on its own capability data, deepening branch-specific expertise and \textit{opening up} the behavioral distance between branches, so that subsequent distillation has substantive new knowledge to transfer. In the \textit{mutual OPD phase}, each branch generates rollouts on the other branch's data and receives token-level supervision from the other, \textit{closing} the behavioral distance and absorbing what the other branch has newly learned with high efficiency, since the two branches remain within each other's effective range. 
The two phases are tightly complementary: RLVR drives the branches apart enough to make distillation informative, while interleaving keeps them close enough for distillation to be absorbable. Mutual OPD then exploits this window to transfer what each branch has learned to the other.
Because all branches start from a shared base and remain tightly coupled through continuous mutual distillation, their parameters do not diverge drastically, and the final unified model can be obtained through simple parameter merging.


We evaluate CoPD in both two-branch and three-branch post-training settings. The two-branch setting unifies text reasoning and multimodal reasoning, with each branch specializing in one capability and serving as an evolving teacher for the other. To demonstrate that CoPD scales beyond the pairwise case, we further evaluate a three-branch setting that jointly consolidates text reasoning, image-text reasoning, and video understanding.
CoPD consistently outperforms single-domain expert, mixed-data RLVR and static (M)OPD pipelines in both the two-branch and three-branch settings. Ablation studies confirm the necessity of bidirectional distillation over unidirectional alternatives and the advantage of continuous co-evolution over one-shot distillation.

Our contributions are as follows:
\begin{itemize}
    \item We identify a key limitation of the prevailing RLVR-then-OPD pipeline through a pilot study showing that OPD gain is inversely related to teacher--student behavioral distance, indicating that distillation suffers when the two drift too far apart. We accordingly argue that teacher and student should \textit{co-evolve} via cross-branch mutual distillation, a principle that naturally extends to multi-expert parallel training.
    \item We propose CoPD, which interleaves branch-specific RLVR with cross-branch mutual OPD throughout training. RLVR drives branch-specific exploration, while mutual OPD continuously transfers cross-domain knowledge. This design supplies sufficient new knowledge while keeping it absorbable at every step, achieving all-in-one consolidation without a separate distillation stage.
    \item Experiments show that CoPD consistently outperforms mixed RLVR, original OPD, MOPD baselines across text, image, video benchmarks. Notably, CoPD breaks the conventional ceiling that a unified student cannot surpass its domain-specific experts, turning cross-domain trade-offs into mutual gains.
\end{itemize}


