\section{Experiment Setup}
In this section, we discuss the experiment setup for post-training on \ours and the subsequent model evaluation.


\input{table/main-results}
\subsection{Post-Training on \data} 
Recent post-training work for video reasoning emphasizes sophisticated RL variants and reward engineering. In contrast, we aim to isolate a different, and arguably more fundamental, bottleneck: \emph{whether data design is the primary limiting factor for knowledge and reasoning intensive video understanding.} Accordingly, we deliberately adopt a standard, widely used SFT$\rightarrow$GRPO pipeline as a controlled scaffold, ensuring that algorithmic complexity does not become a confounder and that observed gains can be attributed more cleanly to the training data. 

Specifically, we use Qwen2.5-VL-7B-Instruct and Qwen3-VL-8B-Instruct as base models to assess whether \data yields consistent gains under various architectural designs and pretraining priors.
For SFT, we fine-tune each base model on \sftdata for one epoch. Starting from the resulting SFT checkpoint, we then run GRPO on \rldata for one epoch. 
For Qwen3-VL-8B-Instruct, we also conduct Zero-RL training by directly running GRPO on \rldata for one epoch.
We set batch size as 32 for both SFT and GRPO. 
For the GRPO accuracy reward, we follow prior video reasoning work~\cite{videor1, videorft} and use ROUGE for open-ended QAs and Exact Match for multiple-choice QAs.
The maximum video token number is 4,096, and the maximum number of frames is 128. The training hyperparameters and details are provided in Appendix~\ref{app:exp-details}.







\subsection{Evaluation Setup}\label{sec:reproduce-analysis}
We next describe our evaluation benchmarks and the standardized protocol for fair, reproducible model comparisons.

\vspace{-8pt}
\paragraph{Evaluation Benchmarks.}
We evaluate models on seven benchmarks
grouped into two categories: 
(1) \emph{General Video Reasoning}, including Video-MME~\cite{videmme}, MVBench~\cite{mvbench} and LongVideoBench~\cite{wu2024longvideobench}, which measure broad video understanding; and (2) \emph{Knowledge-intensive Video Reasoning}, including VideoMMMU~\cite{videommmu}, MMVU~\cite{mmvu},  SciVideoBench~\cite{scivideobench}, and \eval, which focus on domain-specific, expert-level reasoning. 

\vspace{-8pt}
\paragraph{Reproducibility Challenges in Prior Post-Training Work.}
We observe substantial cross-paper inconsistencies in reported results, particularly for base models used as post-training starting points. Based on careful follow-up experiments, we attribute these discrepancies primarily to prompt misalignment: base models are sometimes evaluated under prompt conditions that are misaligned with their intended inference mode. For example, Qwen2.5-VL-Instruct is not a ``reasoning'' model, yet some papers evaluate it with elaborate self-reflection and forced reasoning-trace instructions designed for post-trained `reasoning'' variants.

\vspace{-8pt}
\paragraph{Standardizing Evaluation for Fair Model Comparisons.}
To ensure fair and reproducible comparisons, for each model, we use the official prompt released by the original paper whenever available; otherwise, we adopt the default prompt templates from LMMs-Eval~\cite{lmms_eval}. For input frames, we follow the model-recommended inference configuration whenever it is specified by the model release. 
We run each model three times with independent sampling and report the mean.
All evaluations are performed using the LMMs-Eval framework~\cite{lmms_eval}.
Full evaluation details (\eg model prompts and inference parameters) are provided in Appendix~\ref{app:model-config}.






\section{Experiment Results and Analysis}\label{sec:6}
We next discuss our main findings and ablation analysis.
\subsection{Main Results}
Table~\ref{tab:benchmarks} presents the main experimental results. 
Post-training on \data consistently improves Qwen2.5-VL-7B-Instruct across all evaluated benchmarks, and yields clear gains for Qwen3-VL-8B-Instruct especially on knowledge-intensive video reasoning benchmarks.
The gains are most pronounced on knowledge-intensive tasks: SFT followed by RL on \ours raises the knowledge-intensive average of Qwen2.5-VL-7B from 41.9 to 46.6 (+4.7) and of Qwen3-VL-8B from 48.5 to 51.5 (+3.0), with the largest per-dataset improvements on MMVU and \eval (\eg +4.8 and +8.5 points for Qwen2.5-VL-7B).
Notably, the post-trained Qwen3-VL-8B attains the best knowledge-intensive average among 7/8B-scale models (51.5, versus 50.0 for the strongest competing model, Qwen3-VL-8B-Thinking).
These results underscore the value of our training data, which tightly integrates domain knowledge, visual grounding, and structured reasoning.
Starting from an SFT-initialized model, subsequent RL training consistently delivers higher performance than the SFT-only baseline, highlighting that combining SFT with RL is important for fully leveraging the strengths of \ours data.
Moreover, the RL-only variant also generally outperforms the SFT-only model, indicating that RL alone can induce stronger generalizable reasoning abilities.

\subsection{Case Study}
To better understand how post-training on \ours changes model behavior, we randomly sample 100 examples from \eval and compare outputs from different models. Our case study shows that the post-trained Qwen3-VL-8B model can integrate visual evidence with relevant domain knowledge to perform complex reasoning. It also exhibits ``aha-moment'' reasoning patterns indicative of deeper understanding. Examples are shown in Appendix~\ref{app:case}.


\input{table/frame_scaling}





