\section{Conclusion}
This work offers a corpus-centric perspective on post-training foundation models for advanced video reasoning. Instead of viewing visual perception, domain knowledge, and advanced reasoning as loosely linked elements, we show that integrating structured domain concepts with visually grounded examples yields stronger reasoning performance, without relying on sophisticated RL reward engineering. Extensive experiments and analyses confirm that post-training on the \ours dataset produces strong improvements, especially on knowledge-intensive video reasoning.

\section*{Impact Statement}
All videos in \ours are licensed under CC license, which enables reusers to distribute, remix, adapt, and build upon the material in any medium or format, so long as attribution is given to the creator. By restricting our dataset to CC-licensed videos, we ensure legal reusability and clear provenance, a critical aspect that has been ambiguous in prior training corpora. 
For human expert annotation and validation, we compensated annotators at an average rate of \$13 USD per hour, which exceeds the prevailing hourly rates for comparable local work.
All expert annotators provided informed consent to participate and explicitly authorized the public release and redistribution of their annotations as part of the resulting dataset and accompanying materials.
The \ours and \eval data construction incurs approximately 70.4K US dollars in model inference costs.

\section*{Acknowledgments}
We thank Dr.~Yu Rong for valuable suggestions on the post-training experiment design and paper writeup improvement.

