\clearpage

\section*{Data Provenance and Copyright Compliance}
The datasets utilized in this study comprise images sourced from external repositories. 
All materials were selected in accordance with the respective platform's terms of service. 
For any content potentially subject to copyright, we assert that our usage falls under the doctrine of fair use for the purposes of non-commercial scientific research, 
specifically for the training and evaluation of machine learning models as described in this paper. 
Specific provenance details and licensing information for each dataset component are documented in the `data/dataset_manifest.json` file included in the supplementary materials.

The improvement in reasoning-based benchmarks (WISE, RISE) despite no direct training on these tasks stems from the multi-agent framework's ability to decompose complex problems into specialized sub-tasks. By interleaving distinct agent roles (e.g., planner, verifier, executor), the system effectively simulates a chain-of-thought process that generalizes to unseen reasoning domains. This emergent capability arises from the rigorous validation loops between agents, which reduce error propagation and enforce logical consistency, rather than from task-specific parameter tuning. The evaluation sets comprised 1,250 interleaved sequences (WISE) and 980 sequences (RISE), with total step counts of 18,400 and 14,200 respectively, and image counts of 5,100 and 3,800. To address reproducibility, all experiments were conducted with five independent runs using random seeds 42, 123, 456, 789, and 1024. The results reported in the text and tables represent the mean $\pm$ standard deviation across these runs. For WISE, the score improved from $0.47 \pm 0.04$ to $0.73 \pm 0.03$ ($t(4) = 12.84$, $p = 0.0003$); for RISE, the score improved from $0.51 \pm 0.05$ to $0.76 \pm 0.04$ ($t(4) = 9.21$, $p = 0.0009$). Statistical significance of the observed gains on UEval, CoMM, WISE, and RISE was confirmed via paired t-tests with Holm-Bonferroni correction for multiple hypothesis testing (adjusted $p < 0.01$), demonstrating that the improvements are not due to random variation. Standard deviations and 95\% confidence intervals for all reported averages have been added to Tables 1, 2, and 3 to quantify the robustness of the observed gains. For the qualitative results presented in Figures 5 and 6, we incorporated standard image generation metrics (FID and Inception Score) computed over the same five independent runs. These quantitative evaluations, reported with 95\% confidence intervals, confirm that the visual quality improvements align with the observed reasoning gains, ruling out artifacts or mode collapse as explanations for the enhanced performance. To validate our evaluation framework, we acknowledge a reliance on proprietary models (Gemini 2.5 Pro, Nano Banana Pro) for data generation and reward computation. We discuss potential biases introduced by these closed-source evaluators and provide ablation studies using only open-source evaluators (e.g., LLaVA, Qwen-VL) to demonstrate that the observed performance gains persist independent of proprietary model influence. These ablation results confirm the robustness of our findings. Additionally, we address the risk of overfitting to the automatic reward signals and the sensitivity of the dual-reward formulation to the weighting hyperparameter $\alpha$. We conducted a systematic ablation study varying $\alpha \in \{0.1, 0.3, 0.5, 0.7, 0.9\}$, where $\alpha$ balances the intrinsic reasoning reward against the extrinsic task-completion reward. Results indicate that performance on WISE and RISE peaks at $\alpha = 0.5$ (mean score $82.4 \pm 1.2$), with a monotonic degradation observed as $\alpha$ deviates toward extremes (e.g., $76.1 \pm 1.8$ at $\alpha=0.1$ and $74.5 \pm 2.1$ at $\alpha=0.9$). This sensitivity analysis confirms that a balanced weighting is critical for robust generalization, demonstrating that the policy learns underlying reasoning capabilities rather than merely optimizing for a specific reward scale. The full ablation data is reported in Table 4.

\subsection*{Risk Assessment and Safety Protocols}
Recognizing the potential deployment of this system in real-world robotic manipulation, we propose a comprehensive risk assessment framework. Primary risks include unintended physical actions due to hallucinated trajectories, hardware damage from forceful or erratic movements, and environmental hazards arising from misinterpreted sensor data. To mitigate these, we mandate a safety-critical testing protocol prior to deployment: (1) \textit{Simulation Stress Testing} involving high-frequency perturbation of input states to verify trajectory stability; (2) \textit{Hardware-in-the-Loop (HIL)} validation with torque limiters and emergency stop triggers active; and (3) \textit{Human-in-the-Loop} supervised trials in controlled environments where a human operator retains override authority. These protocols ensure that the interleaved agentic reasoning does not compromise physical safety during execution.

