\section{Related Work}
\label{sec:related}

\noindent\textbf{Computer-use agent benchmarks.}
CUA benchmarks have advanced along two threads. On the GUI side, work evaluates either static element grounding~\citep{li2025screenspotpro,cheng2024seeclick,wu2024osatlas,you2024ferretui,lin2024showui,xu2024aguvis} or end-to-end execution in real OS, mobile, and web environments~\citep{xie2024osworld,bonatti2024waa,rawles2025androidworld,zhou2024webarena,koh2024visualwebarena}. On the CLI/coding side, benchmarks score agents on terminal-driven software-engineering tasks, from repository-level issue resolution~\citep{jimenez2024swebench,openai2024swebenchverified} to interactive terminal control~\citep{terminalbench2025,chu2026terminalworld}. Each thread optimizes for a single modality, so neither exercises an agent's ability to move information across channels, and a high score on one offers no guarantee on workflows that interleave both. This isolation can also blur what a single-channel GUI benchmark measures: on \osworld{}, a pixel-blind CLI agent solves most tasks with accuracy on par with a same-model vision agent while using far fewer interaction steps (Appendix~\ref{app:cli-osworld}), suggesting that some ostensibly GUI-bound tasks may not strictly require the GUI.

\medskip
\noindent\textbf{Multi-interface and hybrid CUA benchmarks.}
Another line of work dispatches agents across GUI and structured channels (CLI, APIs, MCP) within a single scaffold~\citep{song2025coact1,microsoft2025ufo2}, though as standalone task suites rather than deployed agent runtimes. \textsc{MCPWorld}~\citep{yan2025mcpworld}, \textsc{OSWorld-MCP}~\citep{jia2025osworldmcp}, and \textsc{PwP-Bench}~\citep{aggarwal2025pwp} add API or MCP actions on top of GUI task suites, but each task still admits equivalent single-channel solutions, making channel choice a per-step convenience. \textsc{ScienceBoard}~\citep{sun2026scienceboard} requires GUI/CLI co-use on most tasks, but only as a task-level division of labor, and is confined to scientific-discovery workflows.

\medskip
\noindent\textbf{Claw-class benchmarks for deployed CLI agents.}
As deployed CLI agent runtimes such as \textsc{Claude Code}~\citep{anthropic2025claudecode}, \textsc{OpenClaw}~\citep{openclaw2025}, and \textsc{Codex CLI}~\citep{openai2026codexagent} have entered real production use, a new line of benchmarks draws tasks from real user requests submitted to these runtimes. The resulting tasks are naturally in-the-wild, long-horizon, and span multiple files and services, as in \textsc{WildClawBench}~\citep{ding2026wildclawbench} and \textsc{ClawBench}~\citep{zhang2026clawbench}. Concurrent \textsc{CocoaBench}~\citep{hao2026cocoabench} shares this task profile but is browser-only and grades only the final output. None of them evaluate the GUI channel. Our \Eyeson{} inherits this paradigm but (i) promotes GUI to a peer channel of CLI on top of the deployed runtime, (ii) bakes channel non-substitutability into each task, and (iii) grades every trajectory with an agent-as-judge.
