\PassOptionsToPackage{table}{xcolor}
\documentclass[10pt,logo,copyright]{nvidiatechreport}
\linespread{1.15}

\input{packages}

\usepackage{calc}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{etoc}
\usepackage[nameinlink]{cleveref}

\crefname{equation}{Eq.}{Eqs.}
\crefname{figure}{Fig.}{Figs.}
\crefname{section}{Sec.}{Sec.}
\crefname{appendix}{App.}{App.}
\crefname{table}{Tab.}{Tabs.}

\definecolor{darkred}{rgb}{0.7, 0.0, 0.0}

\newcommand{\ours}{SpatialClaw}

\lstdefinestyle{pythonStyle}{%
  language=Python,
  basicstyle=\ttfamily\scriptsize,
  keywordstyle=\color{blue!70!black},
  commentstyle=\color{gray!70!black}\itshape,
  stringstyle=\color{red!60!black},
  showstringspaces=false,
  breaklines=true,
  columns=fullflexible,
  upquote=true,
  aboveskip=4pt,
  belowskip=2pt,
}

\definecolor{deltapos}{RGB}{0,130,0}
\definecolor{deltaneg}{RGB}{200,30,30}
\definecolor{ngreen}{HTML}{76B900}
\colorlet{bestcell}{ngreen!60}
\colorlet{secondcell}{ngreen!20}
\newcommand{\up}[1]{\,{\scriptsize\color{deltapos}(+#1)}}
\newcommand{\dn}[1]{\,{\scriptsize\color{deltaneg}($-$#1)}}
\newcommand{\eq}[1]{\,{\scriptsize\color{gray}(+#1)}}
\newcommand{\paragrapht}{\paragraph}

\newtcolorbox{findingbox}[1][]{%
  enhanced,
  colback=gray!8,
  colframe=gray!60,
  fonttitle=\bfseries,
  coltitle=black,
  colbacktitle=gray!30,
  sharp corners,
  boxrule=0.6pt,
  top=6pt,
  bottom=6pt,
  left=6pt,
  right=6pt,
  title={#1}
}

\newcounter{findingcounter}
\newcommand{\finding}[2]{%
    \refstepcounter{findingcounter}%
    \begin{tcolorbox}[
        colback=white!90!gray,
        colframe=ngreen,
        arc=2pt,
        boxsep=3pt,
        left=10pt,
        right=10pt,
        top=0pt,
        bottom=0pt,
        boxrule=0.8pt,
        enhanced jigsaw
    ]
        \textbf{Finding~\arabic{findingcounter}.}~\textbf{#1} #2
    \end{tcolorbox}
    \vspace{-0.1cm}
}

\title{\ours{}:\\Rethinking Action Interface for Agentic Spatial Reasoning}
\newcommand{\headertitle}{\ours{}: Rethinking Action Interface for Agentic Spatial Reasoning}

\author{Seokju Cho$^{1}$, Ryo Hachiuma, Abhishek Badki, Hang Su, Byung-Kwan Lee, Chan Hee Song, Sifei Liu,
Subhashree Radhakrishnan, Seungryong Kim$^{1}$, Yu-Chiang Frank Wang and Min-Hung Chen\\
{\Affilfont \textbf{Links}:~\href{https://github.com/NVlabs/SpatialClaw}{Code}~\textperiodcentered~\href{https://spatialclaw.github.io/}{Project Page}}\\
{\Affilfont NVIDIA}}
\authornotes{\textsuperscript{1}\,Affiliated with KAIST. Work done during Seokju Cho's internship at NVIDIA.}

\hypersetup{
  pdftitle={SpatialClaw: Rethinking Action Interface for Agentic Spatial Reasoning},
  pdfauthor={Seokju Cho, Ryo Hachiuma, Abhishek Badki, Hang Su, Byung-Kwan Lee, Chan Hee Song, Sifei Liu, Subhashree Radhakrishnan, Seungryong Kim, Yu-Chiang Frank Wang, Min-Hung Chen}
}

\fancyhead[C]{\footerfont \headertitle}

\begin{document}

\etocdepthtag.toc{mainpaper}

\maketitle
\vspace{-6.5mm}

\input{figures/tex/radar}

\input{sections/0_abstract}
\abscontent
\input{sections/1_intro}
\input{sections/3_interfaces}
\input{sections/4_method}
\input{sections/5_experiment}
\input{sections/6_conclusion}

\section*{Acknowledgements}
We are deeply grateful to Valts Blukis, Siyi Chen, Yi Dong, Tsung-Yi Lin, Karan Sapra, Andrew Tao, Bowen Wen, and Zhiding Yu for their valuable discussions, insightful feedback, and generous support throughout this work.

\clearpage
\input{supple}

\clearpage
\setcitestyle{numbers}
\bibliographystyle{plainnat}
\bibliography{main}

\end{document}
