\section{Discussion}
\label{sec:optimize_loop_strategy}
% ================================================================

The per-loop analysis in \autoref{sec:analysis} reveals a consistent
pattern: loop contributions are not uniform, and the interaction between
representational gains and the cost imposed by the CLP offset shifts across
loop index in a way that explains the non-monotonic performance curve.

\paragraph{Loop 2 is the principal site of productive refinement.}
The second loop is the primary site of productive refinement: among the
refinement loops it introduces the most coherent hidden-state update, the
highest inter-loop attention divergence, and the greatest output-distribution
shift, and it is where effective rank peaks.
The first loop establishes the global KV cache used by all subsequent loops,
so the quality of the loop-1 representations sets a ceiling on the
information available to every later loop via the frozen global attention
branch.
The second loop then refines these representations using both fresh local
context and the global cache, producing the largest net improvement.

\paragraph{Beyond loop 2: diminishing gains against a fixed cost.}
Beyond loop 2, the model faces compounding constraints.
The effective rank declines from its loop-2 peak (representations become less
diverse), so the shared block operates on a progressively lower-dimensional
input, reducing the capacity for new computation. Meanwhile, the hidden-state
update becomes oscillatory rather than convergent.
At the same time, the intrinsic offset cost $\Omega^{(r)}$ remains roughly
constant across loops: the CLP shift injects a comparable positional mismatch
at every boundary.
Because per-loop gains shrink rapidly while this offset cost stays fixed, the
mismatch claims an ever-larger share of each loop's net effect, so that beyond
the second loop the cost increasingly outweighs the benefit.

\paragraph{Practical guidelines for loop-count selection.}
These findings suggest several practical guidelines.
The performance-saturation point identified in the analysis corresponds to
a natural operating threshold for PLT deployment: $\loopcount = 2$ captures the
dominant refinement step while incurring only a single additional forward pass.
For finer-grained loop-count selection, the effective-rank trajectory is a
lightweight per-model diagnostic that requires no exhaustive sweep: if effective
rank is still rising at the candidate loop (representational diversity is not
yet saturated), an additional loop may yield genuine refinement, whereas a rank
that has begun to fall signals the onset of narrowing, after which further loops
mostly add the fixed CLP offset cost without compensating gain.
