\vspace{-0.5em}
\section{\system{}}
\label{sec:method}
\vspace{-0.5em}

\vspace{-0.3em}
\subsection{Overview}
\label{sec:overview}
\vspace{-0.3em}

\system{} is organized as a 23-stage pipeline across three phases (Figure~\ref{fig:pipeline}):
\textbf{Discovery} (scoping, literature search, multi-agent hypothesis generation),
\textbf{Experimentation} (self-healing code execution, result analysis, autonomous \textsc{Pivot}/\textsc{Refine} decisions),
and \textbf{Writing} (drafting, multi-agent review, revision, citation verification).
Five mechanisms span all three phases.
Multi-agent debate stress-tests hypotheses and conclusions from complementary perspectives.
Self-healing execution treats experiment failures as diagnostic information rather than termination signals.
Verifiable result reporting enforces that only grounded numbers and verified citations reach the final output.
Human-in-the-loop (HITL) collaboration allows researchers to intervene at high-leverage decision points without managing the full pipeline.
Cross-run evolution converts past failures into reusable safeguards through a persistent, time-decayed lesson store.
Each stage declares a formal input/output contract and supports checkpoint-based resumption; full stage definitions and hardware adaptation details are in Appendix~\ref{app:stages}.

\vspace{-0.3em}
\subsection{Multi-Agent Debate}
\label{sec:debate}
\vspace{-0.3em}

A single LLM agent naturally tends to confirm the hypotheses it generates, because the same model that proposes an idea has no structural incentive to disconfirm it.
\system{} addresses this by instantiating structured debate at two pipeline stages.
Each debate panel uses $K{=}3$ agents with complementary epistemic roles and a synthesizer that integrates their outputs into a single structured artifact.

\noindent \textbf{Hypothesis-stage debate.}
During hypothesis formulation, an \emph{Innovator} proposes high-risk hypotheses that challenge conventional assumptions, a \emph{Pragmatist} evaluates feasibility given hardware and time budgets, and a \emph{Contrarian} actively seeks weaknesses and confounds.
The synthesizer distills these perspectives into 2--4 falsifiable hypotheses, each annotated with testability criteria and required baselines.

\noindent \textbf{Result-stage debate.}
After experiments complete, a second panel evaluates the results.
An \emph{Optimist} surfaces strong findings, a \emph{Skeptic} challenges statistical significance and flags potential confounds, and a \emph{Methodologist} evaluates reproducibility and checks for data leakage.
The synthesizer produces a structured assessment that distinguishes supported claims from unsupported ones before any writing stage begins.

\vspace{-0.3em}
\subsection{Self-Healing Execution}
\label{sec:selfheal}
\vspace{-0.3em}

Experiment failure is common in real research.
Existing autonomous systems treat failure as a termination condition and discard all intermediate progress.
\system{} instead treats failure as diagnostic information: the system identifies what went wrong, decides whether to fix the current approach or change direction, and preserves all recoverable artifacts.

\noindent \textbf{Cascading code generation.}
Research experiments range from single-file scripts to multi-file systems with custom architectures.
A scoring function rates each experiment plan along six dimensions: architectural depth, file count, domain difficulty, dependency chains, historical failure rate, and control-flow complexity, and produces a complexity scalar $c \in [0,1]$.
Experiments above a fixed threshold $\tau$ (set to $0.6$ in all experiments) are dispatched to an external AI coding agent.
Experiments below $\tau$ are handled by a built-in multi-phase code agent that first emits a per-file blueprint, then generates files in dependency order using AST-derived summaries to maintain cross-file consistency.
Static validation gates check for detectable defects including identical ablation implementations and hardcoded metric values before any execution budget is spent.
A dedicated \emph{benchmark agent} handles dataset and baseline discovery; a \emph{figure agent} produces publication-quality visualizations.

\noindent \textbf{Sandboxed execution.}
All generated code runs in Docker containers under a three-phase network policy.
Phase~0 enables network access for dependency installation.
Phase~1 enables network access for data acquisition.
Phase~2 disables network access entirely during experiment execution, preventing both result exfiltration and pre-computed-result downloading.
Metric reporting is handled exclusively through a read-only evaluation harness, so generated code cannot redefine its own measurement infrastructure (Appendix~\ref{app:sandbox}).

\noindent \textbf{\textsc{Pivot}/\textsc{Refine} decisions.}
When an experiment fails or produces degenerate results, an automated repair loop captures the failure signature and generates targeted fixes.
The system then makes one of three decisions: \textsc{Proceed} when evidence supports the hypothesis, \textsc{Refine} when results are weak but the experimental direction is sound, or \textsc{Pivot} when the approach is fundamentally flawed, returning to hypothesis generation with the failure recorded as new evidence.
Systems that terminate on any failure avoid ambitious experiments by design.
By making failure recoverable, \system{} can pursue higher-risk hypotheses that would be abandoned under a brittle execution model.

\vspace{-0.3em}
\subsection{Verifiable Result Reporting}
\label{sec:verification}
\vspace{-0.3em}

LLM-generated papers face two integrity problems: fabricated experimental results and hallucinated citations.
Both arise from the same behavior---the model produces plausible-looking content with no grounding in actual evidence.
\system{} addresses both through deterministic verification gates applied at two granularities.

\noindent \textbf{Numeric registry.}
During execution, the system constructs a \emph{verified registry}: a whitelist of every value produced by experiment runs, storing per-condition means, standard deviations, and individual seed measurements.
At drafting time, pre-built \LaTeX{} tables populated exclusively from the registry are injected into the generation prompt.
After generation, a post-hoc verifier re-extracts every numeric claim and checks it against the registry, scoped per condition to prevent cross-condition false positives.
Claims in strict sections (Abstract, Results, Experiments) that cannot be matched to a registry entry trigger document rejection.
Claims in other sections are replaced with visible placeholders.
The writing agent can read the registry but cannot modify it.

\noindent \textbf{Citation verification.}
Every reference passes through a four-layer pipeline: DOI resolution via CrossRef, fuzzy title matching against OpenAlex, arXiv identifier lookup, and Semantic Scholar as a final fallback.
An LLM-based relevance check then classifies each reference as \textsc{Verified}, \textsc{Suspicious}, or \textsc{Hallucinated}.
References classified as \textsc{Hallucinated} are removed before any draft is finalized.

\vspace{-0.3em}
\subsection{Human-in-the-Loop Collaboration}
\label{sec:hitl}
\vspace{-0.3em}

Full automation reduces output quality at critical junctures where domain judgment matters.
Exhaustive step-by-step oversight eliminates the efficiency gains of automation.
The useful region lies between these extremes: human expertise is most valuable at a small number of high-leverage decision points rather than distributed uniformly across the pipeline.
\system{} provides seven intervention modes that let researchers select their operating point along this spectrum.

\noindent \textbf{Intervention modes.}
\emph{Full-Auto} runs the entire pipeline without human input.
\emph{Gate-Only} pauses at three fixed checkpoints: literature screening, experiment design, and final quality review.
\emph{Thorough} pauses at all phase boundaries, giving researchers visibility without requiring approval at every substep.
\emph{CoPilot} targets six high-leverage decision points, including hypothesis co-creation (\emph{Idea Workshop}), experiment design review (\emph{Baseline Navigator}), and collaborative paper drafting (\emph{Paper Co-Writer}).
\emph{Step-by-Step} requires explicit approval at every stage.
Two further regimes decompose CoPilot for the ablation in Section~\ref{sec:hitl_exp}: \emph{Pre-Experiment} retains intervention only at literature screening, hypothesis generation, and experiment design (early-pipeline), while \emph{Post-Experiment} retains intervention only at result analysis, paper draft, and quality gate (late-pipeline).
Our HITL ablation in Section~\ref{sec:hitl_exp} evaluates all seven modes empirically.

\noindent \textbf{SmartPause.}
Rather than relying on fixed checkpoints, SmartPause monitors the system's estimated uncertainty at each stage.
When uncertainty exceeds a learned threshold, the system pauses and presents the decision to the researcher.
The threshold adapts based on historical approval patterns: stages where the researcher frequently overrides the system are paused more often, while stages with consistently high approval rates proceed autonomously.

\vspace{-0.3em}
\subsection{Cross-Run Evolution}
\label{sec:evolution}
\vspace{-0.3em}

Existing autonomous research systems are stateless across runs: every run begins without knowledge of previous attempts, repeating failures that earlier runs already encountered.
\system{} maintains a persistent lesson store that converts past failures into future safeguards.

At the end of each run, the system extracts structured lessons from repair attempts, \textsc{Pivot}/\textsc{Refine} decisions, HITL gate feedback, and verification results.
Each lesson records a category, a severity score $s(l) \in (0, 1]$, and a recommended mitigation.
When a new run begins, relevant lessons are retrieved by category and ranked by a time-decayed weight:
\begin{equation}
w(l) \;=\; s(l) \cdot \exp\!\left(-\ln 2 \cdot \frac{\Delta t}{T_{1/2}}\right),
\label{eq:lesson-weight}
\end{equation}
where $\Delta t$ is the elapsed time since the lesson was recorded and $T_{1/2}$ is a half-life hyperparameter controlling how quickly older lessons lose influence (default $T_{1/2} = 30$ days).
Lessons are injected into prompts as natural-language overlays, requiring no model retraining and remaining applicable to any LLM backbone.
This design means that recent failures strongly constrain subsequent runs, while lessons from completed, successful lines of work gradually fade from prominence.