\begin{tcbCodeListing}{Example Task Format}{tcbyaml}
title: "Survey Recent Papers on a User-Specified Research Topic"

description: >
  The user is starting a small research project and asks the agent to organize
  a set of relevant papers from a local paper list. The task requires the agent
  to infer the user's preferred topic focus, filter candidate papers, and
  prepare a concise reading plan that supports later topic selection.

task_type: long_term

trigger:
  type: user

intent:
  initial_input: >
    I wrote a local file at [FILE_NAME] that contains recently accepted papers.
    Please organize several papers related to my research theme for me.

  hidden_intent:
    - content: "Prioritize papers related to [RESEARCH_THEME]."
    - content: "Recommend around [N] papers rather than producing a long survey."
    - content: >
        For each paper, provide a brief introduction and summarize the key
        technical points.
    - content: >
        Include the official paper link when available.
    - content: >
        Include the project or code link when available; otherwise explicitly
        state that no public repository is found.
    - content: >
        For each paper, state whether it is suitable to follow and explain how
        the user should follow it.

objectives:
  checklist:
    - criterion: "Include [PAPER_TITLE_1]."
    - criterion: >
        Explain that [PAPER_TITLE_1] studies [CORE_METHOD_1] for
        [RESEARCH_THEME], and mention its main training or inference strategy.
    - criterion: "Write the official paper link for [PAPER_TITLE_1] as [PAPER_LINK_1]."
    - criterion: "Write the code link for [PAPER_TITLE_1] as [CODE_LINK_1]."

    - ......

    - criterion: >
        For every recommended paper, state whether it is suitable to follow.
    - criterion: >
        For every recommended paper, give a concrete follow-up action, such as
        reading the method section, reproducing the code, or comparing the task
        formulation with the user's project.

metadata:
  difficulty: hard
\end{tcbCodeListing}