

% \newcommand{\tcolorboxPromptUserAgentStageOneTitleText}{\fontsize{10}{12}\selectfont}
% \newcommand{\tcolorboxPromptUserAgentStageOneCommonText}{\fontsize{7.5}{9}\selectfont}
% \newcommand{\tcolorboxPromptUserAgentStageOneTitleBeforeVspace}{\vspace{10 pt}}
% \newcommand{\tcolorboxPromptUserAgentStageOneCommonTextLineSkip}{\setlength{\baselineskip}{5 pt}}

\newcommand{\tcolorboxPromptUserAgentStageOneTitleText}{}
\newcommand{\tcolorboxPromptUserAgentStageOneCommonText}{}
\newcommand{\tcolorboxPromptUserAgentStageOneTitleBeforeVspace}{}
\newcommand{\tcolorboxPromptUserAgentStageOneCommonTextLineSkip}{}

\tcolorboxPrompt{Prompt -- Completion Checking
}{
% \tcolorboxPromptUserAgentStageOneCommonTextLineSkip
% \tcolorboxPromptUserAgentStageOneCommonText
\ttfamily
\#\# Assistant Response

<assistant\_message>

\{assistant\_message\}

</assistant\_message> \\

\#\# Files Read Context

Use these file contents as additional context for judging intent satisfaction.

\{files\_context\_xml\} \\

\#\# Hidden Intent Status Snapshot

Judge only the hidden intents listed here.

\{status\_hidden\_intents\_xml\} \\

\#\# Objective

Decide whether the assistant response, together with the files-read context above, already reflects each listed hidden intent. \\

\#\# Evaluation Policy

1. Judge from the assistant response and the files-read context above.

2. Be strict and objective. The assistant must precisely and explicitly hit the hidden intent.

3. A hidden intent is satisfied only if the assistant provides specific, detailed explanations or concrete actions in the response. Vague or generic answers do not count.

4. Fully trust the assistant's wording and the files-read context, but strictly evaluate the level of detail provided.

5. Do not call tools or check factual accuracy beyond the provided context.

6. YES means the response and context precisely address the hidden intent with specific details.

7. NO means the response and context do not precisely hit the intent, or lack specific details. \\

\#\# Output Format

Output only XML blocks in the following shape.

<c1>

<content>

\{hidden\_intent\_content\}

</content>

<decision>

YES or NO

</decision>

</c1>
}