

% \newcommand{\tcolorboxPromptUserAgentStageTwoTitleText}{\fontsize{10}{12}\selectfont}
% \newcommand{\tcolorboxPromptUserAgentStageTwoCommonText}{\fontsize{7.5}{9}\selectfont}
% \newcommand{\tcolorboxPromptUserAgentStageTwoTitleBeforeVspace}{\vspace{10 pt}}
% \newcommand{\tcolorboxPromptUserAgentStageTwoCommonTextLineSkip}{\setlength{\baselineskip}{5 pt}}

\newcommand{\tcolorboxPromptUserAgentStageTwoTitleText}{}
\newcommand{\tcolorboxPromptUserAgentStageTwoCommonText}{}
\newcommand{\tcolorboxPromptUserAgentStageTwoTitleBeforeVspace}{}
\newcommand{\tcolorboxPromptUserAgentStageTwoCommonTextLineSkip}{}

\tcolorboxPrompt{Prompt -- Clarification Checking
}{
% \tcolorboxPromptUserAgentStageTwoCommonTextLineSkip
% \tcolorboxPromptUserAgentStageTwoCommonText
\ttfamily
\#\# Assistant Response

<assistant\_message>

\{assistant\_message\}

</assistant\_message> \\

\#\# Hidden Intents Still Not Provided

Judge only the hidden intents listed here.

\{status\_hidden\_intents\_xml\} \\

\#\# Objective

Decide whether the assistant response contains a clear follow-up question that is specifically about each listed hidden intent. \\

\#\# Evaluation Policy

1. Judge from the assistant response. Consider all follow-up questions, requests, and action suggestions inside it, not only the last sentence.

2. YES means the assistant explicitly asks about that hidden intent, asks a very close confirmation question about it, or proposes concrete next steps that directly correspond to it.

3. NO means the question is missing, vague, generic, or does not clearly target that hidden intent.

4. Generic prompts such as ``anything else?'' or ``do you want to add more?'' must be NO.

5. A question must clearly get the point. Broad topic overlap is not enough. \\

\#\# Output Format

Output only XML blocks in the following shape.

<c1>

<content>

\{hidden\_intent\_content\}

</content>

<decision>

YES or NO

</decision>

</c1>
}