% omit the formula for now.
% use letters to represent things that are used repeatedly (formalization). -- [todo] more needed
% [todo] maybe need to write about the domain "retail", and the tasks' min steps (5/6/7/8/9)
\section{\bench{}}
\label{sec:benchmark}
% overview: (1) what is on the bench; (2) what will be introduced next



\BENCH{} is an interactive benchmark designed to evaluate LLM agents’ ability to explore and plan in massive tool-use environments.
Unlike conventional tool-use settings, \bench{} requires agents to infer implicit sub-goals and discover useful tool-use paths.
This setting reflects real-world agent applications, where models often lack full tool visibility and must explore tools to uncover useful intermediate information.
As demonstrated in Figure~\ref{fig:demo_figure}, \bench{} automatically constructs a broad ecosystem of task-grounded tools and exposes it through a retriever.
The retriever supports bi-directional exploration, enabling forward anticipation from available evidence and backward anticipation from the target outcome.
This design requires agents to explore the tool space while inferring intermediate information.
% Moreover, \bench{} supports evaluation in two modes: a \textbf{default mode}, where the tools retrieved are clean and functional; and a \textbf{block mode}, which introduces environmental noise by replacing selected useful tools with distracting or ineffective alternatives. 
Moreover, \bench{} supports two evaluation modes: a \textbf{default mode} that evaluates tool use under ordinary retrieval noise, with useful tools mixed among distracting alternatives; and a \textbf{block mode}, which further disrupts selected useful paths while preserving solvability.
It therefore tests whether agents can recover and re-plan when a plausible path becomes unreliable.
Each task unfolds through an iterative interaction loop in which the agent observes the state, reasons about its next step, retrieves or invokes tools, and receives executable feedback from the environment. 
We describe the benchmark construction below and discuss its significance and generalization in Appendix~\ref{app:significance-generalization}.



% \jy{check here}
% Formally, we model the benchmark as a typed tool graph $\mathcal{G}=(\mathcal{D}, \mathcal{T})$, where $\mathcal{D}$ is the datatype set and $\mathcal{T}$ is the tool set. 
% A query instance is defined by a natural-language question $q$, an initial datatype set $\mathcal{D}_0 \subseteq \mathcal{D}$, and a target datatype $d^\star \in \mathcal{D}$, where the datatype variables are internal benchmark abstractions used for tool/task construction and runtime verification rather than explicit planning targets presented to the agent. 
% The benchmark supports two evaluation settings: a default setting that measures agent performance in large toolspaces, and a blocked setting that injects retrieval-time noise by replacing selected baseline tools with noisy alternatives. 
% In this section, we describe the benchmark construction in three parts: the data layer, the runtime environment, and the agent-environment interaction protocol.
% \jy{check here}


% \jy{haven't refine the wordings between section 3 and 3.1}

\subsection{Environment Setup}
% \paragraph{Datatypes.}

% Once the datatype set is fixed, it induces a large combinatorial space of candidate tools. For example, if a domain contains $x$ datatypes, then the number of possible $1$-input-$1$-output typed tools is $A_x^2=x(x-1)$, and the number of possible $2$-input-$1$-output typed tools is $A_x^3=x(x-1)(x-2)$, where we exclude cases with repeated input/output roles. 
% Higher-arity tools can be enumerated similarly. 

% Once the datatype set is fixed, it induces a large combinatorial space of candidate tools. \jy{why? you need to specify that we are using reverse engineering here, and our }
% Since the vast majority of tools in our setting produce a single output, we mainly enumerate the family of $k$-input-$1$-output typed tools for different values of $k$. 
% If a domain contains $x$ datatypes, then the number of possible $k$-input-$1$-output candidates is $A_x^{k+1}$ when repeated input/output roles are excluded.
% This datatype-first construction makes the size and structure of the tool ecosystem explicit, which is essential for studying planning under massive toolsets.

% \paragraph{Tool Library.}
% We introduce self-defined datatypes \jy{datatype is LLM generated and filtered.} primarily to define the tool space in a structured and enumerable way. \jy{This is quite confusing: while introducing datatypes can help define the tool space in a structured and enumerable way?}
% For each domain, we first specify a datatype inventory $\mathcal{D}$, where each datatype $d \in \mathcal{D}$ represents a well-defined piece of business information, such as an entity identifier, a status field, or an attribute value. 
% \jy{This example could be more specific, I think entity identifier is just too abstract}
% The final tool library is obtained by filtering this combinatorial candidate space. 
% \jy{the term ``combinatorial candidate space'' appears before actually introducing it}
% Concretely, each candidate tool $\tau$ is a typed transformation from input datatypes $\mathrm{In}(\tau) \subseteq \mathcal{D}$ to an output datatype $\mathrm{Out}(\tau) \in \mathcal{D}$, and we use an LLM \jy{the LLM should have a sign} to remove unreasonable or undesired candidates (detailed filtering criteria provided in Appendix~\ref{app:detailed-filtering-criteria}). 

\paragraph{Tool Library.}
We introduce self-defined datatypes to provide a typed representation of domain-specific information and to support the systematic construction of tools.
For the retail domain, we first define a set of datatypes $\mathcal{D}$.
Each datatype $d \in \mathcal{D}$ represents a concrete type of distinct domain information, such as \texttt{person\_name} or \texttt{purchase\_status}.~The initial datatype inventory is proposed by a generation LLM $M_{\mathrm{gen}}$ and then automatically filtered by another LLM $M_{\mathrm{fil}}$ to remove vague, redundant, unreasonable or unrealistic datatypes. We then construct~candidate tools by considering all pairs of datatype sets $(\mathcal{D}_{\mathrm{in}}, \mathcal{D}_{\mathrm{out}})$ over $\mathcal{D}$, where $|\mathcal{D}_{\mathrm{in}}|{=}m$ and $|\mathcal{D}_{\mathrm{out}}|{=}n$ denote the given numbers of input and output datatypes, respectively.
For each such pair, $M_{\mathrm{gen}}$ proposes a candidate tool functionality whose input schema is given by $\mathcal{D}_{\mathrm{in}}$ and whose output schema is given by $\mathcal{D}_{\mathrm{out}}$.
For each~resulting tool $\tau$, we denote its input and output datatype sets by $\mathcal{D}_{\mathrm{in}}(\tau)$ and $\mathcal{D}_{\mathrm{out}}(\tau)$.
For example,~given
$\mathcal{D}_{\mathrm{in}}{=}\{\texttt{product\_name}, \texttt{store\_name}\}$
and
$\mathcal{D}_{\mathrm{out}}{=}\{\texttt{inventory\_status}\}$,
$M_{\mathrm{gen}}$ may propose a tool functionality that checks whether the specified product is available at the specified store and returns its inventory status. 
The final tool library $\mathcal{T}$ is obtained by applying $M_{\mathrm{fil}}$ to remove unreasonable, redundant, or undesired candidates. 
% We provide tool construction details in Appendix~\ref{app:schema-construction-details} and filtering rules in Appendix~\ref{app:detailed-filtering-criteria}.
We further augment the functional tool library with paired noisy tools to simulate realistic ecosystems where functionally similar tools may include distractors~\citep{huang2024metatoolbenchmarklargelanguage}. These noisy tools are semantically similar to previously constructed tools, but explicitly disclose their unavailability or unreliability in the descriptions, allowing careful agents to reject them after inspection.
We provide executable tool construction details in Appendix~\ref{app:schema-construction-details}, filtering rules in Appendix~\ref{app:detailed-filtering-criteria}, and noisy tool construction details in Appendix~\ref{app:noisy_tool}.

% \paragraph{Tool Response Construction.}

\paragraph{Tool Response Construction.}
We further construct a backend database to support tool responses with concrete data instances. 
% Specifically, while tool specifications define the required datatypes, the backend database instantiates these types with concrete values, enabling tools to return actual information in the benchmark.
% Each row corresponds to a complete retail case and stores the values needed by the tools, such as order, delivery, refund, and payment fields.
% At execution time, a tool uses its input values to find the matching row in the backend database, and then returns the requested output field from that row.
For each retail case, we prompt the generation LLM $M_{\mathrm{gen}}$ to instantiate values for the full datatype set in the retail domain, producing a complete structured record (to ensure the completeness of the backend). 
During execution, a tool $\tau$ matches its input arguments to the corresponding record and returns the values for its output datatype set $\mathcal{D}_{\mathrm{out}}(\tau)$.
We ensure that the instantiated values are non-trivial and cannot be inferred from common sense, making tool use necessary for deriving the requested output values.
Appendix~\ref{app:tools-necessity-details} details backend construction and why answers require tool-derived values.
% \jy{maybe introduce the response from the noisy tools in 1--2 sentences as well.}
For noisy tools, we assign fixed return values at construction time. These responses are shared across models and query instances and are designed to provide no useful evidence for solving the task (details provided in Appendix~\ref{app:noisy_tool}).
% This is sufficient because noisy tools are designed to be unavailable or unreliable, and their responses only need to match their descriptions consistently.


% \qh{1. how to construct the tool response}
% \qh{2. how to ensure that every valid tool invocation lead to valid output?}
% \qh{In the benchmark, all valid inputs must be provided with a valid tool response: briefly introduce how you made it, and details could be put in appendix.}
% \qh{3. ensure that every tool-call cannot be avoided, and briefly introduce why, and details in Appendix~\ref{app:tools-necessity-details}}

% \paragraph{Queries.}
% We construct tasks from the completed tool graph \jy{no tool graph} rather than writing them manually. 
% Internally, a task is defined by an input datatype set and a target datatype, which serves as a structured representation of the missing target information to be recovered. 
% \jy{this sentence is really weird: missing target info??? just say sth like goal state.}
% We enumerate possible task schemas \jy{what is task schema} and keep only those for which the tool graph contains at least one reachable path from the inputs to the target. 
% \jy{no tool graph: we ensure that the queries are all solvable by calculating all the ground-truth tool call sequences using code. details provided in Appendix~\ref{app:groundtruth-details}.}
% We then filter tasks by difficulty-related properties derived from these paths, such as shortest-path length and the functional role of the input datatypes, and finally retain 327 tasks.
% \jy{details should be provided in Appendix~\ref{app:}}
% Queries are obtained by turning retained tasks into natural-language questions and instantiating them with concrete values from the database. 
% \jy{transoformation is done by LLM}
% \jy{how you obtained the ground-truth}
% Formally, for a query instance $x=(q,\mathcal{D}_0,d^\star)$, the agent observes the question $q$, starts with the initial evidence corresponding to $\mathcal{D}_0$, and is expected to recover the target information represented internally by $d^\star$ before producing the final answer.
% \jy{this sentence is repeated by silly gpt, rephrase it in a straightforward way and state the evaluation metric for accuracy, e.g.: exact match? in-accuracy?}
% During instantiation, we require the sampled records to remain consistent with the stored path information, so that each query is not only naturalistic but also executable under the benchmark environment. 
% \jy{don't understand what this is}
% In this way, the final query set reflects both the abstract reachability structure of the tool graph and the instance-level constraints of concrete data.


\paragraph{Queries.}
Figure~\ref{fig:demo_figure} gives an example of this pipeline.
We generate queries through a three-step pipeline.
First, we specify each task internally as $r=(\mathcal{D}_0, \mathcal{Y})$, where $\mathcal{D}_0 \subseteq \mathcal{D}$ denotes the initial input datatype set and $\mathcal{Y} \subseteq \mathcal{D}$ denotes the target datatype set.
Second, we compute the set $\Pi(r)$ of ground-truth tool-call sequences that transform $\mathcal{D}_0$ into $\mathcal{Y}$ via code, and discard any task with $\Pi(r)=\emptyset$.
We formalize this internal reachability structure as a state graph and use it for path enumeration and solvability checks (details are provided in Appendix~\ref{app:state-graph}).
Third, we instantiate each remaining task with concrete entities and attribute values by sampling values from the backend while ensuring that the resulting instance is solvable, and use $M_{\mathrm{gen}}$ to verbalize the instantiated task as a natural-language query $q$.
Given $q$ and step-by-step tool-call instructions, we then prompt $M_{\mathrm{gen}}$ to produce an answer $o^\star$ following one valid sequence $\pi \in \Pi(r)$, and use $o^\star$ as the ground truth.
Finally, we retain only instances whose shortest valid solution requires at least five distinct tool calls.
% \jy{refer to human annotation part, data cases.}
We ensure tool and datatype quality through partial manual inspection, with details in Appendix~\ref{app:human-annotation}, and provide concrete data cases in Appendix~\ref{app:data-case-study}.
Details on selecting $M_{\mathrm{gen}}$ and $M_{\mathrm{fil}}$, query filtering, and ground-truth construction are provided in Appendices~\ref{app:model-choice}, \ref{app:query-construction-details}, and \ref{app:groundtruth-details}, respectively.




% \subsection{Runtime Environment}
% % === There should be sth here ===
% The runtime environment turns the static benchmark data into an executable interaction process. 
% It specifies how the agent interacts with the benchmark during inference, including what evidence is available at each step, what tools can be exposed, and how the environment validates and updates the interaction state after each action.

% \qh{I think the subsection 3.2 and the upside paragraph is not useful.}






% \subsection{Agent–Environment Interaction}
% % runtime's protocol
% % how agent interact with environment
% % stopping criteria
% Each query is executed as a multi-turn interaction between the agent and the environment. 
% At each step, the agent must output exactly one of three actions: retrieving candidate tools, calling a retrieved tool with structured arguments, or returning the final answer. 
% We denote these actions by $a_t \in \{\texttt{retrieve}, \texttt{call}, \texttt{answer}\}$. 
% The environment parses $a_t$, validates its format and semantics, executes the corresponding operation when valid, and returns structured feedback to the agent. 
% \jy{code-level details, don't need to write anything like ``parses xxx''}
% Retrieval feedback includes the matched datatypes, a list of retrieved tools, and an optional note when no exact typed match exists. 
% \jy{here you should explain when will no exact typed match happen and convince others that our benchmark design is robust: it is agent who request a non-existing tool, instead of our benchmark being problematic}
% \jy{What I mean is that, our semantic matching particle in the retriever should be good, it is just the model requesting a non-existing tool}
% Tool-call feedback includes the execution result, obtained from the constructed backend. 
% Here, datatypes are used by the environment as an internal bookkeeping and anti-shortcut mechanism, rather than as explicit planning objects that the agent is instructed to reason over directly. \jy{it is just to prevent the agent to hack the answer}
% The interaction terminates when the agent returns a final answer, exceeds the step budget, or accumulates too many formatting or action errors. 
% \jy{detailed termination condition is provided in Appendix~\ref{app:termination-condition}}
% \jy{too long}

% \jy{Here is an example for another paper: Each task instance in CostBench follows an iterative interaction loop.
% At each step, the agent observes the current state, including the user query, available tools and their costs, previously obtained data, and feedback from earlier actions, then decides which atomic or composite tool to invoke next.
% The environment responds with an updated state and cost-tagged feedback.
% The interaction terminates once the agent produces the final output or reaches the predefined task sequence length.
% Just keep the thing clear and clean (short)}

\subsection{Agent--Environment Interaction}

Each query is executed as a multi-turn interaction between the agent and the environment.
At each step, the agent outputs exactly one of three actions:
retrieving candidate tools, calling a retrieved tool with structured arguments, or returning the final answer, denoted by
$a_t \in \{\texttt{retrieve}, \texttt{tool-call}, \texttt{answer}\}$.
The environment responds according to the chosen action:
\textbf{(1)} for \textbf{\texttt{retrieve}}, it returns candidate tools and an additional note if the requested tool does not exist;
\textbf{(2)} for \textbf{\texttt{tool-call}}, it executes the call on the constructed backend and returns the result;
\textbf{(3)} for \textbf{\texttt{answer}}, the trajectory terminates and the final answer is evaluated.
The trajectory also terminates when a predefined, agent-visible budget $T_{\max}$ is exhausted.
To prevent pure guessing or other shortcuts, we apply datatype checking to ensure that the final answer is derived from a valid \texttt{tool-call} result. 
Please refer to Appendix~\ref{app:runtime-interaction-details} for more details.

\paragraph{Agent State Definition.} At runtime, the environment maintains an explicit agent state. 
We define the state at step $t$ as $s_t = (q, \mathcal{U}_t, \mathcal{D}_t),$ where $q$ is the user query, $\mathcal{U}_t$ is the set of all tools discovered so far in the current query and available for the agent to call, and $\mathcal{D}_t \subseteq \mathcal{D}$ is the set of datatypes already obtained through successful tool calls. 
The initial state is determined by the datatypes used to construct the query, with initial evidence $\mathcal{D}_0$ and $\mathcal{U}_0=\varnothing$. 
Whenever a tool call succeeds and outputs values with datatype set $\mathcal{D}_{\mathrm{out}}(\tau)$, the environment updates $\mathcal{D}_{t+1}=\mathcal{D}_t \cup \mathcal{D}_{\mathrm{out}}(\tau)$.
The agent's objective is to obtain the target information specified in the query and return the correct final answer through tool use.
Note that the state $s_t$ is an environment-maintained latent state which is used to track progress across tool calls, but is not directly visible to the agent.

\paragraph{Tool Retriever.}

Tool-use in \bench{} is supported by a retriever that lets the agent explore the tool space through high-level natural-language queries rather than by inspecting the full tool library. 
At each step $t$, the agent may query the retriever in three complementary modes aligned with bi-directional anticipation:
\textbf{(1) Input-conditioned retrieval}, which supports \textbf{\textit{Forward Anticipation}} by asking what information or tool affordances can be reached from the evidence currently available;
\textbf{(2) Output-conditioned retrieval}, which supports \textbf{\textit{Backward Anticipation}} by asking what tools or prerequisite information may lead to a desired intermediate or final outcome;
and \textbf{(3) Input-output-conditioned retrieval}, which constrains the search by specifying both the available information and the desired result. 
The retriever grounds these queries to tool signatures and adds matched tools to the agent-callable set \(\mathcal{U}_t \subseteq \mathcal{T}\) for subsequent turns.
Implementation details of request matching and retrieval modes are provided in Appendix~\ref{app:retriever-details}.
To ensure that retrieval is reliable under natural-language queries, we conduct a single-step retrieval study and report the results in Appendix~\ref{app:retriever-robustness}.


% \jy{refer to appendix. The choice of Mgen and Mfil should not be here.}

% \paragraph{Retrieval-Time Blocking.}
% % blocking mechanism
% % what is block here: During the retrieval process, some baseline tools are replaced with noisy tools, causing the model to be unable to traverse paths containing the replaced baseline tool.
% Our blocking mechanism is implemented by injecting noisy tools at retrieval time. 
% Let $\tau \in \mathcal{T}$ be a baseline tool that lies on a valid solution path. When $\tau$ would have been exposed to the agent, the environment instead replaces it with one or more noisy tools that are similar to $\tau$ and remain callable under the normal interaction protocol. 
% After retrieval, the agent may still choose to invoke these noisy tools and receive environment feedback. 
% However, because the original baseline edge \jy{no tool graph} has been replaced, the corresponding path is effectively broken for the agent, and is therefore treated as blocked.

% We consider three types of noisy tools. 
% \textit{Return-error} tools closely resemble the baseline tool in both name and apparent function, but return an explicit error when called. 
% \textit{Return-counterfactual} tools also closely resemble the baseline tool, but return a clearly unreasonable value, creating an implicit form of perturbation rather than an explicit failure signal. 
% \textit{Irrelevant} tools have names that resemble the baseline tool, but their functionality is observably different, making them potentially distinguishable at inspection time. 
% The first two types are hard to disambiguate before execution, whereas the third mainly introduces retrieval noise through semantic distraction.


\paragraph{Retrieval-Time Blocking.}

Our benchmark includes an optional blocking module used during retrieval.
Before evaluation, the environment identifies a set of tools $\mathcal{T}_{\mathrm{blk}} \subseteq \mathcal{T}$ that lie on valid solution paths and should be blocked, while keeping this information hidden from the agent.
During retrieval, if a tool $\tau \in \mathcal{T}_{\mathrm{blk}}$ is retrieved as a blocked candidate, the environment replaces it with semantically similar alternatives $\tau'$ under the standard interaction protocol.
Since each blocked tool is unavailable to the agent, any solution path that depends on at least one tool in $\mathcal{T}_{\mathrm{blk}}$ becomes infeasible from the agent's perspective.
By construction, each blocked instance preserves at least one feasible tool-call path, and therefore remains solvable.
The path-preserving selection procedure is described in Appendix~\ref{app:blocked-tool-selection}.
% Appendix~\ref{app:blocker-details} details how we select blocking tools and obtain similar alternatives.
Appendix~\ref{app:blocker-details} details the blocking pipeline, including additional tool construction in Appendix~\ref{app:noisy-tool-construction}.
We consider three types of retrieval-time blocking perturbations:
% \begin{itemize}
%     \item \textbf{Explicit Failure Blocks:}
%     the retrieved noisy tool closely resembles the blocked baseline tool in name and apparent functionality, but returns an explicit error once invoked. 
%     This creates a direct failure signal only after execution.
%     \jy{don't appear expressions like ``noisy tools''. Instead, use sth like ``the env replace the tool with a tool that will actually return explicit error message.'' Keep it short and clear.}
    
%     \item \textbf{Implicit Counterfactual Blocks:}
%     the retrieved noisy tool also closely resembles the baseline tool, but instead of returning an error, it produces a clearly unreasonable or counterfactual value. 
%     This perturbs the agent through misleading feedback rather than an explicit failure.
%     \jy{don't appear expressions like ``noisy tools''. Instead, use sth like ``the env replace the tool with a tool that will return impossible or counterfactual response normal.'' Keep it short and clear.}
    
%     \item \textbf{Semantic Distraction Blocks:}
%     the retrieved noisy tool has a name or description similar to the baseline tool, but its actual functionality is observably different. 
%     Such tools primarily introduce retrieval noise by distracting the agent with superficially relevant but functionally invalid actions.
%     \jy{The fact that the functionality is different is specified in the tool description, and the agent should be able to know it. Keep it short and clear.}
% \end{itemize}

\begin{itemize}[topsep=2pt, partopsep=3pt, leftmargin=*, itemsep=-3pt]
    \item \textbf{Explicit Failure Blocks:}
    the returned tool $\tau'$ produces an explicit error message, such as \texttt{error: endpoint unavailable}.

    \item \textbf{Implicit Failure Blocks:}
    the returned tool $\tau'$ produces an unhelpful response that silently violates its documented behavior.

    \item \textbf{Semantically Misleading Blocks:}
    the returned tool $\tau'$ has related but different functionality, making it appear usable as a substitute.
\end{itemize}

% \qh{Jiayu has changed the name from counterfactual to failure. counterfactual is not accurate. remember to change the ``counterfactual'' thing into failure for the rest of the paper.}

% \qh{And also Jiayu has changed the semantic distraction blocks into semantic misleading blocks}

For example, if~\texttt{get\_refund\_status(order\_id)} is blocked, the retriever replaces it with either an tool returning explicit error, a tool returning an~implicit wrong value such as \texttt{refund\_status=tuna}, or a semantically misleading tool such as \texttt{get\_order\_status(order\_id)} (formal definitions of the blocking mechanism in Appendix~\ref{app:blocking-formalization}).
The mechanism ensures the following:
(1) \textit{Evaluation Fairness}: Within the same setting, all models face the same blocked tools, enabling fair comparison (details are shown in Appendix~\ref{app:evaluation-fairness}).
(2) \textit{Independence Across Instances}: Blocked tools are sampled independently for each task instance, even within the same run.
(3) \textit{Controlled Difficulty}: The benchmark varies the number of replaced baseline tools, the number of injected alternatives, and the type of blocking to control disruption severity.
This setting reflects realistic tool retrieval failures: tools may be explicitly broken, silently unreliable, or semantically similar but functionally wrong. 
\bench{} therefore tests whether agents can detect these failure signals, avoid misleading tools, and recover via alternative paths.
We illustrate a blocking example in Figure~\ref{fig:demo_figure} and compare the three blocking types in Table~\ref{tab:blocking-comparison-tab}.



% \jy{1. evaluation fairness: (1) Independent across instances: tools blocked are different across each task even in the same run (2) Tools blocked are the same across different models in the same setting}
% \jy{2. Significance: one sentence. currently it is too long. }




