\section{Experiment}
% \emre{Takeaways are too general, can we add more interesting and catchy observations? }

\input{Tables/experiment_hyperparameter}

\input{Tables/main_table}

% This is a test sentence before the table.

% This is a test sentence before the table.

% This is a test sentence before the table.

% This is a test sentence before the table.

% This is a test sentence before the table.

\subsection{Settings}

\paragraph{Models.}

We evaluate both proprietary and open-source models to ensure comprehensive assessment of current frontier LLM capabilities.
Proprietary models include 
\textit{GPT}~\citep{GPT-5.4}, and \textit{Gemini}~\citep{Gemini-3.1-Pro}, while open-source models include \textit{Qwen3}~\citep{Qwen3}, \textit{Llama3}~\citep{Llama-3} and \textit{Deepseek}~\citep{DeepSeek-V4}. 
All models use a temperature of 0.0 for deterministic decoding and a max token length of 8192 to prevent truncation.

\paragraph{Metrics.}
 
We evaluate performance using metrics spanning three categories: task completion, exploration behavior and execution quality.
Together, these metrics provide a holistic view of model performance in massive-tool environments.
Full metric definitions and illustrative examples are provided in Appendix~\ref{app:full-metric-definition} and Appendix~\ref{app:metric-illustrative-examples}.

\noindent \textbf{(1) Accuracy (\%):} It measures the proportion of queries with a correct final answer.

% \jy{At evaluation time, the agent receives the question and initial evidence, and its final answer is compared against the ground-truth answer derived from the target datatype. }

\noindent \textbf{(2) Executed Ground-Truth Datatype Precision (EGT Prec.):} This metric measures the fraction of unique datatypes produced by executed tool calls that belong to the ground-truth datatype set, indicating how often execution stays relevant.
% Repeated production of the same datatype is counted only once.

\noindent \textbf{(3) Average Turns (Avg. Turns):} This metric measures the average interaction turns per query.

\noindent \textbf{(4) Mean Explored Datatypes (Mean EDT):} This metric measures the average number of new datatypes uncovered through tool retrieval beyond each query’s initial input datatypes.

\noindent \textbf{(5) Search-to-Call Ratio (S/C Ratio):} This metric quantifies the tool-retrieval/tool-call turn ratio, capturing the exploration–exploitation balance.

% \noindent \textbf{(6) Invalid Tool Call Rate (ITCR) (\%):} This metric measures the invalid tool-call rate, including unretrieved tools, incorrect arguments, etc.

\noindent \textbf{(6) Invalid Tool Call Rate (ITCR) (\%):} This metric measures the fraction of structurally or procedurally invalid calls, such as using unretrieved tools, mismatched arguments, or unavailable inputs.

\noindent \textbf{(7) Untrusted Input Rejection Rate (UIRR) (\%):} This metric measures the fraction of tool-call attempts rejected because at least one argument value is taken from a noisy tool response.


% \qh{add UIRR}

% \jy{Too long. make it shorter.}
% For blocking robustness, we additionally report \textit{Blocked Accuracy}, measured under the path1 blocking setting, and $\mathbf{\Delta}$\textit{Acc.}~(\%), computed as blocked accuracy minus default accuracy.

\paragraph{Prompts and Environment Settings.}
Table~\ref{tab:default_env_settings} summarizes the main benchmark and environment parameters, and Figure~\ref{fig:inference-prompt} shows the inference prompt.
The benchmark contains $56$ datatypes, $327$ evaluation queries, and $1{,}665$ tools.
Here, $L^*$ denotes the shortest valid solution length in tool calls ($5$--$9$), $T_{\max}=100$ denotes the interaction budget, $\Lambda_{\mathrm{ret}}^{\mathrm{cap}}=30$ denotes the maximum number of tools returned per retrieval, and $\sigma_0=42$ denotes the global random seed.




% === For environment setup ===
% You need to specify the hyperparameters you used in the previous settings.
% \jy{Note this}

% === For inference prompt ===
% Our inference prompt is provided in Figure~\ref{fig:inference-prompt}.


\subsection{Results}

% \begin{takeaway}
% Most models struggle with massive-tool planning, and success requires both discovering useful tools and executing them precisely.
% \end{takeaway}

\begin{takeaway}
Planning with massive toolsets remains highly challenging for most models.
\end{takeaway}

% \qh{update}
% \paragraph{Planning with massive toolsets remains highly challenging for most models.}
\noindent Table~\ref{tab:main_unblocked_results_grounded} shows a sharp divide between frontier models and the rest. 
\textit{Gemini-3.1-Pro} achieve the highest accuracy of 77.06\%, while maintaining the highest EGT Precision with about 20 turns. 
In contrast, most other models remain below 60\% accuracy, with \textit{Qwen3-8B} and \textit{Llama-3.1-8B-Instruct} achieving 0\%, revealing the difficulty of long-horizon, exploration-driven planning across large tool ecosystems.
The gaps are also large within model families: larger Qwen and Llama variants outperform their smaller counterparts, and full frontier models far exceed their lightweight versions such as \textit{Gemini-3.5-Flash} and \textit{GPT-5.4-Mini}. 
Together, these results suggest that both model family and scale matter for this task.
Robustness checks, including confidence intervals, are reported in Appendix~\ref{app:evaluation-robustness}.

% \input{Figures/noise_combined}

\begin{takeaway}
Tool discovery requires broad, accurate, and bi-directional exploration.
\end{takeaway}

% \qh{2. second paragraph should be: exploration tendency (measured by Mean EDT) don't always means high acc.}

\paragraph{Exploration tendency strongly relates to task success.}
Broad tool retrieval can expose agents to more potential intermediate information, making exploration tendency strongly associated with task success.
We measure this exploration tendency using Mean EDT, the average number of new datatypes uncovered through retrieval beyond the query's initial datatypes. 
Across models, Mean EDT is strongly correlated with accuracy (Pearson coefficient~\citep{Pearson} $r=0.902$), suggesting that agents uncover more intermediate information are generally more likely to complete the task successfully. 
However, exploration tendency does not explain performance entirely. 
For example, \textit{Llama-3.3-70B-Instruct} achieves a Mean EDT score comparable to \textit{GPT-5.4} (19.20 vs. 20.65), yet the two models differ substantially in accuracy (18.96\% vs. 51.90\%).
This indicates that broad retrieval is an important driver of success, but it remains only part of effective long-horizon tool~use. 

% \paragraph{Weak models exhibit different failure modes in exploration.}
% The exploration metrics show that broader search does not always lead to higher success. 
% Some models explore a large tool and datatype space but still fail to convert this exploration into accurate task completion. 
% For example, \textit{Llama-3.3-70B-Instruct} and \textit{Gemini-3.1-Flash-Lite} have high Mean EDT values (23.88 and 17.43) and elevated search-to-call ratios (2.10 and 4.50), yet their accuracies remain only 44.95\% and 34.56\%. 
% A similar pattern appears within the \textit{DeepSeek} family. 
% \textit{DeepSeek-V4-Flash} explores slightly more broadly than \textit{DeepSeek-V4-Pro} (28.16 vs. 27.84 Mean EDT), but its accuracy is more than 17 points lower.
% Other models fail for the opposite reason. 
% \textit{GPT-5.4-Mini} uses the fewest turns on average (11.85), but reaches only 16.51\% accuracy and 0.165 EGT recall, suggesting that it often stops before finding the right solution path.
% Together, these results point to two common failure modes. 
% Some models explore too little to identify the needed tools, while others explore broadly but fail to organize the retrieved tools into an effective plan.



% \qh{the following 4 paragraph have already been updated}


\paragraph{Frequent retrieval does not guarantee effective exploration.}
A high S/C Ratio and a large number of interaction turns indicate that an agent spends substantial effort on retrieval, but such effort does not necessarily translate into broad discovery of useful intermediate information.
For example, \textit{Gemini-3.5-Flash} has the highest S/C Ratio among all models (10.44) and also uses the most turns on average (57.87).
However, \textit{Gemini-3.1-Pro} achieves a higher Mean EDT while using only about one third of the turns and one seventh of the S/C Ratio.
This contrast shows that frequent searching and long interactions are not sufficient for effective exploration.
An agent may search proactively, but if these searches repeatedly revisit unuseful or uninformative tools, they contribute little to the discovery of new task-relevant datatypes.


% \paragraph{Stronger agents explore more efficiently.}
% In contrast, the stronger agents achieve higher Mean EDT with substantially lower S/C ratios. 
% \textit{Gemini-3.1-Pro} uses only 19.55 turns on average with an S/C Ratio of 1.59, but discovers 27.41 datatypes on average. 
% Similarly, \textit{DeepSeek-V4-Flash} reaches a Mean EDT of 25.34 with 31.41 turns and an S/C Ratio of 2.80. 
% These results show that strong agents do not simply search more; rather, they issue more diverse and broad retrieval requests, better interpret the returned tools, and more effectively expand the useful information frontier.

% \begin{takeaway}
% Effective tool discovery correlates with balancing forward anticipation from evidence and backward anticipation from goals.
% \end{takeaway}

\paragraph{Effective tool discovery requires bi-directional anticipation.}
We define the forward/backward retrieval ratio (F/B ratio) as the total number of input-conditioned retrievals divided by the total number of output-conditioned retrievals across all query traces for a model. 
Across models, agents generally issue more input-conditioned than output-conditioned retrievals, indicating a common preference for forward anticipation. 
Lower-performing models, such as \textit{Llama-3.1-8B-Instruct} and \textit{Qwen3-14B}, rely heavily on input-conditioned retrieval, yielding F/B ratios of 16.56 and 14.18, respectively. 
This suggests that retrieving tools compatible with currently available inputs is often insufficient: these agents may identify executable next steps, but fail to reason backward about which intermediate datatypes must be discovered to reach the final goal.
This pattern is further supported by correlation analysis, where the relative frequency of output-conditioned retrieval is strongly correlated with accuracy (Pearson $r=0.800$). 
Together, these results suggest that effective tool discovery requires agents to combine forward exploration from current evidence with backward anticipation from the desired outcome.

% \paragraph{Forward exploration alone is insufficient.}
% We define the \textit{forward/backward retrieval ratio} (F/B ratio) as the total number of input-conditioned retrievals divided by the total number of output-conditioned retrievals across all query traces for a model.
% Across models, agents generally issue more input-conditioned than output-conditioned retrievals, indicating a common preference for forward anticipation.
% Lower-performing models, such as \textit{Llama-3.1-8B-Instruct} and \textit{Qwen3-14B}, rely heavily on input-conditioned retrieval, yielding F/B ratios of 16.56 and 14.18, respectively.
% This suggests that retrieving tools compatible with currently available inputs is often insufficient: these agents may identify executable next steps, but fail to target the missing intermediate outputs needed to advance toward the final goal.
% Thus, excessive reliance on forward exploration is associated with weaker performance, highlighting the role of backward anticipation in guiding effective tool discovery.

% \paragraph{Stronger agents better balance forward and backward anticipation.}
% Stronger agents make substantially greater use of output-conditioned retrieval, resulting in F/B ratios closer to parity.
% % \qh{gemini-3.1-pro: 1.33; gpt-5.4: 0.86; deepseek-v4-flash: 2.10}
% For example, \textit{Gemini-3.1-Pro} and \textit{GPT-5.4} both achieve an F/B ratio of approximately 1, reflecting a relatively balanced use of forward and backward anticipation.
% This pattern is further supported by the correlation analysis: the relative frequency of output-conditioned retrieval is strongly correlated with accuracy (Pearson $r=0.800$).
% Together, these results suggest that effective tool discovery requires agents to reason in both directions: exploring forward from current evidence while also reasoning backward from the desired outcome to identify missing intermediate information and tools.



\begin{takeaway}
Successful long-horizon tool use requires both effective exploration and accurate exploitation.
\end{takeaway}

\noindent Beyond effective exploration, successful agents must accurately exploit the information they uncover, which EGT Precision captures by measuring whether executed tool calls stay on task-relevant paths. 
EGT Precision is highly correlated with accuracy (Pearson coefficient $r=0.781$), indicating that models are much more likely to succeed when they execute relevant tool-use trajectories.
The strongest models further support this pattern: \textit{Gemini-3.1-Pro} achieve the highest accuracies (77.06\%) while also attaining the highest EGT Precision (91.47\%). 
These results suggest that effective long-horizon tool use requires not only sufficient exploration, but also precise execution over the explored tool space.

\begin{takeaway}
Basic tool-use reliability remains necessary.
\end{takeaway}

\noindent Accurate exploration and exploitation also depend on reliable tool-use. 
Accuracy is negatively correlated with ITCR (Pearson coefficient $r=-0.443$), showing that models that frequently make invalid tool calls are much less likely to complete the task successfully. 
For example, \textit{Llama-3.1-8B-Instruct} has the second highest ITCR and the lowest accuracy, while the strongest model, \textit{Gemini-3.1-Pro}, keep ITCR near zero (0.68\%). 
This suggests that effective long-horizon tool-use requires not only broad exploration and accurate exploitation, but also basic reliability in invoking tools with valid arguments.
The moderate correlation also indicates that invalid tool calls alone do not explain performance differences, and that failures also arise from ineffective exploration and exploitation.
