\section*{Limitations}
\label{sec:limitations}
\paragraph{Teacher-bounded zone.}
The most fundamental limitation of ZPPO is the one we identify in Sec.~\ref{sec:conclusion}: BCQ requires the teacher to succeed on a hard question, so questions on which both teacher and student fail admit no BCQ candidate and leave only NCQ, whose contribution alone is more limited (Tab.~\ref{tab:ablation}, Sec.~\ref{sec:experiments:ablation}). At the upper end of the student-capability range, BCQ's contribution shrinks accordingly and the available \emph{zone} effectively collapses to the NCQ-only regime. Extending the zone beyond what the current teacher covers -- via curated synthetic prompts, ensembles of teachers, or curriculum-aware question selection -- is, in our view, the most important open problem this work leaves behind.

\paragraph{Tension with dynamic sampling.}
Dynamic sampling~\citep{yu2025dapo} simply deletes any question where the student got every rollout correct or every rollout wrong. On the other hand, ZPPO does the opposite: those all-wrong questions are exactly the ones we store in the replay buffer, so that BCQ and NCQ can rewrite the prompt and give the student a fresh chance. Hence, naively combining ZPPO and dynamic sampling is not trivial. A natural fix is to run them in sequence: first identify hard questions and only call dynamic sampling if their BCQ- and NCQ-reformulated rollouts still come back all-correct or all-wrong -- that is, apply dynamic sampling on top of BCQ and NCQ. We leave this hybrid algorithm to future work.

\paragraph{Scope beyond reasoning correctness.}
ZPPO post-trains an already-built compact VLM and is orthogonal to how such small models are first constructed -- via architectural designs that compress visual--linguistic computation into small backbones~\citep{lee-etal-2024-collavo,10.1007/978-3-031-72967-6_16,NEURIPS2024_473a9a75,lee-etal-2024-trol,lee2024phantom} or via large-to-small distillation of strong teachers~\citep{lee2025genrecal,lee2025building,yu2026hide} -- so it could in principle be stacked on any of these as a later post-training stage. Several other axes are deliberately out of scope. Extending the recipe to multi-step and agentic multimodal reasoning~\citep{Lee_2026_CVPR_Recursive,kang2026agent,cho2026spatialclawrethinkingactioninterface} is open; we neither evaluate nor train for robustness, which carries its own substantial literature on robust and causal training~\citep{lee2021towards,NEURIPS2021_8e5e15c4,Lee_2022_CVPR,Kim_2023_CVPR,Lee_2023_ICCV}; we do not consider inference-time efficiency techniques such as visual-token pruning~\citep{kim2026and} that are orthogonal to the training recipe; we do not target conversational or social competence~\citep{leeenhancing}; we do not touch the upstream perception stacks~\citep{lee2020training} that compact models ultimately depend on for embodied deployment; and our evaluation targets single-turn reasoning accuracy rather than complementary axes such as multi-sensor perception~\citep{yu2024spark}, multi-turn dialogue~\citep{Lee_2025_ICCV}, or iterative self-refinement~\citep{lee2026refinebench}. Combining ZPPO with these complementary directions is left to future work.



