\section{Benchmark Statistics}
\label{app:benchmark_statistics}

\subsection{Task and Grader Statistics}
\label{app:user_level_task_grader_statistics}

In this subsection, we summarize the task and grader statistics for the 100 tasks across all users, as shown in Tab.~\ref{tab:benchmark_user_statistics}.
\begin{table}[htbp]
\centering
\small
\setlength{\tabcolsep}{5pt}
\caption{Task and Grader Statistics.}
\label{tab:benchmark_user_statistics}
\begin{tabular}{cccccc}
\toprule
\textbf{User} & \textbf{Task} & \textbf{Session} & \textbf{Hidden Intent} & \textbf{Rubric} & \textbf{\shortstack{Rule}} \\
\midrule
Researcher & 20 & 20 & 84 & 125 & 31 \\
Marketer & 20 & 20 & 74 & 58 & 11 \\
Law Trainee & 20 & 20 & 89 & 131 & 43 \\
Pharmacist & 20 & 20 & 195 & 85 & 25 \\
Financier & 20 & 20 & 82 & 111 & 58 \\
\midrule
\textbf{Total} & \textbf{100} & \textbf{100} & \textbf{524} & \textbf{510} & \textbf{168} \\
\bottomrule
\end{tabular}
\end{table}


\subsection{Tool and Skill Statistics}
\label{app:tool_skill_inventory}

Tabs.~\ref{tab:benchmark_tools} and \ref{tab:benchmark_skills} summarize the tool and skill inventory of the benchmark. In total, the benchmark contains 187 unique tools and 21 skills. 

\par\medskip
\begingroup
\footnotesize
\setlength{\LTpre}{8pt}
\setlength{\LTpost}{0pt}
\setlength{\tabcolsep}{3pt}
\begin{longtable}{@{}p{0.15\linewidth}p{0.27\linewidth}p{0.50\linewidth}@{}}
\caption{Benchmark tools grouped by application domain.}
\label{tab:benchmark_tools}\\
\toprule
\textbf{Domain} & \textbf{Tool} & \textbf{Function} \\
\midrule
\endfirsthead

\toprule
\textbf{Domain} & \textbf{Tool} & \textbf{Function} \\
\midrule
\endhead

\endfoot

\bottomrule
\endlastfoot

Amazon & Add Address & Add a saved address in Amazon. \\
 & Add Product to Cart & Add a product to the cart in Amazon. \\
 & Add Product to Wish List & Add a product to the wish list in Amazon. \\
 & Add to History & Add a product to browsing history in Amazon. \\
 & Clear Cart & Remove all items from the shopping cart in Amazon. \\
 & Log In & Authenticate in Amazon. \\
 & Log Out & Sign out of Amazon. \\
 & Move Cart Item to Wish List & Move a cart item to the wish list in Amazon. \\
 & Move Wish List Item to Cart & Move a wish-list item to the cart in Amazon. \\
 & Place Order & Complete a purchase in Amazon. \\
 & Post Product Question & Post a product question in Amazon. \\
 & Search Product Types & Search product categories in Amazon. \\
 & Search Products & Search products in Amazon. \\
 & Search Sellers & Search sellers in Amazon. \\
 & Update Cart Quantity & Change product quantity in the cart in Amazon. \\
 & View Account & View account information in Amazon. \\
 & View Addresses & View saved addresses in Amazon. \\
 & View Cart & View the shopping cart in Amazon. \\
 & View Last Purchase & View the most recent product purchase in Amazon. \\
 & View Order & View an order in Amazon. \\
 & View Orders & View orders in Amazon. \\
 & View Payment Card & View a stored payment card in Amazon. \\
 & View Payment Cards & View stored payment cards in Amazon. \\
 & View Product & View product in Amazon. \\
 & View Product Options & View available options for a product in Amazon. \\
 & View Product Purchases & View past purchases of a product in Amazon. \\
 & View Product Q\&A & View product questions and answers in Amazon. \\
 & View Product Questions & View product questions in Amazon. \\
 & View Product Reviews & View product reviews in Amazon. \\
 & View Recommendations & View recommended products in Amazon. \\
 & View Seller & View seller information in Amazon. \\
 & View Wish List & View the wish list in Amazon. \\
 & Write Product Review & Write a product review in Amazon. \\
\midrule
Gmail & Archive Thread & Archive an email thread in Gmail. \\
 & Create Draft & Create an email draft in Gmail. \\
 & Forward Thread & Forward an email thread in Gmail. \\
 & Label Thread & Apply a label to an email thread in Gmail. \\
 & Log In & Authenticate in Gmail. \\
 & Log Out & Sign out of Gmail. \\
 & Mark Read & Mark an email thread as read in Gmail. \\
 & Mark Unread & Mark an email thread as unread in Gmail. \\
 & Remove Label & Remove a label from an email thread in Gmail. \\
 & Reply to Email & Reply to an email in Gmail. \\
 & Search Labels & Search labels in Gmail. \\
 & Search Users & Search users in Gmail. \\
 & Send Draft & Send a previously saved draft in Gmail. \\
 & Send Email & Send an email in Gmail. \\
 & Sign Up & Create an account in Gmail. \\
 & Star Thread & Star an email thread in Gmail. \\
 & Update Draft & Edit an email draft in Gmail. \\
 & View Category Sizes & View category sizes in Gmail. \\
 & View Draft & View draft in Gmail. \\
 & View Drafts & View drafts in Gmail. \\
 & View Email & View email in Gmail. \\
 & View Inbox Threads & View inbox threads in Gmail. \\
 & View Outbox Threads & View outbox threads in Gmail. \\
 & View Starred Threads & View starred threads in Gmail. \\
 & View Thread & View thread in Gmail. \\
\midrule
Phone & Add Contact & Add contact in Phone. \\
 & Create Alarm & Create an alarm in Phone. \\
 & Get Date and Time & Retrieve the current date and time from Phone. \\
 & Log In & Authenticate in Phone. \\
 & Search Contacts & Search contacts in Phone. \\
 & Search Messages & Search text messages in Phone. \\
 & Send Text Message & Send a text message in Phone. \\
 & Update Alarm & Update an alarm in Phone. \\
 & Update Contact & Update contact in Phone. \\
 & View Alarm & View an alarm in Phone. \\
 & View Message Thread & View a text-message thread in Phone. \\
 & View Text Message & View a text message in Phone. \\
\midrule
Simple Note & Add Content to Note & Add content to note in Simple Note. \\
 & Create Note & Create note in Simple Note. \\
 & Delete Note & Delete note in Simple Note. \\
 & Log In & Authenticate in Simple Note. \\
 & Log Out & Sign out of Simple Note. \\
 & Search Notes & Search notes in Simple Note. \\
 & Update Note & Update note in Simple Note. \\
 & View Account & View account information in Simple Note. \\
 & View Note & View note in Simple Note. \\
 & View Profile & View profile in Simple Note. \\
\midrule
Splitwise & Add Group Member & Add a member to a group in Splitwise. \\
 & Create Group & Create group in Splitwise. \\
 & Delete Expense & Delete expense in Splitwise. \\
 & Log In & Authenticate in Splitwise. \\
 & Record Expense & Record an expense in Splitwise. \\
 & Search Users & Search users in Splitwise. \\
 & Update Expense & Update expense in Splitwise. \\
 & View Group & View group in Splitwise. \\
 & View Group Balance & View the balance of a group in Splitwise. \\
 & View Group Expenses & View expenses recorded in a group in Splitwise. \\
 & View Groups & View groups in Splitwise. \\
 & View Member Balances & View per-member balances in Splitwise. \\
 & View Person Balance & View the balance of one person in Splitwise. \\
\midrule
Spotify & Add Song to Playlist & Add a song to a playlist in Spotify. \\
 & Add to Queue & Add an item to the playback queue in Spotify. \\
 & Create Playlist & Create playlist in Spotify. \\
 & Follow Artist & Follow an artist in Spotify. \\
 & Log In & Authenticate in Spotify. \\
 & Log Out & Sign out of Spotify. \\
 & Play Music & Start music playback in Spotify. \\
 & Previous Song & Return to the previous song in Spotify. \\
 & Remove Downloaded Song & Remove a downloaded song in Spotify. \\
 & Remove Song from Playlist & Remove a song from a playlist in Spotify. \\
 & Remove Song from Queue & Remove a song from the playback queue in Spotify. \\
 & Review Song & Write a song review in Spotify. \\
 & Save Album & Save an album in Spotify. \\
 & Save Album to Library & Save an album to the library in Spotify. \\
 & Save Playlist & Save a playlist in Spotify. \\
 & Save Song & Save a song in Spotify. \\
 & Save Song to Library & Save a song to the library in Spotify. \\
 & Search Albums & Search albums in Spotify. \\
 & Search Artists & Search artists in Spotify. \\
 & Search Playlists & Search playlists in Spotify. \\
 & Search Songs & Search songs in Spotify. \\
 & Search Users & Search users in Spotify. \\
 & Update Song Review & Edit a song review in Spotify. \\
 & View Album & View album in Spotify. \\
 & View Album Library & View album library in Spotify. \\
 & View Artist & View artist in Spotify. \\
 & View Current Song & View the currently playing song in Spotify. \\
 & View Followed Artists & View artist following in Spotify. \\
 & View Genres & View genres in Spotify. \\
 & View Liked Songs & View liked songs in Spotify. \\
 & View Playlist & View playlist in Spotify. \\
 & View Playlist Library & View playlist library in Spotify. \\
 & View Queue & View the playback queue in Spotify. \\
 & View Song & View song in Spotify. \\
 & View Song Library & View song library in Spotify. \\
 & View Song Review & View a song review in Spotify. \\
 & View Song Reviews & View song reviews in Spotify. \\
\midrule
Todoist & Add Task Label & Add a label to a task in Todoist. \\
 & Assign Task & Assign or unassign a task in Todoist. \\
 & Create Label & Create label in Todoist. \\
 & Create Project & Create project in Todoist. \\
 & Create Section & Create section in Todoist. \\
 & Create Subtask & Create a subtask in Todoist. \\
 & Create Task & Create task in Todoist. \\
 & Delete Label & Delete label in Todoist. \\
 & Delete Notifications & Delete notifications in Todoist. \\
 & Delete Project & Delete project in Todoist. \\
 & Delete Section & Delete section in Todoist. \\
 & Delete Subtask & Delete a subtask in Todoist. \\
 & Delete Task & Delete task in Todoist. \\
 & Delete Task Comment & Delete a task comment in Todoist. \\
 & Log In & Authenticate in Todoist. \\
 & Log Out & Sign out of Todoist. \\
 & Mark Notification & Mark a notification in Todoist. \\
 & Post Task Comment & Post a comment on a task in Todoist. \\
 & Remove Task Label & Remove a label from a task in Todoist. \\
 & Search Labels & Search labels in Todoist. \\
 & Search Users & Search users in Todoist. \\
 & Update Project & Update project in Todoist. \\
 & Update Section & Update section in Todoist. \\
 & Update Subtask & Update a subtask in Todoist. \\
 & Update Task & Update task in Todoist. \\
 & Update Task Comment & Update a task comment in Todoist. \\
 & View Notification Count & View the notification count in Todoist. \\
 & View Notifications & View notifications in Todoist. \\
 & View Project & View project in Todoist. \\
 & View Projects & View projects in Todoist. \\
 & View Sections & View sections in Todoist. \\
 & View Subtasks & View subtasks in Todoist. \\
 & View Task & View task in Todoist. \\
 & View Task Comment & View a task comment in Todoist. \\
 & View Task Comments & View task comments in Todoist. \\
 & View Tasks & View tasks in Todoist. \\
\midrule
File System & Copy Directory & Copy a directory in the File System. \\
 & Copy File & Copy a file in the File System. \\
 & Create Directory & Create a directory in the File System. \\
 & Create File & Create a file in the File System. \\
 & Log In & Authenticate in File System. \\
 & Move Directory & Move a directory in the File System. \\
 & Move File & Move a file in the File System. \\
\midrule
General & Ask User & Pause execution and ask the user for required input. \\
 & Read File & Read text, image, or document content from a file. \\
 & Write File & Write or overwrite file content. \\
 & Edit File & Make targeted edits by replacing matched text in a file. \\
 & List Directory & List directory contents, optionally recursively. \\
 & Glob Search & Find files by glob pattern. \\
 & Grep Search & Search file contents with pattern matching. \\
 & Edit Notebook & Edit cells in a Jupyter notebook. \\
 & Shell Exec & Run a shell command and return its output. \\
 & Send Message & Send a message or file attachment back to the user. \\
 & Spawn Subagent & Launch a background subagent for an independent task. \\
 & Schedule Task & Add, list, or remove scheduled reminders and recurring tasks. \\
\midrule
Web & Web Search & Search the web and return titles, URLs, and snippets. \\
 & Web Fetch & Fetch a URL and extract readable page content. \\

\end{longtable}
\endgroup

\flushbottom

\par\medskip
\begingroup
\footnotesize
\setlength{\LTpre}{8pt}
\setlength{\LTpost}{0pt}
\setlength{\tabcolsep}{4pt}
\begin{longtable}{@{}p{0.32\linewidth}p{0.63\linewidth}@{}}
\caption{Skill information of the benchmark.}
\label{tab:benchmark_skills}\\
\toprule
\textbf{Skill} & \textbf{Function} \\
\midrule
\endfirsthead

\toprule
\textbf{Skill} & \textbf{Function} \\
\midrule
\endhead

\endfoot

\bottomrule
\endlastfoot

Jargon Translator & Converts plain language into workplace jargon and translates jargon back into plain speech. \\
PubMed Search & Retrieves biomedical literature in a PubMed-style workflow for pharmacy and oncology tasks. \\
Local File Grounding & Grounds responses in local task files such as TXT, CSV, JSON, SVG, and Markdown artifacts. \\
PDF Reader & Extracts and grounds answers in local PDF content, including figures, legends, and methods. \\
JSON Translator & Translates text fields inside JSON files, especially description-heavy structured data. \\
Customer Requirement Analysis & Analyzes investor communication materials and turns them into standardized advisory-needs reports. \\
iFinD Finance Data Search & Retrieves market, fund, futures, and macroeconomic data through natural-language finance queries. \\
China Tax Law & Assists with Chinese tax-law research, compliance checks, planning, and dispute analysis. \\
Python Data Analysis & Provides lightweight Python-based data cleaning, statistical analysis, and visualization guidance. \\
Data Analysis & Turns raw data into reports, visualizations, and decision-oriented summaries. \\
Planning & Turns goals and constraints into sequenced execution plans with priorities and time blocks. \\
News Aggregator & Collects and summarizes domestic and international news across multiple topics. \\
Defense Lawyer & Supports criminal-defense analysis, evidence assessment, strategy design, and drafting. \\
Corporate Lawyer & Supports transaction-oriented legal work such as contract review, compliance checks, and risk assessment. \\
Proactive Task Validator & Validates whether hidden intents, dependencies, and checklists are aligned in proactive-task datasets. \\
One-Page Credit Analysis & Structures a one-page credit view for Chinese-market financial analysis and risk review. \\
Financial Analysis & Supports portfolio analysis, risk attribution, and backtesting-oriented financial reporting. \\
Web Browsing & Checks authoritative public webpages when local context alone is insufficient. \\
Local Web Search & Runs targeted web search for product, formulation, and market-comparison tasks. \\
Medical Literature Reader Pro & Reads, compares, and synthesizes biomedical papers and evidence chains. \\
Law Exam Trainer & Builds and explains law-exam practice materials from videos or documents. \\

\end{longtable}
\endgroup
























\subsection{Taxonomy of Benchmark Tasks} % runquan
\label{app:taxonomy_of_benchmark_tasks}

In this subsection, we detail the categorization of the 100 benchmark tasks designed for our evaluation. To systematically assess the multimodal agents across diverse professional and daily scenarios, we group the tasks into 18 fine-grained categories. Rather than relying on broad topical divisions, this classification system strictly emphasizes the specific action intents, reasoning requirements, and underlying workflows inherent to each task. A comprehensive overview of these categories, along with their corresponding descriptions and covered task lists, is presented in Tab.~\ref{tab:task_taxonomy}.

\input{0.1-table/task_category}