\clearpage
\section{Case Study}
\label{app:case_study}

\subsection{Hidden Intents}

\paragraph{Failure case: Researcher -- Claude 4.6 Opus.}
Fig.~\ref{fig:case_zhanghaoran_1_hidden_intent_trigger} and Fig.~\ref{fig:case_zhanghaoran_1_hidden_intent_resolution} show a representative hidden-intent case from the Researcher user episode evaluated with Claude 4.6 Opus.
The session begins from an environment-triggered paper recommendation feed, where the relevant requirement is not stated as a direct user instruction.
The agent first handles the feed as a broad recommendation task, and only later focuses on the OpenClaw-related subset after the user explicitly asks for it.
The key signal is the hidden-intent assignment: several intents are marked as {\provided} rather than {\completed} or {\inferred}, indicating that the user had to surface requirements that the agent did not proactively resolve.
Although the final trajectory has moderate completeness, this case mainly illustrates weak proactive intent resolution under an underspecified trigger.

\input{0.0-figure/case_zhanghaoran_1}

\paragraph{Contrast case: Researcher -- DeepSeek V3.2 and Claude 4.6 Opus.}
Fig.~\ref{fig:case_zhangshunkai_1_deepseek_task} and Fig.~\ref{fig:case_zhangshunkai_1_deepseek_trajectory} show the DeepSeek V3.2 trace, while Fig.~\ref{fig:case_zhangshunkai_1_claude_task} and Fig.~\ref{fig:case_zhangshunkai_1_claude_trajectory} show the Claude 4.6 Opus trace for the same Researcher meal-planning task.
Both runs receive the same initial request, which explicitly mentions only a one-week meal plan and the RMB 20--30 price constraint, while the user's body profile, macro accounting, table structure, and Plan B requirements remain hidden but recoverable from prior context.
We focus on the first assistant turn because it occurs before the user reveals any hidden requirements and therefore best isolates proactive recovery from visible-request following.
DeepSeek V3.2 mostly follows the visible budget and availability cue in this first response; although it recognizes campus-accessible food options, it does not recover the prior-session body profile, table preference, per-meal macro requirement, or Plan B preference, so the user has to reveal several requirements across later turns.
Claude 4.6 Opus captures more task structure immediately: it goes beyond the visible price constraint by imposing a readable table, listing concrete foods for each day and meal, and grounding choices in campus or nearby meal availability.
However, Claude still requires later user turns for macro totals, per-meal P/F/C values, and fallback meals.
The contrast is a stronger-versus-weaker proactivity comparison rather than a binary success-failure case: both trajectories improve after user intervention, but their scores differ according to how much hidden intent is resolved before the user supplies it.

\input{0.0-figure/case_zhangshunkai_1}

\paragraph{Success case: Marketer -- GPT-5.4.}
Fig.~\ref{fig:case_zhangshunkai_2_targeted_followup_task} and Fig.~\ref{fig:case_zhangshunkai_2_targeted_followup_trajectory} show a targeted-followup success case from a Marketer crisis-communication episode evaluated with GPT-5.4.
The initial trigger is a system webhook requesting a final X apology letter after a prior client-alignment session, but it does not restate the approved engineering cause, compensation language, or full incident scale.
Instead of fabricating a publishable statement from incomplete context, the agent asks focused questions about the missing approval facts.
The user-agent tracking marks the matched targeted-followup indexes for the engineering explanation and compensation as {\inferred}, meaning the agent's clarification directly elicited hidden intents that would otherwise have remained unstated.
After the user supplies those two requirements, the next response states the approved cause and compensation while also incorporating the incident-scale facts recovered from the prior context, reaching perfect \textsc{Comp} and \textsc{Proc}.
This case illustrates a successful proactive pattern: a targeted clarification can count as intent inference when it asks for the specific missing fact needed to complete a high-stakes artifact.

\input{0.0-figure/case_zhangshunkai_2}

\paragraph{Success case: Pharmacist -- Kimi K2.5.}
Fig.~\ref{fig:case_zhangshunkai_3_targeted_followup_task} and Fig.~\ref{fig:case_zhangshunkai_3_targeted_followup_trajectory} show a targeted-followup case from a Pharmacist sandbox-Gmail workflow evaluated with Kimi K2.5.
The initial request asks the agent to read a local booking brief, send a simulated Gmail message, and verify the send, but the available tool set reported by the agent does not expose Gmail functions.
In the first turn, the agent asks a focused question about which email-sending method to use, and the user-agent marks the direct-send workflow intent as {\inferred}.
The same response also recovers key brief-derived message requirements by drafting an instrument-availability request with Tuesday/Thursday afternoon availability, the 4-hour LC-MS slot, and the 30-minute sample-preparation note, although later checklist scoring still penalizes exact block content and tool evidence.
This case is useful because it separates hidden-intent proactivity from final task success: the targeted follow-up and partial brief-detail recovery raise \textsc{Proc}, while the final checklist remains low because the Gmail login, direct send, and outbox confirmation are not supported by tool evidence.

\input{0.0-figure/case_zhangshunkai_3}

\subsection{Checklist}
% 找一个适合展示的 task turn + 1 个 checklist 较高的 model output +  1 个 checklist 较低的  model output（同一 task turn 的不同模型 output）
\paragraph{Success case: Law Trainee -- GPT-5.4.}
Fig.~\ref{fig:case_leihaodi_1_checklist_task} shows the shared task and checklist targets, and Fig.~\ref{fig:case_leihaodi_1_gpt54_trajectory} shows the higher-scoring GPT-5.4 trajectory from the Law Trainee user episode.
The GPT-5.4 run shows a high-completeness checklist pattern: after the missing handover details are surfaced during interaction, the agent sends a consolidated SMS that covers the pickup slot, required originals, reply deadline, and risk-flag instruction, while also creating and deleting the temporary Todoist board with the required reminder content.
The remaining misses are mainly strict tool-string checks, so this case illustrates how checklist evaluation can separate strong end-task completion from exact tool-evidence failures.

\paragraph{Failure case: Law Trainee -- Kimi K2.5.}
Fig.~\ref{fig:case_leihaodi_1_kimi_trajectory} shows the lower-scoring Kimi K2.5 trajectory on the same Yuehai court-side handover task.
The Kimi K2.5 trajectory completes some visible actions, including sending SMS messages and briefly using Todoist, but it leaves several checklist requirements unsupported by the final trace.
The handover information is split across messages rather than preserved as one compact phone-readable SMS, the reply deadline and risk-flag instruction are absent, and the Todoist reminder content does not cover the annotated 10:45 packet-completeness and stamped-page checks.
This contrast highlights why checklist-based grading is necessary in addition to surface-level task progress: both agents interact with the same tools, but only the higher-scoring trajectory preserves the operational details needed for a reliable handover.

\input{0.0-figure/case_leihaodi_1}

\subsection{Cross-Session Dependency Design}
% 找一个适合展示的 task group + 1 个 proactivity and completeness 较高的 model trace + 1 个 proactivity and completenes 较低的 model trace（同一 task 的不同模型 trace，包含多条 input / output）
% 其中一个 trace 很好地捕捉到了铺垫任务的 intents 而更好更快地完成了该条测试任务，另一个 trace 则没有很好地完成这一点
\paragraph{Success case: Researcher -- Claude 4.6 Opus.}
Fig.~\ref{fig:case_leihaodi_2_dependency_task} shows a cross-session dependency case from the Researcher user episode, where the current request refers only to the user's research theme.
The relevant theme and output conventions were established in earlier sessions: the user had explored think-with-image multimodal agents, preferred a small batch of papers, and expected links, open-source status, and concrete follow-up paths.
As shown in Fig.~\ref{fig:case_leihaodi_2_claude_trajectory}, Claude 4.6 Opus carries over the research theme and several output conventions, allowing it to complete much of the target paper-selection and metadata checklist within three turns.

\paragraph{Failure case: Researcher -- Kimi K2.5.}
Fig.~\ref{fig:case_leihaodi_2_kimi_trajectory} shows the lower-scoring Kimi K2.5 trajectory on the same task.
Instead of recovering the user's research theme and presentation preferences from prior sessions, the agent repeatedly waits for the user to restate them.
The resulting conversation is longer, and the selected papers drift away from the annotated target set, leading to low proactivity and zero checklist completion despite several rounds of user-provided requirements.

\input{0.0-figure/case_leihaodi_2}

% \subsection{Edge Cases}
% 边界情况怎么处理，例如 agent asks a generic question 是否算 inferred，agent partially satisfies an intent 是否算 completed，agent produces an artifact before asking 是否优先 counted as completed，user reveals intent after irrelevant response 是否 counted as provided
