

\section{Related Work}

\paragraph{Agent safety benchmarks.} As LLMs are increasingly deployed in planning, tool use, and long-horizon execution, safety concerns have expanded from traditional content safety to agentic behavioral safety. In such settings, safety failures may arise from unsafe tool use, incorrect state modifications, long-horizon execution errors, and adversarial interactions with users or environments \citep{ghosh2025safety, rjudge2024, zhang2024agentsafetybench, tur2025safearena, guo2025upwarddeceivers, verma2024operationalizing}. Prior work has proposed agent-oriented risk taxonomies that extend safety analysis to adversarial, ethical, physical, and tool-augmented behavioral risks \citep{agentharm2024, zheng2024ali,luo2025agentauditor}. Building on these foundations, recent benchmarks evaluate interactive agent behaviors from single-turn moderation to multi-step tool-augmented execution \citep{rjudge2024, zhang2024agentsafetybench, os-harm2025, riosworld2025, injecagent2024, toolsword2024, safetoolbench2025, agentdojo2024, tur2025safearena, li2026agentdyn, zhang2025agent, chen2026decodingtrustagentplatformdtapcontrollable}. However, existing benchmarks remain limited in both risk coverage and scalability: most cover only subsets of agentic safety risks, while many evaluation protocols rely heavily on scenario-specific red teaming or manual judgment, making large-scale agentic safety evaluation and reinforcement learning prohibitively expensive.


\paragraph{Safety data and environment for agentic training.} 
Beyond evaluation, agentic safety training requires safety-relevant interaction data that exposes risk-bearing behaviors and provides supervision over multi-step decisions. Real rollout trajectories can capture realistic agent failures, but are costly to collect, difficult to scale, and constrained by privacy and safety considerations. To improve scalability, recent work synthesizes agent trajectories for generic tool-use interactions \citep{li2025domain,toolace2025,kimi2025,auragen2025} and safety-aware scenarios \citep{xie-etal-2025-toolsafety,zhang2025agentalign,mou2026toolsafe}. Beyond static trajectory data, interactive environments provide training signals through executable feedback or simulated interaction dynamics. Code-based sandboxes provide executable tasks and objective reward signals \citep{wang2026agent,song2026envscaler,gao2026self,guo2025genenv}, while LLM-simulated or role-playing environments emulate users, tools, servers, and environment feedback to generate more open-ended trajectories \citep{li2025close,chen2025scaling,li2025simulating}. However, these pipelines are not primarily tailored to safety-oriented data construction, leaving open how to provide systematic risk coverage and structured trajectory-level supervision for scalable safety optimization.


\paragraph{Agent guardrail.} A growing body of works have developed guardrail to detect and mitigate unsafe behaviors in LLMs and agentic systems. Early guardrail models, including LlamaGuard \citep{inan2023llama}, Qwen3Guard \citep{qwen3guard2025}, JoySafety \citep{jd-opensource2025}, PolyGuard \citep{poliguard2025}, and NemoGuard \citep{nemo_guardrails2023}, typically formulate safety supervision as a classification or instruction-following problem, where supervised models assign discrete risk labels to user inputs, model outputs, or dialogue contexts. Recent work extends this paradigm to agentic settings by incorporating tool-use contexts, execution traces, or multi-step interaction histories, as seen in GuardAgent \citep{xiang2024guardagent}, ShieldAgent \citep{shieldagent2025}, SafeEvalAgent \citep{Wang2025safeevalagent}, and AGrail \citep{Luo2025agrail}. Other systems, such as Safiron \citep{huang2025building} and ToolSafe \citep{mou2026toolsafe}, further move toward proactive safety assessment in tool-augmented workflows. However, existing guard models still rely on coarse-grained supervision and are not designed to comprehensively capture trajectory-level failures in agentic tasks. This gap highlights the need for more structured risk representations and trajectory-level data construction to support systematic agentic safety evaluation.
