\section{Introduction}

Building 3D scenes from visual observations is a long-standing goal in computer vision, with broad impact across gaming, robotics, embodied AI, and VR/AR~\citep{ding2025understanding,kong20253d,ye2026world,zhang2025advances}. Two complementary directions have driven progress in this space: reconstruction recovers 3D scenes from real-world captures, while generation synthesizes plausible scenes from limited or even imagined conditions~\citep{zhang2025advances,wang2026feed,li2026thinking}. Together, they form the technological foundation for populating the digital worlds of tomorrow.

3D scene generation and reconstruction have long developed as two separate lines of research. Reconstruction is dominated by feed-forward methods that regress 3D representations directly from multi-view images~\citep{hong2023lrm,charatan2024pixelsplat,szymanowicz2024splatter,xu2025depthsplat,chen2024mvsplat}. Generation has evolved from per-scene optimization with 2D priors via score distillation~\citep{lin2023magic3d,poole2022dreamfusion,tang2023dreamgaussian,wang2023prolificdreamer} to latent-space diffusion as the current mainstream~\citep{yang2025prometheus,go2025splatflow,huang2026gen3r,li2025flashworld}, with recent extensions diffusing in the feature space of pretrained 3D foundation models or representation autoencoders~\citep{gao2026oneworld,sun2026vggt,jang2026repurposing}. A recent work, Gen3R~\citep{huang2026gen3r}, attempts to unify the two tasks by extending the latent-space generation pipeline to also handle reconstruction. However, this design introduces clear limitations: the diffusion objective is defined on intermediate latent features rather than the underlying 3D representation, preventing the 3D output from being directly optimized; in addition, both branches suffer from information loss introduced by the pretrained Variational Autoencoder (VAE) or Representation Autoencoder (RAE), which itself requires additional training.

To address these limitations, we reformulate unification under a pixel-space diffusion paradigm and introduce PixWorld, a single framework that jointly handles 3D reconstruction and generation. This design brings two key benefits. First, for both generation and reconstruction, eliminating the latent stage removes the information loss introduced by latent encoding and the additional training cost of a VAE/RAE, which is especially critical for reconstruction as a fidelity-bound task. Second, for generation, the diffusion objective directly supervises the 3D representation through differentiable rendering (see Fig.~\ref{fig:teaser}), aligning the training signal with the fidelity of the 3D scene rather than with targets in an intermediate latent space. Building on this pixel-space framework, the unification reduces to how to expose both tasks to a single forward pass. We achieve this by partitioning multi-view inputs into clean and noisy subsets: clean views drive reconstruction, while noisy ones are generated conditioned on the clean ones, with both producing a pixel-aligned 3D Gaussian representation~\citep{kerbl20233d}.


Despite this direct pixel-space supervision, image-level photometric and perceptual losses do not fully guarantee geometrically faithful 3D structure.
To address this, we introduce a geometry perception loss that brings rendered views and their ground-truth counterparts close in the geometry-aware feature space of a pretrained 3D foundation model. As this feature space encodes 3D geometric structure beyond 2D appearance, the loss directly encourages the rendered scene to share the same underlying geometry as the ground truth. Extensive experiments on RealEstate10K~\citep{zhou2018stereo}, DL3DV~\citep{ling2024dl3dv} and WorldScore~\citep{duan2025worldscore} demonstrate that a single PixWorld model delivers high-fidelity 3D scene generation and reconstruction, matching state-of-the-art reconstruction methods while outperforming previous latent-space generation approaches.

Our main contributions are summarized as follows:
\begin{itemize}
\item We propose PixWorld, an end-to-end pixel-space diffusion framework that supervises a pixel-aligned 3D Gaussian representation directly through multi-view differentiable rendering, with no intermediate VAE or RAE. This eliminates the information loss and training cost of a latent autoencoder, aligns the diffusion signal with the fidelity of the 3D scene itself, and naturally unifies 3D scene generation and reconstruction in a single model.
\item We introduce a geometry perception loss that aligns rendered views with ground truth in the geometry-aware feature space of a pretrained 3D foundation model, providing 3D structural supervision beyond 2D photometric and perceptual losses.
\item Experiments show that a single PixWorld model achieves superior performance on both 3D scene generation and reconstruction, establishing pixel-space diffusion as a unified paradigm for 3D scene modeling.
\end{itemize}
