\section{Related Work}

\subsection{3D Scene Generation}
\textbf{Diffusion-based Iterative 3D Scene Generation.} 
Early 3D generation methods leverage pretrained 2D generative models~\citep{rombach2022high,podell2023sdxl,zhang2023adding,peebles2023scalable} as generative priors. A representative line employs Score Distillation Sampling (SDS)~\citep{lin2023magic3d,poole2022dreamfusion,tang2023dreamgaussian,wang2023prolificdreamer} to optimize a 3D representation such as 3DGS~\citep{kerbl20233d} or NeRF~\citep{mildenhall2021nerf} by aligning rendered views with a pretrained 2D diffusion model. Without explicit multi-view constraints, these methods suffer from cross-view inconsistency, and the per-scene iterative optimization further limits their scalability.

\textbf{Multi-View Reconstruction-Based 3D Scene Generation.}
A second line first synthesizes multi-view images or videos with pretrained 2D diffusion models, then performs 3D reconstruction from the synthesized views~\citep{chen2024mvsplat360,gao2024cat3d,hao2025gaussvideodreamer,liu2024reconx,liu2023syncdreamer,sargent2023zeronvs,shi2023mvdream,sun2024dimensionx,wu2024reconfusion,zhao2024genxd} or incremental outpainting~\citep{chung2023luciddreamer,fridman2023scenescape,schwarz2025recipe,yu2024wonderjourney,yu2025wonderworld}. While avoiding per-scene optimization, these methods rely on 2D RGB priors without explicit 3D reasoning, often resulting in inconsistent geometry and weak multi-view fidelity.

\textbf{3D Scene Generation in Latent Spaces.}
More recent work moves diffusion into compressed latent spaces and learns a latent-to-3D decoder, typically targeting 3DGS. One line freezes a pretrained image or video VAE (\textit{e.g.}, SD-VAE~\citep{podell2023sdxl}, Wan-VAE~\citep{wan2025wan}) and trains a decoder that maps the resulting latents to 3D Gaussians, with a diffusion backbone fine-tuned in this latent space~\citep{yang2025prometheus,go2025splatflow,li2024director3d,li2025flashworld,dai2025fantasyworld,go2025videorfsplat,go2025vist3a,zhou2026drivinggen,wang2025mmgen}. Following the shift from VAEs to Representation Autoencoders (RAEs)~\citep{zheng2025diffusion,shi2025latent,chen2025aligning,bi2025vision}, a more recent line operates in the representation space of pretrained 3D foundation models~\citep{gao2026oneworld,sun2026vggt,jang2026repurposing}. Building on this latent-space paradigm, Gen3R~\citep{huang2026gen3r} further attempts to unify reconstruction and generation within a single latent-space model.



\subsection{3D Scene Reconstruction}
3D scene reconstruction aims to recover 3D representations from one or more captured views. Classical pipelines rely on multi-view stereo and structure-from-motion to estimate dense geometry, while NeRF~\citep{mildenhall2021nerf} and 3D Gaussian Splatting~\citep{kerbl20233d} popularized per-scene optimization for high-fidelity novel view synthesis. To avoid expensive per-scene fitting, recent feed-forward methods learn to directly map sparse or dense multi-view inputs to a 3D representation in a single forward pass~\citep{hong2023lrm,charatan2024pixelsplat,szymanowicz2024splatter,chen2024mvsplat,xu2025depthsplat}, with LRM~\citep{hong2023lrm} regressing implicit triplanes and pixel-aligned Gaussian regressors~\citep{charatan2024pixelsplat,szymanowicz2024splatter,chen2024mvsplat} predicting per-pixel Gaussians. These methods are deterministic and conditioned on the available observations, and therefore cannot synthesize unseen content beyond the input views.


\subsection{Pixel-Space Generation}
Latent diffusion models~\citep{rombach2022high,podell2023sdxl,peebles2023scalable} dominate large-scale image generation by operating in a compressed VAE latent space, introducing an indirection between the generative target and the pixel output. A complementary line of work performs diffusion directly in pixel space~\citep{dhariwal2021diffusion,jabri2023scalable,hoogeboom2023simple,hoogeboom2025simpler,wang2025pixnerd,chen2025pixelflow,yu2025pixeldit,ma2025deco,chen2025dip,li2025back}, showing that, with sufficient capacity and data, pixel-space generation can match or surpass its latent-space counterparts in fidelity. We extend this perspective to the 3D setting: by performing diffusion directly in pixel space, PixWorld removes the intermediate VAE/RAE of prior latent-space 3D methods and supervises the 3D representation directly through differentiable rendering of the predicted 3DGS.
