% =============================================================================
% 6. RELATED WORK
% =============================================================================
% \section{Related Work}
% \vspace{-0.5ex}
% \label{sec:related}

% \textbf{Mid-training and continued pretraining.}
% A growing line of work identifies a distinct stage between large-scale
% pretraining and task-specific post-training, in which a model is further
% trained on a smaller, carefully curated corpus to install inductive biases
% that are hard to acquire from generic web text or downstream fine-tuning.
% MiniCPM~\citep{minicpm} and OLMo~\citep{olmo} schedule such a stage to
% upweight high-quality or domain-relevant data near the end of pretraining,
% and Code-Llama~\citep{codellama} long-context adaptation can be read as a
% mid-training stage targeting context-length generalization. We adopt the
% same staging philosophy but apply it to a different end: installing an
% \textit{agent-oriented} structural prior through function-aware FIM.

% \textbf{Fill-in-the-middle objectives.}
% The fill-in-the-middle objective~\citep{bavarian2022fim} is a standard
% ingredient in modern code LLM pretraining, with random-span variants used
% in Code-Llama~\citep{codellama}, StarCoder~\citep{starcoder},
% Qwen-Coder~\citep{qwencoder}, and DeepSeek-Coder~\citep{deepseekcoder}.
% Our recipe extends and specializes prior FIM along three axes. First,
% masking targets are chosen at \textit{function granularity} via program
% dependency graph analysis and a complexity--inferability double criterion,
% so the masked region corresponds to a unit of reasoning rather than an
% arbitrary span. Second, we embed an \textit{explicit chain-of-thought}
% inside the FIM middle, supervising the model to produce a rationale
% before the code. Third, the objective lives in a \textit{dedicated
% mid-training stage} rather than dissolved into pretraining.

% \textbf{Coding agent foundation models.}
% Progress on coding agents has been driven by benchmarks such as
% SWE-Bench~\citep{swebench}, scaffolds such as SWE-agent~\citep{sweagent}
% and OpenHands~\citep{openhands}, and trajectory-centric post-training
% pipelines such as SWE-Gym~\citep{swegym}, R2E-Gym~\citep{r2egym}, and
% SWE-Smith~\citep{swesmith}. Our work is complementary: we use R2E-Gym and
% SWE-Smith unmodified on top of a mid-trained checkpoint, and the novelty
% lies one stage earlier---extracting a self-supervised signal from the
% structure already present in source code, which both raises in-domain
% agent metrics and mitigates the off-domain capability erosion that
% trajectory-only post-training otherwise inflicts.

% \textbf{Distillation from frontier models.}
% Our default recipe uses Gemini-3-Flash to generate the chain-of-thought
% rationales embedded in each FIM middle span, placing it within the
% tradition of distilling capability from a strong teacher~\citep{orca,
% selfinstruct, distillwhisper}. We isolate the contribution of FIM
% structure from CoT distillation through a controlled ablation in
% Section~\ref{sec:exp:ablation}: removing the rationale still beats the
% no-mid-training baseline, and replacing the frontier teacher with
% self-generated rationales recovers a sizeable fraction of the gain, so
% the recipe is not distillation-bound.

\section{Related Work}
\vspace{-0.5ex}
\label{sec:related}

\textbf{Mid-training and continued pretraining.}
A growing body of work~\citep{mid-training-survey, gururangan2020dontstop}
identifies a stage between pretraining and post-training in which a model
is further trained on a curated corpus to install inductive biases hard
to acquire from generic web text or fine-tuning. MiniCPM~\citep{minicpm},
OLMo~\citep{olmo}, and DeepSeek-V3~\citep{deepseekv3} schedule such a
stage near the end of pretraining; Code-Llama~\citep{codellama} follows
the same staging philosophy for context-length generalization. Recent
work shows \emph{when} specialized data is introduced matters as much as
how much is used~\citep{remit, frontload-reasoning}. We adopt this
philosophy but target an \emph{agent-oriented} structural prior through
function-aware FIM.

\textbf{Fill-in-the-middle and structure-aware code objectives.}
FIM~\citep{bavarian2022fim, fried2022incoder} is a defining ingredient
of modern code LLM pretraining, used in
Code-Llama~\citep{codellama}, StarCoder-2~\citep{starcoder2},
CodeGen~\citep{codegen}, Qwen-Coder~\citep{qwencoder}, and
DeepSeek-Coder~\citep{deepseekcoder}.
A recent line moves beyond random spans toward \emph{structure-aware}
masking: AST-T5~\citep{ast-t5} and AST-FIM~\citep{ast-fim} mask AST
subtrees, the latter reporting up to $+5$ pts on real-edit FIM
benchmarks; Horizon-Length Prediction~\citep{hlp} adds a planning signal;
Instruction-aware FIM~\citep{ifim} extends the FIM tuple with a
developer-comment slot. Repository-level retrieval methods such as
GraphCoder~\citep{graphcoder} and DRACO~\citep{draco} exploit program
dependencies, but at inference time. Our recipe differs along three
axes: targets are chosen at \emph{function granularity} via PDG analysis
with a complexity--inferability criterion, so the masked region is a
unit of agent-relevant reasoning rather than a syntactic subtree; an
explicit \emph{chain-of-thought} sits inside the FIM middle; and the
objective lives in a \emph{dedicated mid-training stage}.

\textbf{Coding agent foundation models.}
Progress has advanced on three fronts. Benchmarks have moved from
SWE-Bench~\citep{swebench} to broader, contamination-resistant suites
including Multi-SWE-Bench~\citep{multiswebench} and
SWE-Bench-Pro~\citep{swebenchpro}. Scaffolds such as
SWE-agent~\citep{sweagent}, OpenHands~\citep{openhands}, and
Agentless~\citep{agentless} expose different action interfaces.
Trajectory-centric post-training pipelines---SWE-Gym~\citep{swegym},
R2E-Gym~\citep{r2egym}, SWE-Smith~\citep{swesmith},
SWE-Lego~\citep{swelego}, and Skywork-SWE~\citep{skyworkswe}---curate
agent trajectories on top of these scaffolds, with recent extensions
replacing or augmenting SFT with RL~\citep{longcontext-swerl,
selfplay-swerl}. We use R2E-Gym, SWE-Smith, and SWE-Lego unmodified
atop a mid-trained checkpoint; the novelty lies one stage earlier---a
self-supervised signal extracted from structure already present in
source code, which mitigates the off-domain capability erosion
trajectory-only post-training inflicts.

\textbf{Distillation from frontier models.}
Our recipe uses Gemini-3-Flash to generate the CoT rationales embedded
in each FIM middle, placing it within distillation from a strong
teacher~\citep{orca, distillwhisper, magicoder} and CoT
distillation~\citep{ho2023reasoningteachers, hsieh2023distill,
chen2025cotdistill}. Recent work shows diversity and structural
alignment of the rationale matter as much as teacher
strength~\citep{chen2025cotdistill}; consistent with this, our
CoT-source ablation (Section~\ref{sec:exp:ablation}) confirms the
recipe is not distillation-bound.