\section{Dataset and Benchmark Details}
\label{appendix:dataset}

\subsection{BBEH Task Coverage}
We evaluate on all 23 tasks from BBEH~\citep{kazemi2025bigbenchextrahard}.
Table~\ref{tab:bbeh_tasks} lists the tasks and their categories.
For each task, we use all available examples from the BBEH dataset,
yielding 4{,}520 unique problems across all 23 tasks.
Each problem is expanded to $K=8$ beams during generation, for 36{,}160 beam--problem
pairs in total.
These 4{,}520 problems are further partitioned for probe and discriminator training,
stratified per task: within each BBEH task, problems are randomly shuffled and split
into train / validation / test at an $0.8 / 0.1 / 0.1$ ratio, so every task
contributes proportionally to every split (3{,}616 / 452 / 452 problems in aggregate).

\input{tables/bbeh}

Several BBEH tasks build on earlier benchmarks, and we acknowledge the
originating work as requested by the BBEH authors.
BoardGame QA originates from~\citet{kazemi2024boardgameqa}.
Causal Understanding draws on~\citet{nie2024moca} and~\citet{kiciman2023causal}.
Dyck Languages and Word Sorting follow the setup
of~\citet{tyen2023llms}.
Geometric Shapes is based on~\citet{kazemi2023geomverse},
Linguini on~\citet{sanchez2024linguini},
and NYCC builds on~\citet{hessel2022androids}
and~\citet{zhang2024humor}.
Spatial Reasoning follows~\citet{yamada2023evaluating},
Time Arithmetic follows~\citet{fatemi2024test},
Web of Lies follows~\citet{white2024livebench},
and Zebra Puzzles follows~\citet{shah2024causal}.

\subsection{BBEH Answer Extraction}
We use the official BBEH answer extraction logic from
\citet{kazemi2025bigbenchextrahard}, which implements task-specific
regular expression patterns to extract final answers from free-form
model outputs.
\cref{tab:extraction-failure} reports per-LLM parsing statistics over
every generation. Per-beam parsing failure varies by more than an order
of magnitude across the LLMs we evaluate, and the fraction of problems
with at least one parsed beam follows the same ordering. For
Llama-3.3-70B-Instruct nearly every beam emits one of the canonical
answer prefixes, whereas the other LLMs leave a substantially larger
fraction of beams without any of the prefixes the extractor matches
against, and the answer cannot be recovered without other methods.

\input{tables/extraction_failure}

\subsection{Output-Length Statistics}
\label{appendix:output_length}

\cref{tab:output-length-stats,fig:output-length,fig:output-length-per-task}
report character-length statistics of the saved generations across the
five LLMs we evaluate, computed over the entire dataset. Two patterns
emerge. First, GPT-OSS-20B, Skywork-OR1-32B and DS-R1-Qwen-32B exceed
Llama-3.1-8B-Instruct and Llama-3.3-70B-Instruct at the median by
roughly an order of magnitude, and their interquartile bands reach into
the ten-thousand-character range. Second, individual tasks shift the
entire column by a near-constant offset. Hard combinatorial tasks like
Web of Lies, Zebra Puzzles, Multistep Arithmetic and Shuffled Objects
remain at the top of every column, while short answer-format tasks like
NYCC, Linguini, SportQA and SARC Triples remain at the bottom, and
each LLM's relative ordering of tasks is preserved across columns.
Llama-3.3-70B-Instruct produces the tightest distribution of the five,
with both the narrowest IQR and the lowest mean despite a comparable
median to the 8B checkpoint.
The maximum of $818{,}647$ characters for GPT-OSS~20B originates from a single pathological example in the Multistep Arithmetic task. The model emits roughly $7{,}000$ characters of valid reasoning, then enters a runaway indentation loop in which each successive opening parenthesis is preceded by an increasing number of leading spaces, growing quadratically until the $8{,}192$-token generation cap is reached. Approximately $816{,}000$ of the $818{,}647$ characters are whitespace. All eight beams degenerate identically due to deterministic beam search. Evaluation operates at the token level and is unaffected by character length.

\input{tables/output_length_summary}
\input{figures/output_length}
\input{figures/output_length_per_task}

\subsection{BBEH Accuracy of the Source LLMs}
\label{appendix:bbeh_accuracy}

\cref{tab:bbeh-accuracy} reports BBEH correctness of the five LLMs
scored with the official extraction logic of
\citet{kazemi2025bigbenchextrahard} against the canonical task
targets. \textsc{p@1} averages correctness over all beams per problem,
while \textsc{p@8} marks a problem solved if at least one beam scores
correct, and the gap between the two columns quantifies the margin a
verifier could realise from the existing beam set. Across tasks,
Llama-3.3-70B-Instruct, Skywork-OR1-32B and GPT-OSS-20B remain ahead
of the other two on \textsc{p@1} within a few percentage points of
each other, and DS-R1-Qwen-32B records the largest \textsc{p@8} gain
relative to its own \textsc{p@1}, while GPT-OSS-20B records the
smallest such gain.

The bottom rows aggregate the per-task numbers two ways. The micro
average pools all problems across tasks, while the adjusted harmonic
mean uses the BBEH paper's aggregation $N\,/\,\sum_i 1/(a_i+1)$, which
down-weights uneven performance across tasks. For the two LLMs that
overlap with their Table~2 (Llama-3.1-8B-Instruct and DS-R1-Qwen-32B,
reported there at $3.6\%$ and $5.2\%$), our \textsc{p@1} adjusted
harmonic means agree to within the rounding granularity of the
published per-task accuracies. Llama-3.3-70B-Instruct, Skywork-OR1-32B
and GPT-OSS-20B were not evaluated under this aggregation in the
original BBEH release, so the corresponding columns extend the
benchmark's leaderboard to those checkpoints.

\input{tables/bbeh_accuracy}

\subsection{Qualitative Beam Outputs and Embedder Similarity}
\label{appendix:beam_similarity}

To make the per-LLM beam pool concrete and to motivate why the
Output Embedding (\emph{Exact}) candidate of \cref{methodology}
captures the available semantic signal, we inspect pairs of beams from
the same input prompt and contrast a high-similarity pair against a
low-similarity pair. Cosine similarity is taken in the
Llama-Embed-Nemotron-8B space, the same space used by the Pooled and
Exact Output Embedding candidates throughout the paper. Both prompts
are drawn from the held-out test split of \cref{appendix:bbeh_accuracy}.

\begin{table}[h]
\caption{High-similarity pair from example~107 of the Time Arithmetic task. Both beams
walk through the same arithmetic chain and converge on the same answer,
with only the wording of the closing sentence differing.}
\label{tab:beam_sim_high}
\small
\centering
\begin{tabular}{p{0.47\textwidth} p{0.47\textwidth}}
\toprule
\multicolumn{2}{p{0.96\textwidth}}{\textbf{Prompt (truncated).}\
\textit{Question1\ldots If the date for the day after yesterday is
2017-12-02, what is the date 1116 days from now? Question2: A plane
leaves Location A at 01 AM UTC$+$0000 and arrives in Location B
$X'{:}Y'{:}Z'$ later. Location B is in IST(+0530). What time is it in
Location B when the plane lands?}} \\
\midrule
\textbf{Beam 1} & \textbf{Beam 2} \\
\midrule
\textit{To find the date 1116 days from 2017-12-02\ldots
$1116 \div 365 = 3$ years with a remainder of $111$\ldots
$2020$-$12$-$02 + 111$ days $= 2021$-$03$-$23$\ldots
$X' = 3$, $Y' = 24$, $Z' = 9$\ldots
UTC$+$0000\ $01{:}00{:}00 + 03{:}24{:}09 = 04{:}24{:}09 + 5\,$h
$+ 0.5\,$h $=$ UTC$+$0530\ $09{:}54{:}09$.
So, the plane lands on \underline{the same day at 09:54:09 in Location B}.
The answer is: same\_day, 09:54:09}
&
\textit{To find the date 1116 days from 2017-12-02\ldots
$1116 \div 365 = 3$ years with a remainder of $111$\ldots
$2020$-$12$-$02 + 111$ days $= 2021$-$03$-$23$\ldots
$X' = 3$, $Y' = 24$, $Z' = 9$\ldots
UTC$+$0000\ $01{:}00{:}00 + 03{:}24{:}09 = 04{:}24{:}09 + 5\,$h
$+ 0.5\,$h $=$ UTC$+$0530\ $09{:}54{:}09$.
So, the plane lands on \underline{the same day, 09:54:09}.
The answer is: same\_day, 09:54:09}
\\
\midrule
\multicolumn{2}{c}{Embedder-space cosine $\cos = 0.997$} \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\caption{Low-similarity pair from example~113 of the Causal Understanding task. The
left beam walks through the rules in a multi-step argument; the right
beam returns the verdict in one line. The two beams sit in opposite
regions of the embedder space. They also disagree on the final answer:
the long-form beam concludes Yes (matching the ground truth), while the
short beam answers No.}
\label{tab:beam_sim_low}
\small
\centering
\begin{tabular}{p{0.47\textwidth} p{0.47\textwidth}}
\toprule
\multicolumn{2}{p{0.96\textwidth}}{\textbf{Prompt (truncated).}\
\textit{Jim, Carol, Bob, and Nancy share electricity from 8--9 PM. The
breaker fails if three lamps are on at the same time. Jim is not
supposed to use his lamp on Thursdays. Carol, Bob, and Nancy may use
theirs. This Thursday Jim, Carol, and Bob all turn on their lamps at
8 PM and the breaker fails. Did Jim cause the breaker to fail? Reply
Yes / No / Ambiguous.}} \\
\midrule
\textbf{Beam 2} (long-form chain of thought) & \textbf{Beam 7} (short
direct answer) \\
\midrule
\textit{To determine if Jim turning on his lamp at 8 PM caused the
circuit breaker to fail, let's analyze the situation step by step.
1.~The breaker will fail if three people turn on their lamps at the
same time. 2.~Jim is not supposed to use his lamp on Thursdays, but he
turned it on anyway. 3.~Carol and Bob also turned on their lamps at the
same time as Jim. 4.~Since Carol and Bob are allowed to use their lamps
on Thursdays, their actions did not cause the breaker to fail.
5.~The breaker failed because three people (Jim, Carol, and Bob) turned
on their lamps at the same time. Considering these points, the majority
of people would likely say that Jim turning on his lamp at 8 PM caused
the circuit breaker to fail\ldots}
&
\vspace{2em}\textit{The answer is: No.}
\\
\midrule
\textit{Verdict: \textbf{Yes} (correct)} & \textit{Verdict: \textbf{No}} \\
\midrule
\multicolumn{2}{c}{Embedder-space cosine $\cos = 0.063$} \\
\bottomrule
\end{tabular}
\end{table}

The eight beams of the Causal Understanding example~113 cluster into three
regions of the embedder space. These are the long-form chain-of-thought mode and
the two short-answer modes (``Yes'' and ``No''). Within each mode,
beams sit at $\cos \ge 0.98$ from one another. Between the long-form
mode and either short mode, $\cos$ collapses to the $[0.06, 0.10]$
range regardless of whether the two beams agree on the final verdict,
so the low score in \cref{tab:beam_sim_low} is driven by format
divergence rather than answer disagreement. The showcased pair happens
to combine both format divergence and answer disagreement. The long-form beam reaches the correct answer (Yes)
while the short beam answers No, but a long-form Yes paired with a
short-form Yes would yield a cosine in the same low range. The two
short modes sit at $\cos \approx 0.80$ from each other despite
contradicting each other on the verdict. The embedder therefore
separates beams primarily by response format and only secondarily by
the content of the answer itself. This format-first geometry is what
lets the Output Embedding reach the high-purity, high-$\Delta_{\cos}$
region of \cref{appendix:geometric}, and is also why no purely latent
candidate inherits the same separation, since none of them preserves
the surface form that the embedder keys on.

