\section{Extended Related Work}
\label{appendix:extended_related_work}

\textbf{Surveys and theoretical analyses.}
A growing set of surveys systematically categorises the emerging space of latent and continuous reasoning methods. \citet{chen2025reasoninglanguagecomprehensivesurvey} provide a comprehensive taxonomy organising approaches along two axes: token-wise horizontal methods that replace discrete tokens with continuous counterparts, and layer-wise vertical methods that propagate latent state across transformer depth. \citet{zou2026the} offer a theoretical characterisation of the fundamental exploration--execution trade-off between discrete and continuous reasoning, proving that discrete chain-of-thought is forced into a high-certainty regime while continuous representations enable exploration at the cost of amplified noise on computational tasks; they introduce the Symbolic Index as a scalar measure of decisional certainty that governs this trade-off. \citet{mondorf2024accuracyevaluatingreasoningbehavior} survey evaluation practices for reasoning in LLMs, arguing that downstream accuracy conflates reasoning quality with surface-level pattern matching. \citet{sui2025stopoverthinkingsurveyefficient} and \citet{feng2025efficientreasoningmodelssurvey} survey efficient reasoning from the angle of compute cost and length budgets. Together these works catalogue methods and their downstream trade-offs, whereas our framework evaluates the intrinsic representational quality of the intermediate state independently of the decoding strategy or task accuracy.

\textbf{Latent world models, monitoring, and localization.}
Beyond reasoning benchmarks, a parallel line of work demonstrates that LLMs build structured internal models of the world as they process language. \citet{feng2025monitoring} show that latent world states can be extracted as structured propositions via propositional probes, providing direct empirical grounding for our Separability axiom: the hidden states must contain linearly decodable information about the semantic state of the world. \citet{Du_2025} find that human-like object concept representations emerge naturally in multimodal LLMs without explicit supervision, consistent with the view that the internal geometry supports rich semantic structure. Localizing where this structure resides has received substantial attention. \citet{bandarkar2025layerswappingzeroshotcrosslingual} show that swapping transformer layer ranges transfers cross-lingual knowledge across models, and \citet{alabi-etal-2024-hidden} find that adaptation concentrates in early-to-middle adapter layers, both implying semantically meaningful computations are localized at identifiable depths. \citet{sun2025transformerlayerspainters} show that middle layers can be reordered or skipped with minimal accuracy cost, consistent with mid-depth representations sharing structural properties across depth. \citet{ameisen2025circuit} trace computational circuits through residual stream contributions and demonstrate that feature-level computations shift dynamically across layers depending on the prompt, directly motivating why a fixed layer selection strategy for thought representation extraction is insufficient and why our framework tests multiple candidate extraction points.

\textbf{Sentence-level and continuous-space representations.}
A separate family of methods pursues meaning representation at the sentence level rather than the token level. Large Concept Models~\citep{lcm2024} reformulate language modelling as prediction in a sentence embedding space rather than over token vocabularies, leveraging SONAR~\citep{duquenne2023sonarsentencelevelmultimodallanguageagnostic} as the shared multilingual sentence encoder. SONAR-LLM~\citep{dragunov2025sonarllm} extends this further with an autoregressive transformer that reasons in sentence embedding space and decodes back to tokens, operating at a granularity coarser than individual tokens but finer than document-level representations. These methods design representations at a fixed semantic granularity chosen a priori. Our framework is granularity-agnostic and instead asks whether representations produced at any granularity by an existing model satisfy the four axioms, making it applicable to sentence-level methods as a diagnostic tool as well as to the token-level iterative candidates we evaluate here.
