\clearpage
\appendix

\section{Theoretical Analysis}
\label{app:theory}

\subsection{Bayes-Consistent Revision-Policy Learning}
\label{app:bayes-consistency}

Under a fixed, $\theta$-independent corruption proposal, cross-entropy training on per-position oracle revision labels recovers the conditional distribution of the optimal revision action at every position (Theorem~\ref{thm:bayes-consistency}). Theorem~\ref{thm:plugin-excess-risk} converts model-side TV distance into excess 0--1 risk; Proposition~\ref{prop:history-monotonicity} shows the \temporalmethodfull conditioning cannot increase Bayes risk.

\paragraph{Notation.}
Let $\mathcal{V}$ denote the token vocabulary, $\bar{\mathcal{V}} := \mathcal{V} \cup \{\mask\}$, and $E \subseteq \{1, \ldots, N\}$ the set of edit positions. A training sample is drawn as follows. Sample $(\xstar, c) \sim \mathcal{D}$, then a rule $\rho_i \sim \pi_\rho$ over $\{\mathrm{wrong}, \mathrm{mask}\}$ independently at each $i \in E$. For $\rho_i = \mathrm{wrong}$, draw a source token $w_i \sim \nu_{\mathrm{wrong}}(\cdot \mid \xstar_i)$ on $\mathcal{V} \setminus \{\xstar_i\}$ and boundaries $(\beta_i, \mu_i)$ from a fixed $\theta$-independent law $\mathcal{B}$ on $\{1 \le \beta < \mu \le T\}$; for $\rho_i = \mathrm{mask}$, draw a single boundary $\mu_i \sim \mathrm{Uniform}\{1, \ldots, T\}$. The current-step index is $t \sim \mathrm{Uniform}\{0, \ldots, T-1\}$. Given these boundaries, the trajectory $\tilde{x}^{(0:T)}$ is per-position deterministic: $w_i \to \mask \to \xstar_i$ at $(\beta_i, \mu_i)$ for wrong-rule positions, $\mask \to \xstar_i$ at $\mu_i$ for mask-rule positions, and $\xstar_i$ throughout for non-edit positions $i \notin E$.

We write $z := \tilde{x}^{(t)}$ for the current state, $H := \tilde{x}^{(<t)}$ for the trajectory prefix, and $\phi(H, c) \in \mathbb{R}^d$ for a deterministic history feature map (the history component of the \temporalmethod accumulated embedding $a^{(t)}$; see Appendix~\ref{app:rope-engineering}). The accumulated embedding satisfies $a^{(t)} = e^{(t)} + \phi(H, c)$ with $e^{(t)} = \mathrm{wte}(z)$.

Let $X := (z, \phi(H, c), c, E)$ denote the analyst-side conditioning. Our proofs condition on $X$, while the implementation passes only $(a^{(t)}, c)$ to the model and uses $z, E$ to construct the supervision label. We write $p_\theta(\cdot \mid X)_i$ for the model's per-position categorical over $\bar{\mathcal{V}}$. By Proposition~\ref{prop:history-monotonicity}, the Bayes risk under $\sigma(X)$ lower-bounds the Bayes risk achievable from $\sigma(a^{(t)}, c)$, so Theorem~\ref{thm:bayes-consistency}'s population minimum is a lower bound on the implementation's Bayes risk.

\paragraph{Oracle revision rule.}
For each $i \in E$, the deterministic oracle next-action is
\begin{equation}
\tau(z_i, \xstar_i) \;=\; \begin{cases}
\mask & z_i \in \mathcal{V} \setminus \{\xstar_i\} \quad \text{(visible non-target token)} \\
\xstar_i & z_i = \mask \quad \text{(reveal target)} \\
\xstar_i & z_i = \xstar_i \quad \text{(preserve target)}
\end{cases}
\label{eq:oracle-tau}
\end{equation}
This is the single optimal revision action at the current state, not a per-step transition kernel for the trajectory.

\paragraph{Training objective.}
The training loss is the per-position three-state cross-entropy summed over edit positions:
\begin{equation}
\Ltrain(\theta) \;=\; \mathbb{E}_{\mathcal{D} \otimes q}\!\left[\sum_{i \in E} -\log p_\theta\!\left(\tau(z_i, \xstar_i) \,\big|\, X\right)_i\right],
\label{eq:rm-loss}
\end{equation}
where $\mathcal{D} \otimes q$ denotes the joint distribution generated by the sampling order above.

\paragraph{Conditional label distribution.}
For each $i \in E$, define
\begin{equation}
P^*_i(y \mid X) \;:=\; \Pr_{(\xstar, w, \rho, \mathrm{boundaries}, t) \sim (\mathcal{D} \otimes q) \mid X}\!\big[\tau(z_i, \xstar_i) = y\big].
\label{eq:Pstar-def}
\end{equation}
This is the marginal distribution of the oracle label given the model's input. Since $X$ does not include $\xstar$, $P^*_i$ is in general non-degenerate.

\begin{theorem}[Population minimizer recovers the conditional label distribution]
\label{thm:bayes-consistency}
Assume the rich-family idealization: $\{p_\theta\}$ is large enough that for some $\theta$, the model's per-position output $p_\theta(\cdot \mid X)_i$ matches $P^*_i(\cdot \mid X)$ jointly at every $X$ in the support of $\mathcal{D} \otimes q$ and every $i \in E$. Then any population minimizer $\theta^* \in \arg\min_\theta \Ltrain(\theta)$ satisfies
\[
p_{\theta^*}(\cdot \mid X)_i \;=\; P^*_i(\cdot \mid X) \quad \text{for } (\mathcal{D} \otimes q)\text{-a.e.\ } X \text{ and every } i \in E.
\]
\end{theorem}

\begin{proof}
Decompose $\Ltrain$ point-wise in $X$ via the tower property:
\[
\Ltrain(\theta) \;=\; \mathbb{E}_X\!\left[\sum_{i \in E} \mathbb{E}_{Y_i \mid X}\!\left[-\log p_\theta(Y_i \mid X)_i\right]\right],
\]
where $Y_i := \tau(z_i, \xstar_i)$ has conditional distribution $P^*_i(\cdot \mid X)$. \emph{Per-position decoupling.} At fixed $X$, the inner sum has terms depending on $\theta$ only through the per-position output distributions $\{p_\theta(\cdot \mid X)_i\}_{i \in E}$, which the rich-family hypothesis allows $\theta$ to realize independently. Hence minimizing the sum jointly over $\theta$ is equivalent to minimizing each per-position term independently in its corresponding categorical. \emph{Per-position propriety.} By strict propriety of cross-entropy~\cite{gneiting2007strictly}, for any reference distribution $r$ on $\bar{\mathcal{V}}$ the expected score $\mathbb{E}_{y \sim r}[-\log p(y)]$ is uniquely minimized over the categorical simplex at $p = r$. Apply this with $r = P^*_i(\cdot \mid X)$ to each position $i$; the unique per-position minimizer is $p_\theta(\cdot \mid X)_i = P^*_i(\cdot \mid X)$. The almost-everywhere conclusion follows because the support of $\mathcal{D} \otimes q$ has full measure.
\end{proof}

\begin{remark}[well-specification]
If $\{p_\theta\}$ is not rich enough to represent $P^*_i$, the conclusion holds in the function-class limit: the minimizer is the closest representable distribution in expected KL.
\end{remark}

\begin{remark}[loss reweighting]
Theorem~\ref{thm:bayes-consistency} is stated for the unweighted CE objective of Eq.~\eqref{eq:rm-loss}. Adding any constant or $X$-measurable per-position weighting $w_i(X) > 0$ does not change the minimizer, since the per-$X$ minimization scales by $w_i(X)$. Label-dependent weighting tilts the minimizer but is not used in our experiments.
\end{remark}

The general form of Theorem~\ref{thm:bayes-consistency}'s minimizer simplifies sharply at each of the two state types. Define $\alpha_i(X) := \Pr(\xstar_i = z_i \mid X)$, the conditional probability that the visible token equals the gold target.

\begin{proposition}[Visible-token minimizer is a binary calibrated correctness belief]
\label{prop:visible-belief}
At any $X$ with $z_i \neq \mask$,
\[
P^*_i(\cdot \mid X) \;=\; \alpha_i(X) \cdot \delta_{z_i} + (1 - \alpha_i(X)) \cdot \delta_{\mask}.
\]
\end{proposition}

\begin{proof}
Conditional on $z_i \neq \mask$, the value $z_i$ is determined by $X$. The remaining randomness in $\tau(z_i, \xstar_i)$ given $X$ is the indicator $\mathbf{1}\{z_i = \xstar_i\}$, which equals 1 with probability $\alpha_i(X)$ and 0 otherwise. By Eq.~\eqref{eq:oracle-tau}, $\tau = z_i$ in the first case and $\tau = \mask$ in the second.
\end{proof}

\textbf{Interpretation.} At visible-token positions, the population minimizer encodes a calibrated belief about whether the visible token is the gold target. The inference rule \emph{mask iff $p_\theta(\mask \mid X) > p_\theta(z_i \mid X)$} is the Bayes-optimal binary decision based on this belief, since the two probabilities are the atoms of $P^*_i$.

\begin{remark}[degenerate case]
If $\mathbf{1}\{z_i = \xstar_i\}$ is $X$-measurable, then $\alpha_i(X) \in \{0, 1\}$ a.s.\ and $P^*_i$ collapses to $\delta_{\mask}$ or $\delta_{z_i}$. The main results do not require this.
\end{remark}

\begin{proposition}[Calibrated gold belief at MASK positions]
\label{prop:mask-belief}
At any $X$ with $z_i = \mask$,
\[
P^*_i(y \mid X) \;=\; \Pr_{(\xstar, w) \sim (\mathcal{D} \otimes q) \mid X}[\xstar_i = y].
\]
\end{proposition}

\begin{proof}
When $z_i = \mask$, Eq.~\eqref{eq:oracle-tau} gives $\tau(z_i, \xstar_i) = \xstar_i$ tautologically, so the conditional distribution of $\tau$ given $X$ equals the conditional distribution of $\xstar_i$ given $X$.
\end{proof}

\textbf{Interpretation.} At MASK positions the population minimizer becomes a conditional language model over the gold token --- the same target as \cite{nie2025large}'s two-state model. Sampling from $p_\theta(\cdot \mid \mask, \phi(H, c), c)$ recovers the standard mask-prediction sampling step.

\paragraph{Summary.} \method\ trains a single per-position categorical that carries two calibrated beliefs: a binary $\{z_i, \mask\}$ correctness belief at visible-token positions, and a full-vocabulary gold-token belief at MASK positions. The inference rule reads each off the appropriate atoms --- the re-mask check at visible positions and the candidate reveal at MASK positions in Eq.~\ref{eq:remask-rule}.

\subsection{Plug-in Excess-Risk Theorem}
\label{app:plugin-excess-risk}

We analyze the deterministic argmax plug-in policy
\[
\hat{a}_\theta(X)_i \;:=\; \arg\max_{v \in \bar{\mathcal{V}}} p_\theta(v \mid X)_i,
\]
with ties broken by a fixed model-independent rule (e.g., lexicographic order on $\bar{\mathcal{V}}$); the Bayes-optimal classifier $\hat{a}^*(X)_i := \arg\max_v P^*_i(v \mid X)$ is defined under the same tie-breaking rule. Bounds below hold for any such fixed choice. The per-position 0--1 revision risk against the oracle is
\[
R(\theta) \;:=\; \mathbb{E}_{(\mathcal{D} \otimes q)}\!\left[\frac{1}{|E|} \sum_{i \in E} \mathbf{1}\!\left[\hat{a}_\theta(X)_i \neq \tau(z_i, \xstar_i)\right]\right],
\]
and the Bayes risk is $R^* := \inf_p R(p)$. By Propositions~\ref{prop:visible-belief}--\ref{prop:mask-belief}, $R^*$ has irreducible contributions from \emph{both} state types: $1 - \max_y P^*_i(y \mid X)$ at MASK positions, and $\min(\alpha_i(X), 1 - \alpha_i(X))$ at visible-token positions. Both vanish only at degenerate $X$ where $P^*_i$ is a delta.

\begin{theorem}[Plug-in excess Bayes risk]
\label{thm:plugin-excess-risk}
Under the setup above,
\[
R(\theta) - R^{*} \;\leq\; 2 \cdot \mathbb{E}_X\!\left[\frac{1}{|E|} \sum_{i \in E} \TV{p_\theta(\cdot \mid X)_i}{P^*_i(\cdot \mid X)}\right].
\]
\end{theorem}

\begin{proof}
Fix $X$ and $i \in E$; drop the $i$ subscript on $P^*$. By definition $Y := \tau(z_i, \xstar_i)$ has conditional distribution $P^*(\cdot \mid X)$, so $\Pr_{Y \sim P^*}[\hat{a} \neq Y \mid X] = 1 - P^*(\hat{a} \mid X)$ for any deterministic classifier $\hat{a}$ measurable in $X$.

\emph{Case $\hat{a}_\theta = \hat{a}^*$.} Per-$X$ excess risk is zero; the bound is trivial.

\emph{Case $\hat{a}_\theta \neq \hat{a}^*$.} Per-$X$ excess risk equals $P^*(\hat{a}^* \mid X) - P^*(\hat{a}_\theta \mid X)$. Decompose:
\[
P^*(\hat{a}^*) - P^*(\hat{a}_\theta) = \underbrace{[P^*(\hat{a}^*) - p_\theta(\hat{a}^*)]}_{(\mathrm{A})} + \underbrace{[p_\theta(\hat{a}^*) - p_\theta(\hat{a}_\theta)]}_{(\mathrm{B})} + \underbrace{[p_\theta(\hat{a}_\theta) - P^*(\hat{a}_\theta)]}_{(\mathrm{C})}.
\]
$(\mathrm{B}) \leq 0$ since $\hat{a}_\theta$ maximizes $p_\theta$ (under the fixed model-independent tie-break stated above). Bound each of (A) and (C) by absolute value: $(\mathrm{A}) \leq |p_\theta(\hat{a}^*) - P^*(\hat{a}^*)|$ and $(\mathrm{C}) \leq |p_\theta(\hat{a}_\theta) - P^*(\hat{a}_\theta)|$. Since $\hat{a}^* \neq \hat{a}_\theta$, these are two distinct nonnegative terms in the sum $\sum_v |p_\theta(v) - P^*(v)|$, hence $(\mathrm{A}) + (\mathrm{C}) \leq \sum_v |p_\theta(v) - P^*(v)| = 2 \cdot \TV{p_\theta}{P^*}$. Take expectation over $X$ and average over $i$.
\end{proof}

\begin{remark}[on the factor of 2]
This is the standard Devroye--Györfi--Lugosi excess-risk bound~\cite[Theorem~2.2]{devroye2013probabilistic}. The constant 2 is tight: with $P^* = \delta_1$ and $p_\theta = (1/2 - \epsilon, 1/2 + \epsilon)$, excess risk equals 1 while $\mathrm{TV} \to 1/2$.
\end{remark}

\paragraph{Combining Theorems~\ref{thm:bayes-consistency} and \ref{thm:plugin-excess-risk}.}
At the population minimum $p_\theta = P^*$, so $R(\theta^*) = R^*$: the CE-loss minimizer is Bayes-optimal under 0--1 revision risk. The Bayes risk $R^*$ is non-zero by construction (Propositions~\ref{prop:visible-belief}--\ref{prop:mask-belief}); driving CE loss further does not --- and should not --- reduce $R^*$.

\subsection{\temporalmethodfull{} Information Bound}
\label{app:history-monotonicity}

\temporalmethodfull augments the conditioning: the model receives not just $(z, c)$ but $(z, \phi(H, c), c)$. For a random variable (or tuple) $W$, write $\sigma(W)$ for the $\sigma$-algebra generated by $W$; a decision rule $\hat{a}$ is $\sigma(W)$-measurable exactly when $\hat{a}$ can be written as a deterministic function of $W$.

\begin{proposition}[History monotonicity]
\label{prop:history-monotonicity}
Let $\sigma_{\mathrm{base}} := \sigma(z, c)$ and $\sigma_{\mathrm{HR}} := \sigma(z, H, c)$. Then
\[
\inf_{\hat{a} \,\sigma_{\mathrm{HR}}\text{-meas.}} \mathbb{E}\!\left[\mathbf{1}\!\left[\hat{a} \neq \tau\right]\right] \;\leq\; \inf_{\hat{a} \,\sigma_{\mathrm{base}}\text{-meas.}} \mathbb{E}\!\left[\mathbf{1}\!\left[\hat{a} \neq \tau\right]\right].
\]
\end{proposition}

\begin{proof}
$\sigma_{\mathrm{base}} \subseteq \sigma_{\mathrm{HR}}$ since $(z, c)$ is a sub-tuple of $(z, H, c)$. Hence every $\sigma_{\mathrm{base}}$-measurable estimator is also $\sigma_{\mathrm{HR}}$-measurable, so the infimum over a larger family cannot exceed the infimum over a smaller one.
\end{proof}

The same argument applies to any history feature $\phi$ used by the model: $\sigma(z, \phi(H, c), c) \subseteq \sigma(z, H, c)$, so $\phi$ achieves a Bayes risk between the no-history baseline and the full trajectory. Whether the \temporalmethod accumulated embedding strictly decreases Bayes risk relative to the no-history baseline is an empirical question answered in Section~\ref{sec:experiments}.

\paragraph{Information hierarchy.}
The data-processing argument gives a three-level hierarchy --- current state only, current state plus the accumulated embedding, current state plus the full trajectory:
\[
\sigma\!\left(z, c\right) \;\subseteq\; \sigma\!\left(z, \phi(H, c), c\right) \;\subseteq\; \sigma\!\left(z, H, c\right).
\]
The history feature $\phi(H, c)$ is a compressed history signal; distinct trajectories generally produce distinct $\phi(H, c)$, but the mapping is many-to-one.

\subsection{Top-k Variant: Distribution-Shift TV Bound}
\label{app:topk-tv-gap}

Theorem~\ref{thm:bayes-consistency} assumes a fixed corruption $\nu$. In practice we use a non-uniform $\theta$-independent corruption $\nu^{(\mathrm{prac})}$: a frozen pre-trained checkpoint's top-$k$ distribution for text generation, a fixed wrong-digit distribution for Sudoku, and the source-image VQ tokens for image editing. In the image-edit case the VQ token is taken as $w_i$ only when $w_i \neq \xstar_i$; positions where the source already matches the target fall under the keep branch, keeping $\nu^{(\mathrm{prac})}_{\mathrm{wrong}}$ supported on $\mathcal{V} \setminus \{\xstar_i\}$. The loss gap between $\nu_{\mathrm{unif}}$ and $\nu^{(\mathrm{prac})}$ decomposes in three layers.

\begin{assumption}[Bounded log-likelihood]
\label{ass:bounded-rm}
Per-token log-probabilities are clipped: $-\log p_\theta(v \mid X)_i \leq B := -\log\varepsilon$ for every token $v$, position $i$, and input $X$. We use $\varepsilon = 10^{-8}$, giving $B \approx 18$.
\end{assumption}

\paragraph{Layer 1 (joint TV bound).}
For any test function $f$ with $|f| \leq B \cdot |E|$ and any two probability measures $P, Q$ on the joint sample space,
\[
\big|\mathbb{E}_P f - \mathbb{E}_Q f\big| \;\leq\; 2 B \cdot |E| \cdot \mathrm{TV}(P, Q).
\]
Applied to $f = \sum_{i \in E} -\log p_\theta(\tau(z_i, \xstar_i) \mid X)_i$, with $q^{(\mathrm{unif})}$ and $q^{(\mathrm{prac})}$ the joint proposals under $\nu_{\mathrm{unif}}$ and $\nu^{(\mathrm{prac})}$ respectively:
\[
\big|\Ltrain^{(\mathrm{unif})}(\theta) - \Ltrain^{(\mathrm{prac})}(\theta)\big| \;\leq\; 2B \cdot |E| \cdot \mathbb{E}_{(\xstar, c) \sim \mathcal{D}}\!\left[\mathrm{TV}\!\big(q^{(\mathrm{unif})}(\cdot \mid \xstar, c),\; q^{(\mathrm{prac})}(\cdot \mid \xstar, c)\big)\right].
\]

\paragraph{Layer 2 (factorize the joint TV).}
The two proposals share the rule prior $\pi_\rho$, the boundary law $\mathcal{B}$, and the timestep distribution; they differ only in the conditional corruption at $\rho = \mathrm{wrong}$ positions. Define the rule-marginalized corruption $\tilde{\nu}_i(\cdot \mid \xstar_i) := \pi_{\mathrm{wrong}} \cdot \nu_{\mathrm{wrong}}(\cdot \mid \xstar_i) + \pi_{\mathrm{mask}} \cdot \delta_{\mask}$. The proposals agree on the $\pi_{\mathrm{mask}}$ branch, so by the mixture-TV identity $\mathrm{TV}(\alpha P_1 + (1-\alpha) P_2,\, \alpha Q_1 + (1-\alpha) P_2) = \alpha \,\mathrm{TV}(P_1, Q_1)$, the per-position TV between rule-marginalized corruptions equals $\pi_{\mathrm{wrong}} \cdot \mathrm{TV}(\nu^{(\mathrm{unif})}_{\mathrm{wrong}, i}, \nu^{(\mathrm{prac})}_{\mathrm{wrong}, i})$. Combining with the standard product-distribution TV inequality (independence across positions),
\[
\mathrm{TV}\!\big(q^{(\mathrm{unif})},\, q^{(\mathrm{prac})}\big) \;\leq\; \pi_{\mathrm{wrong}} \cdot \sum_{i \in E} \mathrm{TV}\!\big(\nu^{(\mathrm{unif})}_{\mathrm{wrong}, i},\; \nu^{(\mathrm{prac})}_{\mathrm{wrong}, i}\big).
\]

\paragraph{Layer 3 (per-position bound, ready for analysis).}
Combining Layers 1 and 2:
\begin{equation}
\big|\Ltrain^{(\mathrm{unif})}(\theta) - \Ltrain^{(\mathrm{prac})}(\theta)\big| \;\leq\; 2B \cdot |E| \cdot \pi_{\mathrm{wrong}} \cdot \mathbb{E}_{(\xstar, c) \sim \mathcal{D}}\!\left[\sum_{i \in E} \TV{\nu^{(\mathrm{unif})}_{\mathrm{wrong}, i}}{\nu^{(\mathrm{prac})}_{\mathrm{wrong}, i}}\right].
\label{eq:topk-tv-bound}
\end{equation}
Only "wrong"-rule positions contribute; the mask-rule and non-edit branches contribute zero.

\paragraph{Implication.}
Theorem~\ref{thm:bayes-consistency} still applies on the support of $q^{(\mathrm{prac})}$: the population minimizer recovers the conditional label distribution under $\nu^{(\mathrm{prac})}$ rather than under $\nu_{\mathrm{unif}}$. Since inference-time errors are not uniform, $\nu^{(\mathrm{prac})}$ is the more relevant covered region.

\section{Engineering Implementation}
\label{app:rope-engineering}

The relative-lag form of \S\ref{sec:temporal-residual} admits a strict $O(1)$ per-step update of the accumulated embedding $a_i^{(t)}$ (also written $\mathrm{acc}_i$).

\paragraph{Derivation of the recurrence.}
Starting from the closed form (Eq.~\eqref{eq:temporal-embedding})
\[
a_i^{(t)} \;=\; \sum_{k=0}^{t} \gamma^{\,t-k}\, R_{k-t}\, e_i^{(k)},
\qquad
e_i^{(k)} := \mathrm{wte}(\tilde{x}_i^{(k)}),
\]
and the corresponding sum at step $t-1$,
\[
a_i^{(t-1)} \;=\; \sum_{k=0}^{t-1} \gamma^{\,t-1-k}\, R_{k-(t-1)}\, e_i^{(k)},
\]
left-multiplying $a_i^{(t-1)}$ by $\gamma R_{-1}$ and using the rotation composition rule $R_{-1} R_{k-(t-1)} = R_{k-t}$ gives
\[
\gamma\, R_{-1}\, a_i^{(t-1)} \;=\; \sum_{k=0}^{t-1} \gamma^{\,t-k}\, R_{k-t}\, e_i^{(k)}.
\]
Adding the current-step contribution $e_i^{(t)}$ (whose lag from the current step is zero, so $R_0 = I$) recovers $a_i^{(t)}$:
\[
a_i^{(t)} \;=\; e_i^{(t)} \;+\; \gamma\, R_{-1}\, a_i^{(t-1)}.
\]
This is the recurrence used in the implementation.

\paragraph{Memory layout.}
We maintain a single running tensor $a \in \mathbb{R}^{N \times d}$, where $N$ is the sequence length and $d$ the hidden width. The per-step update is one $N \times d$ block-diagonal rotation followed by an element-wise add; storing the full token-level trajectory would cost $O(T \times N \times d)$, while the recurrence collapses this to a single $O(N \times d)$ buffer.

\paragraph{Numerical stability.}
Each $R_\Delta$ is a block-diagonal orthogonal transform, so $R_{-1}$ preserves norms: $\|R_{-1}\, a_i^{(t-1)}\| = \|a_i^{(t-1)}\|$. Combined with the decay factor $\gamma \in (0, 1]$, the running sum's norm grows at most linearly in $t$.

\paragraph{Contrast with an absolute-step formulation.}
A natural alternative assigns each historical token an absolute-step rotation in a fixed (step-zero) reference frame, $\sum_{k=0}^{t} \gamma^{t-k} R_k\, e_i^{(k)}$. This admits the same per-step recurrence but anchors the trajectory at the start of denoising rather than the current step. The relative-lag form of Eq.~\eqref{eq:temporal-embedding} instead anchors the phase at the model's most recent input, and is what we use in all reported experiments.

\paragraph{Loss reweighting.}
Training launchers may attach per-position weights $w_i$ to the CE term. The math launcher uses constant weights; the Sudoku launcher emphasizes edit positions via an $X$-measurable edit-vs-non-edit weight. Both fall under the loss-reweighting remark following Theorem~\ref{thm:bayes-consistency}, so the population minimizer is unchanged.

\paragraph{Bounded-loss floor.}
We clip output probabilities at $\varepsilon = 10^{-8}$, giving $B = -\log\varepsilon \approx 18$ in Assumption~\ref{ass:bounded-rm}.


\section{Training Algorithm}
\label{app:training-algo}

We give the complete per-sample procedure summarized in Section~\ref{sec:training}. Let $T = 6$ denote the trajectory length used in our experiments.

\begin{enumerate}
    \item \textbf{Sample data.} Draw $(\xstar, c) \sim \mathcal{D}$ (target sequence and task condition).

    \item \textbf{Sample per-position rule.} For each $i \in E$, draw $\rho_i \sim \pi_\rho$ over $\{\mathrm{wrong}, \mathrm{mask}\}$ independently. Non-edit positions $i \notin E$ are deterministically labeled "keep"; their state is $\xstar_i$ at every $t$.

    \item \textbf{Sample source token (wrong-rule positions only).} For each $i \in E$ with $\rho_i = \mathrm{wrong}$, sample $w_i \sim \nu_{\mathrm{wrong}}(\cdot \mid \xstar_i)$ on $\mathcal{V} \setminus \{\xstar_i\}$. The rigorous baseline of Appendix~\ref{app:bayes-consistency} uses uniform sampling; the practical variant (Appendix~\ref{app:topk-tv-gap}) uses a task-specific non-uniform $\nu^{(\mathrm{prac})}$. Mask-rule positions ($\rho_i = \mathrm{mask}$) need no source-token sample; they enter the trajectory directly at the $\mask$ state.

    \item \textbf{Sample boundaries.} For each $\rho_i = \mathrm{wrong}$ position, draw $\beta_i \sim \mathrm{Uniform}\{1, \ldots, T-1\}$, then $\mu_i \mid \beta_i \sim \mathrm{Uniform}\{\beta_i + 1, \ldots, T\}$. For each $\rho_i = \mathrm{mask}$ position, draw $\mu_i \sim \mathrm{Uniform}\{1, \ldots, T\}$.

    \item \textbf{Construct the trajectory $\tilde{x}^{(0:T)}$.} Per-position deterministic from the boundaries:
    \begin{itemize}
        \item "wrong" rule: $\tilde{x}^{(t)}_i = w_i$ for $0 \leq t < \beta_i$, $\mask$ for $\beta_i \leq t < \mu_i$, $\xstar_i$ for $\mu_i \leq t \leq T$.
        \item "mask" rule: $\tilde{x}^{(t)}_i = \mask$ for $0 \leq t < \mu_i$, $\xstar_i$ for $\mu_i \leq t \leq T$.
        \item "keep" / non-edit: $\tilde{x}^{(t)}_i = \xstar_i$ for all $t$.
    \end{itemize}
    The wrong-rule ordering $w_i \to \mask \to \xstar_i$ and the mask-rule ordering $\mask \to \xstar_i$ both hold by construction; no post-processing is required.

    \item \textbf{Sample current step.} Draw $t \sim \mathrm{Uniform}\{0, \ldots, T-1\}$ and set $z = \tilde{x}^{(t)}$, $H = \tilde{x}^{(<t)}$.

    \item \textbf{Construct labels.} For each $i \in E$, the per-position label is the oracle action $y_i = \tau(z_i, \xstar_i)$ defined in Eq.~\eqref{eq:oracle-tau}: $\mask$ if $z_i$ is a visible non-target token, $\xstar_i$ if $z_i = \mask$ (reveal target), $\xstar_i$ if $z_i = \xstar_i$ (preserve target).

    \item \textbf{CFG dropout (image editing only).} Draw $\text{text\_drop} \sim \mathrm{Bernoulli}(0.10)$ and $\text{image\_drop} \sim \mathrm{Bernoulli}(0.10)$ independently. If $\text{text\_drop}$, replace the user instruction with its unconditional prefix. If $\text{image\_drop}$, mark image-content positions to have their accumulated embedding overwritten by $\mathrm{wte}(\mask)$ after RMSNorm in step~9. The generic algorithm (text generation, Sudoku) skips this step.

    \item \textbf{Compute the accumulated embedding.} Accumulate via Eq.~\eqref{eq:temporal-embedding},
    \[
    \mathrm{acc}_i \;=\; \textstyle\sum_{k=0}^{t} \gamma^{\,t-k}\, R_{k-t}\, \mathrm{wte}(\tilde{x}_i^{(k)}),
    \]
    or by the recurrence $\mathrm{acc}_i^{(s)} = \mathrm{wte}(\tilde{x}_i^{(s)}) + \gamma R_{-1} \mathrm{acc}_i^{(s-1)}$ with $\mathrm{acc}_i^{(0)} = \mathrm{wte}(\tilde{x}_i^{(0)})$. Apply RMSNorm. For image editing, if $\text{image\_drop}$ from step~8 fires, overwrite $\mathrm{acc}$ at image-content positions with $\mathrm{wte}(\mask)$. The model's input is $(\mathrm{acc}, c)$; the pipeline keeps $z$ aside for label construction and for the analyst-side conditioning of Theorem~\ref{thm:bayes-consistency}.

    \item \textbf{Compute the loss.} Cross-entropy of Eq.~\eqref{eq:total-loss} against the per-position labels $y_i$ from step~7.
\end{enumerate}

\paragraph{Practical top-$k$ proposal.}
For text generation, $\nu^{(\mathrm{prac})}_{\mathrm{wrong}}(\cdot \mid \xstar_i)$ in step~3 is the truncated softmax of a frozen checkpoint $\theta_0$'s top-$k$ output (default $k = 50$), restricted to $\mathcal{V} \setminus \{\xstar_i\}$; sharper conditionals admit smaller $k$. We probe $\theta_0$ once per training sample on $\xstar$ and cache the per-position top-$k$ index sets, drawing each position's wrong token from its own conditional softmax. Since $\theta_0$ is fixed, the proposal is independent of the live parameter $\theta$. Appendix~\ref{app:topk-tv-gap} bounds the objective shift relative to the uniform baseline.

\paragraph{Auxiliary losses (image editing only).}
In addition to the primary cross-entropy loss of Eq.~\eqref{eq:total-loss}, the image-editing pipeline uses an unlikelihood loss on non-edit positions~\cite{welleck2019neural} and a stage-2 ordering loss; their detailed form is deferred to a separate technical companion. These are heuristic regularizers outside the scope of Theorem~\ref{thm:bayes-consistency}.

\input{body/appendix_parts/Image_appendix}



\input{body/appendix_parts/text_appendix}