\section{Related Work}
\label{sec:related-work}

\textbf{Mask diffusion models.}
MDMs~\citep{nie2025large,ye2025dream,lou2023discrete} generate sequences by iteratively denoising masked inputs~\citep{ghazvininejad2019mask,chang2022maskgit,li2022diffusion,gong2024scaling}. Lumina-DiMOO~\citep{xin2025lumina} extends this to a multimodal setting. While these formulations naturally allow tokens to be revisited through masking~\citep{gong2025diffucoder,xie2025dream}, existing approaches primarily focus on one-shot denoising and do not exploit this capability for iterative revision.

\textbf{Editing and revision in autoregressive models.}
Autoregressive models have also been extended to support revision and editing of generated content. Prior work explores editing capabilities by inserting special tokens or performing span-level regeneration, such as insertion-based generation~\citep{stern2019insertion,gu2019levenshtein}, edit-based modeling~\citep{guu2018generating}, and controllable text editing frameworks~\citep{malmi2019encode,mallinson2020felix,mallinson2022edit5,faltings2021text}. However, such mechanisms are not native to autoregressive generation, which fundamentally operates in a forward-only manner. As a result, revising prior outputs typically requires regenerating entire sequences, introducing additional decoding passes. This makes editing indirect and often inefficient, as the model cannot modify earlier decisions in place. In contrast, mask diffusion models naturally support in-place modification through masking, enabling direct and localized revision of generated content. This structural difference provides a more suitable foundation for iterative self-correction, where previous predictions can be selectively revisited and refined without restarting the generation process.

\textbf{RemeDi and re-masking approaches.}
RemeDi~\citep{huang2025don} introduces self-reflective re-masking with a dual-stream architecture and achieves strong performance on from-scratch text generation, highlighting the importance of revising intermediate predictions in mask diffusion language models. However, RemeDi treats re-masking as an additional capability that must be explicitly learned through architectural modifications and auxiliary training objectives. This design increases both training and inference complexity, and limits its adaptability to existing MDM frameworks.

Other works explore variants of re-masking, including mixed noise schedules~\citep{von2025generalized}, predictor--corrector strategies~\citep{wang2025remasking}, and per-step resampling schemes~\citep{song2025seed}. These approaches improve sampling efficiency or stability, but continue to treat masking as an externally driven procedure rather than an intrinsic model behavior. In contrast, we view masking as a native capability of MDMs that can be activated rather than newly introduced, reframing revision as an internal decision process rather than an externally imposed operation.
