\section{Analysis of RL in Agentic Reasoning}
\subsection{Preliminaries}

\paragraph{Agentic Reasoning.}
We study the agentic reasoning of Vision-Language Models (VLMs) $\pi_\theta$ given an input $\vx$ (text question with visual image) and instruction prompt (\ref{app:tools}), interleaving thinking and tool use. At turn $t$, the model emits a thinking segment $\vt_t$ (in \think$\,\ldots\,$\thinkr) and an action $\va_t$ (a tool call formatted as JSON with tool type and arguments, in \toolcall$\,\ldots\,$\toolcallr), then receives an observation $\vo_t = \texttt{exec}(\va_t)$ (in \tool$\,\ldots\,$\toolr). 
We write $\vy_t = \vt_t \va_t \vo_t$ with $\vt_t \va_t \sim \pi_\theta(\cdot \mid \vx, \vy_{<t})$. This continues until the final answer (in \answer$\,\ldots\,$\answerr) is emitted or the turn limit is reached, yielding $\tau = (\vx, \vy_1, \ldots, \vy_T, \vy_{\text{ans}})$. Our tool set covers three standard~\citep{mini-o3, pyvision-rl, deepeyesv2}: a Python interpreter, a web search, and an image zoom-in tool.

\paragraph{RL training with GRPO.}
\label{sec:prelim-rl}

Our starting point is the standard agentic post-training pipeline~\citep{pyvision-rl, deepeyesv2}: supervised fine-tuning (SFT) on expert trajectories followed by reinforcement learning (RL) with verifiable rewards. Given an input $\vx$, the policy $\pi_\theta$ generates a group of $N$ rollouts $\{\tau_i\}_{i=1}^N$, each receiving a binary outcome reward $r_i \in \{0, 1\}$ based on answer correctness. Group Relative Policy Optimization (GRPO)~\citep{GRPO} computes a group-normalized advantage $A_i = (r_i - \mathrm{mean}(\{r_j\}_{j=1}^N)) / \mathrm{std}(\{r_j\}_{j=1}^N),$
which is assigned uniformly to every token in $\tau_i$ and used to update the policy via the PPO-clip surrogate $\mathcal{L}_{\text{clip}}$~\citep{PPO}. We defer the explicit form of $\mathcal{L}_{\text{clip}}$ in \appref{app:training-setup}.

\subsection{Diagnosing the \gap}
\label{sec:gap}
\input{figures/fig3}

We now ground the \gap in three diagnostic measurements collected during RL training (\autoref{fig:analysis}). Our starting point is the SFT checkpoint, trained with GRPO~\citep{GRPO} on outcome rewards.

\paragraph{Tool use is under-attempted.}
A defining property of agentic reasoning, unlike conventional agentic RL~\citep{arpo, ragen2}, is that tool use is \emph{optional}: the model can either produce a \toolcall tag or skip straight to \answer. This optionality makes action fragile under RL.
We track the fraction of rollouts that contain at least one tool call over the course of RL. As shown in~\autoref{fig:analysis-a}, which bins each question by its tool-using rollout count among $N=8$, the corresponding rollout-level tool-use rate is roughly 27\% at step~1 and stays in the 20--35\% range over RL training, well below the thinking-only majority.
GRPO exerts no pressure to grow the tool-using minority, and the behavior that distinguishes agentic from non-agentic reasoning remains under-represented in each group.

\paragraph{When attempted, the tool-using subgroup is prone to being all-wrong.}
Within a GRPO group of $N$ rollouts for one question, we call the rollouts that emit a tool call the \emph{tool-using subgroup} and the rest the \emph{no-tool subgroup}, and the tool-using subgroup is \emph{all-wrong} when none of its rollouts is correct, even if some no-tool rollouts in the same group succeed. Conditional on a question where tool use is attempted, how often does the tool-using subgroup fail together? \autoref{fig:analysis-b} shows this all-wrong rate is $\sim$40\% for tool-using subgroups compared to $\sim$25\% for no-tool subgroups. Under group-normalized advantage, this asymmetry yields non-positive advantage on tool-call tokens in both regimes: in mixed groups they trail the no-tool success signal, and in fully all-wrong groups the numerator $r_i - \operatorname{mean}(\{r_j\})$ collapses to zero. Either way, effective learning signal on tool calls is suppressed on the questions where tool use would matter most.

\paragraph{Tool calls admit substantive resampling at a fixed thinking.}
Resampling at the tool call is only worthwhile if it is a real divergence point under the policy.
If the thinking prefix already commits to the concrete action, resampling at the tool-call boundary changes nothing.
We probe this by decomposing a trajectory at the first tool-call boundary: we fix its thinking prefix $\vt_1$, draw 16 continuations from $\pi_\theta(\cdot \mid \vx, \vt_1)$, and cluster the resulting tool calls into semantic classes by an LLM judge that groups calls by their semantic intent. \autoref{fig:analysis-c} shows that the resampled tool calls do not collapse to a single class. Across all tools, the 16 samples occupy 2.9--3.4 distinct semantic clusters on average. The thinking prefix therefore does not determine the exact tool call and the tool call is an under-explored divergence point whose exploration must be supplied externally.

\begin{takeawaybox}{Takeaways: GRPO under-trains the tool call}
\begin{itemize}[leftmargin=*, itemsep=1pt, topsep=2pt]
    \item Tool use remains a minority behavior under GRPO: $\sim$20--35\% of rollouts across RL steps, well below the thinking-only majority (\autoref{fig:analysis-a}).
    \item When attempted, the tool-using subgroup is all-wrong on $\sim$40\% of questions (vs.\ 25\% for the no-tool subgroup), so tool-call tokens get negative advantage in mixed groups and zero in fully all-wrong groups (\autoref{fig:analysis-b}).
    \item Resampling from a fixed prefix preserves substantive diversity in the tool call (2.9--3.4 semantic clusters per 16 samples in~\autoref{fig:analysis-c}): the tool call is a real divergence point.
\end{itemize}
\end{takeawaybox}


\section{Agent eXplorative Policy Optimization (AXPO)}
\label{sec:method}

\subsection{\Toolresample}
\label{sec:branching}

In agentic reasoning, the policy must learn both thinking and tool use where the latter is the under-trained behavior under GRPO (\autoref{sec:gap}). Rather than scaling rollouts $N$ to generate tool-using successes uniformly, we commit to a thinking prefix that promises a tool call and concentrate stochasticity on the continuation.

Concretely, let $\tau^{\text{src}}$ denote a \textbf{tool-using} rollout from the standard GRPO group of $\vx$, which we call the \emph{source rollout}, and let $\vt_1^{\text{src}}$ be its prefix that includes the opening \toolcall tag, so that decoding from $\vt_1^{\text{src}}$ is constrained to produce a valid tool call.
\emph{\Toolresample} draws $K$ continuations $\{\vy_k^{\text{res}}\}_{k=1}^{K} \sim \pi_\theta(\cdot \mid \vx, \vt_1^{\text{src}})$, executes the tool calls inside each, and rolls every resampled trajectory forward until an \answer tag or the turn limit. Each resampled trajectory $\tau_k^{\text{res}} = (\vt_1^{\text{src}}, \vy_k^{\text{res}})$ therefore shares the fixed thinking prefix with $\tau^{\text{src}}$ and diverges at the tool call, concentrating stochasticity on the action that the thinking was preparing for. Each resampled trajectory receives the same binary outcome reward $r_k^{\text{res}} \in \{0,1\}$ as a standard rollout.

\paragraph{Theoretical motivation.}
\textit{Why grow tool-using rollout coverage?} Recent analysis~\citep{brorl} shows that scaling per-prompt rollouts $N$ yields a policy-improvement guarantee that grows with the correct-token mass the group covers. 
However, our learning target is correct \emph{tool-using} rollouts carry positive gradient signal for tool-call tokens, so coverage at the tool call is bounded by the tool-use rate $q = \Pr_{\pi_\theta}[\texttt{tool use} \mid \vx]$ rather than $N$, and a fraction $1-q$ of any raw sampling budget is wasted on non-tool rollouts that cannot teach the tool call. \Toolresample sidesteps this waste by committing to a prefix $\vt_1^{\text{src}}$ that has already crossed the \toolcall boundary: every resampled continuation from $\vt_1^{\text{src}}$ is tool-using by construction, so the full resampling budget lands on the signal we actually want.

Let $p(\vt_1^{\text{src}}) = \Pr_{\pi_\theta}[R=1 \mid \vx, \vt_1^{\text{src}}]$ be the success probability of a continuation from prefix $\vt_1^{\text{src}}$, and $p^{\text{tool}} = \Pr_{\pi_\theta}[R=1 \mid \vx,\, \texttt{tool use}]$ the per-tool-using-rollout success rate of standard rollout. The following proposition makes the gain precise.

\begin{proposition}[Tool-call resampling dominates raw sampling at recovering tool-using successes]
\label{prop:branching}
Let $\vt_1^{\text{src}}$ be a prefix that has crossed the tool call boundary with $p(\vt_1^{\text{src}}) \geq q\, p^{\text{tool}}$. Then for any $N \geq 1$,
\vspace{-0.05in}
\begin{equation}
    \underbrace{1 - \bigl(1 - p(\vt_1^{\text{src}})\bigr)^N}_{\text{resampling: $N$ tool-using continuations}}
    \;\geq\;
    \underbrace{1 - (1 - q\, p^{\text{tool}})^N}_{\text{raw: $N$ rollouts, $q$-fraction tool-using}},
\end{equation}
with strict inequality whenever $p(\vt_1^{\text{src}}) > q\, p^{\text{tool}}$ and $q\, p^{\text{tool}} \in (0,1)$. Both sides measure the probability that the group contains at least one correct tool-using rollout.
\end{proposition}

The key mechanism is $q \to 1$ by construction: every resampled continuation from $\vt_1^{\text{src}}$ is tool-using, so $(1-q)$ waste factor vanishes and the threshold $q\, p^{\text{tool}}$ sits strictly below the average per-tool-using-rollout success $p^{\text{tool}}$, admitting prefixes at or below the conditional mean. Under GRPO, the left-hand side equals the probability that the tool-using subgroup produces a positive gradient, so resampling strictly raises the probability of recovering an all-wrong tool-using subgroup identified in \autoref{sec:gap} relative to raw sampling at any fixed $N$. Detailed analysis is in \appref{app:coverage}.

\subsection{Resample candidate selection}
\label{sec:selection}

\paragraph{Trigger and budget allocation.}
AXPO resamples only on groups in which the tool-using subgroup is non-empty and entirely wrong (an \emph{all-wrong tool-using subgroup}; no-tool rollouts in the same group may succeed). These are exactly the groups where tool-call tokens receive non-positive advantage under GRPO (\autoref{sec:gap}): zero when the entire group is all-wrong, negative when only no-tool rollouts succeed. A recovered continuation therefore delivers the largest gradient lift per resampled continuation. \autoref{prop:branching} would in principle admit groups with at least one correct tool-using rollout as well, but those already produce positive advantage on tool-call tokens and gain comparatively little from a marginal recovery. Resampling every triggered prefix would add up to $BNK$ continuations on top of $BN$ standard rollouts (batch size $B$), so AXPO caps the extra cost up to $r \cdot BN$ per step with $r < 1$ and allocates breadth-first: every triggered question receives its top-ranked prefix before any receives a second.

\paragraph{Uncertainty-based prefix ranking.}
Not every admitted prefix is worth resampling. \autoref{fig:analysis-c} shows $\sim$30\% of prefixes collapse to a single cluster across 16 resamples. AXPO therefore ranks candidates by the mean policy probability over the tool-call tokens in $\tau^{\text{src}}$ (a tractable uncertainty proxy~\citep{deepconf}) and resamples the lowest-confidence prefixes first (see \appref{app:conf-vs-entropy} for the choice of confidence over predictive entropy).

\subsection{Advantage calculation}
\label{sec:credit}

Given the resampled trajectories from \autoref{sec:selection}, we now specify how their advantages enter the gradient. Because each $\tau_k^{\text{res}}$ shares its prefix with the source rollout $\tau^{\text{src}}$, training both as independent rollouts would assign contradictory advantages to the prefix tokens: the failed source penalizes the prefix while a successful resample rewards it. AXPO decomposes the advantage streams so that every token is updated from exactly one source.

For each selected prefix $\vt_1^{\text{src}}$, the $K$ resampled continuations form an independent advantage group. We compute a per-prefix GRPO advantage
\begin{equation}
\label{eq:axpo-adv}
\hat{A}_k^{\text{res}}(\vt_1^{\text{src}}) \;=\; \frac{r_k^{\text{res}} - \mathrm{mean}\bigl(\{r_j^{\text{res}}\}_{j=1}^K\bigr)}{\mathrm{std}\bigl(\{r_j^{\text{res}}\}_{j=1}^K\bigr)}, \qquad k = 1, \ldots, K,
\end{equation}
which is applied only to the continuation tokens $\vy_k^{\text{res}}$ where the prefix thinking tokens are masked. The source trajectory's prefix tokens are updated through $\tau^{\text{src}}$ with a binary recovery reward that fires whenever resampling recovers at least one correct continuation:
\begin{equation}
\label{eq:axpo-recovery}
r^{\text{prefix}}(\vt_1^{\text{src}}) \;=\; \mathds{1}\!\left[\exists\, k \in \{1,\ldots,K\}:\; r_k^{\text{res}} = 1\right].
\end{equation}
This recovery reward replaces the original (zero) source-rollout reward in the source group's GRPO normalization, yielding a per-prefix advantage that we apply to the source's prefix tokens as follows:
\begin{equation}
\label{eq:axpo-prefix-adv}
\hat{A}^{\text{prefix}}(\vt_1^{\text{src}}) \;=\; \frac{r^{\text{prefix}}(\vt_1^{\text{src}}) - \mathrm{mean}\bigl(\{\tilde r_j\}_{j=1}^N\bigr)}{\mathrm{std}\bigl(\{\tilde r_j\}_{j=1}^N\bigr)},
\qquad
\tilde r_j \;=\; \begin{cases} r^{\text{prefix}}(\vt_1^{\text{src}}), & j = \text{source index}, \\ r_j, & \text{otherwise}, \end{cases}
\end{equation}
and the source continuation is masked. Combining both streams with the standard clipped surrogate from PPO~\citep{PPO} yields the resampling contribution for a selected prefix $\vt_1^{\text{src}}$ to maximize:
\begin{equation}
\label{eq:axpo-loss}
\mathcal{L}_{\text{AXPO}}(\vt_1^{\text{src}}) \;=\; \underbrace{\mathcal{L}_{\text{clip}}\bigl(\vt_1^{\text{src}};\, \hat{A}^{\text{prefix}}(\vt_1^{\text{src}})\bigr)}_{\text{prefix, source only}} \;+\; \sum_{k=1}^{K}\underbrace{\mathcal{L}_{\text{clip}}\bigl(\vy_k^{\text{res}};\, \hat{A}_k^{\text{res}}(\vt_1^{\text{src}})\bigr)}_{\text{resampled continuation}}.
\end{equation}
Non-triggered groups and rollouts not selected for resampling retain the standard GRPO loss from \autoref{sec:prelim-rl}.
\autoref{eq:axpo-loss} is added once per prefix selected in \autoref{sec:selection}.
The recovery-indicator $r^{\text{prefix}}$ avoids the gradient conflict of propagating mixed-sign continuation advantages back to the shared prefix: $\hat{A}^{\text{prefix}}$ is monotone in $r^{\text{prefix}}$, so the prefix is credited positively whenever at least one continuation succeeds, converting the coverage gain into gradient signal on the prefix tokens.

