\section{Related Work}

\subsection{Video Generation Models}
\textbf{Diffusion-based Video Generation.}
The success of Diffusion Models (DMs) \cite{ho2020denoising} in image synthesis has catalyzed their extension to the video domain. Early pioneers like Video Diffusion Models (VDM) {\cite{ho2022video}} first extended the standard 2D U-Net to a 3D structure \cite{wu2026geometry} by replacing 2D convolutions with space-time factorized convolutions. To alleviate the heavy computational burden of 3D operators, many subsequent works {\cite{singer2022make, ho2022imagen, wu2023tune, zhou2022magicvideo, wang2025lavie}} adopted a "spatial-then-temporal" paradigm, inserting 1D temporal attention layers after 2D spatial blocks to capture dynamic dependencies. 
A significant architectural shift occurred with the introduction of Diffusion Transformers (DiT) {\cite{peebles2023scalable}}, which demonstrated superior scalability over U-Net. This has led to the emergence of high-performance open-source video models such as LTX-Video {\cite{hacohen2024ltx}}, which refines the VAE decoder for high-frequency detail reconstruction, and HunyuanVideo {\cite{kong2024hunyuanvideo}}, which integrates Multimodal Large Language Models (MLLMs) as text encoders to enhance text-video alignment. Building upon these advancements, Wan {\cite{wan2025}} meticulously optimizes each critical module—from the autoencoder to the text-video alignment—and provides comprehensive ablation studies to facilitate future video generation research. Furthermore, recent advancements in Flow Matching {\cite{lipman2023flowmatching}} have further optimized the training efficiency and generation quality of these diffusion-based frameworks.

\textbf{Autoregressive Video Generation.} 
Another prominent paradigm treats video generation as a sequence modeling task, analogous to Large Language Models (LLMs). Early works like VideoGPT {\cite{yan2021videogpt}} combined VQ-VAE \cite{van2017neural} with GPT-like architectures \cite{radford2018improving} to autoregressively model discrete latent tokens in a spatio-temporal grid. To extend the duration of generated content, TATS {\cite{ge2022long}} introduced a time-agnostic VQGAN and a time-sensitive transformer to synthesize thousands of frames. While diffusion models have recently dominated the field, autoregressive frameworks remain highly competitive for long-term consistency and streaming generation. For instance, VideoPoet {\cite{kondratyuk2023videopoet}} utilizes a large-scale transformer to unify multiple video-related tasks within a single LLM-style framework. More recently, MAGI-1 {\cite{teng2025magi}} incorporates causal constraints and KV caching to achieve real-time, high-fidelity video synthesis, demonstrating the enduring potential of autoregressive modeling in simulating physical dynamics and causal sequences.

\textbf{Large Video Foundation Models.} 
The landscape of video generation has been revolutionized by the emergence of large-scale foundation models that act as general-purpose world simulators. A landmark moment was the introduction of Sora \href{https://openai.com/index/video-generation-models-as-world-simulators/}{\cite{brooks2024video}}, which demonstrated that Scaling Laws \cite{kaplan2020scaling} previously observed in LLMs also apply to video: increasing parameters and training data significantly enhances the model's understanding of 3D geometry and world dynamics. 
Following this, several powerful industrial models have emerged to push the boundaries of cinematic synthesis. Kling \href{https://klingai.com/}{\cite{kuaishou2024kling}} (now updated to the Kling O3 architecture in 2026) utilizes a 3D-VAE and a computationally efficient full-attention mechanism to support ultra-long, complex human motion synthesis. Similarly, Luma Dream Machine \href{https://lumalabs.ai/dream-machine}{\cite{luma2024dream}} and Runway Gen-3 Alpha \href{https://runwayml.com/blog/introducing-gen-3-alpha/}{\cite{runway2024gen3}} focus on high-fidelity motion and temporal smoothness through massive multi-modal pre-training. 
 Furthermore, ByteDance's Seedance series \href{https://lf3-static.bytednsdoc.com/obj/eden-cn/bdeh7uhpsuht/Seedance%201.0%20Paper.pdf}{\cite{bytedance2025seedance}} has introduced a unified multimodal joint generation architecture. The latest Seedance 2.0 natively supports text, image, audio, and video inputs, leveraging a Mixture-of-Transformer-Experts (MoT) design to balance spatiotemporal consistency with high-fidelity cinematic aesthetics.

\subsection{World Models}
 
World models aim to learn compact representations of environments that enable agents to predict future states and simulate interactions. Early work such as World Models~\cite{ha2018world} demonstrated that agents could be trained entirely within a latent ``dream'' environment produced by a VAE and recurrent dynamics model. This paradigm was further extended by latent dynamics approaches including PlaNet~\cite{hafner2019planet} and the Dreamer series~\cite{hafner2019dream, hafner2020mastering, hafner2023mastering}, which perform reinforcement learning directly within imagined trajectories generated by learned world models.

With the rapid progress of large-scale generative models, the concept of world models has expanded beyond reinforcement learning toward video-based simulation of real-world dynamics. Recent approaches leverage transformer and diffusion architectures to learn rich spatiotemporal representations from large video corpora. For example, UniSim~\cite{yang2023unisim} proposes a neural closed-loop simulator for autonomous driving that generates consistent sensor observations under different actions. Similarly, large-scale video generation models such as Sora~\cite{brooks2024video} have demonstrated the capability to simulate complex physical interactions and long-term temporal consistency, suggesting that generative video models may serve as general-purpose world simulators.

More recent work explores interactive world modeling by conditioning generation on actions \cite{wu2026multiworld} or control signals. GAIA-1~\cite{hu2023gaia} learns a generative world model for autonomous driving that predicts future video conditioned on vehicle controls. Genie~\cite{bruce2024genie} further introduces generative interactive environments capable of synthesizing playable worlds from video data. Recently, Cosmos~\cite{nvidia2025worldsimulationvideofoundation} proposes a large-scale world foundation model designed for robotics and physical AI, enabling realistic simulation of environments conditioned on agent actions. These advances indicate a promising direction toward unified world simulators that combine perception, dynamics modeling, and controllable environment generation.

\subsection{World Action Models}

\textbf{From VLA to WAMs.}
Generalist embodied agents have long relied on Vision-Language-Action (VLA) models that learn reactive, direct observation-to-action mappings without modeling how the physical world evolves under intervention. To overcome the short-sightedness and weak physical grounding of such purely reactive policies, the World Action Model (WAM) paradigm unifies predictive environment dynamics with motor control by jointly targeting the distribution over future states and actions. By internalizing physical laws through future simulation, WAMs enable stronger long-horizon reasoning and spatial awareness than standard VLAs—though this unified objective introduces sharp architectural trade-offs between generation fidelity and inference latency.

\textbf{Cascaded WAMs.}
Cascaded WAMs synthesize future visual representations conditioned on task goals before extracting control actions. Explicit methods first forecast raw 2D RGB frames using inverse dynamics or multimodal conditioning \cite{du2023learning,du2024video,intelligence2026pi}. To mitigate spatial hallucinations inherent in pixel-level generation, subsequent explicit architectures extract intermediate 2D geometric representations like optical flow and point tracks \cite{ko2024learning,bharadhwaj2024track2act,xu2024flow}, or extend into 3D and 4D spaces for rigorous spatial reasoning \cite{zhi20253dflowaction,wang2026mvista}. Despite providing highly interpretable plans, explicit decoding suffers from severe computational latency, rendering high-frequency closed-loop control intractable \cite{yan2026s}. To circumvent this bottleneck, implicit cascaded WAMs bypass high-dimensional image rendering by encoding anticipated futures strictly within continuous or discrete latent spaces. This latent planning involves learning quantized semantic codebooks \cite{ye2025latent}, imposing geometric mask bottlenecks \cite{lou2026mask}, or operating entirely within diffusion representations \cite{huang2024ardup}. To further accelerate reactive control, frameworks extract intermediate network features \cite{hu2024video} or self-distill multi-step generation into a single feed-forward pass  \cite{yan2026s}. Ultimately, while explicit WAMs excel in modularity with off-the-shelf simulators, implicit approaches prioritize the execution scalability and efficiency required for real-time deployment.

\textbf{Joint World-Action Models.}
Departing from the two-stage pipeline, joint WAMs co-model decision-making and physical dynamics within a shared space, splitting into autoregressive and diffusion-based families. The autoregressive family casts states and actions as a unified sequence over a shared vocabulary: causal transformers first predict pixel-space trajectories as implicit planners \cite{wu2024unleashing, cheang2024gr}, then structure a visual chain-of-thought via intermediate subgoal images \cite{zhao2025cot}, and ultimately fuse text, images, and discrete actions into one vocabulary \cite{cen2025worldvla}. Since strict autoregressive action decoding causes error propagation and trajectory drift, remedies include action-attention masking \cite{cen2025worldvla} and a hybrid design with a parallel continuous action head to avoid quantization errors \cite{cen2025rynnvla}; still, sequentially decoding high-dimensional visual states incurs heavy latency that hampers reactive deployment. Conversely, diffusion-based WAMs frame prediction as denoising or flow-matching, jointly optimizing states and actions to expose diverse marginal and conditional distributions for deeper causal understanding and stronger sample efficiency. Early designs fuse all modalities in a single diffusion transformer \cite{guo2024prediction,shen2026videovla,kim2026cosmos}, while later works resolve action-image modality conflicts through separate or bridged token streams \cite{won2025dual,yang2025covar,chen2025unified}. To avoid costly pixel-level reconstruction, several adopt implicit latent world modeling \cite{zheng2025flare,zhao2026frappe,li2026world,yuan2026adaworldpolicy}, and others decouple diffusion timesteps or scale via mixture-of-transformers \cite{zhu2025unified,bi2025motus,team2026motubrain,lyu2026lda}. For inference efficiency, recent frameworks adopt lightweight heads, asynchronous sampling, or speculative decoding \cite{li2025unified,ye2026gigaworld,yuan2026fast,guo2026unified,deng2026dexworldmodel}. Across both families, the central challenge remains balancing high-fidelity joint simulation in training against lightweight, asymmetric modality decoupling at inference.

%\textbf{Kairos within the WAM Landscape.}
%Kairos belongs to the diffusion-based joint WAM family but is built to resolve its training--inference tension, differing on three axes. \textit{Pre-trained priors, not from-scratch co-training:} rather than co-training video and action branches from scratch \cite{guo2024prediction,shen2026videovla,kim2026cosmos}, Kairos natively pre-trains a Video DiT backbone with Hybrid Linear Temporal Attention (SWA, DSWA, GLA), then attaches a lightweight Action DiT that inherits its spatiotemporal priors. \textit{Joint causality at cascaded cost:} an asymmetric masking scheme decodes actions without materializing future video, giving implicit-cascaded efficiency while keeping them grounded in the jointly learned world dynamics. \textit{Cross-embodiment, not curated robot data:} a Cross-Embodiment Data Curriculum bridges open-world video, human demonstrations,and robot interaction, tackling the fragmented-learning bottleneck. Together, these make Kairos not just another joint WAM variant, but a deployment-ready, world-model-native substrate for self-evolving physical AI.

\subsection{Efficient Attention Mechanisms}

The quadratic complexity of the standard self-attention mechanism in Transformers poses a major challenge for modeling long sequences such as high-resolution videos. Given an input sequence of length $N$, vanilla attention requires $\mathcal{O}(N^2)$ time and memory, which quickly becomes prohibitive as the spatial and temporal dimensions increase. To address this limitation, a large body of work has explored efficient attention mechanisms that reduce computational complexity while preserving the modeling capacity of Transformers.

Early approaches focused on approximating the softmax attention through kernelization or low-rank decomposition. Linear Transformers~\cite{katharopoulos2020transformers} reformulated softmax attention using kernel feature maps, enabling attention computation in linear time with respect to sequence length. Similarly, Performer~\cite{choromanski2020rethinking} introduced FAVOR+ random feature approximations to achieve scalable attention with theoretical guarantees. Other works explored sparse attention patterns, restricting interactions to local or predefined structures in order to reduce computational cost \cite{child2019generating,beltagy2020longformer,zaheer2020big}.

More recent research has shifted toward designing sequence models with recurrent or state-space style updates that scale linearly with sequence length. Retentive Networks (RetNet)~\cite{sun2023retentive} replace the softmax attention with a retention mechanism that supports both parallel training and recurrent inference. The Mamba architecture~\cite{gu2024mamba,dao2024transformers,lahoti2026mamba,huang2024localmamba,pei2025efficientvmamba} further demonstrates that selective state space models can achieve strong sequence modeling performance while maintaining linear complexity. 
Building on these developments, recent works further explore gated update mechanisms to improve the stability of long-context modeling. Delta-based architectures such as Gated Delta Networks~\cite{yanggated} introduce gated update rules that mitigate key-value interference during sequence updates. In a similar spirit, Gated Linear Attention (GLA) incorporates gating mechanisms into linear attention to regulate information flow\cite{yang2023gated,li2025minimax}. By introducing learnable gates into the attention update, GLA selectively integrates new key-value information while suppressing outdated context, enabling stable long-range dependency modeling with $\mathcal{O}(N)$ complexity \cite{qin2023hierarchically,qin2024hgrn2,katsch2023gateloop,beck2024xlstm}. Such efficiency is particularly beneficial for tasks involving extremely long token sequences.
