\section{Conclusion}
\label{sec:conclusion}
We introduced \textbf{\ours}, a verifier-grounded framework for building verifiable software worlds for computer-use agents. \ours makes inspectable application state a core design constraint across verifier construction, task synthesis, and benchmark execution. This enables the automatic generation of executable desktop tasks with machine-checkable success criteria while preserving the diversity and realism of real software workflows. In its current form, \ours covers 33 desktop applications and 1,000 finalized tasks across browsers, office tools, creative software, development environments, communication tools, and system utilities.

We show that realistic desktop environments expose important failure modes in current computer-use agents. Frontier agents often make meaningful partial progress, but reliable end-to-end completion remains difficult when success depends on fine-grained application state, persistent files, metadata, or hidden side effects. Our comparison between LLM-as-judge and hard-coded verification further shows that screenshot-based or trajectory-level judgments can miss subtle but consequential errors, whereas executable verifiers can directly inspect the final software state.

More broadly, we view \ours as infrastructure for scaling computer-use research. Progress requires not only stronger models, but also trustworthy environments, grounded rewards, reproducible task construction pipelines, and verifiers that support both evaluation and training. By coupling realistic software worlds with machine-checkable feedback, \ours provides a foundation for studying agent reliability, collecting grounded trajectories, analyzing failures, and improving agents through supervised learning, rejection sampling, or reinforcement learning. We hope this work and the released repository help make future computer-use systems more reliable, measurable, and aligned with real software outcomes.