\section{Experiment}
\label{sec:experiment}
\subsection{Experimental Setup}
\label{sec:experimental-setup}
We design our experiments to evaluate whether OpenComputer provides reliable and challenging
software-world environments for computer-use agents. Our evaluation focuses on two questions:
(1) whether current frontier and open-source agents can complete the synthesized tasks, and
(2) whether our verifier-grounded reward computation can measure both exact success and
partial progress across heterogeneous desktop applications.
\input{figure/benchmark_result}
\paragraph{Benchmark.}
We evaluate agents on the finalized OpenComputer task suite. Each task consists of a natural-language
instruction, an executable sandbox initialization, and a set of machine-checkable success criteria.
The benchmark spans 33 desktop applications. For each task, the agent is placed in a fresh desktop sandbox initialized with the required files, profiles, configuration state, and application artifacts. The agent then interacts with the live GUI
through screenshots and desktop actions until it stops or reaches the step budget. We include OSWorld-Verified~\citep{xie2024osworld} as an external reference to contextualize model performance against a widely used desktop-agent benchmark.

\paragraph{Models.}
We evaluate a mixture of frontier proprietary agents and open-source computer-use models. The
main models include GPT-5.4~\citep{openai2026gpt54}, Claude-Sonnet-4.6~\citep{anthropic2026sonnet46}, Kimi-K2.6~\citep{moonshot2026kimi26}, Gemini-3-Flash~\citep{google2025gemini3flash}, Qwen-3.5-27B~\citep{qwen2026qwen35},
Qwen-3.5-9B~\citep{qwen2026qwen35}, EvoCUA-8B~\citep{xue2026evocua}, and GUI-OWL-1.5-8B~\citep{xu2026mobile}. For Gemini-3-Flash, which does not
provide a built-in desktop action space in our evaluation setting, we prompt the model to emit actions in a Qwen-style computer-use format. All open-source models except Kimi-K2.6 (which we use the official APIs) are deployed with two H100 GPUs.

\paragraph{Metrics.}
We report both task-level and criterion-level metrics. The primary task-level metric is \emph{success
rate}, defined as the fraction of tasks for which all required criteria are satisfied. Because many desktop
tasks contain multiple independent requirements, we also report \emph{average reward}, defined as
the mean fraction of passed verifier checks, this metric gives partial credit when an agent completes some but not all required subtasks. To measure
efficiency, we additionally report the average number of interaction steps and the average wall-clock
time per step.


\subsection{Main Results Analysis}

Table~\ref{tab:model_results} reports the overall performance and efficiency of representative 
% frontier and open-source
computer-use agents on OpenComputer. The results show that OpenComputer is challenging even
for the strongest current agents. GPT-5.4 achieves the best overall performance, with an average
reward of 88.4\% and a task success rate of 68.3\%, but it still fails to completely solve nearly one
third of the benchmark tasks. Claude-Sonnet-4.6 and Kimi-K2.6 follow closely, reaching success
rates of 64.4\% and 58.8\%, respectively. This indicates that frontier models can often make
substantial partial progress, but reliable end-to-end task completion in realistic desktop software
remains far from saturated.

GPT-5.4 is also the most efficient agent in terms of interaction length. It completes tasks in only
19.0 steps on average, substantially fewer than Claude-Sonnet-4.6, Kimi-K2.6, and the open-source
models. One reason is that GPT-5.4 frequently combines multiple low-level operations into a single
computer-control step, reducing the number of interaction rounds needed to complete a task. In
addition, GPT-5.4 does not emit long reasoning traces in our evaluation setting but only the executable actions, which reduces output overhead and improves per-step execution efficiency. This combination of shorter trajectories
and lower textual overhead makes it particularly effective for controlling computer environments.

The external OSWorld-Verified scores provide additional context. Several open-source models have moderate reported OSWorld performance, but their success rates drop substantially on OpenComputer.
For example, GUI-OWL-1.5-8B has a reported OSWorld score of 52.3\%, but achieves only 5.7\%
success on our benchmark; EvoCUA-8B similarly drops from 46.1\% on OSWorld to 10.9\% on
\ours. This gap suggests that these models have limited cross-benchmark generalization,
and that strong performance on existing desktop benchmarks does not necessarily transfer to the
broader and more heterogeneous software settings covered by \ours.
